### 2020-03-15 CoverPage-Evolution.rmd
Setting up repository. Attempting to knit manuscript template on my (EKB) comp.

---
title: A MULTIVARIATE EQUIVALENCY AND GENERAL TEST FOR RENSCH'S RULE: IMPLICATIONS FOR THE EVOLUTION OF SEXUAL SHAPE DIMORPHISM

output:
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: false
    pandoc_args:

csl: evolution.csl
bibliography: bibliography.bib

header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---
\begin{center}
\textbf{Dean C. Adams$^{1,*}$, Elizabeth Glynne$^{1}$ and Antigoni Kaliontzopoulou$^2$}
\end{center}

\begin{center}`r format(Sys.time(), '%d %B, %Y')`\end{center}

$^{1}$Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.

$^{2}$CIBIO Research Centre in Biodiversity and Genetic Resources, InBIO, Universidade do Porto, Campus de Vairão. Vila do Conde, Portugal.

$^{*}$Correspondence: Dean C. Adams <dcadams@iastate.edu>

\hfill\break
**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break

**Short Title**: Multivariate Test of Rensch's Rule \hfill\break

**Author Contributions**: DCA devised the approach, AK provided the empirical example; all authors contributed  to all parts of this manuscript. \hfill\break

**Data Archiving**: Data will be made available on DRYAD upon publication. \hfill\break

**Acknowledgments**: We thank XXX for comments on drafts of this manuscript; Ana Perera, Inés Freitas and Daniele Salvi for the phylogeny reconstruction; and Chiara Ripa and Zélia Pereira for aiding shape data acquisition. This work was  sponsored in part by National Science Foundation Grant DBI-1902511 (to DCA), and by FEDER Funds through the Operational Competitiveness Factors Program - COMPETE and National Funds through FCT - Foundation for Science and Technology within the scope of the project “PTDC/BIA-EVL/28090/2017 - POCI-01-0145-FEDER-028090” (to AK).

### 2020-03-15 Manuscript-template.rmd
Setting up repository. Attempting to knit manuscript template on my (EKB) comp.

---
title:

output:
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: evolution.csl
bibliography: bibliography.bib

header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---
# A multivariate equivalency and general test for Rensch's Rule: Implications for the evolution of sexual shape dimorphism

\hfill\break

**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break

**Short Title**: Multivariate Test of Rensch's Rule \hfill\break

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Rensch's rule, the allometric association between the degree of sexual dimorphism and overall body size, is an important framework for understanding the evolution of trends in phenotypic differentiation between the sexes across taxa. While virtually all recent treatments have explored the degree of sexual size dimorphism, understanding how dimorphism in other phenotypic traits evolves is equally relevant. Indeed, Rensch's original treatment focused not only on size, but also on dimorphism in relative body proportions, coloration, the presence of ornaments, etc. Here, we derive a multivariate equivalency for viewing trends in sexual dimorphism – relative to overall body size – across taxa, and provide a generalized test to determine whether such patterns are consistent with Rensch's rule. For univariate linear traits such as body size, our approach yields equivalent results to those from standard procedures, but our test is also capable of detecting trends in multivariate datasets like shape. Computer simulations reveal the method displays appropriate statistical properties, and an empirical example in Mediterranean lizards illustrates the efficacy of our approach on both univariate and multivariate phenotypes. Our generalized procedure substantially extends the toolkit available for investigating macroevolutionary patterns of sexual dimorphism and seeking a better understanding of the processes that underlie them.

\newpage

# Introduction
It is widely observed that males and females of the same species are not identical, but rather differ -- sometimes substantially -- in their phenotypic characteristics [@Fairbairn1997; @Fairbairn2013]. Indeed, sexual dimorphism is pervasive across many animal clades, and understanding the causes for these patterns has attracted considerable attention in evolutionary biology for well over a century [e.g., @Darwin1871; @Fairbairn2007]. One of the most conspicuous differences between males and females is in their body sizes [@Andersson1994], where sex-specific differences are often hypothesized to be the result of selection pressures related to reproduction, or the distinct ecological roles of males and females in their respective habitat [@Stephens2009; @Littleford-Colquhoun2019; @Tarr2018; @Kaliontzopoulou2015; @Cox2003]. For example, both sexual selection on male body size [@Cox2003; @Garcia-Navas2015; @Horne2020], and fecundity selection on female body size [@Stuart-Fox2008; @Serrano-Meneses2006], are expected to enhance the degree of sexual size dimorphism in taxa over evolutionary time. Likewise, when males and females respond differently to environmental factors, intraspecific competition may generate ecological divergence between them, resulting in varying patterns of sexual dimorphism [@DayanSimberloff2005; @Meiri2014; @Kaliontzopoulou_et_al2010; @Butler_et_al2000; @Bolnick2003]. \hfill\break

At macroevolutionary scales, changes in the relative contribution of the selective forces above may create variation in the degree of sexual dimorphism displayed across closely related taxa. For instance, sister species inhabiting different environments, or possessing distinct ecological adaptations, may exhibit differing degrees of sexual dimorphism, as a result of the varying intensity of sexual selection they endure [@Oestman2011]. Likewise, the interplay between sexual and natural selection that species experience may result in sexual dimorphism of differing magnitudes across taxa [@Kaliontzopoulou2015; @Stuart-Fox2007; @Blanckenhorn2006]. One evolutionary framework that connects adaptive explanations for the evolution of sexual size dimorphism (SSD) with trends in overall body size across taxa is Rensch's rule. Here it is predicted that SSD will increase with species' average body size in species where males are the larger sex (hyperallometry), and likewise will decrease with increasing body size where females are the larger sex (hypoallometry) [@Abouheif1997; @Fairbairn1997]. Put another way, Rensch's rule predicts that the extent of sexual dimorphism increases at more extreme body sizes (both large and small) across taxa. Empirical studies have documented trends in SSD consistent with Rensch's rule in a wide variety of animal taxa [@Dale2007; @Ceballos2013; @Regis2017; @Johnson2017]. However, other clades display the converse trend [@Burbrink2019; @Penalver-Alcazar2019], and some clades display no relationship between SSD and body size [@Astua2010; @Hirst2014; @Johnson2017]. From an evolutionary perspective, the association between SSD and body size has been hypothesized to be driven by variation in ontogenetic or static allometries across taxa, which results from the influence of sexual selection on growth and maturation patterns [@Rensch1960; @Dale2007]. \hfill\break

Interestingly, virtually all evaluations of Rensch's rule have investigated patterns of sexual size dimorphism, yet Rensch's original contributions were concerned with a much wider array of phenotypic traits that can differ between the sexes. These included not only sexual differences in body size, but also in general body proportions (e.g., relative wing and limb lengths), the degree of complexity of different body structures, the presence of sex-specific  ornamentation (e.g., horns, body coloration, etc.), and other sex-specific differences [@Rensch1950; @Rensch1960]. In like manner, some studies have documented patterns of sexual shape dimorphism [e.g., @Berns2013; @Kelly2013; @Kaliontzopoulou2015], sex-specific differences in ornamentation [@Geist1988; @Watson2010; @Emlen2008], or coloration [e.g., @Endler1983]. However, such studies have typically been restricted to one or a few taxa, and have not explained macroevolutionary variation in sexual dimorphism in a manner that relates directly to Rensch's rule. Thus, when viewed from this perspective, Rensch's [-@Rensch1950] vision was far more synthetic, as he sought to explain the extent to which sexual dimorphism -- across a general suite of phenotypic characteristics -- scaled with overall body size in organisms, and why some species displayed a greater degree of sexual dimorphism than did others. Remarkably, patterns of intraspecific static allometry in sexually selected traits establishes a direct link between variation in those traits relative to size, and the evolutionary allometry across taxa in the degree of sexual dimorphism that may be expected [@Reiss1986; @Bonduriansky2007; see also @Rensch1960]. Under this framework, the "exaggeration" of body structures or other phenotypic traits with increasing body size is predicted to translate, at macroevolutionary scales, into increased degrees of sexual dimorphism with increasing body size. However, despite this clear prediction, and Rensch's original observations, the evolutionary scaling of the degree of sexual dimorphism in phenotypic traits other than size, remains largely unexplored. \hfill\break

Another current limitation in our understanding of Rensch's rule is the fact that to date, all studies have been conducted exclusively on univariate data, primarily regarding body size. However, understanding phenotypic evolution is inherently a multivariate endeavor [@Blows2007; @Collyer_et_al2015a; @AdamsCollyer2018a; @AdamsCollyer2019], as processes such as natural and sexual selection simultaneously operate on more than one trait [@Lande1979; @LandeArnold1983], and can affect variation in complex multidimensional traits such as shape. To interrogate such patterns, the analytics used to evaluate evolutionary trends in phenotypic datasets must be sufficiently general to accommodate this empirical fact. Unfortunately, all current analytical approaches for evaluating trends in sexual dimorphism as they relate to overall body size are explicitly univariate; thereby excluding the possibility of evaluating patterns in complex multidimensional traits such as organismal shape. Though is has been extensively documented in sexual size dimorphism, it remains unknown whether Rensch's rule applies equally to sexual shape dimorphism, as determining this requires an analytical framework capable of evaluating trends in sexual dimorphism in both univariate and multivariate datasets. \hfill\break

In this paper, we derive a multivariate equivalency for inspecting variation in sexual dimorphism -- relative to overall body size -- across taxa, and provide a generalized test to determine whether they are consistent with Rensch's rule. Our procedure is appropriate for evaluating trends in univariate traits measured on a linear scale (such as body size or relative limb length), as well as in multivariate datasets representing complex phenotypes (such as sets of body proportions or landmark-based shape data). For univariate data, the approach yields equivalent results to what is normally accomplished for size-based SSD evaluations. However, the approach extends these procedures for the examination of sexual dimorphism in multivariate data, thereby increasing the potential for our conceptual understanding of the consequences of sexual and natural selection on complex organismal phenotypes. We conduct a series of computer simulations demonstrating that the approach displays appropriate statistical properties when evaluating trends in both univariate and multivariate data. We also provide an empirical example in Mediterranean green lizards which illustrates the efficacy of our approach on both univariate and multivariate datasets. Finally, we highlight the new avenues of evolutionary research that our generalization facilitates, and make comments on the prospects for future empirical advances.

# Methods and Results

## *Conceptual Development*

To arrive at a general test for Rensch's rule that can accommodate multivariate data, we must first derive a series of mathematical equivalencies based upon how interspecific patterns of size dimorphism are typically evaluated. For sexual size dimorphism, interspecific patterns are often inspected using a bivariate plot, where the size of one sex is plotted against the size of the other sex [e.g., @Fairbairn1994; @Abouheif1997; @Ceballos2013]. An idealized example is shown in Fig. 1A, with male size plotted against female size. The diagonal line represents no size dimorphism; or a 1:1 size ratio between the sexes across the range of body sizes for species in the clade of interest. In this construction (i.e., with females plotted on the x-axis), deviations from this line represent differential trends in size dimorphism across taxa. To evaluate such patterns statistically, a phylogenetic regression is performed, and in this case, a slope significantly greater than 1.0 (i.e., $\beta_{1}>1.0$) would be treated as evidence that the evolutionary changes in sexual size dimorphism across taxa are consistent with Rensch's rule [@Fairbairn1997]. \hfill\break

An equivalent representation of patterns of sexual dimorphism may also be found using ratios (e.g., male size:female size). When plotted against the average size for each species (Fig. 1B), the same information is conveyed as is found in the more commonly utilized male - female plot of Fig. 1A. However, in this representation, the line designating no sexual dimorphism is horizontal, with a slope of zero (i.e., $\beta_{1}=0.0$) and an intercept of 1.0. As before, a phylogenetic regression may be performed on these data, and in this case, if the slope is significantly steeper than zero (i.e., $\beta_{1}>0.0$), patterns consistent with Rensch's rule are observed. \hfill\break

A third equivalency is shown in Fig. 1C. Here, the Euclidean distance between male and female means for each species ($D_{SD}=\sqrt{(Y_{male}-Y_{female})^2}$) is shown relative to the overall size of the species. This plot is similar to a Bland-Altmann plot [@Altman1983], except that the Euclidean distance, rather than a difference score, is plotted along the y-axis (as distances can accommodate data of different dimensionality). Under Rensch's rule, this *sexual dimorphism distance* may become greater as species are progressively larger [hyperallometry, sensu @Abouheif1997] or may become greater as species are progressively smaller [hypoallometry, sensu @Abouheif1997]. In both cases, patterns correspond to species displaying progressively more sexual dimorphism at extreme body sizes; a trend consistent with Rensch's rule. Note that the inflection point in Fig. 1C represents an intermediate size where there is no sexual dimorphism for the trait, and corresponds to where the male-female regression line of Fig. 1A crosses the line representing no sexual dimorphism. This inflection point may represent the overall mean size across species in the lineage, or may be some other value. For instance, in some lineages [e.g., turtles: @Ceballos2013], nearly all species are female biased, in which case the inflection point for $D_{SD}$ vs. size would be found at the far-right end of the plot. \hfill\break

Finally, from $D_{SD}$ a *sexual dimorphism distance index* ($SDDI$) may be constructed as:

$$SDDI=-1*D_{SD}$$
$$SDDI=+1*D_{SD}$$

Here, $D_{SD}$ is multiplied by either a +1 or -1, based on a comparison of male and female trait values for the species. For instance, when single-valued traits on a linear scale are examined (e.g., leg length or wing extent), $D_{SD}$ is multiplied by -1 for those species where females display larger values than males, and remains unchanged for species where males are larger than females. This logic is analogous to that of Lovich and Gibbons [-@Lovich1992] for their index of overall body size dimorphism. However, from a geometric perspective, this procedure is tantamount to determining which species in Fig. 1A are found below the 1:1 size line (i.e., the line representing no size dimorphism), and multiplying those species' $D_{SD}$ by -1. \hfill\break

Unfortunately, applying the same logic to multidimensional trait data is more complicated, as sex-specific differences are not always guaranteed to align with some *a priori* direction in the multivariate dataspace. Thus, for multivariate datasets, one of two procedures may be used. First, for multivariate sets of traits that are measured on the same linear scale (e.g., length, width, and height measures), one could envision a hyperdimensional version of the plot displayed in Fig. 1A, but where the set of male traits are plotted against the set of female traits. In this construction, because the traits are quantified on the same measured scale, the 1:1 size vector representing no sexual dimorphism does exist, and emanates from the origin of the space extending in a positive direction as the trait values increase. To determine whether male trait values deviate more from this vector than do female trait values, a partial least squares analysis may be performed [@RohlfCorti2000]. Partial least squares (PLS) is a multivariate association procedure which describes the covariation between sets of traits (in this case, between the set of male traits and the set of female traits). Scores along the first PLS axis represent the maximal covariation between male and female traits projected to a single dimension, and species whose male and female PLS scores are identical are those displaying no sexual dimorphism. Thus, deviations from that pattern represent instances of multivariate sexual dimorphism. By plotting these scores, one may determine which species display female scores (on $PLS_1$) that are larger than their corresponding male scores (when negative values on $PLS_1$ correspond to smaller individuals; otherwise, the converse comparison is used). The $D_{SD}$ for these species may then be multiplied by -1 to obtain the $SDDI$ for that species. Note that when PLS is used on univariate data, this procedure obtains values identical to those from the method described previously where male and female sizes are compared directly. Thus, for multivariate traits measured on a linear scale, this is a direct generalization of the standard univariate procedure (see empirical example below). \hfill\break

Alternatively, one might be interested in examining multidimensional traits other than linear biometric measurements, such as color or shape. For instance, one may characterize the shape of males and females using geometric morphometric methods [@Bookstein1991; @MitteroeckerGunz2009; @Adams_et_al2013], where the locations of anatomical landmarks and semilandmarks are used to obtain a set of shape variables. Geometric morphometric methods also result in a multidimensional characterization of phenotypic traits, from which the degree of sexual dimorphism for each species may be quantified using $D_{SD}$. However, because the resulting shape variables do not have a one to one correspondence with linear size measures, there is no *a priori* direction in the multivariate shape space that is equivalent to the 1:1 size vector found with sets of linear traits. In fact, for the case of isometry (i.e., shape does not change as size changes), all observations are represented by a single location in the shape space. Thus, while a PLS of male shapes versus female shapes will describe the covariation between them, this direction is not guaranteed to align with size, because there is no direction in the shape space that may be used to unambiguously determine whether dimorphism is male- or female-biased, as for shape variables there is no *a priori* directionality. Nonetheless, one may still quantitatively determine whether the degree of sexual shape dimorphism covaries with overall body size, by using the sexual dimorphism index as defined by the dimorphism distance (i.e., $SDDI=D_{SD}$) to represent the degree of multivariate shape dimorphism for each species. \hfill\break

Once the sexual dimorphism distance index ($SDDI$) is calculated, it may be evaluated against the average size of each species. Here, data conforming to Rensch's rule will result in significant trends of sexual dimorphism relative to mean body size (Fig. 1D). As with size ratios, the line of no sexual dimorphism is a horizontal line, but this time it is centered on zero, as a distance of zero corresponds to no sexual dimorphism. Thus, a phylogenetic regression of these data whose slope is significantly different from zero (e.g., $\beta_{1}\neq0.0$) provides evidence that the evolution of sexual dimorphism is consistent with patterns expected under Rensch's rule. For cases where traits are measured on a linear scale, patterns expected under Rensch's rule exhibit a slope significantly greater than zero ($\beta_{1}>0.0$), as in the usual formulation. Likewise, when using geometric morphometric shape data (or any other type of mean-centered multidimensional data), hyperallometric patterns consistent with Rensch's rule would also display a slope significantly greater than zero ($\beta_{1}>0.0$), indicating that the degree of sexual shape dimorphism increases with increasing body size. Conversely however, hypoallometric patterns would display the opposite pattern ($\beta_{1}<0.0$), indicating that smaller species display greater sexual shape dimorphism. \hfill\break

## *Simulations*

To ascertain the statistical performance of the approach proposed here, we conducted a series of computer simulations. We evaluated the ability of SDDI to characterize patterns of sexual dimorphism in both univariate and multivariate datasets. For each simulation, we generated a pure-birth phylogeny containing 50 species, on which we simulated datasets under a Brownian motion model of evolution. For univariate data ($p=1$), a single trait was evolved along the phylogeny, and represented the female trait values ($x$). Next, male values ($y$) were simulated in such a manner as to generate a known male:female relationship (plus random error), using the function: $y=\alpha{x}+\mathcal{N}(\mu=0,\sigma=0.1)$. For type I error simulations, no sexual dimorphism was generated, meaning that the slope, $\alpha$, characterized a 1:1 relationship between male and female trait values (i.e., $\alpha=1.0$). For power simulations, we simulated male values with a slope greater than 1.0 to correspond with patterns following Rensch's rule (i.e., $\alpha=1.1$).  A total of 1,000 datasets were generated under each simulation condition. We then evaluated all datasets using standard approaches (i.e., a phylogenetic regression of male values versus female values), as well as a phylogenetic regression of SDDI versus average body size. The proportion of significant datasets was treated as the type I error ($\alpha=1.0$) or power ($\alpha=1.1$) respectively. \hfill\break

For multivariate data ($p=5$), a slightly different procedure was utilized. First, average body size ($s$) was evolved along the phylogeny under Brownian motion. Next, female traits ($\mathbf{X}$) were simulated on the phylogeny under Brownian motion, using a $p\times{p}$ trait covariance matrix with correlations between traits of ($r=0.7$). For simulations where there was no pattern of sexual dimorphism across species (i.e., type I error simulations) male trait values were then generated by incorporating random error to the female trait values, in a manner analogous to that achieved in the univariate simulations: $\mathbf{Y}=\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1$). For simulations where patterns of sexual dimorphism covaried with body size (i.e., power simulations), male trait values were generated by incorporating random error to the female trait values, but with the addition that female trait values were pre-multiplied by a scalar value of body size: $\mathbf{Y}=(0.1*s)*\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1$). This ensured that male and female trait values covaried with one another as expected, and also covaried with size in the desired direction. As before, a total of 1,000 datasets were generated under each simulation condition. We then evaluated all datasets using a phylogenetic regression of SDDI versus average body size. The proportion of significant datasets was treated as the type I error ($\alpha=1.0$) or power ($\alpha=1.1$) respectively. \hfill\break

*Simulation Results*: For the univariate dataset, we found that both a phylogenetic regression of male size on female size, as well as a phylogenetic regression of $SDDI$ on body size, resulted in appropriate type I error rates near the nominal $\alpha\approx0.05$ (Fig. 2). Further, when data were simulated under conditions consistent with Rensch's rule, both procedures were capable of detecting the signal of the pattern, and did so with very high power. Thus, using the $SDDI$ approach proposed here was equivalent to what is normally utilized for univariate datasets. Likewise, for multivariate data, the $SDDI$ approach displayed appropriate type I error ($\alpha\approx0.05$) and high statistical power under the conditions simulated (Fig. 2). Additionally, multiplying $D_{SD}$ by $\pm1$ had little effect on the outcome for multivariate data, as in both cases type I error and power results were virtually identical. Thus, methods for evaluating sexual dimorphism using geometric morphometric data are anticipated to remain robust. Overall results from these simulations imply that using the sexual dimorphism distance index ($SDDI$) proposed here provides an appropriate means of characterizing patterns of sexual dimorphism in both univariate and multivariate traits, and that the procedure is capable of detecting patterns consistent with Rensch's rule when they are present in phenotypic datasets. \hfill\break

## *Empirical Example*

To illustrate the utility of the approach presented here, we conducted an analysis of both size and shape dimorphism in a group of Mediterranean lizards with extensive body size variation. Our dataset consisted of three data types: male and female values of body size (snout-vent length: SVL), male and female values for a series of head measurements (head length, head width, head height, pileus length, and mouth opening: Fig. 3A), and male and female values of head shape (Fig. 3B). A total of 374 specimens from 21 lineages of green lizards, a monophyletic clade that includes two genera (*Lacerta* and *Timon*), were obtained from natural history museum collections. For each lineage, we obtained samples distributed across the entire distribution range, resulting in an initial database of 1062 specimens. In order to take sampling bias into account, we selected the 10 largest individuals per sex, when sufficient specimens were available, for use in subsequent size and shape analyses. A detailed account of the specimens used for this study is available in the Supplemental Material. \hfill\break

From these specimens, we calculated mean body size for each sex and lineage. Likewise, we calculated the mean trait values for each sex and lineage for the set of linear measures. These values were then standardized by SVL; resulting in a set of head measurements proportional to body size for each sex $\times$ lineage combination. Additionally, we quantified lateral head shape variation using geometric morphometric methods [@Adams_et_al2013]. This was accomplished by first digitizing 17 landmarks and semilandmarks from images of the right side of the head (Fig. 3B). The thin-plate spline was used to estimate the position of missing landmarks [@Gunz2009], and the position of the jaw relative to the skull was standardized using the fixed angle method [@Adams1999]. Next, a generalized Procrustes analysis  [@RohlfSlice1990] was performed to eliminate non-shape variation from the landmark data and align the specimens to a common coordinate system. The position of semilandmarks (i.e. 11-16, Fig. 3B) was optimized by minimizing the bending energy [@Bookstein1997]. From the aligned Procrustes coordinates the mean head shape for males and for females was then calculated for each lineage. Finally, the phylogenetic relationships among lineages were estimated using published sequences on five mitochondrial and four nuclear gene fragments, totaling 8383 bp. These were then used following standard phylogenetic inference procedures to generate a dated consensus phylogeny for the lineages of interest (for details see Supplemental Material). \hfill\break

Using these data, we conducted a series of analyses to investigate whether patterns of sexual size dimorphism (SSD) and sexual shape dimorphism (SShD) in this group corresponded to expectations under Rensch's rule. First, we examined patterns of SSD using a phylogenetic regression of male size against female size. Here, a slope significantly greater than 1.0 ($\beta>1.0$) was treated as evidence consistent with Rensch's rule. Additionally, we evaluated SSD patterns using the approach developed here, via a phylogenetic regression of the sexual size dimorphism index against mean body size for each species. Second, we evaluated multivariate patterns of sexual dimorphism in head proportions in relation to body size, using the sexual dimorphism index obtained from the Euclidean distance between male and female means from each species ($D_{SD}$). Partial least squares was used to determine which species displayed female scores (on $PLS_1$) larger than their corresponding male scores, and the sexual dimorphism index was then obtained by multiplying the sexual dimorphism distance by +1 or -1, respectively (i.e., $SDDI=\pm1*D_{SD}$). We then performed a phylogenetic regression of $SDDI$ for head proportions versus mean body size to determine whether patterns of dimorphism corresponded to what is expected under Rensch's rule. Third, we obtained the degree of sexual dimorphism in head shape for each species using the Procrustes distance between male and female means, and from this calculated the sexual shape dimorphism index ($SShDI$) in their original units as described above (i.e., $SShDI=D_{SD}$). We then performed a phylogenetic regression of sexual shape dimorphism ($SShDI$) against the mean body size for each species to determine whether patterns of head shape dimorphism were consistent with predictions from Rensch's rule. Additionally, patterns of head shape dimorphism were further investigated using principal components analysis of the set of mean male and female head shapes across species. Finally, we used phylogenetic correlation [phylogenetic partial least squares: sensu Adams and Felice -@AdamsFelice2014] to determine whether the degree of head shape dimorphism was associated with levels of size dimorphism. All analyses were performed in R 3.6.2 [@R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018], and routines written by the authors. \hfill\break

*Empirical Results*: Using phylogenetic regression, we found a significant relationship between male size and female size, with a slope significantly greater than one ($\beta = 1.39$, $T=2.91$, $P = 0.009$: Fig. 3C). This indicated that the degree of sexual size dimorphism differed systematically with overall body size; a pattern consistent with that expected under Rensch's rule. In this case, both hypoallometry and hyperallometry [*sensu* @Abouheif1997] were displayed (Fig. 3C). Specifically, in male-biased species, the degree of sexual dimorphism was greater with increasing body size, and in female-biased species the degree of sexual dimorphism increased in smaller species (though the pattern was more pronounced in male-biased species). Similarly, when using the sexual size dimorphism index (SSDI), we found a significant and positive evolutionary relationship between SSDI and mean body size, confirming the Rensch's rule pattern identified using the commonly-used procedure ($\beta = 0.43$, $F=22.95$, $Z=1.89$, $P = 0.004$: Fig. 3D).  \hfill\break

When multivariate patterns of sexual dimorphism were evaluated, we found that the degree of dimorphism in head proportions increased as body size increased: a pattern consistent with Rensch's rule ($F=7.20$, $Z=1.41$, $P = 0.022$: Fig. 3E). Likewise, when patterns of head shape dimorphism were examined, we found that the degree of sexual shape dimorphism (SShDI) increased in larger species, and a phylogenetic regression of sexual shape dimorphism versus mean body size was found to be significant ($F=6.05$, $Z=1.35$, $P = 0.03$: Fig. 3F). Thus, for both multivariate datasets, the degree of sexual dimorphism was enhanced in larger species [hyperallometry; sensu @Abouheif1997]: a pattern consistent with what was expected under Rensch's rule. To our knowledge, this is the first empirical demonstration of a pattern following Rensch's rule in multivariate data.  \hfill\break

Visualizing patterns of head shape dimorphism using principal components analysis (Fig. 4), it was evident that males and females occupied distinct regions of morphospace. Further, the head shapes of males and females of the smallest lineage, *Lacerta agilis bosnica*, were distinct from those of the largest species, *Timon lepidus*. When connected, the sexual shape dimorphism vectors for these two species did not coincide with one another in the ordination, implying a difference in sexual shape dimorphism across species. Specifically, the vector for *T. lepidus* was longer than that for *L. agilis bosnica*, which reflected the greater degree of sexual shape dimorphism exhibited by this species. Thin-plate spline deformation grids facilitated a graphical description of these shape differences, where sexual shape dimorphism in *L. agilis bosnica* was best described as a contrast between longer-headed males, with an amplified tympanic area; and short-headed females, with a relatively more reduced tympanic region (Fig. 4, left). By contrast, sexual shape dimorphism in *T. lepidus* was both more prominent in the tympanic area, similarly to what was observed in *L. agilis bosnica*; and also included a modification of the relative lower jaw area. Indeed, males of *T. lepidus* were characterized by both a longer posterior head region and by deeper lower jaw configurations as compared to females of the same species.  \hfill\break

Finally, across taxa, the degree of body size and head shape dimorphism were significantly correlated with one another ($r=0.73$, $p=0.0001$), but not when evaluated in a phylogenetic context ($r_{PPLS} = 0.361$, $Z=1.28$, $P = 0.12$). Part of this difference may be explained by the fact that the largest species (i.e. members of the genus *Timon*) - which displayed the greatest degree of both size and shape dimorphism - form a clade within the broader phylogeny for the group (Fig. 5). Thus, despite the strong association of size and shape dimorphism, there were fewer changes of these patterns across branches of the phylogeny. Generally speaking, the degree of size dimorphism and the degree of shape dimorphism increased in larger taxa, and heat maps of the evolution of size and shape dimorphism across the phylogeny indicated a strong association between the two; particularly in large taxa (Fig. 5). Thus, in this system, it appeared that evolutionary changes in size between the sexes were accompanied by concomitant changes in head shape.

# Discussion

A longstanding question in evolutionary biology is understanding why sexual dimorphism, the degree of differentiation between the sexes, varies across species. A frequently used schema for exploring macroevolutionary patterns of sexual dimorphism is Rensch's rule, which predicts that sexual dimorphism should increase with body size in species where males are the larger sex, and decrease with increasing body size in species where females are the larger sex. Remarkably, virtually all recent studies have focused on macroevolutionary variation in sexual *size* dimorphism, despite the relevance of sexual dimorphism in other traits, as encompassed by Rensch's original treatment [@Rensch1950]. This gap has been driven, in part, by the lack of adequate tools for evaluating Rensch's rule in multivariate traits, such as color, ornament complexity, or shape. In this article, we fill this procedural gap by providing an analytical equivalency that allows the generalization of statistical tests of Rensch's rule for both univariate and multivariate traits. The approach describes the degree of sexual dimorphism as a distance between male and female trait values in phenotype space, thereby facilitating its use on either single-valued traits (e.g., body size), or complex, multivariate traits (e.g. organismal shape). Computer simulations confirm that the approach exhibits appropriate type I error rates and statistical power, thereby providing an adequate procedure for testing whether empirical data comply with the predictions of Rensch's rule. The implementation of our approach on an empirical study of Mediterranean lizards revealed that patterns of sexual dimorphism comply with Rensch's rule not only for body size, but also for relative head proportions and head shape, offering new insights and triggering novel hypotheses about the possible proximate and evolutionary causes that may underlie such patterns. \hfill\break

Our empirical example in Mediterranean green lizards exemplifies a case study investigating variation in sexual dimorphism -- not only in size -- but also in multivariate phenotypes, like shape, and how this method can provide useful insights into the proximate and evolutionary causes underlying such patterns. First, we found that the degree of sexual size dimorphism increased as taxa became larger (Fig. 3C, D). This pattern is compliant with Rensch's rule and corresponds with hyperallometry [@Abouheif1997], which is to be expected when males are the larger sex [@Fairbairn1997]. In this group males are larger than females in most lineages (Fig. 3C), as is commonly observed in lacertids [@Brana1996]. Additionally, as in many lizard species, sexual selection in this group is primarily mediated through territoriality, where male-male competition selects for a larger male body size [@Stamps1983]. Thus, in this system there is an association between male-biased sexual size dimorphism and male-male competition. As such, the patterns of size dimorphism observed in this study are consistent with those expected under a scenario where sexual selection drives the evolution of larger male body size across taxa via male-male competition. \hfill\break

Likewise, our multivariate analyses revealed that head morphology differed between the sexes, and that the degree of sexual dimorphism in multivariate head shape also corresponded to that expected under Rensch's rule (Fig. 3E, 3F). That is, while both small (e.g. *L. a. bosnica*, Fig. 4, left) and large (e.g. *T. lepidus*, Fig. 4, right) species are sexually dimorphic in head shape, the degree of sexual shape dimorphism increased with average body size (Fig. 3F); a pattern consistent with a scenario of hyperallometry. Examination of the resulting shape differences revealed that sexual dimorphism in head shape was associated with an amplification of the posterior head region, and also in the relative size of the lower jaw (in the case of *T. lepidus*). Biomechanically, these are precisely the anatomical regions where the jaw adductor muscles reside [@Haas1973], and as such, an amplification of these areas in males would enhance biting performance [@Herrel1996; @Kaliontzopoulou2012]; a functional trait that has important implications for male competitive capacity [@Lailvaux2007; @HUYGHE2005]. In addition, one might predict that the need for stronger jaw musculature in males becomes disproportionately more intense with increasing body size, as male combats are expected to be more fierce among larger-bodied opponents. Thus, our findings that head shape dimorphism increases with overall body size is also consistent with the hypothesis that sexual selection on structures related to territoriality and male competitive interactions are responsible for the macroevolutionary patterns of sexual shape dimorphism observed in this group. To our knowledge, this is the first example of Rensch's rule in a multivariate phenotype. \hfill\break

With respect to the analytical approach proposed here, our characterization of sexual dimorphism via a *shape distance* and the $SDDI$ statistic is completely general, and may be used to evaluate evolutionary trends in any continuous, quantitative trait that differs between the sexes. Our empirical example highlighted the efficacy of the approach for evaluating trends in univariate traits such as body size, and also several types of complex data where organismal phenotypes are quantified by multiple dimensions. As such, the approach developed here expands our current analytical toolkit to provide a means of evaluating evolutionary trends in sexual dimorphism for a broader class of phenotypic traits than has previously been considered. Indeed, this advance more closely embodies Rensch's [-@Rensch1950] original conceptualization for understanding patterns of sexual dimorphism, as his broader vision was to explain why phenotypic differences *writ large* exist among the sexes. Thus, use of the $SDDI$ will facilitate a greater degree of evolutionary exploration into patterns of sexual dimorphism than has been possible in recent decades. Further, with the inclusion of multivariate traits, the formulation proposed here opens the door for investigations into the evolution of sexual shape dimorphism, and whether such patterns correspond to those expected under Rensch's rule. Thus, the mathematical test proposed here is general, and may be used for any quantitative phenotypic trait under scrutiny. \hfill\break

Finally, because $SDDI$ is derived from a phenotypic distance, one could, in theory, utilize the approach to characterize patterns of sexual dimorphism in non-continuous traits, and determine quantitatively whether the degree of sexual dimorphism is enhanced with increasing body size. Across the tree of life, major phenotypic adaptations are often described as evolutionary innovations or novelties [sensu @Wagner1996a; @Wagner2002; @Peterson2016], where these changes are characterized as discrete shifts from one phenotypic state to another [e.g., @Holliday2012; @OKeefe2011]. In like manner, discrete phenotypic differences are known to exist between the sexes; for instance where males display horns or other ornamentation [@Geist1988; @Emlen2008] that females lack. Over evolutionary time, a series of such discrete phenotypic differences may accumulate in males relative to females, resulting in an elaboration of phenotypic differences between the sexes in suites of such traits. With an appropriate distance measure for discrete traits (e.g., Hamming distance), the approach developed here is capable of characterizing such patterns, and determining whether sets of discrete phenotypic differences are enhanced with increasing body size among species. Thus, our approach is the first to our knowledge to facilitate a formal evaluation of both discrete and continuous phenotypic differences between the sexes, providing a comprehensive means of determining whether sets of phenotypic traits differ in a manner as predicted by Rensch's rule.

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>

\newpage

# Figure Legends

\textbf{Figure 1}. (A) Sexual size dimorphism for males and females across 100 hypothetical species. The pattern displayed exhibits a slope > 1.0, and thus corresponds to Rensch's rule. (B) The same data represented as the ratio of $M/F$ sizes plotted against average species size. (C) Sexual dimorphism represented as the distance between males and females: $SDD = D_{M:F}$. (D) Sexual dimorphism represented as an index of the SD Distance: $SDDI = \pm1*D_{M:F}$. In all panels,  the dashed line corresponds to values representing no sexual dimorphism (the 1:1 line).  \hfill\break

\textbf{Figure 2}. Results of phylogenetic simulations testing the type I error and statistical power for detecting patterns consistent with Rensch's rule. Vertical bars represent the proportion of datasets found to be significant. For type I error analyses, the dashed line marks the standard threshold of $\alpha= 0.05$. For univariate data, both phylogenetic regression of male versus female size (standard approach), and $SDDI$ versus overall body size (new approach) were performed. For multivariate data, only $SDDI$ versus overall body size was examined. \hfill\break

\textbf{Figure 3}. (A) Set of linear measurements used to obtain head proportions. HL: head length, HW: head width, HH: head height, PL: pileus length, MO: mouth opening. (B) Locations of 17 landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head shape. (C) Patterns of sexual size dimorphism represented as male versus female size. (D) Patterns of sexual size dimorphism, represented as sexual size dimorphism index ($SSDI$) versus mean size. (E) Multivariate patterns sexual dimorphism in body proportions ($SShDI$) versus mean size. (F) Multivariate patterns of sexual shape dimorphism ($SShDI$) versus mean size. In panels C, D, E, and F, the red line indicates the regression estimate obtained from the data, while the dashed line corresponds to values representing no sexual dimorphism.  \hfill\break

\textbf{Figure 4}. Principal components plot of male (black squares) and female (red points) species mean head shape: the first two PC axes explain over 60% of the total shape variation. Green lines correspond to the sexual shape dimorphism in one of the smallest species, *Lacerta agilis bosnica*, and in the largest species, *Timon lepidus*. Thin-plate spline deformation grids display the mean head shape for males and females of these species.  \hfill\break

 \textbf{Figure 5}. Heat maps (green: lower values; red: higher values) displaying the evolution of (A) sexual size dimorphism, and (B) sexual shape dimorphism across the phylogeny. For each panel, an index of sexual size dimorphism ($SSDI$) or sexual shape dimorphism ($SShDI$) is displayed, with circles being proportional to the degree of sexual dimorphism. \hfill\break

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\singlespacing \textbf{Figure 1}. (A) Sexual size dimorphism for males and females across 100 hypothetical species. The pattern displayed exhibits a slope > 1.0, and thus corresponds to Rensch's rule. (B) The same data represented as the ratio of $M/F$ sizes plotted against average species size. (C) Sexual dimorphism represented as the distance between males and females: $SDD = D_{M:F}$. (D) Sexual dimorphism represented as an index of the SD Distance: $SDDI = \pm1*D_{M:F}$. In all panels,  the dashed line corresponds to values representing no sexual dimorphism (the 1:1 line).  \hfill\break

\newpage

```{r, echo = FALSE, out.width="70%"}
include_graphics("Fig2.png")

```

\singlespacing \textbf{Figure 2}. Results of phylogenetic simulations testing the type I error and statistical power for detecting patterns consistent with Rensch's rule. Vertical bars represent the proportion of datasets found to be significant. For type I error analyses, the dashed line marks the standard threshold of $\alpha= 0.05$. For univariate data, both phylogenetic regression of male versus female size (standard approach), and $SDDI$ versus overall body size (new approach) were performed. For multivariate data, only $SDDI$ versus overall body size was examined. \hfill\break

\newpage
```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig3.png")

```

\singlespacing \textbf{Figure 3}. (A) Set of linear measurements used to obtain head proportions. HL: head length, HW: head width, HH: head height, PL: pileus length, MO: mouth opening. (B) Locations of 17 landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head shape. (C) Patterns of sexual size dimorphism represented as male versus female size. (D) Patterns of sexual size dimorphism, represented as sexual size dimorphism index ($SSDI$) versus mean size. (E) Multivariate patterns sexual dimorphism in body proportions ($SShDI$) versus mean size. (F) Multivariate patterns of sexual shape dimorphism ($SShDI$) versus mean size. In panels C, D, E, and F, the red line indicates the regression estimate obtained from the data, while the dashed line corresponds to values representing no sexual dimorphism. \hfill\break

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig4.png")

```

\singlespacing \textbf{Figure 4}. Principal components plot of male (black squares) and female (red points) species mean head shape: the first two PC axes explain over 60% of the total shape variation. Green lines correspond to the sexual shape dimorphism in one of the smallest species, *Lacerta agilis bosnica*, and in the largest species, *Timon lepidus*. Thin-plate spline deformation grids display the mean head shape for males and females of these species.  \hfill\break

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig5.png")

```

\singlespacing \textbf{Figure 5}. Heat maps (green: lower values; red: higher values) displaying the evolution of (A) sexual size dimorphism, and (B) sexual shape dimorphism across the phylogeny. For each panel, an index of sexual size dimorphism ($SSDI$) or sexual shape dimorphism ($SShDI$) is displayed, with circles being proportional to the degree of sexual dimorphism. \hfill\break

### 2020-03-15 CoverPage-Evolution_EKB.Rmd
knit attempt - failed

--- 2020-03-15 CoverPage-Evolution.rmd
@@
 ---
-title: A MULTIVARIATE EQUIVALENCY AND GENERAL TEST FOR RENSCH'S RULE: IMPLICATIONS FOR THE EVOLUTION OF SEXUAL SHAPE DIMORPHISM
+title: MISUSES AND INACCURACY PATTERNS OF PAGEL'S LAMBDA ML ESTIMATION
 
@@
 \begin{center}
-\textbf{Dean C. Adams$^{1,*}$, Elizabeth Glynne$^{1}$ and Antigoni Kaliontzopoulou$^2$}
+\textbf{Dean C. Adams$^{1,*}$ and Erica K. Baken$^{1}$}
 \end{center}
@@
 
-$^{2}$CIBIO Research Centre in Biodiversity and Genetic Resources, InBIO, Universidade do Porto, Campus de Vairão. Vila do Conde, Portugal.
-
 $^{*}$Correspondence: Dean C. Adams <dcadams@iastate.edu>
@@
 \hfill\break
-**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break
+**Keywords**: Pagel's lambda, phylogenetic signal
+\hfill\break
 
-**Short Title**: Multivariate Test of Rensch's Rule \hfill\break
+**Short Title**: Problems with Pagel's lambda \hfill\break
 
-**Author Contributions**: DCA devised the approach, AK provided the empirical example; all authors contributed  to all parts of this manuscript. \hfill\break
+**Author Contributions**: DCA ...EKB...; all authors contributed  to all parts of this manuscript. \hfill\break
 
@@
 
-**Acknowledgments**: We thank XXX for comments on drafts of this manuscript; Ana Perera, Inés Freitas and Daniele Salvi for the phylogeny reconstruction; and Chiara Ripa and Zélia Pereira for aiding shape data acquisition.
-This work was  sponsored in part by National Science Foundation Grant DBI-1902511 (to DCA), and by FEDER Funds through the Operational Competitiveness Factors Program - COMPETE and National Funds through FCT - Foundation for Science and Technology within the scope of the project “PTDC/BIA-EVL/28090/2017 - POCI-01-0145-FEDER-028090” (to AK).
+**Acknowledgments**: We thank XXX for comments on drafts of this manuscript; ...... This work was  sponsored in part by National Science Foundation Grant DBI-1902511 (to DCA).

### 2020-03-15 Manuscript-template_EKB.tex
knit attempt - failed

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers
% https://github.com/rstudio/rmarkdown/issues/337
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

% https://github.com/rstudio/rmarkdown/pull/252
\usepackage{titling}
\setlength{\droptitle}{-2em}

\pretitle{\vspace{\droptitle}\centering\huge}
\posttitle{\par}

\preauthor{\centering\large\emph}
\postauthor{\par}

\predate{\centering\large\emph}
\postdate{\par}

\date{}

\begin{document}

\section{A multivariate equivalency and general test for Rensch's Rule:
Implications for the evolution of sexual shape
dimorphism}\label{a-multivariate-equivalency-and-general-test-for-renschs-rule-implications-for-the-evolution-of-sexual-shape-dimorphism}

\hfill\break

\textbf{Keywords}: Pagel's lambda, phylogenetic signal \hfill\break

\textbf{Short Title}: Problems with Pagel's lambda \hfill\break

\section{Abstract}\label{abstract}

Rensch's rule, the allometric association between the degree of sexual
dimorphism and overall body size, is an important framework for
understanding the evolution of trends in phenotypic differentiation
between the sexes across taxa. While virtually all recent treatments
have explored the degree of sexual size dimorphism, understanding how
dimorphism in other phenotypic traits evolves is equally relevant.
Indeed, Rensch's original treatment focused not only on size, but also
on dimorphism in relative body proportions, coloration, the presence of
ornaments, etc. Here, we derive a multivariate equivalency for viewing
trends in sexual dimorphism -- relative to overall body size -- across
taxa, and provide a generalized test to determine whether such patterns
are consistent with Rensch's rule. For univariate linear traits such as
body size, our approach yields equivalent results to those from standard
procedures, but our test is also capable of detecting trends in
multivariate datasets like shape. Computer simulations reveal the method
displays appropriate statistical properties, and an empirical example in
Mediterranean lizards illustrates the efficacy of our approach on both
univariate and multivariate phenotypes. Our generalized procedure
substantially extends the toolkit available for investigating
macroevolutionary patterns of sexual dimorphism and seeking a better
understanding of the processes that underlie them.

\newpage

\section{Introduction}\label{introduction}

It is widely observed that males and females of the same species are not
identical, but rather differ -- sometimes substantially -- in their
phenotypic characteristics (Fairbairn 1997, 2013). Indeed, sexual
dimorphism is pervasive across many animal clades, and understanding the
causes for these patterns has attracted considerable attention in
evolutionary biology for well over a century (e.g., Darwin 1871;
Fairbairn et al. 2007). One of the most conspicuous differences between
males and females is in their body sizes (Andersson 1994), where
sex-specific differences are often hypothesized to be the result of
selection pressures related to reproduction, or the distinct ecological
roles of males and females in their respective habitat (Cox et al. 2003;
Stephens and Wiens 2009; Kaliontzopoulou et al. 2015; Tarr et al. 2018;
Littleford-Colquhoun et al. 2019). For example, both sexual selection on
male body size (Cox et al. 2003; Garcia-Navas et al. 2015; Horne et al.
2020), and fecundity selection on female body size (Serrano-Meneses and
Szekely 2006; Stuart-Fox 2008), are expected to enhance the degree of
sexual size dimorphism in taxa over evolutionary time. Likewise, when
males and females respond differently to environmental factors,
intraspecific competition may generate ecological divergence between
them, resulting in varying patterns of sexual dimorphism (Butler et al.
2000; Bolnick and Doebeli 2003; Dayan and Simberloff 2005;
Kaliontzopoulou et al. 2010; Meiri et al. 2014). \hfill\break

At macroevolutionary scales, changes in the relative contribution of the
selective forces above may create variation in the degree of sexual
dimorphism displayed across closely related taxa. For instance, sister
species inhabiting different environments, or possessing distinct
ecological adaptations, may exhibit differing degrees of sexual
dimorphism, as a result of the varying intensity of sexual selection
they endure (Östman and Stuart-Fox 2011). Likewise, the interplay
between sexual and natural selection that species experience may result
in sexual dimorphism of differing magnitudes across taxa (Blanckenhorn
et al. 2006; Stuart-Fox and Moussalli 2007; Kaliontzopoulou et al.
2015). One evolutionary framework that connects adaptive explanations
for the evolution of sexual size dimorphism (SSD) with trends in overall
body size across taxa is Rensch's rule. Here it is predicted that SSD
will increase with species' average body size in species where males are
the larger sex (hyperallometry), and likewise will decrease with
increasing body size where females are the larger sex (hypoallometry)
(Abouheif and Fairbairn 1997; Fairbairn 1997). Put another way, Rensch's
rule predicts that the extent of sexual dimorphism increases at more
extreme body sizes (both large and small) across taxa. Empirical studies
have documented trends in SSD consistent with Rensch's rule in a wide
variety of animal taxa (Dale et al. 2007; Ceballos et al. 2013; Regis
and Meik 2017; Johnson et al. 2017). However, other clades display the
converse trend (Burbrink and Futterman 2019; Peñalver-Alcázar et al.
2019), and some clades display no relationship between SSD and body size
(Astúa 2010; Hirst and Kiørboe 2014; Johnson et al. 2017). From an
evolutionary perspective, the association between SSD and body size has
been hypothesized to be driven by variation in ontogenetic or static
allometries across taxa, which results from the influence of sexual
selection on growth and maturation patterns (Rensch 1960; Dale et al.
2007). \hfill\break

Interestingly, virtually all evaluations of Rensch's rule have
investigated patterns of sexual size dimorphism, yet Rensch's original
contributions were concerned with a much wider array of phenotypic
traits that can differ between the sexes. These included not only sexual
differences in body size, but also in general body proportions (e.g.,
relative wing and limb lengths), the degree of complexity of different
body structures, the presence of sex-specific ornamentation (e.g.,
horns, body coloration, etc.), and other sex-specific differences
(Rensch 1950, 1960). In like manner, some studies have documented
patterns of sexual shape dimorphism (e.g., Berns and Adams 2013; Kelly
et al. 2013; Kaliontzopoulou et al. 2015), sex-specific differences in
ornamentation (Geist and Bayer 1988; Emlen 2008; Watson and Simmons
2010), or coloration (e.g., Endler 1983). However, such studies have
typically been restricted to one or a few taxa, and have not explained
macroevolutionary variation in sexual dimorphism in a manner that
relates directly to Rensch's rule. Thus, when viewed from this
perspective, Rensch's (1950) vision was far more synthetic, as he sought
to explain the extent to which sexual dimorphism -- across a general
suite of phenotypic characteristics -- scaled with overall body size in
organisms, and why some species displayed a greater degree of sexual
dimorphism than did others. Remarkably, patterns of intraspecific static
allometry in sexually selected traits establishes a direct link between
variation in those traits relative to size, and the evolutionary
allometry across taxa in the degree of sexual dimorphism that may be
expected (see also Rensch 1960; Reiss 1986; Bonduriansky 2007). Under
this framework, the ``exaggeration'' of body structures or other
phenotypic traits with increasing body size is predicted to translate,
at macroevolutionary scales, into increased degrees of sexual dimorphism
with increasing body size. However, despite this clear prediction, and
Rensch's original observations, the evolutionary scaling of the degree
of sexual dimorphism in phenotypic traits other than size, remains
largely unexplored. \hfill\break

Another current limitation in our understanding of Rensch's rule is the
fact that to date, all studies have been conducted exclusively on
univariate data, primarily regarding body size. However, understanding
phenotypic evolution is inherently a multivariate endeavor (Blows 2007;
Collyer et al. 2015; Adams and Collyer 2018, 2019), as processes such as
natural and sexual selection simultaneously operate on more than one
trait (Lande 1979; Lande and Arnold 1983), and can affect variation in
complex multidimensional traits such as shape. To interrogate such
patterns, the analytics used to evaluate evolutionary trends in
phenotypic datasets must be sufficiently general to accommodate this
empirical fact. Unfortunately, all current analytical approaches for
evaluating trends in sexual dimorphism as they relate to overall body
size are explicitly univariate; thereby excluding the possibility of
evaluating patterns in complex multidimensional traits such as
organismal shape. Though is has been extensively documented in sexual
size dimorphism, it remains unknown whether Rensch's rule applies
equally to sexual shape dimorphism, as determining this requires an
analytical framework capable of evaluating trends in sexual dimorphism
in both univariate and multivariate datasets. \hfill\break

In this paper, we derive a multivariate equivalency for inspecting
variation in sexual dimorphism -- relative to overall body size --
across taxa, and provide a generalized test to determine whether they
are consistent with Rensch's rule. Our procedure is appropriate for
evaluating trends in univariate traits measured on a linear scale (such
as body size or relative limb length), as well as in multivariate
datasets representing complex phenotypes (such as sets of body
proportions or landmark-based shape data). For univariate data, the
approach yields equivalent results to what is normally accomplished for
size-based SSD evaluations. However, the approach extends these
procedures for the examination of sexual dimorphism in multivariate
data, thereby increasing the potential for our conceptual understanding
of the consequences of sexual and natural selection on complex
organismal phenotypes. We conduct a series of computer simulations
demonstrating that the approach displays appropriate statistical
properties when evaluating trends in both univariate and multivariate
data. We also provide an empirical example in Mediterranean green
lizards which illustrates the efficacy of our approach on both
univariate and multivariate datasets. Finally, we highlight the new
avenues of evolutionary research that our generalization facilitates,
and make comments on the prospects for future empirical advances.

\section{Methods and Results}\label{methods-and-results}

\subsection{\texorpdfstring{\emph{Conceptual
Development}}{Conceptual Development}}\label{conceptual-development}

To arrive at a general test for Rensch's rule that can accommodate
multivariate data, we must first derive a series of mathematical
equivalencies based upon how interspecific patterns of size dimorphism
are typically evaluated. For sexual size dimorphism, interspecific
patterns are often inspected using a bivariate plot, where the size of
one sex is plotted against the size of the other sex (e.g., Fairbairn
and Preziosi 1994; Abouheif and Fairbairn 1997; Ceballos et al. 2013).
An idealized example is shown in Fig. 1A, with male size plotted against
female size. The diagonal line represents no size dimorphism; or a 1:1
size ratio between the sexes across the range of body sizes for species
in the clade of interest. In this construction (i.e., with females
plotted on the x-axis), deviations from this line represent differential
trends in size dimorphism across taxa. To evaluate such patterns
statistically, a phylogenetic regression is performed, and in this case,
a slope significantly greater than 1.0 (i.e., \(\beta_{1}>1.0\)) would
be treated as evidence that the evolutionary changes in sexual size
dimorphism across taxa are consistent with Rensch's rule (Fairbairn
1997). \hfill\break

An equivalent representation of patterns of sexual dimorphism may also
be found using ratios (e.g., male size:female size). When plotted
against the average size for each species (Fig. 1B), the same
information is conveyed as is found in the more commonly utilized male -
female plot of Fig. 1A. However, in this representation, the line
designating no sexual dimorphism is horizontal, with a slope of zero
(i.e., \(\beta_{1}=0.0\)) and an intercept of 1.0. As before, a
phylogenetic regression may be performed on these data, and in this
case, if the slope is significantly steeper than zero (i.e.,
\(\beta_{1}>0.0\)), patterns consistent with Rensch's rule are observed.
\hfill\break

A third equivalency is shown in Fig. 1C. Here, the Euclidean distance
between male and female means for each species
(\(D_{SD}=\sqrt{(Y_{male}-Y_{female})^2}\)) is shown relative to the
overall size of the species. This plot is similar to a Bland-Altmann
plot (Altman and Bland 1983), except that the Euclidean distance, rather
than a difference score, is plotted along the y-axis (as distances can
accommodate data of different dimensionality). Under Rensch's rule, this
\emph{sexual dimorphism distance} may become greater as species are
progressively larger (hyperallometry, sensu Abouheif and Fairbairn 1997)
or may become greater as species are progressively smaller
(hypoallometry, sensu Abouheif and Fairbairn 1997). In both cases,
patterns correspond to species displaying progressively more sexual
dimorphism at extreme body sizes; a trend consistent with Rensch's rule.
Note that the inflection point in Fig. 1C represents an intermediate
size where there is no sexual dimorphism for the trait, and corresponds
to where the male-female regression line of Fig. 1A crosses the line
representing no sexual dimorphism. This inflection point may represent
the overall mean size across species in the lineage, or may be some
other value. For instance, in some lineages (e.g., turtles: Ceballos et
al. 2013), nearly all species are female biased, in which case the
inflection point for \(D_{SD}\) vs.~size would be found at the far-right
end of the plot. \hfill\break

Finally, from \(D_{SD}\) a \emph{sexual dimorphism distance index}
(\(SDDI\)) may be constructed as:

\[SDDI=-1*D_{SD}\] \[SDDI=+1*D_{SD}\]

Here, \(D_{SD}\) is multiplied by either a +1 or -1, based on a
comparison of male and female trait values for the species. For
instance, when single-valued traits on a linear scale are examined
(e.g., leg length or wing extent), \(D_{SD}\) is multiplied by -1 for
those species where females display larger values than males, and
remains unchanged for species where males are larger than females. This
logic is analogous to that of Lovich and Gibbons (1992) for their index
of overall body size dimorphism. However, from a geometric perspective,
this procedure is tantamount to determining which species in Fig. 1A are
found below the 1:1 size line (i.e., the line representing no size
dimorphism), and multiplying those species' \(D_{SD}\) by -1.
\hfill\break

Unfortunately, applying the same logic to multidimensional trait data is
more complicated, as sex-specific differences are not always guaranteed
to align with some \emph{a priori} direction in the multivariate
dataspace. Thus, for multivariate datasets, one of two procedures may be
used. First, for multivariate sets of traits that are measured on the
same linear scale (e.g., length, width, and height measures), one could
envision a hyperdimensional version of the plot displayed in Fig. 1A,
but where the set of male traits are plotted against the set of female
traits. In this construction, because the traits are quantified on the
same measured scale, the 1:1 size vector representing no sexual
dimorphism does exist, and emanates from the origin of the space
extending in a positive direction as the trait values increase. To
determine whether male trait values deviate more from this vector than
do female trait values, a partial least squares analysis may be
performed (Rohlf and Corti 2000). Partial least squares (PLS) is a
multivariate association procedure which describes the covariation
between sets of traits (in this case, between the set of male traits and
the set of female traits). Scores along the first PLS axis represent the
maximal covariation between male and female traits projected to a single
dimension, and species whose male and female PLS scores are identical
are those displaying no sexual dimorphism. Thus, deviations from that
pattern represent instances of multivariate sexual dimorphism. By
plotting these scores, one may determine which species display female
scores (on \(PLS_1\)) that are larger than their corresponding male
scores (when negative values on \(PLS_1\) correspond to smaller
individuals; otherwise, the converse comparison is used). The \(D_{SD}\)
for these species may then be multiplied by -1 to obtain the \(SDDI\)
for that species. Note that when PLS is used on univariate data, this
procedure obtains values identical to those from the method described
previously where male and female sizes are compared directly. Thus, for
multivariate traits measured on a linear scale, this is a direct
generalization of the standard univariate procedure (see empirical
example below). \hfill\break

Alternatively, one might be interested in examining multidimensional
traits other than linear biometric measurements, such as color or shape.
For instance, one may characterize the shape of males and females using
geometric morphometric methods (Bookstein 1991; Mitteroecker and Gunz
2009; Adams et al. 2013), where the locations of anatomical landmarks
and semilandmarks are used to obtain a set of shape variables. Geometric
morphometric methods also result in a multidimensional characterization
of phenotypic traits, from which the degree of sexual dimorphism for
each species may be quantified using \(D_{SD}\). However, because the
resulting shape variables do not have a one to one correspondence with
linear size measures, there is no \emph{a priori} direction in the
multivariate shape space that is equivalent to the 1:1 size vector found
with sets of linear traits. In fact, for the case of isometry (i.e.,
shape does not change as size changes), all observations are represented
by a single location in the shape space. Thus, while a PLS of male
shapes versus female shapes will describe the covariation between them,
this direction is not guaranteed to align with size, because there is no
direction in the shape space that may be used to unambiguously determine
whether dimorphism is male- or female-biased, as for shape variables
there is no \emph{a priori} directionality. Nonetheless, one may still
quantitatively determine whether the degree of sexual shape dimorphism
covaries with overall body size, by using the sexual dimorphism index as
defined by the dimorphism distance (i.e., \(SDDI=D_{SD}\)) to represent
the degree of multivariate shape dimorphism for each species.
\hfill\break

Once the sexual dimorphism distance index (\(SDDI\)) is calculated, it
may be evaluated against the average size of each species. Here, data
conforming to Rensch's rule will result in significant trends of sexual
dimorphism relative to mean body size (Fig. 1D). As with size ratios,
the line of no sexual dimorphism is a horizontal line, but this time it
is centered on zero, as a distance of zero corresponds to no sexual
dimorphism. Thus, a phylogenetic regression of these data whose slope is
significantly different from zero (e.g., \(\beta_{1}\neq0.0\)) provides
evidence that the evolution of sexual dimorphism is consistent with
patterns expected under Rensch's rule. For cases where traits are
measured on a linear scale, patterns expected under Rensch's rule
exhibit a slope significantly greater than zero (\(\beta_{1}>0.0\)), as
in the usual formulation. Likewise, when using geometric morphometric
shape data (or any other type of mean-centered multidimensional data),
hyperallometric patterns consistent with Rensch's rule would also
display a slope significantly greater than zero (\(\beta_{1}>0.0\)),
indicating that the degree of sexual shape dimorphism increases with
increasing body size. Conversely however, hypoallometric patterns would
display the opposite pattern (\(\beta_{1}<0.0\)), indicating that
smaller species display greater sexual shape dimorphism. \hfill\break

\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}

To ascertain the statistical performance of the approach proposed here,
we conducted a series of computer simulations. We evaluated the ability
of SDDI to characterize patterns of sexual dimorphism in both univariate
and multivariate datasets. For each simulation, we generated a
pure-birth phylogeny containing 50 species, on which we simulated
datasets under a Brownian motion model of evolution. For univariate data
(\(p=1\)), a single trait was evolved along the phylogeny, and
represented the female trait values (\(x\)). Next, male values (\(y\))
were simulated in such a manner as to generate a known male:female
relationship (plus random error), using the function:
\(y=\alpha{x}+\mathcal{N}(\mu=0,\sigma=0.1)\). For type I error
simulations, no sexual dimorphism was generated, meaning that the slope,
\(\alpha\), characterized a 1:1 relationship between male and female
trait values (i.e., \(\alpha=1.0\)). For power simulations, we simulated
male values with a slope greater than 1.0 to correspond with patterns
following Rensch's rule (i.e., \(\alpha=1.1\)). A total of 1,000
datasets were generated under each simulation condition. We then
evaluated all datasets using standard approaches (i.e., a phylogenetic
regression of male values versus female values), as well as a
phylogenetic regression of SDDI versus average body size. The proportion
of significant datasets was treated as the type I error (\(\alpha=1.0\))
or power (\(\alpha=1.1\)) respectively. \hfill\break

For multivariate data (\(p=5\)), a slightly different procedure was
utilized. First, average body size (\(s\)) was evolved along the
phylogeny under Brownian motion. Next, female traits (\(\mathbf{X}\))
were simulated on the phylogeny under Brownian motion, using a
\(p\times{p}\) trait covariance matrix with correlations between traits
of (\(r=0.7\)). For simulations where there was no pattern of sexual
dimorphism across species (i.e., type I error simulations) male trait
values were then generated by incorporating random error to the female
trait values, in a manner analogous to that achieved in the univariate
simulations:
\(\mathbf{Y}=\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)). For
simulations where patterns of sexual dimorphism covaried with body size
(i.e., power simulations), male trait values were generated by
incorporating random error to the female trait values, but with the
addition that female trait values were pre-multiplied by a scalar value
of body size:
\(\mathbf{Y}=(0.1*s)*\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)).
This ensured that male and female trait values covaried with one another
as expected, and also covaried with size in the desired direction. As
before, a total of 1,000 datasets were generated under each simulation
condition. We then evaluated all datasets using a phylogenetic
regression of SDDI versus average body size. The proportion of
significant datasets was treated as the type I error (\(\alpha=1.0\)) or
power (\(\alpha=1.1\)) respectively. \hfill\break

\emph{Simulation Results}: For the univariate dataset, we found that
both a phylogenetic regression of male size on female size, as well as a
phylogenetic regression of \(SDDI\) on body size, resulted in
appropriate type I error rates near the nominal \(\alpha\approx0.05\)
(Fig. 2). Further, when data were simulated under conditions consistent
with Rensch's rule, both procedures were capable of detecting the signal
of the pattern, and did so with very high power. Thus, using the
\(SDDI\) approach proposed here was equivalent to what is normally
utilized for univariate datasets. Likewise, for multivariate data, the
\(SDDI\) approach displayed appropriate type I error
(\(\alpha\approx0.05\)) and high statistical power under the conditions
simulated (Fig. 2). Additionally, multiplying \(D_{SD}\) by \(\pm1\) had
little effect on the outcome for multivariate data, as in both cases
type I error and power results were virtually identical. Thus, methods
for evaluating sexual dimorphism using geometric morphometric data are
anticipated to remain robust. Overall results from these simulations
imply that using the sexual dimorphism distance index (\(SDDI\))
proposed here provides an appropriate means of characterizing patterns
of sexual dimorphism in both univariate and multivariate traits, and
that the procedure is capable of detecting patterns consistent with
Rensch's rule when they are present in phenotypic datasets. \hfill\break

\subsection{\texorpdfstring{\emph{Empirical
Example}}{Empirical Example}}\label{empirical-example}

To illustrate the utility of the approach presented here, we conducted
an analysis of both size and shape dimorphism in a group of
Mediterranean lizards with extensive body size variation. Our dataset
consisted of three data types: male and female values of body size
(snout-vent length: SVL), male and female values for a series of head
measurements (head length, head width, head height, pileus length, and
mouth opening: Fig. 3A), and male and female values of head shape (Fig.
3B). A total of 374 specimens from 21 lineages of green lizards, a
monophyletic clade that includes two genera (\emph{Lacerta} and
\emph{Timon}), were obtained from natural history museum collections.
For each lineage, we obtained samples distributed across the entire
distribution range, resulting in an initial database of 1062 specimens.
In order to take sampling bias into account, we selected the 10 largest
individuals per sex, when sufficient specimens were available, for use
in subsequent size and shape analyses. A detailed account of the
specimens used for this study is available in the Supplemental Material.
\hfill\break

From these specimens, we calculated mean body size for each sex and
lineage. Likewise, we calculated the mean trait values for each sex and
lineage for the set of linear measures. These values were then
standardized by SVL; resulting in a set of head measurements
proportional to body size for each sex \(\times\) lineage combination.
Additionally, we quantified lateral head shape variation using geometric
morphometric methods (Adams et al. 2013). This was accomplished by first
digitizing 17 landmarks and semilandmarks from images of the right side
of the head (Fig. 3B). The thin-plate spline was used to estimate the
position of missing landmarks (Gunz et al. 2009), and the position of
the jaw relative to the skull was standardized using the fixed angle
method (Adams 1999). Next, a generalized Procrustes analysis (Rohlf and
Slice 1990) was performed to eliminate non-shape variation from the
landmark data and align the specimens to a common coordinate system. The
position of semilandmarks (i.e.~11-16, Fig. 3B) was optimized by
minimizing the bending energy (Bookstein 1997). From the aligned
Procrustes coordinates the mean head shape for males and for females was
then calculated for each lineage. Finally, the phylogenetic
relationships among lineages were estimated using published sequences on
five mitochondrial and four nuclear gene fragments, totaling 8383 bp.
These were then used following standard phylogenetic inference
procedures to generate a dated consensus phylogeny for the lineages of
interest (for details see Supplemental Material). \hfill\break

Using these data, we conducted a series of analyses to investigate
whether patterns of sexual size dimorphism (SSD) and sexual shape
dimorphism (SShD) in this group corresponded to expectations under
Rensch's rule. First, we examined patterns of SSD using a phylogenetic
regression of male size against female size. Here, a slope significantly
greater than 1.0 (\(\beta>1.0\)) was treated as evidence consistent with
Rensch's rule. Additionally, we evaluated SSD patterns using the
approach developed here, via a phylogenetic regression of the sexual
size dimorphism index against mean body size for each species. Second,
we evaluated multivariate patterns of sexual dimorphism in head
proportions in relation to body size, using the sexual dimorphism index
obtained from the Euclidean distance between male and female means from
each species (\(D_{SD}\)). Partial least squares was used to determine
which species displayed female scores (on \(PLS_1\)) larger than their
corresponding male scores, and the sexual dimorphism index was then
obtained by multiplying the sexual dimorphism distance by +1 or -1,
respectively (i.e., \(SDDI=\pm1*D_{SD}\)). We then performed a
phylogenetic regression of \(SDDI\) for head proportions versus mean
body size to determine whether patterns of dimorphism corresponded to
what is expected under Rensch's rule. Third, we obtained the degree of
sexual dimorphism in head shape for each species using the Procrustes
distance between male and female means, and from this calculated the
sexual shape dimorphism index (\(SShDI\)) in their original units as
described above (i.e., \(SShDI=D_{SD}\)). We then performed a
phylogenetic regression of sexual shape dimorphism (\(SShDI\)) against
the mean body size for each species to determine whether patterns of
head shape dimorphism were consistent with predictions from Rensch's
rule. Additionally, patterns of head shape dimorphism were further
investigated using principal components analysis of the set of mean male
and female head shapes across species. Finally, we used phylogenetic
correlation (phylogenetic partial least squares: sensu Adams and Felice
2014) to determine whether the degree of head shape dimorphism was
associated with levels of size dimorphism. All analyses were performed
in R 3.6.2 (R Core Team 2019) using the packages \texttt{geomorph}
(Adams and Otárola-Castillo 2013; Adams et al. 2019), \texttt{RRPP}
(Collyer and Adams 2018), and routines written by the authors.
\hfill\break

\emph{Empirical Results}: Using phylogenetic regression, we found a
significant relationship between male size and female size, with a slope
significantly greater than one (\(\beta = 1.39\), \(T=2.91\),
\(P = 0.009\): Fig. 3C). This indicated that the degree of sexual size
dimorphism differed systematically with overall body size; a pattern
consistent with that expected under Rensch's rule. In this case, both
hypoallometry and hyperallometry (\emph{sensu} Abouheif and Fairbairn
1997) were displayed (Fig. 3C). Specifically, in male-biased species,
the degree of sexual dimorphism was greater with increasing body size,
and in female-biased species the degree of sexual dimorphism increased
in smaller species (though the pattern was more pronounced in
male-biased species). Similarly, when using the sexual size dimorphism
index (SSDI), we found a significant and positive evolutionary
relationship between SSDI and mean body size, confirming the Rensch's
rule pattern identified using the commonly-used procedure
(\(\beta = 0.43\), \(F=22.95\), \(Z=1.89\), \(P = 0.004\): Fig. 3D).
\hfill\break

When multivariate patterns of sexual dimorphism were evaluated, we found
that the degree of dimorphism in head proportions increased as body size
increased: a pattern consistent with Rensch's rule (\(F=7.20\),
\(Z=1.41\), \(P = 0.022\): Fig. 3E). Likewise, when patterns of head
shape dimorphism were examined, we found that the degree of sexual shape
dimorphism (SShDI) increased in larger species, and a phylogenetic
regression of sexual shape dimorphism versus mean body size was found to
be significant (\(F=6.05\), \(Z=1.35\), \(P = 0.03\): Fig. 3F). Thus,
for both multivariate datasets, the degree of sexual dimorphism was
enhanced in larger species (hyperallometry; sensu Abouheif and Fairbairn
1997): a pattern consistent with what was expected under Rensch's rule.
To our knowledge, this is the first empirical demonstration of a pattern
following Rensch's rule in multivariate data. \hfill\break

Visualizing patterns of head shape dimorphism using principal components
analysis (Fig. 4), it was evident that males and females occupied
distinct regions of morphospace. Further, the head shapes of males and
females of the smallest lineage, \emph{Lacerta agilis bosnica}, were
distinct from those of the largest species, \emph{Timon lepidus}. When
connected, the sexual shape dimorphism vectors for these two species did
not coincide with one another in the ordination, implying a difference
in sexual shape dimorphism across species. Specifically, the vector for
\emph{T. lepidus} was longer than that for \emph{L. agilis bosnica},
which reflected the greater degree of sexual shape dimorphism exhibited
by this species. Thin-plate spline deformation grids facilitated a
graphical description of these shape differences, where sexual shape
dimorphism in \emph{L. agilis bosnica} was best described as a contrast
between longer-headed males, with an amplified tympanic area; and
short-headed females, with a relatively more reduced tympanic region
(Fig. 4, left). By contrast, sexual shape dimorphism in \emph{T.
lepidus} was both more prominent in the tympanic area, similarly to what
was observed in \emph{L. agilis bosnica}; and also included a
modification of the relative lower jaw area. Indeed, males of \emph{T.
lepidus} were characterized by both a longer posterior head region and
by deeper lower jaw configurations as compared to females of the same
species. \hfill\break

Finally, across taxa, the degree of body size and head shape dimorphism
were significantly correlated with one another (\(r=0.73\),
\(p=0.0001\)), but not when evaluated in a phylogenetic context
(\(r_{PPLS} = 0.361\), \(Z=1.28\), \(P = 0.12\)). Part of this
difference may be explained by the fact that the largest species
(i.e.~members of the genus \emph{Timon}) - which displayed the greatest
degree of both size and shape dimorphism - form a clade within the
broader phylogeny for the group (Fig. 5). Thus, despite the strong
association of size and shape dimorphism, there were fewer changes of
these patterns across branches of the phylogeny. Generally speaking, the
degree of size dimorphism and the degree of shape dimorphism increased
in larger taxa, and heat maps of the evolution of size and shape
dimorphism across the phylogeny indicated a strong association between
the two; particularly in large taxa (Fig. 5). Thus, in this system, it
appeared that evolutionary changes in size between the sexes were
accompanied by concomitant changes in head shape.

\section{Discussion}\label{discussion}

A longstanding question in evolutionary biology is understanding why
sexual dimorphism, the degree of differentiation between the sexes,
varies across species. A frequently used schema for exploring
macroevolutionary patterns of sexual dimorphism is Rensch's rule, which
predicts that sexual dimorphism should increase with body size in
species where males are the larger sex, and decrease with increasing
body size in species where females are the larger sex. Remarkably,
virtually all recent studies have focused on macroevolutionary variation
in sexual \emph{size} dimorphism, despite the relevance of sexual
dimorphism in other traits, as encompassed by Rensch's original
treatment (Rensch 1950). This gap has been driven, in part, by the lack
of adequate tools for evaluating Rensch's rule in multivariate traits,
such as color, ornament complexity, or shape. In this article, we fill
this procedural gap by providing an analytical equivalency that allows
the generalization of statistical tests of Rensch's rule for both
univariate and multivariate traits. The approach describes the degree of
sexual dimorphism as a distance between male and female trait values in
phenotype space, thereby facilitating its use on either single-valued
traits (e.g., body size), or complex, multivariate traits
(e.g.~organismal shape). Computer simulations confirm that the approach
exhibits appropriate type I error rates and statistical power, thereby
providing an adequate procedure for testing whether empirical data
comply with the predictions of Rensch's rule. The implementation of our
approach on an empirical study of Mediterranean lizards revealed that
patterns of sexual dimorphism comply with Rensch's rule not only for
body size, but also for relative head proportions and head shape,
offering new insights and triggering novel hypotheses about the possible
proximate and evolutionary causes that may underlie such patterns.
\hfill\break

Our empirical example in Mediterranean green lizards exemplifies a case
study investigating variation in sexual dimorphism -- not only in size
-- but also in multivariate phenotypes, like shape, and how this method
can provide useful insights into the proximate and evolutionary causes
underlying such patterns. First, we found that the degree of sexual size
dimorphism increased as taxa became larger (Fig. 3C, D). This pattern is
compliant with Rensch's rule and corresponds with hyperallometry
(Abouheif and Fairbairn 1997), which is to be expected when males are
the larger sex (Fairbairn 1997). In this group males are larger than
females in most lineages (Fig. 3C), as is commonly observed in lacertids
(Braña 1996). Additionally, as in many lizard species, sexual selection
in this group is primarily mediated through territoriality, where
male-male competition selects for a larger male body size (Stamps 1983).
Thus, in this system there is an association between male-biased sexual
size dimorphism and male-male competition. As such, the patterns of size
dimorphism observed in this study are consistent with those expected
under a scenario where sexual selection drives the evolution of larger
male body size across taxa via male-male competition. \hfill\break

Likewise, our multivariate analyses revealed that head morphology
differed between the sexes, and that the degree of sexual dimorphism in
multivariate head shape also corresponded to that expected under
Rensch's rule (Fig. 3E, 3F). That is, while both small (e.g. \emph{L. a.
bosnica}, Fig. 4, left) and large (e.g. \emph{T. lepidus}, Fig. 4,
right) species are sexually dimorphic in head shape, the degree of
sexual shape dimorphism increased with average body size (Fig. 3F); a
pattern consistent with a scenario of hyperallometry. Examination of the
resulting shape differences revealed that sexual dimorphism in head
shape was associated with an amplification of the posterior head region,
and also in the relative size of the lower jaw (in the case of \emph{T.
lepidus}). Biomechanically, these are precisely the anatomical regions
where the jaw adductor muscles reside (Haas 1973), and as such, an
amplification of these areas in males would enhance biting performance
(Herrel et al. 1996; Kaliontzopoulou et al. 2012); a functional trait
that has important implications for male competitive capacity (Huyghe et
al. 2005; Lailvaux and Irschick 2007). In addition, one might predict
that the need for stronger jaw musculature in males becomes
disproportionately more intense with increasing body size, as male
combats are expected to be more fierce among larger-bodied opponents.
Thus, our findings that head shape dimorphism increases with overall
body size is also consistent with the hypothesis that sexual selection
on structures related to territoriality and male competitive
interactions are responsible for the macroevolutionary patterns of
sexual shape dimorphism observed in this group. To our knowledge, this
is the first example of Rensch's rule in a multivariate phenotype.
\hfill\break

With respect to the analytical approach proposed here, our
characterization of sexual dimorphism via a \emph{shape distance} and
the \(SDDI\) statistic is completely general, and may be used to
evaluate evolutionary trends in any continuous, quantitative trait that
differs between the sexes. Our empirical example highlighted the
efficacy of the approach for evaluating trends in univariate traits such
as body size, and also several types of complex data where organismal
phenotypes are quantified by multiple dimensions. As such, the approach
developed here expands our current analytical toolkit to provide a means
of evaluating evolutionary trends in sexual dimorphism for a broader
class of phenotypic traits than has previously been considered. Indeed,
this advance more closely embodies Rensch's (1950) original
conceptualization for understanding patterns of sexual dimorphism, as
his broader vision was to explain why phenotypic differences \emph{writ
large} exist among the sexes. Thus, use of the \(SDDI\) will facilitate
a greater degree of evolutionary exploration into patterns of sexual
dimorphism than has been possible in recent decades. Further, with the
inclusion of multivariate traits, the formulation proposed here opens
the door for investigations into the evolution of sexual shape
dimorphism, and whether such patterns correspond to those expected under
Rensch's rule. Thus, the mathematical test proposed here is general, and
may be used for any quantitative phenotypic trait under scrutiny.
\hfill\break

Finally, because \(SDDI\) is derived from a phenotypic distance, one
could, in theory, utilize the approach to characterize patterns of
sexual dimorphism in non-continuous traits, and determine quantitatively
whether the degree of sexual dimorphism is enhanced with increasing body
size. Across the tree of life, major phenotypic adaptations are often
described as evolutionary innovations or novelties (sensu Wagner and
Altenberg 1996; Wagner and Muller 2002; Peterson and Müller 2016), where
these changes are characterized as discrete shifts from one phenotypic
state to another (O'Keefe et al. 2011; e.g., Holliday and Gardner 2012).
In like manner, discrete phenotypic differences are known to exist
between the sexes; for instance where males display horns or other
ornamentation (Geist and Bayer 1988; Emlen 2008) that females lack. Over
evolutionary time, a series of such discrete phenotypic differences may
accumulate in males relative to females, resulting in an elaboration of
phenotypic differences between the sexes in suites of such traits. With
an appropriate distance measure for discrete traits (e.g., Hamming
distance), the approach developed here is capable of characterizing such
patterns, and determining whether sets of discrete phenotypic
differences are enhanced with increasing body size among species. Thus,
our approach is the first to our knowledge to facilitate a formal
evaluation of both discrete and continuous phenotypic differences
between the sexes, providing a comprehensive means of determining
whether sets of phenotypic traits differ in a manner as predicted by
Rensch's rule.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
\setlength{\parskip}{8pt} \noindent

\hypertarget{refs}{}
\hypertarget{ref-Abouheif1997}{}
Abouheif, E., and D. J. Fairbairn. 1997. A comparative analysis of
allometry for sexual size dimorphism: Assessing rensch's rule. The
American Naturalist 149:540--562.

\hypertarget{ref-Adams1999}{}
Adams, D. C. 1999. Methods for shape analysis of landmark data from
articulated structures. Evolutionary Ecology Research 1:959--970.

\hypertarget{ref-AdamsCollyer2018a}{}
Adams, D. C., and M. L. Collyer. 2018. Multivariate phylogenetic
comparative methods: Evaluations, comparisons, and recommendations.
Systematic Biology 67:14--31.

\hypertarget{ref-AdamsCollyer2019}{}
Adams, D. C., and M. L. Collyer. 2019. Phylogenetic comparative methods
and the evolution of multivariate phenotypes. Annual Review of Ecology,
Evolution, and Systematics 50:405--425.

\hypertarget{ref-AdamsFelice2014}{}
Adams, D. C., and R. Felice. 2014. Assessing trait covariation and
morphological integration on phylogenies using evolutionary covariance
matrices. PLoS ONE 9:e94335.

\hypertarget{ref-AdamsOtarola2013}{}
Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
Methods in Ecology and Evolution 4:393--399.

\hypertarget{ref-Adams_et_al2013}{}
Adams, D. C., F. J. Rohlf, and D. E. Slice. 2013. A field comes of age:
Geometric morphometrics in the 21st century. Hystrix 24:7--14.

\hypertarget{ref-AdamsGeomorph}{}
Adams, D., M. Collyer, and A. Kaliontzopoulou. 2019. Geomorph: Software
for geometric morphometric analyses. r package version 3.1.1.

\hypertarget{ref-Altman1983}{}
Altman, D. G., and J. M. Bland. 1983. Measurement in medicine: The
analysis of method comparison studies. The Statistician 32:307--317.

\hypertarget{ref-Andersson1994}{}
Andersson, M. 1994. Sexual selection. Princeton University Press.

\hypertarget{ref-Astua2010}{}
Astúa, D. 2010. Cranial sexual dimorphism in new world marsupials and a
test of renschs rule in didelphidae. Journal of Mammalogy 91:1011--1024.

\hypertarget{ref-Berns2013}{}
Berns, C. M., and D. C. Adams. 2013. Becoming different but staying
alike: Patterns of sexual size and shape dimorphism in bills of
hummingbirds. Evolutionary Biology 40:246--260.

\hypertarget{ref-Blanckenhorn2006}{}
Blanckenhorn, W. U., R. C. Stillwell, K. A. Young, C. W. Fox, and K. G.
Ashton. 2006. When rensch meets bergmann: Does sexual size dimorphism
change systematically with latitude? Evolution 60:2004--2011.

\hypertarget{ref-Blows2007}{}
Blows, M. W. 2007. A tale of two matrices: Multivariate approaches in
evolutionary biology. Journal of Evolutionary Biology 20:1--8.

\hypertarget{ref-Bolnick2003}{}
Bolnick, D. I., and M. Doebeli. 2003. Sexual dimorphism and adaptive
speciation: Two sides of the same ecological coin. Evolution
57:2433--2449.

\hypertarget{ref-Bonduriansky2007}{}
Bonduriansky, R. 2007. Sexual selection and allometry: A critical
appraisal of the evidence and ideas. Evolution 61:838--849.

\hypertarget{ref-Bookstein1997}{}
Bookstein, F. L. 1997. Landmark methods for forms without landmarks:
Morphometrics of group differences in outline shape. Medical Image
Analysis 1:225--243.

\hypertarget{ref-Bookstein1991}{}
Bookstein, F. L. 1991. Morphometric tools for landmark data: Geometry
and biology. Cambridge University Press, Cambridge.

\hypertarget{ref-Brana1996}{}
Braña, F. 1996. Sexual dimorphism in lacertid lizards: Male head
increase vs female abdomen increase? Oikos 75:511--523.

\hypertarget{ref-Burbrink2019}{}
Burbrink, F. T., and I. Futterman. 2019. Female-biased gape and
body-size dimorphism in the new world watersnakes (tribe: Thamnophiini)
oppose predictions from renschs rule. Ecology and Evolution
9:9624--9633.

\hypertarget{ref-Butler_et_al2000}{}
Butler, M. A., T. W. Schoener, and J. B. Losos. 2000. The relationship
between sexual size dimorphism and habitat use in greater antillean
\emph{Anolis} lizards. Evolution 54:259--272.

\hypertarget{ref-Ceballos2013}{}
Ceballos, C. P., D. C. Adams, J. B. Iverson, and N. Valenzuela. 2013.
Phylogenetic patterns of sexual size dimorphism in turtles and their
implications for Rensch's rule. Evolutionary Biology 40:194--208.

\hypertarget{ref-CollyerAdams2018}{}
Collyer, M. L., and D. C. Adams. 2018. RRPP: An R package for fitting
linear models to high-dimensional data using residual randomization.
Methods in Ecology and Evolution 9:1772--1779.

\hypertarget{ref-Collyer_et_al2015a}{}
Collyer, M. L., D. J. Sekora, and D. C. Adams. 2015. A method for
analysis of phenotypic change for phenotypes described by
high-dimensional data. Heredity 115:357--365.

\hypertarget{ref-Cox2003}{}
Cox, R. M., S. L. Skelly, and H. B. John-Alder. 2003. A comparative test
of adaptive hypotheses for sexual size dimorphism in lizards. Evolution
57:1653--1669.

\hypertarget{ref-Dale2007}{}
Dale, J., P. O. Dunn, J. Figuerola, T. Lislevand, T. Székely, and L. A.
Whittingham. 2007. Sexual selection explains renschs rule of allometry
for sexual size dimorphism. Proceedings of the Royal Society B:
Biological Sciences 274:2971--2979.

\hypertarget{ref-Darwin1871}{}
Darwin, C. R. 1871. The descent of man, and selection in relation to
sex. J. Murray, London.

\hypertarget{ref-DayanSimberloff2005}{}
Dayan, T., and D. Simberloff. 2005. Ecological and community-wide
character displacement: The next generation. Ecology Letters 8:875--894.

\hypertarget{ref-Emlen2008}{}
Emlen, D. J. 2008. The evolution of animal weapons. Annual Review of
Ecology, Evolution, and Systematics 39:387--413.

\hypertarget{ref-Endler1983}{}
Endler, J. A. 1983. Natural and sexual selection on color patterns in
poeciliid fishes. Environmental Biology of Fishes 9:173--190.

\hypertarget{ref-Fairbairn1997}{}
Fairbairn, D. J. 1997. Allometry for sexual size dimorphism:Pattern and
process in the coevolution of body size in males and females. Annual
Review of Ecology and Systematics 28:659--687.

\hypertarget{ref-Fairbairn2013}{}
Fairbairn, D. J. 2013. Odd couples - extraordinary differences between
the sexes in the animal kingdom. Princeton University Press.

\hypertarget{ref-Fairbairn1994}{}
Fairbairn, D. J., and R. F. Preziosi. 1994. Sexual selection and the
evolution of allometry for sexual size dimorphism in the water strider,
Aquarius remigis. American Naturalist 144:101--118.

\hypertarget{ref-Fairbairn2007}{}
Fairbairn, D. J., W. U. Blanckenhorn, and T. Székely. 2007. Sex, size
and gender roles. Oxford University Press.

\hypertarget{ref-Garcia-Navas2015}{}
Garcia-Navas, V., T. Bonnet, R. Bonal, and E. Postma. 2015. The role of
fecundity and sexual selection in the evolution of size and sexual size
dimorphism in new world and old world voles (rodentia: Arvicolinae).
Oikos 125:1250--1260.

\hypertarget{ref-Geist1988}{}
Geist, V., and M. Bayer. 1988. Sexual dimorphism in the cervidae and its
relation to habitat. Journal of Zoology 214:45--53.

\hypertarget{ref-Gunz2009}{}
Gunz, P., P. Mitteroecker, S. Neubauer, G. W. Weber, and F. L.
Bookstein. 2009. Principles for the virtual reconstruction of hominin
crania. Journal of Human Evolution 57:48--62.

\hypertarget{ref-Haas1973}{}
Haas, G. 1973. Muscles of the jaws and associated structures in the
rhynchocephalia and squamata. Pp. 285--490 \emph{in} C. Biology of the
reptilia vol. 4. Gans and T. Parsons, eds. Academica Press.

\hypertarget{ref-Herrel1996}{}
Herrel, A., R. V. Damme, and F. D. Vree. 1996. Sexual dimorphism of head
size in podarcis hispanica atrata: Testing the dietary divergence
hypothesis by bite force analysis. Netherlands Journal of Zoology
46:253--262.

\hypertarget{ref-Hirst2014}{}
Hirst, A. G., and T. Kiørboe. 2014. Macroevolutionary patterns of sexual
size dimorphism in copepods. Proceedings of the Royal Society B:
Biological Sciences 281:20140739.

\hypertarget{ref-Holliday2012}{}
Holliday, C. M., and N. M. Gardner. 2012. A new eusuchian crocodyliform
with novel cranial integument and its significance for the origin and
evolution of crocodylia. PLoS ONE 7:e30471.

\hypertarget{ref-Horne2020}{}
Horne, C. R., A. G. Hirst, and D. Atkinson. 2020. Selection for
increased male size predicts variation in sexual size dimorphism among
fish species. Proceedings of the Royal Society B: Biological Sciences
287:20192640.

\hypertarget{ref-HUYGHE2005}{}
Huyghe, K., B. V. Hooydonck, H. Scheers, M. Molina-Borij, and R. V.
Damme. 2005. Morphology, performance and fighting capacity in male
lizards, \emph{Gallotia galloti}. Functional Ecology 19:800--807.

\hypertarget{ref-Johnson2017}{}
Johnson, P. J., M. J. Noonan, A. C. Kitchener, L. A. Harrington, C.
Newman, and D. W. Macdonald. 2017. Rensching cats and dogs: Feeding
ecology and fecundity trends explain variation in the allometry of
sexual size dimorphism. Royal Society Open Science 4:170453.

\hypertarget{ref-Kaliontzopoulou2012}{}
Kaliontzopoulou, A., D. C. Adams, A. van der Meijden, A. Perera, and M.
A. Carretero. 2012. Relationships between head morphology, bite
performance and ecology in two species of podarcis wall lizards.
Evolutionary Ecology 26:825--845.

\hypertarget{ref-Kaliontzopoulou2015}{}
Kaliontzopoulou, A., M. A. Carretero, and D. C. Adams. 2015.
Ecomorphological variation in male and female wall lizards and the
macroevolution of sexual dimorphism in relation to habitat use. Journal
of Evolutionary Biology 28:80--94.

\hypertarget{ref-Kaliontzopoulou_et_al2010}{}
Kaliontzopoulou, A., M. A. Carretero, and G. A. Llorente. 2010.
Intraspecific ecomorphological variation: Linear and geometric
morphometrics reveal habitat-related patterns within \emph{podarcis
bocagei} wall lizards. Journal of Evolutionary Biology 23:1234--1244.

\hypertarget{ref-Kelly2013}{}
Kelly, C. D., K. E. Folinsbee, D. C. Adams, and M. D. Jennions. 2013.
Intraspecific sexual size and shape dimorphism in an australian
freshwater fish differs with respect to a biogeographic barrier and
latitude. Evolutionary Biology 40:408--419.

\hypertarget{ref-Lailvaux2007}{}
Lailvaux, S. P., and D. J. Irschick. 2007. The evolution of
performance-based male fighting ability in caribbean anolis lizards. The
American Naturalist 170:573--586.

\hypertarget{ref-Lande1979}{}
Lande, R. 1979. Quantitative genetic analysis of multivariate evolution,
applied to brain: Body size allometry. Evolution 33:402--416.

\hypertarget{ref-LandeArnold1983}{}
Lande, R., and S. J. Arnold. 1983. The measurement of selection on
correlated characters. Evolution 37:1210--1226.

\hypertarget{ref-Littleford-Colquhoun2019}{}
Littleford-Colquhoun, B. L., C. Clemente, G. Thompson, R. H. Cristescu,
N. Peterson, K. Strickland, D. Stuart-Fox, and C. H. Frere. 2019. How
sexual and natural selection shape sexual size dimorphism: Evidence from
multiple evolutionary scales. Functional Ecology 33:1446--1458.

\hypertarget{ref-Lovich1992}{}
Lovich, J. E., and J. W. Gibbons. 1992. A review of techniques for
quantifying sexual size dimorphism. Growth, Development, and Aging
56:269--281.

\hypertarget{ref-Meiri2014}{}
Meiri, S., A. E. Kadison, M. Novosolov, P. Pafilis, J. Foufopoulos, Y.
Itescu, P. Raia, and D. Pincheira-Donoso. 2014. The number of competitor
species is unlinked to sexual dimorphism. Journal of Animal Ecology
83:1302--1312.

\hypertarget{ref-MitteroeckerGunz2009}{}
Mitteroecker, P., and P. Gunz. 2009. Advances in geometric
morphometrics. Evolutionary Biology 36:235--247.

\hypertarget{ref-OKeefe2011}{}
O'Keefe, F. R., H. P. Street, B. C. Wilhelm, C. D. Richards, and H. Zhu.
2011. A new skeleton of the cryptoclidid plesiosaurTatenectes
laramiensisreveals a novel body shape among plesiosaurs. Journal of
Vertebrate Paleontology 31:330--339.

\hypertarget{ref-Oestman2011}{}
Östman, O., and D. Stuart-Fox. 2011. Sexual selection is positively
associated with ecological generalism among agamid lizards. Journal of
Evolutionary Biology 24:733--740.

\hypertarget{ref-Penalver-Alcazar2019}{}
Peñalver-Alcázar, M., P. Galán, and P. Aragón. 2019. Assessing rensch's
rule in a newt: Roles of primary productivity and conspecific density in
interpopulation variation of sexual size dimorphism. Journal of
Biogeography 46:2558--2569.

\hypertarget{ref-Peterson2016}{}
Peterson, T., and G. B. Müller. 2016. Phenotypic novelty in EvoDevo: The
distinction between continuous and discontinuous variation and its
importance in evolutionary theory. Evolutionary Biology 43:314--335.

\hypertarget{ref-R-Base}{}
R Core Team. 2019. R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria.

\hypertarget{ref-Regis2017}{}
Regis, K. W., and J. M. Meik. 2017. Allometry of sexual size dimorphism
in turtles: A comparison of mass and length data. PeerJ 5:e2914.

\hypertarget{ref-Reiss1986}{}
Reiss, M. J. 1986. Sexual dimorphism in body size: Are larger species
more dimorphic? Journal of Theoretical Biology 121:163--172.

\hypertarget{ref-Rensch1950}{}
Rensch, B. 1950. Die abhangigkeit der relativen sexualdifferenz von der
korpergrosse. Bonner Zoologische Beitraege 1:58--69.

\hypertarget{ref-Rensch1960}{}
Rensch, B. 1960. Evolution above the species level. Columbia University
Press.

\hypertarget{ref-RohlfCorti2000}{}
Rohlf, F. J., and M. Corti. 2000. The use of partial least-squares to
study covariation in shape. Systematic Biology 49:740--753.

\hypertarget{ref-RohlfSlice1990}{}
Rohlf, F. J., and D. E. Slice. 1990. Extensions of the procrustes method
for the optimal superimposition of landmarks. Systematic Zoology
39:40--59.

\hypertarget{ref-Serrano-Meneses2006}{}
Serrano-Meneses, M.-A., and T. Szekely. 2006. Sexual size dimorphism in
seabirds: Sexual selection, fecundity selection and differential
niche-utilisation. Oikos 113:385--394.

\hypertarget{ref-Stamps1983}{}
Stamps, J. 1983. Sexual selection, sexual dimorphism and territoriality.
Pp. 169--204 \emph{in} Lizard ecology: studies of a model organism. Huey
R.B., Pianka E.R., and T. Schoener, eds. Harvard University Press.

\hypertarget{ref-Stephens2009}{}
Stephens, P. R., and J. J. Wiens. 2009. Evolution of sexual size
dimorphisms in emydid turtles: Ecological dimorphism, rensch's rule, and
sympatric divergence. Evolution 63:910--925.

\hypertarget{ref-Stuart-Fox2008}{}
Stuart-Fox, D. 2008. A test of rensch's rule in dwarf chameleons
(bradypodion spp.), a group with female-biased sexual size dimorphism.
Evolutionary Ecology 23:425--433.

\hypertarget{ref-Stuart-Fox2007}{}
Stuart-Fox, D., and A. Moussalli. 2007. Sex-specific ecomorphological
variation and the evolution of sexual dimorphism in dwarf chameleons
(bradypodion spp.). Journal of Evolutionary Biology 20:1073--1081.

\hypertarget{ref-Tarr2018}{}
Tarr, S., S. Meiri, J. J. Hicks, and A. C. Algar. 2018. A biogeographic
reversal in sexual size dimorphism along a continental temperature
gradient. Ecography 42:706--716.

\hypertarget{ref-Wagner1996a}{}
Wagner, G. P., and L. Altenberg. 1996. Perspective: Complex adaptations
and the evolution of evolvability. Evolution 50:967.

\hypertarget{ref-Wagner2002}{}
Wagner, G. P., and G. B. Muller. 2002. Evolutionary innovations overcome
ancestral constraints: A re-examination of character evolution in male
sepsid flies (diptera: Sepsidae). Evolution and Development 4:1--6.

\hypertarget{ref-Watson2010}{}
Watson, N. L., and L. W. Simmons. 2010. Mate choice in the dung beetle
onthophagus sagittarius: Are female horns ornaments? Behavioral Ecology
21:424--430.

\newpage

\section{Figure Legends}\label{figure-legends}

\textbf{Figure 1}. (A) Sexual size dimorphism for males and females
across 100 hypothetical species. The pattern displayed exhibits a slope
\textgreater{} 1.0, and thus corresponds to Rensch's rule. (B) The same
data represented as the ratio of \(M/F\) sizes plotted against average
species size. (C) Sexual dimorphism represented as the distance between
males and females: \(SDD = D_{M:F}\). (D) Sexual dimorphism represented
as an index of the SD Distance: \(SDDI = \pm1*D_{M:F}\). In all panels,
the dashed line corresponds to values representing no sexual dimorphism
(the 1:1 line). \hfill\break

\textbf{Figure 2}. Results of phylogenetic simulations testing the type
I error and statistical power for detecting patterns consistent with
Rensch's rule. Vertical bars represent the proportion of datasets found
to be significant. For type I error analyses, the dashed line marks the
standard threshold of \(\alpha= 0.05\). For univariate data, both
phylogenetic regression of male versus female size (standard approach),
and \(SDDI\) versus overall body size (new approach) were performed. For
multivariate data, only \(SDDI\) versus overall body size was examined.
\hfill\break

\textbf{Figure 3}. (A) Set of linear measurements used to obtain head
proportions. HL: head length, HW: head width, HH: head height, PL:
pileus length, MO: mouth opening. (B) Locations of 17 landmarks (1-10;
17) and semilandmarks (11-16) used to quantify head shape. (C) Patterns
of sexual size dimorphism represented as male versus female size. (D)
Patterns of sexual size dimorphism, represented as sexual size
dimorphism index (\(SSDI\)) versus mean size. (E) Multivariate patterns
sexual dimorphism in body proportions (\(SShDI\)) versus mean size. (F)
Multivariate patterns of sexual shape dimorphism (\(SShDI\)) versus mean
size. In panels C, D, E, and F, the red line indicates the regression
estimate obtained from the data, while the dashed line corresponds to
values representing no sexual dimorphism. \hfill\break

\textbf{Figure 4}. Principal components plot of male (black squares) and
female (red points) species mean head shape: the first two PC axes
explain over 60\% of the total shape variation. Green lines correspond
to the sexual shape dimorphism in one of the smallest species,
\emph{Lacerta agilis bosnica}, and in the largest species, \emph{Timon
lepidus}. Thin-plate spline deformation grids display the mean head
shape for males and females of these species. \hfill\break

\textbf{Figure 5}. Heat maps (green: lower values; red: higher values)
displaying the evolution of (A) sexual size dimorphism, and (B) sexual
shape dimorphism across the phylogeny. For each panel, an index of
sexual size dimorphism (\(SSDI\)) or sexual shape dimorphism (\(SShDI\))
is displayed, with circles being proportional to the degree of sexual
dimorphism. \hfill\break

\newpage

\includegraphics[width=0.95\linewidth]{Fig1}

\singlespacing \textbf{Figure 1}. (A) Sexual size dimorphism for males
and females across 100 hypothetical species. The pattern displayed
exhibits a slope \textgreater{} 1.0, and thus corresponds to Rensch's
rule. (B) The same data represented as the ratio of \(M/F\) sizes
plotted against average species size. (C) Sexual dimorphism represented
as the distance between males and females: \(SDD = D_{M:F}\). (D) Sexual
dimorphism represented as an index of the SD Distance:
\(SDDI = \pm1*D_{M:F}\). In all panels, the dashed line corresponds to
values representing no sexual dimorphism (the 1:1 line). \hfill\break

\newpage

\includegraphics[width=0.7\linewidth]{Fig2}

\singlespacing \textbf{Figure 2}. Results of phylogenetic simulations
testing the type I error and statistical power for detecting patterns
consistent with Rensch's rule. Vertical bars represent the proportion of
datasets found to be significant. For type I error analyses, the dashed
line marks the standard threshold of \(\alpha= 0.05\). For univariate
data, both phylogenetic regression of male versus female size (standard
approach), and \(SDDI\) versus overall body size (new approach) were
performed. For multivariate data, only \(SDDI\) versus overall body size
was examined. \hfill\break

\newpage

\includegraphics[width=0.95\linewidth]{Fig3}

\singlespacing \textbf{Figure 3}. (A) Set of linear measurements used to
obtain head proportions. HL: head length, HW: head width, HH: head
height, PL: pileus length, MO: mouth opening. (B) Locations of 17
landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head
shape. (C) Patterns of sexual size dimorphism represented as male versus
female size. (D) Patterns of sexual size dimorphism, represented as
sexual size dimorphism index (\(SSDI\)) versus mean size. (E)
Multivariate patterns sexual dimorphism in body proportions (\(SShDI\))
versus mean size. (F) Multivariate patterns of sexual shape dimorphism
(\(SShDI\)) versus mean size. In panels C, D, E, and F, the red line
indicates the regression estimate obtained from the data, while the
dashed line corresponds to values representing no sexual dimorphism.
\hfill\break

\newpage

\includegraphics[width=0.95\linewidth]{Fig4}

\singlespacing \textbf{Figure 4}. Principal components plot of male
(black squares) and female (red points) species mean head shape: the
first two PC axes explain over 60\% of the total shape variation. Green
lines correspond to the sexual shape dimorphism in one of the smallest
species, \emph{Lacerta agilis bosnica}, and in the largest species,
\emph{Timon lepidus}. Thin-plate spline deformation grids display the
mean head shape for males and females of these species. \hfill\break

\newpage

\includegraphics[width=0.95\linewidth]{Fig5}

\singlespacing \textbf{Figure 5}. Heat maps (green: lower values; red:
higher values) displaying the evolution of (A) sexual size dimorphism,
and (B) sexual shape dimorphism across the phylogeny. For each panel, an
index of sexual size dimorphism (\(SSDI\)) or sexual shape dimorphism
(\(SShDI\)) is displayed, with circles being proportional to the degree
of sexual dimorphism. \hfill\break

\end{document}

### 2020-03-15 CoverPage-Evolution.rmd
kint attempt - success
moved rench's rule docs into Example folder

rmd files in main folder are to be edited for lambda manuscript

@@
 ---
-title: A MULTIVARIATE EQUIVALENCY AND GENERAL TEST FOR RENSCH'S RULE: IMPLICATIONS FOR THE EVOLUTION OF SEXUAL SHAPE DIMORPHISM
+title: MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS
 
@@
 \begin{center}
-\textbf{Dean C. Adams$^{1,*}$, Elizabeth Glynne$^{1}$ and Antigoni Kaliontzopoulou$^2$}
+\textbf{Dean C. Adams$^{1*}$ and Erica K. Baken$^{1}$}
 \end{center}
@@
 \hfill\break
-**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break
+**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break
 
-**Short Title**: Multivariate Test of Rensch's Rule \hfill\break
+**Short Title**: Problems with Pagel's lambda \hfill\break
 
-**Author Contributions**: DCA devised the approach, AK provided the empirical example; all authors contributed  to all parts of this manuscript. \hfill\break
+**Author Contributions**: DCA ...EKB...; all authors contributed  to all parts of this manuscript. \hfill\break
 

### 2020-03-15 Manuscript-template.rmd
kint attempt - success
moved rench's rule docs into Example folder

rmd files in main folder are to be edited for lambda manuscript

@@
 ---
-# A multivariate equivalency and general test for Rensch's Rule: Implications for the evolution of sexual shape dimorphism
+# Misuses and Inaccuracies of Pagel's Lambda ML Estimation Methods
 
@@
 
-**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break
+**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break
 
-**Short Title**: Multivariate Test of Rensch's Rule \hfill\break
+**Short Title**: Problems with Pagel's lambda \hfill\break
 

### 2020-03-24 Manuscript-template.rmd
Organizing data, running fixed simulation scripts for Regressions analyses. Not yet finished: reg tree size 512 and 1028, and all ANOVA analyses.

@@
 Indeed, Rensch's original treatment focused not only on size, but also on dimorphism in relative body proportions, coloration, the presence of ornaments, etc.
-Here, we derive a multivariate equivalency for viewing trends in sexual dimorphism – relative to overall body size – across taxa, and provide a generalized test to determine whether such patterns are consistent with Rensch's rule.
+<!-- I think this can be rephrased --> Here, we derive a multivariate equivalency for viewing trends in sexual dimorphism – relative to overall body size – across taxa, and provide a generalized test to determine whether such patterns are consistent with Rensch's rule.
 For univariate linear traits such as body size, our approach yields equivalent results to those from standard procedures, but our test is also capable of detecting trends in multivariate datasets like shape.

### 2020-03-24 Data_Analyses/Simulation_script-PB-ANOVA.R

@@ -1,0 +1,157 @@
+# Simulating pgls inference errors from estimated lambdas
+
+# look at Revell ~2009 where he investigated Y~X simulating error in X and in Y or one or the other under BM
+setwd("/Users/ericabaken/Documents/School/Projects/Lambda_Eval/Manuscript/2020-Adams_and_Baken/")
+
+library(phytools)
+library(geomorph)
+library(geiger)
+library(svMisc)
+library(caper)
+
+treesizes <-  2^(5:10) # 32,64,128,256,512,1024
+n <- treesizes[5] # number of tips (not simulations)
+nsim <- 50 # number of simulations
+lambdas <- seq(0,1,.05) # starting with just 3 lambda values
+lambdas_expanded <- rep(lambdas, each = nsim)
+beta <- c(0, .25, .5, .75, 1) # true slope
+
+DataTable_lambda_0 <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+
+DataTable_lambda_ml <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+
+DataTable_lambda_0[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
+DataTable_lambda_ml[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
+
+Start.time<-Sys.time()
+Start.time
+for (j in 1:(length(lambdas_expanded))) {
+
+  while(TRUE){
+ 
+    tree<-pbtree(n=n,scale=1) 
+    
+    if(lambdas_expanded[j] > 0) {
+    tree_scaled <- rescale(tree, model = "lambda", lambda = lambdas_expanded[j])
+    }
+    
+    if(lambdas_expanded[j] == 0) { # without this bit, lambda always estimated as 0.5, and it doesn't make sense
+    tree_scaled <- rescale(tree, model = "lambda", lambda = 0.01) 
+    }
+    
+    X <- rTraitDisc(tree_scaled, model = "ER", states = c(0,1))
[…]

### 2020-03-24 Data_Analyses/Simulation_script-PB-Reg.R

@@ -1,0 +1,125 @@
+# Simulating pgls inference errors from estimated lambdas
+
+# look at Revell ~2009 where he investigated Y~X simulating error in X and in Y or one or the other under BM
+setwd("/Users/ericabaken/Documents/School/Projects/Lambda_Eval/Manuscript/2020-Adams_and_Baken/")
+
+library(phytools)
+library(geomorph)
+library(geiger)
+library(svMisc)
+library(caper)
+
+treesizes <-  2^(5:10) # 32,64,128,256,512,1024
+nsim <- 50 # number of simulations
+lambdas <- seq(0,1,.05) # starting with just 3 lambda values
+lambdas_expanded <- rep(lambdas, each = nsim)
+beta <- c(0, .25, .5, .75, 1) # true slope
+fullsimlength <- length(lambdas_expanded)
+
+
+n <- treesizes[5] # number of tips (not simulations)
+
+DataTable_lambda_0 <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+
+DataTable_lambda_ml <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+
+DataTable_lambda_0[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
+DataTable_lambda_ml[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
+
+Start.time<-Sys.time()
+Start.time
+for (j in 1:(length(lambdas_expanded))) {
+
+  while(TRUE){
+ 
+    tree<-pbtree(n=n,scale=1) 
+    
+    if(lambdas_expanded[j] > 0) {
+    tree_scaled <- rescale(tree, model = "lambda", lambda = lambdas_expanded[j])
+    }
+    
+    if(lambdas_expanded[j] == 0) { # without this bit, lambda always estimated as 0.5, and it doesn't make sense
+    tree_scaled <- rescale(tree, model = "lambda", lambda = 0.01) 
[…]

### 2020-03-24 VersionDiff.r

@@ -1,0 +1,8 @@
+# Script to compare two Rmd files
+ #Note, requires installation of 'latexdiff' in your tex components & a version of Perl
+
+# NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
+
+library(latexdiffr)
+latexdiff("2020-Evol-MultiRR_Vsn6.Rmd", "2020-Evol-MultiRR_Vsn6_AK.Rmd", clean = TRUE)
+

### 2020-03-26 Manuscript.Rmd
All regression analyses done except 1024.
Very rough beginning to manuscript.

--- 2020-03-15 Manuscript-template.rmd
@@
 ---
-# A multivariate equivalency and general test for Rensch's Rule: Implications for the evolution of sexual shape dimorphism
+# Pagel's Lambda Estimates are Often Inaccurate
 
@@
 
-**Keywords**: Sexual dimorphism, multivariate, morphological evolution \hfill\break
+**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break
 
-**Short Title**: Multivariate Test of Rensch's Rule \hfill\break
+**Short Title**: Inaccuracies in Pagel's Lambda \hfill\break
 
@@
 
-Rensch's rule, the allometric association between the degree of sexual dimorphism and overall body size, is an important framework for understanding the evolution of trends in phenotypic differentiation between the sexes across taxa.
-While virtually all recent treatments have explored the degree of sexual size dimorphism, understanding how dimorphism in other phenotypic traits evolves is equally relevant.
-Indeed, Rensch's original treatment focused not only on size, but also on dimorphism in relative body proportions, coloration, the presence of ornaments, etc.
-Here, we derive a multivariate equivalency for viewing trends in sexual dimorphism – relative to overall body size – across taxa, and provide a generalized test to determine whether such patterns are consistent with Rensch's rule.
-For univariate linear traits such as body size, our approach yields equivalent results to those from standard procedures, but our test is also capable of detecting trends in multivariate datasets like shape.
-Computer simulations reveal the method displays appropriate statistical properties, and an empirical example in Mediterranean lizards illustrates the efficacy of our approach on both univariate and multivariate phenotypes.
-Our generalized procedure substantially extends the toolkit available for investigating macroevolutionary patterns of sexual dimorphism and seeking a better understanding of the processes that underlie them.
+ conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)
 
@@
 # Introduction
-It is widely observed that males and females of the same species are not identical, but rather differ -- sometimes substantially -- in their phenotypic characteristics [@Fairbairn1997; @Fairbairn2013].
-Indeed, sexual dimorphism is pervasive across many animal clades, and understanding the causes for these patterns has attracted considerable attention in evolutionary biology for well over a century [e.g., @Darwin1871; @Fairbairn2007].
-One of the most conspicuous differences between males and females is in their body sizes [@Andersson1994], where sex-specific differences are often hypothesized to be the result of selection pressures related to reproduction, or the distinct ecological roles of males and females in their respective habitat [@Stephens2009; @Littleford-Colquhoun2019; @Tarr2018; @Kaliontzopoulou2015; @Cox2003].
-For example, both sexual selection on male body size [@Cox2003; @Garcia-Navas2015; @Horne2020], and fecundity selection on female body size [@Stuart-Fox2008; @Serrano-Meneses2006], are expected to enhance the degree of sexual size dimorphism in taxa over evolutionary time.
-Likewise, when males and females respond differently to environmental factors, intraspecific competition may generate ecological divergence between them, resulting in varying patterns of sexual dimorphism [@DayanSimberloff2005; @Meiri2014; @Kaliontzopoulou_et_al2010; @Butler_et_al2000; @Bolnick2003].
+Investigating macroevolutionary patterns requires a phylogenetic approach as species are non-independent by nature of their shared ancestry.
+Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; [@Felsenstein1987]), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
+[@AdamsNason2019]).
+Understanding the degree of phylogenetic signal present in a dataset is paramount, and identifies the mode with which a trait has evolved.
+High measures of phylogenetic signal indicate a Brownian motion process, whereas lower levels of phylogenetic signal indicate natural selection or some other evolutionary force has influenced the traits evolutionary history.
 \hfill\break
 
-At macroevolutionary scales, changes in the relative contribution of the selective forces above may create variation in the degree of sexual dimorphism displayed across closely related taxa.
-For instance, sister species inhabiting different environments, or possessing distinct ecological adaptations, may exhibit differing degrees of sexual dimorphism, as a result of the varying intensity of sexual selection they endure [@Oestman2011].
-Likewise, the interplay between sexual and natural selection that species experience may result in sexual dimorphism of differing magnitudes across taxa [@Kaliontzopoulou2015; @Stuart-Fox2007; @Blanckenhorn2006].
-One evolutionary framework that connects adaptive explanations for the evolution of sexual size dimorphism (SSD) with trends in overall body size across taxa is Rensch's rule.
-Here it is predicted that SSD will increase with species' average body size in species where males are the larger sex (hyperallometry), and likewise will decrease with increasing body size where females are the larger sex (hypoallometry) [@Abouheif1997; @Fairbairn1997].
-Put another way, Rensch's rule predicts that the extent of sexual dimorphism increases at more extreme body sizes (both large and small) across taxa.
-Empirical studies have documented trends in SSD consistent with Rensch's rule in a wide variety of animal taxa [@Dale2007; @Ceballos2013; @Regis2017; @Johnson2017].
-However, other clades display the converse trend [@Burbrink2019; @Penalver-Alcazar2019], and some clades display no relationship between SSD and body size [@Astua2010; @Hirst2014; @Johnson2017].
-From an evolutionary perspective, the association between SSD and body size has been hypothesized to be driven by variation in ontogenetic or static allometries across taxa, which results from the influence of sexual selection on growth and maturation patterns [@Rensch1960; @Dale2007].
+Several approaches to quantify phylogenetic signal exist.
+For continuous data, the most common parameters used in the literature include Pagel's lambda [@Pagel1997] and Blomberg's kappa [@Blomberg1900].
+Pagel's lambda has the advantage of being couched in the likelihood framework and thus has also been utilized to encorporate phylogenetic signal while doing phylogenetic regressions and ANOVA.
+However, the accuracy of the lambda estiamtion methods have not been fully evaluated, and thus it remains unknown the degree to which lambda estimates appropriately represent degree of phylogenetic signal.
 \hfill\break
 
-Interestingly, virtually all evaluations of Rensch's rule have investigated patterns of sexual size dimorphism, yet Rensch's original contributions were concerned with a much wider array of phenotypic traits that can differ between the sexes.
-These included not only sexual differences in body size, but also in general body proportions (e.g., relative wing and limb lengths), the degree of complexity of different body structures, the presence of sex-specific  ornamentation (e.g., horns, body coloration, etc.), and other sex-specific differences [@Rensch1950; @Rensch1960].
-In like manner, some studies have documented patterns of sexual shape dimorphism [e.g., @Berns2013; @Kelly2013; @Kaliontzopoulou2015], sex-specific differences in ornamentation [@Geist1988; @Watson2010; @Emlen2008], or coloration [e.g., @Endler1983].
-However, such studies have typically been restricted to one or a few taxa, and have not explained macroevolutionary variation in sexual dimorphism in a manner that relates directly to Rensch's rule.
-Thus, when viewed from this perspective, Rensch's [-@Rensch1950] vision was far more synthetic, as he sought to explain the extent to which sexual dimorphism -- across a general suite of phenotypic characteristics -- scaled with overall body size in organisms, and why some species displayed a greater degree of sexual dimorphism than did others.
-Remarkably, patterns of intraspecific static allometry in sexually selected traits establishes a direct link between variation in those traits relative to size, and the evolutionary allometry across taxa in the degree of sexual dimorphism that may be expected [@Reiss1986; @Bonduriansky2007; see also @Rensch1960].
-Under this framework, the "exaggeration" of body structures or other phenotypic traits with increasing body size is predicted to translate, at macroevolutionary scales, into increased degrees of sexual dimorphism with increasing body size.
-However, despite this clear prediction, and Rensch's original observations, the evolutionary scaling of the degree of sexual dimorphism in phenotypic traits other than size, remains largely unexplored.
-\hfill\break
-
-Another current limitation in our understanding of Rensch's rule is the fact that to date, all studies have been conducted exclusively on univariate data, primarily regarding body size.
-However, understanding phenotypic evolution is inherently a multivariate endeavor [@Blows2007; @Collyer_et_al2015a; @AdamsCollyer2018a; @AdamsCollyer2019], as processes such as natural and sexual selection simultaneously operate on more than one trait [@Lande1979; @LandeArnold1983], and can affect variation in complex multidimensional traits such as shape.
-To interrogate such patterns, the analytics used to evaluate evolutionary trends in phenotypic datasets must be sufficiently general to accommodate this empirical fact.
-Unfortunately, all current analytical approaches for evaluating trends in sexual dimorphism as they relate to overall body size are explicitly univariate; thereby excluding the possibility of evaluating patterns in complex multidimensional traits such as organismal shape.
-Though is has been extensively documented in sexual size dimorphism, it remains unknown whether Rensch's rule applies equally to sexual shape dimorphism, as determining this requires an analytical framework capable of evaluating trends in sexual dimorphism in both univariate and multivariate datasets.
-\hfill\break
-
-In this paper, we derive a multivariate equivalency for inspecting variation in sexual dimorphism -- relative to overall body size -- across taxa, and provide a generalized test to determine whether they are consistent with Rensch's rule.
-Our procedure is appropriate for evaluating trends in univariate traits measured on a linear scale (such as body size or relative limb length), as well as in multivariate datasets representing complex phenotypes (such as sets of body proportions or landmark-based shape data).
-For univariate data, the approach yields equivalent results to what is normally accomplished for size-based SSD evaluations.
-However, the approach extends these procedures for the examination of sexual dimorphism in multivariate data, thereby increasing the potential for our conceptual understanding of the consequences of sexual and natural selection on complex organismal phenotypes.
-We conduct a series of computer simulations demonstrating that the approach displays appropriate statistical properties when evaluating trends in both univariate and multivariate data.
-We also provide an empirical example in Mediterranean green lizards which illustrates the efficacy of our approach on both univariate and multivariate datasets.
-Finally, we highlight the new avenues of evolutionary research that our generalization facilitates, and make comments on the prospects for future empirical advances.
+An earlier study[@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies could be using  estimation methods for various parameters.
+That paper concluded that a measure of power must be considered when quantifying Pagel's lambda.
+Here we take a more comprehensive approach to demonstrate the scenarios under which estimated lambdas accurately reflect known lambdas as well as the effect of these at times dubious estimation methods on significance testing when used in a pgls framework.
 
@@
 
-## *Conceptual Development*
+## *Simulated trait*
 
-To arrive at a general test for Rensch's rule that can accommodate multivariate data, we must first derive a series of mathematical equivalencies based upon how interspecific patterns of size dimorphism are typically evaluated.
-For sexual size dimorphism, interspecific patterns are often inspected using a bivariate plot, where the size of one sex is plotted against the size of the other sex [e.g., @Fairbairn1994; @Abouheif1997; @Ceballos2013].
-An idealized example is shown in Fig.
-1A, with male size plotted against female size.
-The diagonal line represents no size dimorphism; or a 1:1 size ratio between the sexes across the range of body sizes for species in the clade of interest.
-In this construction (i.e., with females plotted on the x-axis), deviations from this line represent differential trends in size dimorphism across taxa.
-To evaluate such patterns statistically, a phylogenetic regression is performed, and in this case, a slope significantly greater than 1.0 (i.e., $\beta_{1}>1.0$) would be treated as evidence that the evolutionary changes in sexual size dimorphism across taxa are consistent with Rensch's rule [@Fairbairn1997].
+To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies with known lambdas.
+We ____ by scaling simulated phylogeny with the scaling parameter and .
+<!-- obviously this paragraph sucks -->  We also did this with other tree shapes (symmetrical and ladder).
 \hfill\break
 
-An equivalent representation of patterns of sexual dimorphism may also be found using ratios (e.g., male size:female size).
-When plotted against the average size for each species (Fig.
-1B), the same information is conveyed as is found in the more commonly utilized male - female plot of Fig.
-1A.
-However, in this representation, the line designating no sexual dimorphism is horizontal, with a slope of zero (i.e., $\beta_{1}=0.0$) and an intercept of 1.0.
-As before, a phylogenetic regression may be performed on these data, and in this case, if the slope is significantly steeper than zero (i.e., $\beta_{1}>0.0$), patterns consistent with Rensch's rule are observed.
-\hfill\break
+## *Simulated ANOVA and Regressions*
 
-A third equivalency is shown in Fig.
-1C.
-Here, the Euclidean distance between male and female means for each species ($D_{SD}=\sqrt{(Y_{male}-Y_{female})^2}$) is shown relative to the overall size of the species.
-This plot is similar to a Bland-Altmann plot [@Altman1983], except that the Euclidean distance, rather than a difference score, is plotted along the y-axis (as distances can accommodate data of different dimensionality).
-Under Rensch's rule, this *sexual dimorphism distance* may become greater as species are progressively larger [hyperallometry, sensu @Abouheif1997] or may become greater as species are progressively smaller [hypoallometry, sensu @Abouheif1997].
-In both cases, patterns correspond to species displaying progressively more sexual dimorphism at extreme body sizes; a trend consistent with Rensch's rule.
-Note that the inflection point in Fig.
-1C represents an intermediate size where there is no sexual dimorphism for the trait, and corresponds to where the male-female regression line of Fig.
-1A crosses the line representing no sexual dimorphism.
-This inflection point may represent the overall mean size across species in the lineage, or may be some other value.
-For instance, in some lineages [e.g., turtles: @Ceballos2013], nearly all species are female biased, in which case the inflection point for $D_{SD}$ vs.
-size would be found at the far-right end of the plot.
-\hfill\break
+To ascertain the statistical performance of pgls  \hfill\break
 
-Finally, from $D_{SD}$ a *sexual dimorphism distance index* ($SDDI$) may be constructed as:
+## *Meta-Analysis of Empirical Results*
 
-$$SDDI=-1*D_{SD}$$
-$$SDDI=+1*D_{SD}$$
+Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so. \hfill\break
 
-Here, $D_{SD}$ is multiplied by either a +1 or -1, based on a comparison of male and female trait values for the species.
-For instance, when single-valued traits on a linear scale are examined (e.g., leg length or wing extent), $D_{SD}$ is multiplied by -1 for those species where females display larger values than males, and remains unchanged for species where males are larger than females.
-This logic is analogous to that of Lovich and Gibbons [-@Lovich1992] for their index of overall body size dimorphism.
-However, from a geometric perspective, this procedure is tantamount to determining which species in Fig.
-1A are found below the 1:1 size line (i.e., the line representing no size dimorphism), and multiplying those species' $D_{SD}$ by -1.
-\hfill\break
-
-Unfortunately, applying the same logic to multidimensional trait data is more complicated, as sex-specific differences are not always guaranteed to align with some *a priori* direction in the multivariate dataspace.
-Thus, for multivariate datasets, one of two procedures may be used.
-First, for multivariate sets of traits that are measured on the same linear scale (e.g., length, width, and height measures), one could envision a hyperdimensional version of the plot displayed in Fig.
-1A, but where the set of male traits are plotted against the set of female traits.
-In this construction, because the traits are quantified on the same measured scale, the 1:1 size vector representing no sexual dimorphism does exist, and emanates from the origin of the space extending in a positive direction as the trait values increase.
-To determine whether male trait values deviate more from this vector than do female trait values, a partial least squares analysis may be performed [@RohlfCorti2000].
-Partial least squares (PLS) is a multivariate association procedure which describes the covariation between sets of traits (in this case, between the set of male traits and the set of female traits).
-Scores along the first PLS axis represent the maximal covariation between male and female traits projected to a single dimension, and species whose male and female PLS scores are identical are those displaying no sexual dimorphism.
-Thus, deviations from that pattern represent instances of multivariate sexual dimorphism.
-By plotting these scores, one may determine which species display female scores (on $PLS_1$) that are larger than their corresponding male scores (when negative values on $PLS_1$ correspond to smaller individuals; otherwise, the converse comparison is used).
-The $D_{SD}$ for these species may then be multiplied by -1 to obtain the $SDDI$ for that species.
-Note that when PLS is used on univariate data, this procedure obtains values identical to those from the method described previously where male and female sizes are compared directly.
-Thus, for multivariate traits measured on a linear scale, this is a direct generalization of the standard univariate procedure (see empirical example below).
-\hfill\break
-
-Alternatively, one might be interested in examining multidimensional traits other than linear biometric measurements, such as color or shape.
-For instance, one may characterize the shape of males and females using geometric morphometric methods [@Bookstein1991; @MitteroeckerGunz2009; @Adams_et_al2013], where the locations of anatomical landmarks and semilandmarks are used to obtain a set of shape variables.
-Geometric morphometric methods also result in a multidimensional characterization of phenotypic traits, from which the degree of sexual dimorphism for each species may be quantified using $D_{SD}$.
-However, because the resulting shape variables do not have a one to one correspondence with linear size measures, there is no *a priori* direction in the multivariate shape space that is equivalent to the 1:1 size vector found with sets of linear traits.
-In fact, for the case of isometry (i.e., shape does not change as size changes), all observations are represented by a single location in the shape space.
-Thus, while a PLS of male shapes versus female shapes will describe the covariation between them, this direction is not guaranteed to align with size, because there is no direction in the shape space that may be used to unambiguously determine whether dimorphism is male- or female-biased, as for shape variables there is no *a priori* directionality.
-Nonetheless, one may still quantitatively determine whether the degree of sexual shape dimorphism covaries with overall body size, by using the sexual dimorphism index as defined by the dimorphism distance (i.e., $SDDI=D_{SD}$) to represent the degree of multivariate shape dimorphism for each species.
-\hfill\break
-
-Once the sexual dimorphism distance index ($SDDI$) is calculated, it may be evaluated against the average size of each species.
-Here, data conforming to Rensch's rule will result in significant trends of sexual dimorphism relative to mean body size (Fig.
-1D).
-As with size ratios, the line of no sexual dimorphism is a horizontal line, but this time it is centered on zero, as a distance of zero corresponds to no sexual dimorphism.
-Thus, a phylogenetic regression of these data whose slope is significantly different from zero (e.g., $\beta_{1}\neq0.0$) provides evidence that the evolution of sexual dimorphism is consistent with patterns expected under Rensch's rule.
-For cases where traits are measured on a linear scale, patterns expected under Rensch's rule exhibit a slope significantly greater than zero ($\beta_{1}>0.0$), as in the usual formulation.
-Likewise, when using geometric morphometric shape data (or any other type of mean-centered multidimensional data), hyperallometric patterns consistent with Rensch's rule would also display a slope significantly greater than zero ($\beta_{1}>0.0$), indicating that the degree of sexual shape dimorphism increases with increasing body size.
-Conversely however, hypoallometric patterns would display the opposite pattern ($\beta_{1}<0.0$), indicating that smaller species display greater sexual shape dimorphism.
-\hfill\break
-
-## *Simulations*
-
-To ascertain the statistical performance of the approach proposed here, we conducted a series of computer simulations.
-We evaluated the ability of SDDI to characterize patterns of sexual dimorphism in both univariate and multivariate datasets.
-For each simulation, we generated a pure-birth phylogeny containing 50 species, on which we simulated datasets under a Brownian motion model of evolution.
-For univariate data ($p=1$), a single trait was evolved along the phylogeny, and represented the female trait values ($x$).
-Next, male values ($y$) were simulated in such a manner as to generate a known male:female relationship (plus random error), using the function: $y=\alpha{x}+\mathcal{N}(\mu=0,\sigma=0.1)$.
-For type I error simulations, no sexual dimorphism was generated, meaning that the slope, $\alpha$, characterized a 1:1 relationship between male and female trait values (i.e., $\alpha=1.0$).
-For power simulations, we simulated male values with a slope greater than 1.0 to correspond with patterns following Rensch's rule (i.e., $\alpha=1.1$).
-A total of 1,000 datasets were generated under each simulation condition.
-We then evaluated all datasets using standard approaches (i.e., a phylogenetic regression of male values versus female values), as well as a phylogenetic regression of SDDI versus average body size.
-The proportion of significant datasets was treated as the type I error ($\alpha=1.0$) or power ($\alpha=1.1$) respectively.
-\hfill\break
-
-For multivariate data ($p=5$), a slightly different procedure was utilized.
-First, average body size ($s$) was evolved along the phylogeny under Brownian motion.
-Next, female traits ($\mathbf{X}$) were simulated on the phylogeny under Brownian motion, using a $p\times{p}$ trait covariance matrix with correlations between traits of ($r=0.7$).
-For simulations where there was no pattern of sexual dimorphism across species (i.e., type I error simulations) male trait values were then generated by incorporating random error to the female trait values, in a manner analogous to that achieved in the univariate simulations: $\mathbf{Y}=\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1$).
-For simulations where patterns of sexual dimorphism covaried with body size (i.e., power simulations), male trait values were generated by incorporating random error to the female trait values, but with the addition that female trait values were pre-multiplied by a scalar value of body size: $\mathbf{Y}=(0.1*s)*\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1$).
-This ensured that male and female trait values covaried with one another as expected, and also covaried with size in the desired direction.
-As before, a total of 1,000 datasets were generated under each simulation condition.
-We then evaluated all datasets using a phylogenetic regression of SDDI versus average body size.
-The proportion of significant datasets was treated as the type I error ($\alpha=1.0$) or power ($\alpha=1.1$) respectively.
-\hfill\break
-
-*Simulation Results*: For the univariate dataset, we found that both a phylogenetic regression of male size on female size, as well as a phylogenetic regression of $SDDI$ on body size, resulted in appropriate type I error rates near the nominal $\alpha\approx0.05$ (Fig.
-2).
-Further, when data were simulated under conditions consistent with Rensch's rule, both procedures were capable of detecting the signal of the pattern, and did so with very high power.
-Thus, using the $SDDI$ approach proposed here was equivalent to what is normally utilized for univariate datasets.
-Likewise, for multivariate data, the $SDDI$ approach displayed appropriate type I error ($\alpha\approx0.05$) and high statistical power under the conditions simulated (Fig.
-2).
-Additionally, multiplying $D_{SD}$ by $\pm1$ had little effect on the outcome for multivariate data, as in both cases type I error and power results were virtually identical.
-Thus, methods for evaluating sexual dimorphism using geometric morphometric data are anticipated to remain robust.
-Overall results from these simulations imply that using the sexual dimorphism distance index ($SDDI$) proposed here provides an appropriate means of characterizing patterns of sexual dimorphism in both univariate and multivariate traits, and that the procedure is capable of detecting patterns consistent with Rensch's rule when they are present in phenotypic datasets.
-\hfill\break
-
-## *Empirical Example*
-
-To illustrate the utility of the approach presented here, we conducted an analysis of both size and shape dimorphism in a group of Mediterranean lizards with extensive body size variation.
-Our dataset consisted of three data types: male and female values of body size (snout-vent length: SVL), male and female values for a series of head measurements (head length, head width, head height, pileus length, and mouth opening: Fig.
-3A), and male and female values of head shape (Fig.
-3B).
-A total of 374 specimens from 21 lineages of green lizards, a monophyletic clade that includes two genera (*Lacerta* and *Timon*), were obtained from natural history museum collections.
-For each lineage, we obtained samples distributed across the entire distribution range, resulting in an initial database of 1062 specimens.
-In order to take sampling bias into account, we selected the 10 largest individuals per sex, when sufficient specimens were available, for use in subsequent size and shape analyses.
-A detailed account of the specimens used for this study is available in the Supplemental Material.
-\hfill\break
-
-From these specimens, we calculated mean body size for each sex and lineage.
-Likewise, we calculated the mean trait values for each sex and lineage for the set of linear measures.
-These values were then standardized by SVL; resulting in a set of head measurements proportional to body size for each sex $\times$ lineage combination.
-Additionally, we quantified lateral head shape variation using geometric morphometric methods [@Adams_et_al2013].
-This was accomplished by first digitizing 17 landmarks and semilandmarks from images of the right side of the head (Fig.
-3B).
-The thin-plate spline was used to estimate the position of missing landmarks [@Gunz2009], and the position of the jaw relative to the skull was standardized using the fixed angle method [@Adams1999].
-Next, a generalized Procrustes analysis  [@RohlfSlice1990] was performed to eliminate non-shape variation from the landmark data and align the specimens to a common coordinate system.
-The position of semilandmarks (i.e.
-11-16, Fig.
-3B) was optimized by minimizing the bending energy [@Bookstein1997].
-From the aligned Procrustes coordinates the mean head shape for males and for females was then calculated for each lineage.
-Finally, the phylogenetic relationships among lineages were estimated using published sequences on five mitochondrial and four nuclear gene fragments, totaling 8383 bp.
-These were then used following standard phylogenetic inference procedures to generate a dated consensus phylogeny for the lineages of interest (for details see Supplemental Material).
-\hfill\break
-
-Using these data, we conducted a series of analyses to investigate whether patterns of sexual size dimorphism (SSD) and sexual shape dimorphism (SShD) in this group corresponded to expectations under Rensch's rule.
-First, we examined patterns of SSD using a phylogenetic regression of male size against female size.
-Here, a slope significantly greater than 1.0 ($\beta>1.0$) was treated as evidence consistent with Rensch's rule.
-Additionally, we evaluated SSD patterns using the approach developed here, via a phylogenetic regression of the sexual size dimorphism index against mean body size for each species.
-Second, we evaluated multivariate patterns of sexual dimorphism in head proportions in relation to body size, using the sexual dimorphism index obtained from the Euclidean distance between male and female means from each species ($D_{SD}$).
-Partial least squares was used to determine which species displayed female scores (on $PLS_1$) larger than their corresponding male scores, and the sexual dimorphism index was then obtained by multiplying the sexual dimorphism distance by +1 or -1, respectively (i.e., $SDDI=\pm1*D_{SD}$).
-We then performed a phylogenetic regression of $SDDI$ for head proportions versus mean body size to determine whether patterns of dimorphism corresponded to what is expected under Rensch's rule.
-Third, we obtained the degree of sexual dimorphism in head shape for each species using the Procrustes distance between male and female means, and from this calculated the sexual shape dimorphism index ($SShDI$) in their original units as described above (i.e., $SShDI=D_{SD}$).
-We then performed a phylogenetic regression of sexual shape dimorphism ($SShDI$) against the mean body size for each species to determine whether patterns of head shape dimorphism were consistent with predictions from Rensch's rule.
-Additionally, patterns of head shape dimorphism were further investigated using principal components analysis of the set of mean male and female head shapes across species.
-Finally, we used phylogenetic correlation [phylogenetic partial least squares: sensu Adams and Felice -@AdamsFelice2014] to determine whether the degree of head shape dimorphism was associated with levels of size dimorphism.
-All analyses were performed in R 3.6.2 [@R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018], and routines written by the authors.
-\hfill\break
-
-*Empirical Results*: Using phylogenetic regression, we found a significant relationship between male size and female size, with a slope significantly greater than one ($\beta = 1.39$, $T=2.91$, $P = 0.009$: Fig.
-3C).
-This indicated that the degree of sexual size dimorphism differed systematically with overall body size; a pattern consistent with that expected under Rensch's rule.
-In this case, both hypoallometry and hyperallometry [*sensu* @Abouheif1997] were displayed (Fig.
-3C).
-Specifically, in male-biased species, the degree of sexual dimorphism was greater with increasing body size, and in female-biased species the degree of sexual dimorphism increased in smaller species (though the pattern was more pronounced in male-biased species).
-Similarly, when using the sexual size dimorphism index (SSDI), we found a significant and positive evolutionary relationship between SSDI and mean body size, confirming the Rensch's rule pattern identified using the commonly-used procedure ($\beta = 0.43$, $F=22.95$, $Z=1.89$, $P = 0.004$: Fig.
-3D).
-\hfill\break
-
-When multivariate patterns of sexual dimorphism were evaluated, we found that the degree of dimorphism in head proportions increased as body size increased: a pattern consistent with Rensch's rule ($F=7.20$, $Z=1.41$, $P = 0.022$: Fig.
-3E).
-Likewise, when patterns of head shape dimorphism were examined, we found that the degree of sexual shape dimorphism (SShDI) increased in larger species, and a phylogenetic regression of sexual shape dimorphism versus mean body size was found to be significant ($F=6.05$, $Z=1.35$, $P = 0.03$: Fig.
-3F).
-Thus, for both multivariate datasets, the degree of sexual dimorphism was enhanced in larger species [hyperallometry; sensu @Abouheif1997]: a pattern consistent with what was expected under Rensch's rule.
-To our knowledge, this is the first empirical demonstration of a pattern following Rensch's rule in multivariate data.
-\hfill\break
-
-Visualizing patterns of head shape dimorphism using principal components analysis (Fig.
-4), it was evident that males and females occupied distinct regions of morphospace.
-Further, the head shapes of males and females of the smallest lineage, *Lacerta agilis bosnica*, were distinct from those of the largest species, *Timon lepidus*.
-When connected, the sexual shape dimorphism vectors for these two species did not coincide with one another in the ordination, implying a difference in sexual shape dimorphism across species.
-Specifically, the vector for *T.
-lepidus* was longer than that for *L.
-agilis bosnica*, which reflected the greater degree of sexual shape dimorphism exhibited by this species.
-Thin-plate spline deformation grids facilitated a graphical description of these shape differences, where sexual shape dimorphism in *L.
-agilis bosnica* was best described as a contrast between longer-headed males, with an amplified tympanic area; and short-headed females, with a relatively more reduced tympanic region (Fig.
-4, left).
-By contrast, sexual shape dimorphism in *T.
-lepidus* was both more prominent in the tympanic area, similarly to what was observed in *L.
-agilis bosnica*; and also included a modification of the relative lower jaw area.
-Indeed, males of *T.
-lepidus* were characterized by both a longer posterior head region and by deeper lower jaw configurations as compared to females of the same species.
-\hfill\break
-
-Finally, across taxa, the degree of body size and head shape dimorphism were significantly correlated with one another ($r=0.73$, $p=0.0001$), but not when evaluated in a phylogenetic context ($r_{PPLS} = 0.361$, $Z=1.28$, $P = 0.12$).
-Part of this difference may be explained by the fact that the largest species (i.e.
-members of the genus *Timon*) - which displayed the greatest degree of both size and shape dimorphism - form a clade within the broader phylogeny for the group (Fig.
-5).
-Thus, despite the strong association of size and shape dimorphism, there were fewer changes of these patterns across branches of the phylogeny.
-Generally speaking, the degree of size dimorphism and the degree of shape dimorphism increased in larger taxa, and heat maps of the evolution of size and shape dimorphism across the phylogeny indicated a strong association between the two; particularly in large taxa (Fig.
-5).
-Thus, in this system, it appeared that evolutionary changes in size between the sexes were accompanied by concomitant changes in head shape.
+All analyses were performed in R 3.6.2 [@R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018]. \hfill\break
 
 # Discussion
+
+Using the estimated lambda values from pgls are not useful. The questions of whether or not signal exists is appropriate, but inferring more from lambda magnitude is inappropriate. \hfill\break
 
@@
 
-\textbf{Figure 4}.
-Principal components plot of male (black squares) and female (red points) species mean head shape: the first two PC axes explain over 60% of the total shape variation.
-Green lines correspond to the sexual shape dimorphism in one of the smallest species, *Lacerta agilis bosnica*, and in the largest species, *Timon lepidus*.
-Thin-plate spline deformation grids display the mean head shape for males and females of these species.
-\hfill\break
-
- \textbf{Figure 5}.
-Heat maps (green: lower values; red: higher values) displaying the evolution of (A) sexual size dimorphism, and (B) sexual shape dimorphism across the phylogeny.
-For each panel, an index of sexual size dimorphism ($SSDI$) or sexual shape dimorphism ($SShDI$) is displayed, with circles being proportional to the degree of sexual dimorphism.
-\hfill\break
-
 \newpage
@@
 ```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+include_graphics("Figures/Fig1.png")
 
@@
 
-\singlespacing \textbf{Figure 1}.
-(A) Sexual size dimorphism for males and females across 100 hypothetical species.
-The pattern displayed exhibits a slope > 1.0, and thus corresponds to Rensch's rule.
-(B) The same data represented as the ratio of $M/F$ sizes plotted against average species size.
-(C) Sexual dimorphism represented as the distance between males and females: $SDD = D_{M:F}$.
-(D) Sexual dimorphism represented as an index of the SD Distance: $SDDI = \pm1*D_{M:F}$.
-In all panels,  the dashed line corresponds to values representing no sexual dimorphism (the 1:1 line).
-\hfill\break
+\singlespacing \textbf{Figure 1}.  \hfill\break
 
@@
 ```{r, echo = FALSE, out.width="70%"}
-include_graphics("Fig2.png")
+include_graphics("Figures/Fig2.png")
 
@@
 
-\singlespacing \textbf{Figure 2}.
-Results of phylogenetic simulations testing the type I error and statistical power for detecting patterns consistent with Rensch's rule.
-Vertical bars represent the proportion of datasets found to be significant.
-For type I error analyses, the dashed line marks the standard threshold of $\alpha= 0.05$.
-For univariate data, both phylogenetic regression of male versus female size (standard approach), and $SDDI$ versus overall body size (new approach) were performed.
-For multivariate data, only $SDDI$ versus overall body size was examined.
-\hfill\break
+\singlespacing \textbf{Figure 2}.  \hfill\break
 
@@
 ```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig3.png")
+include_graphics("Figures/Fig3.png")
 
@@
 
-\singlespacing \textbf{Figure 3}.
-(A) Set of linear measurements used to obtain head proportions.
-HL: head length, HW: head width, HH: head height, PL: pileus length, MO: mouth opening.
-(B) Locations of 17 landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head shape.
-(C) Patterns of sexual size dimorphism represented as male versus female size.
-(D) Patterns of sexual size dimorphism, represented as sexual size dimorphism index ($SSDI$) versus mean size.
-(E) Multivariate patterns sexual dimorphism in body proportions ($SShDI$) versus mean size.
-(F) Multivariate patterns of sexual shape dimorphism ($SShDI$) versus mean size.
-In panels C, D, E, and F, the red line indicates the regression estimate obtained from the data, while the dashed line corresponds to values representing no sexual dimorphism.
-\hfill\break
-
-\newpage
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig4.png")
-
-```
-
-\singlespacing \textbf{Figure 4}.
-Principal components plot of male (black squares) and female (red points) species mean head shape: the first two PC axes explain over 60% of the total shape variation.
-Green lines correspond to the sexual shape dimorphism in one of the smallest species, *Lacerta agilis bosnica*, and in the largest species, *Timon lepidus*.
-Thin-plate spline deformation grids display the mean head shape for males and females of these species.
-\hfill\break
-
-\newpage
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig5.png")
-
-```
-
-\singlespacing \textbf{Figure 5}.
-Heat maps (green: lower values; red: higher values) displaying the evolution of (A) sexual size dimorphism, and (B) sexual shape dimorphism across the phylogeny.
-For each panel, an index of sexual size dimorphism ($SSDI$) or sexual shape dimorphism ($SShDI$) is displayed, with circles being proportional to the degree of sexual dimorphism.
-\hfill\break
+\singlespacing \textbf{Figure 3}.  \hfill\break

### 2020-03-26 Data_Analyses/Simulation_script-PB-ANOVA.R

@@ -98,19 +98,7 @@ for (i in 3:length(lambdas_expanded)) {
 DataTableMat_lambda_ml <- as.data.frame(DataTableMat_lambda_ml)
 DataTableMat_lambda_ml$beta <- rep(beta,(length(lambdas)*nsim))
 
-
-
-
-
-
-
-
 plot(DataTableMat_lambda_ml$lambda.est ~ DataTableMat_lambda_ml$lambda.input, pch = 19)
-
-
-
-
-
 
 # writing files
 file_name <- paste("Data_Analyses/Sim_Data/PB_lambda_ml-", n, "-ANOVA.csv", sep = "")
@@ -118,40 +106,3 @@ write.csv(DataTableMat_lambda_ml, file_name, row.names = F)
 
 file_name <- paste("Data_Analyses/Sim_Data/PB_lambda_0-", n, "-ANOVA.csv", sep = "")
 write.csv(DataTableMat_lambda_0, file_name, row.names = F)
-
-
-# 32 
-# 64 
-# 128  
-# 256 
-# 512 
-# 1024 
-
-data <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512-ANOVA.csv") # run another example where you make lmanba input onto y instead of x
-# go reread liams paper on lambda input on both x and y and both
-data <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32.csv") # this is producing same correlation, fix
-
-data_beta <- subset(data, data$beta == 0)
-
-plot(data_beta$lambda.est ~ data_beta$lambda.input, pch = 19)
-
-cor(data_beta$lambda.est, data_beta$lambda.input)
-
-
-plot(data$slope ~ data$beta, pch = 19)
-abline(a=c(0,1)) # parameter estimate in the phylo regression outputs an expected value, with broad range of accuracy. 
-# that distribution decreases with increasing phylo size
-# look at how the range mean std of slope estimates changes with ml vs 0 lambda inputs
-# incorporating lambda does not affect inference in the 
-# error rates (x axis = f in pgls lambda 0, y axis = f in pgls labmda ml)
-
-data0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32.csv")
-dataml <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32.csv")
-
-plot(dataml$'F'~data0$'F')
-plot(dataml$'P'~data0$'P')
-plot(dataml$slope~data0$slope)
[…]

### 2020-03-26 Data_Analyses/Simulation_script-PB-Reg.R

@@ -17,7 +17,7 @@ beta <- c(0, .25, .5, .75, 1) # true slope
 fullsimlength <- length(lambdas_expanded)
 
 
-n <- treesizes[5] # number of tips (not simulations)
+n <- treesizes[5] # number of tips: adjust each iteration
 
 DataTable_lambda_0 <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
                    dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
@@ -96,16 +96,7 @@ for (i in 3:length(lambdas_expanded)) {
 DataTableMat_lambda_ml <- as.data.frame(DataTableMat_lambda_ml)
 DataTableMat_lambda_ml$beta <- rep(beta,(length(lambdas)*nsim))
 
-
-
 plot(DataTableMat_lambda_ml$lambda.est ~ DataTableMat_lambda_ml$lambda.input, pch = 19)
-
-
-
-
-
-
-
 
 # writing files
 file_name <- paste("Data_Analyses/Sim_Data/PB_lambda_ml-", n, ".csv", sep = "")
@@ -113,13 +104,3 @@ write.csv(DataTableMat_lambda_ml, file_name, row.names = F)
 
 file_name <- paste("Data_Analyses/Sim_Data/PB_lambda_0-", n, ".csv", sep = "")
 write.csv(DataTableMat_lambda_0, file_name, row.names = F)
-
-
-# 32 done
-# 64 done
-# 128 done
-# 256 done
-# 512 
-# 1024 
-
-plot(DataTableMat_lambda_ml$slope ~ DataTableMat_lambda_ml$beta, pch = 19)

### 2020-03-26 Figures/Figures_Script.R

@@ -1,0 +1,70 @@
+# Figures
+
+data_reg_32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32.csv")
+data_reg_64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-64.csv")
+data_reg_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-128.csv")
+data_reg_256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-256.csv")
+data_reg_512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512.csv")
+data_reg_1028 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1028.csv")
+
+data_anova_32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32-ANOVA.csv")
+data_anova_64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-64-ANOVA.csv")
+data_anova_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-128-ANOVA.csv")
+data_anova_256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-256-ANOVA.csv")
+data_anova_512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512-ANOVA.csv")
+data_anova_1028 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1028-ANOVA.csv")
+
+data_reg_32_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32.csv")
+data_reg_64_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-64.csv")
+data_reg_128_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-128.csv")
+data_reg_256_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-256.csv")
+data_reg_512_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-512.csv")
+data_reg_1028_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1028.csv")
+
+data_anova_32_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32-ANOVA.csv")
[…]

### 2020-03-30 Manuscript/Manuscript.Rmd
All tree simulations completed except the 1028, still running on the lab server.
Rough intro of manuscript developed slightly.

--- 2020-03-26 Manuscript.Rmd
@@
 Investigating macroevolutionary patterns requires a phylogenetic approach as species are non-independent by nature of their shared ancestry.
-Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; [@Felsenstein1987]), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
-[@AdamsNason2019]).
+Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; [Felsenstein1985]), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
+[@Grafen1989] [@HarveyPagel1991] [@Rohlf2001]  [@AdamsNason2019]).
 Understanding the degree of phylogenetic signal present in a dataset is paramount, and identifies the mode with which a trait has evolved.
@@
 For continuous data, the most common parameters used in the literature include Pagel's lambda [@Pagel1997] and Blomberg's kappa [@Blomberg1900].
-Pagel's lambda has the advantage of being couched in the likelihood framework and thus has also been utilized to encorporate phylogenetic signal while doing phylogenetic regressions and ANOVA.
-However, the accuracy of the lambda estiamtion methods have not been fully evaluated, and thus it remains unknown the degree to which lambda estimates appropriately represent degree of phylogenetic signal.
+Pagel's lambda has the advantage of being nested in the likelihood framework and thus has also been utilized to encorporate phylogenetic signal while doing phylogenetic regressions and ANOVAs.
+At the development of these estimation methods, clear recommendations on the use and interpretation of this parameter were made based on  data simulations and error evaluations [@Revell2010].
+However, the accuracy of the lambda estiamtion methods have not been fully evaluated, and consequently many studies have substantially misused these methods.
 \hfill\break
 
-An earlier study[@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies could be using  estimation methods for various parameters.
-That paper concluded that a measure of power must be considered when quantifying Pagel's lambda.
-Here we take a more comprehensive approach to demonstrate the scenarios under which estimated lambdas accurately reflect known lambdas as well as the effect of these at times dubious estimation methods on significance testing when used in a pgls framework.
+An earlier study [@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies could be for  estimating various evolutionary parameters, such as Pagel's lambda and ____.
+However, it remains unknown the degree to which lambda estimates appropriately represent degree of phylogenetic signal, and thus we fully explicate the extent to which Pagel's lambda estimation methods have been misused in recent macroevolutionary studies.
+Here we take a comprehensive approach to demonstrate the scenarios under which estimated Pagel's lambdas accurately reflect known lambdas.
+We then demonstrate the effect of these at times dubious estimation methods on significance testing when used in a pgls framework.
+Finally, we perform a meta-analysis of research published in 2019 employing Pagel's lambda estimation methods to demonstrate the uses and misuses of this method in the literature.
+This investigation provides evidence for the circumstances under which estimating Pagel's lambda can be informative, and which interpretations or uses are misleading with the hope of moving the field towards more appropriate employment of our methodological toolbox.
 
@@
 
-All analyses were performed in R 3.6.2 [@R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018]. \hfill\break
+All analyses were performed in R 3.6.2 [R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018]. \hfill\break
 
@@
 
-Using the estimated lambda values from pgls are not useful. The questions of whether or not signal exists is appropriate, but inferring more from lambda magnitude is inappropriate. \hfill\break
+Using the estimated lambda values from pgls are not useful. The questions of whether or not signal exists is appropriate, but inferring more from lambda *magnitude* is inappropriate. \hfill\break
 
-A longstanding question in evolutionary biology is understanding why sexual dimorphism, the degree of differentiation between the sexes, varies across species.
-A frequently used schema for exploring macroevolutionary patterns of sexual dimorphism is Rensch's rule, which predicts that sexual dimorphism should increase with body size in species where males are the larger sex, and decrease with increasing body size in species where females are the larger sex.
-Remarkably, virtually all recent studies have focused on macroevolutionary variation in sexual *size* dimorphism, despite the relevance of sexual dimorphism in other traits, as encompassed by Rensch's original treatment [@Rensch1950].
-This gap has been driven, in part, by the lack of adequate tools for evaluating Rensch's rule in multivariate traits, such as color, ornament complexity, or shape.
-In this article, we fill this procedural gap by providing an analytical equivalency that allows the generalization of statistical tests of Rensch's rule for both univariate and multivariate traits.
-The approach describes the degree of sexual dimorphism as a distance between male and female trait values in phenotype space, thereby facilitating its use on either single-valued traits (e.g., body size), or complex, multivariate traits (e.g.
-organismal shape).
-Computer simulations confirm that the approach exhibits appropriate type I error rates and statistical power, thereby providing an adequate procedure for testing whether empirical data comply with the predictions of Rensch's rule.
-The implementation of our approach on an empirical study of Mediterranean lizards revealed that patterns of sexual dimorphism comply with Rensch's rule not only for body size, but also for relative head proportions and head shape, offering new insights and triggering novel hypotheses about the possible proximate and evolutionary causes that may underlie such patterns.
-\hfill\break
-
-Our empirical example in Mediterranean green lizards exemplifies a case study investigating variation in sexual dimorphism -- not only in size -- but also in multivariate phenotypes, like shape, and how this method can provide useful insights into the proximate and evolutionary causes underlying such patterns.
-First, we found that the degree of sexual size dimorphism increased as taxa became larger (Fig.
-3C, D).
-This pattern is compliant with Rensch's rule and corresponds with hyperallometry [@Abouheif1997], which is to be expected when males are the larger sex [@Fairbairn1997].
-In this group males are larger than females in most lineages (Fig.
-3C), as is commonly observed in lacertids [@Brana1996].
-Additionally, as in many lizard species, sexual selection in this group is primarily mediated through territoriality, where male-male competition selects for a larger male body size [@Stamps1983].
-Thus, in this system there is an association between male-biased sexual size dimorphism and male-male competition.
-As such, the patterns of size dimorphism observed in this study are consistent with those expected under a scenario where sexual selection drives the evolution of larger male body size across taxa via male-male competition.
-\hfill\break
-
-Likewise, our multivariate analyses revealed that head morphology differed between the sexes, and that the degree of sexual dimorphism in multivariate head shape also corresponded to that expected under Rensch's rule (Fig.
-3E, 3F).
-That is, while both small (e.g.
-*L.
-a.
-bosnica*, Fig.
-4, left) and large (e.g.
-*T.
-lepidus*, Fig.
-4, right) species are sexually dimorphic in head shape, the degree of sexual shape dimorphism increased with average body size (Fig.
-3F); a pattern consistent with a scenario of hyperallometry.
-Examination of the resulting shape differences revealed that sexual dimorphism in head shape was associated with an amplification of the posterior head region, and also in the relative size of the lower jaw (in the case of *T.
-lepidus*).
-Biomechanically, these are precisely the anatomical regions where the jaw adductor muscles reside [@Haas1973], and as such, an amplification of these areas in males would enhance biting performance [@Herrel1996; @Kaliontzopoulou2012]; a functional trait that has important implications for male competitive capacity [@Lailvaux2007; @HUYGHE2005].
-In addition, one might predict that the need for stronger jaw musculature in males becomes disproportionately more intense with increasing body size, as male combats are expected to be more fierce among larger-bodied opponents.
-Thus, our findings that head shape dimorphism increases with overall body size is also consistent with the hypothesis that sexual selection on structures related to territoriality and male competitive interactions are responsible for the macroevolutionary patterns of sexual shape dimorphism observed in this group.
-To our knowledge, this is the first example of Rensch's rule in a multivariate phenotype.
-\hfill\break
-
-With respect to the analytical approach proposed here, our characterization of sexual dimorphism via a *shape distance* and the $SDDI$ statistic is completely general, and may be used to evaluate evolutionary trends in any continuous, quantitative trait that differs between the sexes.
-Our empirical example highlighted the efficacy of the approach for evaluating trends in univariate traits such as body size, and also several types of complex data where organismal phenotypes are quantified by multiple dimensions.
-As such, the approach developed here expands our current analytical toolkit to provide a means of evaluating evolutionary trends in sexual dimorphism for a broader class of phenotypic traits than has previously been considered.
-Indeed, this advance more closely embodies Rensch's [-@Rensch1950] original conceptualization for understanding patterns of sexual dimorphism, as his broader vision was to explain why phenotypic differences *writ large* exist among the sexes.
-Thus, use of the $SDDI$ will facilitate a greater degree of evolutionary exploration into patterns of sexual dimorphism than has been possible in recent decades.
-Further, with the inclusion of multivariate traits, the formulation proposed here opens the door for investigations into the evolution of sexual shape dimorphism, and whether such patterns correspond to those expected under Rensch's rule.
-Thus, the mathematical test proposed here is general, and may be used for any quantitative phenotypic trait under scrutiny.
-\hfill\break
-
-Finally, because $SDDI$ is derived from a phenotypic distance, one could, in theory, utilize the approach to characterize patterns of sexual dimorphism in non-continuous traits, and determine quantitatively whether the degree of sexual dimorphism is enhanced with increasing body size.
-Across the tree of life, major phenotypic adaptations are often described as evolutionary innovations or novelties [sensu @Wagner1996a; @Wagner2002; @Peterson2016], where these changes are characterized as discrete shifts from one phenotypic state to another [e.g., @Holliday2012; @OKeefe2011].
-In like manner, discrete phenotypic differences are known to exist between the sexes; for instance where males display horns or other ornamentation [@Geist1988; @Emlen2008] that females lack.
-Over evolutionary time, a series of such discrete phenotypic differences may accumulate in males relative to females, resulting in an elaboration of phenotypic differences between the sexes in suites of such traits.
-With an appropriate distance measure for discrete traits (e.g., Hamming distance), the approach developed here is capable of characterizing such patterns, and determining whether sets of discrete phenotypic differences are enhanced with increasing body size among species.
-Thus, our approach is the first to our knowledge to facilitate a formal evaluation of both discrete and continuous phenotypic differences between the sexes, providing a comprehensive means of determining whether sets of phenotypic traits differ in a manner as predicted by Rensch's rule.
+More discussion paragraphs
 
@@
 \textbf{Figure 1}.
-(A) Sexual size dimorphism for males and females across 100 hypothetical species.
-The pattern displayed exhibits a slope > 1.0, and thus corresponds to Rensch's rule.
-(B) The same data represented as the ratio of $M/F$ sizes plotted against average species size.
-(C) Sexual dimorphism represented as the distance between males and females: $SDD = D_{M:F}$.
-(D) Sexual dimorphism represented as an index of the SD Distance: $SDDI = \pm1*D_{M:F}$.
-In all panels,  the dashed line corresponds to values representing no sexual dimorphism (the 1:1 line).
+Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
+As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 256 tips.
 \hfill\break
 
-\textbf{Figure 2}.
-Results of phylogenetic simulations testing the type I error and statistical power for detecting patterns consistent with Rensch's rule.
-Vertical bars represent the proportion of datasets found to be significant.
-For type I error analyses, the dashed line marks the standard threshold of $\alpha= 0.05$.
-For univariate data, both phylogenetic regression of male versus female size (standard approach), and $SDDI$ versus overall body size (new approach) were performed.
-For multivariate data, only $SDDI$ versus overall body size was examined.
-\hfill\break
+\textbf{Figure 2}. Figure 2 legend here\hfill\break
 
-\textbf{Figure 3}.
-(A) Set of linear measurements used to obtain head proportions.
-HL: head length, HW: head width, HH: head height, PL: pileus length, MO: mouth opening.
-(B) Locations of 17 landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head shape.
-(C) Patterns of sexual size dimorphism represented as male versus female size.
-(D) Patterns of sexual size dimorphism, represented as sexual size dimorphism index ($SSDI$) versus mean size.
-(E) Multivariate patterns sexual dimorphism in body proportions ($SShDI$) versus mean size.
-(F) Multivariate patterns of sexual shape dimorphism ($SShDI$) versus mean size.
-In panels C, D, E, and F, the red line indicates the regression estimate obtained from the data, while the dashed line corresponds to values representing no sexual dimorphism.
+\textbf{Figure 3}. (A) Figure 3 legend (B) Second part of legend. \hfill\break
+
+\newpage
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig1.png")
+
+```
+
+\singlespacing \textbf{Figure 1}.
+Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
+As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 256 tips.
 \hfill\break
@@
 
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Figures/Fig1.png")
-
-```
-
-\singlespacing \textbf{Figure 1}.  \hfill\break
-
-\newpage
-
-```{r, echo = FALSE, out.width="70%"}
-include_graphics("Figures/Fig2.png")
-
-```
-
-\singlespacing \textbf{Figure 2}.  \hfill\break
-
-\newpage
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Figures/Fig3.png")
-
-```
-
-\singlespacing \textbf{Figure 3}.  \hfill\break
+# Other Figures Here

### 2020-03-30 Manuscript/Manuscript.tex
All tree simulations completed except the 1028, still running on the lab server.
Rough intro of manuscript developed slightly.

--- 2020-03-15 Manuscript-template_EKB.tex
@@
 
-\section{A multivariate equivalency and general test for Rensch's Rule:
-Implications for the evolution of sexual shape
-dimorphism}\label{a-multivariate-equivalency-and-general-test-for-renschs-rule-implications-for-the-evolution-of-sexual-shape-dimorphism}
+\section{Pagel's Lambda Estimates are Often
+Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}
 
@@
 
-\textbf{Short Title}: Problems with Pagel's lambda \hfill\break
+\textbf{Short Title}: Inaccuracies in Pagel's Lambda \hfill\break
 
@@
 
-Rensch's rule, the allometric association between the degree of sexual
-dimorphism and overall body size, is an important framework for
-understanding the evolution of trends in phenotypic differentiation
-between the sexes across taxa. While virtually all recent treatments
-have explored the degree of sexual size dimorphism, understanding how
-dimorphism in other phenotypic traits evolves is equally relevant.
-Indeed, Rensch's original treatment focused not only on size, but also
-on dimorphism in relative body proportions, coloration, the presence of
-ornaments, etc. Here, we derive a multivariate equivalency for viewing
-trends in sexual dimorphism -- relative to overall body size -- across
-taxa, and provide a generalized test to determine whether such patterns
-are consistent with Rensch's rule. For univariate linear traits such as
-body size, our approach yields equivalent results to those from standard
-procedures, but our test is also capable of detecting trends in
-multivariate datasets like shape. Computer simulations reveal the method
-displays appropriate statistical properties, and an empirical example in
-Mediterranean lizards illustrates the efficacy of our approach on both
-univariate and multivariate phenotypes. Our generalized procedure
-substantially extends the toolkit available for investigating
-macroevolutionary patterns of sexual dimorphism and seeking a better
-understanding of the processes that underlie them.
+conclusion holds: interpreting the regression is not appreciably
+different (in terms of slopes and f values)
 
@@
 
-It is widely observed that males and females of the same species are not
-identical, but rather differ -- sometimes substantially -- in their
-phenotypic characteristics (Fairbairn 1997, 2013). Indeed, sexual
-dimorphism is pervasive across many animal clades, and understanding the
-causes for these patterns has attracted considerable attention in
-evolutionary biology for well over a century (e.g., Darwin 1871;
-Fairbairn et al. 2007). One of the most conspicuous differences between
-males and females is in their body sizes (Andersson 1994), where
-sex-specific differences are often hypothesized to be the result of
-selection pressures related to reproduction, or the distinct ecological
-roles of males and females in their respective habitat (Cox et al. 2003;
-Stephens and Wiens 2009; Kaliontzopoulou et al. 2015; Tarr et al. 2018;
-Littleford-Colquhoun et al. 2019). For example, both sexual selection on
-male body size (Cox et al. 2003; Garcia-Navas et al. 2015; Horne et al.
-2020), and fecundity selection on female body size (Serrano-Meneses and
-Szekely 2006; Stuart-Fox 2008), are expected to enhance the degree of
-sexual size dimorphism in taxa over evolutionary time. Likewise, when
-males and females respond differently to environmental factors,
-intraspecific competition may generate ecological divergence between
-them, resulting in varying patterns of sexual dimorphism (Butler et al.
-2000; Bolnick and Doebeli 2003; Dayan and Simberloff 2005;
-Kaliontzopoulou et al. 2010; Meiri et al. 2014). \hfill\break
+Investigating macroevolutionary patterns requires a phylogenetic
+approach as species are non-independent by nature of their shared
+ancestry. Since the first appropriate method was introduced by
+Felsenstein (phylogenetic independent contrasts; {[}Felsenstein1987{]}),
+dozens of other methods have been developed and applied to increasingly
+complex questions in macroevolutionary biology (e.g.
+{[}AdamsNason2019{]}). Understanding the degree of phylogenetic signal
+present in a dataset is paramount, and identifies the mode with which a
+trait has evolved. High measures of phylogenetic signal indicate a
+Brownian motion process, whereas lower levels of phylogenetic signal
+indicate natural selection or some other evolutionary force has
+influenced the traits evolutionary history. \hfill\break
 
-At macroevolutionary scales, changes in the relative contribution of the
-selective forces above may create variation in the degree of sexual
-dimorphism displayed across closely related taxa. For instance, sister
-species inhabiting different environments, or possessing distinct
-ecological adaptations, may exhibit differing degrees of sexual
-dimorphism, as a result of the varying intensity of sexual selection
-they endure (Östman and Stuart-Fox 2011). Likewise, the interplay
-between sexual and natural selection that species experience may result
-in sexual dimorphism of differing magnitudes across taxa (Blanckenhorn
-et al. 2006; Stuart-Fox and Moussalli 2007; Kaliontzopoulou et al.
-2015). One evolutionary framework that connects adaptive explanations
-for the evolution of sexual size dimorphism (SSD) with trends in overall
-body size across taxa is Rensch's rule. Here it is predicted that SSD
-will increase with species' average body size in species where males are
-the larger sex (hyperallometry), and likewise will decrease with
-increasing body size where females are the larger sex (hypoallometry)
-(Abouheif and Fairbairn 1997; Fairbairn 1997). Put another way, Rensch's
-rule predicts that the extent of sexual dimorphism increases at more
-extreme body sizes (both large and small) across taxa. Empirical studies
-have documented trends in SSD consistent with Rensch's rule in a wide
-variety of animal taxa (Dale et al. 2007; Ceballos et al. 2013; Regis
-and Meik 2017; Johnson et al. 2017). However, other clades display the
-converse trend (Burbrink and Futterman 2019; Peñalver-Alcázar et al.
-2019), and some clades display no relationship between SSD and body size
-(Astúa 2010; Hirst and Kiørboe 2014; Johnson et al. 2017). From an
-evolutionary perspective, the association between SSD and body size has
-been hypothesized to be driven by variation in ontogenetic or static
-allometries across taxa, which results from the influence of sexual
-selection on growth and maturation patterns (Rensch 1960; Dale et al.
-2007). \hfill\break
+Several approaches to quantify phylogenetic signal exist. For continuous
+data, the most common parameters used in the literature include Pagel's
+lambda {[}Pagel1997{]} and Blomberg's kappa {[}Blomberg1900{]}. Pagel's
+lambda has the advantage of being couched in the likelihood framework
+and thus has also been utilized to encorporate phylogenetic signal while
+doing phylogenetic regressions and ANOVA. However, the accuracy of the
+lambda estiamtion methods have not been fully evaluated, and thus it
+remains unknown the degree to which lambda estimates appropriately
+represent degree of phylogenetic signal. \hfill\break
 
-Interestingly, virtually all evaluations of Rensch's rule have
-investigated patterns of sexual size dimorphism, yet Rensch's original
-contributions were concerned with a much wider array of phenotypic
-traits that can differ between the sexes. These included not only sexual
-differences in body size, but also in general body proportions (e.g.,
-relative wing and limb lengths), the degree of complexity of different
-body structures, the presence of sex-specific ornamentation (e.g.,
-horns, body coloration, etc.), and other sex-specific differences
-(Rensch 1950, 1960). In like manner, some studies have documented
-patterns of sexual shape dimorphism (e.g., Berns and Adams 2013; Kelly
-et al. 2013; Kaliontzopoulou et al. 2015), sex-specific differences in
-ornamentation (Geist and Bayer 1988; Emlen 2008; Watson and Simmons
-2010), or coloration (e.g., Endler 1983). However, such studies have
-typically been restricted to one or a few taxa, and have not explained
-macroevolutionary variation in sexual dimorphism in a manner that
-relates directly to Rensch's rule. Thus, when viewed from this
-perspective, Rensch's (1950) vision was far more synthetic, as he sought
-to explain the extent to which sexual dimorphism -- across a general
-suite of phenotypic characteristics -- scaled with overall body size in
-organisms, and why some species displayed a greater degree of sexual
-dimorphism than did others. Remarkably, patterns of intraspecific static
-allometry in sexually selected traits establishes a direct link between
-variation in those traits relative to size, and the evolutionary
-allometry across taxa in the degree of sexual dimorphism that may be
-expected (see also Rensch 1960; Reiss 1986; Bonduriansky 2007). Under
-this framework, the ``exaggeration'' of body structures or other
-phenotypic traits with increasing body size is predicted to translate,
-at macroevolutionary scales, into increased degrees of sexual dimorphism
-with increasing body size. However, despite this clear prediction, and
-Rensch's original observations, the evolutionary scaling of the degree
-of sexual dimorphism in phenotypic traits other than size, remains
-largely unexplored. \hfill\break
-
-Another current limitation in our understanding of Rensch's rule is the
-fact that to date, all studies have been conducted exclusively on
-univariate data, primarily regarding body size. However, understanding
-phenotypic evolution is inherently a multivariate endeavor (Blows 2007;
-Collyer et al. 2015; Adams and Collyer 2018, 2019), as processes such as
-natural and sexual selection simultaneously operate on more than one
-trait (Lande 1979; Lande and Arnold 1983), and can affect variation in
-complex multidimensional traits such as shape. To interrogate such
-patterns, the analytics used to evaluate evolutionary trends in
-phenotypic datasets must be sufficiently general to accommodate this
-empirical fact. Unfortunately, all current analytical approaches for
-evaluating trends in sexual dimorphism as they relate to overall body
-size are explicitly univariate; thereby excluding the possibility of
-evaluating patterns in complex multidimensional traits such as
-organismal shape. Though is has been extensively documented in sexual
-size dimorphism, it remains unknown whether Rensch's rule applies
-equally to sexual shape dimorphism, as determining this requires an
-analytical framework capable of evaluating trends in sexual dimorphism
-in both univariate and multivariate datasets. \hfill\break
-
-In this paper, we derive a multivariate equivalency for inspecting
-variation in sexual dimorphism -- relative to overall body size --
-across taxa, and provide a generalized test to determine whether they
-are consistent with Rensch's rule. Our procedure is appropriate for
-evaluating trends in univariate traits measured on a linear scale (such
-as body size or relative limb length), as well as in multivariate
-datasets representing complex phenotypes (such as sets of body
-proportions or landmark-based shape data). For univariate data, the
-approach yields equivalent results to what is normally accomplished for
-size-based SSD evaluations. However, the approach extends these
-procedures for the examination of sexual dimorphism in multivariate
-data, thereby increasing the potential for our conceptual understanding
-of the consequences of sexual and natural selection on complex
-organismal phenotypes. We conduct a series of computer simulations
-demonstrating that the approach displays appropriate statistical
-properties when evaluating trends in both univariate and multivariate
-data. We also provide an empirical example in Mediterranean green
-lizards which illustrates the efficacy of our approach on both
-univariate and multivariate datasets. Finally, we highlight the new
-avenues of evolutionary research that our generalization facilitates,
-and make comments on the prospects for future empirical advances.
+An earlier study {[}BoettigerEtAl2012{]}, breifly addressed this topic
+by showing how uninformative smaller phylogenies could be using
+estimation methods for various parameters. That paper concluded that a
+measure of power must be considered when quantifying Pagel's lambda.
+Here we take a more comprehensive approach to demonstrate the scenarios
+under which estimated lambdas accurately reflect known lambdas as well
+as the effect of these at times dubious estimation methods on
+significance testing when used in a pgls framework.
 
@@
 
-\subsection{\texorpdfstring{\emph{Conceptual
-Development}}{Conceptual Development}}\label{conceptual-development}
+\subsection{\texorpdfstring{\emph{Simulated
+trait}}{Simulated trait}}\label{simulated-trait}
 
-To arrive at a general test for Rensch's rule that can accommodate
-multivariate data, we must first derive a series of mathematical
-equivalencies based upon how interspecific patterns of size dimorphism
-are typically evaluated. For sexual size dimorphism, interspecific
-patterns are often inspected using a bivariate plot, where the size of
-one sex is plotted against the size of the other sex (e.g., Fairbairn
-and Preziosi 1994; Abouheif and Fairbairn 1997; Ceballos et al. 2013).
-An idealized example is shown in Fig. 1A, with male size plotted against
-female size. The diagonal line represents no size dimorphism; or a 1:1
-size ratio between the sexes across the range of body sizes for species
-in the clade of interest. In this construction (i.e., with females
-plotted on the x-axis), deviations from this line represent differential
-trends in size dimorphism across taxa. To evaluate such patterns
-statistically, a phylogenetic regression is performed, and in this case,
-a slope significantly greater than 1.0 (i.e., \(\beta_{1}>1.0\)) would
-be treated as evidence that the evolutionary changes in sexual size
-dimorphism across taxa are consistent with Rensch's rule (Fairbairn
-1997). \hfill\break
+To assess the accuracy of Pagel's lambda estimations, we simulated
+pure-birth phylogenies with known lambdas. We \_\_\_\_ by scaling
+simulated phylogeny with the scaling parameter and . We also did this
+with other tree shapes (symmetrical and ladder). \hfill\break
 
-An equivalent representation of patterns of sexual dimorphism may also
-be found using ratios (e.g., male size:female size). When plotted
-against the average size for each species (Fig. 1B), the same
-information is conveyed as is found in the more commonly utilized male -
-female plot of Fig. 1A. However, in this representation, the line
-designating no sexual dimorphism is horizontal, with a slope of zero
-(i.e., \(\beta_{1}=0.0\)) and an intercept of 1.0. As before, a
-phylogenetic regression may be performed on these data, and in this
-case, if the slope is significantly steeper than zero (i.e.,
-\(\beta_{1}>0.0\)), patterns consistent with Rensch's rule are observed.
-\hfill\break
+\subsection{\texorpdfstring{\emph{Simulated ANOVA and
+Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}
 
-A third equivalency is shown in Fig. 1C. Here, the Euclidean distance
-between male and female means for each species
-(\(D_{SD}=\sqrt{(Y_{male}-Y_{female})^2}\)) is shown relative to the
-overall size of the species. This plot is similar to a Bland-Altmann
-plot (Altman and Bland 1983), except that the Euclidean distance, rather
-than a difference score, is plotted along the y-axis (as distances can
-accommodate data of different dimensionality). Under Rensch's rule, this
-\emph{sexual dimorphism distance} may become greater as species are
-progressively larger (hyperallometry, sensu Abouheif and Fairbairn 1997)
-or may become greater as species are progressively smaller
-(hypoallometry, sensu Abouheif and Fairbairn 1997). In both cases,
-patterns correspond to species displaying progressively more sexual
-dimorphism at extreme body sizes; a trend consistent with Rensch's rule.
-Note that the inflection point in Fig. 1C represents an intermediate
-size where there is no sexual dimorphism for the trait, and corresponds
-to where the male-female regression line of Fig. 1A crosses the line
-representing no sexual dimorphism. This inflection point may represent
-the overall mean size across species in the lineage, or may be some
-other value. For instance, in some lineages (e.g., turtles: Ceballos et
-al. 2013), nearly all species are female biased, in which case the
-inflection point for \(D_{SD}\) vs.~size would be found at the far-right
-end of the plot. \hfill\break
+To ascertain the statistical performance of pgls \hfill\break
 
-Finally, from \(D_{SD}\) a \emph{sexual dimorphism distance index}
-(\(SDDI\)) may be constructed as:
+\subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
+Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}
 
-\[SDDI=-1*D_{SD}\] \[SDDI=+1*D_{SD}\]
+Despite the urging of Boettiger and colleagues to publish confidence
+intervals with all lambda parameter estimates, only 18\% of papers
+published in 2019 do so. \hfill\break
 
-Here, \(D_{SD}\) is multiplied by either a +1 or -1, based on a
-comparison of male and female trait values for the species. For
-instance, when single-valued traits on a linear scale are examined
-(e.g., leg length or wing extent), \(D_{SD}\) is multiplied by -1 for
-those species where females display larger values than males, and
-remains unchanged for species where males are larger than females. This
-logic is analogous to that of Lovich and Gibbons (1992) for their index
-of overall body size dimorphism. However, from a geometric perspective,
-this procedure is tantamount to determining which species in Fig. 1A are
-found below the 1:1 size line (i.e., the line representing no size
-dimorphism), and multiplying those species' \(D_{SD}\) by -1.
-\hfill\break
-
-Unfortunately, applying the same logic to multidimensional trait data is
-more complicated, as sex-specific differences are not always guaranteed
-to align with some \emph{a priori} direction in the multivariate
-dataspace. Thus, for multivariate datasets, one of two procedures may be
-used. First, for multivariate sets of traits that are measured on the
-same linear scale (e.g., length, width, and height measures), one could
-envision a hyperdimensional version of the plot displayed in Fig. 1A,
-but where the set of male traits are plotted against the set of female
-traits. In this construction, because the traits are quantified on the
-same measured scale, the 1:1 size vector representing no sexual
-dimorphism does exist, and emanates from the origin of the space
-extending in a positive direction as the trait values increase. To
-determine whether male trait values deviate more from this vector than
-do female trait values, a partial least squares analysis may be
-performed (Rohlf and Corti 2000). Partial least squares (PLS) is a
-multivariate association procedure which describes the covariation
-between sets of traits (in this case, between the set of male traits and
-the set of female traits). Scores along the first PLS axis represent the
-maximal covariation between male and female traits projected to a single
-dimension, and species whose male and female PLS scores are identical
-are those displaying no sexual dimorphism. Thus, deviations from that
-pattern represent instances of multivariate sexual dimorphism. By
-plotting these scores, one may determine which species display female
-scores (on \(PLS_1\)) that are larger than their corresponding male
-scores (when negative values on \(PLS_1\) correspond to smaller
-individuals; otherwise, the converse comparison is used). The \(D_{SD}\)
-for these species may then be multiplied by -1 to obtain the \(SDDI\)
-for that species. Note that when PLS is used on univariate data, this
-procedure obtains values identical to those from the method described
-previously where male and female sizes are compared directly. Thus, for
-multivariate traits measured on a linear scale, this is a direct
-generalization of the standard univariate procedure (see empirical
-example below). \hfill\break
-
-Alternatively, one might be interested in examining multidimensional
-traits other than linear biometric measurements, such as color or shape.
-For instance, one may characterize the shape of males and females using
-geometric morphometric methods (Bookstein 1991; Mitteroecker and Gunz
-2009; Adams et al. 2013), where the locations of anatomical landmarks
-and semilandmarks are used to obtain a set of shape variables. Geometric
-morphometric methods also result in a multidimensional characterization
-of phenotypic traits, from which the degree of sexual dimorphism for
-each species may be quantified using \(D_{SD}\). However, because the
-resulting shape variables do not have a one to one correspondence with
-linear size measures, there is no \emph{a priori} direction in the
-multivariate shape space that is equivalent to the 1:1 size vector found
-with sets of linear traits. In fact, for the case of isometry (i.e.,
-shape does not change as size changes), all observations are represented
-by a single location in the shape space. Thus, while a PLS of male
-shapes versus female shapes will describe the covariation between them,
-this direction is not guaranteed to align with size, because there is no
-direction in the shape space that may be used to unambiguously determine
-whether dimorphism is male- or female-biased, as for shape variables
-there is no \emph{a priori} directionality. Nonetheless, one may still
-quantitatively determine whether the degree of sexual shape dimorphism
-covaries with overall body size, by using the sexual dimorphism index as
-defined by the dimorphism distance (i.e., \(SDDI=D_{SD}\)) to represent
-the degree of multivariate shape dimorphism for each species.
-\hfill\break
-
-Once the sexual dimorphism distance index (\(SDDI\)) is calculated, it
-may be evaluated against the average size of each species. Here, data
-conforming to Rensch's rule will result in significant trends of sexual
-dimorphism relative to mean body size (Fig. 1D). As with size ratios,
-the line of no sexual dimorphism is a horizontal line, but this time it
-is centered on zero, as a distance of zero corresponds to no sexual
-dimorphism. Thus, a phylogenetic regression of these data whose slope is
-significantly different from zero (e.g., \(\beta_{1}\neq0.0\)) provides
-evidence that the evolution of sexual dimorphism is consistent with
-patterns expected under Rensch's rule. For cases where traits are
-measured on a linear scale, patterns expected under Rensch's rule
-exhibit a slope significantly greater than zero (\(\beta_{1}>0.0\)), as
-in the usual formulation. Likewise, when using geometric morphometric
-shape data (or any other type of mean-centered multidimensional data),
-hyperallometric patterns consistent with Rensch's rule would also
-display a slope significantly greater than zero (\(\beta_{1}>0.0\)),
-indicating that the degree of sexual shape dimorphism increases with
-increasing body size. Conversely however, hypoallometric patterns would
-display the opposite pattern (\(\beta_{1}<0.0\)), indicating that
-smaller species display greater sexual shape dimorphism. \hfill\break
-
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}
-
-To ascertain the statistical performance of the approach proposed here,
-we conducted a series of computer simulations. We evaluated the ability
-of SDDI to characterize patterns of sexual dimorphism in both univariate
-and multivariate datasets. For each simulation, we generated a
-pure-birth phylogeny containing 50 species, on which we simulated
-datasets under a Brownian motion model of evolution. For univariate data
-(\(p=1\)), a single trait was evolved along the phylogeny, and
-represented the female trait values (\(x\)). Next, male values (\(y\))
-were simulated in such a manner as to generate a known male:female
-relationship (plus random error), using the function:
-\(y=\alpha{x}+\mathcal{N}(\mu=0,\sigma=0.1)\). For type I error
-simulations, no sexual dimorphism was generated, meaning that the slope,
-\(\alpha\), characterized a 1:1 relationship between male and female
-trait values (i.e., \(\alpha=1.0\)). For power simulations, we simulated
-male values with a slope greater than 1.0 to correspond with patterns
-following Rensch's rule (i.e., \(\alpha=1.1\)). A total of 1,000
-datasets were generated under each simulation condition. We then
-evaluated all datasets using standard approaches (i.e., a phylogenetic
-regression of male values versus female values), as well as a
-phylogenetic regression of SDDI versus average body size. The proportion
-of significant datasets was treated as the type I error (\(\alpha=1.0\))
-or power (\(\alpha=1.1\)) respectively. \hfill\break
-
-For multivariate data (\(p=5\)), a slightly different procedure was
-utilized. First, average body size (\(s\)) was evolved along the
-phylogeny under Brownian motion. Next, female traits (\(\mathbf{X}\))
-were simulated on the phylogeny under Brownian motion, using a
-\(p\times{p}\) trait covariance matrix with correlations between traits
-of (\(r=0.7\)). For simulations where there was no pattern of sexual
-dimorphism across species (i.e., type I error simulations) male trait
-values were then generated by incorporating random error to the female
-trait values, in a manner analogous to that achieved in the univariate
-simulations:
-\(\mathbf{Y}=\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)). For
-simulations where patterns of sexual dimorphism covaried with body size
-(i.e., power simulations), male trait values were generated by
-incorporating random error to the female trait values, but with the
-addition that female trait values were pre-multiplied by a scalar value
-of body size:
-\(\mathbf{Y}=(0.1*s)*\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)).
-This ensured that male and female trait values covaried with one another
-as expected, and also covaried with size in the desired direction. As
-before, a total of 1,000 datasets were generated under each simulation
-condition. We then evaluated all datasets using a phylogenetic
-regression of SDDI versus average body size. The proportion of
-significant datasets was treated as the type I error (\(\alpha=1.0\)) or
-power (\(\alpha=1.1\)) respectively. \hfill\break
-
-\emph{Simulation Results}: For the univariate dataset, we found that
-both a phylogenetic regression of male size on female size, as well as a
-phylogenetic regression of \(SDDI\) on body size, resulted in
-appropriate type I error rates near the nominal \(\alpha\approx0.05\)
-(Fig. 2). Further, when data were simulated under conditions consistent
-with Rensch's rule, both procedures were capable of detecting the signal
-of the pattern, and did so with very high power. Thus, using the
-\(SDDI\) approach proposed here was equivalent to what is normally
-utilized for univariate datasets. Likewise, for multivariate data, the
-\(SDDI\) approach displayed appropriate type I error
-(\(\alpha\approx0.05\)) and high statistical power under the conditions
-simulated (Fig. 2). Additionally, multiplying \(D_{SD}\) by \(\pm1\) had
-little effect on the outcome for multivariate data, as in both cases
-type I error and power results were virtually identical. Thus, methods
-for evaluating sexual dimorphism using geometric morphometric data are
-anticipated to remain robust. Overall results from these simulations
-imply that using the sexual dimorphism distance index (\(SDDI\))
-proposed here provides an appropriate means of characterizing patterns
-of sexual dimorphism in both univariate and multivariate traits, and
-that the procedure is capable of detecting patterns consistent with
-Rensch's rule when they are present in phenotypic datasets. \hfill\break
-
-\subsection{\texorpdfstring{\emph{Empirical
-Example}}{Empirical Example}}\label{empirical-example}
-
-To illustrate the utility of the approach presented here, we conducted
-an analysis of both size and shape dimorphism in a group of
-Mediterranean lizards with extensive body size variation. Our dataset
-consisted of three data types: male and female values of body size
-(snout-vent length: SVL), male and female values for a series of head
-measurements (head length, head width, head height, pileus length, and
-mouth opening: Fig. 3A), and male and female values of head shape (Fig.
-3B). A total of 374 specimens from 21 lineages of green lizards, a
-monophyletic clade that includes two genera (\emph{Lacerta} and
-\emph{Timon}), were obtained from natural history museum collections.
-For each lineage, we obtained samples distributed across the entire
-distribution range, resulting in an initial database of 1062 specimens.
-In order to take sampling bias into account, we selected the 10 largest
-individuals per sex, when sufficient specimens were available, for use
-in subsequent size and shape analyses. A detailed account of the
-specimens used for this study is available in the Supplemental Material.
-\hfill\break
-
-From these specimens, we calculated mean body size for each sex and
-lineage. Likewise, we calculated the mean trait values for each sex and
-lineage for the set of linear measures. These values were then
-standardized by SVL; resulting in a set of head measurements
-proportional to body size for each sex \(\times\) lineage combination.
-Additionally, we quantified lateral head shape variation using geometric
-morphometric methods (Adams et al. 2013). This was accomplished by first
-digitizing 17 landmarks and semilandmarks from images of the right side
-of the head (Fig. 3B). The thin-plate spline was used to estimate the
-position of missing landmarks (Gunz et al. 2009), and the position of
-the jaw relative to the skull was standardized using the fixed angle
-method (Adams 1999). Next, a generalized Procrustes analysis (Rohlf and
-Slice 1990) was performed to eliminate non-shape variation from the
-landmark data and align the specimens to a common coordinate system. The
-position of semilandmarks (i.e.~11-16, Fig. 3B) was optimized by
-minimizing the bending energy (Bookstein 1997). From the aligned
-Procrustes coordinates the mean head shape for males and for females was
-then calculated for each lineage. Finally, the phylogenetic
-relationships among lineages were estimated using published sequences on
-five mitochondrial and four nuclear gene fragments, totaling 8383 bp.
-These were then used following standard phylogenetic inference
-procedures to generate a dated consensus phylogeny for the lineages of
-interest (for details see Supplemental Material). \hfill\break
-
-Using these data, we conducted a series of analyses to investigate
-whether patterns of sexual size dimorphism (SSD) and sexual shape
-dimorphism (SShD) in this group corresponded to expectations under
-Rensch's rule. First, we examined patterns of SSD using a phylogenetic
-regression of male size against female size. Here, a slope significantly
-greater than 1.0 (\(\beta>1.0\)) was treated as evidence consistent with
-Rensch's rule. Additionally, we evaluated SSD patterns using the
-approach developed here, via a phylogenetic regression of the sexual
-size dimorphism index against mean body size for each species. Second,
-we evaluated multivariate patterns of sexual dimorphism in head
-proportions in relation to body size, using the sexual dimorphism index
-obtained from the Euclidean distance between male and female means from
-each species (\(D_{SD}\)). Partial least squares was used to determine
-which species displayed female scores (on \(PLS_1\)) larger than their
-corresponding male scores, and the sexual dimorphism index was then
-obtained by multiplying the sexual dimorphism distance by +1 or -1,
-respectively (i.e., \(SDDI=\pm1*D_{SD}\)). We then performed a
-phylogenetic regression of \(SDDI\) for head proportions versus mean
-body size to determine whether patterns of dimorphism corresponded to
-what is expected under Rensch's rule. Third, we obtained the degree of
-sexual dimorphism in head shape for each species using the Procrustes
-distance between male and female means, and from this calculated the
-sexual shape dimorphism index (\(SShDI\)) in their original units as
-described above (i.e., \(SShDI=D_{SD}\)). We then performed a
-phylogenetic regression of sexual shape dimorphism (\(SShDI\)) against
-the mean body size for each species to determine whether patterns of
-head shape dimorphism were consistent with predictions from Rensch's
-rule. Additionally, patterns of head shape dimorphism were further
-investigated using principal components analysis of the set of mean male
-and female head shapes across species. Finally, we used phylogenetic
-correlation (phylogenetic partial least squares: sensu Adams and Felice
-2014) to determine whether the degree of head shape dimorphism was
-associated with levels of size dimorphism. All analyses were performed
-in R 3.6.2 (R Core Team 2019) using the packages \texttt{geomorph}
-(Adams and Otárola-Castillo 2013; Adams et al. 2019), \texttt{RRPP}
-(Collyer and Adams 2018), and routines written by the authors.
-\hfill\break
-
-\emph{Empirical Results}: Using phylogenetic regression, we found a
-significant relationship between male size and female size, with a slope
-significantly greater than one (\(\beta = 1.39\), \(T=2.91\),
-\(P = 0.009\): Fig. 3C). This indicated that the degree of sexual size
-dimorphism differed systematically with overall body size; a pattern
-consistent with that expected under Rensch's rule. In this case, both
-hypoallometry and hyperallometry (\emph{sensu} Abouheif and Fairbairn
-1997) were displayed (Fig. 3C). Specifically, in male-biased species,
-the degree of sexual dimorphism was greater with increasing body size,
-and in female-biased species the degree of sexual dimorphism increased
-in smaller species (though the pattern was more pronounced in
-male-biased species). Similarly, when using the sexual size dimorphism
-index (SSDI), we found a significant and positive evolutionary
-relationship between SSDI and mean body size, confirming the Rensch's
-rule pattern identified using the commonly-used procedure
-(\(\beta = 0.43\), \(F=22.95\), \(Z=1.89\), \(P = 0.004\): Fig. 3D).
-\hfill\break
-
-When multivariate patterns of sexual dimorphism were evaluated, we found
-that the degree of dimorphism in head proportions increased as body size
-increased: a pattern consistent with Rensch's rule (\(F=7.20\),
-\(Z=1.41\), \(P = 0.022\): Fig. 3E). Likewise, when patterns of head
-shape dimorphism were examined, we found that the degree of sexual shape
-dimorphism (SShDI) increased in larger species, and a phylogenetic
-regression of sexual shape dimorphism versus mean body size was found to
-be significant (\(F=6.05\), \(Z=1.35\), \(P = 0.03\): Fig. 3F). Thus,
-for both multivariate datasets, the degree of sexual dimorphism was
-enhanced in larger species (hyperallometry; sensu Abouheif and Fairbairn
-1997): a pattern consistent with what was expected under Rensch's rule.
-To our knowledge, this is the first empirical demonstration of a pattern
-following Rensch's rule in multivariate data. \hfill\break
-
-Visualizing patterns of head shape dimorphism using principal components
-analysis (Fig. 4), it was evident that males and females occupied
-distinct regions of morphospace. Further, the head shapes of males and
-females of the smallest lineage, \emph{Lacerta agilis bosnica}, were
-distinct from those of the largest species, \emph{Timon lepidus}. When
-connected, the sexual shape dimorphism vectors for these two species did
-not coincide with one another in the ordination, implying a difference
-in sexual shape dimorphism across species. Specifically, the vector for
-\emph{T. lepidus} was longer than that for \emph{L. agilis bosnica},
-which reflected the greater degree of sexual shape dimorphism exhibited
-by this species. Thin-plate spline deformation grids facilitated a
-graphical description of these shape differences, where sexual shape
-dimorphism in \emph{L. agilis bosnica} was best described as a contrast
-between longer-headed males, with an amplified tympanic area; and
-short-headed females, with a relatively more reduced tympanic region
-(Fig. 4, left). By contrast, sexual shape dimorphism in \emph{T.
-lepidus} was both more prominent in the tympanic area, similarly to what
-was observed in \emph{L. agilis bosnica}; and also included a
-modification of the relative lower jaw area. Indeed, males of \emph{T.
-lepidus} were characterized by both a longer posterior head region and
-by deeper lower jaw configurations as compared to females of the same
-species. \hfill\break
-
-Finally, across taxa, the degree of body size and head shape dimorphism
-were significantly correlated with one another (\(r=0.73\),
-\(p=0.0001\)), but not when evaluated in a phylogenetic context
-(\(r_{PPLS} = 0.361\), \(Z=1.28\), \(P = 0.12\)). Part of this
-difference may be explained by the fact that the largest species
-(i.e.~members of the genus \emph{Timon}) - which displayed the greatest
-degree of both size and shape dimorphism - form a clade within the
-broader phylogeny for the group (Fig. 5). Thus, despite the strong
-association of size and shape dimorphism, there were fewer changes of
-these patterns across branches of the phylogeny. Generally speaking, the
-degree of size dimorphism and the degree of shape dimorphism increased
-in larger taxa, and heat maps of the evolution of size and shape
-dimorphism across the phylogeny indicated a strong association between
-the two; particularly in large taxa (Fig. 5). Thus, in this system, it
-appeared that evolutionary changes in size between the sexes were
-accompanied by concomitant changes in head shape.
+All analyses were performed in R 3.6.2 {[}R-Base{]} using the packages
+\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2019),
+\texttt{RRPP} (Collyer and Adams 2018). \hfill\break
 
@@
 
-A longstanding question in evolutionary biology is understanding why
-sexual dimorphism, the degree of differentiation between the sexes,
-varies across species. A frequently used schema for exploring
-macroevolutionary patterns of sexual dimorphism is Rensch's rule, which
-predicts that sexual dimorphism should increase with body size in
-species where males are the larger sex, and decrease with increasing
-body size in species where females are the larger sex. Remarkably,
-virtually all recent studies have focused on macroevolutionary variation
-in sexual \emph{size} dimorphism, despite the relevance of sexual
-dimorphism in other traits, as encompassed by Rensch's original
-treatment (Rensch 1950). This gap has been driven, in part, by the lack
-of adequate tools for evaluating Rensch's rule in multivariate traits,
-such as color, ornament complexity, or shape. In this article, we fill
-this procedural gap by providing an analytical equivalency that allows
-the generalization of statistical tests of Rensch's rule for both
-univariate and multivariate traits. The approach describes the degree of
-sexual dimorphism as a distance between male and female trait values in
-phenotype space, thereby facilitating its use on either single-valued
-traits (e.g., body size), or complex, multivariate traits
-(e.g.~organismal shape). Computer simulations confirm that the approach
-exhibits appropriate type I error rates and statistical power, thereby
-providing an adequate procedure for testing whether empirical data
-comply with the predictions of Rensch's rule. The implementation of our
-approach on an empirical study of Mediterranean lizards revealed that
-patterns of sexual dimorphism comply with Rensch's rule not only for
-body size, but also for relative head proportions and head shape,
-offering new insights and triggering novel hypotheses about the possible
-proximate and evolutionary causes that may underlie such patterns.
-\hfill\break
+Using the estimated lambda values from pgls are not useful. The
+questions of whether or not signal exists is appropriate, but inferring
+more from lambda \emph{magnitude} is inappropriate. \hfill\break
 
-Our empirical example in Mediterranean green lizards exemplifies a case
-study investigating variation in sexual dimorphism -- not only in size
--- but also in multivariate phenotypes, like shape, and how this method
-can provide useful insights into the proximate and evolutionary causes
-underlying such patterns. First, we found that the degree of sexual size
-dimorphism increased as taxa became larger (Fig. 3C, D). This pattern is
-compliant with Rensch's rule and corresponds with hyperallometry
-(Abouheif and Fairbairn 1997), which is to be expected when males are
-the larger sex (Fairbairn 1997). In this group males are larger than
-females in most lineages (Fig. 3C), as is commonly observed in lacertids
-(Braña 1996). Additionally, as in many lizard species, sexual selection
-in this group is primarily mediated through territoriality, where
-male-male competition selects for a larger male body size (Stamps 1983).
-Thus, in this system there is an association between male-biased sexual
-size dimorphism and male-male competition. As such, the patterns of size
-dimorphism observed in this study are consistent with those expected
-under a scenario where sexual selection drives the evolution of larger
-male body size across taxa via male-male competition. \hfill\break
-
-Likewise, our multivariate analyses revealed that head morphology
-differed between the sexes, and that the degree of sexual dimorphism in
-multivariate head shape also corresponded to that expected under
-Rensch's rule (Fig. 3E, 3F). That is, while both small (e.g. \emph{L. a.
-bosnica}, Fig. 4, left) and large (e.g. \emph{T. lepidus}, Fig. 4,
-right) species are sexually dimorphic in head shape, the degree of
-sexual shape dimorphism increased with average body size (Fig. 3F); a
-pattern consistent with a scenario of hyperallometry. Examination of the
-resulting shape differences revealed that sexual dimorphism in head
-shape was associated with an amplification of the posterior head region,
-and also in the relative size of the lower jaw (in the case of \emph{T.
-lepidus}). Biomechanically, these are precisely the anatomical regions
-where the jaw adductor muscles reside (Haas 1973), and as such, an
-amplification of these areas in males would enhance biting performance
-(Herrel et al. 1996; Kaliontzopoulou et al. 2012); a functional trait
-that has important implications for male competitive capacity (Huyghe et
-al. 2005; Lailvaux and Irschick 2007). In addition, one might predict
-that the need for stronger jaw musculature in males becomes
-disproportionately more intense with increasing body size, as male
-combats are expected to be more fierce among larger-bodied opponents.
-Thus, our findings that head shape dimorphism increases with overall
-body size is also consistent with the hypothesis that sexual selection
-on structures related to territoriality and male competitive
-interactions are responsible for the macroevolutionary patterns of
-sexual shape dimorphism observed in this group. To our knowledge, this
-is the first example of Rensch's rule in a multivariate phenotype.
-\hfill\break
-
-With respect to the analytical approach proposed here, our
-characterization of sexual dimorphism via a \emph{shape distance} and
-the \(SDDI\) statistic is completely general, and may be used to
-evaluate evolutionary trends in any continuous, quantitative trait that
-differs between the sexes. Our empirical example highlighted the
-efficacy of the approach for evaluating trends in univariate traits such
-as body size, and also several types of complex data where organismal
-phenotypes are quantified by multiple dimensions. As such, the approach
-developed here expands our current analytical toolkit to provide a means
-of evaluating evolutionary trends in sexual dimorphism for a broader
-class of phenotypic traits than has previously been considered. Indeed,
-this advance more closely embodies Rensch's (1950) original
-conceptualization for understanding patterns of sexual dimorphism, as
-his broader vision was to explain why phenotypic differences \emph{writ
-large} exist among the sexes. Thus, use of the \(SDDI\) will facilitate
-a greater degree of evolutionary exploration into patterns of sexual
-dimorphism than has been possible in recent decades. Further, with the
-inclusion of multivariate traits, the formulation proposed here opens
-the door for investigations into the evolution of sexual shape
-dimorphism, and whether such patterns correspond to those expected under
-Rensch's rule. Thus, the mathematical test proposed here is general, and
-may be used for any quantitative phenotypic trait under scrutiny.
-\hfill\break
-
-Finally, because \(SDDI\) is derived from a phenotypic distance, one
-could, in theory, utilize the approach to characterize patterns of
-sexual dimorphism in non-continuous traits, and determine quantitatively
-whether the degree of sexual dimorphism is enhanced with increasing body
-size. Across the tree of life, major phenotypic adaptations are often
-described as evolutionary innovations or novelties (sensu Wagner and
-Altenberg 1996; Wagner and Muller 2002; Peterson and Müller 2016), where
-these changes are characterized as discrete shifts from one phenotypic
-state to another (O'Keefe et al. 2011; e.g., Holliday and Gardner 2012).
-In like manner, discrete phenotypic differences are known to exist
-between the sexes; for instance where males display horns or other
-ornamentation (Geist and Bayer 1988; Emlen 2008) that females lack. Over
-evolutionary time, a series of such discrete phenotypic differences may
-accumulate in males relative to females, resulting in an elaboration of
-phenotypic differences between the sexes in suites of such traits. With
-an appropriate distance measure for discrete traits (e.g., Hamming
-distance), the approach developed here is capable of characterizing such
-patterns, and determining whether sets of discrete phenotypic
-differences are enhanced with increasing body size among species. Thus,
-our approach is the first to our knowledge to facilitate a formal
-evaluation of both discrete and continuous phenotypic differences
-between the sexes, providing a comprehensive means of determining
-whether sets of phenotypic traits differ in a manner as predicted by
-Rensch's rule.
+More discussion paragraphs
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Abouheif1997}{}
-Abouheif, E., and D. J. Fairbairn. 1997. A comparative analysis of
-allometry for sexual size dimorphism: Assessing rensch's rule. The
-American Naturalist 149:540--562.
-
-\hypertarget{ref-Adams1999}{}
-Adams, D. C. 1999. Methods for shape analysis of landmark data from
-articulated structures. Evolutionary Ecology Research 1:959--970.
-
-\hypertarget{ref-AdamsCollyer2018a}{}
-Adams, D. C., and M. L. Collyer. 2018. Multivariate phylogenetic
-comparative methods: Evaluations, comparisons, and recommendations.
-Systematic Biology 67:14--31.
-
-\hypertarget{ref-AdamsCollyer2019}{}
-Adams, D. C., and M. L. Collyer. 2019. Phylogenetic comparative methods
-and the evolution of multivariate phenotypes. Annual Review of Ecology,
-Evolution, and Systematics 50:405--425.
-
-\hypertarget{ref-AdamsFelice2014}{}
-Adams, D. C., and R. Felice. 2014. Assessing trait covariation and
-morphological integration on phylogenies using evolutionary covariance
-matrices. PLoS ONE 9:e94335.
-
 \hypertarget{ref-AdamsOtarola2013}{}
@@
 
-\hypertarget{ref-Adams_et_al2013}{}
-Adams, D. C., F. J. Rohlf, and D. E. Slice. 2013. A field comes of age:
-Geometric morphometrics in the 21st century. Hystrix 24:7--14.
-
 \hypertarget{ref-AdamsGeomorph}{}
@@
 for geometric morphometric analyses. r package version 3.1.1.
-
-\hypertarget{ref-Altman1983}{}
-Altman, D. G., and J. M. Bland. 1983. Measurement in medicine: The
-analysis of method comparison studies. The Statistician 32:307--317.
-
-\hypertarget{ref-Andersson1994}{}
-Andersson, M. 1994. Sexual selection. Princeton University Press.
-
-\hypertarget{ref-Astua2010}{}
-Astúa, D. 2010. Cranial sexual dimorphism in new world marsupials and a
-test of renschs rule in didelphidae. Journal of Mammalogy 91:1011--1024.
-
-\hypertarget{ref-Berns2013}{}
-Berns, C. M., and D. C. Adams. 2013. Becoming different but staying
-alike: Patterns of sexual size and shape dimorphism in bills of
-hummingbirds. Evolutionary Biology 40:246--260.
-
-\hypertarget{ref-Blanckenhorn2006}{}
-Blanckenhorn, W. U., R. C. Stillwell, K. A. Young, C. W. Fox, and K. G.
-Ashton. 2006. When rensch meets bergmann: Does sexual size dimorphism
-change systematically with latitude? Evolution 60:2004--2011.
-
-\hypertarget{ref-Blows2007}{}
-Blows, M. W. 2007. A tale of two matrices: Multivariate approaches in
-evolutionary biology. Journal of Evolutionary Biology 20:1--8.
-
-\hypertarget{ref-Bolnick2003}{}
-Bolnick, D. I., and M. Doebeli. 2003. Sexual dimorphism and adaptive
-speciation: Two sides of the same ecological coin. Evolution
-57:2433--2449.
-
-\hypertarget{ref-Bonduriansky2007}{}
-Bonduriansky, R. 2007. Sexual selection and allometry: A critical
-appraisal of the evidence and ideas. Evolution 61:838--849.
-
-\hypertarget{ref-Bookstein1997}{}
-Bookstein, F. L. 1997. Landmark methods for forms without landmarks:
-Morphometrics of group differences in outline shape. Medical Image
-Analysis 1:225--243.
-
-\hypertarget{ref-Bookstein1991}{}
-Bookstein, F. L. 1991. Morphometric tools for landmark data: Geometry
-and biology. Cambridge University Press, Cambridge.
-
-\hypertarget{ref-Brana1996}{}
-Braña, F. 1996. Sexual dimorphism in lacertid lizards: Male head
-increase vs female abdomen increase? Oikos 75:511--523.
-
-\hypertarget{ref-Burbrink2019}{}
-Burbrink, F. T., and I. Futterman. 2019. Female-biased gape and
-body-size dimorphism in the new world watersnakes (tribe: Thamnophiini)
-oppose predictions from renschs rule. Ecology and Evolution
-9:9624--9633.
-
-\hypertarget{ref-Butler_et_al2000}{}
-Butler, M. A., T. W. Schoener, and J. B. Losos. 2000. The relationship
-between sexual size dimorphism and habitat use in greater antillean
-\emph{Anolis} lizards. Evolution 54:259--272.
-
-\hypertarget{ref-Ceballos2013}{}
-Ceballos, C. P., D. C. Adams, J. B. Iverson, and N. Valenzuela. 2013.
-Phylogenetic patterns of sexual size dimorphism in turtles and their
-implications for Rensch's rule. Evolutionary Biology 40:194--208.
 
@@
 
-\hypertarget{ref-Collyer_et_al2015a}{}
-Collyer, M. L., D. J. Sekora, and D. C. Adams. 2015. A method for
-analysis of phenotypic change for phenotypes described by
-high-dimensional data. Heredity 115:357--365.
-
-\hypertarget{ref-Cox2003}{}
-Cox, R. M., S. L. Skelly, and H. B. John-Alder. 2003. A comparative test
-of adaptive hypotheses for sexual size dimorphism in lizards. Evolution
-57:1653--1669.
-
-\hypertarget{ref-Dale2007}{}
-Dale, J., P. O. Dunn, J. Figuerola, T. Lislevand, T. Székely, and L. A.
-Whittingham. 2007. Sexual selection explains renschs rule of allometry
-for sexual size dimorphism. Proceedings of the Royal Society B:
-Biological Sciences 274:2971--2979.
-
-\hypertarget{ref-Darwin1871}{}
-Darwin, C. R. 1871. The descent of man, and selection in relation to
-sex. J. Murray, London.
-
-\hypertarget{ref-DayanSimberloff2005}{}
-Dayan, T., and D. Simberloff. 2005. Ecological and community-wide
-character displacement: The next generation. Ecology Letters 8:875--894.
-
-\hypertarget{ref-Emlen2008}{}
-Emlen, D. J. 2008. The evolution of animal weapons. Annual Review of
-Ecology, Evolution, and Systematics 39:387--413.
-
-\hypertarget{ref-Endler1983}{}
-Endler, J. A. 1983. Natural and sexual selection on color patterns in
-poeciliid fishes. Environmental Biology of Fishes 9:173--190.
-
-\hypertarget{ref-Fairbairn1997}{}
-Fairbairn, D. J. 1997. Allometry for sexual size dimorphism:Pattern and
-process in the coevolution of body size in males and females. Annual
-Review of Ecology and Systematics 28:659--687.
-
-\hypertarget{ref-Fairbairn2013}{}
-Fairbairn, D. J. 2013. Odd couples - extraordinary differences between
-the sexes in the animal kingdom. Princeton University Press.
-
-\hypertarget{ref-Fairbairn1994}{}
-Fairbairn, D. J., and R. F. Preziosi. 1994. Sexual selection and the
-evolution of allometry for sexual size dimorphism in the water strider,
-Aquarius remigis. American Naturalist 144:101--118.
-
-\hypertarget{ref-Fairbairn2007}{}
-Fairbairn, D. J., W. U. Blanckenhorn, and T. Székely. 2007. Sex, size
-and gender roles. Oxford University Press.
-
-\hypertarget{ref-Garcia-Navas2015}{}
-Garcia-Navas, V., T. Bonnet, R. Bonal, and E. Postma. 2015. The role of
-fecundity and sexual selection in the evolution of size and sexual size
-dimorphism in new world and old world voles (rodentia: Arvicolinae).
-Oikos 125:1250--1260.
-
-\hypertarget{ref-Geist1988}{}
-Geist, V., and M. Bayer. 1988. Sexual dimorphism in the cervidae and its
-relation to habitat. Journal of Zoology 214:45--53.
-
-\hypertarget{ref-Gunz2009}{}
-Gunz, P., P. Mitteroecker, S. Neubauer, G. W. Weber, and F. L.
-Bookstein. 2009. Principles for the virtual reconstruction of hominin
-crania. Journal of Human Evolution 57:48--62.
-
-\hypertarget{ref-Haas1973}{}
-Haas, G. 1973. Muscles of the jaws and associated structures in the
-rhynchocephalia and squamata. Pp. 285--490 \emph{in} C. Biology of the
-reptilia vol. 4. Gans and T. Parsons, eds. Academica Press.
-
-\hypertarget{ref-Herrel1996}{}
-Herrel, A., R. V. Damme, and F. D. Vree. 1996. Sexual dimorphism of head
-size in podarcis hispanica atrata: Testing the dietary divergence
-hypothesis by bite force analysis. Netherlands Journal of Zoology
-46:253--262.
-
-\hypertarget{ref-Hirst2014}{}
-Hirst, A. G., and T. Kiørboe. 2014. Macroevolutionary patterns of sexual
-size dimorphism in copepods. Proceedings of the Royal Society B:
-Biological Sciences 281:20140739.
-
-\hypertarget{ref-Holliday2012}{}
-Holliday, C. M., and N. M. Gardner. 2012. A new eusuchian crocodyliform
-with novel cranial integument and its significance for the origin and
-evolution of crocodylia. PLoS ONE 7:e30471.
-
-\hypertarget{ref-Horne2020}{}
-Horne, C. R., A. G. Hirst, and D. Atkinson. 2020. Selection for
-increased male size predicts variation in sexual size dimorphism among
-fish species. Proceedings of the Royal Society B: Biological Sciences
-287:20192640.
-
-\hypertarget{ref-HUYGHE2005}{}
-Huyghe, K., B. V. Hooydonck, H. Scheers, M. Molina-Borij, and R. V.
-Damme. 2005. Morphology, performance and fighting capacity in male
-lizards, \emph{Gallotia galloti}. Functional Ecology 19:800--807.
-
-\hypertarget{ref-Johnson2017}{}
-Johnson, P. J., M. J. Noonan, A. C. Kitchener, L. A. Harrington, C.
-Newman, and D. W. Macdonald. 2017. Rensching cats and dogs: Feeding
-ecology and fecundity trends explain variation in the allometry of
-sexual size dimorphism. Royal Society Open Science 4:170453.
-
-\hypertarget{ref-Kaliontzopoulou2012}{}
-Kaliontzopoulou, A., D. C. Adams, A. van der Meijden, A. Perera, and M.
-A. Carretero. 2012. Relationships between head morphology, bite
-performance and ecology in two species of podarcis wall lizards.
-Evolutionary Ecology 26:825--845.
-
-\hypertarget{ref-Kaliontzopoulou2015}{}
-Kaliontzopoulou, A., M. A. Carretero, and D. C. Adams. 2015.
-Ecomorphological variation in male and female wall lizards and the
-macroevolution of sexual dimorphism in relation to habitat use. Journal
-of Evolutionary Biology 28:80--94.
-
-\hypertarget{ref-Kaliontzopoulou_et_al2010}{}
-Kaliontzopoulou, A., M. A. Carretero, and G. A. Llorente. 2010.
-Intraspecific ecomorphological variation: Linear and geometric
-morphometrics reveal habitat-related patterns within \emph{podarcis
-bocagei} wall lizards. Journal of Evolutionary Biology 23:1234--1244.
-
-\hypertarget{ref-Kelly2013}{}
-Kelly, C. D., K. E. Folinsbee, D. C. Adams, and M. D. Jennions. 2013.
-Intraspecific sexual size and shape dimorphism in an australian
-freshwater fish differs with respect to a biogeographic barrier and
-latitude. Evolutionary Biology 40:408--419.
-
-\hypertarget{ref-Lailvaux2007}{}
-Lailvaux, S. P., and D. J. Irschick. 2007. The evolution of
-performance-based male fighting ability in caribbean anolis lizards. The
-American Naturalist 170:573--586.
-
-\hypertarget{ref-Lande1979}{}
-Lande, R. 1979. Quantitative genetic analysis of multivariate evolution,
-applied to brain: Body size allometry. Evolution 33:402--416.
-
-\hypertarget{ref-LandeArnold1983}{}
-Lande, R., and S. J. Arnold. 1983. The measurement of selection on
-correlated characters. Evolution 37:1210--1226.
-
-\hypertarget{ref-Littleford-Colquhoun2019}{}
-Littleford-Colquhoun, B. L., C. Clemente, G. Thompson, R. H. Cristescu,
-N. Peterson, K. Strickland, D. Stuart-Fox, and C. H. Frere. 2019. How
-sexual and natural selection shape sexual size dimorphism: Evidence from
-multiple evolutionary scales. Functional Ecology 33:1446--1458.
-
-\hypertarget{ref-Lovich1992}{}
-Lovich, J. E., and J. W. Gibbons. 1992. A review of techniques for
-quantifying sexual size dimorphism. Growth, Development, and Aging
-56:269--281.
-
-\hypertarget{ref-Meiri2014}{}
-Meiri, S., A. E. Kadison, M. Novosolov, P. Pafilis, J. Foufopoulos, Y.
-Itescu, P. Raia, and D. Pincheira-Donoso. 2014. The number of competitor
-species is unlinked to sexual dimorphism. Journal of Animal Ecology
-83:1302--1312.
-
-\hypertarget{ref-MitteroeckerGunz2009}{}
-Mitteroecker, P., and P. Gunz. 2009. Advances in geometric
-morphometrics. Evolutionary Biology 36:235--247.
-
-\hypertarget{ref-OKeefe2011}{}
-O'Keefe, F. R., H. P. Street, B. C. Wilhelm, C. D. Richards, and H. Zhu.
-2011. A new skeleton of the cryptoclidid plesiosaurTatenectes
-laramiensisreveals a novel body shape among plesiosaurs. Journal of
-Vertebrate Paleontology 31:330--339.
-
-\hypertarget{ref-Oestman2011}{}
-Östman, O., and D. Stuart-Fox. 2011. Sexual selection is positively
-associated with ecological generalism among agamid lizards. Journal of
-Evolutionary Biology 24:733--740.
-
-\hypertarget{ref-Penalver-Alcazar2019}{}
-Peñalver-Alcázar, M., P. Galán, and P. Aragón. 2019. Assessing rensch's
-rule in a newt: Roles of primary productivity and conspecific density in
-interpopulation variation of sexual size dimorphism. Journal of
-Biogeography 46:2558--2569.
-
-\hypertarget{ref-Peterson2016}{}
-Peterson, T., and G. B. Müller. 2016. Phenotypic novelty in EvoDevo: The
-distinction between continuous and discontinuous variation and its
-importance in evolutionary theory. Evolutionary Biology 43:314--335.
-
-\hypertarget{ref-R-Base}{}
-R Core Team. 2019. R: A language and environment for statistical
-computing. R Foundation for Statistical Computing, Vienna, Austria.
-
-\hypertarget{ref-Regis2017}{}
-Regis, K. W., and J. M. Meik. 2017. Allometry of sexual size dimorphism
-in turtles: A comparison of mass and length data. PeerJ 5:e2914.
-
-\hypertarget{ref-Reiss1986}{}
-Reiss, M. J. 1986. Sexual dimorphism in body size: Are larger species
-more dimorphic? Journal of Theoretical Biology 121:163--172.
-
-\hypertarget{ref-Rensch1950}{}
-Rensch, B. 1950. Die abhangigkeit der relativen sexualdifferenz von der
-korpergrosse. Bonner Zoologische Beitraege 1:58--69.
-
-\hypertarget{ref-Rensch1960}{}
-Rensch, B. 1960. Evolution above the species level. Columbia University
-Press.
-
-\hypertarget{ref-RohlfCorti2000}{}
-Rohlf, F. J., and M. Corti. 2000. The use of partial least-squares to
-study covariation in shape. Systematic Biology 49:740--753.
-
-\hypertarget{ref-RohlfSlice1990}{}
-Rohlf, F. J., and D. E. Slice. 1990. Extensions of the procrustes method
-for the optimal superimposition of landmarks. Systematic Zoology
-39:40--59.
-
-\hypertarget{ref-Serrano-Meneses2006}{}
-Serrano-Meneses, M.-A., and T. Szekely. 2006. Sexual size dimorphism in
-seabirds: Sexual selection, fecundity selection and differential
-niche-utilisation. Oikos 113:385--394.
-
-\hypertarget{ref-Stamps1983}{}
-Stamps, J. 1983. Sexual selection, sexual dimorphism and territoriality.
-Pp. 169--204 \emph{in} Lizard ecology: studies of a model organism. Huey
-R.B., Pianka E.R., and T. Schoener, eds. Harvard University Press.
-
-\hypertarget{ref-Stephens2009}{}
-Stephens, P. R., and J. J. Wiens. 2009. Evolution of sexual size
-dimorphisms in emydid turtles: Ecological dimorphism, rensch's rule, and
-sympatric divergence. Evolution 63:910--925.
-
-\hypertarget{ref-Stuart-Fox2008}{}
-Stuart-Fox, D. 2008. A test of rensch's rule in dwarf chameleons
-(bradypodion spp.), a group with female-biased sexual size dimorphism.
-Evolutionary Ecology 23:425--433.
-
-\hypertarget{ref-Stuart-Fox2007}{}
-Stuart-Fox, D., and A. Moussalli. 2007. Sex-specific ecomorphological
-variation and the evolution of sexual dimorphism in dwarf chameleons
-(bradypodion spp.). Journal of Evolutionary Biology 20:1073--1081.
-
-\hypertarget{ref-Tarr2018}{}
-Tarr, S., S. Meiri, J. J. Hicks, and A. C. Algar. 2018. A biogeographic
-reversal in sexual size dimorphism along a continental temperature
-gradient. Ecography 42:706--716.
-
-\hypertarget{ref-Wagner1996a}{}
-Wagner, G. P., and L. Altenberg. 1996. Perspective: Complex adaptations
-and the evolution of evolvability. Evolution 50:967.
-
-\hypertarget{ref-Wagner2002}{}
-Wagner, G. P., and G. B. Muller. 2002. Evolutionary innovations overcome
-ancestral constraints: A re-examination of character evolution in male
-sepsid flies (diptera: Sepsidae). Evolution and Development 4:1--6.
-
-\hypertarget{ref-Watson2010}{}
-Watson, N. L., and L. W. Simmons. 2010. Mate choice in the dung beetle
-onthophagus sagittarius: Are female horns ornaments? Behavioral Ecology
-21:424--430.
-
 \newpage
@@
 
-\textbf{Figure 1}. (A) Sexual size dimorphism for males and females
-across 100 hypothetical species. The pattern displayed exhibits a slope
-\textgreater{} 1.0, and thus corresponds to Rensch's rule. (B) The same
-data represented as the ratio of \(M/F\) sizes plotted against average
-species size. (C) Sexual dimorphism represented as the distance between
-males and females: \(SDD = D_{M:F}\). (D) Sexual dimorphism represented
-as an index of the SD Distance: \(SDDI = \pm1*D_{M:F}\). In all panels,
-the dashed line corresponds to values representing no sexual dimorphism
-(the 1:1 line). \hfill\break
+\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known
+lambda inputs on various tree sizes. As trees increase in size, the
+estimates more closely resemble the input lambdas, however considerable
+and concerning variation is apparent in trees smaller than those with
+256 tips. \hfill\break
 
-\textbf{Figure 2}. Results of phylogenetic simulations testing the type
-I error and statistical power for detecting patterns consistent with
-Rensch's rule. Vertical bars represent the proportion of datasets found
-to be significant. For type I error analyses, the dashed line marks the
-standard threshold of \(\alpha= 0.05\). For univariate data, both
-phylogenetic regression of male versus female size (standard approach),
-and \(SDDI\) versus overall body size (new approach) were performed. For
-multivariate data, only \(SDDI\) versus overall body size was examined.
+\textbf{Figure 2}. Figure 2 legend here\hfill\break
+
+\textbf{Figure 3}. (A) Figure 3 legend (B) Second part of legend.
 \hfill\break
-
-\textbf{Figure 3}. (A) Set of linear measurements used to obtain head
-proportions. HL: head length, HW: head width, HH: head height, PL:
-pileus length, MO: mouth opening. (B) Locations of 17 landmarks (1-10;
-17) and semilandmarks (11-16) used to quantify head shape. (C) Patterns
-of sexual size dimorphism represented as male versus female size. (D)
-Patterns of sexual size dimorphism, represented as sexual size
-dimorphism index (\(SSDI\)) versus mean size. (E) Multivariate patterns
-sexual dimorphism in body proportions (\(SShDI\)) versus mean size. (F)
-Multivariate patterns of sexual shape dimorphism (\(SShDI\)) versus mean
-size. In panels C, D, E, and F, the red line indicates the regression
-estimate obtained from the data, while the dashed line corresponds to
-values representing no sexual dimorphism. \hfill\break
-
-\textbf{Figure 4}. Principal components plot of male (black squares) and
-female (red points) species mean head shape: the first two PC axes
-explain over 60\% of the total shape variation. Green lines correspond
-to the sexual shape dimorphism in one of the smallest species,
-\emph{Lacerta agilis bosnica}, and in the largest species, \emph{Timon
-lepidus}. Thin-plate spline deformation grids display the mean head
-shape for males and females of these species. \hfill\break
-
-\textbf{Figure 5}. Heat maps (green: lower values; red: higher values)
-displaying the evolution of (A) sexual size dimorphism, and (B) sexual
-shape dimorphism across the phylogeny. For each panel, an index of
-sexual size dimorphism (\(SSDI\)) or sexual shape dimorphism (\(SShDI\))
-is displayed, with circles being proportional to the degree of sexual
-dimorphism. \hfill\break
 
@@
 
-\singlespacing \textbf{Figure 1}. (A) Sexual size dimorphism for males
-and females across 100 hypothetical species. The pattern displayed
-exhibits a slope \textgreater{} 1.0, and thus corresponds to Rensch's
-rule. (B) The same data represented as the ratio of \(M/F\) sizes
-plotted against average species size. (C) Sexual dimorphism represented
-as the distance between males and females: \(SDD = D_{M:F}\). (D) Sexual
-dimorphism represented as an index of the SD Distance:
-\(SDDI = \pm1*D_{M:F}\). In all panels, the dashed line corresponds to
-values representing no sexual dimorphism (the 1:1 line). \hfill\break
+\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations
+across known lambda inputs on various tree sizes. As trees increase in
+size, the estimates more closely resemble the input lambdas, however
+considerable and concerning variation is apparent in trees smaller than
+those with 256 tips. \hfill\break
 
@@
 
-\includegraphics[width=0.7\linewidth]{Fig2}
-
-\singlespacing \textbf{Figure 2}. Results of phylogenetic simulations
-testing the type I error and statistical power for detecting patterns
-consistent with Rensch's rule. Vertical bars represent the proportion of
-datasets found to be significant. For type I error analyses, the dashed
-line marks the standard threshold of \(\alpha= 0.05\). For univariate
-data, both phylogenetic regression of male versus female size (standard
-approach), and \(SDDI\) versus overall body size (new approach) were
-performed. For multivariate data, only \(SDDI\) versus overall body size
-was examined. \hfill\break
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig3}
-
-\singlespacing \textbf{Figure 3}. (A) Set of linear measurements used to
-obtain head proportions. HL: head length, HW: head width, HH: head
-height, PL: pileus length, MO: mouth opening. (B) Locations of 17
-landmarks (1-10; 17) and semilandmarks (11-16) used to quantify head
-shape. (C) Patterns of sexual size dimorphism represented as male versus
-female size. (D) Patterns of sexual size dimorphism, represented as
-sexual size dimorphism index (\(SSDI\)) versus mean size. (E)
-Multivariate patterns sexual dimorphism in body proportions (\(SShDI\))
-versus mean size. (F) Multivariate patterns of sexual shape dimorphism
-(\(SShDI\)) versus mean size. In panels C, D, E, and F, the red line
-indicates the regression estimate obtained from the data, while the
-dashed line corresponds to values representing no sexual dimorphism.
-\hfill\break
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig4}
-
-\singlespacing \textbf{Figure 4}. Principal components plot of male
-(black squares) and female (red points) species mean head shape: the
-first two PC axes explain over 60\% of the total shape variation. Green
-lines correspond to the sexual shape dimorphism in one of the smallest
-species, \emph{Lacerta agilis bosnica}, and in the largest species,
-\emph{Timon lepidus}. Thin-plate spline deformation grids display the
-mean head shape for males and females of these species. \hfill\break
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig5}
-
-\singlespacing \textbf{Figure 5}. Heat maps (green: lower values; red:
-higher values) displaying the evolution of (A) sexual size dimorphism,
-and (B) sexual shape dimorphism across the phylogeny. For each panel, an
-index of sexual size dimorphism (\(SSDI\)) or sexual shape dimorphism
-(\(SShDI\)) is displayed, with circles being proportional to the degree
-of sexual dimorphism. \hfill\break
+\section{Other Figures Here}\label{other-figures-here}
 

### 2020-03-30 Data_Analyses/Simulation_script-PB-ANOVA.R

@@ -15,7 +15,7 @@ lambdas <- seq(0,1,.05) # starting with just 3 lambda values
 lambdas_expanded <- rep(lambdas, each = nsim)
 beta <- c(0, .25, .5, .75, 1) # true slope
 
-n <- treesizes[4] # number of tips (not simulations)
+n <- treesizes[1] # number of tips (not simulations)
 
 DataTable_lambda_0 <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
                    dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
@@ -43,9 +43,6 @@ for (j in 1:(length(lambdas_expanded))) {
     }
     
     X <- rTraitDisc(tree_scaled, model = "ER", states = c(0,1))
-    #X <- sim.char(trees_scaled, par = 1, nsim = 1)[,,1] # simmap for binary? email from dean, luke harmon
-           # We need to simulate a discrete variable:https://lukejharmon.github.io/pcm/chapter7_introdiscrete/
-            # Threshold model: http://blog.phytools.org/2013/09/threshold-character-evolution.html
     #comp_data_x <- comparative.data(tree, data.frame(X=X, Species = names(X)), Species, vcv = F, vcv.dim = 2)
     
     #pgls_x_lambda_0 <- try(pgls(X ~ 1, comp_data_x, lambda=0.01), silent = T)
@@ -63,11 +60,12 @@ for (j in 1:(length(lambdas_expanded))) {
     if(#!is(pgls_x_lambda_0, 'try-error') & !is(pgls_x_lambda_ml, 'try-error') &
        !is(pgls_xy_lambda_0, 'try-error') & !is(pgls_xy_lambda_ml, 'try-error')) break  }
   
+  
+DataTable_lambda_0[,5,j] <- unlist(lapply(1:length(beta), function(i) pgls_xy_lambda_0[[i]]$param[2])) # estimated lambdas of x and y
+DataTable_lambda_ml[,5,j] <- unlist(lapply(1:length(beta), function(i) pgls_xy_lambda_ml[[i]]$param[2])) # estimated lambdas of x and y
+ 
 DataTable_lambda_0[,6,j] <- unlist(lapply(1:length(beta), function(i) pgls_xy_lambda_0[[i]]$model$coef[2,1])) # slopes
[…]

### 2020-03-30 Data_Analyses/Simulation_script-PB-Reg.R

@@ -1,6 +1,6 @@
 # Simulating pgls inference errors from estimated lambdas
 
-# look at Revell ~2009 where he investigated Y~X simulating error in X and in Y or one or the other under BM
+# Set up
 setwd("/Users/ericabaken/Documents/School/Projects/Lambda_Eval/Manuscript/2020-Adams_and_Baken/")
 
 library(phytools)
@@ -16,17 +16,19 @@ lambdas_expanded <- rep(lambdas, each = nsim)
 beta <- c(0, .25, .5, .75, 1) # true slope
 fullsimlength <- length(lambdas_expanded)
 
-
 n <- treesizes[5] # number of tips: adjust each iteration
 
-DataTable_lambda_0 <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
-                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+# New Rep #####
+DataTable_lambda_0 <- array(NA, dim=c(length(beta), 7, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est.x", "lambda.est.xy", "slope")))
 
-DataTable_lambda_ml <- array(NA, dim=c(length(beta), 6, length(lambdas)*nsim), 
-                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est", "slope")))
+DataTable_lambda_ml <- array(NA, dim=c(length(beta), 7, length(lambdas)*nsim), 
+                   dimnames = list(beta, c("Rsq", "F", "P", "lambda.input", "lambda.est.x", "lambda.est.xy", "slope")))
 
 DataTable_lambda_0[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
 DataTable_lambda_ml[,4,] <- rep(lambdas_expanded, each = length(beta)) # putting in input lambda
+
+pb <- txtProgressBar(0, fullsimlength, style=1)
 
 Start.time<-Sys.time()
 Start.time
@@ -62,11 +64,14 @@ for (j in 1:(length(lambdas_expanded))) {
[…]

### 2020-03-30 Figures/Figures_Script.R

@@ -33,14 +33,15 @@ treesizes <-  2^(5:10)
 ## Figure 1 ######
 
 library(cowplot)
+library(ggplot2)
 
-DataArray <- array(data_reg_32, data_reg_64, data_reg_128,
+DataArray <- list(data_reg_32, data_reg_64, data_reg_128,
                    data_reg_256, data_reg_512, data_reg_1028)
 
 PlotList <- lapply(1:6, function(j){
    ggplot(DataArray[[j]], aes(x = lambda.input, y = lambda.est)) +
       geom_point() + theme(legend.position= "none", panel.background = element_rect("transparent")) + 
-      xlab("Input Lambda") + ylab("Estimated Lambda") + ggtitle(TreeSizeLabel)
+      xlab("Input Lambda") + ylab("Estimated Lambda") + ggtitle(treesizes[[j]])
 })
 
 jpeg("Figures/Fig1.jpeg", res = 120, quality = 100, width = 1000, height = 660)
@@ -48,19 +49,60 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
+jpeg("Manuscript/Fig1.jpeg", res = 120, quality = 100, width = 1000, height = 660)
+plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
+          PlotList[[4]], PlotList[[5]], PlotList[[6]])
+dev.off()
 
-## Figure 2 #######
+## Figure 2 ######
+data_reg_32_combo <- cbind(data_reg_32, data_reg_32_0)
+colnames(data_reg_32_combo) <- c(colnames(data_reg_32), "Rsq_0", "F_0", "P_0", "lambda.input_0", "lambda.est_0", "slope_0", "beta_0")
+
+data_reg_64_combo <- cbind(data_reg_64, data_reg_64_0)
+colnames(data_reg_64_combo) <- c(colnames(data_reg_64), "Rsq_0", "F_0", "P_0", "lambda.input_0", "lambda.est_0", "slope_0", "beta_0")
+
+data_reg_128_combo <- cbind(data_reg_128, data_reg_128_0)
+colnames(data_reg_128_combo) <- c(colnames(data_reg_128), "Rsq_0", "F_0", "P_0", "lambda.input_0", "lambda.est_0", "slope_0", "beta_0")
+
[…]

### 2020-04-07 Manuscript/Manuscript.Rmd
Draft 1 of manuscript still being worked on, but general intro finished. Figures 1, 2, and 3 constructed and included in manuscript.
Regression analyses for tree 1024 still running on lab server. All other data munged together, still needs fixing once lab server scripts are done.

@@
 
- conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)
+{conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)}
 
@@
 Investigating macroevolutionary patterns requires a phylogenetic approach as species are non-independent by nature of their shared ancestry.
-Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; [Felsenstein1985]), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
-[@Grafen1989] [@HarveyPagel1991] [@Rohlf2001]  [@AdamsNason2019]).
-Understanding the degree of phylogenetic signal present in a dataset is paramount, and identifies the mode with which a trait has evolved.
-High measures of phylogenetic signal indicate a Brownian motion process, whereas lower levels of phylogenetic signal indicate natural selection or some other evolutionary force has influenced the traits evolutionary history.
+Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; [@Felsenstein1985]), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
+[@Grafen1989]; [@HarveyPagel1991]; [@Rohlf2001]; [@AdamsNason2019]).
+A particularly useful measure in this field is phylogenetic signal, a quantification of a trait's correspondence with the clade's phylogenic history.
+Understanding the degree of phylogenetic signal present in a dataset is paramount and identifies the mode with which a trait has evolved; high measures of phylogenetic signal indicate a Brownian motion process, whereas lower levels of phylogenetic signal indicate natural selection or some other evolutionary force has influenced the traits evolutionary history.
 \hfill\break
@@
 For continuous data, the most common parameters used in the literature include Pagel's lambda [@Pagel1997] and Blomberg's kappa [@Blomberg1900].
-Pagel's lambda has the advantage of being nested in the likelihood framework and thus has also been utilized to encorporate phylogenetic signal while doing phylogenetic regressions and ANOVAs.
+Pagel's lambda has the advantage of being nested in the likelihood framework and thus has also been utilized to simultaneously estimate and account for phylogenetic signal while doing phylogenetic regressions or ANOVAs.
 At the development of these estimation methods, clear recommendations on the use and interpretation of this parameter were made based on data simulations and error evaluations [@Revell2010].
@@
 
-An earlier study [@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies could be for  estimating various evolutionary parameters, such as Pagel's lambda and ____.
-However, it remains unknown the degree to which lambda estimates appropriately represent degree of phylogenetic signal, and thus we fully explicate the extent to which Pagel's lambda estimation methods have been misused in recent macroevolutionary studies.
+An earlier study [@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies are when  estimating various evolutionary parameters, such as Pagel's lambda.
+However, it remains unknown the degree to which lambda estimates accurately represent degree of phylogenetic signal, and thus we fully explicate the extent to which Pagel's lambda estimation methods can and have been misused in recent macroevolutionary studies.
 Here we take a comprehensive approach to demonstrate the scenarios under which estimated Pagel's lambdas accurately reflect known lambdas.
-We then demonstrate the effect of these at times dubious estimation methods on significance testing when used in a pgls framework.
+We then demonstrate the effect of these, at times, dubious estimation methods on significance testing when used in a pgls framework.
 Finally, we perform a meta-analysis of research published in 2019 employing Pagel's lambda estimation methods to demonstrate the uses and misuses of this method in the literature.
@@
 
-To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies with known lambdas.
-We ____ by scaling simulated phylogeny with the scaling parameter and .
-<!-- obviously this paragraph sucks -->  We also did this with other tree shapes (symmetrical and ladder).
+To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies of increasing size, ranging in tip number from 32 to 1028.
+We then scaled the phylogenies by known lambda values ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size) and generated trait data by simulating a single continuous variable on each scaled phylogeny under Brownian motion.
+With the generated datasets, we used phylogenetic generalized least squares (PGLS) to estimate lambda values and extracted the model fit likelihood values.
+Finally, we calculated model fit likelihood values of PGLS models using the known lambdas as the input lambdas in the model and using ordinary least squares models (OLS; lambda = 0).
 \hfill\break
+
+To visualize the accuracy of the lambda estimation methods, we first plotted known lambdas (input lambdas) against the estimated lambdas.
+We also calculated rates of model misspecification for each phylogeny size, quantifying the rate at which models with known lambdas of 0 incorrectly define a significant phylogenetic signal as well as the rate at which models with known lambdas > 0 incorrectly estimate lambda as not significantly different from 0.
+\hfill\break
+
+Results from these data (Figure 1).
 
@@
 
-To ascertain the statistical performance of pgls  \hfill\break
+To ascertain the statistical performance of PGLS lambda estimation methods, we generated datasets with two trait variables across a range of known lambda values and correlation strengths.
+First, the independent variable (both continuous and binary) was generated in the same manner as the simulation procedures above.
+For each independent variable, we then generated a continuous dependent variable with input slopes ranging from 0 to 1 (0.25 intervals).
+From these data, used PGLS to estimate slope as well as F-values to assess the accuracy and error rates of these regression and ANOVA analyses while estimating lambdas.
+Finally, we again fit the dependent variable to the independent variable, estimated the slope and F-values using PGLS with the known lambdas as the input lambdas as well as under OLS.
+\hfill\break
+
+We calculated error rates ... \hfill\break
+
+Scaling and trait generation procedures for both simulation methods was repeated with symmetrical and ladder phylogeny shapes.
+Results of these phylogeny shapes were consistent with pure-birth phylogeny results and can be found in the Supplemental Materials.
+All analyses were performed in R v3.6.0 [@RCoreTeam] using the packages `geiger` [@HarmonEtAl2019] and `caper` [@Orme2011], and the corresponding scripts can be found in the Supplemental Materials.
+\hfill\break
+
+Results from these data (Figure 2).
 
@@
 
-All analyses were performed in R 3.6.2 [R-Base] using the packages `geomorph` [@AdamsOtarola2013; @AdamsGeomorph], `RRPP` [@CollyerAdams2018]. \hfill\break
+More methods paragraphs \hfill\break
+
+Results from these dta (Figure 3).
 
@@
 As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 256 tips.
-\hfill\break
 
-\textbf{Figure 2}. Figure 2 legend here\hfill\break
+\textbf{Figure 2}.
+Estimated ANOVA slopes under PGLS.
+Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
+However, for trees with fewer than 512 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
 
-\textbf{Figure 3}. (A) Figure 3 legend (B) Second part of legend. \hfill\break
+\textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The vast majority of these values were close to 0 or 1.
 
@@
 \newpage
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig2.png")
 
-# Other Figures Here
+```
+
+\singlespacing \textbf{Figure 2}.
+Estimated ANOVA slopes under PGLS.
+Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
+However, for trees with fewer than 512 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
+\hfill\break
+
+\newpage
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig3.png")
+
+```
+
+\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The vast majority of these values were close to 0 or 1.

### 2020-04-07 Manuscript/Manuscript.tex
Draft 1 of manuscript still being worked on, but general intro finished. Figures 1, 2, and 3 constructed and included in manuscript.
Regression analyses for tree 1024 still running on lab server. All other data munged together, still needs fixing once lab server scripts are done.

@@
 
-conclusion holds: interpreting the regression is not appreciably
-different (in terms of slopes and f values)
+\{conclusion holds: interpreting the regression is not appreciably
+different (in terms of slopes and f values)\}
 
@@
 ancestry. Since the first appropriate method was introduced by
-Felsenstein (phylogenetic independent contrasts; {[}Felsenstein1987{]}),
+Felsenstein (phylogenetic independent contrasts; (Felsenstein 1985)),
 dozens of other methods have been developed and applied to increasingly
-complex questions in macroevolutionary biology (e.g.
-{[}AdamsNason2019{]}). Understanding the degree of phylogenetic signal
-present in a dataset is paramount, and identifies the mode with which a
-trait has evolved. High measures of phylogenetic signal indicate a
-Brownian motion process, whereas lower levels of phylogenetic signal
-indicate natural selection or some other evolutionary force has
-influenced the traits evolutionary history. \hfill\break
+complex questions in macroevolutionary biology (e.g. (Grafen 1989);
+(Harvey and Pagel 1991); (Rohlf 2001); ({\textbf{???}})). A particularly
+useful measure in this field is phylogenetic signal, a quantification of
+a trait's correspondence with the clade's phylogenic history.
+Understanding the degree of phylogenetic signal present in a dataset is
+paramount and identifies the mode with which a trait has evolved; high
+measures of phylogenetic signal indicate a Brownian motion process,
+whereas lower levels of phylogenetic signal indicate natural selection
+or some other evolutionary force has influenced the traits evolutionary
+history. \hfill\break
 
@@
 data, the most common parameters used in the literature include Pagel's
-lambda {[}Pagel1997{]} and Blomberg's kappa {[}Blomberg1900{]}. Pagel's
-lambda has the advantage of being couched in the likelihood framework
-and thus has also been utilized to encorporate phylogenetic signal while
-doing phylogenetic regressions and ANOVA. However, the accuracy of the
-lambda estiamtion methods have not been fully evaluated, and thus it
-remains unknown the degree to which lambda estimates appropriately
-represent degree of phylogenetic signal. \hfill\break
+lambda ({\textbf{???}}) and Blomberg's kappa ({\textbf{???}}). Pagel's
+lambda has the advantage of being nested in the likelihood framework and
+thus has also been utilized to simultaneously estimate and account for
+phylogenetic signal while doing phylogenetic regressions or ANOVAs. At
+the development of these estimation methods, clear recommendations on
+the use and interpretation of this parameter were made based on data
+simulations and error evaluations (Revell 2010). However, the accuracy
+of the lambda estiamtion methods have not been fully evaluated, and
+consequently many studies have substantially misused these methods.
+\hfill\break
 
-An earlier study {[}BoettigerEtAl2012{]}, breifly addressed this topic
-by showing how uninformative smaller phylogenies could be using
-estimation methods for various parameters. That paper concluded that a
-measure of power must be considered when quantifying Pagel's lambda.
-Here we take a more comprehensive approach to demonstrate the scenarios
-under which estimated lambdas accurately reflect known lambdas as well
-as the effect of these at times dubious estimation methods on
-significance testing when used in a pgls framework.
+An earlier study ({\textbf{???}}), breifly addressed this topic by
+showing how uninformative smaller phylogenies are when estimating
+various evolutionary parameters, such as Pagel's lambda. However, it
+remains unknown the degree to which lambda estimates accurately
+represent degree of phylogenetic signal, and thus we fully explicate the
+extent to which Pagel's lambda estimation methods can and have been
+misused in recent macroevolutionary studies. Here we take a
+comprehensive approach to demonstrate the scenarios under which
+estimated Pagel's lambdas accurately reflect known lambdas. We then
+demonstrate the effect of these, at times, dubious estimation methods on
+significance testing when used in a pgls framework. Finally, we perform
+a meta-analysis of research published in 2019 employing Pagel's lambda
+estimation methods to demonstrate the uses and misuses of this method in
+the literature. This investigation provides evidence for the
+circumstances under which estimating Pagel's lambda can be informative,
+and which interpretations or uses are misleading with the hope of moving
+the field towards more appropriate employment of our methodological
+toolbox.
 
@@
 To assess the accuracy of Pagel's lambda estimations, we simulated
-pure-birth phylogenies with known lambdas. We \_\_\_\_ by scaling
-simulated phylogeny with the scaling parameter and . We also did this
-with other tree shapes (symmetrical and ladder). \hfill\break
+pure-birth phylogenies of increasing size, ranging in tip number from 32
+to 1028. We then scaled the phylogenies by known lambda values ranging
+from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size)
+and generated trait data by simulating a single continuous variable on
+each scaled phylogeny under Brownian motion. With the generated
+datasets, we used phylogenetic generalized least squares (PGLS) to
+estimate lambda values and extracted the model fit likelihood values.
+Finally, we calculated model fit likelihood values of PGLS models using
+the known lambdas as the input lambdas in the model and using ordinary
+least squares models (OLS; lambda = 0). \hfill\break
+
+To visualize the accuracy of the lambda estimation methods, we first
+plotted known lambdas (input lambdas) against the estimated lambdas. We
+also calculated rates of model misspecification for each phylogeny size,
+quantifying the rate at which models with known lambdas of 0 incorrectly
+define a significant phylogenetic signal as well as the rate at which
+models with known lambdas \textgreater{} 0 incorrectly estimate lambda
+as not significantly different from 0. \hfill\break
+
+Results from these data (Figure 1).
 
@@
 
-To ascertain the statistical performance of pgls \hfill\break
+To ascertain the statistical performance of PGLS lambda estimation
+methods, we generated datasets with two trait variables across a range
+of known lambda values and correlation strengths. First, the independent
+variable (both continuous and binary) was generated in the same manner
+as the simulation procedures above. For each independent variable, we
+then generated a continuous dependent variable with input slopes ranging
+from 0 to 1 (0.25 intervals). From these data, used PGLS to estimate
+slope as well as F-values to assess the accuracy and error rates of
+these regression and ANOVA analyses while estimating lambdas. Finally,
+we again fit the dependent variable to the independent variable,
+estimated the slope and F-values using PGLS with the known lambdas as
+the input lambdas as well as under OLS. \hfill\break
+
+We calculated error rates \ldots{} \hfill\break
+
+Scaling and trait generation procedures for both simulation methods was
+repeated with symmetrical and ladder phylogeny shapes. Results of these
+phylogeny shapes were consistent with pure-birth phylogeny results and
+can be found in the Supplemental Materials. All analyses were performed
+in R v3.6.0 ({\textbf{???}}) using the packages \texttt{geiger}
+({\textbf{???}}) and \texttt{caper} ({\textbf{???}}), and the
+corresponding scripts can be found in the Supplemental Materials.
+\hfill\break
+
+Results from these data (Figure 2).
 
@@
 
-All analyses were performed in R 3.6.2 {[}R-Base{]} using the packages
-\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2019),
-\texttt{RRPP} (Collyer and Adams 2018). \hfill\break
+More methods paragraphs \hfill\break
+
+Results from these dta (Figure 3).
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-AdamsOtarola2013}{}
-Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
-the collection and analysis of geometric morphometric shape data.
-Methods in Ecology and Evolution 4:393--399.
+\hypertarget{ref-Felsenstein1985}{}
+Felsenstein, J. 1985. Phylogenies and the comparative method. American
+Naturalist 125:1--15.
 
-\hypertarget{ref-AdamsGeomorph}{}
-Adams, D., M. Collyer, and A. Kaliontzopoulou. 2019. Geomorph: Software
-for geometric morphometric analyses. r package version 3.1.1.
+\hypertarget{ref-Grafen1989}{}
+Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
+of the Royal Society of London B, Biological Sciences 326:119--157.
 
-\hypertarget{ref-CollyerAdams2018}{}
-Collyer, M. L., and D. C. Adams. 2018. RRPP: An R package for fitting
-linear models to high-dimensional data using residual randomization.
-Methods in Ecology and Evolution 9:1772--1779.
+\hypertarget{ref-HarveyPagel1991}{}
+Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
+evolutionary biology. Oxford University Press, Oxford.
+
+\hypertarget{ref-Revell2010}{}
+Revell, L. J. 2010. Phylogenetic signal and linear regression on species
+data. Methods in Ecology and Evolution 1:319--329.
+
+\hypertarget{ref-Rohlf2001}{}
+Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
+variables: Geometric interpretations. Evolution 55:2143--2160.
 
@@
 and concerning variation is apparent in trees smaller than those with
-256 tips. \hfill\break
+256 tips.
 
-\textbf{Figure 2}. Figure 2 legend here\hfill\break
+\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes,
+the mean estimated slope matches the input slope, and as trees increase
+in size, the variance around this mean estimate decreases. However, for
+trees with fewer than 512 tips, the error around the estimated slope is
+considerable, where these analyses frequently estimate slopes in the
+opposite direction of the known pattern.
 
-\textbf{Figure 3}. (A) Figure 3 legend (B) Second part of legend.
-\hfill\break
+\textbf{Figure 3}. Frequency of estimated lambda values published in
+manuscripts in 2019. The vast majority of these values were close to 0
+or 1.
 
@@
 
-\section{Other Figures Here}\label{other-figures-here}
+\includegraphics[width=0.95\linewidth]{Fig2}
+
+\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS.
+Across tree sizes, the mean estimated slope matches the input slope, and
+as trees increase in size, the variance around this mean estimate
+decreases. However, for trees with fewer than 512 tips, the error around
+the estimated slope is considerable, where these analyses frequently
+estimate slopes in the opposite direction of the known pattern.
+\hfill\break
+
+\newpage
+
+\includegraphics[width=0.95\linewidth]{Fig3}
+
+\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
+published in manuscripts in 2019. The vast majority of these values were
+close to 0 or 1.
 

### 2020-04-07 Data_Analyses/Munging_script.R

@@ -1,0 +1,73 @@
+# Munging Data
+tree.sizes <- 2^(5:10)
+
+# Regression data
+
+data_reg_32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32.csv")
+data_reg_64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-64.csv")
+data_reg_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-128.csv")
+data_reg_256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-256.csv")
+data_reg_512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512.csv")
+data_reg_1024_1 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1024-1.csv")
+data_reg_1024_2 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1024-2.csv")
+data_reg_1024_3 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1024-3.csv")
+data_reg_1024_4 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1024-4.csv")
+data_reg_1024_5 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1024-5.csv")
+
+data_reg_32_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32.csv")
+data_reg_64_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-64.csv")
+data_reg_128_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-128.csv")
+data_reg_256_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-256.csv")
+data_reg_512_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-512.csv")
+data_reg_1024_0_1 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1024-1.csv")
+data_reg_1024_0_2 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1024-2.csv")
+data_reg_1024_0_3 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1024-3.csv")
+data_reg_1024_0_4 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1024-4.csv")
[…]

### 2020-04-07 Data_Analyses/Simulation_script-PB-Reg.R

@@ -6,7 +6,7 @@ setwd("/Users/ericabaken/Documents/School/Projects/Lambda_Eval/Manuscript/2020-A
 library(phytools)
 library(geomorph)
 library(geiger)
-library(svMisc)
+#library(svMisc)
 library(caper)
 
 treesizes <-  2^(5:10) # 32,64,128,256,512,1024

### 2020-04-07 Figures/Figures_Script.R

@@ -1,45 +1,20 @@
 # Figures
-
-data_reg_32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32.csv")
-data_reg_64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-64.csv")
-data_reg_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-128.csv")
-data_reg_256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-256.csv")
-data_reg_512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512.csv")
-data_reg_1028 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1028.csv")
-
-data_anova_32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-32-ANOVA.csv")
-data_anova_64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-64-ANOVA.csv")
-data_anova_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-128-ANOVA.csv")
-data_anova_256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-256-ANOVA.csv")
-data_anova_512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-512-ANOVA.csv")
-data_anova_1028 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_ml-1028-ANOVA.csv")
-
-data_reg_32_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32.csv")
-data_reg_64_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-64.csv")
-data_reg_128_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-128.csv")
-data_reg_256_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-256.csv")
-data_reg_512_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-512.csv")
-data_reg_1028_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-1028.csv")
-
-data_anova_32_0 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_0-32-ANOVA.csv")
[…]

### 2020-04-08 Manuscript/Manuscript.tex
Finishing very first rough draft of manuscript to send towards Dean. Figure 2 generated with alternative options. Supplemental Figures also available in Figures folder.
Discussion not yet started.

@@
 ancestry. Since the first appropriate method was introduced by
-Felsenstein (phylogenetic independent contrasts; (Felsenstein 1985)),
+Felsenstein (phylogenetic independent contrasts; Felsenstein (1985)),
 dozens of other methods have been developed and applied to increasingly
-complex questions in macroevolutionary biology (e.g. (Grafen 1989);
-(Harvey and Pagel 1991); (Rohlf 2001); ({\textbf{???}})). A particularly
+complex questions in macroevolutionary biology (e.g. Grafen (1989);
+Harvey and Pagel (1991); Rohlf (2001); ({\textbf{???}})). A particularly
 useful measure in this field is phylogenetic signal, a quantification of
@@
 Understanding the degree of phylogenetic signal present in a dataset is
-paramount and identifies the mode with which a trait has evolved; high
+paramount and identifies the mode under which a trait has evolved; high
 measures of phylogenetic signal indicate a Brownian motion process,
@@
 demonstrate the effect of these, at times, dubious estimation methods on
-significance testing when used in a pgls framework. Finally, we perform
+significance testing when used in a PGLS framework. Finally, we perform
 a meta-analysis of research published in 2019 employing Pagel's lambda
@@
 and which interpretations or uses are misleading with the hope of moving
-the field towards more appropriate employment of our methodological
-toolbox.
+the field towards more appropriate employment of phylogenetic
+comparative methods.
 
@@
 To assess the accuracy of Pagel's lambda estimations, we simulated
-pure-birth phylogenies of increasing size, ranging in tip number from 32
-to 1028. We then scaled the phylogenies by known lambda values ranging
-from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size)
-and generated trait data by simulating a single continuous variable on
-each scaled phylogeny under Brownian motion. With the generated
-datasets, we used phylogenetic generalized least squares (PGLS) to
-estimate lambda values and extracted the model fit likelihood values.
-Finally, we calculated model fit likelihood values of PGLS models using
-the known lambdas as the input lambdas in the model and using ordinary
-least squares models (OLS; lambda = 0). \hfill\break
+pure-birth phylogenies of variable size, ranging in tip number from 32
+to 1024. We then scaled the simulated phylogenies by lambda values
+ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree
+size) with which we generated trait data with known lambdas by
+simulating a continuous variable on each scaled phylogeny under Brownian
+motion. We then estimated lambda values from these data using
+phylogenetic generalized least squares (PGLS) to compare against the
+known lambda values. \hfill\break
 
 To visualize the accuracy of the lambda estimation methods, we first
-plotted known lambdas (input lambdas) against the estimated lambdas. We
-also calculated rates of model misspecification for each phylogeny size,
-quantifying the rate at which models with known lambdas of 0 incorrectly
-define a significant phylogenetic signal as well as the rate at which
-models with known lambdas \textgreater{} 0 incorrectly estimate lambda
-as not significantly different from 0. \hfill\break
+plotted known lambdas (input lambdas) against the estimated lambdas
+(Figure 1). This plot demonstrates the rampant inaccuracy of estimating
+lambdas on phylogenies with fewer than 200 tips, as the spread of data
+in the upper panels in Figure 1 is remarkably wide. We also see that the
+widest spread of estimated lambdas is observed near the center of each
+plot corresponding to intermediate values of known lambda. Lastly, we
+see a slight tendency for the PGLS estimation methods to underestimate
+lambda, especially for known lambda values below 0.5 as can be seen by
+the numerous data points along the x axis for the smaller phylogenies
+analyzed (n tips \textless{} 200). \hfill\break
 
-Results from these data (Figure 1).
+{[}insert Figure 1 here{]}
 
@@
 
-To ascertain the statistical performance of PGLS lambda estimation
-methods, we generated datasets with two trait variables across a range
-of known lambda values and correlation strengths. First, the independent
-variable (both continuous and binary) was generated in the same manner
-as the simulation procedures above. For each independent variable, we
-then generated a continuous dependent variable with input slopes ranging
-from 0 to 1 (0.25 intervals). From these data, used PGLS to estimate
-slope as well as F-values to assess the accuracy and error rates of
-these regression and ANOVA analyses while estimating lambdas. Finally,
-we again fit the dependent variable to the independent variable,
-estimated the slope and F-values using PGLS with the known lambdas as
-the input lambdas as well as under OLS. \hfill\break
+To measure the statistical performance of PGLS lambda estimation methods
+when applied to ANOVA and regression analyses, we used the above
+generated data (independent variables) to simulate a second set of trait
+variables (depedendent variables) across the range of correlation
+strengths (betas ranging from 0 to 1 at intervales of 0.25). We then
+used PGLS to estimate phylogenetic signal of the dependent variable and
+the slope coefficient for the regression between the dependent and
+independent variables. We also calculated F- and p-values from the
+regression analyses. Finally, we again fit the dependent variable to the
+independent variable while holding the lambda value at 0 to approximate
+an ordinary least squares (OLS) apprach. \hfill\break
 
-We calculated error rates \ldots{} \hfill\break
+We compared the estimated slope coefficients across variable input beta
+values to assess the ability of PGLS to identify significant
+correlations (Figure 2). All distribution groups center around the 1:1
+relationship between input beta values and estimated slopes. However the
+variance around that mean is substantial, with estimates ranging from
+approximately -2 to 4 in datasets with 32 taxa and a known relationship
+of (beta = 1). This variance of the estimated slopes only becomes
+reasonable when phylogenies have over 200 tips, similarly to what we saw
+in the first simulation analysis in this study. \hfill\break
 
-Scaling and trait generation procedures for both simulation methods was
-repeated with symmetrical and ladder phylogeny shapes. Results of these
-phylogeny shapes were consistent with pure-birth phylogeny results and
-can be found in the Supplemental Materials. All analyses were performed
-in R v3.6.0 ({\textbf{???}}) using the packages \texttt{geiger}
-({\textbf{???}}) and \texttt{caper} ({\textbf{???}}), and the
-corresponding scripts can be found in the Supplemental Materials.
-\hfill\break
+{[}insert Figure 2 here{]} \hfill\break
 
-Results from these data (Figure 2).
+Surprisingly, the estimated lambda values of the dependent variable do
+not correspond with input lambda values that characterize the
+independent variable (Figure S1). This disconnect then resulted in no
+appreciable difference of slope estimates across input lambda values
+(Figure S2), nor do we see substantial differences in slope estimates
+between the PGLS and OLS analyses (results not shown). Thus, we show
+that phylogenetic signal present in one variable does not translate to
+phylogenetic signal in a second, even highly correlated, dependent
+variable. \hfill\break
+
+Scaling and trait generation procedures for all simulation methods were
+repeated with symmetrical and ladder phylogenies of variable size.
+Results generated using these variable phylogeny shapes were consistent
+with the pure-birth phylogeny results presented above and can be found
+in the Supplemental Materials. All analyses were performed in R v3.6.0
+({\textbf{???}}) using the packages \texttt{geiger} ({\textbf{???}}) and
+\texttt{caper} ({\textbf{???}}), and the corresponding scripts can be
+found in the Supplemental Materials.
 
@@
 
-Despite the urging of Boettiger and colleagues to publish confidence
-intervals with all lambda parameter estimates, only 18\% of papers
-published in 2019 do so. \hfill\break
+To understand the extent of this problematic estimation method in
+application, we performed a meta-analysis of studies published in 2019
+that cited Pagel's 1997 manuscript ({\textbf{???}}). The list of
+manuscripts was compiled through Google Scholar on Jan 23, 2020 and
+totaled 341 manuscripts. For each study, we extracted any published
+lambda estimates, along with the size of the phylogeny used in the
+analysis. We also noted whether authors reported confidence intervals,
+significance tests assessing difference of the lambda estimate from 0 or
+1, and whether authors interpreted biological meaning from the magnitude
+of the estimated lambda. For studies that reported more than one lambda
+estimate, we also noted if the authors compared the lambdas against one
+another, and whether that was accompanied with an appropriate
+statistical test between the estimated lambda values. \hfill\break
 
-More methods paragraphs \hfill\break
+We found 182 manuscripts from 2019 that estimated and reported Pagel's
+lambda values using PGLS methods. These papers averaged 8.527 lambda
+values, ranging from a single lambda estimate up to 71 estimated
+lambdas. Almost exactly half of the published lambda estimates were
+either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3). 73.32\% of
+the published lambdas were estimated using phylogenies with fewer than
+200 tips, and 348 lambda estimates (8.57\% of all published estimates)
+came from phylogenies with fewer than 30 tips. \hfill\break
 
-Results from these dta (Figure 3).
+{[}insert Figure 3 here{]} \hfill\break
+
+Many of the reviewed manuscripts liberally interpreted the magnitude of
+the estimate lambda, using phrases such as ``strong'' or ``weak''
+phylogentic signal when statistically, all that was clear was a
+difference between the estimated lambda and 0 or 1 respectively. We
+estimated that about 20.49\% of the manuscripts revealed some sort of
+biological interpretation of the magnitude of estimated phylogenetic
+signal that overreached the statistical findings. We also identified
+seven manuscripts as having inappropriately interpreted differences in
+lambda values, indicating that some traits had stronger or weaker signal
+than other traits without the appropriate statistical tests.
+\hfill\break
+
+As is evidenced by macroevolutionary papers published in 2019 papers,
+Pagel's lambda estimation methods are often misused and
+over-interpretted. Despite the urging of Boettiger and colleagues to
+publish confidence intervals with all lambda parameter estimates, only
+18\% of papers published in 2019 do so.
 
@@
 
-Using the estimated lambda values from pgls are not useful. The
-questions of whether or not signal exists is appropriate, but inferring
-more from lambda \emph{magnitude} is inappropriate. \hfill\break
+{[}This part is obviously not written yet{]} \hfill\break
+
+General conclusions :Using the estimated lambda values from pgls are not
+useful. The questions of whether or not signal exists is appropriate,
+but inferring more from lambda \emph{magnitude} is inappropriate.
+\hfill\break
 
@@
 and concerning variation is apparent in trees smaller than those with
-256 tips.
+200 tips. \hfill\break
 
@@
 in size, the variance around this mean estimate decreases. However, for
-trees with fewer than 512 tips, the error around the estimated slope is
+trees with fewer than 200 tips, the error around the estimated slope is
 considerable, where these analyses frequently estimate slopes in the
-opposite direction of the known pattern.
+opposite direction of the known pattern. \hfill\break
 
 \textbf{Figure 3}. Frequency of estimated lambda values published in
-manuscripts in 2019. The vast majority of these values were close to 0
-or 1.
+manuscripts in 2019. The majority of these values were close to 0 or 1,
+and from phylogenies with fewer than 200 taxa.
 
@@
 considerable and concerning variation is apparent in trees smaller than
-those with 256 tips. \hfill\break
+those with 200 tips. \hfill\break
 
@@
 as trees increase in size, the variance around this mean estimate
-decreases. However, for trees with fewer than 512 tips, the error around
+decreases. However, for trees with fewer than 200 tips, the error around
 the estimated slope is considerable, where these analyses frequently
@@
 \singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
-published in manuscripts in 2019. The vast majority of these values were
-close to 0 or 1.
+published in manuscripts in 2019. The majority of these values were
+close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 

### 2020-04-08 Data_Analyses/Analysis_script.R

@@ -1,0 +1,105 @@
+# Analyses
+
+treesizes <-  2^(5:10)
+
+ANOVA.data <- read.csv("Data_Analyses/Munged_Data/ANOVA.fulldataset.csv")
+ANOVA.data <- as.data.frame(ANOVA.data)
+
+Regression.data <- read.csv("Data_Analyses/Munged_Data/Regression.fulldataset.csv")
+Regression.data <- as.data.frame(Regression.data)
+
+# Literature numbers ####
+library(readxl)
+review <- read_excel("Literature_Review/1Summary.xlsx", sheet = "Since 2019 Lambda Vals")
+colnames(review)
+length(unique(review$Reference)) #182
+
+# Published lambda values
+published.lambdas <- review[,c(1,3,4)]
+published.lambdas.pruned <- published.lambdas[-which(published.lambdas$`Estimated Lambdas Simple` > 1 | published.lambdas$`Estimated Lambdas Simple` < 0),]
+published.lambdas.df <- as.data.frame(published.lambdas.pruned)
+nrow(published.lambdas.df) # 1552
+1552/182 # average number of reported lambda values per manuscript
+colnames(published.lambdas.df) <- c("Refs", "Published_Lambdas", "Taxa")
+published.lambdas.df$Refs <- as.factor(published.lambdas.df$Refs)
+
+min(table(published.lambdas.df$Refs)) # 1: lowest number of reported lambdas per paper
+max(table(published.lambdas.df$Refs)) # 71: highest number of reported lambdas per paper
+
+length(which(published.lambdas.df$Published_Lambdas < 0.05))/nrow(published.lambdas.df)*100 # 25.43%
+length(which(published.lambdas.df$Published_Lambdas > 0.90))/nrow(published.lambdas.df)*100 # 24.74%
+
+length(which(published.lambdas.df$Taxa < 30))/nrow(published.lambdas.df)*100 # 75.32%
+length(which(published.lambdas.df$Taxa < 30)) # 348
+
+
+
+
+# Numerical summaries of all papers
+review.fullsheet <- read_excel("Literature_Review/1Summary.xlsx", sheet = "Since 2019")
+review.fullsheet <- as.data.frame(review.fullsheet)
+
+all.papers <- review.fullsheet[,c("Reference", "Estimated lambdas")]
+all.papers$YN <- all.papers$`Estimated lambdas`
+
[…]

### 2020-04-08 Data_Analyses/Munging_script.R

@@ -73,6 +73,6 @@ anova.fulldata <- rbind(data_anova_32, data_anova_64, data_anova_128, data_anova
 anova.fulldata <- as.data.frame(anova.fulldata)
 
 anova.fulldata$tree.size <- rep(rep(as.character(tree.sizes), each = 5250), 2)
-reg.fulldata$method <- rep(c("ml", "0"), each = length(reg.fulldata$tree.size)/2)
+anova.fulldata$method <- rep(c("ml", "0"), each = length(anova.fulldata$tree.size)/2)
 
 write.csv(anova.fulldata, "Data_Analyses/Munged_Data/ANOVA.fulldataset.csv", row.names = F)

### 2020-04-08 Figures/Figures_Script.R

@@ -30,6 +30,7 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
+
 ## Figure 2 ######
 
 ANOVA.data$beta <- as.factor(ANOVA.data$beta)
@@ -53,7 +54,23 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# other plot options for Figure 2 
+# other plot options for Figure 2 - alt1
+
+
+PlotList <- lapply(1:6, function(j){
+   reduced.data <- ANOVA.data[which(ANOVA.data$tree.size == treesizes[[j]]),]
+   ggplot(reduced.data, aes(x = beta, y = slope, fill = method)) + #geom_abline(intercept = 0, slope = .2) +
+      geom_boxplot() + theme(legend.position= "none", panel.background = element_rect("transparent")) + 
+      xlab("Input Slope") + ylab("Estimated Slope") + ggtitle(treesizes[[j]]) + expand_limits(y=c(-3.5,4))
+})
+
+jpeg("Figures/Fig2_alt.jpeg", res = 120, quality = 100, width = 1000, height = 660)
+plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
+          PlotList[[4]], PlotList[[5]], PlotList[[6]])
+dev.off()
+
+
+# another plot options for Figure 2 -alt2
 
 ANOVA.data.0 <- ANOVA.data[which(ANOVA.data$method == "0"),]
  
@@ -69,8 +86,11 @@ PlotList <- lapply(1:6, function(j){
       xlab("F from PGLS (est lambda)") + ylab("F from OLS") + ggtitle(treesizes[[j]]) 
 })
 
+jpeg("Figures/Fig2_alt2.jpeg", res = 120, quality = 100, width = 1000, height = 660)
 plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
+dev.off()
+
 
 
 ## Figure 3 #####
@@ -142,6 +162,28 @@ PlotList <- lapply(1:6, function(j){
 })
 
[…]

### 2020-04-14 Manuscript/Manuscript-1-DCA.Rmd
paper vsn 1-DCA

---
title:

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: evolution.csl
bibliography: Lambda-Refs.bib

header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---
# Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate

\hfill\break

**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break

**Short Title**: Inaccuracies in Pagel's Lambda \hfill\break

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

{conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)}

\newpage

# Introduction
Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species generates statistical non-independence [@Felsenstein1985; @HarveyPagel1991]. Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs); a suite of analytical tools that condition the data on the phylogeny through the course of statisical evaluations of phenotypic trends [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004]. The past several decades have witnessed a rapid expansion in the development of PCMs to address an ever-growing set of macroevolutionary hypotheses [@MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @Adams2014a; @AdamsCollyer2018b]. These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Munkemuller_et_al2012]. Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [see @Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008]. \hfill\break

In many macroevolutionary studies, it is often of interest to quantify the degree to which phylogenetic signal is displayed in continuous traits. Several analytical tools have been developed for this purpose [e.g., @Gittleman1990; @Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Klingenberg2010; @Adams2014a], which differ primarily in how they characterize the phylogenetic dependency of trait variation among taxa. One commonly used statistical measure, *Kappa* ($K$), expresses the strength of phylogenetic signal as the ratio of observed trait variation to the trait variation conditioned on the phylogeny; scaled by what is expected under Brownian motion given the phylogeny's size and shape [@Blomberg_et_al2003; also @Adams2014a]. Another approach, Pagel's $\lambda$ [@Pagel1999], uses maximum likelihood to fit the data to the phylogeny under some model of evolutionary change (typically Brownian motion). The inclusion of a scaling parameter, $\lambda$,  transforms the lengths of the internal branches of the phylogeny to improve the fit, and this parameter describes the degree of phylogenetic signal in the dataset [@Pagel1999; @Freckleton_et_al2002]. Pagel's $\lambda$ also has the advantage that it may be included when estimating the association of traits in a phylogenetic context, meaning that one may account for the degree of phylogenetic signal while conducting phylogenetic regression or ANOVA [see @Freckleton_et_al2002]. \hfill\break

Several studies have investigated the statistical properties of methods for estimating phylogenetic signal under various conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010]. These have largely focused on the ability of methods to detect the presence of phylogenetic signal (i.e., type I and type II error rates) under complex models of evolutionary change, across a range of phylogeny sizes, and with varying degrees of phylogenetic uncertainty or unresolved topologies. In terms of parameter estimation, Revell [-@Revell2010] confirmed that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al. [-@Munkemuller_et_al2012] found that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger. However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012]. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation. However, this conclusion assumes that parameter estimation remains equally precise across its range of values ($\lambda = 0 \to 1$); an assumption that to date, has not been investigated. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.  \hfill\break

In this study, we evaluate the precision of Pagel's $\lambda$ to estimate known levels of phylogenetic signal in phenotypic data. First we use computer simulations across differing numbers of species, and differing input levels of phylogenetic signal to explore when $\lambda$ captures known levels of phylogenetic signal, and under what circumstances. Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic ANOVA and regression (i.e., PGLS) affects parameter estimation, and whether estimates of phylogenetic signal in PGLS are accurate. We then survey the recent macroevolutionary literature for published papers containing estimates of $\lambda$ from empirical datasets, and compare these empirical estimates to patterns gleaned from our computer simulations. In general we find that while PGLS parameters are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not. Thus, empirical studies should interpret the degree of phylogenetic signal in phenotypic data with caution when using this parameter.  **something more?  Let's wait for final data to decide.**

# Methods and Results

## *Simulated trait*

To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies of variable size, ranging in tip number from 32 to 1024. We then scaled the simulated phylogenies by lambda values ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size) with which we generated trait data with known lambdas by simulating a continuous variable on each scaled phylogeny under Brownian motion. We then estimated lambda values from these data using phylogenetic generalized least squares (PGLS)  to compare against the known lambda values. \hfill\break

To visualize the accuracy of the lambda estimation methods, we first plotted known lambdas (input lambdas) against the estimated lambdas (Figure 1). This plot demonstrates the rampant inaccuracy of estimating lambdas on phylogenies with fewer than 200 tips, as the spread of data in the upper panels in Figure 1 is remarkably wide. We also see that the widest spread of estimated lambdas is observed near the center of each plot corresponding to intermediate values of known lambda. Lastly, we see a slight tendency for the PGLS estimation methods to underestimate lambda, especially for known lambda values below 0.5 as can be seen by the numerous data points along the x axis for the smaller phylogenies analyzed (n tips < 200). \hfill\break

[insert Figure 1 here]

## *Simulated ANOVA and Regressions*

To measure the statistical performance of PGLS lambda estimation methods when applied to ANOVA and regression analyses, we used the above generated data (independent variables) to simulate a second set of trait variables (depedendent variables) across the range of correlation strengths (betas ranging from 0 to 1 at intervales of 0.25). We then used PGLS to estimate phylogenetic signal of the dependent variable and the slope coefficient for the regression between the dependent and independent variables. We also calculated F- and p-values from the regression analyses. Finally, we again fit the dependent variable to the independent variable while holding the lambda value at 0 to approximate an ordinary least squares (OLS) apprach. \hfill\break

We compared the estimated slope coefficients across variable input beta values to assess the ability of PGLS to identify significant correlations (Figure 2). All distribution groups center around the 1:1 relationship between input beta values and estimated slopes. However the variance around that mean is substantial, with estimates ranging from approximately -2 to 4 in datasets with 32 taxa and a known relationship of (beta = 1). This variance of the estimated slopes only becomes reasonable when phylogenies have over 200 tips, similarly to what we saw in the first simulation analysis in this study. \hfill\break

[insert Figure 2 here] \hfill\break

Surprisingly, the estimated lambda values of the dependent variable do not correspond with input lambda values that characterize the independent variable (Figure S1). This disconnect then resulted in no appreciable difference of slope estimates across input lambda values (Figure S2), nor do we see substantial differences in slope estimates between the PGLS and OLS analyses (results not shown). Thus, we show that phylogenetic signal present in one variable does not translate to phylogenetic signal in a second, even highly correlated, dependent variable. \hfill\break

Scaling and trait generation procedures for all simulation methods were repeated with symmetrical and ladder phylogenies of variable size. Results generated using these variable phylogeny shapes were consistent with the pure-birth phylogeny results presented above and can be found in the Supplemental Materials. All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008] and `caper` [@Orme2013], and the corresponding scripts can be found in the Supplemental Materials.

## *Meta-Analysis of Empirical Results*

To understand the extent of this problematic estimation method in application, we performed a meta-analysis of studies published in 2019 that cited Pagel's 1997 manuscript [@Pagel1999]. The list of manuscripts was compiled through Google Scholar on Jan 23, 2020 and totaled 341 manuscripts. For each study, we extracted any published lambda estimates, along with the size of the phylogeny used in the analysis. We also noted whether authors reported confidence intervals, significance tests assessing difference of the lambda estimate from 0 or 1, and whether authors interpreted biological meaning from the magnitude of the estimated lambda. For studies that reported more than one lambda estimate, we also noted if the authors compared the lambdas against one another, and whether that was accompanied with an appropriate statistical test between the estimated lambda values. \hfill\break

We found 182 manuscripts from 2019 that estimated and reported Pagel's lambda values using PGLS methods. These papers averaged 8.527 lambda values, ranging from a single lambda estimate up to 71 estimated lambdas. Almost exactly half of the published lambda estimates were either below 0.05 (25.32%) or above 0.9 (24.74%; Figure 3). 73.32% of the published lambdas were estimated using phylogenies with fewer than 200 tips, and 348 lambda estimates (8.57% of all published estimates) came from phylogenies with fewer than 30 tips.  \hfill\break

[insert Figure 3 here] \hfill\break

Many of the reviewed manuscripts liberally interpreted the magnitude of the estimate lambda, using phrases such as "strong" or "weak" phylogentic signal when statistically, all that was clear was a difference between the estimated lambda and 0 or 1 respectively. We estimated that about 20.49% of the manuscripts revealed some sort of biological interpretation of the magnitude of estimated phylogenetic signal that overreached the statistical findings. We also identified seven manuscripts as having inappropriately interpreted differences in lambda values, indicating that some traits had stronger or weaker signal than other traits without the appropriate statistical tests. \hfill\break

As is evidenced by macroevolutionary papers published in 2019 papers, Pagel's lambda estimation methods are often misused and over-interpretted. Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so.

# Discussion

[This part is obviously not written yet] \hfill\break

General conclusions :Using the estimated lambda values from pgls are not useful. The questions of whether or not signal exists is appropriate, but inferring more from lambda *magnitude* is inappropriate. \hfill\break

More discussion paragraphs

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>

\newpage

# Figure Legends

\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips. \hfill\break

\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases. However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern. \hfill\break

\textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips. \hfill\break

\newpage
```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig2.png")

```

\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases. However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern. \hfill\break

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig3.png")

```

\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

### 2020-04-14 Manuscript/Manuscript-1-DCA.tex
paper vsn 1-DCA

--- 2020-04-08 Manuscript/Manuscript.tex
@@
-\documentclass[]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{setspace}\doublespacing
 \usepackage{lineno}\linenumbers
-% https://github.com/rstudio/rmarkdown/issues/337
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
 
-% https://github.com/rstudio/rmarkdown/pull/252
-\usepackage{titling}
-\setlength{\droptitle}{-2em}
-
-\pretitle{\vspace{\droptitle}\centering\huge}
-\posttitle{\par}
-
-\preauthor{\centering\large\emph}
-\postauthor{\par}
-
-\predate{\centering\large\emph}
-\postdate{\par}
-
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 
-\section{Pagel's Lambda Estimates are Often
-Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}
+\hypertarget{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}{%
+\section{Estimates of Phylogenetic Signal Based on Lambda are Often
+Inaccurate}\label{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}}
 
@@
 
-\section{Abstract}\label{abstract}
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
 
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Investigating macroevolutionary patterns requires a phylogenetic
-approach as species are non-independent by nature of their shared
-ancestry. Since the first appropriate method was introduced by
-Felsenstein (phylogenetic independent contrasts; Felsenstein (1985)),
-dozens of other methods have been developed and applied to increasingly
-complex questions in macroevolutionary biology (e.g. Grafen (1989);
-Harvey and Pagel (1991); Rohlf (2001); ({\textbf{???}})). A particularly
-useful measure in this field is phylogenetic signal, a quantification of
-a trait's correspondence with the clade's phylogenic history.
-Understanding the degree of phylogenetic signal present in a dataset is
-paramount and identifies the mode under which a trait has evolved; high
-measures of phylogenetic signal indicate a Brownian motion process,
-whereas lower levels of phylogenetic signal indicate natural selection
-or some other evolutionary force has influenced the traits evolutionary
-history. \hfill\break
+Investigating macroevolutionary patterns of trait variation requires a
+phylogenetic perspective, because the shared ancestry among species
+generates statistical non-independence (Felsenstein 1985; Harvey and
+Pagel 1991). Accounting for this evolutionary non-independence is the
+purview of \emph{phylogenetic comparative methods} (PCMs); a suite of
+analytical tools that condition the data on the phylogeny through the
+course of statisical evaluations of phenotypic trends (e.g., Grafen
+1989; Garland and Ives 2000; Rohlf 2001; Butler and King 2004). The past
+several decades have witnessed a rapid expansion in the development of
+PCMs to address an ever-growing set of macroevolutionary hypotheses
+(Martins and Hansen 1997; O'Meara et al. 2006; Revell and Harmon 2008;
+Beaulieu et al. 2012; Adams 2014b,a; Adams and Collyer 2018). These
+methods are predicated on the notion that phylogenetic signal -- the
+tendancy for closely related species to display similar trait values --
+is present in cross-species datasets (Abouheif 1999; Pagel 1999;
+Blomberg et al. 2003; Munkemuller et al. 2012). Indeed, under numerous
+evolutionary models, phylogenetic signal is to be expected, as
+stochastic character change along the hierarchical structure of the tree
+of life generates trait covaration among related taxa (see Felsenstein
+1985; Blomberg et al. 2003; Revell et al. 2008). \hfill\break
 
-Several approaches to quantify phylogenetic signal exist. For continuous
-data, the most common parameters used in the literature include Pagel's
-lambda ({\textbf{???}}) and Blomberg's kappa ({\textbf{???}}). Pagel's
-lambda has the advantage of being nested in the likelihood framework and
-thus has also been utilized to simultaneously estimate and account for
-phylogenetic signal while doing phylogenetic regressions or ANOVAs. At
-the development of these estimation methods, clear recommendations on
-the use and interpretation of this parameter were made based on data
-simulations and error evaluations (Revell 2010). However, the accuracy
-of the lambda estiamtion methods have not been fully evaluated, and
-consequently many studies have substantially misused these methods.
-\hfill\break
+In many macroevolutionary studies, it is often of interest to quantify
+the degree to which phylogenetic signal is displayed in continuous
+traits. Several analytical tools have been developed for this purpose
+(e.g., Gittleman and Kot 1990; Abouheif 1999; Pagel 1999; Blomberg et
+al. 2003; Klingenberg and Gidaszewski 2010; Adams 2014a), which differ
+primarily in how they characterize the phylogenetic dependency of trait
+variation among taxa. One commonly used statistical measure,
+\emph{Kappa} (\(K\)), expresses the strength of phylogenetic signal as
+the ratio of observed trait variation to the trait variation conditioned
+on the phylogeny; scaled by what is expected under Brownian motion given
+the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
+Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
+likelihood to fit the data to the phylogeny under some model of
+evolutionary change (typically Brownian motion). The inclusion of a
+scaling parameter, \(\lambda\), transforms the lengths of the internal
+branches of the phylogeny to improve the fit, and this parameter
+describes the degree of phylogenetic signal in the dataset (Pagel 1999;
+Freckleton et al. 2002). Pagel's \(\lambda\) also has the advantage that
+it may be included when estimating the association of traits in a
+phylogenetic context, meaning that one may account for the degree of
+phylogenetic signal while conducting phylogenetic regression or ANOVA
+(see Freckleton et al. 2002). \hfill\break
 
-An earlier study ({\textbf{???}}), breifly addressed this topic by
-showing how uninformative smaller phylogenies are when estimating
-various evolutionary parameters, such as Pagel's lambda. However, it
-remains unknown the degree to which lambda estimates accurately
-represent degree of phylogenetic signal, and thus we fully explicate the
-extent to which Pagel's lambda estimation methods can and have been
-misused in recent macroevolutionary studies. Here we take a
-comprehensive approach to demonstrate the scenarios under which
-estimated Pagel's lambdas accurately reflect known lambdas. We then
-demonstrate the effect of these, at times, dubious estimation methods on
-significance testing when used in a PGLS framework. Finally, we perform
-a meta-analysis of research published in 2019 employing Pagel's lambda
-estimation methods to demonstrate the uses and misuses of this method in
-the literature. This investigation provides evidence for the
-circumstances under which estimating Pagel's lambda can be informative,
-and which interpretations or uses are misleading with the hope of moving
-the field towards more appropriate employment of phylogenetic
-comparative methods.
+Several studies have investigated the statistical properties of methods
+for estimating phylogenetic signal under various conditions (e.g.,
+Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
+2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
+Revell 2010). These have largely focused on the ability of methods to
+detect the presence of phylogenetic signal (i.e., type I and type II
+error rates) under complex models of evolutionary change, across a range
+of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
+or unresolved topologies. In terms of parameter estimation, Revell
+(2010) confirmed that regression parameters were accurately estimated
+when \(\lambda\) was included during phylogenetic regression, and
+Munkmuller et al.~(2012) found that estimates of phylogenetic signal
+obtained using various measures generally increased when input levels of
+phylogenetic signal were stronger. However, the precision of those
+estimates could not be determined, because the input levels of
+phylogenetic signal were simulated via a scaling factor (\(w\)) that was
+not directly comparable to the measures of phylogenetic signal being
+compared (see Munkemuller et al. 2012). One study (Boettiger et al.
+2012) found that estimates of Pagel's \(\lambda\) displayed less
+variation when data were simulated on a large phylogeny (\(N=281\)) as
+compared to a small one (\(N=13\)), concluding that insufficient data
+(i.e., the number of species) was the underlying cause of the lack of
+precision in parameter estimation. However, this conclusion assumes that
+parameter estimation remains equally precise across its range of values
+(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+investigated. Thus, despite widespread use of Pagel's (1999) \(\lambda\)
+in macroevolutionary studies, at present, we still lack a general
+understanding of the precision with which \(\lambda\) can estimate
+levels of phylogenetic signal in phenotypic datasets. \hfill\break
 
-\section{Methods and Results}\label{methods-and-results}
+In this study, we evaluate the precision of Pagel's \(\lambda\) to
+estimate known levels of phylogenetic signal in phenotypic data. First
+we use computer simulations across differing numbers of species, and
+differing input levels of phylogenetic signal to explore when
+\(\lambda\) captures known levels of phylogenetic signal, and under what
+circumstances. Additionally, we use simulations to determine how the
+inclusion of \(\lambda\) in phylogenetic ANOVA and regression (i.e.,
+PGLS) affects parameter estimation, and whether estimates of
+phylogenetic signal in PGLS are accurate. We then survey the recent
+macroevolutionary literature for published papers containing estimates
+of \(\lambda\) from empirical datasets, and compare these empirical
+estimates to patterns gleaned from our computer simulations. In general
+we find that while PGLS parameters are accurately estimated with the
+inclusion of phylogenetic signal, estimates of \(\lambda\) are not.
+Thus, empirical studies should interpret the degree of phylogenetic
+signal in phenotypic data with caution when using this parameter.
+\textbf{something more? Let's wait for final data to decide.}
 
+\hypertarget{methods-and-results}{%
+\section{Methods and Results}\label{methods-and-results}}
+
+\hypertarget{simulated-trait}{%
 \subsection{\texorpdfstring{\emph{Simulated
-trait}}{Simulated trait}}\label{simulated-trait}
+trait}}{Simulated trait}}\label{simulated-trait}}
 
@@
 
+\hypertarget{simulated-anova-and-regressions}{%
 \subsection{\texorpdfstring{\emph{Simulated ANOVA and
-Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}
+Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}}
 
@@
 in the Supplemental Materials. All analyses were performed in R v3.6.0
-({\textbf{???}}) using the packages \texttt{geiger} ({\textbf{???}}) and
-\texttt{caper} ({\textbf{???}}), and the corresponding scripts can be
-found in the Supplemental Materials.
+(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
+2008) and \texttt{caper} (Orme et al. 2013), and the corresponding
+scripts can be found in the Supplemental Materials.
 
+\hypertarget{meta-analysis-of-empirical-results}{%
 \subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
-Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}
+Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}}
 
@@
 application, we performed a meta-analysis of studies published in 2019
-that cited Pagel's 1997 manuscript ({\textbf{???}}). The list of
-manuscripts was compiled through Google Scholar on Jan 23, 2020 and
-totaled 341 manuscripts. For each study, we extracted any published
-lambda estimates, along with the size of the phylogeny used in the
-analysis. We also noted whether authors reported confidence intervals,
-significance tests assessing difference of the lambda estimate from 0 or
-1, and whether authors interpreted biological meaning from the magnitude
-of the estimated lambda. For studies that reported more than one lambda
+that cited Pagel's 1997 manuscript (Pagel 1999). The list of manuscripts
+was compiled through Google Scholar on Jan 23, 2020 and totaled 341
+manuscripts. For each study, we extracted any published lambda
+estimates, along with the size of the phylogeny used in the analysis. We
+also noted whether authors reported confidence intervals, significance
+tests assessing difference of the lambda estimate from 0 or 1, and
+whether authors interpreted biological meaning from the magnitude of the
+estimated lambda. For studies that reported more than one lambda
 estimate, we also noted if the authors compared the lambdas against one
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Felsenstein1985}{}
+\leavevmode\hypertarget{ref-Abouheif1999}{}%
+Abouheif, E. 1999. A method for testing the assumption of phylogenetic
+independence in comparative data. Evolutionary Ecology Research
+1:895--909.
+
+\leavevmode\hypertarget{ref-Adams2014a}{}%
+Adams, D. C. 2014a. A generalized Kappa statistic for estimating
+phylogenetic signal from shape and other high-dimensional dultivariate
+data. Systematic Biology 63:685--697.
+
+\leavevmode\hypertarget{ref-Adams2014b}{}%
+Adams, D. C. 2014b. A method for assessing phylogenetic least squares
+models for shape and other high-dimensional multivariate data. Evolution
+68:2675--2688.
+
+\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
+Adams, D. C., and M. L. Collyer. 2018. Phylogenetic anova: Group-clade
+aggregation, biological challenges, and a refined permutation procedure.
+Evolution 72:1204--1215.
+
+\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
+Beaulieu, J. M., D. C. Jhwueng, C. Boettiger, and B. C. O'Meara. 2012.
+Modeling stabilizing selection: Expanding the ornstein-uhlenbeck model
+of adaptive evolution. Evolution 66:2369--2383.
+
+\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
+Blomberg, S. P., T. Garland, and A. R. Ives. 2003. Testing for
+phylogenetic signal in comparative data: Behavioral traits are more
+labile. Evolution 57:717--745.
+
+\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
+Boettiger, C., G. Coop, and P. Ralph. 2012. Is your phylogeny
+informative? Measuring the power of comparative methods. Evolution
+67:2240--2251.
+
+\leavevmode\hypertarget{ref-ButlerKing2004}{}%
+Butler, M. A., and A. A. King. 2004. Phylogenetic comparative analysis:
+A modeling approach for adaptive evolution. American Naturalist
+164:683--695.
+
+\leavevmode\hypertarget{ref-DinizFilho2012}{}%
+Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
+comparison of metrics for estimating phylogenetic signal under
+alternative evolutionary models. Genetics and molecular biology
+35:673--679. SciELO Brasil.
+
+\leavevmode\hypertarget{ref-Felsenstein1985}{}%
 Felsenstein, J. 1985. Phylogenies and the comparative method. American
@@
 
-\hypertarget{ref-Grafen1989}{}
+\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
+Freckleton, R. P., P. H. Harvey, and M. Pagel. 2002. Phylogenetic
+analysis and comparative data: A test and review of evidence. American
+Naturalist 160:712--726.
+
+\leavevmode\hypertarget{ref-GarlandIves2000}{}%
+Garland, T. J., and A. R. Ives. 2000. Using the past to predict the
+present: Confidence intervals for regression equations in phylogenetic
+comparative methods. American Naturalist 155:346--364.
+
+\leavevmode\hypertarget{ref-Gittleman1990}{}%
+Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
+model for estimating phylogenetic effects. Systematic Zoology
+39:227--241. Society of Systematic Zoology.
+
+\leavevmode\hypertarget{ref-Grafen1989}{}%
 Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
@@
 
-\hypertarget{ref-HarveyPagel1991}{}
+\leavevmode\hypertarget{ref-Harmon2008}{}%
+Harmon, L. J., J. T. Weir, C. D. Brock, R. E. Glor, and W. Challenger.
+2008. GEIGER: Investigating evolutionary radiations. Bioinformatics
+24:129--131.
+
+\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
 Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
@@
 
-\hypertarget{ref-Revell2010}{}
+\leavevmode\hypertarget{ref-Klingenberg2010}{}%
+Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
+phylogenetic signals and homoplasy in morphometric data. Systematic
+biology 59:245--261. Oxford University Press.
+
+\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
+Martins, E. P., and T. F. Hansen. 1997. Phylogenies and the comparative
+method: A general approach to incorporating phylogenetic information
+into the analysis of interspecific data. American Naturalist
+149:646--667.
+
+\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
+Molina-Venegas, R., and M. A. Rodriguez. 2017. Revisiting phylogenetic
+signal; strong or negligible impacts of polytomies and branch length
+information? BMC evolutionary biology 17:53. BioMed Central.
+
+\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
+Munkemuller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
+Schiffers, and W. Thuiller. 2012. How to measure and test phylogenetic
+signal. Methods in Ecology and Evolution 3:743--756.
+
+\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
+O'Meara, B. C., C. Ane, M. J. Sanderson, and P. C. Wainwright. 2006.
+Testing for different rates of continuous trait evolution using
+likelihood. Evolution 60:922--933.
+
+\leavevmode\hypertarget{ref-Orme2013}{}%
+Orme, D., R. P. Freckleton, G. H. Thomas, T. Petzoldt, S. A. Fritz, and
+N. Isaac. 2013. CAPER: Comparative analyses of phylogenetics and
+evolution in r. Methods in Ecology and Evolution 3:145--151.
+
+\leavevmode\hypertarget{ref-Pagel1999}{}%
+Pagel, M. D. 1999. Inferring the historical patterns of biological
+evolution. Nature 401:877--884.
+
+\leavevmode\hypertarget{ref-Pavoine2012}{}%
+Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
+biological traits: The ubiquity of cross-product statistics. Evolution:
+International Journal of Organic Evolution 67:828--840. Wiley Online
+Library.
+
+\leavevmode\hypertarget{ref-RCT}{}%
+R Core Team. 2019. R: A language and environment for statistical
+computing. R Foundation for Statistical Computing, Vienna, Austria.
+
+\leavevmode\hypertarget{ref-Revell2010}{}%
 Revell, L. J. 2010. Phylogenetic signal and linear regression on species
@@
 
-\hypertarget{ref-Rohlf2001}{}
+\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
+Revell, L. J., and L. J. Harmon. 2008. Testing quantitative genetic
+hypotheses about the evolutionary rate matrix for continuous characters.
+Evolutionary Ecology Research 10:311--331.
+
+\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
+Revell, L. J., L. J. Harmon, and D. C. Collar. 2008. Phylogenetic
+signal, evolutionary process, and rate. Systematic Biology 57:591--601.
+
+\leavevmode\hypertarget{ref-Rohlf2001}{}%
 Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
@@
 
-\section{Figure Legends}\label{figure-legends}
+\hypertarget{figure-legends}{%
+\section{Figure Legends}\label{figure-legends}}
 

### 2020-04-16 Manuscript/Manuscript-1-DCA.Rmd
vsn 1-dca

@@
 The past several decades have witnessed a rapid expansion in the development of PCMs to address an ever-growing set of macroevolutionary hypotheses [@MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @Adams2014a; @AdamsCollyer2018b].
-These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Munkemuller_et_al2012].
+These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
 Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [see @Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
@@
 
-In many macroevolutionary studies, it is often of interest to quantify the degree to which phylogenetic signal is displayed in continuous traits.
+For many macroevolutionary analyses, it is often of interest to quantify the degree to which phylogenetic signal is displayed in continuous traits.
 Several analytical tools have been developed for this purpose [e.g., @Gittleman1990; @Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Klingenberg2010; @Adams2014a], which differ primarily in how they characterize the phylogenetic dependency of trait variation among taxa.
@@
 These have largely focused on the ability of methods to detect the presence of phylogenetic signal (i.e., type I and type II error rates) under complex models of evolutionary change, across a range of phylogeny sizes, and with varying degrees of phylogenetic uncertainty or unresolved topologies.
-In terms of parameter estimation, Revell [-@Revell2010] confirmed that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al.
-[-@Munkemuller_et_al2012] found that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger.
+In terms of parameter estimation, Revell [-@Revell2010] found that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al.
+[-@Munkemuller_et_al2012] demonstrated that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger.
 However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012].
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation.
-However, this conclusion assumes that parameter estimation remains equally precise across its range of values ($\lambda = 0 \to 1$); an assumption that to date, has not been investigated.
+However, this conclusion assumes that the precision of parameter estimation remains constant across the range of values ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
 Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
@@
   In this study, we evaluate the precision of Pagel's $\lambda$ to estimate known levels of phylogenetic signal in phenotypic data.
-First we use computer simulations across differing numbers of species, and differing input levels of phylogenetic signal to explore when $\lambda$ captures known levels of phylogenetic signal, and under what circumstances.
-Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic ANOVA and regression (i.e., PGLS) affects parameter estimation, and whether estimates of phylogenetic signal in PGLS are accurate.
+First we use computer simulations across differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
+For comparison, we also evaluate estimates of $Kappa$ and its empirically-derived effect size ($Z_K$) for this purpose.
+Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic regression and ANOVA (i.e., PGLS) affects parameter estimation, and whether levels of phylogenetic signal estimated in a PGLS framework are accurate.
 We then survey the recent macroevolutionary literature for published papers containing estimates of $\lambda$ from empirical datasets, and compare these empirical estimates to patterns gleaned from our computer simulations.
-In general we find that while PGLS parameters are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
-Thus, empirical studies should interpret the degree of phylogenetic signal in phenotypic data with caution when using this parameter.
-**something more?
-Let's wait for final data to decide.**
+In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
+Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, while effect sizes ($Z_K$) obtained from $Kappa$ display far less variation, and increase in a steady (**uniform?**) manner with increasing phylogenetic signal.
+We conclude that Pagel's $\lambda$ should not be used for characterizing and comparing the degree of phylogenetic signal in phenotypic datasets, while comparisons of effect sizes obtained from $Kappa$ hold considerable promise in this regard.
 
-# Methods and Results
+# Materials and Methods
 
-## *Simulated trait*
+## *Simulations*
 
-To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies of variable size, ranging in tip number from 32 to 1024.
-We then scaled the simulated phylogenies by lambda values ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size) with which we generated trait data with known lambdas by simulating a continuous variable on each scaled phylogeny under Brownian motion.
-We then estimated lambda values from these data using phylogenetic generalized least squares (PGLS)  to compare against the known lambda values.
+We conducted a series of computer simulations to evaluate the ability of Pagel's [-@Pagel1999] $\lambda$ to estimate known levels of phylogenetic signal.
+Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine the extent to which tree shape affected our findings (see Supplemental Material).
+Our first set of simulations evaluated the extent to which values of $\lambda$ estimated from the data corresponded with actual input levels of phylogenetic signal.
+For these simulations we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$).
+Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda$.
+We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 - 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
+Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with known and differing levels of phylogenetic signal.
+These varied from datasets with no phylogenetic signal (when $\lambda_{in} =0$) to datsets with phylogenetic signal corresponding to what was expected under Brownian motion (when $\lambda_{in} =1$).
 \hfill\break
 
-To visualize the accuracy of the lambda estimation methods, we first plotted known lambdas (input lambdas) against the estimated lambdas (Figure 1).
-This plot demonstrates the rampant inaccuracy of estimating lambdas on phylogenies with fewer than 200 tips, as the spread of data in the upper panels in Figure 1 is remarkably wide.
-We also see that the widest spread of estimated lambdas is observed near the center of each plot corresponding to intermediate values of known lambda.
-Lastly, we see a slight tendency for the PGLS estimation methods to underestimate lambda, especially for known lambda values below 0.5 as can be seen by the numerous data points along the x axis for the smaller phylogenies analyzed (n tips < 200).
+Using the simulated phylogenies we estimated the phylogenetic signal in each dataset using Pagel's $\lambda$.
+Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression.
+Additionally, the precision of $\lambda_{est}$ was approximated by observing both the range and the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
+Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
+Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019].
+Because $Kappa$ and $Z_K$ differ in scale from $\lambda$, both measures were linearly normalized to a standard uniform distribution ($0\rightarrow1$), and the range and variation in the set of normalized $Kappa$ and $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were estimated for comparison.
 \hfill\break
 
-[insert Figure 1 here]
+Our second set of simulations evaluated the extent to which values of $\lambda$ estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$) corresponded with actual input levels of phylogenetic signal in the response variable.
+As before we generated 50 pure-birth phylogenies at each of six levels of species richness ($n=2^5 - 2^{10}$), and rescaled each with a set of input $\lambda$ values ($\lambda_{in} = 0.0 - 1.0$; in 21 intervals of 0.05 units).
+Next, an independent variable $X$ was simulated on each phylogeny under a Brownian motion model of evolution (for PGLS regression).
+For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny.
+Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal.
+This was accomplished as: $Y=\beta{X}+\epsilon$.
+Here, the association between $Y$ and $X$  was modeled using a range of values ($\beta=(0,0.25, 0.5, 0.75,1)$), and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution [$\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: see @Revell2010 for a similar simulation design].
+For each dataset, the fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained and evaluated as above.
+All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
+R-scripts are found in the Supplemental Material.
 
-## *Simulated ANOVA and Regressions*
+## *Literature Survey*
 
-To measure the statistical performance of PGLS lambda estimation methods when applied to ANOVA and regression analyses, we used the above generated data (independent variables) to simulate a second set of trait variables (depedendent variables) across the range of correlation strengths (betas ranging from 0 to 1 at intervales of 0.25).
-We then used PGLS to estimate phylogenetic signal of the dependent variable and the slope coefficient for the regression between the dependent and independent variables.
-We also calculated F- and p-values from the regression analyses.
-Finally, we again fit the dependent variable to the independent variable while holding the lambda value at 0 to approximate an ordinary least squares (OLS) apprach.
+To determine how Pagel's $\lambda$ is commonly utilized in empirical studies, we conducted a literature survey.
+From Google.scholar we obtained a list of all papers published in 2019 that used $\lambda$; resulting in 341 studies.
+For each study, we extracted all $\lambda_{est}$, the size of the phylogeny ($n$), and noted whether authors reported confidence intervals or performed significance tests assessing difference of $\lambda_{est}$ from either zero or one.
+We also noted whether biological interpretations based on $\lambda_{est}$ were made, and for studies that reported more than one $\lambda_{est}$, we also noted whether these were compared in some manner, and whether such comparisons were accompanied with statistical tests between $\lambda_{est}$.
 \hfill\break
 
-We compared the estimated slope coefficients across variable input beta values to assess the ability of PGLS to identify significant correlations (Figure 2).
-All distribution groups center around the 1:1 relationship between input beta values and estimated slopes.
-However the variance around that mean is substantial, with estimates ranging from approximately -2 to 4 in datasets with 32 taxa and a known relationship of (beta = 1).
-This variance of the estimated slopes only becomes reasonable when phylogenies have over 200 tips, similarly to what we saw in the first simulation analysis in this study.
+# Results
+
+## *Simulations*
+
+lambda: (Fig 1) not good predictor of input physignal (regression), particularly at low sample sizes.
+Also, similar to early results of Boettiger (but greatly expanded) range and variation of lambda very large at low N, and generally decreases with larger N.
+However, the precision was not uniform across the range of lambda-in.
+Variation in lambda-est was greatest at intermediate levels of physig, and descreased at lower and higher values.
+Thus, the overall dispersion across the range of input values resembled an asymmetric lens (**NAME?**), with less variation at lambda-in close to 1, greatest variation in middle, and less (but substantial) variation near lambda-in of zero.
+Notably,  even at very large species richness (n=1024) the range of lam-est was quite large: e.g., l-in 0.5 has range of 45% of all possible values.
+Thus, lambda-est is not very precise in terms of estimating the true l-in parameter, meaning that using lambda to biologically interpret the 'strength' of phylogenetic signal in a dataset, or compare between datasets, is fraught with difficulty.
 \hfill\break
+
+[insert Figure 1 here] \hfill\break
+
+By contrast,  (FIg 2) Kappa.
+Much smaller range and variation as compared with lambda, but with larger l-in, variation in kappa does increase.
+However, this is not the case with Z-K.
+Z-K increases predictably (linearly, progressively...) with L-in, and importantly variation in Z-K remains nearly constant across the range of input values.
+this means that Z-K is equally precise in estimating the strength of phylogenetic signal, and may be more adept at facilitating useful statements on teh relative strength of phylogenetic signal within and across datasets.
+\hfill\break
+
+**NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K, Z-k-normalized.**
 
@@
 
-Surprisingly, the estimated lambda values of the dependent variable do not correspond with input lambda values that characterize the independent variable (Figure S1).
-This disconnect then resulted in no appreciable difference of slope estimates across input lambda values (Figure S2), nor do we see substantial differences in slope estimates between the PGLS and OLS analyses (results not shown).
-Thus, we show that phylogenetic signal present in one variable does not translate to phylogenetic signal in a second, even highly correlated, dependent variable.
-\hfill\break
-
-Scaling and trait generation procedures for all simulation methods were repeated with symmetrical and ladder phylogenies of variable size.
-Results generated using these variable phylogeny shapes were consistent with the pure-birth phylogeny results presented above and can be found in the Supplemental Materials.
-All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008] and `caper` [@Orme2013], and the corresponding scripts can be found in the Supplemental Materials.
-
-## *Meta-Analysis of Empirical Results*
-
-To understand the extent of this problematic estimation method in application, we performed a meta-analysis of studies published in 2019 that cited Pagel's 1997 manuscript [@Pagel1999].
-The list of manuscripts was compiled through Google Scholar on Jan 23, 2020 and totaled 341 manuscripts.
-For each study, we extracted any published lambda estimates, along with the size of the phylogeny used in the analysis.
-We also noted whether authors reported confidence intervals, significance tests assessing difference of the lambda estimate from 0 or 1, and whether authors interpreted biological meaning from the magnitude of the estimated lambda.
-For studies that reported more than one lambda estimate, we also noted if the authors compared the lambdas against one another, and whether that was accompanied with an appropriate statistical test between the estimated lambda values.
-\hfill\break
+## *Literature Survey*
 
@@
 
+# Results
+
 # Discussion
 
-[This part is obviously not written yet] \hfill\break
+1: summary paragraph
 
-General conclusions :Using the estimated lambda values from pgls are not useful.
-The questions of whether or not signal exists is appropriate, but inferring more from lambda *magnitude* is inappropriate.
+2: expand on Lambda..
+lambda innacurate, not precise, level of precision varies with input physig (worse in mid-range). Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
+
+3: By contrast, effect size Z-K useful, equally precise across range of values.
+Can be used to characterize the strength of physignal, and because robust to input levels, etc.
+may be used to compare across datasets.
+
+Somewhere, recognize that this is somewhat 'backwards' from prior recommendations where Kappa had somewhat lower performance in terms of type I and type II error (which??
+I forget).
+However, recall that those studies did not examine the precision of the estimates.
+Nor was Z-k included, because it was not yet invented.
+So Use of Z-k should make good sense here.
+
+Closing paragraph.
+
 \hfill\break

### 2020-04-16 Manuscript/Manuscript-1-DCA.tex
vsn 1-dca

@@
 tendancy for closely related species to display similar trait values --
-is present in cross-species datasets (Abouheif 1999; Pagel 1999;
-Blomberg et al. 2003; Munkemuller et al. 2012). Indeed, under numerous
-evolutionary models, phylogenetic signal is to be expected, as
-stochastic character change along the hierarchical structure of the tree
-of life generates trait covaration among related taxa (see Felsenstein
-1985; Blomberg et al. 2003; Revell et al. 2008). \hfill\break
+is present in cross-species datasets (Felsenstein 1985; Pagel 1999;
+Blomberg et al. 2003). Indeed, under numerous evolutionary models,
+phylogenetic signal is to be expected, as stochastic character change
+along the hierarchical structure of the tree of life generates trait
+covaration among related taxa (see Felsenstein 1985; Blomberg et al.
+2003; Revell et al. 2008). \hfill\break
 
-In many macroevolutionary studies, it is often of interest to quantify
+For many macroevolutionary analyses, it is often of interest to quantify
 the degree to which phylogenetic signal is displayed in continuous
@@
 or unresolved topologies. In terms of parameter estimation, Revell
-(2010) confirmed that regression parameters were accurately estimated
-when \(\lambda\) was included during phylogenetic regression, and
-Munkmuller et al.~(2012) found that estimates of phylogenetic signal
+(2010) found that regression parameters were accurately estimated when
+\(\lambda\) was included during phylogenetic regression, and Munkmuller
+et al.~(2012) demonstrated that estimates of phylogenetic signal
 obtained using various measures generally increased when input levels of
@@
 precision in parameter estimation. However, this conclusion assumes that
-parameter estimation remains equally precise across its range of values
-(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
-investigated. Thus, despite widespread use of Pagel's (1999) \(\lambda\)
-in macroevolutionary studies, at present, we still lack a general
-understanding of the precision with which \(\lambda\) can estimate
-levels of phylogenetic signal in phenotypic datasets. \hfill\break
+the precision of parameter estimation remains constant across the range
+of values (\(\lambda = 0 \to 1\)); an assumption that to date, has not
+been verified. Thus, despite widespread use of Pagel's (1999)
+\(\lambda\) in macroevolutionary studies, at present, we still lack a
+general understanding of the precision with which \(\lambda\) can
+estimate levels of phylogenetic signal in phenotypic datasets.
+\hfill\break
 
@@
 estimate known levels of phylogenetic signal in phenotypic data. First
-we use computer simulations across differing numbers of species, and
-differing input levels of phylogenetic signal to explore when
-\(\lambda\) captures known levels of phylogenetic signal, and under what
-circumstances. Additionally, we use simulations to determine how the
-inclusion of \(\lambda\) in phylogenetic ANOVA and regression (i.e.,
-PGLS) affects parameter estimation, and whether estimates of
-phylogenetic signal in PGLS are accurate. We then survey the recent
+we use computer simulations across differing numbers of species,
+differently shaped phylogenies, and differing input levels of
+phylogenetic signal, to explore the degree to which \(\lambda\)
+correctly identifies known levels of phylogenetic signal, and under what
+circumstances. For comparison, we also evaluate estimates of \(Kappa\)
+and its empirically-derived effect size (\(Z_K\)) for this purpose.
+Additionally, we use simulations to determine how the inclusion of
+\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
+parameter estimation, and whether levels of phylogenetic signal
+estimated in a PGLS framework are accurate. We then survey the recent
 macroevolutionary literature for published papers containing estimates
@@
 estimates to patterns gleaned from our computer simulations. In general
-we find that while PGLS parameters are accurately estimated with the
-inclusion of phylogenetic signal, estimates of \(\lambda\) are not.
-Thus, empirical studies should interpret the degree of phylogenetic
-signal in phenotypic data with caution when using this parameter.
-\textbf{something more? Let's wait for final data to decide.}
+we find that while PGLS parameters (e.g., \(\beta\)) are accurately
+estimated with the inclusion of phylogenetic signal, estimates of
+\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
+widely for a given input value of phylogenetic signal, while effect
+sizes (\(Z_K\)) obtained from \(Kappa\) display far less variation, and
+increase in a steady (\textbf{uniform?}) manner with increasing
+phylogenetic signal. We conclude that Pagel's \(\lambda\) should not be
+used for characterizing and comparing the degree of phylogenetic signal
+in phenotypic datasets, while comparisons of effect sizes obtained from
+\(Kappa\) hold considerable promise in this regard.
 
-\hypertarget{methods-and-results}{%
-\section{Methods and Results}\label{methods-and-results}}
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
-\hypertarget{simulated-trait}{%
-\subsection{\texorpdfstring{\emph{Simulated
-trait}}{Simulated trait}}\label{simulated-trait}}
+\hypertarget{simulations}{%
+\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}
 
-To assess the accuracy of Pagel's lambda estimations, we simulated
-pure-birth phylogenies of variable size, ranging in tip number from 32
-to 1024. We then scaled the simulated phylogenies by lambda values
-ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree
-size) with which we generated trait data with known lambdas by
-simulating a continuous variable on each scaled phylogeny under Brownian
-motion. We then estimated lambda values from these data using
-phylogenetic generalized least squares (PGLS) to compare against the
-known lambda values. \hfill\break
+We conducted a series of computer simulations to evaluate the ability of
+Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
+signal. Our primary simulations were based on pure-birth phylogenies;
+however, we also evaluated patterns on both balanced and pectinate trees
+to determine the extent to which tree shape affected our findings (see
+Supplemental Material). Our first set of simulations evaluated the
+extent to which values of \(\lambda\) estimated from the data
+corresponded with actual input levels of phylogenetic signal. For these
+simulations we generated 50 pure-birth phylogenies at each of six
+different tree sizes, ranging from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)).
+Next, we rescaled the simulated phylogenies by multiplying the internal
+branches by \(\lambda\). We used a set of values that encompassed the
+entire range of \(\lambda\): (\(\lambda_{in} = 0.0 - 1.0\); in 21
+intervals of 0.05 units), resulting in 1050 scaled phylogenies at each
+level of species richness (\(n\)). Continuous traits were then simulated
+on each phylogeny under a Brownian motion model of evolution to obtain
+datasets with known and differing levels of phylogenetic signal. These
+varied from datasets with no phylogenetic signal (when
+\(\lambda_{in} =0\)) to datsets with phylogenetic signal corresponding
+to what was expected under Brownian motion (when \(\lambda_{in} =1\)).
+\hfill\break
 
-To visualize the accuracy of the lambda estimation methods, we first
-plotted known lambdas (input lambdas) against the estimated lambdas
-(Figure 1). This plot demonstrates the rampant inaccuracy of estimating
-lambdas on phylogenies with fewer than 200 tips, as the spread of data
-in the upper panels in Figure 1 is remarkably wide. We also see that the
-widest spread of estimated lambdas is observed near the center of each
-plot corresponding to intermediate values of known lambda. Lastly, we
-see a slight tendency for the PGLS estimation methods to underestimate
-lambda, especially for known lambda values below 0.5 as can be seen by
-the numerous data points along the x axis for the smaller phylogenies
-analyzed (n tips \textless{} 200). \hfill\break
+Using the simulated phylogenies we estimated the phylogenetic signal in
+each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
+each level of species richness (\(n\)), we assessed the relationship
+between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
+Additionally, the precision of \(\lambda_{est}\) was approximated by
+observing both the range and the variation of \(\lambda_{est}\) obtained
+at each input level of phylogenetic signal (\(\lambda_{in}\)). Finally,
+for comparison we estimated the degree of phylogenetic signal in each
+dataset using \(Kappa\) (Blomberg et al. 2003; Adams 2014a), and
+calculated its effect size (\(Z_K\): sensu Adams and Collyer 2016,
+2019a) for each dataset. Here, the effect size was estimated as:
+\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
+observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
+standard deviation of the empirical sampling distribution of values
+obtained from the permutation distribution (for discussion see: Adams
+and Collyer 2019b). Because \(Kappa\) and \(Z_K\) differ in scale from
+\(\lambda\), both measures were linearly normalized to a standard
+uniform distribution (\(0\rightarrow1\)), and the range and variation in
+the set of normalized \(Kappa\) and \(Z_K\) values at each input level
+of phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
+\hfill\break
 
-{[}insert Figure 1 here{]}
+Our second set of simulations evaluated the extent to which values of
+\(\lambda\) estimated in PGLS regression and ANOVA (i.e., \(Y\sim{X}\))
+corresponded with actual input levels of phylogenetic signal in the
+response variable. As before we generated 50 pure-birth phylogenies at
+each of six levels of species richness (\(n=2^5 - 2^{10}\)), and
+rescaled each with a set of input \(\lambda\) values
+(\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units). Next, an
+independent variable \(X\) was simulated on each phylogeny under a
+Brownian motion model of evolution (for PGLS regression). For
+phylogenetic ANOVA, random groups (\(X\)) were obtained by simulating a
+discrete (binary) character on each phylogeny. Next, the dependent
+variable was simulated in such a manner as to contain a known
+relationship with \(X\) plus random error containing phylogenetic
+signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
+association between \(Y\) and \(X\) was modeled using a range of values
+(\(\beta=(0,0.25, 0.5, 0.75,1)\)), and the residual error was modeled to
+contain phylogenetic signal simulated under a Brownian motion model of
+evolution (\(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): see Revell
+2010 for a similar simulation design). For each dataset, the fit of the
+phylogenetic regression was estimated using maximum likelihood, and
+parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
+obtained and evaluated as above. All analyses were performed in R v3.6.0
+(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
+2008), \texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell
+2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
+al. 2020). R-scripts are found in the Supplemental Material.
 
-\hypertarget{simulated-anova-and-regressions}{%
-\subsection{\texorpdfstring{\emph{Simulated ANOVA and
-Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}}
+\hypertarget{literature-survey}{%
+\subsection{\texorpdfstring{\emph{Literature
+Survey}}{Literature Survey}}\label{literature-survey}}
 
-To measure the statistical performance of PGLS lambda estimation methods
-when applied to ANOVA and regression analyses, we used the above
-generated data (independent variables) to simulate a second set of trait
-variables (depedendent variables) across the range of correlation
-strengths (betas ranging from 0 to 1 at intervales of 0.25). We then
-used PGLS to estimate phylogenetic signal of the dependent variable and
-the slope coefficient for the regression between the dependent and
-independent variables. We also calculated F- and p-values from the
-regression analyses. Finally, we again fit the dependent variable to the
-independent variable while holding the lambda value at 0 to approximate
-an ordinary least squares (OLS) apprach. \hfill\break
+To determine how Pagel's \(\lambda\) is commonly utilized in empirical
+studies, we conducted a literature survey. From Google.scholar we
+obtained a list of all papers published in 2019 that used \(\lambda\);
+resulting in 341 studies. For each study, we extracted all
+\(\lambda_{est}\), the size of the phylogeny (\(n\)), and noted whether
+authors reported confidence intervals or performed significance tests
+assessing difference of \(\lambda_{est}\) from either zero or one. We
+also noted whether biological interpretations based on \(\lambda_{est}\)
+were made, and for studies that reported more than one
+\(\lambda_{est}\), we also noted whether these were compared in some
+manner, and whether such comparisons were accompanied with statistical
+tests between \(\lambda_{est}\). \hfill\break
 
-We compared the estimated slope coefficients across variable input beta
-values to assess the ability of PGLS to identify significant
-correlations (Figure 2). All distribution groups center around the 1:1
-relationship between input beta values and estimated slopes. However the
-variance around that mean is substantial, with estimates ranging from
-approximately -2 to 4 in datasets with 32 taxa and a known relationship
-of (beta = 1). This variance of the estimated slopes only becomes
-reasonable when phylogenies have over 200 tips, similarly to what we saw
-in the first simulation analysis in this study. \hfill\break
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{simulations-1}{%
+\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations-1}}
+
+lambda: (Fig 1) not good predictor of input physignal (regression),
+particularly at low sample sizes. Also, similar to early results of
+Boettiger (but greatly expanded) range and variation of lambda very
+large at low N, and generally decreases with larger N. However, the
+precision was not uniform across the range of lambda-in. Variation in
+lambda-est was greatest at intermediate levels of physig, and descreased
+at lower and higher values. Thus, the overall dispersion across the
+range of input values resembled an asymmetric lens (\textbf{NAME?}),
+with less variation at lambda-in close to 1, greatest variation in
+middle, and less (but substantial) variation near lambda-in of zero.
+Notably, even at very large species richness (n=1024) the range of
+lam-est was quite large: e.g., l-in 0.5 has range of 45\% of all
+possible values. Thus, lambda-est is not very precise in terms of
+estimating the true l-in parameter, meaning that using lambda to
+biologically interpret the `strength' of phylogenetic signal in a
+dataset, or compare between datasets, is fraught with difficulty.
+\hfill\break
+
+{[}insert Figure 1 here{]} \hfill\break
+
+By contrast, (FIg 2) Kappa. Much smaller range and variation as compared
+with lambda, but with larger l-in, variation in kappa does increase.
+However, this is not the case with Z-K. Z-K increases predictably
+(linearly, progressively\ldots) with L-in, and importantly variation in
+Z-K remains nearly constant across the range of input values. this means
+that Z-K is equally precise in estimating the strength of phylogenetic
+signal, and may be more adept at facilitating useful statements on teh
+relative strength of phylogenetic signal within and across datasets.
+\hfill\break
+
+\textbf{NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K,
+Z-k-normalized.}
 
@@
 
-Surprisingly, the estimated lambda values of the dependent variable do
-not correspond with input lambda values that characterize the
-independent variable (Figure S1). This disconnect then resulted in no
-appreciable difference of slope estimates across input lambda values
-(Figure S2), nor do we see substantial differences in slope estimates
-between the PGLS and OLS analyses (results not shown). Thus, we show
-that phylogenetic signal present in one variable does not translate to
-phylogenetic signal in a second, even highly correlated, dependent
-variable. \hfill\break
-
-Scaling and trait generation procedures for all simulation methods were
-repeated with symmetrical and ladder phylogenies of variable size.
-Results generated using these variable phylogeny shapes were consistent
-with the pure-birth phylogeny results presented above and can be found
-in the Supplemental Materials. All analyses were performed in R v3.6.0
-(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
-2008) and \texttt{caper} (Orme et al. 2013), and the corresponding
-scripts can be found in the Supplemental Materials.
-
-\hypertarget{meta-analysis-of-empirical-results}{%
-\subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
-Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}}
-
-To understand the extent of this problematic estimation method in
-application, we performed a meta-analysis of studies published in 2019
-that cited Pagel's 1997 manuscript (Pagel 1999). The list of manuscripts
-was compiled through Google Scholar on Jan 23, 2020 and totaled 341
-manuscripts. For each study, we extracted any published lambda
-estimates, along with the size of the phylogeny used in the analysis. We
-also noted whether authors reported confidence intervals, significance
-tests assessing difference of the lambda estimate from 0 or 1, and
-whether authors interpreted biological meaning from the magnitude of the
-estimated lambda. For studies that reported more than one lambda
-estimate, we also noted if the authors compared the lambdas against one
-another, and whether that was accompanied with an appropriate
-statistical test between the estimated lambda values. \hfill\break
+\hypertarget{literature-survey-1}{%
+\subsection{\texorpdfstring{\emph{Literature
+Survey}}{Literature Survey}}\label{literature-survey-1}}
 
@@
 
+\hypertarget{results-1}{%
+\section{Results}\label{results-1}}
+
 \hypertarget{discussion}{%
@@
 
-{[}This part is obviously not written yet{]} \hfill\break
+1: summary paragraph
 
-General conclusions :Using the estimated lambda values from pgls are not
-useful. The questions of whether or not signal exists is appropriate,
-but inferring more from lambda \emph{magnitude} is inappropriate.
+2: expand on Lambda.. lambda innacurate, not precise, level of precision
+varies with input physig (worse in mid-range). Patterns worse with PGLS,
+though beta still estimated properly. Conclusion, lambda not overly
+useful.
+
+3: By contrast, effect size Z-K useful, equally precise across range of
+values. Can be used to characterize the strength of physignal, and
+because robust to input levels, etc. may be used to compare across
+datasets.
+
+Somewhere, recognize that this is somewhat `backwards' from prior
+recommendations where Kappa had somewhat lower performance in terms of
+type I and type II error (which?? I forget). However, recall that those
+studies did not examine the precision of the estimates. Nor was Z-k
+included, because it was not yet invented. So Use of Z-k should make
+good sense here.
+
+Closing paragraph.
+
 \hfill\break
@@
 
+\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
+Adams, D. C., and M. L. Collyer. 2019a. Comparing the strength of
+modular signal, and evaluating alternative modular hypotheses, using
+covariance ratio effect sizes with morphometric data. Evolution
+73:2352--2367.
+
+\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
+Adams, D. C., and M. L. Collyer. 2016. On the comparison of the strength
+of morphological integration across morphometric datasets. Evolution
+70:2623--2631.
+
 \leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
@@
 Evolution 72:1204--1215.
+
+\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
+Adams, D. C., and M. L. Collyer. 2019b. Phylogenetic comparative methods
+and the evolution of multivariate phenotypes. Annual Review of Ecology,
+Evolution, and Systematics 50:405--425.
+
+\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
+Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
+Software for geometric morphometric analyses. R package version 3.2.1.
+
+\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
+Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
+the collection and analysis of geometric morphometric shape data.
+Methods in Ecology and Evolution 4:393--399.
 
@@
 comparison of metrics for estimating phylogenetic signal under
-alternative evolutionary models. Genetics and molecular biology
-35:673--679. SciELO Brasil.
+alternative evolutionary models. Genetics and Molecular Biology
+35:673--679.
 
@@
 model for estimating phylogenetic effects. Systematic Zoology
-39:227--241. Society of Systematic Zoology.
+39:227--241.
 
@@
 phylogenetic signals and homoplasy in morphometric data. Systematic
-biology 59:245--261. Oxford University Press.
+biology 59:245--261.
 
@@
 signal; strong or negligible impacts of polytomies and branch length
-information? BMC evolutionary biology 17:53. BioMed Central.
+information? BMC evolutionary biology 17:53.
 
@@
 biological traits: The ubiquity of cross-product statistics. Evolution:
-International Journal of Organic Evolution 67:828--840. Wiley Online
-Library.
+International Journal of Organic Evolution 67:828--840.
 
@@
 data. Methods in Ecology and Evolution 1:319--329.
+
+\leavevmode\hypertarget{ref-Revell2012}{}%
+Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
+biology (and other things). Methods in Ecology and Evolution 3:217--223.
 

### 2020-04-20 Manuscript/Manuscript-1-DCA.Rmd
vsn 1

@@
 In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
-Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, while effect sizes ($Z_K$) obtained from $Kappa$ display far less variation, and increase in a steady (**uniform?**) manner with increasing phylogenetic signal.
-We conclude that Pagel's $\lambda$ should not be used for characterizing and comparing the degree of phylogenetic signal in phenotypic datasets, while comparisons of effect sizes obtained from $Kappa$ hold considerable promise in this regard.
+Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
+Alternatively, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more uniform across the range of input values, and estimates of phylogenetic signal with this measure increase in a uniform manner with increasing phylogenetic signal.
+We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution.
+By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and comparing the strength of phylogenetic signal across datasets.
 
@@
 Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression.
-Additionally, the precision of $\lambda_{est}$ was approximated by observing both the range and the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
+Additionally, the precision of $\lambda_{est}$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
 Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
 Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019].
-Because $Kappa$ and $Z_K$ differ in scale from $\lambda$, both measures were linearly normalized to a standard uniform distribution ($0\rightarrow1$), and the range and variation in the set of normalized $Kappa$ and $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were estimated for comparison.
+Because $Kappa$ and $Z_K$ differ in scale from $\lambda$, both measures were linearly normalized to a standard uniform distribution ($0\rightarrow1$), and the variation in the set of normalized $Kappa$ and $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were estimated for comparison.
 \hfill\break
@@
 This was accomplished as: $Y=\beta{X}+\epsilon$.
-Here, the association between $Y$ and $X$  was modeled using a range of values ($\beta=(0,0.25, 0.5, 0.75,1)$), and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution [$\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: see @Revell2010 for a similar simulation design].
+Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution: $\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: [see @Revell2010 for a similar simulation design].
 For each dataset, the fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained and evaluated as above.

### 2020-04-20 Manuscript/Manuscript-1-DCA.tex
vsn 1

@@
 \(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
-widely for a given input value of phylogenetic signal, while effect
-sizes (\(Z_K\)) obtained from \(Kappa\) display far less variation, and
-increase in a steady (\textbf{uniform?}) manner with increasing
-phylogenetic signal. We conclude that Pagel's \(\lambda\) should not be
-used for characterizing and comparing the degree of phylogenetic signal
-in phenotypic datasets, while comparisons of effect sizes obtained from
-\(Kappa\) hold considerable promise in this regard.
+widely for a given input value of phylogenetic signal, and that this
+variation is not constant across the range of input signal, with
+decreased precision when phylogenetic signal is of intermediate
+strength. Alternatively, variation across effect sizes (\(Z_K\))
+obtained from \(Kappa\) is far more uniform across the range of input
+values, and estimates of phylogenetic signal with this measure increase
+in a uniform manner with increasing phylogenetic signal. We conclude
+that estimates of phylogenetic signal using Pagel's \(\lambda\) are
+often innacurate, and thus interpreting strength of phylogenetic signal
+in phenotypic datasets based on this measure should be treated with
+caution. By contrast, effect sizes obtained from \(Kappa\) hold promise
+for characterizing phylogenetic signal, and comparing the strength of
+phylogenetic signal across datasets.
 
@@
 Additionally, the precision of \(\lambda_{est}\) was approximated by
-observing both the range and the variation of \(\lambda_{est}\) obtained
-at each input level of phylogenetic signal (\(\lambda_{in}\)). Finally,
-for comparison we estimated the degree of phylogenetic signal in each
-dataset using \(Kappa\) (Blomberg et al. 2003; Adams 2014a), and
-calculated its effect size (\(Z_K\): sensu Adams and Collyer 2016,
-2019a) for each dataset. Here, the effect size was estimated as:
+observing the variation of \(\lambda_{est}\) obtained at each input
+level of phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison
+we estimated the degree of phylogenetic signal in each dataset using
+\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and calculated its effect
+size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each dataset.
+Here, the effect size was estimated as:
 \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
@@
 \(\lambda\), both measures were linearly normalized to a standard
-uniform distribution (\(0\rightarrow1\)), and the range and variation in
-the set of normalized \(Kappa\) and \(Z_K\) values at each input level
-of phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
+uniform distribution (\(0\rightarrow1\)), and the variation in the set
+of normalized \(Kappa\) and \(Z_K\) values at each input level of
+phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
 \hfill\break
@@
 signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
-association between \(Y\) and \(X\) was modeled using a range of values
-(\(\beta=(0,0.25, 0.5, 0.75,1)\)), and the residual error was modeled to
-contain phylogenetic signal simulated under a Brownian motion model of
-evolution (\(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): see Revell
-2010 for a similar simulation design). For each dataset, the fit of the
-phylogenetic regression was estimated using maximum likelihood, and
-parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
+association between \(Y\) and \(X\) was modeled using a range of values:
+\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
+to contain phylogenetic signal simulated under a Brownian motion model
+of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
+Revell 2010 for a similar simulation design). For each dataset, the fit
+of the phylogenetic regression was estimated using maximum likelihood,
+and parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
 obtained and evaluated as above. All analyses were performed in R v3.6.0

### 2020-04-20 Data_Analyses/Analysis_script-DCA.R

@@ -1,0 +1,273 @@
+# Analyses
+
+treesizes <-  2^(5:10)
+
+ANOVA.data <- read.csv("Munged_Data/ANOVA.fulldataset.csv")
+ANOVA.data <- as.data.frame(ANOVA.data)
+
+Regression.data <- read.csv("Munged_Data/Regression.fulldataset.csv")
+Regression.data <- as.data.frame(Regression.data)
+
+
+# Slopes, Ranges, and Variances for Estimated Lambdas, Kappas, and Kappa Zs #####
+
+kappa_data_32 <- read.csv("Sim_Data/PB_lambda_kappacomparison_32.csv")
+kappa_data_64 <- read.csv("Sim_Data/PB_lambda_kappacomparison_64.csv")
+kappa_data_128 <- read.csv("Sim_Data/PB_lambda_kappacomparison_128.csv")
+kappa_data_256 <- read.csv("Sim_Data/PB_lambda_kappacomparison_256.csv")
+kappa_data_512 <- read.csv("Sim_Data/PB_lambda_kappacomparison_512.csv")
+kappa_data_1024 <- read.csv("Sim_Data/PB_lambda_kappacomparison_512.csv")
+
+# Getting normalized K and Z vals
+
+kappa_data_32$kappa.norm <- (kappa_data_32$kappa-min(kappa_data_32$kappa))/(range(kappa_data_32$kappa)[2] - range(kappa_data_32$kappa)[1])
+kappa_data_32$kappa.z.norm <- (kappa_data_32$kappa.z-min(kappa_data_32$kappa.z))/(range(kappa_data_32$kappa.z)[2] - range(kappa_data_32$kappa.z)[1])
+
+kappa_data_64$kappa.norm <- (kappa_data_64$kappa-min(kappa_data_64$kappa))/(range(kappa_data_64$kappa)[2] - range(kappa_data_64$kappa)[1])
+kappa_data_64$kappa.z.norm <- (kappa_data_64$kappa.z-min(na.omit(kappa_data_64$kappa.z)))/(range(na.omit(kappa_data_64$kappa.z))[2] - range(na.omit(kappa_data_64$kappa.z))[1])
+
+kappa_data_128$kappa.norm <- (kappa_data_128$kappa-min(kappa_data_128$kappa))/(range(kappa_data_128$kappa)[2] - range(kappa_data_128$kappa)[1])
[…]

### 2020-04-20 Manuscript/Manuscript-1-DCA.Rmd
new edits

@@
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation.
-However, this conclusion assumes that the precision of parameter estimation remains constant across the range of values ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
+However, this conclusion assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
 Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
@@
 
-  In this study, we evaluate the precision of Pagel's $\lambda$ to estimate known levels of phylogenetic signal in phenotypic data.
+  In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data.
 First we use computer simulations across differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
@@
 Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
-Alternatively, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more uniform across the range of input values, and estimates of phylogenetic signal with this measure increase in a uniform manner with increasing phylogenetic signal.
+By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and that the precision of estimating $Z_K$ is preferable to that observed using $\lambda$.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution.
@@
 We conducted a series of computer simulations to evaluate the ability of Pagel's [-@Pagel1999] $\lambda$ to estimate known levels of phylogenetic signal.
-Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine the extent to which tree shape affected our findings (see Supplemental Material).
+Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information).
 Our first set of simulations evaluated the extent to which values of $\lambda$ estimated from the data corresponded with actual input levels of phylogenetic signal.
@@
 Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda$.
-We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 - 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
+We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
 Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with known and differing levels of phylogenetic signal.
@@
 Additionally, the precision of $\lambda_{est}$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
-Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
+Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and from this calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
 Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019].
-Because $Kappa$ and $Z_K$ differ in scale from $\lambda$, both measures were linearly normalized to a standard uniform distribution ($0\rightarrow1$), and the variation in the set of normalized $Kappa$ and $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were estimated for comparison.
+Because $Z_K$ differs in scale from $\lambda$, we used a linear normalization to standardize $Z_K$ to a uniform distribution ($0\rightarrow1$), and variation in the set of normalized $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were then estimated for comparison.
 \hfill\break
@@
 \textbf{Figure 1}.
-Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
-As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips.
+Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
 \hfill\break
@@
 \textbf{Figure 2}.
-Estimated ANOVA slopes under PGLS.
-Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
-However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
+Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
+**A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+**D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 \hfill\break
 
-\textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
+\textbf{Figure 3}.
+Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
+\hfill\break
+
+\textbf{Figure 4}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 
@@
 \singlespacing \textbf{Figure 1}.
-Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
-As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips.
-\hfill\break
+Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
 
@@
 \singlespacing \textbf{Figure 2}.
-Estimated ANOVA slopes under PGLS.
-Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
-However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
-\hfill\break
+Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
+**A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+**D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 
@@
 \singlespacing \textbf{Figure 3}.
-Frequency of estimated lambda values published in manuscripts in 2019.
+Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
+
+\newpage
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig4.png")
+
+```
+
+\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values published in manuscripts in 2019.
 The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

### 2020-04-20 Manuscript/Manuscript-1-DCA.tex
new edits

@@
 precision in parameter estimation. However, this conclusion assumes that
-the precision of parameter estimation remains constant across the range
-of values (\(\lambda = 0 \to 1\)); an assumption that to date, has not
-been verified. Thus, despite widespread use of Pagel's (1999)
-\(\lambda\) in macroevolutionary studies, at present, we still lack a
-general understanding of the precision with which \(\lambda\) can
-estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
+the precision in estimating \(\lambda\) remains constant across its
+range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
+macroevolutionary studies, at present, we still lack a general
+understanding of the precision with which \(\lambda\) can estimate
+levels of phylogenetic signal in phenotypic datasets. \hfill\break
 
-In this study, we evaluate the precision of Pagel's \(\lambda\) to
-estimate known levels of phylogenetic signal in phenotypic data. First
+In this study, we evaluate the precision of Pagel's \(\lambda\) in
+estimating known levels of phylogenetic signal in phenotypic data. First
 we use computer simulations across differing numbers of species,
@@
 decreased precision when phylogenetic signal is of intermediate
-strength. Alternatively, variation across effect sizes (\(Z_K\))
-obtained from \(Kappa\) is far more uniform across the range of input
-values, and estimates of phylogenetic signal with this measure increase
-in a uniform manner with increasing phylogenetic signal. We conclude
-that estimates of phylogenetic signal using Pagel's \(\lambda\) are
-often innacurate, and thus interpreting strength of phylogenetic signal
-in phenotypic datasets based on this measure should be treated with
-caution. By contrast, effect sizes obtained from \(Kappa\) hold promise
-for characterizing phylogenetic signal, and comparing the strength of
-phylogenetic signal across datasets.
+strength. By contrast, variation across effect sizes (\(Z_K\)) obtained
+from \(Kappa\) is far more consistent across the range of input values,
+and that the precision of estimating \(Z_K\) is preferable to that
+observed using \(\lambda\). We conclude that estimates of phylogenetic
+signal using Pagel's \(\lambda\) are often innacurate, and thus
+interpreting strength of phylogenetic signal in phenotypic datasets
+based on this measure should be treated with caution. By contrast,
+effect sizes obtained from \(Kappa\) hold promise for characterizing
+phylogenetic signal, and comparing the strength of phylogenetic signal
+across datasets.
 
@@
 however, we also evaluated patterns on both balanced and pectinate trees
-to determine the extent to which tree shape affected our findings (see
-Supplemental Material). Our first set of simulations evaluated the
-extent to which values of \(\lambda\) estimated from the data
-corresponded with actual input levels of phylogenetic signal. For these
-simulations we generated 50 pure-birth phylogenies at each of six
-different tree sizes, ranging from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)).
-Next, we rescaled the simulated phylogenies by multiplying the internal
-branches by \(\lambda\). We used a set of values that encompassed the
-entire range of \(\lambda\): (\(\lambda_{in} = 0.0 - 1.0\); in 21
-intervals of 0.05 units), resulting in 1050 scaled phylogenies at each
-level of species richness (\(n\)). Continuous traits were then simulated
-on each phylogeny under a Brownian motion model of evolution to obtain
-datasets with known and differing levels of phylogenetic signal. These
-varied from datasets with no phylogenetic signal (when
-\(\lambda_{in} =0\)) to datsets with phylogenetic signal corresponding
-to what was expected under Brownian motion (when \(\lambda_{in} =1\)).
-\hfill\break
+to determine whether tree shape affected our findings (see Supporting
+Information). Our first set of simulations evaluated the extent to which
+values of \(\lambda\) estimated from the data corresponded with actual
+input levels of phylogenetic signal. For these simulations we generated
+50 pure-birth phylogenies at each of six different tree sizes, ranging
+from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the
+simulated phylogenies by multiplying the internal branches by
+\(\lambda\). We used a set of values that encompassed the entire range
+of \(\lambda\): (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05
+units), resulting in 1050 scaled phylogenies at each level of species
+richness (\(n\)). Continuous traits were then simulated on each
+phylogeny under a Brownian motion model of evolution to obtain datasets
+with known and differing levels of phylogenetic signal. These varied
+from datasets with no phylogenetic signal (when \(\lambda_{in} =0\)) to
+datsets with phylogenetic signal corresponding to what was expected
+under Brownian motion (when \(\lambda_{in} =1\)). \hfill\break
 
@@
 we estimated the degree of phylogenetic signal in each dataset using
-\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and calculated its effect
-size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each dataset.
-Here, the effect size was estimated as:
+\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and from this calculated
+its effect size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each
+dataset. Here, the effect size was estimated as:
 \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
@@
 obtained from the permutation distribution (for discussion see: Adams
-and Collyer 2019b). Because \(Kappa\) and \(Z_K\) differ in scale from
-\(\lambda\), both measures were linearly normalized to a standard
-uniform distribution (\(0\rightarrow1\)), and the variation in the set
-of normalized \(Kappa\) and \(Z_K\) values at each input level of
-phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
-\hfill\break
+and Collyer 2019b). Because \(Z_K\) differs in scale from \(\lambda\),
+we used a linear normalization to standardize \(Z_K\) to a uniform
+distribution (\(0\rightarrow1\)), and variation in the set of normalized
+\(Z_K\) values at each input level of phylogenetic signal
+(\(\lambda_{in}\)) were then estimated for comparison. \hfill\break
 
@@
 
-\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. As trees increase in size, the
-estimates more closely resemble the input lambdas, however considerable
-and concerning variation is apparent in trees smaller than those with
-200 tips. \hfill\break
+\textbf{Figure 1}. Precision of Pagel's \(\lambda\) across known levels
+of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
+various sizes. As phylogenies increase in size, variation in
+\(\lambda_{in}\) decreases; however the precision is not constant across
+the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
+intermediate levels of phylogenetic signal. \hfill\break
 
-\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes,
-the mean estimated slope matches the input slope, and as trees increase
-in size, the variance around this mean estimate decreases. However, for
-trees with fewer than 200 tips, the error around the estimated slope is
-considerable, where these analyses frequently estimate slopes in the
-opposite direction of the known pattern. \hfill\break
+\textbf{Figure 2}. Variation in estimates of phylogenetic signal across
+input levels of phylogenetic signal. \textbf{A}: Estimates of Pagel's
+\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
+\textbf{B} Estimates of \(Z_K\) for data simulated on phylogenies with
+128 taxa (\(n=128\)), \textbf{C} Variance in the variation of
+\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
+on phylogenies containing differing numbers of species. \textbf{D}
+Variance in the variation of \(Z_K\) across input levels of phylogenetic
+signal, estimated on phylogenies containing differing numbers of
+species. \hfill\break
 
-\textbf{Figure 3}. Frequency of estimated lambda values published in
+\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
+phylogenetic regression (\(|Y\sim X\)), across known levels of input
+phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
+As phylogenies increase in size, variation in \(\lambda_{in}\)
+decreases; however the precision is not constant across the range of
+input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
+levels of phylogenetic signal. \hfill\break
+
+\textbf{Figure 4}. Frequency of estimated lambda values published in
 manuscripts in 2019. The majority of these values were close to 0 or 1,
@@
 
-\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations
-across known lambda inputs on various tree sizes. As trees increase in
-size, the estimates more closely resemble the input lambdas, however
-considerable and concerning variation is apparent in trees smaller than
-those with 200 tips. \hfill\break
+\singlespacing \textbf{Figure 1}. Precision of Pagel's \(\lambda\)
+across known levels of input phylogenetic signal (\(\lambda_{in}\)) on
+phylogenies of various sizes. As phylogenies increase in size, variation
+in \(\lambda_{in}\) decreases; however the precision is not constant
+across the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is
+highest at intermediate levels of phylogenetic signal.
 
@@
 
-\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS.
-Across tree sizes, the mean estimated slope matches the input slope, and
-as trees increase in size, the variance around this mean estimate
-decreases. However, for trees with fewer than 200 tips, the error around
-the estimated slope is considerable, where these analyses frequently
-estimate slopes in the opposite direction of the known pattern.
-\hfill\break
+\singlespacing \textbf{Figure 2}. Variation in estimates of phylogenetic
+signal across input levels of phylogenetic signal. \textbf{A}: Estimates
+of Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
+(\(n=128\)), \textbf{B} Estimates of \(Z_K\) for data simulated on
+phylogenies with 128 taxa (\(n=128\)), \textbf{C} Variance in the
+variation of \(\lambda_{est}\) across input levels of phylogenetic
+signal, estimated on phylogenies containing differing numbers of
+species. \textbf{D} Variance in the variation of \(Z_K\) across input
+levels of phylogenetic signal, estimated on phylogenies containing
+differing numbers of species.
 
@@
 
-\includegraphics[width=0.95\linewidth]{Fig3}
+\includegraphics[width=0.95\linewidth]{Fig1}
 
-\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
+\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
+incorporated in phylogenetic regression (\(|Y\sim X\)), across known
+levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
+various sizes. As phylogenies increase in size, variation in
+\(\lambda_{in}\) decreases; however the precision is not constant across
+the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
+intermediate levels of phylogenetic signal.
+
+\newpage
+
+\includegraphics[width=0.95\linewidth]{Fig4}
+
+\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values
 published in manuscripts in 2019. The majority of these values were

### 2020-04-20 Manuscript/SupportingInfo.Rmd
new edits

---
title:

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: evolution.csl
bibliography: Lambda-Refs.bib

---
# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate

**FOR DCA**

1: balanced/symmetric trees
2: Kappa and K_Z
3: Beta from PGLS

Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.

## Simulations on differently shaped phylogenies

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees. As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units). For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution. Using these, we estimated the degree of phylogenetic signal using $\lambda$. \hfill\break

*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break

## Literature Survey

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>

### 2020-04-20 Manuscript/SupportingInfo.tex
new edits

--- 2020-04-16 Manuscript/Manuscript-1-DCA.tex
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{setspace}\doublespacing
-\usepackage{lineno}\linenumbers
 
@@
 
-\hypertarget{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}{%
-\section{Estimates of Phylogenetic Signal Based on Lambda are Often
-Inaccurate}\label{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}}
+\hypertarget{supporting-information-for-estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}{%
+\section{Supporting Information for: Estimates of Phylogenetic Signal
+Based on Lambda are Often
+Inaccurate}\label{supporting-information-for-estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}}
 
-\hfill\break
+Here we provide additional supporting information referenced in the main
+document: additional analyses, and simulation results across a wider set
+of input conditions.
 
-\textbf{Keywords}: Pagel's lambda, phylogenetic signal \hfill\break
+\hypertarget{simulations-on-differently-shaped-phylogenies}{%
+\subsection{Simulations on differently shaped
+phylogenies}\label{simulations-on-differently-shaped-phylogenies}}
 
-\textbf{Short Title}: Inaccuracies in Pagel's Lambda \hfill\break
+In addition to using pure-birth phylogenies, we explored the effect of
+tree shape on our findings using both balanced and pectinate trees. As
+before, simulations were conducted on a range of tree sizes
+(\(n=2^5 - 2^{10}\)), and across a range of input levels of phylogenetic
+signal (\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units).
+For each \(n\) and \(\lambda_{in}\) combination, 50 replicates of a
+continuous trait were simulated using a Brownian motion model of
+evolution. Using these, we estimated the degree of phylogenetic signal
+using \(\lambda\). \hfill\break
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\emph{Results}. As found with pure-birth phylogenies, estimates of
+\(\lambda\) varied \ldots.. ** FINISH THIS** (Fig. S1, S2).
 
-\{conclusion holds: interpreting the regression is not appreciably
-different (in terms of slopes and f values)\}
+\includegraphics[width=0.75\linewidth]{Fig1}
 
-\newpage
+\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known
+lambda inputs on various tree sizes. Results obtained using balanced
+trees. \hfill\break
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\includegraphics[width=0.75\linewidth]{Fig1}
 
-Investigating macroevolutionary patterns of trait variation requires a
-phylogenetic perspective, because the shared ancestry among species
-generates statistical non-independence (Felsenstein 1985; Harvey and
-Pagel 1991). Accounting for this evolutionary non-independence is the
-purview of \emph{phylogenetic comparative methods} (PCMs); a suite of
-analytical tools that condition the data on the phylogeny through the
-course of statisical evaluations of phenotypic trends (e.g., Grafen
-1989; Garland and Ives 2000; Rohlf 2001; Butler and King 2004). The past
-several decades have witnessed a rapid expansion in the development of
-PCMs to address an ever-growing set of macroevolutionary hypotheses
-(Martins and Hansen 1997; O'Meara et al. 2006; Revell and Harmon 2008;
-Beaulieu et al. 2012; Adams 2014b,a; Adams and Collyer 2018). These
-methods are predicated on the notion that phylogenetic signal -- the
-tendancy for closely related species to display similar trait values --
-is present in cross-species datasets (Felsenstein 1985; Pagel 1999;
-Blomberg et al. 2003). Indeed, under numerous evolutionary models,
-phylogenetic signal is to be expected, as stochastic character change
-along the hierarchical structure of the tree of life generates trait
-covaration among related taxa (see Felsenstein 1985; Blomberg et al.
-2003; Revell et al. 2008). \hfill\break
-
-For many macroevolutionary analyses, it is often of interest to quantify
-the degree to which phylogenetic signal is displayed in continuous
-traits. Several analytical tools have been developed for this purpose
-(e.g., Gittleman and Kot 1990; Abouheif 1999; Pagel 1999; Blomberg et
-al. 2003; Klingenberg and Gidaszewski 2010; Adams 2014a), which differ
-primarily in how they characterize the phylogenetic dependency of trait
-variation among taxa. One commonly used statistical measure,
-\emph{Kappa} (\(K\)), expresses the strength of phylogenetic signal as
-the ratio of observed trait variation to the trait variation conditioned
-on the phylogeny; scaled by what is expected under Brownian motion given
-the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
-Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
-likelihood to fit the data to the phylogeny under some model of
-evolutionary change (typically Brownian motion). The inclusion of a
-scaling parameter, \(\lambda\), transforms the lengths of the internal
-branches of the phylogeny to improve the fit, and this parameter
-describes the degree of phylogenetic signal in the dataset (Pagel 1999;
-Freckleton et al. 2002). Pagel's \(\lambda\) also has the advantage that
-it may be included when estimating the association of traits in a
-phylogenetic context, meaning that one may account for the degree of
-phylogenetic signal while conducting phylogenetic regression or ANOVA
-(see Freckleton et al. 2002). \hfill\break
-
-Several studies have investigated the statistical properties of methods
-for estimating phylogenetic signal under various conditions (e.g.,
-Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
-2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
-Revell 2010). These have largely focused on the ability of methods to
-detect the presence of phylogenetic signal (i.e., type I and type II
-error rates) under complex models of evolutionary change, across a range
-of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
-or unresolved topologies. In terms of parameter estimation, Revell
-(2010) found that regression parameters were accurately estimated when
-\(\lambda\) was included during phylogenetic regression, and Munkmuller
-et al.~(2012) demonstrated that estimates of phylogenetic signal
-obtained using various measures generally increased when input levels of
-phylogenetic signal were stronger. However, the precision of those
-estimates could not be determined, because the input levels of
-phylogenetic signal were simulated via a scaling factor (\(w\)) that was
-not directly comparable to the measures of phylogenetic signal being
-compared (see Munkemuller et al. 2012). One study (Boettiger et al.
-2012) found that estimates of Pagel's \(\lambda\) displayed less
-variation when data were simulated on a large phylogeny (\(N=281\)) as
-compared to a small one (\(N=13\)), concluding that insufficient data
-(i.e., the number of species) was the underlying cause of the lack of
-precision in parameter estimation. However, this conclusion assumes that
-the precision of parameter estimation remains constant across the range
-of values (\(\lambda = 0 \to 1\)); an assumption that to date, has not
-been verified. Thus, despite widespread use of Pagel's (1999)
-\(\lambda\) in macroevolutionary studies, at present, we still lack a
-general understanding of the precision with which \(\lambda\) can
-estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
-
-In this study, we evaluate the precision of Pagel's \(\lambda\) to
-estimate known levels of phylogenetic signal in phenotypic data. First
-we use computer simulations across differing numbers of species,
-differently shaped phylogenies, and differing input levels of
-phylogenetic signal, to explore the degree to which \(\lambda\)
-correctly identifies known levels of phylogenetic signal, and under what
-circumstances. For comparison, we also evaluate estimates of \(Kappa\)
-and its empirically-derived effect size (\(Z_K\)) for this purpose.
-Additionally, we use simulations to determine how the inclusion of
-\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
-parameter estimation, and whether levels of phylogenetic signal
-estimated in a PGLS framework are accurate. We then survey the recent
-macroevolutionary literature for published papers containing estimates
-of \(\lambda\) from empirical datasets, and compare these empirical
-estimates to patterns gleaned from our computer simulations. In general
-we find that while PGLS parameters (e.g., \(\beta\)) are accurately
-estimated with the inclusion of phylogenetic signal, estimates of
-\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
-widely for a given input value of phylogenetic signal, while effect
-sizes (\(Z_K\)) obtained from \(Kappa\) display far less variation, and
-increase in a steady (\textbf{uniform?}) manner with increasing
-phylogenetic signal. We conclude that Pagel's \(\lambda\) should not be
-used for characterizing and comparing the degree of phylogenetic signal
-in phenotypic datasets, while comparisons of effect sizes obtained from
-\(Kappa\) hold considerable promise in this regard.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{simulations}{%
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}
-
-We conducted a series of computer simulations to evaluate the ability of
-Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
-signal. Our primary simulations were based on pure-birth phylogenies;
-however, we also evaluated patterns on both balanced and pectinate trees
-to determine the extent to which tree shape affected our findings (see
-Supplemental Material). Our first set of simulations evaluated the
-extent to which values of \(\lambda\) estimated from the data
-corresponded with actual input levels of phylogenetic signal. For these
-simulations we generated 50 pure-birth phylogenies at each of six
-different tree sizes, ranging from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)).
-Next, we rescaled the simulated phylogenies by multiplying the internal
-branches by \(\lambda\). We used a set of values that encompassed the
-entire range of \(\lambda\): (\(\lambda_{in} = 0.0 - 1.0\); in 21
-intervals of 0.05 units), resulting in 1050 scaled phylogenies at each
-level of species richness (\(n\)). Continuous traits were then simulated
-on each phylogeny under a Brownian motion model of evolution to obtain
-datasets with known and differing levels of phylogenetic signal. These
-varied from datasets with no phylogenetic signal (when
-\(\lambda_{in} =0\)) to datsets with phylogenetic signal corresponding
-to what was expected under Brownian motion (when \(\lambda_{in} =1\)).
-\hfill\break
-
-Using the simulated phylogenies we estimated the phylogenetic signal in
-each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
-each level of species richness (\(n\)), we assessed the relationship
-between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
-Additionally, the precision of \(\lambda_{est}\) was approximated by
-observing both the range and the variation of \(\lambda_{est}\) obtained
-at each input level of phylogenetic signal (\(\lambda_{in}\)). Finally,
-for comparison we estimated the degree of phylogenetic signal in each
-dataset using \(Kappa\) (Blomberg et al. 2003; Adams 2014a), and
-calculated its effect size (\(Z_K\): sensu Adams and Collyer 2016,
-2019a) for each dataset. Here, the effect size was estimated as:
-\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
-observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
-standard deviation of the empirical sampling distribution of values
-obtained from the permutation distribution (for discussion see: Adams
-and Collyer 2019b). Because \(Kappa\) and \(Z_K\) differ in scale from
-\(\lambda\), both measures were linearly normalized to a standard
-uniform distribution (\(0\rightarrow1\)), and the range and variation in
-the set of normalized \(Kappa\) and \(Z_K\) values at each input level
-of phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
-\hfill\break
-
-Our second set of simulations evaluated the extent to which values of
-\(\lambda\) estimated in PGLS regression and ANOVA (i.e., \(Y\sim{X}\))
-corresponded with actual input levels of phylogenetic signal in the
-response variable. As before we generated 50 pure-birth phylogenies at
-each of six levels of species richness (\(n=2^5 - 2^{10}\)), and
-rescaled each with a set of input \(\lambda\) values
-(\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units). Next, an
-independent variable \(X\) was simulated on each phylogeny under a
-Brownian motion model of evolution (for PGLS regression). For
-phylogenetic ANOVA, random groups (\(X\)) were obtained by simulating a
-discrete (binary) character on each phylogeny. Next, the dependent
-variable was simulated in such a manner as to contain a known
-relationship with \(X\) plus random error containing phylogenetic
-signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
-association between \(Y\) and \(X\) was modeled using a range of values
-(\(\beta=(0,0.25, 0.5, 0.75,1)\)), and the residual error was modeled to
-contain phylogenetic signal simulated under a Brownian motion model of
-evolution (\(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): see Revell
-2010 for a similar simulation design). For each dataset, the fit of the
-phylogenetic regression was estimated using maximum likelihood, and
-parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
-obtained and evaluated as above. All analyses were performed in R v3.6.0
-(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
-2008), \texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell
-2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
-al. 2020). R-scripts are found in the Supplemental Material.
+\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known
+lambda inputs on various tree sizes. Results obtained using pectinate
+trees. \hfill\break
 
 \hypertarget{literature-survey}{%
-\subsection{\texorpdfstring{\emph{Literature
-Survey}}{Literature Survey}}\label{literature-survey}}
-
-To determine how Pagel's \(\lambda\) is commonly utilized in empirical
-studies, we conducted a literature survey. From Google.scholar we
-obtained a list of all papers published in 2019 that used \(\lambda\);
-resulting in 341 studies. For each study, we extracted all
-\(\lambda_{est}\), the size of the phylogeny (\(n\)), and noted whether
-authors reported confidence intervals or performed significance tests
-assessing difference of \(\lambda_{est}\) from either zero or one. We
-also noted whether biological interpretations based on \(\lambda_{est}\)
-were made, and for studies that reported more than one
-\(\lambda_{est}\), we also noted whether these were compared in some
-manner, and whether such comparisons were accompanied with statistical
-tests between \(\lambda_{est}\). \hfill\break
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{simulations-1}{%
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations-1}}
-
-lambda: (Fig 1) not good predictor of input physignal (regression),
-particularly at low sample sizes. Also, similar to early results of
-Boettiger (but greatly expanded) range and variation of lambda very
-large at low N, and generally decreases with larger N. However, the
-precision was not uniform across the range of lambda-in. Variation in
-lambda-est was greatest at intermediate levels of physig, and descreased
-at lower and higher values. Thus, the overall dispersion across the
-range of input values resembled an asymmetric lens (\textbf{NAME?}),
-with less variation at lambda-in close to 1, greatest variation in
-middle, and less (but substantial) variation near lambda-in of zero.
-Notably, even at very large species richness (n=1024) the range of
-lam-est was quite large: e.g., l-in 0.5 has range of 45\% of all
-possible values. Thus, lambda-est is not very precise in terms of
-estimating the true l-in parameter, meaning that using lambda to
-biologically interpret the `strength' of phylogenetic signal in a
-dataset, or compare between datasets, is fraught with difficulty.
-\hfill\break
-
-{[}insert Figure 1 here{]} \hfill\break
-
-By contrast, (FIg 2) Kappa. Much smaller range and variation as compared
-with lambda, but with larger l-in, variation in kappa does increase.
-However, this is not the case with Z-K. Z-K increases predictably
-(linearly, progressively\ldots) with L-in, and importantly variation in
-Z-K remains nearly constant across the range of input values. this means
-that Z-K is equally precise in estimating the strength of phylogenetic
-signal, and may be more adept at facilitating useful statements on teh
-relative strength of phylogenetic signal within and across datasets.
-\hfill\break
-
-\textbf{NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K,
-Z-k-normalized.}
-
-{[}insert Figure 2 here{]} \hfill\break
-
-\hypertarget{literature-survey-1}{%
-\subsection{\texorpdfstring{\emph{Literature
-Survey}}{Literature Survey}}\label{literature-survey-1}}
-
-We found 182 manuscripts from 2019 that estimated and reported Pagel's
-lambda values using PGLS methods. These papers averaged 8.527 lambda
-values, ranging from a single lambda estimate up to 71 estimated
-lambdas. Almost exactly half of the published lambda estimates were
-either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3). 73.32\% of
-the published lambdas were estimated using phylogenies with fewer than
-200 tips, and 348 lambda estimates (8.57\% of all published estimates)
-came from phylogenies with fewer than 30 tips. \hfill\break
-
-{[}insert Figure 3 here{]} \hfill\break
-
-Many of the reviewed manuscripts liberally interpreted the magnitude of
-the estimate lambda, using phrases such as ``strong'' or ``weak''
-phylogentic signal when statistically, all that was clear was a
-difference between the estimated lambda and 0 or 1 respectively. We
-estimated that about 20.49\% of the manuscripts revealed some sort of
-biological interpretation of the magnitude of estimated phylogenetic
-signal that overreached the statistical findings. We also identified
-seven manuscripts as having inappropriately interpreted differences in
-lambda values, indicating that some traits had stronger or weaker signal
-than other traits without the appropriate statistical tests.
-\hfill\break
-
-As is evidenced by macroevolutionary papers published in 2019 papers,
-Pagel's lambda estimation methods are often misused and
-over-interpretted. Despite the urging of Boettiger and colleagues to
-publish confidence intervals with all lambda parameter estimates, only
-18\% of papers published in 2019 do so.
-
-\hypertarget{results-1}{%
-\section{Results}\label{results-1}}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-1: summary paragraph
-
-2: expand on Lambda.. lambda innacurate, not precise, level of precision
-varies with input physig (worse in mid-range). Patterns worse with PGLS,
-though beta still estimated properly. Conclusion, lambda not overly
-useful.
-
-3: By contrast, effect size Z-K useful, equally precise across range of
-values. Can be used to characterize the strength of physignal, and
-because robust to input levels, etc. may be used to compare across
-datasets.
-
-Somewhere, recognize that this is somewhat `backwards' from prior
-recommendations where Kappa had somewhat lower performance in terms of
-type I and type II error (which?? I forget). However, recall that those
-studies did not examine the precision of the estimates. Nor was Z-k
-included, because it was not yet invented. So Use of Z-k should make
-good sense here.
-
-Closing paragraph.
-
-\hfill\break
-
-More discussion paragraphs
+\subsection{Literature Survey}\label{literature-survey}}
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abouheif1999}{}%
-Abouheif, E. 1999. A method for testing the assumption of phylogenetic
-independence in comparative data. Evolutionary Ecology Research
-1:895--909.
-
-\leavevmode\hypertarget{ref-Adams2014a}{}%
-Adams, D. C. 2014a. A generalized Kappa statistic for estimating
-phylogenetic signal from shape and other high-dimensional dultivariate
-data. Systematic Biology 63:685--697.
-
-\leavevmode\hypertarget{ref-Adams2014b}{}%
-Adams, D. C. 2014b. A method for assessing phylogenetic least squares
-models for shape and other high-dimensional multivariate data. Evolution
-68:2675--2688.
-
-\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
-Adams, D. C., and M. L. Collyer. 2019a. Comparing the strength of
-modular signal, and evaluating alternative modular hypotheses, using
-covariance ratio effect sizes with morphometric data. Evolution
-73:2352--2367.
-
-\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
-Adams, D. C., and M. L. Collyer. 2016. On the comparison of the strength
-of morphological integration across morphometric datasets. Evolution
-70:2623--2631.
-
-\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
-Adams, D. C., and M. L. Collyer. 2018. Phylogenetic anova: Group-clade
-aggregation, biological challenges, and a refined permutation procedure.
-Evolution 72:1204--1215.
-
-\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
-Adams, D. C., and M. L. Collyer. 2019b. Phylogenetic comparative methods
-and the evolution of multivariate phenotypes. Annual Review of Ecology,
-Evolution, and Systematics 50:405--425.
-
-\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
-Software for geometric morphometric analyses. R package version 3.2.1.
-
-\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
-Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
-the collection and analysis of geometric morphometric shape data.
-Methods in Ecology and Evolution 4:393--399.
-
-\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
-Beaulieu, J. M., D. C. Jhwueng, C. Boettiger, and B. C. O'Meara. 2012.
-Modeling stabilizing selection: Expanding the ornstein-uhlenbeck model
-of adaptive evolution. Evolution 66:2369--2383.
-
-\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
-Blomberg, S. P., T. Garland, and A. R. Ives. 2003. Testing for
-phylogenetic signal in comparative data: Behavioral traits are more
-labile. Evolution 57:717--745.
-
-\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
-Boettiger, C., G. Coop, and P. Ralph. 2012. Is your phylogeny
-informative? Measuring the power of comparative methods. Evolution
-67:2240--2251.
-
-\leavevmode\hypertarget{ref-ButlerKing2004}{}%
-Butler, M. A., and A. A. King. 2004. Phylogenetic comparative analysis:
-A modeling approach for adaptive evolution. American Naturalist
-164:683--695.
-
-\leavevmode\hypertarget{ref-DinizFilho2012}{}%
-Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
-comparison of metrics for estimating phylogenetic signal under
-alternative evolutionary models. Genetics and Molecular Biology
-35:673--679.
-
-\leavevmode\hypertarget{ref-Felsenstein1985}{}%
-Felsenstein, J. 1985. Phylogenies and the comparative method. American
-Naturalist 125:1--15.
-
-\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
-Freckleton, R. P., P. H. Harvey, and M. Pagel. 2002. Phylogenetic
-analysis and comparative data: A test and review of evidence. American
-Naturalist 160:712--726.
-
-\leavevmode\hypertarget{ref-GarlandIves2000}{}%
-Garland, T. J., and A. R. Ives. 2000. Using the past to predict the
-present: Confidence intervals for regression equations in phylogenetic
-comparative methods. American Naturalist 155:346--364.
-
-\leavevmode\hypertarget{ref-Gittleman1990}{}%
-Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
-model for estimating phylogenetic effects. Systematic Zoology
-39:227--241.
-
-\leavevmode\hypertarget{ref-Grafen1989}{}%
-Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
-of the Royal Society of London B, Biological Sciences 326:119--157.
-
-\leavevmode\hypertarget{ref-Harmon2008}{}%
-Harmon, L. J., J. T. Weir, C. D. Brock, R. E. Glor, and W. Challenger.
-2008. GEIGER: Investigating evolutionary radiations. Bioinformatics
-24:129--131.
-
-\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
-Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
-evolutionary biology. Oxford University Press, Oxford.
-
-\leavevmode\hypertarget{ref-Klingenberg2010}{}%
-Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
-phylogenetic signals and homoplasy in morphometric data. Systematic
-biology 59:245--261.
-
-\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
-Martins, E. P., and T. F. Hansen. 1997. Phylogenies and the comparative
-method: A general approach to incorporating phylogenetic information
-into the analysis of interspecific data. American Naturalist
-149:646--667.
-
-\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
-Molina-Venegas, R., and M. A. Rodriguez. 2017. Revisiting phylogenetic
-signal; strong or negligible impacts of polytomies and branch length
-information? BMC evolutionary biology 17:53.
-
-\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
-Munkemuller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
-Schiffers, and W. Thuiller. 2012. How to measure and test phylogenetic
-signal. Methods in Ecology and Evolution 3:743--756.
-
-\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
-O'Meara, B. C., C. Ane, M. J. Sanderson, and P. C. Wainwright. 2006.
-Testing for different rates of continuous trait evolution using
-likelihood. Evolution 60:922--933.
-
-\leavevmode\hypertarget{ref-Orme2013}{}%
-Orme, D., R. P. Freckleton, G. H. Thomas, T. Petzoldt, S. A. Fritz, and
-N. Isaac. 2013. CAPER: Comparative analyses of phylogenetics and
-evolution in r. Methods in Ecology and Evolution 3:145--151.
-
-\leavevmode\hypertarget{ref-Pagel1999}{}%
-Pagel, M. D. 1999. Inferring the historical patterns of biological
-evolution. Nature 401:877--884.
-
-\leavevmode\hypertarget{ref-Pavoine2012}{}%
-Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
-biological traits: The ubiquity of cross-product statistics. Evolution:
-International Journal of Organic Evolution 67:828--840.
-
-\leavevmode\hypertarget{ref-RCT}{}%
-R Core Team. 2019. R: A language and environment for statistical
-computing. R Foundation for Statistical Computing, Vienna, Austria.
-
-\leavevmode\hypertarget{ref-Revell2010}{}%
-Revell, L. J. 2010. Phylogenetic signal and linear regression on species
-data. Methods in Ecology and Evolution 1:319--329.
-
-\leavevmode\hypertarget{ref-Revell2012}{}%
-Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
-biology (and other things). Methods in Ecology and Evolution 3:217--223.
-
-\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
-Revell, L. J., and L. J. Harmon. 2008. Testing quantitative genetic
-hypotheses about the evolutionary rate matrix for continuous characters.
-Evolutionary Ecology Research 10:311--331.
-
-\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
-Revell, L. J., L. J. Harmon, and D. C. Collar. 2008. Phylogenetic
-signal, evolutionary process, and rate. Systematic Biology 57:591--601.
-
-\leavevmode\hypertarget{ref-Rohlf2001}{}%
-Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
-variables: Geometric interpretations. Evolution 55:2143--2160.
-
-\newpage
-
-\hypertarget{figure-legends}{%
-\section{Figure Legends}\label{figure-legends}}
-
-\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. As trees increase in size, the
-estimates more closely resemble the input lambdas, however considerable
-and concerning variation is apparent in trees smaller than those with
-200 tips. \hfill\break
-
-\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes,
-the mean estimated slope matches the input slope, and as trees increase
-in size, the variance around this mean estimate decreases. However, for
-trees with fewer than 200 tips, the error around the estimated slope is
-considerable, where these analyses frequently estimate slopes in the
-opposite direction of the known pattern. \hfill\break
-
-\textbf{Figure 3}. Frequency of estimated lambda values published in
-manuscripts in 2019. The majority of these values were close to 0 or 1,
-and from phylogenies with fewer than 200 taxa.
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig1}
-
-\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations
-across known lambda inputs on various tree sizes. As trees increase in
-size, the estimates more closely resemble the input lambdas, however
-considerable and concerning variation is apparent in trees smaller than
-those with 200 tips. \hfill\break
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig2}
-
-\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS.
-Across tree sizes, the mean estimated slope matches the input slope, and
-as trees increase in size, the variance around this mean estimate
-decreases. However, for trees with fewer than 200 tips, the error around
-the estimated slope is considerable, where these analyses frequently
-estimate slopes in the opposite direction of the known pattern.
-\hfill\break
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig3}
-
-\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
-published in manuscripts in 2019. The majority of these values were
-close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 

### 2020-04-20 Manuscript/Manuscript-1-DCA.Rmd
vsn 1

@@
 Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
-By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and that the precision of estimating $Z_K$ is preferable to that observed using $\lambda$.
+By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and the precision of estimating $Z_K$ is preferable to that observed using $\lambda$.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution.
@@
 
-We conducted a series of computer simulations to evaluate the ability of Pagel's [-@Pagel1999] $\lambda$ to estimate known levels of phylogenetic signal.
+We conducted a series of computer simulations to evaluate the ability of Pagel's $\lambda$ to estimate known levels of phylogenetic signal.
 Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information).
@@
 Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression.
-Additionally, the precision of $\lambda_{est}$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
+Additionally, the precision of $\lambda$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
 Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and from this calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
@@
 All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
-R-scripts are found in the Supplemental Material.
+R-scripts are found in the Supporting Information.
 
@@
 
-lambda: (Fig 1) not good predictor of input physignal (regression), particularly at low sample sizes.
-Also, similar to early results of Boettiger (but greatly expanded) range and variation of lambda very large at low N, and generally decreases with larger N.
-However, the precision was not uniform across the range of lambda-in.
-Variation in lambda-est was greatest at intermediate levels of physig, and descreased at lower and higher values.
-Thus, the overall dispersion across the range of input values resembled an asymmetric lens (**NAME?**), with less variation at lambda-in close to 1, greatest variation in middle, and less (but substantial) variation near lambda-in of zero.
-Notably,  even at very large species richness (n=1024) the range of lam-est was quite large: e.g., l-in 0.5 has range of 45% of all possible values.
-Thus, lambda-est is not very precise in terms of estimating the true l-in parameter, meaning that using lambda to biologically interpret the 'strength' of phylogenetic signal in a dataset, or compare between datasets, is fraught with difficulty.
+Our primary series of simulations revealed several patterns.
+First, the relationship between $\lambda_{est}$ and input levels of phylogenetic signal was $\sim\beta=1.0$ at each level of species richness ($n=32, 64, 128, 256, 512, 1024$).
+This implied that the mean $\lambda_{est}$ across datasets for a given set of simulation conditions reflected input levels of phylogenetic signal.
+However, as shown in Figure 1, the precision of those estimates varied widely, and in several interesting ways.
+Predictably, precision was worst at low levels of species richness, where in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{est}=0.5$: range = $0.0\to0.985$).
+Second, as species richness increased, variation across estimates of $\lambda$ decreased (Figure 1).
+This confirmed the pattern identified by Boettiger et al.
+[-@Boettiger_et_al2012] on  a small ($n=13$) and large ($n=281$) phylogeny, and demonstrated that this trend was general across a wider range of species richness.
+Importantly however, our broader set of simulations revealed that the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, and that there was a particular pattern to the imprecision of estimating this phylogenetic parameter.
+Specifically, we found that variation in $\lambda_{est}$ was highest at intermediate levels of phylogenetic signal, and decreased at both lower and higher levels of input signal (Figure 1).
+This implied that the precision in estimating $\lambda$ was worst at intermediate levels of phylogenetic signal; a pattern that remained consistent across all levels of species richness evaluated here (Figure 1).
+Notably, even at very large levels of species richness, the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=1024$; $\lambda_{est}=0.5$: range = $0.31\to 0.68$).
+From this we conclude that $\lambda_{est}$ is not an accurate estimator of the strength of phylogenetic signal in phenotypic datasets, and that biological interpretations of the level of phylogenetic signal based on $\lambda$ should be treated with caution.
 \hfill\break
@@
 
-By contrast,  (FIg 2) Kappa.
+**STOPPED HERE**By contrast,  (FIg 2) Kappa.
 Much smaller range and variation as compared with lambda, but with larger l-in, variation in kappa does increase.
@@
 2: expand on Lambda..
-lambda innacurate, not precise, level of precision varies with input physig (worse in mid-range). Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
+lambda innacurate, not precise, level of precision varies with input physig (worse in mid-range).
+NEW RESULT.
+We are first to show this.
+NOTE: pattern is obvious with reflection.
+Since it is a 'bounded' parameter estimation should be best at the extremes...
+(state this?)..
+hmm.
+
+Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
 
@@
 Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-**A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-**D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 \hfill\break
@@
 Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-**A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-**D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 

### 2020-04-20 Manuscript/Manuscript-1-DCA.tex
vsn 1

@@
 from \(Kappa\) is far more consistent across the range of input values,
-and that the precision of estimating \(Z_K\) is preferable to that
-observed using \(\lambda\). We conclude that estimates of phylogenetic
-signal using Pagel's \(\lambda\) are often innacurate, and thus
-interpreting strength of phylogenetic signal in phenotypic datasets
-based on this measure should be treated with caution. By contrast,
-effect sizes obtained from \(Kappa\) hold promise for characterizing
-phylogenetic signal, and comparing the strength of phylogenetic signal
-across datasets.
+and the precision of estimating \(Z_K\) is preferable to that observed
+using \(\lambda\). We conclude that estimates of phylogenetic signal
+using Pagel's \(\lambda\) are often innacurate, and thus interpreting
+strength of phylogenetic signal in phenotypic datasets based on this
+measure should be treated with caution. By contrast, effect sizes
+obtained from \(Kappa\) hold promise for characterizing phylogenetic
+signal, and comparing the strength of phylogenetic signal across
+datasets.
 
@@
 We conducted a series of computer simulations to evaluate the ability of
-Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
-signal. Our primary simulations were based on pure-birth phylogenies;
-however, we also evaluated patterns on both balanced and pectinate trees
-to determine whether tree shape affected our findings (see Supporting
+Pagel's \(\lambda\) to estimate known levels of phylogenetic signal. Our
+primary simulations were based on pure-birth phylogenies; however, we
+also evaluated patterns on both balanced and pectinate trees to
+determine whether tree shape affected our findings (see Supporting
 Information). Our first set of simulations evaluated the extent to which
@@
 between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
-Additionally, the precision of \(\lambda_{est}\) was approximated by
-observing the variation of \(\lambda_{est}\) obtained at each input
-level of phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison
-we estimated the degree of phylogenetic signal in each dataset using
+Additionally, the precision of \(\lambda\) was approximated by observing
+the variation of \(\lambda_{est}\) obtained at each input level of
+phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison we
+estimated the degree of phylogenetic signal in each dataset using
 \(Kappa\) (Blomberg et al. 2003; Adams 2014a), and from this calculated
@@
 2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
-al. 2020). R-scripts are found in the Supplemental Material.
+al. 2020). R-scripts are found in the Supporting Information.
 
@@
 
-lambda: (Fig 1) not good predictor of input physignal (regression),
-particularly at low sample sizes. Also, similar to early results of
-Boettiger (but greatly expanded) range and variation of lambda very
-large at low N, and generally decreases with larger N. However, the
-precision was not uniform across the range of lambda-in. Variation in
-lambda-est was greatest at intermediate levels of physig, and descreased
-at lower and higher values. Thus, the overall dispersion across the
-range of input values resembled an asymmetric lens (\textbf{NAME?}),
-with less variation at lambda-in close to 1, greatest variation in
-middle, and less (but substantial) variation near lambda-in of zero.
-Notably, even at very large species richness (n=1024) the range of
-lam-est was quite large: e.g., l-in 0.5 has range of 45\% of all
-possible values. Thus, lambda-est is not very precise in terms of
-estimating the true l-in parameter, meaning that using lambda to
-biologically interpret the `strength' of phylogenetic signal in a
-dataset, or compare between datasets, is fraught with difficulty.
-\hfill\break
+Our primary series of simulations revealed several patterns. First, the
+relationship between \(\lambda_{est}\) and input levels of phylogenetic
+signal was \(\sim\beta=1.0\) at each level of species richness
+(\(n=32, 64, 128, 256, 512, 1024\)). This implied that the mean
+\(\lambda_{est}\) across datasets for a given set of simulation
+conditions reflected input levels of phylogenetic signal. However, as
+shown in Figure 1, the precision of those estimates varied widely, and
+in several interesting ways. Predictably, precision was worst at low
+levels of species richness, where in many cases the set of
+\(\lambda_{est}\) spanned nearly the entire range of possible values
+(e.g., \(n=32\); \(\lambda_{est}=0.5\): range = \(0.0\to0.985\)).
+Second, as species richness increased, variation across estimates of
+\(\lambda\) decreased (Figure 1). This confirmed the pattern identified
+by Boettiger et al.~(2012) on a small (\(n=13\)) and large (\(n=281\))
+phylogeny, and demonstrated that this trend was general across a wider
+range of species richness. Importantly however, our broader set of
+simulations revealed that the precision of \(\lambda_{est}\) was not
+uniform across all levels of phylogenetic signal, and that there was a
+particular pattern to the imprecision of estimating this phylogenetic
+parameter. Specifically, we found that variation in \(\lambda_{est}\)
+was highest at intermediate levels of phylogenetic signal, and decreased
+at both lower and higher levels of input signal (Figure 1). This implied
+that the precision in estimating \(\lambda\) was worst at intermediate
+levels of phylogenetic signal; a pattern that remained consistent across
+all levels of species richness evaluated here (Figure 1). Notably, even
+at very large levels of species richness, the range of \(\lambda_{est}\)
+still encompassed a substantial portion of possible values (e.g.,
+\(n=1024\); \(\lambda_{est}=0.5\): range = \(0.31\to 0.68\)). From this
+we conclude that \(\lambda_{est}\) is not an accurate estimator of the
+strength of phylogenetic signal in phenotypic datasets, and that
+biological interpretations of the level of phylogenetic signal based on
+\(\lambda\) should be treated with caution. \hfill\break
 
@@
 2: expand on Lambda.. lambda innacurate, not precise, level of precision
-varies with input physig (worse in mid-range). Patterns worse with PGLS,
-though beta still estimated properly. Conclusion, lambda not overly
-useful.
+varies with input physig (worse in mid-range). NEW RESULT. We are first
+to show this. NOTE: pattern is obvious with reflection. Since it is a
+`bounded' parameter estimation should be best at the extremes\ldots{}
+(state this?).. hmm.
+
+Patterns worse with PGLS, though beta still estimated properly.
+Conclusion, lambda not overly useful.
 
@@
 \textbf{Figure 2}. Variation in estimates of phylogenetic signal across
-input levels of phylogenetic signal. \textbf{A}: Estimates of Pagel's
+input levels of phylogenetic signal. (A) Estimates of Pagel's
 \(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
-\textbf{B} Estimates of \(Z_K\) for data simulated on phylogenies with
-128 taxa (\(n=128\)), \textbf{C} Variance in the variation of
-\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
-on phylogenies containing differing numbers of species. \textbf{D}
-Variance in the variation of \(Z_K\) across input levels of phylogenetic
-signal, estimated on phylogenies containing differing numbers of
-species. \hfill\break
+(B) Estimates of \(Z_K\) for data simulated on phylogenies with 128 taxa
+(\(n=128\)), (C) Variance in the variation of \(\lambda_{est}\) across
+input levels of phylogenetic signal, estimated on phylogenies containing
+differing numbers of species. (D) Variance in the variation of \(Z_K\)
+across input levels of phylogenetic signal, estimated on phylogenies
+containing differing numbers of species. \hfill\break
 
@@
 \singlespacing \textbf{Figure 2}. Variation in estimates of phylogenetic
-signal across input levels of phylogenetic signal. \textbf{A}: Estimates
-of Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
-(\(n=128\)), \textbf{B} Estimates of \(Z_K\) for data simulated on
-phylogenies with 128 taxa (\(n=128\)), \textbf{C} Variance in the
-variation of \(\lambda_{est}\) across input levels of phylogenetic
-signal, estimated on phylogenies containing differing numbers of
-species. \textbf{D} Variance in the variation of \(Z_K\) across input
-levels of phylogenetic signal, estimated on phylogenies containing
-differing numbers of species.
+signal across input levels of phylogenetic signal. (A) Estimates of
+Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
+(\(n=128\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
+with 128 taxa (\(n=128\)), (C) Variance in the variation of
+\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
+on phylogenies containing differing numbers of species. (D) Variance in
+the variation of \(Z_K\) across input levels of phylogenetic signal,
+estimated on phylogenies containing differing numbers of species.
 
@@
 
-\includegraphics[width=0.95\linewidth]{Fig1}
+\includegraphics[width=0.95\linewidth]{Fig3}
 

### 2020-04-21 Manuscript/Manuscript-1-DCA.Rmd
vsn 1 edits

@@
 The inclusion of a scaling parameter, $\lambda$,  transforms the lengths of the internal branches of the phylogeny to improve the fit, and this parameter describes the degree of phylogenetic signal in the dataset [@Pagel1999; @Freckleton_et_al2002].
-Pagel's $\lambda$ also has the advantage that it may be included when estimating the association of traits in a phylogenetic context, meaning that one may account for the degree of phylogenetic signal while conducting phylogenetic regression or ANOVA [see @Freckleton_et_al2002].
+Pagel's $\lambda$ also has the appeal that it may be included when estimating the association of traits in a phylogenetic context, meaning that one may account for the degree of phylogenetic signal while conducting phylogenetic regression or ANOVA [see @Freckleton_et_al2002].
 \hfill\break
@@
 However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012].
-One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation.
+One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates.
 However, this conclusion assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
@@
 In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data.
-First we use computer simulations across differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
-For comparison, we also evaluate estimates of $Kappa$ and its empirically-derived effect size ($Z_K$) for this purpose.
+We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
+We also evaluate the precision of a standardized effect size ($Z_K$) based on $Kappa$ for comparison.
 Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic regression and ANOVA (i.e., PGLS) affects parameter estimation, and whether levels of phylogenetic signal estimated in a PGLS framework are accurate.
@@
 In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
-Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
-By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and the precision of estimating $Z_K$ is preferable to that observed using $\lambda$.
-We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution.
-By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and comparing the strength of phylogenetic signal across datasets.
+We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
+In addition, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
+By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and the precision of $Z_K$ is generally preferable to that of $\lambda_{est}$.
+Further, because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable, and may be compared across datasets.
+We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
+By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
 
@@
 Additionally, the precision of $\lambda$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
-Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and from this calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset.
-Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019].
-Because $Z_K$ differs in scale from $\lambda$, we used a linear normalization to standardize $Z_K$ to a uniform distribution ($0\rightarrow1$), and variation in the set of normalized $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were then estimated for comparison.
+Finally, for comparison we characterized the strength of phylogenetic signal in each dataset using a standardized effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] based on $Kappa$.
+As suggested by Adams and Collyer [-@AdamsCollyer2019], the effect size for phylogenetic signal may be estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution.
+$Z_K$ describes the strength of phylogenetic signal as a standard deviate from its sampling distribution, and thus directly measures the strength of signal in a manner that is comparable across datasets.
+Variation in the set of $Z_K$ at each input level of phylogenetic signal was then calculated as an estimate of precision in $Z_K$.
+However, because $Z_K$ differs in scale from $\lambda$, we used a linear normalization to standardize $Z_K$ to a uniform distribution ($0\rightarrow1$), and estimated the precision of $Z_K$ from the normalized values.
 \hfill\break
@@
 We also noted whether biological interpretations based on $\lambda_{est}$ were made, and for studies that reported more than one $\lambda_{est}$, we also noted whether these were compared in some manner, and whether such comparisons were accompanied with statistical tests between $\lambda_{est}$.
-\hfill\break
 
@@
 
-Our primary series of simulations revealed several patterns.
-First, the relationship between $\lambda_{est}$ and input levels of phylogenetic signal was $\sim\beta=1.0$ at each level of species richness ($n=32, 64, 128, 256, 512, 1024$).
-This implied that the mean $\lambda_{est}$ across datasets for a given set of simulation conditions reflected input levels of phylogenetic signal.
+Our first set of simulations revealed several patterns.
+First, the relationship between $\lambda_{est}$ and $\lambda_{in}$ across the range input levels of phylogenetic signal was $\sim\beta=1.0$; indicating that the average $\lambda_{est}$ across datasets for a given set of simulation condition correctly reflected input levels of phylogenetic signal.
 However, as shown in Figure 1, the precision of those estimates varied widely, and in several interesting ways.
-Predictably, precision was worst at low levels of species richness, where in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{est}=0.5$: range = $0.0\to0.985$).
+Predictably, precision was worse at low levels of species richness, where in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: range of $\lambda_{est}= 0.0\to 0.985$).
 Second, as species richness increased, variation across estimates of $\lambda$ decreased (Figure 1).
 This confirmed the pattern identified by Boettiger et al.
-[-@Boettiger_et_al2012] on  a small ($n=13$) and large ($n=281$) phylogeny, and demonstrated that this trend was general across a wider range of species richness.
-Importantly however, our broader set of simulations revealed that the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, and that there was a particular pattern to the imprecision of estimating this phylogenetic parameter.
+[-@Boettiger_et_al2012] on  a small ($n=13$) versus a large ($n=281$) phylogeny, and demonstrated that the trend of increasing precision with higher species richness was general.
+Importantly however, our broader set of simulations revealed that the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal.
 Specifically, we found that variation in $\lambda_{est}$ was highest at intermediate levels of phylogenetic signal, and decreased at both lower and higher levels of input signal (Figure 1).
-This implied that the precision in estimating $\lambda$ was worst at intermediate levels of phylogenetic signal; a pattern that remained consistent across all levels of species richness evaluated here (Figure 1).
-Notably, even at very large levels of species richness, the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=1024$; $\lambda_{est}=0.5$: range = $0.31\to 0.68$).
-From this we conclude that $\lambda_{est}$ is not an accurate estimator of the strength of phylogenetic signal in phenotypic datasets, and that biological interpretations of the level of phylogenetic signal based on $\lambda$ should be treated with caution.
+This implied that the precision in estimating $\lambda$ was worse at intermediate values, and improved as the levels of phylogenetic signal were closer to $\lambda_{in}=0$ (no phylogenetic signal) or $\lambda_{in}=1$ (Brownian motion).
+Notably, even at large levels of species richness, the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: range of $\lambda_{est} = 0.32\to 0.68$).
+Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; range of $\lambda_{in} = 0.25 \to 0.65$).
+Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
 \hfill\break
@@
 
-**STOPPED HERE**By contrast,  (FIg 2) Kappa.
-Much smaller range and variation as compared with lambda, but with larger l-in, variation in kappa does increase.
-However, this is not the case with Z-K.
-Z-K increases predictably (linearly, progressively...) with L-in, and importantly variation in Z-K remains nearly constant across the range of input values.
-this means that Z-K is equally precise in estimating the strength of phylogenetic signal, and may be more adept at facilitating useful statements on teh relative strength of phylogenetic signal within and across datasets.
+By contrast, when characterizing phylogenetic signal with the standardized effect size ($Z_K$) of $Kappa$, we found that the precision of $Z_K$ was more stable, as variation across datasets was far more consistent across the range of input values.
+For example, when $n=128$ the precision of $\lambda_{est}$ (Figure 2A) varied considerably more across input levels of phylogenetic signal than did the precision of $Z_K$ for the same datasets (Figure 2B).
+Further, for a given $n$, the variance in precision estimates of $Z_K$ was considerably smaller than the variance in precision estimates of $\lambda_{est}$ (Fig.
+2C,D); implying that estimates of phylogenetic signal were more reliable and robust when using $Z_K$ (for additional results see Supporting Information).
+Finally, it should also be recognized that because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable when using this measure, as $Z_K$ expresses the strength of phylogenetic signal in standard deviation units relative to the mean [see @AdamsCollyer2019].
+This further implies that comparisons of the strength of phylogenetic signal among phenotypic traits are possible, and may be accomplished statistically via a two-sample test that formally compares $Z_K$ across datasets [for comparisons of multivariate effect sizes see: @AdamsCollyer2016; @AdamsCollyer2019b].
 \hfill\break
 
-**NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K, Z-k-normalized.**
+[insert Figure 2 here] \hfill\break
 
-[insert Figure 2 here] \hfill\break
+When $\lambda$ was incorporated in PGLS regression and ANOVA (i.e., $Y\sim{X}$), we found much the same pattern as in the previous simulations.
+Namely, the precision of $\lambda_{est}$ varied inversely with species richness; where greater precision was obtained at higher levels of species richness (Figure 3).
+Likewise, the precision in estimating $\lambda$ was worse at intermediate values (Figure 3), and improved as the levels of phylogenetic signal were closer to $\lambda_{in}=0$ (no phylogenetic signal) or $\lambda_{in}=1$ (Brownian motion).
+Generally, the spread of estimates was slightly broader when $\lambda$  was co-estimated with regression parameters in PGLS regression (Figure 3), as compared to when $\lambda$ was estimated in cases only the dependent variable (Y) was considered (Figure 1).
+Finally, found that regression parameters ($\beta$) were accurately estimated when $\lambda$ was included during phylogenetic regression; a result which confirmed earlier findings of Revell [-@Revell2010] (see Supporting Information).
+\hfill\break
+
+[insert Figure 3 here]
 
@@
 
-[insert Figure 3 here] \hfill\break
+[insert Figure 4 here] \hfill\break
 

### 2020-04-21 Manuscript/Manuscript-1-DCA.tex
vsn 1 edits

@@
 describes the degree of phylogenetic signal in the dataset (Pagel 1999;
-Freckleton et al. 2002). Pagel's \(\lambda\) also has the advantage that
-it may be included when estimating the association of traits in a
+Freckleton et al. 2002). Pagel's \(\lambda\) also has the appeal that it
+may be included when estimating the association of traits in a
 phylogenetic context, meaning that one may account for the degree of
@@
 compared to a small one (\(N=13\)), concluding that insufficient data
-(i.e., the number of species) was the underlying cause of the lack of
-precision in parameter estimation. However, this conclusion assumes that
-the precision in estimating \(\lambda\) remains constant across its
+(i.e., the number of species) was the underlying cause of the increased
+variation across parameter estimates. However, this conclusion assumes
+that the precision in estimating \(\lambda\) remains constant across its
 range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
@@
 In this study, we evaluate the precision of Pagel's \(\lambda\) in
-estimating known levels of phylogenetic signal in phenotypic data. First
-we use computer simulations across differing numbers of species,
-differently shaped phylogenies, and differing input levels of
-phylogenetic signal, to explore the degree to which \(\lambda\)
-correctly identifies known levels of phylogenetic signal, and under what
-circumstances. For comparison, we also evaluate estimates of \(Kappa\)
-and its empirically-derived effect size (\(Z_K\)) for this purpose.
-Additionally, we use simulations to determine how the inclusion of
-\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
-parameter estimation, and whether levels of phylogenetic signal
-estimated in a PGLS framework are accurate. We then survey the recent
-macroevolutionary literature for published papers containing estimates
-of \(\lambda\) from empirical datasets, and compare these empirical
-estimates to patterns gleaned from our computer simulations. In general
-we find that while PGLS parameters (e.g., \(\beta\)) are accurately
-estimated with the inclusion of phylogenetic signal, estimates of
-\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
-widely for a given input value of phylogenetic signal, and that this
-variation is not constant across the range of input signal, with
-decreased precision when phylogenetic signal is of intermediate
-strength. By contrast, variation across effect sizes (\(Z_K\)) obtained
-from \(Kappa\) is far more consistent across the range of input values,
-and the precision of estimating \(Z_K\) is preferable to that observed
-using \(\lambda\). We conclude that estimates of phylogenetic signal
-using Pagel's \(\lambda\) are often innacurate, and thus interpreting
-strength of phylogenetic signal in phenotypic datasets based on this
-measure should be treated with caution. By contrast, effect sizes
-obtained from \(Kappa\) hold promise for characterizing phylogenetic
-signal, and comparing the strength of phylogenetic signal across
-datasets.
+estimating known levels of phylogenetic signal in phenotypic data. We
+use computer simulations with differing numbers of species, differently
+shaped phylogenies, and differing input levels of phylogenetic signal,
+to explore the degree to which \(\lambda\) correctly identifies known
+levels of phylogenetic signal, and under what circumstances. We also
+evaluate the precision of a standardized effect size (\(Z_K\)) based on
+\(Kappa\) for comparison. Additionally, we use simulations to determine
+how the inclusion of \(\lambda\) in phylogenetic regression and ANOVA
+(i.e., PGLS) affects parameter estimation, and whether levels of
+phylogenetic signal estimated in a PGLS framework are accurate. We then
+survey the recent macroevolutionary literature for published papers
+containing estimates of \(\lambda\) from empirical datasets, and compare
+these empirical estimates to patterns gleaned from our computer
+simulations. In general we find that while PGLS parameters (e.g.,
+\(\beta\)) are accurately estimated with the inclusion of phylogenetic
+signal, estimates of \(\lambda\) are not. We also find that estimates of
+\(\lambda\) vary widely for a given input value of phylogenetic signal,
+and that this variation is not constant across the range of input
+signal, with decreased precision when phylogenetic signal is of
+intermediate strength. In addition, the same \(\lambda_{est}\) may be
+obtained from datasets containing vastly different input levels of
+phylogenetic signal. By contrast, variation across effect sizes
+(\(Z_K\)) obtained from \(Kappa\) is far more consistent across the
+range of input values, and the precision of \(Z_K\) is generally
+preferable to that of \(\lambda_{est}\). Further, because \(Z_K\) is a
+standardized effect size, the strength of phylogenetic signal is more
+readily interpretable, and may be compared across datasets. We conclude
+that estimates of phylogenetic signal using Pagel's \(\lambda\) are
+often innacurate, and thus interpreting strength of phylogenetic signal
+in phenotypic datasets based on this measure is compromised. By
+contrast, effect sizes obtained from \(Kappa\) hold promise for
+characterizing phylogenetic signal, and for comparing the strength of
+phylogenetic signal across datasets.
 
@@
 phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison we
-estimated the degree of phylogenetic signal in each dataset using
-\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and from this calculated
-its effect size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each
-dataset. Here, the effect size was estimated as:
+characterized the strength of phylogenetic signal in each dataset using
+a standardized effect size (\(Z_K\): sensu Adams and Collyer 2016,
+2019a) based on \(Kappa\). As suggested by Adams and Collyer (2019b),
+the effect size for phylogenetic signal may be estimated as:
 \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
@@
 standard deviation of the empirical sampling distribution of values
-obtained from the permutation distribution (for discussion see: Adams
-and Collyer 2019b). Because \(Z_K\) differs in scale from \(\lambda\),
-we used a linear normalization to standardize \(Z_K\) to a uniform
-distribution (\(0\rightarrow1\)), and variation in the set of normalized
-\(Z_K\) values at each input level of phylogenetic signal
-(\(\lambda_{in}\)) were then estimated for comparison. \hfill\break
+obtained from the permutation distribution. \(Z_K\) describes the
+strength of phylogenetic signal as a standard deviate from its sampling
+distribution, and thus directly measures the strength of signal in a
+manner that is comparable across datasets. Variation in the set of
+\(Z_K\) at each input level of phylogenetic signal was then calculated
+as an estimate of precision in \(Z_K\). However, because \(Z_K\) differs
+in scale from \(\lambda\), we used a linear normalization to standardize
+\(Z_K\) to a uniform distribution (\(0\rightarrow1\)), and estimated the
+precision of \(Z_K\) from the normalized values. \hfill\break
 
@@
 manner, and whether such comparisons were accompanied with statistical
-tests between \(\lambda_{est}\). \hfill\break
+tests between \(\lambda_{est}\).
 
@@
 
-Our primary series of simulations revealed several patterns. First, the
-relationship between \(\lambda_{est}\) and input levels of phylogenetic
-signal was \(\sim\beta=1.0\) at each level of species richness
-(\(n=32, 64, 128, 256, 512, 1024\)). This implied that the mean
-\(\lambda_{est}\) across datasets for a given set of simulation
-conditions reflected input levels of phylogenetic signal. However, as
-shown in Figure 1, the precision of those estimates varied widely, and
-in several interesting ways. Predictably, precision was worst at low
-levels of species richness, where in many cases the set of
-\(\lambda_{est}\) spanned nearly the entire range of possible values
-(e.g., \(n=32\); \(\lambda_{est}=0.5\): range = \(0.0\to0.985\)).
-Second, as species richness increased, variation across estimates of
-\(\lambda\) decreased (Figure 1). This confirmed the pattern identified
-by Boettiger et al.~(2012) on a small (\(n=13\)) and large (\(n=281\))
-phylogeny, and demonstrated that this trend was general across a wider
-range of species richness. Importantly however, our broader set of
+Our first set of simulations revealed several patterns. First, the
+relationship between \(\lambda_{est}\) and \(\lambda_{in}\) across the
+range input levels of phylogenetic signal was \(\sim\beta=1.0\);
+indicating that the average \(\lambda_{est}\) across datasets for a
+given set of simulation condition correctly reflected input levels of
+phylogenetic signal. However, as shown in Figure 1, the precision of
+those estimates varied widely, and in several interesting ways.
+Predictably, precision was worse at low levels of species richness,
+where in many cases the set of \(\lambda_{est}\) spanned nearly the
+entire range of possible values (e.g., \(n=32\); \(\lambda_{in}=0.5\):
+range of \(\lambda_{est}= 0.0\to 0.985\)). Second, as species richness
+increased, variation across estimates of \(\lambda\) decreased (Figure
+1). This confirmed the pattern identified by Boettiger et al.~(2012) on
+a small (\(n=13\)) versus a large (\(n=281\)) phylogeny, and
+demonstrated that the trend of increasing precision with higher species
+richness was general. Importantly however, our broader set of
 simulations revealed that the precision of \(\lambda_{est}\) was not
-uniform across all levels of phylogenetic signal, and that there was a
-particular pattern to the imprecision of estimating this phylogenetic
-parameter. Specifically, we found that variation in \(\lambda_{est}\)
-was highest at intermediate levels of phylogenetic signal, and decreased
-at both lower and higher levels of input signal (Figure 1). This implied
-that the precision in estimating \(\lambda\) was worst at intermediate
-levels of phylogenetic signal; a pattern that remained consistent across
-all levels of species richness evaluated here (Figure 1). Notably, even
-at very large levels of species richness, the range of \(\lambda_{est}\)
+uniform across all levels of phylogenetic signal. Specifically, we found
+that variation in \(\lambda_{est}\) was highest at intermediate levels
+of phylogenetic signal, and decreased at both lower and higher levels of
+input signal (Figure 1). This implied that the precision in estimating
+\(\lambda\) was worse at intermediate values, and improved as the levels
+of phylogenetic signal were closer to \(\lambda_{in}=0\) (no
+phylogenetic signal) or \(\lambda_{in}=1\) (Brownian motion). Notably,
+even at large levels of species richness, the range of \(\lambda_{est}\)
 still encompassed a substantial portion of possible values (e.g.,
-\(n=1024\); \(\lambda_{est}=0.5\): range = \(0.31\to 0.68\)). From this
-we conclude that \(\lambda_{est}\) is not an accurate estimator of the
-strength of phylogenetic signal in phenotypic datasets, and that
-biological interpretations of the level of phylogenetic signal based on
-\(\lambda\) should be treated with caution. \hfill\break
+\(n=512\); \(\lambda_{in}=0.5\): range of
+\(\lambda_{est} = 0.32\to 0.68\)). Likewise, the same \(\lambda_{est}\)
+could be obtained from datasets containing vastly different input levels
+of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\); range
+of \(\lambda_{in} = 0.25 \to 0.65\)). Taken together, these findings
+reveal that \(\lambda_{est}\) does not precisely characterize observed
+levels of phylogenetic signal in phenotypic datasets, and that
+biological interpretations of the strength of phylogenetic signal based
+on \(\lambda\) may be highly inaccurate. \hfill\break
 
@@
 
-By contrast, (FIg 2) Kappa. Much smaller range and variation as compared
-with lambda, but with larger l-in, variation in kappa does increase.
-However, this is not the case with Z-K. Z-K increases predictably
-(linearly, progressively\ldots) with L-in, and importantly variation in
-Z-K remains nearly constant across the range of input values. this means
-that Z-K is equally precise in estimating the strength of phylogenetic
-signal, and may be more adept at facilitating useful statements on teh
-relative strength of phylogenetic signal within and across datasets.
-\hfill\break
-
-\textbf{NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K,
-Z-k-normalized.}
+By contrast, when characterizing phylogenetic signal with the
+standardized effect size (\(Z_K\)) of \(Kappa\), we found that the
+precision of \(Z_K\) was more stable, as variation across datasets was
+far more consistent across the range of input values. For example, when
+\(n=128\) the precision of \(\lambda_{est}\) (Figure 2A) varied
+considerably more across input levels of phylogenetic signal than did
+the precision of \(Z_K\) for the same datasets (Figure 2B). Further, for
+a given \(n\), the variance in precision estimates of \(Z_K\) was
+considerably smaller than the variance in precision estimates of
+\(\lambda_{est}\) (Fig. 2C,D); implying that estimates of phylogenetic
+signal were more reliable and robust when using \(Z_K\) (for additional
+results see Supporting Information). Finally, it should also be
+recognized that because \(Z_K\) is a standardized effect size, the
+strength of phylogenetic signal is more readily interpretable when using
+this measure, as \(Z_K\) expresses the strength of phylogenetic signal
+in standard deviation units relative to the mean (see Adams and Collyer
+2019b). This further implies that comparisons of the strength of
+phylogenetic signal among phenotypic traits are possible, and may be
+accomplished statistically via a two-sample test that formally compares
+\(Z_K\) across datasets (for comparisons of multivariate effect sizes
+see: Adams and Collyer 2016, 2019a). \hfill\break
 
 {[}insert Figure 2 here{]} \hfill\break
+
+When \(\lambda\) was incorporated in PGLS regression and ANOVA (i.e.,
+\(Y\sim{X}\)), we found much the same pattern as in the previous
+simulations. Namely, the precision of \(\lambda_{est}\) varied inversely
+with species richness; where greater precision was obtained at higher
+levels of species richness (Figure 3). Likewise, the precision in
+estimating \(\lambda\) was worse at intermediate values (Figure 3), and
+improved as the levels of phylogenetic signal were closer to
+\(\lambda_{in}=0\) (no phylogenetic signal) or \(\lambda_{in}=1\)
+(Brownian motion). Generally, the spread of estimates was slightly
+broader when \(\lambda\) was co-estimated with regression parameters in
+PGLS regression (Figure 3), as compared to when \(\lambda\) was
+estimated in cases only the dependent variable (Y) was considered
+(Figure 1). Finally, found that regression parameters (\(\beta\)) were
+accurately estimated when \(\lambda\) was included during phylogenetic
+regression; a result which confirmed earlier findings of Revell (2010)
+(see Supporting Information). \hfill\break
+
+{[}insert Figure 3 here{]}
 
@@
 
-{[}insert Figure 3 here{]} \hfill\break
+{[}insert Figure 4 here{]} \hfill\break
 

### 2020-04-21 Manuscript/Manuscript-1-DCA.Rmd
vsn 1

@@
 Several analytical tools have been developed for this purpose [e.g., @Gittleman1990; @Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Klingenberg2010; @Adams2014a], which differ primarily in how they characterize the phylogenetic dependency of trait variation among taxa.
-One commonly used statistical measure, *Kappa* ($K$), expresses the strength of phylogenetic signal as the ratio of observed trait variation to the trait variation conditioned on the phylogeny; scaled by what is expected under Brownian motion given the phylogeny's size and shape [@Blomberg_et_al2003; also @Adams2014a].
+One commonly used statistical measure, *Kappa* ($K$), expresses the strength of phylogenetic signal as the ratio of observed trait variation to the trait variation conditioned on the phylogeny; which is then scaled by what is expected under Brownian motion given the phylogeny's size and shape [@Blomberg_et_al2003; also @Adams2014a].
 Another approach, Pagel's $\lambda$ [@Pagel1999], uses maximum likelihood to fit the data to the phylogeny under some model of evolutionary change (typically Brownian motion).
@@
 However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012].
+Indeed, understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates.
@@
 We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
-In addition, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
+Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
 By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and the precision of $Z_K$ is generally preferable to that of $\lambda_{est}$.
 Further, because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable, and may be compared across datasets.
-We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
+We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
 By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
@@
 Finally, for comparison we characterized the strength of phylogenetic signal in each dataset using a standardized effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] based on $Kappa$.
-As suggested by Adams and Collyer [-@AdamsCollyer2019], the effect size for phylogenetic signal may be estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution.
+As suggested by Adams and Collyer [-@AdamsCollyer2019], an effect size for phylogenetic signal may be estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution.
 $Z_K$ describes the strength of phylogenetic signal as a standard deviate from its sampling distribution, and thus directly measures the strength of signal in a manner that is comparable across datasets.
@@
 Our first set of simulations revealed several patterns.
-First, the relationship between $\lambda_{est}$ and $\lambda_{in}$ across the range input levels of phylogenetic signal was $\sim\beta=1.0$; indicating that the average $\lambda_{est}$ across datasets for a given set of simulation condition correctly reflected input levels of phylogenetic signal.
+First, the relationship between $\lambda_{est}$ and $\lambda_{in}$ across the range input levels of phylogenetic signal was $\beta\sim1.0$; indicating that the average $\lambda_{est}$ across datasets for a given simulation condition correctly reflected the input levels of phylogenetic signal.
 However, as shown in Figure 1, the precision of those estimates varied widely, and in several interesting ways.
@@
 This confirmed the pattern identified by Boettiger et al.
-[-@Boettiger_et_al2012] on  a small ($n=13$) versus a large ($n=281$) phylogeny, and demonstrated that the trend of increasing precision with higher species richness was general.
+[-@Boettiger_et_al2012] on  a small ($n=13$) versus a large ($n=281$) phylogeny, and demonstrated that the trend of increasing precision with higher species richness was general; adhering to parametric statistical theory.
 Importantly however, our broader set of simulations revealed that the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal.
@@
 
-When $\lambda$ was incorporated in PGLS regression and ANOVA (i.e., $Y\sim{X}$), we found much the same pattern as in the previous simulations.
-Namely, the precision of $\lambda_{est}$ varied inversely with species richness; where greater precision was obtained at higher levels of species richness (Figure 3).
+When $\lambda$ was incorporated in PGLS regression and ANOVA (i.e., $Y\sim{X}$), we found much the same pattern as in our earlier simulations.
+Namely, the precision of $\lambda_{est}$ covaried with species richness; where greater precision was obtained at higher levels of species richness (Figure 3).
 Likewise, the precision in estimating $\lambda$ was worse at intermediate values (Figure 3), and improved as the levels of phylogenetic signal were closer to $\lambda_{in}=0$ (no phylogenetic signal) or $\lambda_{in}=1$ (Brownian motion).
-Generally, the spread of estimates was slightly broader when $\lambda$  was co-estimated with regression parameters in PGLS regression (Figure 3), as compared to when $\lambda$ was estimated in cases only the dependent variable (Y) was considered (Figure 1).
-Finally, found that regression parameters ($\beta$) were accurately estimated when $\lambda$ was included during phylogenetic regression; a result which confirmed earlier findings of Revell [-@Revell2010] (see Supporting Information).
+Generally, the spread of estimates was slightly broader when $\lambda$  was co-estimated with regression parameters in PGLS regression (Figure 3), as compared to when $\lambda$ was estimated for only the dependent variable (Figure 1).
+Finally, we found that regression parameters ($\beta$) were accurately estimated when $\lambda$ was included during phylogenetic regression; a result which confirmed earlier findings of Revell [-@Revell2010] (see Supporting Information).
 \hfill\break

### 2020-04-23 Data_Analyses/Analysis_script.R
More lit review comparison, looking at lambdas between .25 and .75

@@ -229,22 +229,28 @@ length(unique(review$Reference)) # 182
 published.lambdas <- review[,c(1,3,4)]
 published.lambdas.pruned <- published.lambdas[-which(published.lambdas$`Estimated Lambdas Simple` > 1 | published.lambdas$`Estimated Lambdas Simple` < 0),]
 published.lambdas.df <- as.data.frame(published.lambdas.pruned)
-nrow(published.lambdas.df) # 1552
-1552/182 # average number of reported lambda values per manuscript
+nrow(published.lambdas.df) # 1556
+1556/182 # average number of reported lambda values per manuscript = 8.549451
 colnames(published.lambdas.df) <- c("Refs", "Published_Lambdas", "Taxa")
 published.lambdas.df$Refs <- as.factor(published.lambdas.df$Refs)
 
 min(table(published.lambdas.df$Refs)) # 1: lowest number of reported lambdas per paper
 max(table(published.lambdas.df$Refs)) # 71: highest number of reported lambdas per paper
 
-length(which(published.lambdas.df$Published_Lambdas < 0.05))/nrow(published.lambdas.df)*100 # 25.43%
-length(which(published.lambdas.df$Published_Lambdas > 0.90))/nrow(published.lambdas.df)*100 # 24.74%
+length(which(published.lambdas.df$Published_Lambdas < 0.05))/nrow(published.lambdas.df)*100 # 25.32%
+length(which(published.lambdas.df$Published_Lambdas > 0.90))/nrow(published.lambdas.df)*100 # 24.94%
 
-length(which(published.lambdas.df$Taxa < 30))/nrow(published.lambdas.df)*100 # 75.32%
+length(which(published.lambdas.df$Taxa < 30))/nrow(published.lambdas.df)*100 # 22.36504%
 length(which(published.lambdas.df$Taxa < 30)) # 348
 
+review_25to75 <- review[which(review$`Estimated Lambdas Simple` >=.25 & review$`Estimated Lambdas Simple` < .7501),]
+colnames(review_25to75) <- c("Reference", "Estimated Lambda", "Estimated Lambdas Simple", "Taxa", "__1", "Interpreted", "Quote", "Extra")
 
+review_25to75[which(review_25to75$Interpreted == "Kinda"),6] <- "Yes"
+review_25to75$Interpreted <- as.factor(review_25to75$Interpreted)
 
+length(unique(review_25to75$Reference)) # 108 manuscripts published lambdas between .25 and .75
[…]

### 2020-04-23 Data_Analyses/Simulation_script-PB-KappaComparison.R

@@ -16,7 +16,7 @@ lambdas <- seq(0,1,.05) # starting with just 3 lambda values
 lambdas_expanded <- rep(lambdas, each = nsim)
 fullsimlength <- length(lambdas_expanded)
 
-n <- treesizes[1] # number of tips: adjust each iteration
+n <- treesizes[5] # number of tips: adjust each iteration
 
 # New Rep #####
 DataTable_lambda <- data.frame(lambda.input = lambdas_expanded, 
@@ -26,12 +26,11 @@ DataTable_lambda <- data.frame(lambda.input = lambdas_expanded,
                                CI.lower = rep(NA, nsim*length(lambdas)),
                                CI.upper = rep(NA, nsim*length(lambdas))) 
 
-pb <- txtProgressBar(1, fullsimlength, style=1)
+pb <- txtProgressBar(499, fullsimlength, style=1)
 
 Start.time<-Sys.time()
 Start.time
-for (j in 1:fullsimlength) {
-
+for (j in 499:fullsimlength) {
   while(TRUE){
  
     tree<-pbtree(n=n,scale=1) 
@@ -55,24 +54,26 @@ for (j in 1:fullsimlength) {
 
 DataTable_lambda[j,2] <- pgls_est$param[2]
 DataTable_lambda[j,3] <- kappa_est$phy.signal 
+DataTable_lambda[j,4] <- kappa_est$Z
 
-
-kappa_sum <- capture.output(summary(kappa_est)) # this capture.output was necessary to silence the print function that cluttered my console. not analytically necessary.
-text_z <- kappa_sum[11]
-
-while(TRUE){
-placeholder <- try(as.numeric(gsub("[^0-9.-]", "", text_z)),silent=T) # gettin rid of text
-if(!is(placeholder, 'try-error')) break}
-
-DataTable_lambda[j,4] <- placeholder
-
-DataTable_lambda[j,5] <- pgls_est$djdjdj # confidence intervals....
-DataTable_lambda[j,6] <- pgls_est$djdjdj # confidence intervals....
+DataTable_lambda[j,5] <- pgls_est$param.CI$lambda$ci.val[1] # lower CI
+DataTable_lambda[j,6] <- pgls_est$param.CI$lambda$ci.val[2] # upper CI
 
 setTxtProgressBar(pb, j)
 } 
 Sys.time()-Start.time
 
[…]

### 2020-04-23 Manuscript/Manuscript-2.Rmd
vsn 2

--- 2020-04-21 Manuscript/Manuscript-1-DCA.Rmd
@@
 ---
-# Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate
+# A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate
 
@@
 
-**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break
+**Keywords**: phylogenetic signal, effect size, Pagel's lambda \hfill\break
 
-**Short Title**: Inaccuracies in Pagel's Lambda \hfill\break
+**Short Title**: An Effect Size for Phylogenetic Signal \hfill\break
 
@@
 
-For many macroevolutionary analyses, it is often of interest to quantify the degree to which phylogenetic signal is displayed in continuous traits.
-Several analytical tools have been developed for this purpose [e.g., @Gittleman1990; @Abouheif1999; @Pagel1999; @Blomberg_et_al2003; @Klingenberg2010; @Adams2014a], which differ primarily in how they characterize the phylogenetic dependency of trait variation among taxa.
-One commonly used statistical measure, *Kappa* ($K$), expresses the strength of phylogenetic signal as the ratio of observed trait variation to the trait variation conditioned on the phylogeny; which is then scaled by what is expected under Brownian motion given the phylogeny's size and shape [@Blomberg_et_al2003; also @Adams2014a].
-Another approach, Pagel's $\lambda$ [@Pagel1999], uses maximum likelihood to fit the data to the phylogeny under some model of evolutionary change (typically Brownian motion).
-The inclusion of a scaling parameter, $\lambda$,  transforms the lengths of the internal branches of the phylogeny to improve the fit, and this parameter describes the degree of phylogenetic signal in the dataset [@Pagel1999; @Freckleton_et_al2002].
-Pagel's $\lambda$ also has the appeal that it may be included when estimating the association of traits in a phylogenetic context, meaning that one may account for the degree of phylogenetic signal while conducting phylogenetic regression or ANOVA [see @Freckleton_et_al2002].
+Several analytical tools have been developed to quantify phylogenetic signal in phenotypic datasets, including measures of serial independence [$\mathbf{C}$: @Abouheif1999], autocorrelation estimates [$I$: @Gittleman1990], statistical ratios of trait variation relative to what is expected given the phylogeny [$Kappa$: @Blomberg_et_al2003; @Adams2014a], and scaling parameters used in maximum likelihood fitting of the data to the phylogeny [$\lambda$: @Pagel1999], among others [e.g., @Klingenberg2010].
+The statistical properties of these methods -- namely type I error rates and power -- have also been investigated to determine when phylogenetic signal can be detected and under what conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @Adams2014a; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010].
+One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999].
+Here, maximum likelihood is used to fit the data to the phylogeny under a Brownian motion model of evolution.
+A parameter ($\lambda$) is included, which transforms the lengths of the internal branches of the phylogeny to improve the fit [@Pagel1999; @Freckleton_et_al2002].
+Pagel's $\lambda$, which ranges from $0\to1$, is then treated as a measure of phylogenetic signal; with values close to 0 signifying little phylogenetic signal, and values close to 1 indicating that trait variation accumulates according to Brownian motion.
+Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic regression  (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
 \hfill\break
 
-Several studies have investigated the statistical properties of methods for estimating phylogenetic signal under various conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010].
-These have largely focused on the ability of methods to detect the presence of phylogenetic signal (i.e., type I and type II error rates) under complex models of evolutionary change, across a range of phylogeny sizes, and with varying degrees of phylogenetic uncertainty or unresolved topologies.
-In terms of parameter estimation, Revell [-@Revell2010] found that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al.
-[-@Munkemuller_et_al2012] demonstrated that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger.
-However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012].
-Indeed, understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
-One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates.
-However, this conclusion assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
+Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
+This is often accomplished by interpreting empirical estimates of $\lambda$; with $\lambda\approx0$ signifying 'weak' phylogenetic signal, and $\lambda\approx1$ being interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
+Other approaches for interpreting $\lambda$ are more statistical, through the use of confidence intervals [@Vandelook2019] or likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
+Likewise, qualitative comparisons of $\lambda$ across multiple phenotypic traits have also been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are well known ($\lambda=0$ is no phylogenetic signal, while $\lambda=1$ corresponds to Brownian motion).
+However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully examined.
+\hfill\break
+
+First, it presumes that values of $\lambda$ can be precisely estimated, and that parameter estimates of $\lambda$ are representative of the underlying phylogenetic signal.
+Understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
+One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
+From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
+Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise with greater sample sizes [@Cohen1988].
+However, this conclusion also assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
 Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
+\hfill\break
+
+Second, while $\lambda$ is observed on a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data.
+For this to be the case, $\lambda$ must be a statistical effect size.
+Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
+Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets [e.g., meta-analysis: @HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared across datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
+Unfortunatley, not all statistical parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
+As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
+To date, it has not been determined whether $\lambda$ is an effect size for phylogenetic signal.
 \hfill\break
@@
 We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
-We also evaluate the precision of a standardized effect size ($Z_K$) based on $Kappa$ for comparison.
-Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic regression and ANOVA (i.e., PGLS) affects parameter estimation, and whether levels of phylogenetic signal estimated in a PGLS framework are accurate.
-We then survey the recent macroevolutionary literature for published papers containing estimates of $\lambda$ from empirical datasets, and compare these empirical estimates to patterns gleaned from our computer simulations.
+We then derive a standardized effect size for phylogenetic signal, and apply the concept to two common measures of phylogenetic signal: $\lambda$ ($Z_{\lambda}$) and $Kappa$ ($Z_K$).
 In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
@@
 Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
-By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and the precision of $Z_K$ is generally preferable to that of $\lambda_{est}$.
+Likewise, the precision of $Z_{\lambda}$ is less consistent than that observed with $Z_K$, across the range of input values, implying that $Z_K$ is a more robust effect size measure.
 Further, because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable, and may be compared across datasets.
@@
 By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
+
+# Methods and Results
+
+-Lambda is not precise: holds in estimates, and PGLSregression
+  -Precision varies with N (duh) but also WITH INPUT!
+  -same input on different N, different lambda.est
+  -same lambda.est from different input
+  -Boettiger example: add small tips, greatly change lambda...
+-Conclusion: Lambda does not quantify strength of phylogenetic signal as is often assumed
+
+-Empirical examples: how interpretation can get folks ito trouble.
+
+-A proposed Effect size for the strength of physig
+  -Derivation of effect size (expand on Adams and Collyer 2019 suggestion that an effect size for phylogenetic signal is needed.
+The ‘gold standard’ of these is a standaredized effect size: (following the meta-analytic literature beginning with Glass 1976).
+In other words, a Z-score.
+So, how does Z work when based on lambda?
+On kappa?
+  - Z-score. could be Lambda or Kappa.  Show it is Kappa
+-Comparing the strength of physig
+  - two sample Z-score
+
+-Conclusions and Implications
+
+## $\lambda$ is not Precise
+
+## Lambda does not measure phylogenetic signal
+
+## An effect size for Physig
 

### 2020-04-23 Manuscript/Manuscript-2.tex
vsn 2

--- 2020-04-21 Manuscript/Manuscript-1-DCA.tex
@@
 
-\hypertarget{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}{%
-\section{Estimates of Phylogenetic Signal Based on Lambda are Often
-Inaccurate}\label{estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}}
+\hypertarget{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}{%
+\section{A Standardized Effect Size for Evaluating the Strength of
+Phylogenetic Signal, and Why Lambda is not
+Appropriate}\label{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}}
 
@@
 
-\textbf{Keywords}: Pagel's lambda, phylogenetic signal \hfill\break
+\textbf{Keywords}: phylogenetic signal, effect size, Pagel's lambda
+\hfill\break
 
-\textbf{Short Title}: Inaccuracies in Pagel's Lambda \hfill\break
+\textbf{Short Title}: An Effect Size for Phylogenetic Signal
+\hfill\break
 
@@
 
-For many macroevolutionary analyses, it is often of interest to quantify
-the degree to which phylogenetic signal is displayed in continuous
-traits. Several analytical tools have been developed for this purpose
-(e.g., Gittleman and Kot 1990; Abouheif 1999; Pagel 1999; Blomberg et
-al. 2003; Klingenberg and Gidaszewski 2010; Adams 2014a), which differ
-primarily in how they characterize the phylogenetic dependency of trait
-variation among taxa. One commonly used statistical measure,
-\emph{Kappa} (\(K\)), expresses the strength of phylogenetic signal as
-the ratio of observed trait variation to the trait variation conditioned
-on the phylogeny; scaled by what is expected under Brownian motion given
-the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
-Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
-likelihood to fit the data to the phylogeny under some model of
-evolutionary change (typically Brownian motion). The inclusion of a
-scaling parameter, \(\lambda\), transforms the lengths of the internal
-branches of the phylogeny to improve the fit, and this parameter
-describes the degree of phylogenetic signal in the dataset (Pagel 1999;
-Freckleton et al. 2002). Pagel's \(\lambda\) also has the appeal that it
-may be included when estimating the association of traits in a
-phylogenetic context, meaning that one may account for the degree of
-phylogenetic signal while conducting phylogenetic regression or ANOVA
-(see Freckleton et al. 2002). \hfill\break
+Several analytical tools have been developed to quantify phylogenetic
+signal in phenotypic datasets, including measures of serial independence
+(\(\mathbf{C}\): Abouheif 1999), autocorrelation estimates (\(I\):
+Gittleman and Kot 1990), statistical ratios of trait variation relative
+to what is expected given the phylogeny (\(Kappa\): Blomberg et al.
+2003; Adams 2014a), and scaling parameters used in maximum likelihood
+fitting of the data to the phylogeny (\(\lambda\): Pagel 1999), among
+others (e.g., Klingenberg and Gidaszewski 2010). The statistical
+properties of these methods -- namely type I error rates and power --
+have also been investigated to determine when phylogenetic signal can be
+detected and under what conditions (e.g., Munkemuller et al. 2012;
+Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
+Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008; Revell
+2010). One of the most widely used methods for characterizing
+phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
+(Pagel 1999). Here, maximum likelihood is used to fit the data to the
+phylogeny under a Brownian motion model of evolution. A parameter
+(\(\lambda\)) is included, which transforms the lengths of the internal
+branches of the phylogeny to improve the fit (Pagel 1999; Freckleton et
+al. 2002). Pagel's \(\lambda\), which ranges from \(0\to1\), is then
+treated as a measure of phylogenetic signal; with values close to 0
+signifying little phylogenetic signal, and values close to 1 indicating
+that trait variation accumulates according to Brownian motion. Pagel's
+\(\lambda\) also has the appeal that it may be included in phylogenetic
+regression (PGLS) to account for the degree of phylogenetic signal in
+comparative analyses (see Freckleton et al. 2002). \hfill\break
 
-Several studies have investigated the statistical properties of methods
-for estimating phylogenetic signal under various conditions (e.g.,
-Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
-2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
-Revell 2010). These have largely focused on the ability of methods to
-detect the presence of phylogenetic signal (i.e., type I and type II
-error rates) under complex models of evolutionary change, across a range
-of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
-or unresolved topologies. In terms of parameter estimation, Revell
-(2010) found that regression parameters were accurately estimated when
-\(\lambda\) was included during phylogenetic regression, and Munkmuller
-et al.~(2012) demonstrated that estimates of phylogenetic signal
-obtained using various measures generally increased when input levels of
-phylogenetic signal were stronger. However, the precision of those
-estimates could not be determined, because the input levels of
-phylogenetic signal were simulated via a scaling factor (\(w\)) that was
-not directly comparable to the measures of phylogenetic signal being
-compared (see Munkemuller et al. 2012). One study (Boettiger et al.
-2012) found that estimates of Pagel's \(\lambda\) displayed less
-variation when data were simulated on a large phylogeny (\(N=281\)) as
-compared to a small one (\(N=13\)), concluding that insufficient data
-(i.e., the number of species) was the underlying cause of the increased
-variation across parameter estimates. However, this conclusion assumes
-that the precision in estimating \(\lambda\) remains constant across its
-range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+Evolutionary biologists commonly seek to describe the relative strength
+of phylogenetic signal in phenotypic traits, to determine the extent to
+which shared evolutionary history has influenced trait covariation among
+taxa. This is often accomplished by interpreting empirical estimates of
+\(\lambda\); with \(\lambda\approx0\) signifying `weak' phylogenetic
+signal, and \(\lambda\approx1\) being interpreted as `strong'
+phylogenetic signal (e.g., De Meester et al. 2019; Pintanel et al. 2019;
+Su et al. 2019). Other approaches for interpreting \(\lambda\) are more
+statistical, through the use of confidence intervals (Vandelook et al.
+2019) or likelihood ratio tests that compare the observed model fit to
+that obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
+2002; Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative
+comparisons of \(\lambda\) across multiple phenotypic traits have also
+been used to infer whether the strength of phylogenetic signal is
+greater in one trait as compared to another (e.g., Liu et al. 2019; Bai
+et al. 2019). Indeed, it seems intuitive to interpret the strength of
+phylogenetic signal in this manner, as \(\lambda\) is a parameter on a
+bounded scale (\(0\to1\)) for which interpretation of its extremal
+points are well known (\(\lambda=0\) is no phylogenetic signal, while
+\(\lambda=1\) corresponds to Brownian motion). However, equating values
+of \(\lambda\) directly to the strength of phylogenetic signal presumes
+two important statistical properties that have not been fully examined.
+\hfill\break
+
+First, it presumes that values of \(\lambda\) can be precisely
+estimated, and that parameter estimates of \(\lambda\) are
+representative of the underlying phylogenetic signal. Understanding the
+precision in estimating levels of phylogenetic signal is paramount,
+because biological inferences regarding the strength of phylogenetic
+signal in phenotypic datasets requires high precision of its estimation.
+One study (Boettiger et al. 2012) found that estimates of Pagel's
+\(\lambda\) displayed less variation (i.e., greater precision) when data
+were simulated on a large phylogeny (\(N=281\)) as compared to a small
+one (\(N=13\)). From this observation it was concluded that insufficient
+data (i.e., the number of species) was the underlying cause of the
+increased variation across parameter estimates (Boettiger et al. 2012).
+Indeed, such a pattern is common with statistical estimators, as summary
+statistics and parameters are often more precise with greater sample
+sizes (Cohen 1988). However, this conclusion also assumes that the
+precision in estimating \(\lambda\) remains constant across its range
+(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
 verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
@@
 levels of phylogenetic signal in phenotypic datasets. \hfill\break
+
+Second, while \(\lambda\) is observed on a bounded scale (\(0\to1\)),
+this does not \emph{de-facto} imply that the estimated values of this
+parameter correspond to the actual strength of the underlying input
+signal in the data. For this to be the case, \(\lambda\) must be a
+statistical effect size. Effect sizes are a measure the magnitude of a
+statistical effect in data, represented on a common scale (Glass 1976;
+Cohen 1988). Effect sizes have widespread use in many areas of the
+quantiative sciences, as they represent measures that may be readily
+summarized across datasets (e.g., meta-analysis: Hedges and Olkin 1985;
+Glass 1976; Arnqvist and Wooster 1995), or compared across datasets
+(e.g., Adams and Collyer 2016, 2019a). Unfortunatley, not all
+statistical parameters and test statistics are effect sizes, and thus
+many summary measures must first be converted to standardized units
+(i.e., an effect size) for meaningful comparison (see Rosenthal 1994).
+As a consequence, it follows that only if \(\lambda\) is a statistical
+effect size can comparisons of estimates across datasets be
+interpretable. To date, it has not been determined whether \(\lambda\)
+is an effect size for phylogenetic signal. \hfill\break
 
@@
 to explore the degree to which \(\lambda\) correctly identifies known
-levels of phylogenetic signal, and under what circumstances. We also
-evaluate the precision of a standardized effect size (\(Z_K\)) based on
-\(Kappa\) for comparison. Additionally, we use simulations to determine
-how the inclusion of \(\lambda\) in phylogenetic regression and ANOVA
-(i.e., PGLS) affects parameter estimation, and whether levels of
-phylogenetic signal estimated in a PGLS framework are accurate. We then
-survey the recent macroevolutionary literature for published papers
-containing estimates of \(\lambda\) from empirical datasets, and compare
-these empirical estimates to patterns gleaned from our computer
-simulations. In general we find that while PGLS parameters (e.g.,
-\(\beta\)) are accurately estimated with the inclusion of phylogenetic
-signal, estimates of \(\lambda\) are not. We also find that estimates of
-\(\lambda\) vary widely for a given input value of phylogenetic signal,
-and that this variation is not constant across the range of input
-signal, with decreased precision when phylogenetic signal is of
-intermediate strength. In addition, the same \(\lambda_{est}\) may be
-obtained from datasets containing vastly different input levels of
-phylogenetic signal. By contrast, variation across effect sizes
-(\(Z_K\)) obtained from \(Kappa\) is far more consistent across the
-range of input values, and the precision of \(Z_K\) is generally
-preferable to that of \(\lambda_{est}\). Further, because \(Z_K\) is a
-standardized effect size, the strength of phylogenetic signal is more
-readily interpretable, and may be compared across datasets. We conclude
-that estimates of phylogenetic signal using Pagel's \(\lambda\) are
-often innacurate, and thus interpreting strength of phylogenetic signal
-in phenotypic datasets based on this measure is compromised. By
-contrast, effect sizes obtained from \(Kappa\) hold promise for
-characterizing phylogenetic signal, and for comparing the strength of
-phylogenetic signal across datasets.
+levels of phylogenetic signal, and under what circumstances. We then
+derive a standardized effect size for phylogenetic signal, and apply the
+concept to two common measures of phylogenetic signal: \(\lambda\)
+(\(Z_{\lambda}\)) and \(Kappa\) (\(Z_K\)). In general we find that while
+PGLS parameters (e.g., \(\beta\)) are accurately estimated with the
+inclusion of phylogenetic signal, estimates of \(\lambda\) are not. We
+also find that estimates of \(\lambda\) vary widely for a given input
+value of phylogenetic signal, and that this variation is not constant
+across the range of input signal, with decreased precision when
+phylogenetic signal is of intermediate strength. Additionally, the same
+\(\lambda_{est}\) may be obtained from datasets containing vastly
+different input levels of phylogenetic signal. Likewise, the precision
+of \(Z_{\lambda}\) is less consistent than that observed with \(Z_K\),
+across the range of input values, implying that \(Z_K\) is a more robust
+effect size measure. Further, because \(Z_K\) is a standardized effect
+size, the strength of phylogenetic signal is more readily interpretable,
+and may be compared across datasets. We conclude that estimates of
+phylogenetic signal using Pagel's \(\lambda\) are often inaccurate, and
+thus interpreting strength of phylogenetic signal in phenotypic datasets
+based on this measure is compromised. By contrast, effect sizes obtained
+from \(Kappa\) hold promise for characterizing phylogenetic signal, and
+for comparing the strength of phylogenetic signal across datasets.
+
+\hypertarget{methods-and-results}{%
+\section{Methods and Results}\label{methods-and-results}}
+
+-Lambda is not precise: holds in estimates, and PGLSregression
+-Precision varies with N (duh) but also WITH INPUT! -same input on
+different N, different lambda.est -same lambda.est from different input
+-Boettiger example: add small tips, greatly change lambda\ldots{}
+-Conclusion: Lambda does not quantify strength of phylogenetic signal as
+is often assumed
+
+-Empirical examples: how interpretation can get folks ito trouble.
+
+-A proposed Effect size for the strength of physig -Derivation of effect
+size (expand on Adams and Collyer 2019 suggestion that an effect size
+for phylogenetic signal is needed. The `gold standard' of these is a
+standaredized effect size: (following the meta-analytic literature
+beginning with Glass 1976). In other words, a Z-score. So, how does Z
+work when based on lambda? On kappa? - Z-score. could be Lambda or
+Kappa. Show it is Kappa -Comparing the strength of physig - two sample
+Z-score
+
+-Conclusions and Implications
+
+\hypertarget{lambda-is-not-precise}{%
+\subsection{\texorpdfstring{\(\lambda\) is not
+Precise}{\textbackslash lambda is not Precise}}\label{lambda-is-not-precise}}
+
+\hypertarget{lambda-does-not-measure-phylogenetic-signal}{%
+\subsection{Lambda does not measure phylogenetic
+signal}\label{lambda-does-not-measure-phylogenetic-signal}}
+
+\hypertarget{an-effect-size-for-physig}{%
+\subsection{An effect size for Physig}\label{an-effect-size-for-physig}}
 
@@
 a standardized effect size (\(Z_K\): sensu Adams and Collyer 2016,
-2019a) based on \(Kappa\). As suggested by Adams and Collyer (2019b),
-the effect size for phylogenetic signal may be estimated as:
+2019a) based on \(Kappa\). As suggested by Adams and Collyer (2019b), an
+effect size for phylogenetic signal may be estimated as:
 \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
@@
 relationship between \(\lambda_{est}\) and \(\lambda_{in}\) across the
-range input levels of phylogenetic signal was \(\sim\beta=1.0\);
+range input levels of phylogenetic signal was \(\beta\sim1.0\);
 indicating that the average \(\lambda_{est}\) across datasets for a
-given set of simulation condition correctly reflected input levels of
+given simulation condition correctly reflected the input levels of
 phylogenetic signal. However, as shown in Figure 1, the precision of
@@
 demonstrated that the trend of increasing precision with higher species
-richness was general. Importantly however, our broader set of
-simulations revealed that the precision of \(\lambda_{est}\) was not
-uniform across all levels of phylogenetic signal. Specifically, we found
-that variation in \(\lambda_{est}\) was highest at intermediate levels
-of phylogenetic signal, and decreased at both lower and higher levels of
-input signal (Figure 1). This implied that the precision in estimating
-\(\lambda\) was worse at intermediate values, and improved as the levels
-of phylogenetic signal were closer to \(\lambda_{in}=0\) (no
-phylogenetic signal) or \(\lambda_{in}=1\) (Brownian motion). Notably,
-even at large levels of species richness, the range of \(\lambda_{est}\)
-still encompassed a substantial portion of possible values (e.g.,
-\(n=512\); \(\lambda_{in}=0.5\): range of
-\(\lambda_{est} = 0.32\to 0.68\)). Likewise, the same \(\lambda_{est}\)
-could be obtained from datasets containing vastly different input levels
-of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\); range
-of \(\lambda_{in} = 0.25 \to 0.65\)). Taken together, these findings
-reveal that \(\lambda_{est}\) does not precisely characterize observed
-levels of phylogenetic signal in phenotypic datasets, and that
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) may be highly inaccurate. \hfill\break
+richness was general; adhering to parametric statistical theory.
+Importantly however, our broader set of simulations revealed that the
+precision of \(\lambda_{est}\) was not uniform across all levels of
+phylogenetic signal. Specifically, we found that variation in
+\(\lambda_{est}\) was highest at intermediate levels of phylogenetic
+signal, and decreased at both lower and higher levels of input signal
+(Figure 1). This implied that the precision in estimating \(\lambda\)
+was worse at intermediate values, and improved as the levels of
+phylogenetic signal were closer to \(\lambda_{in}=0\) (no phylogenetic
+signal) or \(\lambda_{in}=1\) (Brownian motion). Notably, even at large
+levels of species richness, the range of \(\lambda_{est}\) still
+encompassed a substantial portion of possible values (e.g., \(n=512\);
+\(\lambda_{in}=0.5\): range of \(\lambda_{est} = 0.32\to 0.68\)).
+Likewise, the same \(\lambda_{est}\) could be obtained from datasets
+containing vastly different input levels of phylogenetic signal (e.g.,
+\(n=512\); \(\lambda_{est} = 0.5\); range of
+\(\lambda_{in} = 0.25 \to 0.65\)). Taken together, these findings reveal
+that \(\lambda_{est}\) does not precisely characterize observed levels
+of phylogenetic signal in phenotypic datasets, and that biological
+interpretations of the strength of phylogenetic signal based on
+\(\lambda\) may be highly inaccurate. \hfill\break
 
@@
 When \(\lambda\) was incorporated in PGLS regression and ANOVA (i.e.,
-\(Y\sim{X}\)), we found much the same pattern as in the previous
-simulations. Namely, the precision of \(\lambda_{est}\) varied inversely
-with species richness; where greater precision was obtained at higher
-levels of species richness (Figure 3). Likewise, the precision in
-estimating \(\lambda\) was worse at intermediate values (Figure 3), and
-improved as the levels of phylogenetic signal were closer to
-\(\lambda_{in}=0\) (no phylogenetic signal) or \(\lambda_{in}=1\)
-(Brownian motion). Generally, the spread of estimates was slightly
-broader when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression (Figure 3), as compared to when \(\lambda\) was
-estimated in cases only the dependent variable (Y) was considered
-(Figure 1). Finally, found that regression parameters (\(\beta\)) were
-accurately estimated when \(\lambda\) was included during phylogenetic
-regression; a result which confirmed earlier findings of Revell (2010)
-(see Supporting Information). \hfill\break
+\(Y\sim{X}\)), we found much the same pattern as in our earlier
+simulations. Namely, the precision of \(\lambda_{est}\) covaried with
+species richness; where greater precision was obtained at higher levels
+of species richness (Figure 3). Likewise, the precision in estimating
+\(\lambda\) was worse at intermediate values (Figure 3), and improved as
+the levels of phylogenetic signal were closer to \(\lambda_{in}=0\) (no
+phylogenetic signal) or \(\lambda_{in}=1\) (Brownian motion). Generally,
+the spread of estimates was slightly broader when \(\lambda\) was
+co-estimated with regression parameters in PGLS regression (Figure 3),
+as compared to when \(\lambda\) was estimated for only the dependent
+variable (Figure 1). Finally, we found that regression parameters
+(\(\beta\)) were accurately estimated when \(\lambda\) was included
+during phylogenetic regression; a result which confirmed earlier
+findings of Revell (2010) (see Supporting Information). \hfill\break
 
@@
 
+\leavevmode\hypertarget{ref-Arnqvist1995}{}%
+Arnqvist, G., and D. Wooster. 1995. Meta-analysis: Synthesizing research
+findings in ecology and evolution. Trends in Ecology and Evolution
+10:236--240.
+
+\leavevmode\hypertarget{ref-Bai2019}{}%
+Bai, K., S. Lv, S. Ning, D. Zeng, Y. Guo, and B. Wang. 2019. Leaf
+nutrient concentrations associated with phylogeny, leaf habit and soil
+chemistry in tropical karst seasonal rainforest tree species. Plant and
+Soil 434:305--326.
+
 \leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
@@
 
+\leavevmode\hypertarget{ref-Bose2019}{}%
+Bose, R., B. R. Ramesh, R. Pélissier, and F. Munoz. 2019. Phylogenetic
+diversity in the western ghats biodiversity hotspot reflects
+environmental filtering and past niche diversification of trees. Journal
+of Biogeography 46:145--157.
+
 \leavevmode\hypertarget{ref-ButlerKing2004}{}%
@@
 164:683--695.
+
+\leavevmode\hypertarget{ref-Cohen1988}{}%
+Cohen, J. 1988. Statistical power analysis for the behavioral sciences.
+Routledge.
+
+\leavevmode\hypertarget{ref-Cooper2010}{}%
+Cooper, N., W. Jetz, and R. P. Freckleton. 2010. Phylogenetic
+comparative approaches for studying niche conservatism. Journal of
+Evolutionary Biology 23:2529--2539.
+
+\leavevmode\hypertarget{ref-DeMeester2019}{}%
+De Meester, G., K. Huyghe, and R. Van Damme. 2019. Brain size, ecology
+and sociality: A reptilian perspective. Biological Journal of the
+Linnean Society 126:381--391.
 
@@
 
+\leavevmode\hypertarget{ref-Glass1976}{}%
+Glass, G. V. 1976. Primary, secondary, and meta-analysis of research.
+Educational Researcher 5:3--8.
+
 \leavevmode\hypertarget{ref-Grafen1989}{}%
@@
 
+\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
+Hedges, L. V., and I. Olkin. 1985. Statistical methods for
+meta-analysis. Elsevier.
+
 \leavevmode\hypertarget{ref-Klingenberg2010}{}%
@@
 biology 59:245--261.
+
+\leavevmode\hypertarget{ref-Liu2019}{}%
+Liu, H., C. P. Osborne, D. Yin, R. P. Freckleton, G. Jiang, and M. Liu.
+2019. Phylogeny and ecological processes influence grass coexistence at
+different spatial scales within the steppe biome. Oecologia 191:25--38.
 
@@
 
+\leavevmode\hypertarget{ref-Pintanel2019}{}%
+Pintanel, P., M. Tejedo, S. R. Ron, G. A. Llorente, and A.
+Merino-Viteri. 2019. Elevational and microclimatic drivers of thermal
+tolerance in andean pristimantis frogs. Journal of Biogeography
+46:1664--1675.
+
 \leavevmode\hypertarget{ref-RCT}{}%
@@
 variables: Geometric interpretations. Evolution 55:2143--2160.
+
+\leavevmode\hypertarget{ref-Rosenthal1994}{}%
+Rosenthal, R. 1994. The handbook of research synthesis. Pp. 231--244
+\emph{in} L. V. Cooper H Hedges, ed. Russell Sage Foundation.
+
+\leavevmode\hypertarget{ref-Su2019}{}%
+Su, G., S. Villéger, and S. Brosse. 2019. Morphological diversity of
+freshwater fishes differs between realms, but morphologically extreme
+species are widespread. Global ecology and biogeography 28:211--221.
+
+\leavevmode\hypertarget{ref-Vandelook2019}{}%
+Vandelook, F., S. Janssens, P. Gijbels, E. Fischer, W. Van den Ende, O.
+Honnay, and S. Abrahamczyk. 2019. Nectar traits differ between
+pollination syndromes in balsaminaceae. Annals of Botany 124:269--279.
 

### 2020-04-23 Data_Analyses/Lambda.Effect.r

@@ -1,0 +1,68 @@
+#Playing with CI for lambda effect size
+
+library(geiger)
+library(phytools)
+library(caper)
+library(geomorph)
+
+ntaxa <-  2^(5:10) # 32,64,128,256,512,1024
+nsim <- 25 # number of simulations
+lambdas <- seq(0,1,.05); lambdas[1] <- 0.005
+n <- ntaxa[3]
+
+tree<-pbtree(n=n,scale=1) 
+tree.scaled <- lapply(1:length(lambdas), function(j) rescale(tree, 
+            model = "lambda", lambda = lambdas[j]))
+lambda.est <- Z.lambda <- k.val <- Z.k <- matrix(NA, nrow=nsim, ncol=length(lambdas))
+
+X <- lapply(1:length(lambdas), function(j) sim.char(tree.scaled[[j]], par = 1, nsim = nsim) )
+for (i in 21:length(lambdas)){
+  data_x <- lapply(1:nsim, function(j) comparative.data(tree, data.frame(X.df=X[[i]][,,j], 
+                                                                       Species = names(X[[i]][,,j])), Species, vcv = F, vcv.dim = 2))
+  pgls_est <- lapply(1:nsim, function(j) pgls(X.df ~ 1, data_x[[j]], lambda='ML'))
+  K <- lapply(1:nsim, function(j) physignal(X[[i]][,,j],tree, print.progress = FALSE))
+  
+  lambda.est[,i] <- unlist(lapply(1:nsim, function(j) pgls_est[[j]]$param[2]))
+  CI <- lapply(1:nsim, function(j) pgls_est[[j]]$param.CI$lambda$ci.val)
+  
+  lambda.std <- unlist(lapply(1:nsim, function(j) {
+    if(is.na(CI[[j]])[1]==TRUE){lambda.std <- (CI[[j]][2]-lambda.est[j,i])*sqrt(n)/qnorm(0.975)} 
+    else {lambda.std <- (lambda.est[j,i]-CI[[j]][1])*sqrt(n)/qnorm(0.975)}
+  }))
+
+  #lambda.std <- (lambda.est[,i]-CI.lo)*sqrt(n)/qnorm(0.975)  #simple version unlooped
+  Z.lambda[,i] <- lambda.est[,i]/lambda.std
+  
[…]

### 2020-04-23 NewFig2.R

@@ -1,0 +1,30 @@
+# Compare variation across runs
+
+treesizes <-  2^(5:10)
+
+lambda <- read.csv("Data_Analyses/Munged_Data/Lambda_est_vars.csv", header=TRUE)
+Z.k <- read.csv("Data_Analyses/Munged_Data/Kappa_Z_norm_vars.csv", header=TRUE)
+ colnames(lambda) <- colnames(Z.k) <- c("lambda.in","32","64","128","256","512","1024")
+kappa_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_128.csv")
+
+#Plot for SM
+library(latex2exp)
+jpeg("Fig2.new.jpeg", res = 120, quality = 100, width = 1000, height = 1000)
+par(mfrow=c(2,2))
+plot(kappa_128$lambda.est ~ kappa_128$lambda.input, 
+     xlab = TeX('$\\lambda_{in}$'), ylab = TeX('$\\lambda_{est}$'), pch = 19)
+title("A", adj=0)
+text(min(kappa_128$lambda.input),max(kappa_128$lambda.est),
+     "n=128",adj= c(0.1,1))
+plot(kappa_128$kappa.z ~ kappa_128$lambda.input, 
+     xlab = TeX('$\\lambda_{in}$'), ylab = TeX('$Z_K$'),pch = 19)
+title("B", adj=0)
+text(min(kappa_128$lambda.input),max(kappa_128$kappa.z),
+     "n=128",adj= c(0.1,1))
+boxplot(lambda[,-1],ylim=c(0,.1),ylab = TeX('$\\sigma^2$'), xlab = "n")
+title("C", adj=0)
+boxplot(Z.k[,-1],ylim=c(0,.1),ylab = TeX('$\\sigma^2$'), xlab = "n")
+title("D", adj=0)
+par(mfrow=c(1,1))
+
+dev.off()

### 2020-04-23 Manuscript/Manuscript-2.Rmd
vsn 2

@@
 Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are well known ($\lambda=0$ is no phylogenetic signal, while $\lambda=1$ corresponds to Brownian motion).
-However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully examined.
+However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
 \hfill\break
@@
 First, it presumes that values of $\lambda$ can be precisely estimated, and that parameter estimates of $\lambda$ are representative of the underlying phylogenetic signal.
-Understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
+However, for such applications, understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
@@
 As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
-To date, it has not been determined whether $\lambda$ is an effect size for phylogenetic signal.
+For the case of $\lambda$, this has not yet been explored.
 \hfill\break
@@
 We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
-We then derive a standardized effect size for phylogenetic signal, and apply the concept to two common measures of phylogenetic signal: $\lambda$ ($Z_{\lambda}$) and $Kappa$ ($Z_K$).
-In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
-We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
+We find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
+We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
 Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
-Likewise, the precision of $Z_{\lambda}$ is less consistent than that observed with $Z_K$, across the range of input values, implying that $Z_K$ is a more robust effect size measure.
-Further, because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable, and may be compared across datasets.
+Thus, $\lambda$ is not a reliable effect size for measuring the strength of phylogenetic signal.
+We subsequently derive a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
+Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
+Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.

### 2020-04-23 Manuscript/Manuscript-2.tex
vsn 2

@@
 of \(\lambda\) directly to the strength of phylogenetic signal presumes
-two important statistical properties that have not been fully examined.
+two important statistical properties that have not been fully explored.
 \hfill\break
@@
 estimated, and that parameter estimates of \(\lambda\) are
-representative of the underlying phylogenetic signal. Understanding the
-precision in estimating levels of phylogenetic signal is paramount,
-because biological inferences regarding the strength of phylogenetic
-signal in phenotypic datasets requires high precision of its estimation.
-One study (Boettiger et al. 2012) found that estimates of Pagel's
-\(\lambda\) displayed less variation (i.e., greater precision) when data
-were simulated on a large phylogeny (\(N=281\)) as compared to a small
-one (\(N=13\)). From this observation it was concluded that insufficient
-data (i.e., the number of species) was the underlying cause of the
-increased variation across parameter estimates (Boettiger et al. 2012).
-Indeed, such a pattern is common with statistical estimators, as summary
-statistics and parameters are often more precise with greater sample
-sizes (Cohen 1988). However, this conclusion also assumes that the
-precision in estimating \(\lambda\) remains constant across its range
-(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
-verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
-macroevolutionary studies, at present, we still lack a general
-understanding of the precision with which \(\lambda\) can estimate
-levels of phylogenetic signal in phenotypic datasets. \hfill\break
+representative of the underlying phylogenetic signal. However, for such
+applications, understanding the precision in estimating levels of
+phylogenetic signal is paramount, because biological inferences
+regarding the strength of phylogenetic signal in phenotypic datasets
+requires high precision of its estimation. One study (Boettiger et al.
+2012) found that estimates of Pagel's \(\lambda\) displayed less
+variation (i.e., greater precision) when data were simulated on a large
+phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
+observation it was concluded that insufficient data (i.e., the number of
+species) was the underlying cause of the increased variation across
+parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
+common with statistical estimators, as summary statistics and parameters
+are often more precise with greater sample sizes (Cohen 1988). However,
+this conclusion also assumes that the precision in estimating
+\(\lambda\) remains constant across its range (\(\lambda = 0 \to 1\));
+an assumption that to date, has not been verified. Thus, despite
+widespread use of Pagel's (1999) \(\lambda\) in macroevolutionary
+studies, at present, we still lack a general understanding of the
+precision with which \(\lambda\) can estimate levels of phylogenetic
+signal in phenotypic datasets. \hfill\break
 
@@
 effect size can comparisons of estimates across datasets be
-interpretable. To date, it has not been determined whether \(\lambda\)
-is an effect size for phylogenetic signal. \hfill\break
+interpretable. For the case of \(\lambda\), this has not yet been
+explored. \hfill\break
 
@@
 to explore the degree to which \(\lambda\) correctly identifies known
-levels of phylogenetic signal, and under what circumstances. We then
-derive a standardized effect size for phylogenetic signal, and apply the
-concept to two common measures of phylogenetic signal: \(\lambda\)
-(\(Z_{\lambda}\)) and \(Kappa\) (\(Z_K\)). In general we find that while
-PGLS parameters (e.g., \(\beta\)) are accurately estimated with the
-inclusion of phylogenetic signal, estimates of \(\lambda\) are not. We
-also find that estimates of \(\lambda\) vary widely for a given input
-value of phylogenetic signal, and that this variation is not constant
-across the range of input signal, with decreased precision when
-phylogenetic signal is of intermediate strength. Additionally, the same
-\(\lambda_{est}\) may be obtained from datasets containing vastly
-different input levels of phylogenetic signal. Likewise, the precision
-of \(Z_{\lambda}\) is less consistent than that observed with \(Z_K\),
-across the range of input values, implying that \(Z_K\) is a more robust
-effect size measure. Further, because \(Z_K\) is a standardized effect
-size, the strength of phylogenetic signal is more readily interpretable,
-and may be compared across datasets. We conclude that estimates of
-phylogenetic signal using Pagel's \(\lambda\) are often inaccurate, and
-thus interpreting strength of phylogenetic signal in phenotypic datasets
-based on this measure is compromised. By contrast, effect sizes obtained
-from \(Kappa\) hold promise for characterizing phylogenetic signal, and
-for comparing the strength of phylogenetic signal across datasets.
+levels of phylogenetic signal, and under what circumstances. We find
+that while PGLS parameters (e.g., \(\beta\)) are accurately estimated
+with the inclusion of phylogenetic signal, estimates of \(\lambda\) are
+not. We also find that estimates of \(\lambda\) vary widely for a given
+input value of phylogenetic signal, and that the precision in estimating
+\(\lambda\) is not constant across the range of input signal, with
+decreased precision when phylogenetic signal is of intermediate
+strength. Additionally, the same \(\lambda_{est}\) may be obtained from
+datasets containing vastly different input levels of phylogenetic
+signal. Thus, \(\lambda\) is not a reliable effect size for measuring
+the strength of phylogenetic signal. We subsequently derive a
+standardized effect size for measuring the strength of phylogenetic
+signal in phenotypic datasets, and apply the concept to two common
+measures of phylogenetic signal: \(\lambda\) and \(Kappa\). Through
+simulations across a wide range of conditions, we find that the
+precision of effect sizes based on \(\lambda\) (\(Z_{\lambda}\)) are
+less reliable than that those based on \(Kappa\) (\(Z_K\)), implying
+that \(Z_K\) is a more robust effect size measure. Additionally, we
+propose a two-sample test statistic that may be used to compare the
+strength of phylogenetic signal among datasets. We conclude that
+estimates of phylogenetic signal using Pagel's \(\lambda\) are often
+inaccurate, and thus interpreting strength of phylogenetic signal in
+phenotypic datasets based on this measure is compromised. By contrast,
+effect sizes obtained from \(Kappa\) hold promise for characterizing
+phylogenetic signal, and for comparing the strength of phylogenetic
+signal across datasets.
 

### 2020-04-24 Manuscript/Manuscript-2.Rmd
vsn 2

@@
 A parameter ($\lambda$) is included, which transforms the lengths of the internal branches of the phylogeny to improve the fit [@Pagel1999; @Freckleton_et_al2002].
-Pagel's $\lambda$, which ranges from $0\to1$, is then treated as a measure of phylogenetic signal; with values close to 0 signifying little phylogenetic signal, and values close to 1 indicating that trait variation accumulates according to Brownian motion.
+Pagel's $\lambda$ ranges from $0\to1$, with larger values signifying a greater dependence of observed trait variation on the phylogeny.
 Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic regression  (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
@@
 Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-This is often accomplished by interpreting empirical estimates of $\lambda$; with $\lambda\approx0$ signifying 'weak' phylogenetic signal, and $\lambda\approx1$ being interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
+This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
 Other approaches for interpreting $\lambda$ are more statistical, through the use of confidence intervals [@Vandelook2019] or likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
 Likewise, qualitative comparisons of $\lambda$ across multiple phenotypic traits have also been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
-Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are well known ($\lambda=0$ is no phylogenetic signal, while $\lambda=1$ corresponds to Brownian motion).
+Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion).
 However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
@@
 
-First, it presumes that values of $\lambda$ can be precisely estimated, and that parameter estimates of $\lambda$ are representative of the underlying phylogenetic signal.
-However, for such applications, understanding the precision in estimating levels of phylogenetic signal is paramount, because biological inferences regarding the strength of phylogenetic signal in phenotypic datasets requires high precision of its estimation.
+First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
+Therefore, understanding the precision in estimating $\lambda$ is paramount.
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
@@
 
-Second, while $\lambda$ is observed on a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data.
+Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data.
 For this to be the case, $\lambda$ must be a statistical effect size.
@@
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets [e.g., meta-analysis: @HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared across datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
-Unfortunatley, not all statistical parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
+Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
 As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
@@
 
--Lambda is not precise: holds in estimates, and PGLSregression
-  -Precision varies with N (duh) but also WITH INPUT!
-  -same input on different N, different lambda.est
-  -same lambda.est from different input
-  -Boettiger example: add small tips, greatly change lambda...
--Conclusion: Lambda does not quantify strength of phylogenetic signal as is often assumed
+## The Precision of $\lambda$ is Variable
+
+We conducted a series of computer simulations to evaluate the precision of Pagel's $\lambda$.
+Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information).
+First we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$).
+Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in} = 0.0 \to 1.0$ using 21 intervals of 0.05 units, resulting in 1050 scaled phylogenies at each level of species richness ($n$).
+Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal corresponding reflecting Brownian motion (when $\lambda_{in} =1$).
+We then calculated the precision (variance: $\sigma^2_\lambda$) of $\lambda$ at each input level of phylogenetic signal and at each level of species richness ($n$).
+\hfill\break
+
+We also evaluated the precision of $\lambda$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$).
+Here, an independent variable $X$ was simulated on each phylogeny under a Brownian motion model of evolution (for PGLS regression).
+For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny.
+Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal.
+This was accomplished as: $Y=\beta{X}+\epsilon$.
+Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution: $\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: [see @Revell2010 for a similar simulation design].
+The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
+Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and at each level of species richness ($n$) were then observed.
+All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
+R-scripts are found in the Supporting Information.
+\hfill\break
+
+*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions.
+Predictably, precision improved as the number of species increased (Figure 1).
+This confirmed earlier findings of Boettiger et al.
+[-@Boettiger_et_al2012], and adhered to parametric statistical theory.
+However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal.
+Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approaches 0 and 1 (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$).
+Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values.
+Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$).
+Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$).
+Results were similar when $\lambda$ was co-estimated with regression parameters in PGLS regression (Figure 2).
+Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell 2010 [-@Revell2010] (see Supporting Information).
+However, estimates of phylogenetic signal were not, and the spread of $\lambda_{est}$ was even broader tham that observed when $\lambda$ was estimated for only the dependent variable.
+Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
+\hfill\break
+
+[insert Figure 1 here] \hfill\break
+
+[insert Figure 2 here] \hfill\break
+
+The above findings have several important implications for empirical studies.
 
@@
 
--A proposed Effect size for the strength of physig
+## A Standardized Effect Size for Phylogenetic Signal
+
   -Derivation of effect size (expand on Adams and Collyer 2019 suggestion that an effect size for phylogenetic signal is needed.
@@
 
-## $\lambda$ is not Precise
-
-## Lambda does not measure phylogenetic signal
-
-## An effect size for Physig
-
-# Materials and Methods
-
-## *Simulations*
-
-We conducted a series of computer simulations to evaluate the ability of Pagel's $\lambda$ to estimate known levels of phylogenetic signal.
-Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information).
-Our first set of simulations evaluated the extent to which values of $\lambda$ estimated from the data corresponded with actual input levels of phylogenetic signal.
-For these simulations we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$).
-Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda$.
-We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
-Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with known and differing levels of phylogenetic signal.
-These varied from datasets with no phylogenetic signal (when $\lambda_{in} =0$) to datsets with phylogenetic signal corresponding to what was expected under Brownian motion (when $\lambda_{in} =1$).
-\hfill\break
-
-Using the simulated phylogenies we estimated the phylogenetic signal in each dataset using Pagel's $\lambda$.
-Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression.
-Additionally, the precision of $\lambda$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$).
 Finally, for comparison we characterized the strength of phylogenetic signal in each dataset using a standardized effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] based on $Kappa$.
@@
 \hfill\break
-
-Our second set of simulations evaluated the extent to which values of $\lambda$ estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$) corresponded with actual input levels of phylogenetic signal in the response variable.
-As before we generated 50 pure-birth phylogenies at each of six levels of species richness ($n=2^5 - 2^{10}$), and rescaled each with a set of input $\lambda$ values ($\lambda_{in} = 0.0 - 1.0$; in 21 intervals of 0.05 units).
-Next, an independent variable $X$ was simulated on each phylogeny under a Brownian motion model of evolution (for PGLS regression).
-For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny.
-Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal.
-This was accomplished as: $Y=\beta{X}+\epsilon$.
-Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution: $\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: [see @Revell2010 for a similar simulation design].
-For each dataset, the fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained and evaluated as above.
-All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
-R-scripts are found in the Supporting Information.
 
@@
 
-Our first set of simulations revealed several patterns.
-First, the relationship between $\lambda_{est}$ and $\lambda_{in}$ across the range input levels of phylogenetic signal was $\beta\sim1.0$; indicating that the average $\lambda_{est}$ across datasets for a given simulation condition correctly reflected the input levels of phylogenetic signal.
-However, as shown in Figure 1, the precision of those estimates varied widely, and in several interesting ways.
-Predictably, precision was worse at low levels of species richness, where in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: range of $\lambda_{est}= 0.0\to 0.985$).
-Second, as species richness increased, variation across estimates of $\lambda$ decreased (Figure 1).
-This confirmed the pattern identified by Boettiger et al.
-[-@Boettiger_et_al2012] on  a small ($n=13$) versus a large ($n=281$) phylogeny, and demonstrated that the trend of increasing precision with higher species richness was general; adhering to parametric statistical theory.
-Importantly however, our broader set of simulations revealed that the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal.
-Specifically, we found that variation in $\lambda_{est}$ was highest at intermediate levels of phylogenetic signal, and decreased at both lower and higher levels of input signal (Figure 1).
-This implied that the precision in estimating $\lambda$ was worse at intermediate values, and improved as the levels of phylogenetic signal were closer to $\lambda_{in}=0$ (no phylogenetic signal) or $\lambda_{in}=1$ (Brownian motion).
-Notably, even at large levels of species richness, the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: range of $\lambda_{est} = 0.32\to 0.68$).
-Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; range of $\lambda_{in} = 0.25 \to 0.65$).
-Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
-\hfill\break
-
-[insert Figure 1 here] \hfill\break
-
 By contrast, when characterizing phylogenetic signal with the standardized effect size ($Z_K$) of $Kappa$, we found that the precision of $Z_K$ was more stable, as variation across datasets was far more consistent across the range of input values.
@@
 [insert Figure 2 here] \hfill\break
-
-When $\lambda$ was incorporated in PGLS regression and ANOVA (i.e., $Y\sim{X}$), we found much the same pattern as in our earlier simulations.
-Namely, the precision of $\lambda_{est}$ covaried with species richness; where greater precision was obtained at higher levels of species richness (Figure 3).
-Likewise, the precision in estimating $\lambda$ was worse at intermediate values (Figure 3), and improved as the levels of phylogenetic signal were closer to $\lambda_{in}=0$ (no phylogenetic signal) or $\lambda_{in}=1$ (Brownian motion).
-Generally, the spread of estimates was slightly broader when $\lambda$  was co-estimated with regression parameters in PGLS regression (Figure 3), as compared to when $\lambda$ was estimated for only the dependent variable (Figure 1).
-Finally, we found that regression parameters ($\beta$) were accurately estimated when $\lambda$ was included during phylogenetic regression; a result which confirmed earlier findings of Revell [-@Revell2010] (see Supporting Information).
-\hfill\break
-
-[insert Figure 3 here]
 
@@
 \textbf{Figure 2}.
+Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
+\hfill\break
+
+\textbf{Figure 3}.
 Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
@@
 (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-\hfill\break
-
-\textbf{Figure 3}.
-Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
-As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
 \hfill\break
@@
 \singlespacing \textbf{Figure 2}.
-Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
 
@@
 \singlespacing \textbf{Figure 3}.
-Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
-As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
+Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
+(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 

### 2020-04-24 Manuscript/Manuscript-2.tex
vsn 2

@@
 branches of the phylogeny to improve the fit (Pagel 1999; Freckleton et
-al. 2002). Pagel's \(\lambda\), which ranges from \(0\to1\), is then
-treated as a measure of phylogenetic signal; with values close to 0
-signifying little phylogenetic signal, and values close to 1 indicating
-that trait variation accumulates according to Brownian motion. Pagel's
-\(\lambda\) also has the appeal that it may be included in phylogenetic
-regression (PGLS) to account for the degree of phylogenetic signal in
-comparative analyses (see Freckleton et al. 2002). \hfill\break
+al. 2002). Pagel's \(\lambda\) ranges from \(0\to1\), with larger values
+signifying a greater dependence of observed trait variation on the
+phylogeny. Pagel's \(\lambda\) also has the appeal that it may be
+included in phylogenetic regression (PGLS) to account for the degree of
+phylogenetic signal in comparative analyses (see Freckleton et al.
+2002). \hfill\break
 
@@
 taxa. This is often accomplished by interpreting empirical estimates of
-\(\lambda\); with \(\lambda\approx0\) signifying `weak' phylogenetic
-signal, and \(\lambda\approx1\) being interpreted as `strong'
-phylogenetic signal (e.g., De Meester et al. 2019; Pintanel et al. 2019;
-Su et al. 2019). Other approaches for interpreting \(\lambda\) are more
-statistical, through the use of confidence intervals (Vandelook et al.
-2019) or likelihood ratio tests that compare the observed model fit to
-that obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
-2002; Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative
-comparisons of \(\lambda\) across multiple phenotypic traits have also
-been used to infer whether the strength of phylogenetic signal is
-greater in one trait as compared to another (e.g., Liu et al. 2019; Bai
-et al. 2019). Indeed, it seems intuitive to interpret the strength of
-phylogenetic signal in this manner, as \(\lambda\) is a parameter on a
-bounded scale (\(0\to1\)) for which interpretation of its extremal
-points are well known (\(\lambda=0\) is no phylogenetic signal, while
-\(\lambda=1\) corresponds to Brownian motion). However, equating values
-of \(\lambda\) directly to the strength of phylogenetic signal presumes
-two important statistical properties that have not been fully explored.
-\hfill\break
+\(\lambda\); with smaller values signifying `weak' phylogenetic signal,
+while larger values are interpreted as `strong' phylogenetic signal
+(e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
+Other approaches for interpreting \(\lambda\) are more statistical,
+through the use of confidence intervals (Vandelook et al. 2019) or
+likelihood ratio tests that compare the observed model fit to that
+obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al. 2002;
+Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative comparisons
+of \(\lambda\) across multiple phenotypic traits have also been used to
+infer whether the strength of phylogenetic signal is greater in one
+trait as compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
+Indeed, it seems intuitive to interpret the strength of phylogenetic
+signal in this manner, as \(\lambda\) is a parameter on a bounded scale
+(\(0\to1\)) for which interpretation of its extremal points are
+understood (\(\lambda=0\) represents no phylogenetic signal, while
+\(\lambda=1\) is phylogenetic signal as expected under Brownian motion).
+However, equating values of \(\lambda\) directly to the strength of
+phylogenetic signal presumes two important statistical properties that
+have not been fully explored. \hfill\break
 
 First, it presumes that values of \(\lambda\) can be precisely
-estimated, and that parameter estimates of \(\lambda\) are
-representative of the underlying phylogenetic signal. However, for such
-applications, understanding the precision in estimating levels of
-phylogenetic signal is paramount, because biological inferences
-regarding the strength of phylogenetic signal in phenotypic datasets
-requires high precision of its estimation. One study (Boettiger et al.
-2012) found that estimates of Pagel's \(\lambda\) displayed less
-variation (i.e., greater precision) when data were simulated on a large
-phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
-observation it was concluded that insufficient data (i.e., the number of
-species) was the underlying cause of the increased variation across
-parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
-common with statistical estimators, as summary statistics and parameters
-are often more precise with greater sample sizes (Cohen 1988). However,
-this conclusion also assumes that the precision in estimating
-\(\lambda\) remains constant across its range (\(\lambda = 0 \to 1\));
-an assumption that to date, has not been verified. Thus, despite
-widespread use of Pagel's (1999) \(\lambda\) in macroevolutionary
-studies, at present, we still lack a general understanding of the
-precision with which \(\lambda\) can estimate levels of phylogenetic
-signal in phenotypic datasets. \hfill\break
+estimated, as biological inferences regarding the strength of
+phylogenetic signal depend on high accuracy in its estimation.
+Therefore, understanding the precision in estimating \(\lambda\) is
+paramount. One study (Boettiger et al. 2012) found that estimates of
+Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
+when data were simulated on a large phylogeny (\(N=281\)) as compared to
+a small one (\(N=13\)). From this observation it was concluded that
+insufficient data (i.e., the number of species) was the underlying cause
+of the increased variation across parameter estimates (Boettiger et al.
+2012). Indeed, such a pattern is common with statistical estimators, as
+summary statistics and parameters are often more precise with greater
+sample sizes (Cohen 1988). However, this conclusion also assumes that
+the precision in estimating \(\lambda\) remains constant across its
+range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
+macroevolutionary studies, at present, we still lack a general
+understanding of the precision with which \(\lambda\) can estimate
+levels of phylogenetic signal in phenotypic datasets. \hfill\break
 
-Second, while \(\lambda\) is observed on a bounded scale (\(0\to1\)),
-this does not \emph{de-facto} imply that the estimated values of this
-parameter correspond to the actual strength of the underlying input
-signal in the data. For this to be the case, \(\lambda\) must be a
-statistical effect size. Effect sizes are a measure the magnitude of a
-statistical effect in data, represented on a common scale (Glass 1976;
-Cohen 1988). Effect sizes have widespread use in many areas of the
-quantiative sciences, as they represent measures that may be readily
-summarized across datasets (e.g., meta-analysis: Hedges and Olkin 1985;
-Glass 1976; Arnqvist and Wooster 1995), or compared across datasets
-(e.g., Adams and Collyer 2016, 2019a). Unfortunatley, not all
-statistical parameters and test statistics are effect sizes, and thus
-many summary measures must first be converted to standardized units
-(i.e., an effect size) for meaningful comparison (see Rosenthal 1994).
-As a consequence, it follows that only if \(\lambda\) is a statistical
-effect size can comparisons of estimates across datasets be
-interpretable. For the case of \(\lambda\), this has not yet been
-explored. \hfill\break
+Second, while estimates of \(\lambda\) are within a bounded scale
+(\(0\to1\)), this does not \emph{de-facto} imply that the estimated
+values of this parameter correspond to the actual strength of the
+underlying input signal in the data. For this to be the case,
+\(\lambda\) must be a statistical effect size. Effect sizes are a
+measure the magnitude of a statistical effect in data, represented on a
+common scale (Glass 1976; Cohen 1988). Effect sizes have widespread use
+in many areas of the quantiative sciences, as they represent measures
+that may be readily summarized across datasets (e.g., meta-analysis:
+Hedges and Olkin 1985; Glass 1976; Arnqvist and Wooster 1995), or
+compared across datasets (e.g., Adams and Collyer 2016, 2019a).
+Unfortunatley, not all model parameters and test statistics are effect
+sizes, and thus many summary measures must first be converted to
+standardized units (i.e., an effect size) for meaningful comparison (see
+Rosenthal 1994). As a consequence, it follows that only if \(\lambda\)
+is a statistical effect size can comparisons of estimates across
+datasets be interpretable. For the case of \(\lambda\), this has not yet
+been explored. \hfill\break
 
@@
 
--Lambda is not precise: holds in estimates, and PGLSregression
--Precision varies with N (duh) but also WITH INPUT! -same input on
-different N, different lambda.est -same lambda.est from different input
--Boettiger example: add small tips, greatly change lambda\ldots{}
--Conclusion: Lambda does not quantify strength of phylogenetic signal as
-is often assumed
+\hypertarget{the-precision-of-lambda-is-variable}{%
+\subsection{\texorpdfstring{The Precision of \(\lambda\) is
+Variable}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
+
+We conducted a series of computer simulations to evaluate the precision
+of Pagel's \(\lambda\). Our primary simulations were based on pure-birth
+phylogenies; however, we also evaluated patterns on both balanced and
+pectinate trees to determine whether tree shape affected our findings
+(see Supporting Information). First we generated 50 pure-birth
+phylogenies at each of six different tree sizes, ranging from 32 to 1024
+taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the simulated phylogenies
+by multiplying the internal branches by \(\lambda_{in} = 0.0 \to 1.0\)
+using 21 intervals of 0.05 units, resulting in 1050 scaled phylogenies
+at each level of species richness (\(n\)). Continuous traits were then
+simulated on each phylogeny under a Brownian motion model of evolution
+to obtain datasets with differing levels of phylogenetic signal, that
+ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
+phylogenetic signal corresponding reflecting Brownian motion (when
+\(\lambda_{in} =1\)). We then calculated the precision (variance:
+\(\sigma^2_\lambda\)) of \(\lambda\) at each input level of phylogenetic
+signal and at each level of species richness (\(n\)). \hfill\break
+
+We also evaluated the precision of \(\lambda\) when estimated in PGLS
+regression and ANOVA (i.e., \(Y\sim{X}\)). Here, an independent variable
+\(X\) was simulated on each phylogeny under a Brownian motion model of
+evolution (for PGLS regression). For phylogenetic ANOVA, random groups
+(\(X\)) were obtained by simulating a discrete (binary) character on
+each phylogeny. Next, the dependent variable was simulated in such a
+manner as to contain a known relationship with \(X\) plus random error
+containing phylogenetic signal. This was accomplished as:
+\(Y=\beta{X}+\epsilon\). Here, the association between \(Y\) and \(X\)
+was modeled using a range of values:
+\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
+to contain phylogenetic signal simulated under a Brownian motion model
+of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
+Revell 2010 for a similar simulation design). The fit of the
+phylogenetic regression was estimated using maximum likelihood, and
+parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
+obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
+of phylogenetic signal and at each level of species richness (\(n\))
+were then observed. All analyses were performed in R v3.6.0 (R Core Team
+2019) using the packages \texttt{geiger} (Harmon et al. 2008),
+\texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell 2012), and
+\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2020).
+R-scripts are found in the Supporting Information. \hfill\break
+
+\emph{Results.} We found that the precision of \(\lambda_{est}\) varied
+widely across simulation conditions. Predictably, precision improved as
+the number of species increased (Figure 1). This confirmed earlier
+findings of Boettiger et al.~(2012), and adhered to parametric
+statistical theory. However, in many cases the set of \(\lambda_{est}\)
+spanned nearly the entire range of possible values (e.g., \(n=32\);
+\(\lambda_{in}=0.5\): \(\lambda_{est}= 0.0\to 0.985\)), revealing that
+estimates of \(\lambda\) were not a reliable indicator of input
+phylogenetic signal. Importantly, the precision of \(\lambda_{est}\) was
+not uniform across all levels of phylogenetic signal, with the worst
+precision at intermediate levels of signal (\(\lambda_{in}\approx0.5\)),
+and improved precision as input levels approaches 0 and 1 (i.e.,
+\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
+estimates of \(\lambda\) were least reflective of the true input signal
+at intermediate values. Additionally, even at large levels of species
+richness, we found that the range of \(\lambda_{est}\) still encompassed
+a substantial portion of possible values (e.g., \(n=512\);
+\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
+same \(\lambda_{est}\) could be obtained from datasets containing vastly
+different input levels of phylogenetic signal (e.g., \(n=512\);
+\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). Results were
+similar when \(\lambda\) was co-estimated with regression parameters in
+PGLS regression (Figure 2). Here, regression parameters (\(\beta\)) were
+accurately estimated, confirming earlier findings of Revell 2010 (2010)
+(see Supporting Information). However, estimates of phylogenetic signal
+were not, and the spread of \(\lambda_{est}\) was even broader tham that
+observed when \(\lambda\) was estimated for only the dependent variable.
+Taken together, these findings reveal that \(\lambda_{est}\) does not
+precisely characterize observed levels of phylogenetic signal in
+phenotypic datasets, and that biological interpretations of the strength
+of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
+\hfill\break
+
+{[}insert Figure 1 here{]} \hfill\break
+
+{[}insert Figure 2 here{]} \hfill\break
+
+The above findings have several important implications for empirical
+studies.
 
@@
 
--A proposed Effect size for the strength of physig -Derivation of effect
-size (expand on Adams and Collyer 2019 suggestion that an effect size
-for phylogenetic signal is needed. The `gold standard' of these is a
-standaredized effect size: (following the meta-analytic literature
-beginning with Glass 1976). In other words, a Z-score. So, how does Z
-work when based on lambda? On kappa? - Z-score. could be Lambda or
-Kappa. Show it is Kappa -Comparing the strength of physig - two sample
-Z-score
+\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
+\subsection{A Standardized Effect Size for Phylogenetic
+Signal}\label{a-standardized-effect-size-for-phylogenetic-signal}}
+
+-Derivation of effect size (expand on Adams and Collyer 2019 suggestion
+that an effect size for phylogenetic signal is needed. The `gold
+standard' of these is a standaredized effect size: (following the
+meta-analytic literature beginning with Glass 1976). In other words, a
+Z-score. So, how does Z work when based on lambda? On kappa? - Z-score.
+could be Lambda or Kappa. Show it is Kappa -Comparing the strength of
+physig - two sample Z-score
 
@@
 
-\hypertarget{lambda-is-not-precise}{%
-\subsection{\texorpdfstring{\(\lambda\) is not
-Precise}{\textbackslash lambda is not Precise}}\label{lambda-is-not-precise}}
-
-\hypertarget{lambda-does-not-measure-phylogenetic-signal}{%
-\subsection{Lambda does not measure phylogenetic
-signal}\label{lambda-does-not-measure-phylogenetic-signal}}
-
-\hypertarget{an-effect-size-for-physig}{%
-\subsection{An effect size for Physig}\label{an-effect-size-for-physig}}
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{simulations}{%
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}
-
-We conducted a series of computer simulations to evaluate the ability of
-Pagel's \(\lambda\) to estimate known levels of phylogenetic signal. Our
-primary simulations were based on pure-birth phylogenies; however, we
-also evaluated patterns on both balanced and pectinate trees to
-determine whether tree shape affected our findings (see Supporting
-Information). Our first set of simulations evaluated the extent to which
-values of \(\lambda\) estimated from the data corresponded with actual
-input levels of phylogenetic signal. For these simulations we generated
-50 pure-birth phylogenies at each of six different tree sizes, ranging
-from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the
-simulated phylogenies by multiplying the internal branches by
-\(\lambda\). We used a set of values that encompassed the entire range
-of \(\lambda\): (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05
-units), resulting in 1050 scaled phylogenies at each level of species
-richness (\(n\)). Continuous traits were then simulated on each
-phylogeny under a Brownian motion model of evolution to obtain datasets
-with known and differing levels of phylogenetic signal. These varied
-from datasets with no phylogenetic signal (when \(\lambda_{in} =0\)) to
-datsets with phylogenetic signal corresponding to what was expected
-under Brownian motion (when \(\lambda_{in} =1\)). \hfill\break
-
-Using the simulated phylogenies we estimated the phylogenetic signal in
-each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
-each level of species richness (\(n\)), we assessed the relationship
-between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
-Additionally, the precision of \(\lambda\) was approximated by observing
-the variation of \(\lambda_{est}\) obtained at each input level of
-phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison we
-characterized the strength of phylogenetic signal in each dataset using
-a standardized effect size (\(Z_K\): sensu Adams and Collyer 2016,
-2019a) based on \(Kappa\). As suggested by Adams and Collyer (2019b), an
-effect size for phylogenetic signal may be estimated as:
-\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
-observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
-standard deviation of the empirical sampling distribution of values
+Finally, for comparison we characterized the strength of phylogenetic
+signal in each dataset using a standardized effect size (\(Z_K\): sensu
+Adams and Collyer 2016, 2019a) based on \(Kappa\). As suggested by Adams
+and Collyer (2019b), an effect size for phylogenetic signal may be
+estimated as: \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\)
+was the observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean
+and standard deviation of the empirical sampling distribution of values
 obtained from the permutation distribution. \(Z_K\) describes the
@@
 precision of \(Z_K\) from the normalized values. \hfill\break
-
-Our second set of simulations evaluated the extent to which values of
-\(\lambda\) estimated in PGLS regression and ANOVA (i.e., \(Y\sim{X}\))
-corresponded with actual input levels of phylogenetic signal in the
-response variable. As before we generated 50 pure-birth phylogenies at
-each of six levels of species richness (\(n=2^5 - 2^{10}\)), and
-rescaled each with a set of input \(\lambda\) values
-(\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units). Next, an
-independent variable \(X\) was simulated on each phylogeny under a
-Brownian motion model of evolution (for PGLS regression). For
-phylogenetic ANOVA, random groups (\(X\)) were obtained by simulating a
-discrete (binary) character on each phylogeny. Next, the dependent
-variable was simulated in such a manner as to contain a known
-relationship with \(X\) plus random error containing phylogenetic
-signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
-association between \(Y\) and \(X\) was modeled using a range of values:
-\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
-to contain phylogenetic signal simulated under a Brownian motion model
-of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
-Revell 2010 for a similar simulation design). For each dataset, the fit
-of the phylogenetic regression was estimated using maximum likelihood,
-and parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
-obtained and evaluated as above. All analyses were performed in R v3.6.0
-(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
-2008), \texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell
-2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
-al. 2020). R-scripts are found in the Supporting Information.
 
@@
 
-\hypertarget{simulations-1}{%
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations-1}}
-
-Our first set of simulations revealed several patterns. First, the
-relationship between \(\lambda_{est}\) and \(\lambda_{in}\) across the
-range input levels of phylogenetic signal was \(\beta\sim1.0\);
-indicating that the average \(\lambda_{est}\) across datasets for a
-given simulation condition correctly reflected the input levels of
-phylogenetic signal. However, as shown in Figure 1, the precision of
-those estimates varied widely, and in several interesting ways.
-Predictably, precision was worse at low levels of species richness,
-where in many cases the set of \(\lambda_{est}\) spanned nearly the
-entire range of possible values (e.g., \(n=32\); \(\lambda_{in}=0.5\):
-range of \(\lambda_{est}= 0.0\to 0.985\)). Second, as species richness
-increased, variation across estimates of \(\lambda\) decreased (Figure
-1). This confirmed the pattern identified by Boettiger et al.~(2012) on
-a small (\(n=13\)) versus a large (\(n=281\)) phylogeny, and
-demonstrated that the trend of increasing precision with higher species
-richness was general; adhering to parametric statistical theory.
-Importantly however, our broader set of simulations revealed that the
-precision of \(\lambda_{est}\) was not uniform across all levels of
-phylogenetic signal. Specifically, we found that variation in
-\(\lambda_{est}\) was highest at intermediate levels of phylogenetic
-signal, and decreased at both lower and higher levels of input signal
-(Figure 1). This implied that the precision in estimating \(\lambda\)
-was worse at intermediate values, and improved as the levels of
-phylogenetic signal were closer to \(\lambda_{in}=0\) (no phylogenetic
-signal) or \(\lambda_{in}=1\) (Brownian motion). Notably, even at large
-levels of species richness, the range of \(\lambda_{est}\) still
-encompassed a substantial portion of possible values (e.g., \(n=512\);
-\(\lambda_{in}=0.5\): range of \(\lambda_{est} = 0.32\to 0.68\)).
-Likewise, the same \(\lambda_{est}\) could be obtained from datasets
-containing vastly different input levels of phylogenetic signal (e.g.,
-\(n=512\); \(\lambda_{est} = 0.5\); range of
-\(\lambda_{in} = 0.25 \to 0.65\)). Taken together, these findings reveal
-that \(\lambda_{est}\) does not precisely characterize observed levels
-of phylogenetic signal in phenotypic datasets, and that biological
-interpretations of the strength of phylogenetic signal based on
-\(\lambda\) may be highly inaccurate. \hfill\break
-
-{[}insert Figure 1 here{]} \hfill\break
+\hypertarget{simulations}{%
+\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}
 
@@
 {[}insert Figure 2 here{]} \hfill\break
-
-When \(\lambda\) was incorporated in PGLS regression and ANOVA (i.e.,
-\(Y\sim{X}\)), we found much the same pattern as in our earlier
-simulations. Namely, the precision of \(\lambda_{est}\) covaried with
-species richness; where greater precision was obtained at higher levels
-of species richness (Figure 3). Likewise, the precision in estimating
-\(\lambda\) was worse at intermediate values (Figure 3), and improved as
-the levels of phylogenetic signal were closer to \(\lambda_{in}=0\) (no
-phylogenetic signal) or \(\lambda_{in}=1\) (Brownian motion). Generally,
-the spread of estimates was slightly broader when \(\lambda\) was
-co-estimated with regression parameters in PGLS regression (Figure 3),
-as compared to when \(\lambda\) was estimated for only the dependent
-variable (Figure 1). Finally, we found that regression parameters
-(\(\beta\)) were accurately estimated when \(\lambda\) was included
-during phylogenetic regression; a result which confirmed earlier
-findings of Revell (2010) (see Supporting Information). \hfill\break
-
-{[}insert Figure 3 here{]}
 
@@
 
-\textbf{Figure 2}. Variation in estimates of phylogenetic signal across
+\textbf{Figure 2}. Precision of Pagel's \(\lambda\) when incorporated in
+phylogenetic regression (\(|Y\sim X\)), across known levels of input
+phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
+As phylogenies increase in size, variation in \(\lambda_{in}\)
+decreases; however the precision is not constant across the range of
+input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
+levels of phylogenetic signal. \hfill\break
+
+\textbf{Figure 3}. Variation in estimates of phylogenetic signal across
 input levels of phylogenetic signal. (A) Estimates of Pagel's
@@
 containing differing numbers of species. \hfill\break
-
-\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
-phylogenetic regression (\(|Y\sim X\)), across known levels of input
-phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
-As phylogenies increase in size, variation in \(\lambda_{in}\)
-decreases; however the precision is not constant across the range of
-input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
-levels of phylogenetic signal. \hfill\break
 
@@
 
-\singlespacing \textbf{Figure 2}. Variation in estimates of phylogenetic
+\singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\) when
+incorporated in phylogenetic regression (\(|Y\sim X\)), across known
+levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
+various sizes. As phylogenies increase in size, variation in
+\(\lambda_{in}\) decreases; however the precision is not constant across
+the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
+intermediate levels of phylogenetic signal.
+
+\newpage
+
+\includegraphics[width=0.95\linewidth]{Fig3}
+
+\singlespacing \textbf{Figure 3}. Variation in estimates of phylogenetic
 signal across input levels of phylogenetic signal. (A) Estimates of
@@
 
-\includegraphics[width=0.95\linewidth]{Fig3}
-
-\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
-incorporated in phylogenetic regression (\(|Y\sim X\)), across known
-levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
-various sizes. As phylogenies increase in size, variation in
-\(\lambda_{in}\) decreases; however the precision is not constant across
-the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
-intermediate levels of phylogenetic signal.
-
-\newpage
-
 \includegraphics[width=0.95\linewidth]{Fig4}

### 2020-04-24 Manuscript/Manuscript-2.Rmd
vsn 2

@@
 From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
-Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise with greater sample sizes [@Cohen1988].
-However, this conclusion also assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
+Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
+However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
 Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
@@
 Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
-Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets [e.g., meta-analysis: @HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared across datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
+Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analysis [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
 Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
@@
 Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
-Thus, $\lambda$ is not a reliable effect size for measuring the strength of phylogenetic signal.
+Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data.
 We subsequently derive a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
 Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
-Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets.
+Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
@@
 
-## The Precision of $\lambda$ is Variable
+## *The Precision of $\lambda$ is Variable*
 
@@
 First we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$).
-Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in} = 0.0 \to 1.0$ using 21 intervals of 0.05 units, resulting in 1050 scaled phylogenies at each level of species richness ($n$).
+Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
 Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal corresponding reflecting Brownian motion (when $\lambda_{in} =1$).
-We then calculated the precision (variance: $\sigma^2_\lambda$) of $\lambda$ at each input level of phylogenetic signal and at each level of species richness ($n$).
+For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the precision of $\lambda$ using the variance ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness.
 \hfill\break
@@
 The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
-Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and at each level of species richness ($n$) were then observed.
+Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness were then observed.
+\hfill\break
+
 All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
@@
 However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal.
-Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approaches 0 and 1 (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$).
+Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approached the extremes of its range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$).
 Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values.
@@
 Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell 2010 [-@Revell2010] (see Supporting Information).
-However, estimates of phylogenetic signal were not, and the spread of $\lambda_{est}$ was even broader tham that observed when $\lambda$ was estimated for only the dependent variable.
+However, estimates of phylogenetic signal were not, and the spread of $\lambda_{est}$ was even broader than that observed when $\lambda$ was estimated for only the dependent variable.
 Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
@@
 
-The above findings have several important implications for empirical studies.
+## *A Standardized Effect Size for Phylogenetic Signal*
 
--Empirical examples: how interpretation can get folks ito trouble.
+The results above demonstate that $\lambda$ is not a reliable estimate of the phylogenetic signal in phenotypic data.
+As such, biological interpretations of the strength of phylogenetic signal, and comparisons of the magnitude of such effects across datasets, are severely compromised when based on this parameter.
+As an alternative, we propose that summary estimates of phylogenetic signal be converted to effect sizes for these purposes.
+A standardized effect size is found as:
 
-## A Standardized Effect Size for Phylogenetic Signal
+\begin{align}
+    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
+\end{align}
 
-  -Derivation of effect size (expand on Adams and Collyer 2019 suggestion that an effect size for phylogenetic signal is needed.
-The ‘gold standard’ of these is a standaredized effect size: (following the meta-analytic literature beginning with Glass 1976).
-In other words, a Z-score.
-So, how does Z work when based on lambda?
-On kappa?
+where $\theta_{obs}$ is the observed test statistic, $E(\theta)$ is its expected value under the null hypothesis, and $\sigma_\theta$ is its standard error [@Glass1976; @Cohen1988;@Rosenthal1994].
+$Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to a standard normal deviate [@Glass1976; @Kelley2012].
+Here, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
+However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution obtained from permutation procedures.
+\hfill\break
+
+Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented an effect size, obtained from $Kappa$ and its empirical sampling distribution from permutation.
+Here we formalize that suggestion, and find an effect size as:
+
+\begin{align}
+    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
+\end{align}
+
+Similarly, an effect size based on $\lambda$ could be envisioned as:
+
+\begin{align}
+    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
+\end{align}
+
+Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002].
+
   - Z-score. could be Lambda or Kappa.  Show it is Kappa
@@
 Finally, for comparison we characterized the strength of phylogenetic signal in each dataset using a standardized effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] based on $Kappa$.
-As suggested by Adams and Collyer [-@AdamsCollyer2019], an effect size for phylogenetic signal may be estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution.
-$Z_K$ describes the strength of phylogenetic signal as a standard deviate from its sampling distribution, and thus directly measures the strength of signal in a manner that is comparable across datasets.
+
 Variation in the set of $Z_K$ at each input level of phylogenetic signal was then calculated as an estimate of precision in $Z_K$.

### 2020-04-24 Manuscript/Manuscript-2.tex
vsn 2

@@
 2012). Indeed, such a pattern is common with statistical estimators, as
-summary statistics and parameters are often more precise with greater
+summary statistics and parameters are often more precise at greater
 sample sizes (Cohen 1988). However, this conclusion also assumes that
-the precision in estimating \(\lambda\) remains constant across its
-range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+the precision of \(\lambda\) remains constant across its range
+(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
 verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
@@
 in many areas of the quantiative sciences, as they represent measures
-that may be readily summarized across datasets (e.g., meta-analysis:
-Hedges and Olkin 1985; Glass 1976; Arnqvist and Wooster 1995), or
-compared across datasets (e.g., Adams and Collyer 2016, 2019a).
+that may be readily summarized across datasets as in meta-analysis
+(Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster 1995), or
+compared among datasets (e.g., Adams and Collyer 2016, 2019a).
 Unfortunatley, not all model parameters and test statistics are effect
@@
 datasets containing vastly different input levels of phylogenetic
-signal. Thus, \(\lambda\) is not a reliable effect size for measuring
-the strength of phylogenetic signal. We subsequently derive a
+signal. Thus, \(\lambda\) is not a reliable estimate of the strength of
+phylogenetic signal in phenotypic data. We subsequently derive a
 standardized effect size for measuring the strength of phylogenetic
@@
 propose a two-sample test statistic that may be used to compare the
-strength of phylogenetic signal among datasets. We conclude that
-estimates of phylogenetic signal using Pagel's \(\lambda\) are often
-inaccurate, and thus interpreting strength of phylogenetic signal in
-phenotypic datasets based on this measure is compromised. By contrast,
-effect sizes obtained from \(Kappa\) hold promise for characterizing
-phylogenetic signal, and for comparing the strength of phylogenetic
-signal across datasets.
+strength of phylogenetic signal among datasets, and provide an empirical
+example to demonstrate its use. We conclude that estimates of
+phylogenetic signal using Pagel's \(\lambda\) are often inaccurate, and
+thus interpreting strength of phylogenetic signal in phenotypic datasets
+based on this measure is compromised. By contrast, effect sizes obtained
+from \(Kappa\) hold promise for characterizing phylogenetic signal, and
+for comparing the strength of phylogenetic signal across datasets.
 
@@
 \hypertarget{the-precision-of-lambda-is-variable}{%
-\subsection{\texorpdfstring{The Precision of \(\lambda\) is
-Variable}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
+\subsection{\texorpdfstring{\emph{The Precision of \(\lambda\) is
+Variable}}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
 
@@
 taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the simulated phylogenies
-by multiplying the internal branches by \(\lambda_{in} = 0.0 \to 1.0\)
-using 21 intervals of 0.05 units, resulting in 1050 scaled phylogenies
+by multiplying the internal branches by \(\lambda_{in}\), using 21
+intervals of 0.05 units across its range
+(\(\lambda_{in} = 0.0 \to 1.0\)), resulting in 1050 scaled phylogenies
 at each level of species richness (\(n\)). Continuous traits were then
@@
 phylogenetic signal corresponding reflecting Brownian motion (when
-\(\lambda_{in} =1\)). We then calculated the precision (variance:
-\(\sigma^2_\lambda\)) of \(\lambda\) at each input level of phylogenetic
-signal and at each level of species richness (\(n\)). \hfill\break
+\(\lambda_{in} =1\)). For each dataset we then estimated phylogenetic
+signal (\(\lambda_{est}\)), and calculated the precision of \(\lambda\)
+using the variance (\(\sigma^2_\lambda\)) across datasets at each input
+level of phylogenetic signal and level of species richness. \hfill\break
 
@@
 obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
-of phylogenetic signal and at each level of species richness (\(n\))
-were then observed. All analyses were performed in R v3.6.0 (R Core Team
-2019) using the packages \texttt{geiger} (Harmon et al. 2008),
-\texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell 2012), and
-\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2020).
-R-scripts are found in the Supporting Information. \hfill\break
+of phylogenetic signal and level of species richness were then observed.
+\hfill\break
+
+All analyses were performed in R v3.6.0 (R Core Team 2019) using the
+packages \texttt{geiger} (Harmon et al. 2008), \texttt{caper} (Orme et
+al. 2013), \texttt{phytools} (Revell 2012), and \texttt{geomorph} (Adams
+and Otárola-Castillo 2013; Adams et al. 2020). R-scripts are found in
+the Supporting Information. \hfill\break
 
@@
 precision at intermediate levels of signal (\(\lambda_{in}\approx0.5\)),
-and improved precision as input levels approaches 0 and 1 (i.e.,
-\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
-estimates of \(\lambda\) were least reflective of the true input signal
-at intermediate values. Additionally, even at large levels of species
-richness, we found that the range of \(\lambda_{est}\) still encompassed
-a substantial portion of possible values (e.g., \(n=512\);
-\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
-same \(\lambda_{est}\) could be obtained from datasets containing vastly
-different input levels of phylogenetic signal (e.g., \(n=512\);
-\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). Results were
-similar when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression (Figure 2). Here, regression parameters (\(\beta\)) were
-accurately estimated, confirming earlier findings of Revell 2010 (2010)
-(see Supporting Information). However, estimates of phylogenetic signal
-were not, and the spread of \(\lambda_{est}\) was even broader tham that
-observed when \(\lambda\) was estimated for only the dependent variable.
-Taken together, these findings reveal that \(\lambda_{est}\) does not
+and improved precision as input levels approached the extremes of its
+range (i.e., \(\lambda_{in}\rightarrow0\) \&
+\(\lambda_{in}\rightarrow1\)). Thus, estimates of \(\lambda\) were least
+reflective of the true input signal at intermediate values.
+Additionally, even at large levels of species richness, we found that
+the range of \(\lambda_{est}\) still encompassed a substantial portion
+of possible values (e.g., \(n=512\); \(\lambda_{in}=0.5\):
+\(\lambda_{est} = 0.32\to 0.68\)). Likewise, the same \(\lambda_{est}\)
+could be obtained from datasets containing vastly different input levels
+of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\);
+\(\lambda_{in} = 0.25 \to 0.65\)). Results were similar when \(\lambda\)
+was co-estimated with regression parameters in PGLS regression (Figure
+2). Here, regression parameters (\(\beta\)) were accurately estimated,
+confirming earlier findings of Revell 2010 (2010) (see Supporting
+Information). However, estimates of phylogenetic signal were not, and
+the spread of \(\lambda_{est}\) was even broader than that observed when
+\(\lambda\) was estimated for only the dependent variable. Taken
+together, these findings reveal that \(\lambda_{est}\) does not
 precisely characterize observed levels of phylogenetic signal in
@@
 
-The above findings have several important implications for empirical
-studies.
+\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
+\subsection{\texorpdfstring{\emph{A Standardized Effect Size for
+Phylogenetic
+Signal}}{A Standardized Effect Size for Phylogenetic Signal}}\label{a-standardized-effect-size-for-phylogenetic-signal}}
 
--Empirical examples: how interpretation can get folks ito trouble.
+The results above demonstate that \(\lambda\) is not a reliable estimate
+of the phylogenetic signal in phenotypic data. As such, biological
+interpretations of the strength of phylogenetic signal, and comparisons
+of the magnitude of such effects across datasets, are severely
+compromised when based on this parameter. As an alternative, we propose
+that summary estimates of phylogenetic signal be converted to effect
+sizes for these purposes. A standardized effect size is found as:
 
-\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
-\subsection{A Standardized Effect Size for Phylogenetic
-Signal}\label{a-standardized-effect-size-for-phylogenetic-signal}}
+\begin{align}
+    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
+\end{align}
 
--Derivation of effect size (expand on Adams and Collyer 2019 suggestion
-that an effect size for phylogenetic signal is needed. The `gold
-standard' of these is a standaredized effect size: (following the
-meta-analytic literature beginning with Glass 1976). In other words, a
-Z-score. So, how does Z work when based on lambda? On kappa? - Z-score.
-could be Lambda or Kappa. Show it is Kappa -Comparing the strength of
-physig - two sample Z-score
+where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
+its expected value under the null hypothesis, and \(\sigma_\theta\) is
+its standard error (Glass 1976; Cohen 1988; Rosenthal 1994).
+\(Z_{\theta}\) expresses the magnitude of the effect in \(\theta_{obs}\)
+by transforming the original test statistic to a standard normal deviate
+(Glass 1976; Kelley and Preacher 2012). Here, \(\theta_{obs}\) and
+\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
+obtained from the distribution of \(\theta\) derived from parametric
+theory. However, recent advances in resampling theory (Collyer et al.
+2015; Adams and Collyer 2016, 2019a) have shown that \(E(\theta)\) and
+\(\sigma_\theta\) may also be obtained from an empirical sampling
+distribution obtained from permutation procedures. \hfill\break
+
+Adams and Collyer (2019b) recently suggested that the strength of
+phylogenetic signal could be represented an effect size, obtained from
+\(Kappa\) and its empirical sampling distribution from permutation. Here
+we formalize that suggestion, and find an effect size as:
+
+\begin{align}
+    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
+\end{align}
+
+Similarly, an effect size based on \(\lambda\) could be envisioned as:
+
+\begin{align}
+    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
+\end{align}
+
+Note that under the null hypothesis, \(E(\lambda)=0\), a no phylogenetic
+signal is expected under this condition (Freckleton et al. 2002).
+
+\begin{itemize}
+\tightlist
+\item
+  Z-score. could be Lambda or Kappa. Show it is Kappa -Comparing the
+  strength of physig
+\item
+  two sample Z-score
+\end{itemize}
 
@@
 signal in each dataset using a standardized effect size (\(Z_K\): sensu
-Adams and Collyer 2016, 2019a) based on \(Kappa\). As suggested by Adams
-and Collyer (2019b), an effect size for phylogenetic signal may be
-estimated as: \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\)
-was the observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean
-and standard deviation of the empirical sampling distribution of values
-obtained from the permutation distribution. \(Z_K\) describes the
-strength of phylogenetic signal as a standard deviate from its sampling
-distribution, and thus directly measures the strength of signal in a
-manner that is comparable across datasets. Variation in the set of
-\(Z_K\) at each input level of phylogenetic signal was then calculated
-as an estimate of precision in \(Z_K\). However, because \(Z_K\) differs
-in scale from \(\lambda\), we used a linear normalization to standardize
-\(Z_K\) to a uniform distribution (\(0\rightarrow1\)), and estimated the
-precision of \(Z_K\) from the normalized values. \hfill\break
+Adams and Collyer 2016, 2019a) based on \(Kappa\).
+
+Variation in the set of \(Z_K\) at each input level of phylogenetic
+signal was then calculated as an estimate of precision in \(Z_K\).
+However, because \(Z_K\) differs in scale from \(\lambda\), we used a
+linear normalization to standardize \(Z_K\) to a uniform distribution
+(\(0\rightarrow1\)), and estimated the precision of \(Z_K\) from the
+normalized values. \hfill\break
 
@@
 
+\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
+Collyer, M. L., D. J. Sekora, and D. C. Adams. 2015. A method for
+analysis of phenotypic change for phenotypes described by
+high-dimensional data. Heredity 115:357--365.
+
 \leavevmode\hypertarget{ref-Cooper2010}{}%
@@
 meta-analysis. Elsevier.
+
+\leavevmode\hypertarget{ref-Kelley2012}{}%
+Kelley, K., and K. J. Preacher. 2012. On effect size. Psychological
+Methods 17:137--152.
 
@@
 \textbf{Figure 2}. Precision of Pagel's \(\lambda\) when incorporated in
-phylogenetic regression (\(|Y\sim X\)), across known levels of input
+phylogenetic regression (\(Y\sim X\)), across known levels of input
 phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
@@
 \singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\) when
-incorporated in phylogenetic regression (\(|Y\sim X\)), across known
+incorporated in phylogenetic regression (\(Y\sim X\)), across known
 levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of

### 2020-04-24 Manuscript/Manuscript-2.Rmd
vsn 2

@@
 As such, biological interpretations of the strength of phylogenetic signal, and comparisons of the magnitude of such effects across datasets, are severely compromised when based on this parameter.
-As an alternative, we propose that summary estimates of phylogenetic signal be converted to effect sizes for these purposes.
-A standardized effect size is found as:
+As an alternative, we propose that summary estimates of phylogenetic signal be converted to effect sizes for interpretation and comparison.
+Statistically, a standardized effect size may be found as:
 
@@
 $Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to a standard normal deviate [@Glass1976; @Kelley2012].
-Here, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
-However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution obtained from permutation procedures.
+Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
+However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures.
 \hfill\break
@@
 
-Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002].
+Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002]. \hfill\break
 
-  - Z-score. could be Lambda or Kappa.  Show it is Kappa
--Comparing the strength of physig
-  - two sample Z-score
-
--Conclusions and Implications
-
-Finally, for comparison we characterized the strength of phylogenetic signal in each dataset using a standardized effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] based on $Kappa$.
-
-Variation in the set of $Z_K$ at each input level of phylogenetic signal was then calculated as an estimate of precision in $Z_K$.
-However, because $Z_K$ differs in scale from $\lambda$, we used a linear normalization to standardize $Z_K$ to a uniform distribution ($0\rightarrow1$), and estimated the precision of $Z_K$ from the normalized values.
+To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
+Results are found in Figure 3.
+Here two things are evident.
+First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than to estimates of $Z_\lambda$.
+Second, the precision of $Z_K$ is considerably more stable, as coefficients of variation for the set of $\sigma^2_{Z_K}$ across input levels of phylogenetic signal for the same sample size was an order of magnitude smaller than that observed for $\sigma^2_{Z_\lambda}$ (Figure 3).
+This implies that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
 \hfill\break
 
-## *Literature Survey*
+[insert Figure 3 here]
+
+## *Statistical Comparisons of Phylogenetic Signal*
+
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, it may be of interest to compare such measures across datasets.
+This is useful, for instance, to determine whether the strength of phylogenetic signal is greater in one phenotypic trait as compared with another.
+As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:
+
+\begin{align}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}}
+\end{align}
+
+where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2.
+Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
+As with other two-sample tests, $\hat{Z}_{12}$ is typically considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [see @AdamsCollyer2016; @AdamsCollyer2019b].
+
+## *Empirical Example*
+
+# Conclusions and Implications
+
+## *Literature Survey... WHERE DOES THIS GO???*
 
@@
 
-# Results
-
-## *Simulations*
-
-By contrast, when characterizing phylogenetic signal with the standardized effect size ($Z_K$) of $Kappa$, we found that the precision of $Z_K$ was more stable, as variation across datasets was far more consistent across the range of input values.
-For example, when $n=128$ the precision of $\lambda_{est}$ (Figure 2A) varied considerably more across input levels of phylogenetic signal than did the precision of $Z_K$ for the same datasets (Figure 2B).
-Further, for a given $n$, the variance in precision estimates of $Z_K$ was considerably smaller than the variance in precision estimates of $\lambda_{est}$ (Fig.
-2C,D); implying that estimates of phylogenetic signal were more reliable and robust when using $Z_K$ (for additional results see Supporting Information).
-Finally, it should also be recognized that because $Z_K$ is a standardized effect size, the strength of phylogenetic signal is more readily interpretable when using this measure, as $Z_K$ expresses the strength of phylogenetic signal in standard deviation units relative to the mean [see @AdamsCollyer2019].
-This further implies that comparisons of the strength of phylogenetic signal among phenotypic traits are possible, and may be accomplished statistically via a two-sample test that formally compares $Z_K$ across datasets [for comparisons of multivariate effect sizes see: @AdamsCollyer2016; @AdamsCollyer2019b].
-\hfill\break
-
-[insert Figure 2 here] \hfill\break
-
-## *Literature Survey*
-
+*Results*.
 We found 182 manuscripts from 2019 that estimated and reported Pagel's lambda values using PGLS methods.
@@
 Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so.
-
-# Results
 

### 2020-04-24 Manuscript/Manuscript-2.tex
vsn 2

@@
 that summary estimates of phylogenetic signal be converted to effect
-sizes for these purposes. A standardized effect size is found as:
+sizes for interpretation and comparison. Statistically, a standardized
+effect size may be found as:
 
@@
 by transforming the original test statistic to a standard normal deviate
-(Glass 1976; Kelley and Preacher 2012). Here, \(\theta_{obs}\) and
+(Glass 1976; Kelley and Preacher 2012). Typically, \(\theta_{obs}\) and
 \(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
@@
 \(\sigma_\theta\) may also be obtained from an empirical sampling
-distribution obtained from permutation procedures. \hfill\break
+distribution of \(\theta\) obtained from permutation procedures.
+\hfill\break
 
@@
 signal is expected under this condition (Freckleton et al. 2002).
+\hfill\break
 
-\begin{itemize}
-\tightlist
-\item
-  Z-score. could be Lambda or Kappa. Show it is Kappa -Comparing the
-  strength of physig
-\item
-  two sample Z-score
-\end{itemize}
+To evaluate the utility of \(Z_K\) and \(Z_\lambda\) we calculated both
+effect sizes for the simulated datasets generated above, and summarized
+the precision of each using its variance (\(\sigma^2_{Z_K}\) and
+\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 3. Here two
+things are evident. First, estimates of \(Z_K\) track the input
+phylogenetic signal in a more linear fashion than to estimates of
+\(Z_\lambda\). Second, the precision of \(Z_K\) is considerably more
+stable, as coefficients of variation for the set of \(\sigma^2_{Z_K}\)
+across input levels of phylogenetic signal for the same sample size was
+an order of magnitude smaller than that observed for
+\(\sigma^2_{Z_\lambda}\) (Figure 3). This implies that estimates of the
+strength of phylogenetic signal were more reliable and robust when using
+\(Z_K\) as compared with \(Z_\lambda\). \hfill\break
 
--Conclusions and Implications
+{[}insert Figure 3 here{]}
 
-Finally, for comparison we characterized the strength of phylogenetic
-signal in each dataset using a standardized effect size (\(Z_K\): sensu
-Adams and Collyer 2016, 2019a) based on \(Kappa\).
+\hypertarget{statistical-comparisons-of-phylogenetic-signal}{%
+\subsection{\texorpdfstring{\emph{Statistical Comparisons of
+Phylogenetic
+Signal}}{Statistical Comparisons of Phylogenetic Signal}}\label{statistical-comparisons-of-phylogenetic-signal}}
 
-Variation in the set of \(Z_K\) at each input level of phylogenetic
-signal was then calculated as an estimate of precision in \(Z_K\).
-However, because \(Z_K\) differs in scale from \(\lambda\), we used a
-linear normalization to standardize \(Z_K\) to a uniform distribution
-(\(0\rightarrow1\)), and estimated the precision of \(Z_K\) from the
-normalized values. \hfill\break
+Once the magnitude of phylogenetic signal is characterized using
+\(Z_K\), it may be of interest to compare such measures across datasets.
+This is useful, for instance, to determine whether the strength of
+phylogenetic signal is greater in one phenotypic trait as compared with
+another. As with other effect sizes derived from permutation
+distributions (e.g., Adams and Collyer 2016, 2019a), a two-sample test
+statistic may be found as:
 
-\hypertarget{literature-survey}{%
-\subsection{\texorpdfstring{\emph{Literature
-Survey}}{Literature Survey}}\label{literature-survey}}
+\begin{align}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}}
+\end{align}
+
+where \(K_1\), \(K_2\), \(\hat\mu_{K_1}\), \(\hat\mu_{K_2}\),
+\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above for
+equation 2. Estimates of significance of \(\hat{Z}_{12}\) may be
+obtained from a standard normal distribution. As with other two-sample
+tests, \(\hat{Z}_{12}\) is typically considered a two-tailed test,
+however directional (one-tailed) tests may be specified should the
+empirical situation require it (see Adams and Collyer 2016, 2019a).
+
+\hypertarget{empirical-example}{%
+\subsection{\texorpdfstring{\emph{Empirical
+Example}}{Empirical Example}}\label{empirical-example}}
+
+\hypertarget{conclusions-and-implications}{%
+\section{Conclusions and
+Implications}\label{conclusions-and-implications}}
+
+\hypertarget{literature-survey-where-does-this-go}{%
+\subsection{\texorpdfstring{\emph{Literature Survey\ldots{} WHERE DOES
+THIS
+GO???}}{Literature Survey\ldots{} WHERE DOES THIS GO???}}\label{literature-survey-where-does-this-go}}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{simulations}{%
-\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}
-
-By contrast, when characterizing phylogenetic signal with the
-standardized effect size (\(Z_K\)) of \(Kappa\), we found that the
-precision of \(Z_K\) was more stable, as variation across datasets was
-far more consistent across the range of input values. For example, when
-\(n=128\) the precision of \(\lambda_{est}\) (Figure 2A) varied
-considerably more across input levels of phylogenetic signal than did
-the precision of \(Z_K\) for the same datasets (Figure 2B). Further, for
-a given \(n\), the variance in precision estimates of \(Z_K\) was
-considerably smaller than the variance in precision estimates of
-\(\lambda_{est}\) (Fig. 2C,D); implying that estimates of phylogenetic
-signal were more reliable and robust when using \(Z_K\) (for additional
-results see Supporting Information). Finally, it should also be
-recognized that because \(Z_K\) is a standardized effect size, the
-strength of phylogenetic signal is more readily interpretable when using
-this measure, as \(Z_K\) expresses the strength of phylogenetic signal
-in standard deviation units relative to the mean (see Adams and Collyer
-2019b). This further implies that comparisons of the strength of
-phylogenetic signal among phenotypic traits are possible, and may be
-accomplished statistically via a two-sample test that formally compares
-\(Z_K\) across datasets (for comparisons of multivariate effect sizes
-see: Adams and Collyer 2016, 2019a). \hfill\break
-
-{[}insert Figure 2 here{]} \hfill\break
-
-\hypertarget{literature-survey-1}{%
-\subsection{\texorpdfstring{\emph{Literature
-Survey}}{Literature Survey}}\label{literature-survey-1}}
-
-We found 182 manuscripts from 2019 that estimated and reported Pagel's
-lambda values using PGLS methods. These papers averaged 8.527 lambda
-values, ranging from a single lambda estimate up to 71 estimated
-lambdas. Almost exactly half of the published lambda estimates were
-either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3). 73.32\% of
-the published lambdas were estimated using phylogenies with fewer than
-200 tips, and 348 lambda estimates (8.57\% of all published estimates)
-came from phylogenies with fewer than 30 tips. \hfill\break
+\emph{Results}. We found 182 manuscripts from 2019 that estimated and
+reported Pagel's lambda values using PGLS methods. These papers averaged
+8.527 lambda values, ranging from a single lambda estimate up to 71
+estimated lambdas. Almost exactly half of the published lambda estimates
+were either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3).
+73.32\% of the published lambdas were estimated using phylogenies with
+fewer than 200 tips, and 348 lambda estimates (8.57\% of all published
+estimates) came from phylogenies with fewer than 30 tips. \hfill\break
 
@@
 18\% of papers published in 2019 do so.
-
-\hypertarget{results-1}{%
-\section{Results}\label{results-1}}
 

### 2020-04-25 Manuscript/Manuscript-2.tex
vsn 2

@@
 (e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
-Other approaches for interpreting \(\lambda\) are more statistical,
-through the use of confidence intervals (Vandelook et al. 2019) or
-likelihood ratio tests that compare the observed model fit to that
-obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al. 2002;
-Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative comparisons
-of \(\lambda\) across multiple phenotypic traits have also been used to
-infer whether the strength of phylogenetic signal is greater in one
-trait as compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
-Indeed, it seems intuitive to interpret the strength of phylogenetic
-signal in this manner, as \(\lambda\) is a parameter on a bounded scale
-(\(0\to1\)) for which interpretation of its extremal points are
-understood (\(\lambda=0\) represents no phylogenetic signal, while
-\(\lambda=1\) is phylogenetic signal as expected under Brownian motion).
-However, equating values of \(\lambda\) directly to the strength of
-phylogenetic signal presumes two important statistical properties that
-have not been fully explored. \hfill\break
+Other approaches for interpreting \(\lambda\) are more statistical. For
+instance, some have evaluated whether the observed \(\lambda\) differs
+from some expected value through the use of confidence intervals
+(Vandelook et al. 2019) or by performing likelihood ratio tests that
+compare the observed model fit to that obtained when \(\lambda=0\) or
+\(\lambda=1\) (Freckleton et al. 2002; Cooper et al. 2010; Bose et al.
+2019). Additionally, qualitative comparisons of \(\lambda\) estimates
+obtained from multiple phenotypic traits have been used to infer whether
+the strength of phylogenetic signal is greater in one trait as compared
+to another (e.g., Liu et al. 2019; Bai et al. 2019). Indeed, statements
+regarding the strength of phylogenetic signal based on \(\lambda\) are
+rather common in the evolutionary literature. We conducted a literature
+survey in Google.scholar and found that of the 204 papers published in
+2019 that estimated and reported Pagel's \(\lambda\), 40\% interpreted
+the strength of phylogenetic signal for at least one phenotypic trait.
+Additionally, nearly 30\% of the 421 \(\lambda\) estimates between 0.25
+and 0.75 were assigned a strength of signal, and because nearly half of
+the 1,572 values reported were near the limits of the parameter (Figure
+1), this percentage is even higher, as biological interpretation of
+phylogenetic signal at the limits of \(\lambda\) are known. \hfill\break
 
-First, it presumes that values of \(\lambda\) can be precisely
-estimated, as biological inferences regarding the strength of
-phylogenetic signal depend on high accuracy in its estimation.
-Therefore, understanding the precision in estimating \(\lambda\) is
-paramount. One study (Boettiger et al. 2012) found that estimates of
-Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
-when data were simulated on a large phylogeny (\(N=281\)) as compared to
-a small one (\(N=13\)). From this observation it was concluded that
-insufficient data (i.e., the number of species) was the underlying cause
-of the increased variation across parameter estimates (Boettiger et al.
-2012). Indeed, such a pattern is common with statistical estimators, as
-summary statistics and parameters are often more precise at greater
-sample sizes (Cohen 1988). However, this conclusion also assumes that
-the precision of \(\lambda\) remains constant across its range
-(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
-verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
-macroevolutionary studies, at present, we still lack a general
-understanding of the precision with which \(\lambda\) can estimate
-levels of phylogenetic signal in phenotypic datasets. \hfill\break
+{[}insert Figure 1 here{]} \hfill\break
+
+It seems intuitive to interpret the strength of phylogenetic signal
+based on the value of \(\lambda\), as \(\lambda\) is a parameter on a
+bounded scale (\(0\to1\)) for which interpretation of its extremal
+points are understood. Specifically, \(\lambda=0\) represents no
+phylogenetic signal, while \(\lambda=1\) is phylogenetic signal as
+expected under Brownian motion. However, equating values of \(\lambda\)
+directly to the strength of phylogenetic signal presumes two important
+statistical properties that have not been fully explored. First, it
+presumes that values of \(\lambda\) can be precisely estimated, as
+biological inferences regarding the strength of phylogenetic signal
+depend on high accuracy in its estimation. Therefore, understanding the
+precision in estimating \(\lambda\) is paramount. One study (Boettiger
+et al. 2012) found that estimates of Pagel's \(\lambda\) displayed less
+variation (i.e., greater precision) when data were simulated on a large
+phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
+observation it was concluded that insufficient data (i.e., the number of
+species) was the underlying cause of the increased variation across
+parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
+common with statistical estimators, as summary statistics and parameters
+are often more precise at greater sample sizes (Cohen 1988). However,
+this conclusion also assumes that the precision of \(\lambda\) remains
+constant across its range (\(\lambda = 0 \to 1\)); an assumption that to
+date, has not been verified. Thus, despite widespread use of Pagel's
+(1999) \(\lambda\) in macroevolutionary studies, at present, we still
+lack a general understanding of the precision with which \(\lambda\) can
+estimate levels of phylogenetic signal in phenotypic datasets.
+\hfill\break
 
@@
 widely across simulation conditions. Predictably, precision improved as
-the number of species increased (Figure 1). This confirmed earlier
+the number of species increased (Figure 2). This confirmed earlier
 findings of Boettiger et al.~(2012), and adhered to parametric
@@
 was co-estimated with regression parameters in PGLS regression (Figure
-2). Here, regression parameters (\(\beta\)) were accurately estimated,
+3). Here, regression parameters (\(\beta\)) were accurately estimated,
 confirming earlier findings of Revell 2010 (2010) (see Supporting
@@
 
-{[}insert Figure 1 here{]} \hfill\break
+{[}insert Figure 2 here{]} \hfill\break
 
-{[}insert Figure 2 here{]} \hfill\break
+{[}insert Figure 3 here{]} \hfill\break
 
@@
 Adams and Collyer (2019b) recently suggested that the strength of
-phylogenetic signal could be represented an effect size, obtained from
-\(Kappa\) and its empirical sampling distribution from permutation. Here
-we formalize that suggestion, and find an effect size as:
+phylogenetic signal could be represented as an effect size, based on the
+\(Kappa\) statistic and its empirical sampling distribution from
+permutation. Here we formalize that suggestion, and find an effect size
+as:
 
@@
 
+where \(K_{obs}\) is the observed phylogenetic signal, and \(\hat\mu_K\)
+and \(\hat\sigma_K\) are the mean and standard deviation of the
+empirical sampling distribution of \(Kappa\) obtained via permutation.
 Similarly, an effect size based on \(\lambda\) could be envisioned as:
@@
 \begin{align}
-    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
+    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
 \end{align}
 
-Note that under the null hypothesis, \(E(\lambda)=0\), a no phylogenetic
-signal is expected under this condition (Freckleton et al. 2002).
-\hfill\break
+In this case, \(\lambda_{obs}\) and \(\hat\sigma_\lambda\) are
+empirically derived using maximum likelihood. Note also that under the
+null hypothesis, \(E(\lambda)=0\), a no phylogenetic signal is expected
+under this condition (Freckleton et al. 2002). \hfill\break
 
@@
 the precision of each using its variance (\(\sigma^2_{Z_K}\) and
-\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 3. Here two
+\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 4. Here two
 things are evident. First, estimates of \(Z_K\) track the input
-phylogenetic signal in a more linear fashion than to estimates of
+phylogenetic signal in a more linear fashion than do estimates of
 \(Z_\lambda\). Second, the precision of \(Z_K\) is considerably more
-stable, as coefficients of variation for the set of \(\sigma^2_{Z_K}\)
-across input levels of phylogenetic signal for the same sample size was
-an order of magnitude smaller than that observed for
-\(\sigma^2_{Z_\lambda}\) (Figure 3). This implies that estimates of the
+stable as compared with \(Z_\lambda\), as coefficients of variation for
+the set of \(\sigma^2_{Z_K}\) across input levels of phylogenetic signal
+were an order of magnitude smaller for than was observed for
+\(\sigma^2_{Z_\lambda}\) (Figure 4). This implied that estimates of the
 strength of phylogenetic signal were more reliable and robust when using
@@
 
-{[}insert Figure 3 here{]}
+{[}insert Figure 4 here{]}
 
@@
 Implications}\label{conclusions-and-implications}}
-
-\hypertarget{literature-survey-where-does-this-go}{%
-\subsection{\texorpdfstring{\emph{Literature Survey\ldots{} WHERE DOES
-THIS
-GO???}}{Literature Survey\ldots{} WHERE DOES THIS GO???}}\label{literature-survey-where-does-this-go}}
-
-To determine how Pagel's \(\lambda\) is commonly utilized in empirical
-studies, we conducted a literature survey. From Google.scholar we
-obtained a list of all papers published in 2019 that used \(\lambda\);
-resulting in 341 studies. For each study, we extracted all
-\(\lambda_{est}\), the size of the phylogeny (\(n\)), and noted whether
-authors reported confidence intervals or performed significance tests
-assessing difference of \(\lambda_{est}\) from either zero or one. We
-also noted whether biological interpretations based on \(\lambda_{est}\)
-were made, and for studies that reported more than one
-\(\lambda_{est}\), we also noted whether these were compared in some
-manner, and whether such comparisons were accompanied with statistical
-tests between \(\lambda_{est}\).
-
-\emph{Results}. We found 182 manuscripts from 2019 that estimated and
-reported Pagel's lambda values using PGLS methods. These papers averaged
-8.527 lambda values, ranging from a single lambda estimate up to 71
-estimated lambdas. Almost exactly half of the published lambda estimates
-were either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3).
-73.32\% of the published lambdas were estimated using phylogenies with
-fewer than 200 tips, and 348 lambda estimates (8.57\% of all published
-estimates) came from phylogenies with fewer than 30 tips. \hfill\break
-
-{[}insert Figure 4 here{]} \hfill\break
-
-Many of the reviewed manuscripts liberally interpreted the magnitude of
-the estimate lambda, using phrases such as ``strong'' or ``weak''
-phylogentic signal when statistically, all that was clear was a
-difference between the estimated lambda and 0 or 1 respectively. We
-estimated that about 20.49\% of the manuscripts revealed some sort of
-biological interpretation of the magnitude of estimated phylogenetic
-signal that overreached the statistical findings. We also identified
-seven manuscripts as having inappropriately interpreted differences in
-lambda values, indicating that some traits had stronger or weaker signal
-than other traits without the appropriate statistical tests.
-\hfill\break
-
-As is evidenced by macroevolutionary papers published in 2019 papers,
-Pagel's lambda estimation methods are often misused and
-over-interpretted. Despite the urging of Boettiger and colleagues to
-publish confidence intervals with all lambda parameter estimates, only
-18\% of papers published in 2019 do so.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
 
@@
 
-\textbf{Figure 1}. Precision of Pagel's \(\lambda\) across known levels
+\textbf{Figure 1}. Frequency distribution of \(\lambda\) estimates
+published in 2019. The majority of these values were close to 0 or 1,
+and from phylogenies with fewer than 200 taxa. \hfill\break
+
+\textbf{Figure 2}. Precision of Pagel's \(\lambda\) across known levels
 of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
@@
 
-\textbf{Figure 2}. Precision of Pagel's \(\lambda\) when incorporated in
+\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
 phylogenetic regression (\(Y\sim X\)), across known levels of input
@@
 
-\textbf{Figure 3}. Variation in estimates of phylogenetic signal across
+\textbf{Figure 4}. Variation in estimates of phylogenetic signal across
 input levels of phylogenetic signal. (A) Estimates of Pagel's
@@
 across input levels of phylogenetic signal, estimated on phylogenies
-containing differing numbers of species. \hfill\break
-
-\textbf{Figure 4}. Frequency of estimated lambda values published in
-manuscripts in 2019. The majority of these values were close to 0 or 1,
-and from phylogenies with fewer than 200 taxa.
+containing differing numbers of species.
 
@@
 
-\singlespacing \textbf{Figure 1}. Precision of Pagel's \(\lambda\)
+\singlespacing \textbf{Figure 1}. Frequency distribution of \(\lambda\)
+estimates published in 2019. The majority of these values were close to
+0 or 1, and from phylogenies with fewer than 200 taxa.
+
+\newpage
+
+\includegraphics[width=0.95\linewidth]{Fig2}
+
+\singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\)
 across known levels of input phylogenetic signal (\(\lambda_{in}\)) on
@@
 
-\includegraphics[width=0.95\linewidth]{Fig2}
+\includegraphics[width=0.95\linewidth]{Fig3}
 
-\singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\) when
+\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
 incorporated in phylogenetic regression (\(Y\sim X\)), across known
@@
 
-\includegraphics[width=0.95\linewidth]{Fig3}
+\includegraphics[width=0.95\linewidth]{Fig4}
 
-\singlespacing \textbf{Figure 3}. Variation in estimates of phylogenetic
+\singlespacing \textbf{Figure 4}. Variation in estimates of phylogenetic
 signal across input levels of phylogenetic signal. (A) Estimates of
@@
 
-\newpage
-
-\includegraphics[width=0.95\linewidth]{Fig4}
-
-\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values
-published in manuscripts in 2019. The majority of these values were
-close to 0 or 1, and from phylogenies with fewer than 200 taxa.
-
 \end{document}

### 2020-04-27 Data_Analyses/Analysis_script.R
Updating all Figures to new version matching Manuscript2 (all previous figures deleted).

@@ -352,6 +352,32 @@ source("Data_Analyses/CompareZ.r")
 sav_vs_bwl <- compare.Z(kappa_sav, kappa_bwl)
 sav_vs_bwl$pairwise.P
 
+# Figure 5 B and C
+
+Z_SDs <- (sav_vs_bwl$sample.r.sd.pk.stand)/sqrt(length(phylo_pruned$tip.label))
+Interval <- Z_SDs*1.96
+
+png("Figures/Fig5_BC.png", width = 600, height = 300)
+par(mfrow=c(1,2), mar = c(5,4,2,2))
+
+hist(sav_named, main = "SA:V and Body Width Data", xlab = "Value", xlim = c(0,3.1), ylim = c(0,90), yaxs="i")
+hist(bwl_named, main = "SA:V and Body Width Data", xlab = "Value", add= T, col = "black")
+legend("topright",c("SA:V","BW"),fill=c("white","black"), inset = .03)
+box()
+
+plotCI(x = c(1,2), y = sav_vs_bwl$sample.z, xlim = c(.8, 2.2), 
+       xlab = "Morphological Traits", yaxs="i", axes = F,
+       ylim = c(0,10), pch = 19, ylab = "Effect Sizes",
+       uiw = Interval)
+axis(2, at=0:10, labels=0:10)
+axis(1, at=c(1,2), labels = c("SA:V", "Body Width"))
+box()
+
+dev.off()
+
+
+
+
 
 # tropical sav vs temperate sav; z scores comparison here too
 

### 2020-04-27 Data_Analyses/Simulation_script-PB-KappaComparison.R

@@ -77,3 +77,13 @@ length(which(is.na(DataTable_lambda$kappa.z)=="TRUE"))
 file_name <- paste("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_", n, ".csv", sep = "")
 write.csv(DataTable_lambda, file_name, row.names = F)
 
+
+
+# figure: 3 panels
+# A
+# first slide, dump top drawing
+# keep only svl and bwl
+# slide 4: top two panels go right below those. SA:V under the tubes
+
+# B emperical histograms of both traits overlaying 
+# C 2 dots for z scores with CIs

### 2020-04-27 Figures/Figures_Script.R

@@ -1,81 +1,18 @@
 # Figures
-treesizes <-  2^(5:10)
-
-ANOVA.data <- read.csv("Data_Analyses/Munged_Data/ANOVA.fulldataset.csv")
-ANOVA.data <- as.data.frame(ANOVA.data)
-
-Regression.data <- read.csv("Data_Analyses/Munged_Data/Regression.fulldataset.csv")
-Regression.data <- as.data.frame(Regression.data)
-
-## Figure 1 ######
-
-library(cowplot)
-library(ggplot2)
-
-Data32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_32.csv")
-Data64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_64.csv")
-Data128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_128.csv")
-Data256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_256.csv")
-Data512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_512.csv")
-Data1024 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_1024.csv")
-
-data_list <- list(Data32, Data64, Data128,
-                  Data256, Data512, Data1024)
-
-PlotList <- lapply(1:6, function(j){
-   ggplot(data_list[[j]], aes(x = lambda.input, y = lambda.est)) +
-      geom_point() + theme(legend.position= "none", panel.background = element_rect("transparent")) + 
-      xlab("Input Lambda") + ylab("Estimated Lambda") + ggtitle(treesizes[[j]]) 
-})
-
-jpeg("Figures/Fig1.jpeg", res = 120, quality = 100, width = 1000, height = 660)
-plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
-          PlotList[[4]], PlotList[[5]], PlotList[[6]])
-dev.off()
-
-jpeg("Manuscript/Fig1.jpeg", res = 120, quality = 100, width = 1000, height = 660)
-plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
-          PlotList[[4]], PlotList[[5]], PlotList[[6]])
-dev.off()
 
 
-## Figure 2 ####
-kappa_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_128.csv")
+# Fig 1: lit hist ####
 
[…]

### 2020-04-27 Manuscript/Manuscript-3.Rmd
vsn 3

--- 2020-04-24 Manuscript/Manuscript-2.Rmd
@@
 This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
-Other approaches for interpreting $\lambda$ are more statistical, through the use of confidence intervals [@Vandelook2019] or likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
-Likewise, qualitative comparisons of $\lambda$ across multiple phenotypic traits have also been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
-Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion).
-However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
+Other approaches for interpreting $\lambda$ are more statistical.
+For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
+Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
+We conducted a literature survey in Google.scholar and found that of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$, 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.
 \hfill\break
 
+[insert Figure 1 here] \hfill\break
+
+It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
+Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion.
+However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
 First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
@@
 We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
-Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
+Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
 Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data.
@@
 *Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions.
-Predictably, precision improved as the number of species increased (Figure 1).
+Predictably, precision improved as the number of species increased (Figure 2).
 This confirmed earlier findings of Boettiger et al.
@@
 Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$).
-Results were similar when $\lambda$ was co-estimated with regression parameters in PGLS regression (Figure 2).
-Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell 2010 [-@Revell2010] (see Supporting Information).
-However, estimates of phylogenetic signal were not, and the spread of $\lambda_{est}$ was even broader than that observed when $\lambda$ was estimated for only the dependent variable.
+Note that these findings were particularly unsettling when considered in light of our literature survey.
+Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
+Of those, 30% were interpreted directly in terms of the strength of phylogenetic signal that they represented (e.g., small, intermediate, strong signal).
+\hfill\break
+
+Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression, the results of our simulations were quite similar.
+Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
+Estimates of phylogenetic signal were less precise, and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable (Figure 3).
 Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
@@
 
-[insert Figure 1 here] \hfill\break
+[insert Figure 2 here] \hfill\break
 
-[insert Figure 2 here] \hfill\break
+[insert Figure 3 here] \hfill\break
 
@@
 
-Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented an effect size, obtained from $Kappa$ and its empirical sampling distribution from permutation.
-Here we formalize that suggestion, and find an effect size as:
+Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
+Here we formalize that suggestion, resulting in an effect size of:
 
@@
 
+where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $Kappa$ obtained via permutation.
 Similarly, an effect size based on $\lambda$ could be envisioned as:
@@
 \begin{align}
-    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
+    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
 \end{align}
 
-Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002]. \hfill\break
+In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood.
+Note that under the null hypothesis no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition.
+\hfill\break
 
 To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
-Results are found in Figure 3.
+Results are found in Figure 4.
 Here two things are evident.
-First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than to estimates of $Z_\lambda$.
-Second, the precision of $Z_K$ is considerably more stable, as coefficients of variation for the set of $\sigma^2_{Z_K}$ across input levels of phylogenetic signal for the same sample size was an order of magnitude smaller than that observed for $\sigma^2_{Z_\lambda}$ (Figure 3).
-This implies that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
+First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B).
+Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$.
+Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$.
+This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
+Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C,D).
+This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
 \hfill\break
 
-[insert Figure 3 here]
+[insert Figure 4 here]
 
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, it may be of interest to compare such measures across datasets.
-This is useful, for instance, to determine whether the strength of phylogenetic signal is greater in one phenotypic trait as compared with another.
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared with another.
 As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:
@@
 Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
-As with other two-sample tests, $\hat{Z}_{12}$ is typically considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [see @AdamsCollyer2016; @AdamsCollyer2019b].
+Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [see @AdamsCollyer2016; @AdamsCollyer2019b].
+\hfill\break
+
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during their calculation [**S&R??**].
+As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation.
+
+###
+Conceptually, this is akin to the conversion of data to their standard normal deviates prior to their use in multivariate analysis, when the original variables are expressed in different units or scales [see ].
+###
+
+####
+
+mathematical transformation or standardization of the trait dimensions, so
+the variables are expressed in similar units
+n the variables in a multivariate data set are on different scales,
+
+Everitt, Brian; Hothorn, Torsten J (2011), An Introduction to Applied Multivariate Analysis with R, Springer,
+
+Also, Legendre... Adams and Collyer. 
+####
+
+Because of this standardization, the strength of phylogenetic signal across datasets has meaning
+
+something about comparisons having meaning...
+
+As such, characterizing the strength of phylogenetic signal with $Z_K$ facilitates a formal statistical comparison of values across continuous traits of different units, scale, or even dimensionality (e.g., univariate traits using $Kappa$, and multivariate traits using $K_{mult}$).
 
@@
 # Conclusions and Implications
-
-## *Literature Survey... WHERE DOES THIS GO???*
-
-To determine how Pagel's $\lambda$ is commonly utilized in empirical studies, we conducted a literature survey.
-From Google.scholar we obtained a list of all papers published in 2019 that used $\lambda$; resulting in 341 studies.
-For each study, we extracted all $\lambda_{est}$, the size of the phylogeny ($n$), and noted whether authors reported confidence intervals or performed significance tests assessing difference of $\lambda_{est}$ from either zero or one.
-We also noted whether biological interpretations based on $\lambda_{est}$ were made, and for studies that reported more than one $\lambda_{est}$, we also noted whether these were compared in some manner, and whether such comparisons were accompanied with statistical tests between $\lambda_{est}$.
-
-*Results*.
-We found 182 manuscripts from 2019 that estimated and reported Pagel's lambda values using PGLS methods.
-These papers averaged 8.527 lambda values, ranging from a single lambda estimate up to 71 estimated lambdas.
-Almost exactly half of the published lambda estimates were either below 0.05 (25.32%) or above 0.9 (24.74%; Figure 3).
-73.32% of the published lambdas were estimated using phylogenies with fewer than 200 tips, and 348 lambda estimates (8.57% of all published estimates) came from phylogenies with fewer than 30 tips.
-\hfill\break
-
-[insert Figure 4 here] \hfill\break
-
-Many of the reviewed manuscripts liberally interpreted the magnitude of the estimate lambda, using phrases such as "strong" or "weak" phylogentic signal when statistically, all that was clear was a difference between the estimated lambda and 0 or 1 respectively.
-We estimated that about 20.49% of the manuscripts revealed some sort of biological interpretation of the magnitude of estimated phylogenetic signal that overreached the statistical findings.
-We also identified seven manuscripts as having inappropriately interpreted differences in lambda values, indicating that some traits had stronger or weaker signal than other traits without the appropriate statistical tests.
-\hfill\break
-
-As is evidenced by macroevolutionary papers published in 2019 papers, Pagel's lambda estimation methods are often misused and over-interpretted.
-Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so.
-
-# Discussion
 
@@
 
-\textbf{Figure 1}.
+\textbf{Figure 1}. Frequency distribution of $\lambda$ estimates published in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa. \hfill\break
+
+\textbf{Figure 2}.
 Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
@@
 
-\textbf{Figure 2}.
+\textbf{Figure 3}.
 Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
@@
 
-\textbf{Figure 3}.
+\textbf{Figure 4}.
 Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
@@
 (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-\hfill\break
-
-\textbf{Figure 4}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 
@@
 
-\singlespacing \textbf{Figure 1}.
-Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
-As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
+\singlespacing \textbf{Figure 1}.  Frequency distribution of $\lambda$ estimates published in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 
 \newpage
+
 ```{r, echo = FALSE, out.width="95%"}
@@
 \singlespacing \textbf{Figure 2}.
-Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
 As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
@@
 \singlespacing \textbf{Figure 3}.
-Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes.
+As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.
 
@@
 
-\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values published in manuscripts in 2019.
-The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
+\singlespacing \textbf{Figure 4}.
+Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
+(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.

### 2020-04-27 Manuscript/Manuscript-3.tex
vsn 3

--- 2020-04-25 Manuscript/Manuscript-2.tex
@@
 the strength of phylogenetic signal for at least one phenotypic trait.
-Additionally, nearly 30\% of the 421 \(\lambda\) estimates between 0.25
-and 0.75 were assigned a strength of signal, and because nearly half of
-the 1,572 values reported were near the limits of the parameter (Figure
-1), this percentage is even higher, as biological interpretation of
-phylogenetic signal at the limits of \(\lambda\) are known. \hfill\break
+Additionally, because nearly half of the 1,572 \(\lambda\) values
+reported were near the limits of the parameter (Figure 1), this
+percentage is even higher, as biological interpretation of phylogenetic
+signal at the limits of \(\lambda\) are known. \hfill\break
 
@@
 decreased precision when phylogenetic signal is of intermediate
-strength. Additionally, the same \(\lambda_{est}\) may be obtained from
-datasets containing vastly different input levels of phylogenetic
-signal. Thus, \(\lambda\) is not a reliable estimate of the strength of
-phylogenetic signal in phenotypic data. We subsequently derive a
-standardized effect size for measuring the strength of phylogenetic
-signal in phenotypic datasets, and apply the concept to two common
-measures of phylogenetic signal: \(\lambda\) and \(Kappa\). Through
-simulations across a wide range of conditions, we find that the
+strength. Additionally, the same estimated values of \(\lambda\) may be
+obtained from datasets containing vastly different input levels of
+phylogenetic signal. Thus, \(\lambda\) is not a reliable estimate of the
+strength of phylogenetic signal in phenotypic data. We subsequently
+derive a standardized effect size for measuring the strength of
+phylogenetic signal in phenotypic datasets, and apply the concept to two
+common measures of phylogenetic signal: \(\lambda\) and \(Kappa\).
+Through simulations across a wide range of conditions, we find that the
 precision of effect sizes based on \(\lambda\) (\(Z_{\lambda}\)) are
@@
 of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\);
-\(\lambda_{in} = 0.25 \to 0.65\)). Results were similar when \(\lambda\)
-was co-estimated with regression parameters in PGLS regression (Figure
-3). Here, regression parameters (\(\beta\)) were accurately estimated,
-confirming earlier findings of Revell 2010 (2010) (see Supporting
-Information). However, estimates of phylogenetic signal were not, and
-the spread of \(\lambda_{est}\) was even broader than that observed when
-\(\lambda\) was estimated for only the dependent variable. Taken
-together, these findings reveal that \(\lambda_{est}\) does not
-precisely characterize observed levels of phylogenetic signal in
-phenotypic datasets, and that biological interpretations of the strength
-of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
-\hfill\break
+\(\lambda_{in} = 0.25 \to 0.65\)). Note that these findings were
+particularly unsettling when considered in light of our literature
+survey. Over one quarter of the \(\lambda\) estimates obtained in
+empirical studies (421 of 1,572) were between \(\lambda=0.25\) and
+\(\lambda=0.75\) (Figure 1). Of those, 30\% were interpreted directly in
+terms of the strength of phylogenetic signal that they represented
+(e.g., small, intermediate, strong signal). \hfill\break
+
+Finally, when \(\lambda\) was co-estimated with regression parameters in
+PGLS regression, the results of our simulations were quite similar.
+Here, regression parameters (\(\beta\)) were accurately estimated,
+confirming earlier findings of Revell (2010) (see Supporting
+Information). Additionally, estimates of phylogenetic signal were less
+precise, and the spread of \(\lambda_{est}\) was similar to that
+observed when \(\lambda\) was estimated for only the dependent variable
+(Figure 3). Taken together, these findings reveal that \(\lambda_{est}\)
+does not precisely characterize observed levels of phylogenetic signal
+in phenotypic datasets, and that biological interpretations of the
+strength of phylogenetic signal based on \(\lambda\) may be highly
+inaccurate. \hfill\break
 
@@
 \(Kappa\) statistic and its empirical sampling distribution from
-permutation. Here we formalize that suggestion, and find an effect size
-as:
+permutation. Here we formalize that suggestion, resulting in an effect
+size of:
 
@@
 In this case, \(\lambda_{obs}\) and \(\hat\sigma_\lambda\) are
-empirically derived using maximum likelihood. Note also that under the
-null hypothesis, \(E(\lambda)=0\), a no phylogenetic signal is expected
-under this condition (Freckleton et al. 2002). \hfill\break
+empirically derived using maximum likelihood. Note that under the null
+hypothesis no phylogenetic signal is expected (Freckleton et al. 2002),
+and thus \(E(\lambda)=0\) under this condition. \hfill\break
 
@@
 phylogenetic signal in a more linear fashion than do estimates of
-\(Z_\lambda\). Second, the precision of \(Z_K\) is considerably more
-stable as compared with \(Z_\lambda\), as coefficients of variation for
-the set of \(\sigma^2_{Z_K}\) across input levels of phylogenetic signal
-were an order of magnitude smaller for than was observed for
-\(\sigma^2_{Z_\lambda}\) (Figure 4). This implied that estimates of the
-strength of phylogenetic signal were more reliable and robust when using
-\(Z_K\) as compared with \(Z_\lambda\). \hfill\break
+\(Z_\lambda\) (Figure 4A,B). Thus, actual changes in the strength of
+phylogenetic signal are reflected more evenly in the corresponding
+values of the effect size \(Z_K\). Second, the precision of \(Z_K\) is
+considerably more stable as compared with \(Z_\lambda\). This may be
+seen by calculating the coefficients of variation for the set of
+precision estimates (i.e., \(\sigma^2_{Z_K}\) and
+\(\sigma^2_{Z_\lambda}\)) across input levels of phylogenetic signal.
+Here coefficients of variation in the precision of \(Z_K\) were an order
+of magnitude smaller for than for \(Z_\lambda\) (Figure 4C,D). This
+implied that estimates of the strength of phylogenetic signal were more
+reliable and robust when using \(Z_K\) as compared with \(Z_\lambda\).
+\hfill\break
 
@@
 Once the magnitude of phylogenetic signal is characterized using
-\(Z_K\), it may be of interest to compare such measures across datasets.
-This is useful, for instance, to determine whether the strength of
-phylogenetic signal is greater in one phenotypic trait as compared with
-another. As with other effect sizes derived from permutation
-distributions (e.g., Adams and Collyer 2016, 2019a), a two-sample test
-statistic may be found as:
+\(Z_K\), one may wish to compare such measures across datasets, to
+determine whether the strength of phylogenetic signal is significantly
+greater in one phenotypic trait as compared with another. As with other
+effect sizes derived from permutation distributions (e.g., Adams and
+Collyer 2016, 2019a), a two-sample test statistic may be found as:
 
@@
 equation 2. Estimates of significance of \(\hat{Z}_{12}\) may be
-obtained from a standard normal distribution. As with other two-sample
-tests, \(\hat{Z}_{12}\) is typically considered a two-tailed test,
-however directional (one-tailed) tests may be specified should the
-empirical situation require it (see Adams and Collyer 2016, 2019a).
+obtained from a standard normal distribution. Typically,
+\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
+(one-tailed) tests may be specified should the empirical situation
+require it (see Adams and Collyer 2016, 2019a). \hfill\break
+
+One important advantage of the approach advocated here is that the
+resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
+measurement cancel out during their calculation {[}\textbf{S\&R??}{]}.
+As a consequence, the effect sizes represent the strength of
+phylogenetic signal on a common and comparable scale, measured in
+standard deviation units, regardless of the initial units and scale of
+the original phenotypic variables under investigation.
+
+\hypertarget{section}{%
+\subsubsection{}\label{section}}
+
+Conceptually, this is akin to the conversion of data to their standard
+normal deviates prior to their use in multivariate analysis, when the
+original variables are expressed in different units or scales {[}see
+{]}. \#\#\#
+
+\hypertarget{section-1}{%
+\paragraph{}\label{section-1}}
+
+mathematical transformation or standardization of the trait dimensions,
+so the variables are expressed in similar units n the variables in a
+multivariate data set are on different scales,
+
+Everitt, Brian; Hothorn, Torsten J (2011), An Introduction to Applied
+Multivariate Analysis with R, Springer,
+
+Also, Legendre\ldots{} Adams and Collyer. \#\#\#\#
+
+Because of this standardization, the strength of phylogenetic signal
+across datasets has meaning
+
+something about comparisons having meaning\ldots{}
+
+As such, characterizing the strength of phylogenetic signal with \(Z_K\)
+facilitates a formal statistical comparison of values across continuous
+traits of different units, scale, or even dimensionality (e.g.,
+univariate traits using \(Kappa\), and multivariate traits using
+\(K_{mult}\)).
 

### 2020-04-27 Data_Analyses/EmpiricalExample/Foot.SVL.Physig.R

@@ -1,0 +1,52 @@
+# Phylogenetic Signal of Foot Shape and SVL
+
+#TRY A RATIO
+
+# Libraries #### 
+library(ape)
+library(geomorph) # Morphological analyses
+library(geiger) # for pruning trees and datasets to each other (function datatree)
+
+Tree <- read.tree("Bonett_and_Blair_Phylo.tre") # Full Tree
+LinMeas <- read.csv("LinMeas.SpeciesSummaries.csv")
+
+row.names(LinMeas) <- LinMeas$Species
+TreeDataPruned <- treedata(Tree, LinMeas) 
+TreePruned <- TreeDataPruned$phy
+LinMeasOrdered <- LinMeas[match(TreePruned$tip.label, LinMeas$Species),]
+SVL <- LinMeasOrdered$SVL; names(SVL) <- rownames(LinMeasOrdered)
+BWL <- LinMeasOrdered$BWL /SVL; names(BWL) <- rownames(LinMeasOrdered)
+SE <- LinMeasOrdered$SE  /SVL; names(SE) <- rownames(LinMeasOrdered)
+HL <- LinMeasOrdered$HL  /SVL; names(HL) <- rownames(LinMeasOrdered)
+FLL <- LinMeasOrdered$FLL  /SVL; names(FLL) <- rownames(LinMeasOrdered)
+HLL <- LinMeasOrdered$HLL  /SVL; names(HLL) <- rownames(LinMeasOrdered)
+TL <- LinMeasOrdered$TL  /SVL; names(TL) <- rownames(LinMeasOrdered)
+
+
+physignal(A = SVL, phy = TreePruned)
+physignal(A = BWL, phy = TreePruned)
+physignal(A = SE, phy = TreePruned)
+physignal(A = HL, phy = TreePruned)
+physignal(A = FLL, phy = TreePruned)
+physignal(A = HLL, phy = TreePruned)
+physignal(A = TL, phy = TreePruned)
+
+res.BWL <-physignal(A = BWL, phy = TreePruned)
+summary(res.BWL)
+plot(res.BWL)
+length(BWL)            
+
+# Foot Shape Analyses ####
+Tree <- read.tree("Bonett_and_Blair_Phylo.tre") # Full Tree
+Microhabitats <- read.csv("Microhabitats.csv")  # Full Tree Microhabitat Data
+FootLMs <- readland.tps("FootShape.SpeciesSummaries.tps", specID = "ID")
[…]

### 2020-04-27 Data_Analyses/EmpiricalExample/SA.V.physignal.R

@@ -1,0 +1,23 @@
+# Phylogenetic signal of SA:V
+library(stringr)
+library(geiger)
+library(geomorph)
+library(ape)
+
+DF <- read.csv("SAV.Disparity.csv")
+row.names(DF) <- DF$X
+Tree <- read.tree("BB.PrunedToLMData.tre")
+  Tree$tip.label <- str_replace_all(Tree$tip.label, "_", " ")
+DATA <- treedata(Tree, DF, warnings= T)
+  Tree.SAtoV <- DATA$phy
+  Data.SAtoV <- DATA$data[match(Tree.SAtoV$tip.label, row.names(DATA$data)),-1]
+SAtoV <- as.numeric(Data.SAtoV[,1])
+names(SAtoV) <- row.names(Data.SAtoV)
+
+res.SAToV<- physignal(A = SAtoV, phy = Tree.SAtoV)
+summary(res.SAToV)
+plot(res.SAToV)
+
+length(Tree.SAtoV$tip.label)
+plot(Tree.SAtoV)
+

### 2020-04-27 Figures/Figures_Script-DCA.R

@@ -1,0 +1,375 @@
+# Figures
+
+
+# Fig 1: lit hist ####
+
+library(readxl)
+
+tree.sizes <-  treesizes <- 2^(5:10)
+
+review <- read_excel("Literature_Review/1Summary.xlsx", sheet = "Since 2019 Lambda Vals", col_types = "numeric")
+colnames(review)
+published.lambdas <- review[,c(3,4)]
+published.lambdas.pruned <- published.lambdas[-which(published.lambdas > 1 | published.lambdas < 0),]
+published.lambdas.df <- as.data.frame(published.lambdas.pruned)
+colnames(published.lambdas.df) <- c("Published_Lambdas", "Taxa")
+
+published.lambdas.df$Taxa_Bins <- as.character(published.lambdas.df$Taxa)
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa<=tree.sizes[[1]])] <- "Up To 32"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[1]] & published.lambdas.df$Taxa<=tree.sizes[[2]])] <- "33 to 64"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[2]] & published.lambdas.df$Taxa<=tree.sizes[[3]])] <- "65 to 128"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[3]] & published.lambdas.df$Taxa<=tree.sizes[[4]])] <- "129 to 256"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[4]] & published.lambdas.df$Taxa<=tree.sizes[[5]])] <- "257 to 512"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[5]] & published.lambdas.df$Taxa<=tree.sizes[[6]])] <- "513 to 1024"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[6]])] <- "Over 1024"
+
+published.lambdas.df$Taxa_Bins <- as.factor(published.lambdas.df$Taxa_Bins)
+published.lambdas.df$Taxa_Bins <- relevel(published.lambdas.df$Taxa_Bins, "513 to 1024")
+published.lambdas.df$Taxa_Bins <- relevel(published.lambdas.df$Taxa_Bins, "257 to 512")
[…]

### 2020-04-27 Data_Analyses/Analysis_script.R
Updated Fig5_bc.png

@@ -348,33 +348,42 @@ names(bwl_named) <- species_sums_bwl$Species
 kappa_bwl <- physignal(bwl_named, phy = phylo_pruned, iter = 999, print.progress = F)
 kappa_bwl # kappa = 0.2515; z = 2.3597; p = 0.001
 
+faketrait <- sim.char(phy = phylo_pruned, par = 1, nsim = 1)[,,1]
+kappa_fake <- physignal(faketrait, phy = phylo_pruned)
+plot(kappa_fake)
+
 source("Data_Analyses/CompareZ.r")
 sav_vs_bwl <- compare.Z(kappa_sav, kappa_bwl)
 sav_vs_bwl$pairwise.P
 
 # Figure 5 B and C
-
-Z_SDs <- (sav_vs_bwl$sample.r.sd.pk.stand)/sqrt(length(phylo_pruned$tip.label))
-Interval <- Z_SDs*1.96
+Nulls <- cbind(kappa_sav$random.K[-1], kappa_bwl$random.K[-1])
+colnames(Nulls) <- c("SAV.k.null", "BWL.k.null")
+Interval<-abs(qnorm(sav_vs_bwl$sample.r.sd.pk.stand)) # CI are represented without 
 
 png("Figures/Fig5_BC.png", width = 600, height = 300)
 par(mfrow=c(1,2), mar = c(5,4,2,2))
-
-hist(sav_named, main = "SA:V and Body Width Data", xlab = "Value", xlim = c(0,3.1), ylim = c(0,90), yaxs="i")
-hist(bwl_named, main = "SA:V and Body Width Data", xlab = "Value", add= T, col = "black")
-legend("topright",c("SA:V","BW"),fill=c("white","black"), inset = .03)
+ # null distribution of kappas with observed kappas
+hist(Nulls[,1], col = alpha("red",0.5), xlim = c(0,.9), border = "red", breaks = 15, yaxs = "i", xlab = "Null and Observed Kappas", main = "")
+hist(Nulls[,2], col = alpha("black", 0.5), border = "black", xlim = c(0,.9), breaks = 15, yaxs = "i", add = T)
+abline(v = kappa_sav$random.K[1], col = "red", lwd = 3)
+abline(v = kappa_bwl$random.K[1], col = "black", lwd = 3)
+legend("topright",c("SA:V","BW Rel"),fill=c("red","black"), inset = .03)
 box()
 
[…]

### 2020-04-27 Figures/Figures_Script-DCA.R

@@ -149,6 +149,7 @@ Data1024 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_1024.csv"
 ###LOOP THIS OVER EACH LEVEL OF N. 
         # calculating z scores for lambdas 
 
+# add loop here over all tree sizes
         Z.lambda <- matrix(NA, nrow=50, ncol = 21)
         n <- 128
         nsim <- 50
@@ -166,6 +167,8 @@ Data1024 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_1024.csv"
         }
         
         Data128$lambda.est.z <- as.vector(Z.lambda)
+        
+        # end loop
 
 #I THINK for Panel A & B, use N=32
 
@@ -186,10 +189,10 @@ PanelB <- ggplot(Data128, aes(x = lambda.input, y = kappa.z)) +
 # go to all 21 lambda inputs and get variance of lambda z scores, then calculate coefficient of variation of those variances
       
 # LOOP THIS OVER EACH LEVEL OF N.  SO obtain 21 lambda.z.vars at each
-      lambda.z.vars <- apply(Z.lambda,2,var)
+      lambda.z.vars <- apply(Z.lambda,2,var) # loop this over all tree sizes
 
 #Then for each N do:
-      CV.k <- sd(apply(Z.k,2,var))/   mean(apply(Z.k,2,var))*100
+      CV.k <- sd(apply(Z.k,2,var))/   mean(apply(Z.k,2,var))*100 # loop that over all tree sizes
 
 # THEN PLOT all 6 CV for K and all 6 CV for lambda.  (Hmm, perhaps this is a single plot?)      
 
@@ -215,16 +218,18 @@ title(main = list("C", cex = 1, font = 1), adj = 0)
       Z.k <- matrix(Data128$kappa.z, nrow = 50, ncol = 21)
       colnames(Z.k) <- lambdas
       
-      lambda.k.vars <- apply(Z.k,2,var)
+      lambda.k.vars <- apply(Z.k,2,var) # loop over all tree sizes
 
 barplot(lambda.k.vars, col = "darkgray", space = -0.01, axis.lty = 1, ylim = c(0,max(lambda.k.vars)),
         xlab = "Lambda Input", ylab = "Lambda Kappa Score Variances")
 box()
 title(main = list("D", cex = 1, font = 1), adj = 0)
       
[…]

### 2020-04-28 Manuscript/CoverPage-Evolution.Rmd
vsn 3

--- 2020-03-15 CoverPage-Evolution_EKB.Rmd
@@
 ---
-title: MISUSES AND INACCURACY PATTERNS OF PAGEL'S LAMBDA ML ESTIMATION
+title: A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate
 
@@
 \hfill\break
-**Keywords**: Pagel's lambda, phylogenetic signal
-\hfill\break
 
-**Short Title**: Problems with Pagel's lambda \hfill\break
+**Keywords**: phylogenetic signal, effect size, Pagel's lambda \hfill\break
+
+**Short Title**: An Effect Size for Phylogenetic Signal \hfill\break
 

### 2020-04-28 Manuscript/Manuscript-3.Rmd
vsn 3

@@
 
-**Keywords**: phylogenetic signal, effect size, Pagel's lambda \hfill\break
-
-**Short Title**: An Effect Size for Phylogenetic Signal \hfill\break
-
 ```{r setup, include=FALSE}
@@
 # Abstract
-
-{conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)}
 
@@
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-We conducted a literature survey in Google.scholar and found that of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$, 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
 Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.
@@
 
-In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data.
+In this study, we evaluate the precision of Pagel's $\lambda$ for estimating known levels of phylogenetic signal in phenotypic data.
 We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
-We find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not.
-We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength.
+We find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across its range.
+Rather, there is decreased precision when input levels of phylogenetic signal are of intermediate strength.
 Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
 Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data.
-We subsequently derive a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
+We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
 Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
-Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
+We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
@@
 
-All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` [@AdamsOtarola2013; @AdamsGeomorph].
+All analyses were performed in R v3.6.0 [@RCT] using the packages `geiger` [@Harmon2008], `caper` [@Orme2013], `phytools` [@Revell2012], and `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
 R-scripts are found in the Supporting Information.
@@
 However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal.
-Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approached the extremes of its range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$).
+Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal.
+The worst precision was observed at intermediate levels of phylogenetic signal ($\lambda_{in}\approx0.5$), while precision improved as input levels approached the extremes of $\lambda$'s range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$).
 Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values.
@@
 Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$).
-Note that these findings were particularly unsettling when considered in light of our literature survey.
+These findings were particularly unsettling when considered in light of our literature survey.
 Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
-Of those, 30% were interpreted directly in terms of the strength of phylogenetic signal that they represented (e.g., small, intermediate, strong signal).
+This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
 \hfill\break
@@
 Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
-Estimates of phylogenetic signal were less precise, and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable (Figure 3).
+However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
 Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
@@
 The results above demonstate that $\lambda$ is not a reliable estimate of the phylogenetic signal in phenotypic data.
-As such, biological interpretations of the strength of phylogenetic signal, and comparisons of the magnitude of such effects across datasets, are severely compromised when based on this parameter.
-As an alternative, we propose that summary estimates of phylogenetic signal be converted to effect sizes for interpretation and comparison.
+As such, biological interpretations of the strength of phylogenetic signal, and comparisons of the magnitude of such effects across datasets are severely compromised when based on this parameter.
+As an alternative, we propose that summary estimates of phylogenetic signal be converted to effect sizes for interpretation and comparison of the relative strength of phylogenetic signal in phenotypic datasets.
 Statistically, a standardized effect size may be found as:
@@
 where $\theta_{obs}$ is the observed test statistic, $E(\theta)$ is its expected value under the null hypothesis, and $\sigma_\theta$ is its standard error [@Glass1976; @Cohen1988;@Rosenthal1994].
-$Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to a standard normal deviate [@Glass1976; @Kelley2012].
+$Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to its standard normal deviate [@Glass1976; @Kelley2012].
 Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
@@
 
-Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
+Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
 Here we formalize that suggestion, resulting in an effect size of:
@@
 In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood.
-Note that under the null hypothesis no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition.
+Note that under the null hypothesis, no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition.
 \hfill\break
@@
 To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
-Results are found in Figure 4.
+Results are found in Figure 4 (additional results are found in the Supporting Information).
 Here two things are evident.
@@
 
-One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during their calculation [**S&R??**].
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
 As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation.
-
-###
-Conceptually, this is akin to the conversion of data to their standard normal deviates prior to their use in multivariate analysis, when the original variables are expressed in different units or scales [see ].
-###
-
-####
-
-mathematical transformation or standardization of the trait dimensions, so
-the variables are expressed in similar units
-n the variables in a multivariate data set are on different scales,
-
-Everitt, Brian; Hothorn, Torsten J (2011), An Introduction to Applied Multivariate Analysis with R, Springer,
-
-Also, Legendre... Adams and Collyer. 
-####
-
-Because of this standardization, the strength of phylogenetic signal across datasets has meaning
-
-something about comparisons having meaning...
-
-As such, characterizing the strength of phylogenetic signal with $Z_K$ facilitates a formal statistical comparison of values across continuous traits of different units, scale, or even dimensionality (e.g., univariate traits using $Kappa$, and multivariate traits using $K_{mult}$).
+This means that conceivably, the strength of phylogenetic signal may be compared across datsets for phenotypic traits in different units and scale (e.g., behavioral and morphological traits), so long as they are continuous variables.
+Note that this is directly related to the common practice of converting data to their standard normal deviates prior to their use in multivariate analysis, if the original variables are expressed in different units or scales [see @Everitt2011; @LegendreLegendre2012; @AdamsCollyer2019].
 
 ## *Empirical Example*
+
+To demonstrate the utility of $\hat{Z}_{12}$ we performed an analysis of the strength of phylogenetic signal in two phenotypic datasets from species of plethodontid salamander.
+The data were part of a series of studies examining macroevolutionary trends in phenotypic diversification in this group [@Baken2019; @Baken2020].
+Our dataset contained surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) for 305 species (Figure 5A).
+For SA:V, 11 linear body measurements from 2,781 individuals were taken, from which estimates of the surface area and volume of the head, body, and tail were calculated and subsequently combined to arrive at the SA:V for each indiviual [for mathematical details see @Baken2020].
+Species means were then obtained.
+Likewise, body size (SVL) and body width (BW) measurements were taken from 3,371 individuals, and species means of relative body width ($\frac{BW}{SVL}$) were calculated [data from @Baken2019].
+A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the dataset, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species.
+The phylogenetic signal in each trait was then characterized using $Kappa$, which was converted to its effect size ($Z_K$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
+Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**).
+\hfill\break
+
+*Results.* Both SA:V and relative body width displayed significant phylogenetic signal ($Kappa_{SA:V}=0.7608$; $P=0.001$; $Kappa_{BW/SVL}=0.2515$; $P=0.001$).
+For both phenotypic traits, $K_{obs}$ differed markedly from their corresponding permutation distributions, which were found to overlap almost perfectly (Figure 5B).
+However, while both traits displayed significant phylogenetic signal, there was nearly a four-fold difference in the magnitude of their effect sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C).
+Using the two-sample test statistic above, this difference was found to be highly significant ($\hat{Z}_{12}=4.13$; $P=0.000036$).
+Thus it may be concluded that SA:V displays significantly stronger phylogenetic signal than does relative body width, and that shared evolutionary history has strongly influenced trait covariation among taxa for SA:V.
+Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020].
+Thus, because of this macroevolutionary association, strong phylogenetic signal in SA:V is to be expected.
 

### 2020-04-28 Manuscript/Manuscript-3.tex
vsn 3

@@
 
-\textbf{Keywords}: phylogenetic signal, effect size, Pagel's lambda
-\hfill\break
-
-\textbf{Short Title}: An Effect Size for Phylogenetic Signal
-\hfill\break
-
 \hypertarget{abstract}{%
 \section{Abstract}\label{abstract}}
-
-\{conclusion holds: interpreting the regression is not appreciably
-different (in terms of slopes and f values)\}
 
@@
 regarding the strength of phylogenetic signal based on \(\lambda\) are
-rather common in the evolutionary literature. We conducted a literature
-survey in Google.scholar and found that of the 204 papers published in
-2019 that estimated and reported Pagel's \(\lambda\), 40\% interpreted
-the strength of phylogenetic signal for at least one phenotypic trait.
-Additionally, because nearly half of the 1,572 \(\lambda\) values
-reported were near the limits of the parameter (Figure 1), this
-percentage is even higher, as biological interpretation of phylogenetic
-signal at the limits of \(\lambda\) are known. \hfill\break
+rather common in the evolutionary literature. For instance, of the 204
+papers published in 2019 that estimated and reported Pagel's \(\lambda\)
+(found in Google.scholar), 40\% interpreted the strength of phylogenetic
+signal for at least one phenotypic trait. Additionally, because nearly
+half of the 1,572 \(\lambda\) values reported were near the limits of
+the parameter (Figure 1), this percentage is even higher, as biological
+interpretation of phylogenetic signal at the limits of \(\lambda\) are
+known. \hfill\break
 
@@
 
-In this study, we evaluate the precision of Pagel's \(\lambda\) in
+In this study, we evaluate the precision of Pagel's \(\lambda\) for
 estimating known levels of phylogenetic signal in phenotypic data. We
@@
 levels of phylogenetic signal, and under what circumstances. We find
-that while PGLS parameters (e.g., \(\beta\)) are accurately estimated
-with the inclusion of phylogenetic signal, estimates of \(\lambda\) are
-not. We also find that estimates of \(\lambda\) vary widely for a given
-input value of phylogenetic signal, and that the precision in estimating
-\(\lambda\) is not constant across the range of input signal, with
-decreased precision when phylogenetic signal is of intermediate
-strength. Additionally, the same estimated values of \(\lambda\) may be
-obtained from datasets containing vastly different input levels of
-phylogenetic signal. Thus, \(\lambda\) is not a reliable estimate of the
-strength of phylogenetic signal in phenotypic data. We subsequently
-derive a standardized effect size for measuring the strength of
-phylogenetic signal in phenotypic datasets, and apply the concept to two
-common measures of phylogenetic signal: \(\lambda\) and \(Kappa\).
-Through simulations across a wide range of conditions, we find that the
-precision of effect sizes based on \(\lambda\) (\(Z_{\lambda}\)) are
-less reliable than that those based on \(Kappa\) (\(Z_K\)), implying
-that \(Z_K\) is a more robust effect size measure. Additionally, we
-propose a two-sample test statistic that may be used to compare the
-strength of phylogenetic signal among datasets, and provide an empirical
-example to demonstrate its use. We conclude that estimates of
-phylogenetic signal using Pagel's \(\lambda\) are often inaccurate, and
-thus interpreting strength of phylogenetic signal in phenotypic datasets
-based on this measure is compromised. By contrast, effect sizes obtained
-from \(Kappa\) hold promise for characterizing phylogenetic signal, and
-for comparing the strength of phylogenetic signal across datasets.
+that estimates of \(\lambda\) vary widely for a given input value of
+phylogenetic signal, and that the precision in estimating \(\lambda\) is
+not constant across its range. Rather, there is decreased precision when
+input levels of phylogenetic signal are of intermediate strength.
+Additionally, the same estimated values of \(\lambda\) may be obtained
+from datasets containing vastly different input levels of phylogenetic
+signal. Thus, \(\lambda\) is not a reliable estimate of the strength of
+phylogenetic signal in phenotypic data. We then describe a standardized
+effect size for measuring the strength of phylogenetic signal in
+phenotypic datasets, and apply the concept to two common measures of
+phylogenetic signal: \(\lambda\) and \(Kappa\). Through simulations
+across a wide range of conditions, we find that the precision of effect
+sizes based on \(\lambda\) (\(Z_{\lambda}\)) are less reliable than that
+those based on \(Kappa\) (\(Z_K\)), implying that \(Z_K\) is a more
+robust effect size measure. We also propose a two-sample test statistic
+that may be used to compare the strength of phylogenetic signal among
+datasets, and provide an empirical example to demonstrate its use. We
+conclude that estimates of phylogenetic signal using Pagel's \(\lambda\)
+are often inaccurate, and thus interpreting strength of phylogenetic
+signal in phenotypic datasets based on this measure is compromised. By
+contrast, effect sizes obtained from \(Kappa\) hold promise for
+characterizing phylogenetic signal, and for comparing the strength of
+phylogenetic signal across datasets.
 
@@
 packages \texttt{geiger} (Harmon et al. 2008), \texttt{caper} (Orme et
-al. 2013), \texttt{phytools} (Revell 2012), and \texttt{geomorph} (Adams
-and Otárola-Castillo 2013; Adams et al. 2020). R-scripts are found in
-the Supporting Information. \hfill\break
+al. 2013), \texttt{phytools} (Revell 2012), and \texttt{geomorph} 3.2.1
+(Adams and Otárola-Castillo 2013; Adams et al. 2020). R-scripts are
+found in the Supporting Information. \hfill\break
 
@@
 phylogenetic signal. Importantly, the precision of \(\lambda_{est}\) was
-not uniform across all levels of phylogenetic signal, with the worst
-precision at intermediate levels of signal (\(\lambda_{in}\approx0.5\)),
-and improved precision as input levels approached the extremes of its
-range (i.e., \(\lambda_{in}\rightarrow0\) \&
-\(\lambda_{in}\rightarrow1\)). Thus, estimates of \(\lambda\) were least
-reflective of the true input signal at intermediate values.
-Additionally, even at large levels of species richness, we found that
-the range of \(\lambda_{est}\) still encompassed a substantial portion
-of possible values (e.g., \(n=512\); \(\lambda_{in}=0.5\):
-\(\lambda_{est} = 0.32\to 0.68\)). Likewise, the same \(\lambda_{est}\)
-could be obtained from datasets containing vastly different input levels
-of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\);
-\(\lambda_{in} = 0.25 \to 0.65\)). Note that these findings were
-particularly unsettling when considered in light of our literature
-survey. Over one quarter of the \(\lambda\) estimates obtained in
-empirical studies (421 of 1,572) were between \(\lambda=0.25\) and
-\(\lambda=0.75\) (Figure 1). Of those, 30\% were interpreted directly in
-terms of the strength of phylogenetic signal that they represented
-(e.g., small, intermediate, strong signal). \hfill\break
+not uniform across all levels of phylogenetic signal. The worst
+precision was observed at intermediate levels of phylogenetic signal
+(\(\lambda_{in}\approx0.5\)), while precision improved as input levels
+approached the extremes of \(\lambda\)'s range (i.e.,
+\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
+estimates of \(\lambda\) were least reflective of the true input signal
+at intermediate values. Additionally, even at large levels of species
+richness, we found that the range of \(\lambda_{est}\) still encompassed
+a substantial portion of possible values (e.g., \(n=512\);
+\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
+same \(\lambda_{est}\) could be obtained from datasets containing vastly
+different input levels of phylogenetic signal (e.g., \(n=512\);
+\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). These
+findings were particularly unsettling when considered in light of our
+literature survey. Over one quarter of the \(\lambda\) estimates
+obtained in empirical studies (421 of 1,572) were between
+\(\lambda=0.25\) and \(\lambda=0.75\) (Figure 1). This range reflected
+the region that our simulations identified as being the least reliable
+in terms of accurately characterizing levels of phylogenetic signal, yet
+30\% of these empirical estimates were explicitly interpreted in terms
+of the strength of phylogenetic signal that they represented (i.e.,
+weak, intermediate, strong phylogenetic signal). \hfill\break
 
@@
 confirming earlier findings of Revell (2010) (see Supporting
-Information). Additionally, estimates of phylogenetic signal were less
-precise, and the spread of \(\lambda_{est}\) was similar to that
-observed when \(\lambda\) was estimated for only the dependent variable
-(Figure 3). Taken together, these findings reveal that \(\lambda_{est}\)
-does not precisely characterize observed levels of phylogenetic signal
-in phenotypic datasets, and that biological interpretations of the
-strength of phylogenetic signal based on \(\lambda\) may be highly
-inaccurate. \hfill\break
+Information). However, estimates of phylogenetic signal (\(\lambda\))
+were less precise (Figure 3), and the spread of \(\lambda_{est}\) was
+similar to that observed when \(\lambda\) was estimated for only the
+dependent variable, as in Figure 2. Taken together, these findings
+reveal that \(\lambda_{est}\) does not precisely characterize observed
+levels of phylogenetic signal in phenotypic datasets, and that
+biological interpretations of the strength of phylogenetic signal based
+on \(\lambda\) may be highly inaccurate. \hfill\break
 
@@
 interpretations of the strength of phylogenetic signal, and comparisons
-of the magnitude of such effects across datasets, are severely
+of the magnitude of such effects across datasets are severely
 compromised when based on this parameter. As an alternative, we propose
 that summary estimates of phylogenetic signal be converted to effect
-sizes for interpretation and comparison. Statistically, a standardized
-effect size may be found as:
+sizes for interpretation and comparison of the relative strength of
+phylogenetic signal in phenotypic datasets. Statistically, a
+standardized effect size may be found as:
 
@@
 \(Z_{\theta}\) expresses the magnitude of the effect in \(\theta_{obs}\)
-by transforming the original test statistic to a standard normal deviate
-(Glass 1976; Kelley and Preacher 2012). Typically, \(\theta_{obs}\) and
-\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
-obtained from the distribution of \(\theta\) derived from parametric
-theory. However, recent advances in resampling theory (Collyer et al.
-2015; Adams and Collyer 2016, 2019a) have shown that \(E(\theta)\) and
-\(\sigma_\theta\) may also be obtained from an empirical sampling
-distribution of \(\theta\) obtained from permutation procedures.
-\hfill\break
+by transforming the original test statistic to its standard normal
+deviate (Glass 1976; Kelley and Preacher 2012). Typically,
+\(\theta_{obs}\) and \(\sigma_\theta\) are estimated from the data,
+while \(E(\theta)\) is obtained from the distribution of \(\theta\)
+derived from parametric theory. However, recent advances in resampling
+theory (Collyer et al. 2015; Adams and Collyer 2016, 2019a) have shown
+that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from an
+empirical sampling distribution of \(\theta\) obtained from permutation
+procedures. \hfill\break
 
-Adams and Collyer (2019b) recently suggested that the strength of
-phylogenetic signal could be represented as an effect size, based on the
-\(Kappa\) statistic and its empirical sampling distribution from
-permutation. Here we formalize that suggestion, resulting in an effect
-size of:
+Adams and Collyer (2019b) suggested that the strength of phylogenetic
+signal could be represented as an effect size, based on the \(Kappa\)
+statistic and its empirical sampling distribution from permutation. Here
+we formalize that suggestion, resulting in an effect size of:
 
@@
 empirically derived using maximum likelihood. Note that under the null
-hypothesis no phylogenetic signal is expected (Freckleton et al. 2002),
+hypothesis, no phylogenetic signal is expected (Freckleton et al. 2002),
 and thus \(E(\lambda)=0\) under this condition. \hfill\break
@@
 the precision of each using its variance (\(\sigma^2_{Z_K}\) and
-\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 4. Here two
-things are evident. First, estimates of \(Z_K\) track the input
-phylogenetic signal in a more linear fashion than do estimates of
-\(Z_\lambda\) (Figure 4A,B). Thus, actual changes in the strength of
-phylogenetic signal are reflected more evenly in the corresponding
-values of the effect size \(Z_K\). Second, the precision of \(Z_K\) is
-considerably more stable as compared with \(Z_\lambda\). This may be
-seen by calculating the coefficients of variation for the set of
-precision estimates (i.e., \(\sigma^2_{Z_K}\) and
-\(\sigma^2_{Z_\lambda}\)) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of \(Z_K\) were an order
-of magnitude smaller for than for \(Z_\lambda\) (Figure 4C,D). This
-implied that estimates of the strength of phylogenetic signal were more
-reliable and robust when using \(Z_K\) as compared with \(Z_\lambda\).
-\hfill\break
+\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 4 (additional
+results are found in the Supporting Information). Here two things are
+evident. First, estimates of \(Z_K\) track the input phylogenetic signal
+in a more linear fashion than do estimates of \(Z_\lambda\) (Figure
+4A,B). Thus, actual changes in the strength of phylogenetic signal are
+reflected more evenly in the corresponding values of the effect size
+\(Z_K\). Second, the precision of \(Z_K\) is considerably more stable as
+compared with \(Z_\lambda\). This may be seen by calculating the
+coefficients of variation for the set of precision estimates (i.e.,
+\(\sigma^2_{Z_K}\) and \(\sigma^2_{Z_\lambda}\)) across input levels of
+phylogenetic signal. Here coefficients of variation in the precision of
+\(Z_K\) were an order of magnitude smaller for than for \(Z_\lambda\)
+(Figure 4C,D). This implied that estimates of the strength of
+phylogenetic signal were more reliable and robust when using \(Z_K\) as
+compared with \(Z_\lambda\). \hfill\break
 
@@
 resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
-measurement cancel out during their calculation {[}\textbf{S\&R??}{]}.
-As a consequence, the effect sizes represent the strength of
+measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
+2012). As a consequence, the effect sizes represent the strength of
 phylogenetic signal on a common and comparable scale, measured in
 standard deviation units, regardless of the initial units and scale of
-the original phenotypic variables under investigation.
-
-\hypertarget{section}{%
-\subsubsection{}\label{section}}
-
-Conceptually, this is akin to the conversion of data to their standard
-normal deviates prior to their use in multivariate analysis, when the
-original variables are expressed in different units or scales {[}see
-{]}. \#\#\#
-
-\hypertarget{section-1}{%
-\paragraph{}\label{section-1}}
-
-mathematical transformation or standardization of the trait dimensions,
-so the variables are expressed in similar units n the variables in a
-multivariate data set are on different scales,
-
-Everitt, Brian; Hothorn, Torsten J (2011), An Introduction to Applied
-Multivariate Analysis with R, Springer,
-
-Also, Legendre\ldots{} Adams and Collyer. \#\#\#\#
-
-Because of this standardization, the strength of phylogenetic signal
-across datasets has meaning
-
-something about comparisons having meaning\ldots{}
-
-As such, characterizing the strength of phylogenetic signal with \(Z_K\)
-facilitates a formal statistical comparison of values across continuous
-traits of different units, scale, or even dimensionality (e.g.,
-univariate traits using \(Kappa\), and multivariate traits using
-\(K_{mult}\)).
+the original phenotypic variables under investigation. This means that
+conceivably, the strength of phylogenetic signal may be compared across
+datsets for phenotypic traits in different units and scale (e.g.,
+behavioral and morphological traits), so long as they are continuous
+variables. Note that this is directly related to the common practice of
+converting data to their standard normal deviates prior to their use in
+multivariate analysis, if the original variables are expressed in
+different units or scales (see Everitt and Hothorn 2011; Legendre and
+Legendre 2012; Adams and Collyer 2019b).
 
@@
 Example}}{Empirical Example}}\label{empirical-example}}
+
+To demonstrate the utility of \(\hat{Z}_{12}\) we performed an analysis
+of the strength of phylogenetic signal in two phenotypic datasets from
+species of plethodontid salamander. The data were part of a series of
+studies examining macroevolutionary trends in phenotypic diversification
+in this group (Baken and Adams 2019; Baken et al. 2020). Our dataset
+contained surface area to volume ratios (SA:V) and relative body width
+(\(\frac{BW}{SVL}\)) for 305 species (Figure 5A). For SA:V, 11 linear
+body measurements from 2,781 individuals were taken, from which
+estimates of the surface area and volume of the head, body, and tail
+were calculated and subsequently combined to arrive at the SA:V for each
+indiviual (for mathematical details see Baken et al. 2020). Species
+means were then obtained. Likewise, body size (SVL) and body width (BW)
+measurements were taken from 3,371 individuals, and species means of
+relative body width (\(\frac{BW}{SVL}\)) were calculated (data from
+Baken and Adams 2019). A time-dated molecular phylogeny for the group
+(Bonett and Blair 2017) was then pruned to match the species in the
+dataset, resulting in a phylogeny and corresponding phenotypic dataset
+containing 305 species. The phylogenetic signal in each trait was then
+characterized using \(Kappa\), which was converted to its effect size
+(\(Z_K\)) using \texttt{geomorph} 3.2.1 (Adams and Otárola-Castillo
+2013; Adams et al. 2020). Finally, the strength of phylogenetic signal
+was compared across traits using \(\hat{Z}_{12}\) as described above
+(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
+acceptance}). \hfill\break
+
+\emph{Results.} Both SA:V and relative body width displayed significant
+phylogenetic signal (\(Kappa_{SA:V}=0.7608\); \(P=0.001\);
+\(Kappa_{BW/SVL}=0.2515\); \(P=0.001\)). For both phenotypic traits,
+\(K_{obs}\) differed markedly from their corresponding permutation
+distributions, which were found to overlap almost perfectly (Figure 5B).
+However, while both traits displayed significant phylogenetic signal,
+there was nearly a four-fold difference in the magnitude of their effect
+sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C).
+Using the two-sample test statistic above, this difference was found to
+be highly significant (\(\hat{Z}_{12}=4.13\); \(P=0.000036\)). Thus it
+may be concluded that SA:V displays significantly stronger phylogenetic
+signal than does relative body width, and that shared evolutionary
+history has strongly influenced trait covariation among taxa for SA:V.
+Biologically, this observation corresponds with the fact that tropical
+species -- which form a monophyletic group within plethodontids --
+display greater variation in SA:V which covaries with disparity in their
+climatic niches (Baken et al. 2020). Thus, because of this
+macroevolutionary association, strong phylogenetic signal in SA:V is to
+be expected.
 
@@
 
+\leavevmode\hypertarget{ref-Baken2019}{}%
+Baken, E. K., and D. C. Adams. 2019. Macroevolution of arboreality in
+salamanders. Ecology and Evolution 9:7005--7016.
+
+\leavevmode\hypertarget{ref-Baken2020}{}%
+Baken, E. K., L. E. Mellenthin, and D. C. Adams. 2020. Macroevolution of
+desiccation‐related morphology in plethodontid salamanders as inferred
+from a novel surface area to volume ratio estimation approach. Evolution
+74:476--486.
+
 \leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
@@
 67:2240--2251.
+
+\leavevmode\hypertarget{ref-Bonett2017}{}%
+Bonett, R. M., and A. L. Blair. 2017. Evidence for complex life cycle
+constraints on salamander body form diversification. Proceedings of the
+National Academy of Sciences, U.S.A. 114:9936--9941.
 
@@
 35:673--679.
+
+\leavevmode\hypertarget{ref-Everitt2011}{}%
+Everitt, B., and T. Hothorn. 2011. An introduction to applied
+multivariate analysis with r. Springer Science \& Business Media.
 
@@
 biology 59:245--261.
+
+\leavevmode\hypertarget{ref-LegendreLegendre2012}{}%
+Legendre, P., and L. Legendre. 2012. Numerical ecology. 3rd ed.
+Elsevier, Amsterdam.
 
@@
 \emph{in} L. V. Cooper H Hedges, ed. Russell Sage Foundation.
+
+\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
+Sokal, R. R., and F. J. Rohlf. 2012. Biometry. 4th ed. W.H. Freeman \&
+Co., San Francisco.
 

### 2020-04-28 Manuscript/Manuscript-3.Rmd
vsn 3

@@
 \textbf{Figure 4}.
-Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+Variation in effect size estimates of phylogenetic signal across input levels of phylogenetic signal.
+(A) Estimates $Z_\lambda$ for data simulated on phylogenies with 32 taxa ($n=32$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 32 taxa ($n=32$), (C) Coefficients of variation of precision estimates of $Z_\lambda$ and $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+
+\textbf{Figure 5}.
+(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
 
@@
 \singlespacing \textbf{Figure 4}.
-Variation in estimates of phylogenetic signal across input levels of phylogenetic signal.
-(A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
-(D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+Variation in effect size estimates of phylogenetic signal across input levels of phylogenetic signal.
+(A) Estimates $Z_\lambda$ for data simulated on phylogenies with 32 taxa ($n=32$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 32 taxa ($n=32$), (C) Coefficients of variation of precision estimates of $Z_\lambda$ and $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig5.png")
+
+```
+
+\singlespacing \textbf{Figure 5}.
+(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).

### 2020-04-30 Manuscript/Manuscript-4.Rmd
vsn 4

--- 2020-04-28 Manuscript/Manuscript-3.Rmd
@@
 
+Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits.
+However, analytical tools for such comparisons have largely remained underdeveloped.
+In this study, we evaluated the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal.
+We find that the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised.
+We then propose a standardized effect size based on $Kappa$ ($Z_K$), which measures the strength of phylogenetic signal, and places it on a common scale for statistical comparison.
+Tests based on $Z_K$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
+Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal various phenotypic traits, even when those traits are found in different evolutionary lineages or of in different units or scale.
+
 \newpage
@@
 For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
-Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.
+Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
+
 \hfill\break
@@
 Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
-However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
-Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
+However, this conclusion also implies that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
+Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
 \hfill\break
@@
 Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
-Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analysis [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
+Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
 Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
@@
 Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
-Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data.
+Thus, $\lambda$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data.
 We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
-Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
+Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
 We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
@@
 Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
-Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal corresponding reflecting Brownian motion (when $\lambda_{in} =1$).
-For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the precision of $\lambda$ using the variance ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness.
+Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal reflecting Brownian motion (when $\lambda_{in} =1$).
+For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the variance of $\lambda$ ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness as an estimate of precision.
 \hfill\break
@@
 We also evaluated the precision of $\lambda$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$).
-Here, an independent variable $X$ was simulated on each phylogeny under a Brownian motion model of evolution (for PGLS regression).
+Here, an independent variable $X$ was simulated on each rescaled phylogeny under a Brownian motion model of evolution (for PGLS regression).
 For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny.
@@
 This was accomplished as: $Y=\beta{X}+\epsilon$.
-Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution: $\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: [see @Revell2010 for a similar simulation design].
+Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution on each rescaled phylogeny: $\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$: [see @Revell2010 for a similar simulation design].
 The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
-Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness were then observed.
+We then calculated precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness.
 \hfill\break
@@
 Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
-This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
+This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
 \hfill\break
 
-Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression, the results of our simulations were quite similar.
+Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression and ANOVA, the results of our simulations were quite similar.
 Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
-However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
+However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; see also Supporting Information), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
 Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
@@
 
-Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
+Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
 Here we formalize that suggestion, resulting in an effect size of:
@@
 \begin{align}
-    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
+    Z_{K}=\frac{log(K_{obs})-\hat\mu_{log(K)}}{\hat\sigma_{log(K)}}
 \end{align}
 
-where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $Kappa$ obtained via permutation.
+where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation.
+Here the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018].
 Similarly, an effect size based on $\lambda$ could be envisioned as:
@@
 
-In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood.
-Note that under the null hypothesis, no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition.
+In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood, as permutation approaches have not been developed for evaluating $\lambda$.
+Note also that under the null hypothesis, no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition.
 \hfill\break
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C,D).
+Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C).
 This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared with another.
-As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared to another.
+As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:
 
@@
 \hfill\break
-
-One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation.
-This means that conceivably, the strength of phylogenetic signal may be compared across datsets for phenotypic traits in different units and scale (e.g., behavioral and morphological traits), so long as they are continuous variables.
-Note that this is directly related to the common practice of converting data to their standard normal deviates prior to their use in multivariate analysis, if the original variables are expressed in different units or scales [see @Everitt2011; @LegendreLegendre2012; @AdamsCollyer2019].
 
@@
 
-# Conclusions and Implications
+# Discussion
 
-1: summary paragraph
-
-2: expand on Lambda..
-lambda innacurate, not precise, level of precision varies with input physig (worse in mid-range).
-NEW RESULT.
-We are first to show this.
-NOTE: pattern is obvious with reflection.
-Since it is a 'bounded' parameter estimation should be best at the extremes...
-(state this?)..
-hmm.
-
-Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
-
-3: By contrast, effect size Z-K useful, equally precise across range of values.
-Can be used to characterize the strength of physignal, and because robust to input levels, etc.
-may be used to compare across datasets.
-
-Somewhere, recognize that this is somewhat 'backwards' from prior recommendations where Kappa had somewhat lower performance in terms of type I and type II error (which??
-I forget).
-However, recall that those studies did not examine the precision of the estimates.
-Nor was Z-k included, because it was not yet invented.
-So Use of Z-k should make good sense here.
-
-Closing paragraph.
-
+It is common in comparative evolutionary studies to characterize the phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has generated trait covariation among taxa.
+However, while numerous analytical approaches may be used to quantify phylogenetic signal [e.g., @Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999], methods that explicitly measure the strength of phylogenetic signal, or facilitate comparisons among datasets, have remained underdeveloped.
+In this study, we evaluated the precision of one common measure, Pagel's $\lambda$, and explored its efficacy for characterizing the strength of phylogenetic signal in phenotypic data.
+Using computer simulations, we found that the precision of $\lambda$ increased with increasing sample sizes; a pattern noted previously [@Boettiger_et_al2012], and one that conformed with parametric statistical theory [@Cohen1988].
+However, we also found that vastly different $\lambda$ estimates could be obtained from data containing the same level of phylogenetic signal, and that similar $\lambda$ estimates may be obtained from data containing differing levels of phylogenetic signal.
+Further, the precision of $\lambda$ varied with the strength of phylogenetic signal, where lower precision was observed when in data whose phylogenetic signal was of intemediate strength.
+From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and that biological interpretations of the strength of signal based on this parameter may innacurately characterize such effects.
 \hfill\break
 
-More discussion paragraphs
+As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal.
+$Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
+We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
+First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal.
+Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomite changes in the values of $Z_K$.
+Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
+We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect.
+\hfill\break
+
+Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another.
+Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003].
+However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
+Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale.
+By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+Thus when viewed from this perspective, the approach developed here is statistically aligned with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons.
+As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
+\hfill\break
+
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
+Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$.
+For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
+Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.
+\hfill\break
+
+** Finally... Need Closing paragraph. **
 

### 2020-05-01 Manuscript/SupportingInfo-EKB.Rmd
Supporting figures mostly finished, supporting information markdown started.

--- 2020-04-20 Manuscript/SupportingInfo.Rmd
@@
 
-**FOR DCA**
-
-1: balanced/symmetric trees
-2: Kappa and K_Z
-3: Beta from PGLS
-
 Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.
@@
 
-*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).
+*Results*.
+As found with pure-birth phylogenies, estimates of $\lambda$ varied most dramatically in simulations with fewer species and at intermediate values of lambda (Fig.
+S1, S2).
+Pectinate trees showed an interesting tendency to underestimate $/lambda$ as 0 for input $/lambda$ values.
+Most dramatically, some simulations on the largest pectinate phylogeny ($n = 1024$) estimated $/lambda = 0$ for input values as high as 0.75 (Fig.
+S2).
 
 ```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+include_graphics("FigS1.png")
 
@@
 ```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+include_graphics("FigS2.png")
 
@@
 
-## Literature Survey
+## Simulations of phylogenetic regression and ANOVA
+
+```{r setup, include=FALSE}
+  library(knitr)
+  knitr::opts_chunk$set(echo = TRUE)
+```
+
+We analyzed the reliability of parameter estimations when used in phylogenetic regression and ANOVA.
+This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$).
+For each $\lambda$ input value, independent variables were then generated with a known relationship ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.
+
+*Results* As with the results from the regression analyses in the main text, we found limmited $\lambda$ estimation precision in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig.
+S3).
+Slope was reliably estimated across tree sizes and known $\lambda$ values (Fig.
+S4).
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS3.png")
+
+```
+
+\textbf{Figure S3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic ANOVA. Results are from input $\beta$ values of $0.5$. \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS4.png")
+
+```
+
+\textbf{Figure S4}. Precision of slope estimates across known input $\beta$ values of phylogenetic regression. Results shown include all input $\lambda$ values. \hfill\break
 

### 2020-05-01 Data_Analyses/Simulation_script-Bal-KappaComparison.R

@@ -15,44 +15,13 @@ lambdas_expanded <- rep(lambdas, each = nsim)
 fullsimlength <- length(lambdas_expanded)
 
 # Making all the trees ####
-library(TreeSimGM)
+phy.list <- lapply(1:length(treesizes), function(i) stree(n = treesizes[i], type = "balanced"))
+for (i in 1:length(treesizes)) {
+  phy.list[[i]]$edge.length <- rep(1, (length(phy.list[[i]]$tip.label) - 1)*2)
+}
 
-asymmetric_tree <- sim.age(age=63, numbsim=1, waitsp=function()2, waitext=function()2.5, symmetric=FALSE)
-length(asymmetric_tree[[1]]$tip.label)
-tree_ladder_32 <- force.ultrametric(asymmetric_tree[[1]], method = "extend")
-for (i in 5:12) {tree_ladder_32[[5]] <- NULL}
-
-asymmetric_tree <- sim.age(age=127, numbsim=1, waitsp=function()2, waitext=function()2.5, symmetric=FALSE)
-length(asymmetric_tree[[1]]$tip.label)
-tree_ladder_64 <- force.ultrametric(asymmetric_tree[[1]], method = "extend")
-for (i in 5:12) {tree_ladder_64[[5]] <- NULL}
-
-asymmetric_tree <- sim.age(age=255, numbsim=1, waitsp=function()2, waitext=function()2.5, symmetric=FALSE)
-length(asymmetric_tree[[1]]$tip.label)
-tree_ladder_128 <- force.ultrametric(asymmetric_tree[[1]], method = "extend")
-for (i in 5:12) {tree_ladder_128[[5]] <- NULL}
-
-asymmetric_tree <- sim.age(age=511, numbsim=1, waitsp=function()2, waitext=function()2.5, symmetric=FALSE)
-length(asymmetric_tree[[1]]$tip.label)
-tree_ladder_256 <- force.ultrametric(asymmetric_tree[[1]], method = "extend")
-for (i in 5:12) {tree_ladder_256[[5]] <- NULL}
-
-asymmetric_tree <- sim.age(age=1023, numbsim=1, waitsp=function()2, waitext=function()2.5, symmetric=FALSE)
-length(asymmetric_tree[[1]]$tip.label)
-tree_ladder_512 <- force.ultrametric(asymmetric_tree[[1]], method = "extend")
-for (i in 5:12) {tree_ladder_512[[5]] <- NULL}
[…]

### 2020-05-01 Figures/Figures_Script.R

@@ -94,7 +94,7 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# Fig 3: same thing with regression beta = 0.5 ####
+# Fig 3: lambda_est~input with regression beta = 0.5 ####
 Data32 <- read.csv("Data_Analyses/Sim_Data/PB_reg-32.csv")
 Data64 <- read.csv("Data_Analyses/Sim_Data/PB_reg-64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/PB_reg-128.csv")
@@ -138,7 +138,7 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# Fig 4: Edited April 28 2020 ####
+# Fig 4: lambda_z~input and kappa_z~input at n = 32, coefficiences of variance ####
 
   Data32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_32.csv")
   Data64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_64.csv")
@@ -235,31 +235,26 @@ dev.off()
 
 # Panel A made in ppt, panels B and C made in Analysis_script
 
-# Fig S1: kappa~lambda_input 6 panel #####
-
+# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 512 OR 1024)####
 library(cowplot)
 library(ggplot2)
 
-Data32 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_32.csv")
-Data64 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_64.csv")
-Data128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_128.csv")
-Data256 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_256.csv")
-Data512 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_512.csv")
-Data1024_1 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_1024_1.csv")
[…]

### 2020-05-02 Manuscript/SupportingInfo-EKB.Rmd
Balanced analyses up through n = 512 finished.

@@
 # Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate
+
+*COPY FIGURES FROM FIGURES FOLDER INTO MANUSCRIPT FOLDER IN ORDER TO KNIT*
 
@@
 
+*still needs heavy edits*
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS5.png")
+
+```
+
+\textbf{Figure S5}. lambda_Z~lambda_input across tree sizes \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS6.png")
+
+```
+
+\textbf{Figure S6}. kappa_Z~lambda_input across tree sizes \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS7.png")
+
+```
+
+\textbf{Figure S7}. kappa~lambda_input across tree sizes \hfill\break
+
+\newpage
+
 # References

### 2020-05-02 Data_Analyses/Simulation_script-Pect-KappaComparison.R

@@ -62,11 +62,11 @@ DataTable_lambda <- data.frame(lambda.input = lambdas_expanded,
                                CI.lower = rep(NA, nsim*length(lambdas)),
                                CI.upper = rep(NA, nsim*length(lambdas))) 
 
-pb <- txtProgressBar(1, fullsimlength, style=1)
+pb <- txtProgressBar(153, fullsimlength, style=1)
 
 Start.time<-Sys.time()
 Start.time
-for (j in 1:fullsimlength) {
+for (j in 153:fullsimlength) {
   while(TRUE){
 
     if(lambdas_expanded[j] > 0) {
@@ -112,5 +112,5 @@ anyNA(DataTable_lambda) # want FALSE
 plot(DataTable_lambda$lambda.est~DataTable_lambda$lambda.input, pch = 19)
 
 # Writing Output Files ####
-file_name <- paste("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_", n, "-1to152.csv", sep = "")
+file_name <- paste("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_", n, "-153andUp.csv", sep = "")
 write.csv(DataTable_lambda, file_name, row.names = F)

### 2020-05-02 Figures/Figures_Script.R

@@ -235,7 +235,7 @@ dev.off()
 
 # Panel A made in ppt, panels B and C made in Analysis_script
 
-# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 512 OR 1024)####
+# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 1024)####
 library(cowplot)
 library(ggplot2)
 
@@ -243,7 +243,7 @@ Data32 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_32.csv")
 Data64 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") 
-Data512 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") # REPLACE WHEN SIM DONE
+Data512 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_512.csv") 
 Data1024 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") # REPLACE WHEN SIM DONE
 
 data_list <- list(Data32, Data64, Data128,
@@ -301,13 +301,15 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# Fig S3: lambda_est~lambda_input (same as fig 3 but with ANOVA; NOT YET FINISHED 512 AND 1024) #####
+# Fig S3: lambda_est~lambda_input (same as fig 3 but with ANOVA; NOT YET FINISHED 1024) #####
 
 Data32 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-32.csv")
 Data64 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-256.csv") 
[…]

### 2020-05-02 Data_Analyses/Simulation_script-Bal-KappaComparison.R
Balanced analyses up through n = 512 finished.

@@ -20,8 +20,8 @@ for (i in 1:length(treesizes)) {
   phy.list[[i]]$edge.length <- rep(1, (length(phy.list[[i]]$tip.label) - 1)*2)
 }
 
-tree<- phy.list[[4]] # BALANCED
-n <- treesizes[4]
+tree<- phy.list[[5]] # BALANCED
+n <- treesizes[5]
 
 # New Rep #####
 DataTable_lambda <- data.frame(lambda.input = lambdas_expanded, 

### 2020-05-02 Figures/Figures_Script.R

@@ -235,7 +235,7 @@ dev.off()
 
 # Panel A made in ppt, panels B and C made in Analysis_script
 
-# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 512 OR 1024)####
+# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 1024)####
 library(cowplot)
 library(ggplot2)
 
@@ -243,7 +243,7 @@ Data32 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_32.csv")
 Data64 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") 
-Data512 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") # REPLACE WHEN SIM DONE
+Data512 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_512.csv") 
 Data1024 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") # REPLACE WHEN SIM DONE
 
 data_list <- list(Data32, Data64, Data128,
@@ -301,13 +301,15 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# Fig S3: lambda_est~lambda_input (same as fig 3 but with ANOVA; NOT YET FINISHED 512 AND 1024) #####
+# Fig S3: lambda_est~lambda_input (same as fig 3 but with ANOVA; NOT YET FINISHED 1024) #####
 
 Data32 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-32.csv")
 Data64 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/PB_ANOVA-256.csv") 
[…]

### 2020-05-05 Manuscript/SupportingInfo-EKB.Rmd
Draft of Supporting Information ready for edits. All robustness analyses have been finished, and corresponding supporting figures have been generated. One has to copy said figures over to manuscript folder in order to knit SupportingInfo-EKB.Rmd.

@@
 
+\newpage
+
 ## Simulations of phylogenetic regression and ANOVA
-
-```{r setup, include=FALSE}
-  library(knitr)
-  knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$).
-For each $\lambda$ input value, independent variables were then generated with a known relationship ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.
+For each $\lambda$ input value, independent variables were then generated with a known relationship to the simulated dependent variable ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.
 
-*Results* As with the results from the regression analyses in the main text, we found limmited $\lambda$ estimation precision in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig.
+*Results* As with the results from the regression analyses in the main text, we found poor precision of $\lambda$ estimation in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig.
 S3).
@@
 
-*still needs heavy edits*
+## Effect Sizes of Pagel's Lambda and Kappa
+
+In this manuscript, we propose the use of effect size (Z-scores) over raw metrics of phylogenetic signal.
+We quantified Z-scores for estimated values of Pagel's $\lambda$ and Kappa of simulated datasets to demonstrate the variable behavior of these effect sizes across tree size and input values of phylogenetic signal.
+Simulation procedures can be found in the main text.
+
+*Results* Z scores of Pagel's $\lambda$ show a dramatically non-linear pattern across values of known $\lambda$ across all tree sizes (Figure S5).
+On the other hand, variation in the Kappa Z-scores is consistent across input $\lambda$s and shows a nearly linear relationship.
+Furthermore the variation of $Z_K$ is fairly consistent across values of input $\lambda$ indicating that $Z_K$ is an accurate and reliable statistic and can be appropriately used when comparing values of phylogenetic signal across clades or traits.
 
@@
 
-\textbf{Figure S5}. lambda_Z~lambda_input across tree sizes \hfill\break
+\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's $\lambda$ across input $\lambda$ values on various tree sizes.  \hfill\break
 
@@
 
-\textbf{Figure S6}. kappa_Z~lambda_input across tree sizes \hfill\break
+\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break
 
@@
 
-\textbf{Figure S7}. kappa~lambda_input across tree sizes \hfill\break
+\textbf{Figure S7}. Accuracy of Kappa estimations across known lambda inputs on various tree sizes.  \hfill\break
 

### 2020-05-05 Data_Analyses/Simulation_script-Pect-KappaComparison.R

@@ -62,11 +62,11 @@ DataTable_lambda <- data.frame(lambda.input = lambdas_expanded,
                                CI.lower = rep(NA, nsim*length(lambdas)),
                                CI.upper = rep(NA, nsim*length(lambdas))) 
 
-pb <- txtProgressBar(585, fullsimlength, style=1)
+pb <- txtProgressBar(1, fullsimlength, style=1)
 
 Start.time<-Sys.time()
 Start.time
-for (j in 585:fullsimlength) {
+for (j in 1:fullsimlength) {
   while(TRUE){
 
     if(lambdas_expanded[j] > 0) {
@@ -104,15 +104,10 @@ DataTable_lambda[which(is.na(DataTable_lambda$CI.upper)=="TRUE"),6] <- 1
 
 anyNA(DataTable_lambda) # want FALSE
 
-#DataTable_lambda[which(is.na(DataTable_lambda$kappa.z)=="TRUE"),]
-#length(which(is.na(DataTable_lambda$kappa.z)=="TRUE"))
-
 # Test Plot ####
 
 plot(DataTable_lambda$lambda.est~DataTable_lambda$lambda.input, pch = 19)
 
 # Writing Output Files ####
-file_name <- paste("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_", n, "585toEnd.csv", sep = "")
+file_name <- paste("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_", n, ".csv", sep = "")
 write.csv(DataTable_lambda, file_name, row.names = F)
-
-

### 2020-05-05 Figures/Figures_Script.R

@@ -235,7 +235,7 @@ dev.off()
 
 # Panel A made in ppt, panels B and C made in Analysis_script
 
-# Fig S1: lambda_est~input at each sample size, balanced (NOT YET FINISHED 1024)####
+# Fig S1: lambda_est~input at each sample size, balanced ####
 library(cowplot)
 library(ggplot2)
 
@@ -244,7 +244,9 @@ Data64 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") 
 Data512 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_512.csv") 
-Data1024 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_256.csv") # REPLACE WHEN SIM DONE
+Data1024_1 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_1024-1.csv")
+Data1024_2 <- read.csv("Data_Analyses/Sim_Data/Bal_lambda_kappacomparison_1024-2.csv")
+Data1024 <- rbind(Data1024_1, Data1024_2)
 
 data_list <- list(Data32, Data64, Data128,
                   Data256, Data512, Data1024)
@@ -268,7 +270,7 @@ plot_grid(PlotList[[1]], PlotList[[2]], PlotList[[3]],
           PlotList[[4]], PlotList[[5]], PlotList[[6]])
 dev.off()
 
-# Fig S2: lambda_est~input at each sample size, pectinate (NOT YET FINISHED 1024) ####
+# Fig S2: lambda_est~input at each sample size, pectinate ####
 library(cowplot)
 library(ggplot2)
 
@@ -277,7 +279,7 @@ Data64 <- read.csv("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_64.csv")
 Data128 <- read.csv("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_128.csv")
 Data256 <- read.csv("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_256.csv")
 Data512 <- read.csv("Data_Analyses/Sim_Data/Pect_lambda_kappacomparison_512.csv")
[…]

### 2020-05-06 Manuscript/Manuscript-5.Rmd
Supp INfo

--- 2020-04-30 Manuscript/Manuscript-4.Rmd
@@
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
 Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
@@
 For this to be the case, $\lambda$ must be a statistical effect size.
-Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
+Effect sizes are a measure of the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
@@
 These findings were particularly unsettling when considered in light of our literature survey.
-Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
+Over one quarter of the $\lambda$ estimates published in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
 This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
@@
 Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
-However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures.
+However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures.
 \hfill\break
@@
 where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation.
-Here the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018].
-Similarly, an effect size based on $\lambda$ could be envisioned as:
+Note that the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018].
+\hfill\break
+
+An effect size based on $\lambda$ could be envisioned, which is found as:
 
@@
 
-To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
-Results are found in Figure 4 (additional results are found in the Supporting Information).
+To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$, Figure 4: additional results in the Supporting Information).
 Here two things are evident.
-First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B).
+First, estimates of $Z_K$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A,B).
 Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$.
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C).
-This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
+Here coefficients of variation in the precision of $Z_K$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$.
 \hfill\break
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared to another.
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait than another.
 As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:
@@
 
-To demonstrate the utility of $\hat{Z}_{12}$ we performed an analysis of the strength of phylogenetic signal in two phenotypic datasets from species of plethodontid salamander.
+To demonstrate the utility of $\hat{Z}_{12}$ we quantified and compared the strength of phylogenetic signal of two phenotypic traits across species of plethodontid salamander.
 The data were part of a series of studies examining macroevolutionary trends in phenotypic diversification in this group [@Baken2019; @Baken2020].
@@
 Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020].
-Thus, because of this macroevolutionary association, strong phylogenetic signal in SA:V is to be expected.
+We hypothesize that because of this macroevolutionary association, strong phylogenetic signal in SA:V is observed.
 
@@
 First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomite changes in the values of $Z_K$.
+Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_K$.
 Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
@@
 By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
-Thus when viewed from this perspective, the approach developed here is statistically aligned with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons.
+Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons.
 As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
@@
 For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
+In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
 Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
-Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.
+Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
 \hfill\break
 
-** Finally... Need Closing paragraph. **
+**Finally... Need Closing paragraph.**
 

### 2020-05-06 Manuscript/SupportingInfo.Rmd
Supp INfo

@@
 ---
-# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate
-
-**FOR DCA**
-
-1: balanced/symmetric trees
-2: Kappa and K_Z
-3: Beta from PGLS
+# Supporting Information for: A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate
 
@@
 
-*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).
+*Results*.
+As found with pure-birth phylogenies, estimates of $\lambda$ varied most dramatically in simulations with fewer species and at intermediate values of lambda (Fig.
+S1, S2).
+Pectinate trees showed an interesting tendency to underestimate $\lambda$ across input values.
+Most dramatically, some simulations on the largest pectinate phylogeny ($n = 1024$) estimated $\lambda = 0$ for input values as high as 0.5 (Fig.
+S2).
 
 ```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+include_graphics("FigS1.png")
 
@@
 
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break
+\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees.
 
-## Literature Survey
+## Simulations of phylogenetic regression and ANOVA
 
-\newpage
+We analyzed the precision of parameter estimation when parameters were obtained in phylogenetic regression and ANOVA.
+This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$).
+For each $\lambda$ input value, independent variables were then generated with a known relationship to the simulated dependent variable ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.
 
-# References
+*Results* As with the results from the regression analyses in the main text, we found poor precision of $\lambda$ estimation in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig.
+S3).
+Slope was reliably estimated across tree sizes and known $\lambda$ values (Fig.
+S4).
 
- \setlength{\parindent}{-0.25in}
- \setlength{\leftskip}{0.25in}
- \setlength{\parskip}{8pt}
- \noindent
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS3.png")
 
-<div id="refs"></div>
+```
+
+\textbf{Figure S3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic ANOVA. Results are from input $\beta$ values of $0.5$. \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS4.png")
+
+```
+
+\textbf{Figure S4}. Precision of slope estimates across known input $\beta$ values of phylogenetic regression. Results shown include all input $\lambda$ values.
+
+## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_K$)
+
+We proposed the use of effect sizes (Z-scores) as a measure of the strength of phylogenetic signal.
+These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $Kappa$.
+The precision of effect sizes $Z_\lambda$ and $Z_K$ are summarized in the main text, primarily for simulations where $n=32$.
+Here we present results from additional simulations.
+
+*Results* Effect sizes for Pagel's $\lambda$ ($Z_\lambda$) scale nonlinearly with input levels of phylogenetic signal, and this pattern is consistent across all tree sizes examined (Figure S5).
+Further, the precision of $Z_\lambda$ varied widely across input levels of phylogenetic signal, with decidely greater variation (less precision) at stronger input levels of phylogenetic signal (Figure S5).
+By contrast, effect sizes for *Kappa* ($Z_K$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6).
+Additionally, the precision of $Z_K$ remained constant across input levels of phylogenetic signal (Figure S6).
+Together these results imply that $Z_K$ is a more reliable estimate of the strength of phylogenetic signal.
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS5.png")
+
+```
+
+\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's $\lambda$ across input $\lambda$ values on various tree sizes.  \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS6.png")
+
+```
+
+\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break
+
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("FigS7.png")
+
+```
+
+\textbf{Figure S7}.
+Estimates of phylogenetic signal using $Kappa$ across known lambda inputs on various tree sizes.
+Plot shows that while $Kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies.
+Thus, conversion to a standardized effect size ($Z_K$) is preferred.

### 2020-05-09 Manuscript/Manuscript-6-DCA.Rmd
vsn 6

--- 2020-05-06 Manuscript/Manuscript-5.Rmd
@@
 We find that the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised.
-We then propose a standardized effect size based on $Kappa$ ($Z_K$), which measures the strength of phylogenetic signal, and places it on a common scale for statistical comparison.
-Tests based on $Z_K$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
-Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal various phenotypic traits, even when those traits are found in different evolutionary lineages or of in different units or scale.
+We then propose a standardized effect size based on $\kappa$ ($Z_\kappa$), which measures the strength of phylogenetic signal, and places it on a common scale for statistical comparison.
+Tests based on $Z_\kappa$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
+Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scale.
 
@@
 # Introduction
-Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species generates statistical non-independence [@Felsenstein1985; @HarveyPagel1991].
-Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs); a suite of analytical tools that condition the data on the phylogeny through the course of statisical evaluations of phenotypic trends [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004].
+Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates an assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
+Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004].
 The past several decades have witnessed a rapid expansion in the development of PCMs to address an ever-growing set of macroevolutionary hypotheses [@MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @Adams2014a; @AdamsCollyer2018b].
@@
 
-Several analytical tools have been developed to quantify phylogenetic signal in phenotypic datasets, including measures of serial independence [$\mathbf{C}$: @Abouheif1999], autocorrelation estimates [$I$: @Gittleman1990], statistical ratios of trait variation relative to what is expected given the phylogeny [$Kappa$: @Blomberg_et_al2003; @Adams2014a], and scaling parameters used in maximum likelihood fitting of the data to the phylogeny [$\lambda$: @Pagel1999], among others [e.g., @Klingenberg2010].
+Several analytical tools have been developed to quantify phylogenetic signal in phenotypic datasets, including measures of serial independence [$\mathbf{C}$: @Abouheif1999], autocorrelation estimates [$I$: @Gittleman1990], statistical ratios of trait variation relative to what is expected given the phylogeny [$\kappa$: @Blomberg_et_al2003; @Adams2014a], and scaling parameters used in maximum likelihood fitting of the data to the phylogeny [$\lambda$: @Pagel1999], among others [e.g., @Klingenberg2010].
 The statistical properties of these methods -- namely type I error rates and power -- have also been investigated to determine when phylogenetic signal can be detected and under what conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @Adams2014a; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010].
 One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999].
-Here, maximum likelihood is used to fit the data to the phylogeny under a Brownian motion model of evolution.
-A parameter ($\lambda$) is included, which transforms the lengths of the internal branches of the phylogeny to improve the fit [@Pagel1999; @Freckleton_et_al2002].
+The ($\lambda$) parameter transforms the lengths of the internal branches of the phylogeny to improve a fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
 Pagel's $\lambda$ ranges from $0\to1$, with larger values signifying a greater dependence of observed trait variation on the phylogeny.
-Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic regression  (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
+Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic generalized least-squares regression (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
 \hfill\break
 
-Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
-Other approaches for interpreting $\lambda$ are more statistical.
-For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
-Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
+The appeal of $\lambda$ as a descriptive statistic for evolutionary biologists is a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\lambda$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
@@
 Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
-
 \hfill\break
@@
 [insert Figure 1 here] \hfill\break
+
+Various other approaches use $\lambda$ as a parameter that can be varied for inferences akin to sensitivity analysis.
+For instance, some have performed likelihood ratio tests that compare observed model fits to those obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or evaluated whether observed $\lambda$ differs from an expected $\lambda$, based on confidence intervals generated for the expected value [@Vandelook2019].
+Qualitative comparisons of $\lambda$ estimates have also been performed for multiple traits on the same phylogenetic tree to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+\hfill\break
 
@@
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
-Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
+Unfortunatley, not all model parameters and descriptive statistics are effect sizes, and thus many summary measures must first be converted to statistics with standardized units (i.e., conversion to an effect size) for meaningful comparison [see @Rosenthal1994].
 As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
@@
 Thus, $\lambda$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data.
-We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
-Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
+We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $\kappa$.
+Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $\kappa$ ($Z_\kappa$), implying that $Z_\kappa$ is a more robust effect size measure.
 We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
-By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
+By contrast, effect sizes obtained from $\kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
 
@@
 For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the variance of $\lambda$ ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness as an estimate of precision.
+We verified that the variance of traits simulated had no effect on phylogenetic signal estimation.
 \hfill\break
@@
 Here, an independent variable $X$ was simulated on each rescaled phylogeny under a Brownian motion model of evolution (for PGLS regression).
-For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny.
+For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary, 0 or 1) character on each phylogeny.
 Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal.
 This was accomplished as: $Y=\beta{X}+\epsilon$.
-Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution on each rescaled phylogeny: $\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$: [see @Revell2010 for a similar simulation design].
+The association between $Y$ and $X$ was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error ($\epsilon$) was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution on each rescaled phylogeny: $\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$: [see @Revell2010 for a similar simulation design].
 The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
 We then calculated precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness.
+We verified that the amount of residual variance simulated had no effect on $\sigma^2_\lambda$ but did influence the precision of coefficients estimated from the linear model (precision increased with smaller $\epsilon$, as expected).
 \hfill\break
@@
 Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression and ANOVA, the results of our simulations were quite similar.
-Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
+Regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
 However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; see also Supporting Information), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
@@
 
-Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
+Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size based on the $\kappa$ statistic and its empirical sampling distribution from permutation.
 Here we formalize that suggestion, resulting in an effect size of:
@@
 \begin{align}
-    Z_{K}=\frac{log(K_{obs})-\hat\mu_{log(K)}}{\hat\sigma_{log(K)}}
+    Z_\kappa=\frac{log(\kappa_{obs})-\hat\mu_{log(\kappa)}}{\hat\sigma_{log(\kappa)}}
 \end{align}
 
-where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation.
-Note that the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018].
+where $\kappa_{obs}$ is the observed phylogenetic signal, and $\hat\mu_\kappa$ and $\hat\sigma_\kappa$ are the mean and standard deviation of the empirical sampling distribution of $log(\kappa)$ obtained via permutation.
+Note that the logarithm was used because $\kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018].
 \hfill\break
@@
 
-To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$, Figure 4: additional results in the Supporting Information).
+To evaluate the utility of $Z_\kappa$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_\kappa}$ and $\sigma^2_{Z_\lambda}$, Figure 4: additional results in the Supporting Information).
 Here two things are evident.
-First, estimates of $Z_K$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A,B).
-Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$.
-Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$.
-This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of $Z_K$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$.
+First, estimates of $Z_\kappa$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A, B).
+Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_\kappa$.
+Second, the precision of $Z_\kappa$ is considerably more stable as compared with $Z_\lambda$.
+This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_\kappa}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
+Coefficients of variation in the precision of $Z_\kappa$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_\kappa$.
 \hfill\break
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait than another.
+Once the magnitude of phylogenetic signal is characterized using $Z_\kappa$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait than another.
 As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:
@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}}
+  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
 \end{align}
 
-where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2.
+where $\kappa_1$, $\kappa_2$, $\hat\mu_{\kappa_1}$, $\hat\mu_{\kappa_2}$, $\hat\sigma_{\kappa_1}$, and $\hat\sigma_{\kappa_2}$ are as defined above for equation 2.
+The right side of the equation illustrates that if $Z_\kappa$ has already been calculated for two sampling distributions as in equation 2, the sampling distributions have unit variance for each of the $Z_\kappa$ statistics.
 Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
@@
 A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the dataset, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species.
-The phylogenetic signal in each trait was then characterized using $Kappa$, which was converted to its effect size ($Z_K$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
+The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
 Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**).
@@
 
-*Results.* Both SA:V and relative body width displayed significant phylogenetic signal ($Kappa_{SA:V}=0.7608$; $P=0.001$; $Kappa_{BW/SVL}=0.2515$; $P=0.001$).
-For both phenotypic traits, $K_{obs}$ differed markedly from their corresponding permutation distributions, which were found to overlap almost perfectly (Figure 5B).
+*Results.* Both SA:V and relative body width displayed significant phylogenetic signal ($\kappa_{SA:V}=0.7608$; $P=0.001$; $\kappa_{BW/SVL}=0.2515$; $P=0.001$).
+For both phenotypic traits, $\kappa_{obs}$ differed markedly from their corresponding permutation distributions, which were found to overlap almost perfectly (Figure 5B).
 However, while both traits displayed significant phylogenetic signal, there was nearly a four-fold difference in the magnitude of their effect sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C).
@@
 $Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
-We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
-First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_K$.
-Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
-We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect.
+We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
+First, $Z_\kappa$ was more precise than $Z_{\lambda}$, and precision was more consistent across the range of input levels of phylogenetic signal.
+Additionally, values of $Z_\kappa$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_\kappa$.
+Thus, $Z_\kappa$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
+We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_\kappa$ as a statistical measure of this effect.
 \hfill\break
 
-Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another.
-Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003].
-However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
-Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale.
-By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
-Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons.
+Based on the effect size $Z_\kappa$, we then proposed a two-sample test, which provides means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another, via a hypothesis test.
+Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $\kappa$ [e.g.,  @Blomberg_et_al2003].
+However, $\kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
+Thus, $\kappa$ should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
+By contrast, standardizing $\kappa$ ($Z_\kappa$, via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
 As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
@@
 
-One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation.
-This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$.
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
+Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
 For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
 In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
-Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
 Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
@@
 
-**Finally... Need Closing paragraph.**
+Phylogenetic signal can be thought of as both an attribute to be measured in the data and a paramter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
+As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
+However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a useful  statistic to describe the amount of phylogenetic signal in the data.
+Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
+Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
+By contrast, $\kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_\kappa$).
+Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
+With low coefficients of variation for $Z_\kappa$ (Figure 4), it is difficult to imagine that a hypothesis test can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
 
@@
 Variation in effect size estimates of phylogenetic signal across input levels of phylogenetic signal.
-(A) Estimates $Z_\lambda$ for data simulated on phylogenies with 32 taxa ($n=32$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 32 taxa ($n=32$), (C) Coefficients of variation of precision estimates of $Z_\lambda$ and $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
+(A) Estimates $Z_\lambda$ for data simulated on phylogenies with 32 taxa ($n=32$), (B) Estimates of $Z_\kappa$ for data simulated on phylogenies with 32 taxa ($n=32$), (C) Coefficients of variation of precision estimates of $Z_\lambda$ and $Z_\kappa$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
 
@@
 (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
+(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
 
@@
 (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
+(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).

### 2020-05-09 Manuscript/Manuscript-6-DCA.tex
vsn 6

--- 2020-04-28 Manuscript/Manuscript-3.tex
@@
 
+Macroevolutionary studies frequently characterize the phylogenetic
+signal in phenotypes, and wish to compare the strength of that signal
+across traits. However, analytical tools for such comparisons have
+largely remained underdeveloped. In this study, we evaluated the
+efficacy of one commonly used parameter (Pagel's \(\lambda\)) to
+estimate the strength of phylogenetic signal in phenotypic traits, and
+evaluate the degree to which \(\lambda\) correctly identifies known
+levels of phylogenetic signal. We find that the precision of \(\lambda\)
+in estimating actual levels of phylogenetic signal is often inaccurate,
+and that biological interpretations of the strength of phylogenetic
+signal based on \(\lambda\) are therefore compromised. We then propose a
+standardized effect size based on \(\kappa\) (\(Z_\kappa\)), which
+measures the strength of phylogenetic signal, and places it on a common
+scale for statistical comparison. Tests based on \(Z_\kappa\) provide a
+mechanism for formally comparing the strength of phylogenetic signal
+across datasets, in much the same manner as effect sizes may be used to
+summarize patterns in quantitative meta-analysis. Our approach extends
+the phylogenetic comparative toolkit to address hypotheses that compare
+the strength of phylogenetic signal between various phenotypic traits,
+even when those traits are found in different evolutionary lineages or
+have different units or scale.
+
 \newpage
@@
 phylogenetic perspective, because the shared ancestry among species
-generates statistical non-independence (Felsenstein 1985; Harvey and
-Pagel 1991). Accounting for this evolutionary non-independence is the
-purview of \emph{phylogenetic comparative methods} (PCMs); a suite of
-analytical tools that condition the data on the phylogeny through the
-course of statisical evaluations of phenotypic trends (e.g., Grafen
-1989; Garland and Ives 2000; Rohlf 2001; Butler and King 2004). The past
-several decades have witnessed a rapid expansion in the development of
-PCMs to address an ever-growing set of macroevolutionary hypotheses
-(Martins and Hansen 1997; O'Meara et al. 2006; Revell and Harmon 2008;
-Beaulieu et al. 2012; Adams 2014b,a; Adams and Collyer 2018). These
-methods are predicated on the notion that phylogenetic signal -- the
-tendancy for closely related species to display similar trait values --
-is present in cross-species datasets (Felsenstein 1985; Pagel 1999;
-Blomberg et al. 2003). Indeed, under numerous evolutionary models,
-phylogenetic signal is to be expected, as stochastic character change
-along the hierarchical structure of the tree of life generates trait
-covaration among related taxa (see Felsenstein 1985; Blomberg et al.
-2003; Revell et al. 2008). \hfill\break
+violates an assumption of independence among trait values that is common
+for statistical tests (Felsenstein 1985; Harvey and Pagel 1991).
+Accounting for this evolutionary non-independence is the purview of
+\emph{phylogenetic comparative methods} (PCMs): a suite of analytical
+tools that condition trends in the data on the phylogenetic relatedness
+of observations (e.g., Grafen 1989; Garland and Ives 2000; Rohlf 2001;
+Butler and King 2004). The past several decades have witnessed a rapid
+expansion in the development of PCMs to address an ever-growing set of
+macroevolutionary hypotheses (Martins and Hansen 1997; O'Meara et al.
+2006; Revell and Harmon 2008; Beaulieu et al. 2012; Adams 2014b,a; Adams
+and Collyer 2018). These methods are predicated on the notion that
+phylogenetic signal -- the tendancy for closely related species to
+display similar trait values -- is present in cross-species datasets
+(Felsenstein 1985; Pagel 1999; Blomberg et al. 2003). Indeed, under
+numerous evolutionary models, phylogenetic signal is to be expected, as
+stochastic character change along the hierarchical structure of the tree
+of life generates trait covaration among related taxa (see Felsenstein
+1985; Blomberg et al. 2003; Revell et al. 2008). \hfill\break
 
@@
 Gittleman and Kot 1990), statistical ratios of trait variation relative
-to what is expected given the phylogeny (\(Kappa\): Blomberg et al.
+to what is expected given the phylogeny (\(\kappa\): Blomberg et al.
 2003; Adams 2014a), and scaling parameters used in maximum likelihood
@@
 phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
-(Pagel 1999). Here, maximum likelihood is used to fit the data to the
-phylogeny under a Brownian motion model of evolution. A parameter
-(\(\lambda\)) is included, which transforms the lengths of the internal
-branches of the phylogeny to improve the fit (Pagel 1999; Freckleton et
-al. 2002). Pagel's \(\lambda\) ranges from \(0\to1\), with larger values
-signifying a greater dependence of observed trait variation on the
-phylogeny. Pagel's \(\lambda\) also has the appeal that it may be
-included in phylogenetic regression (PGLS) to account for the degree of
-phylogenetic signal in comparative analyses (see Freckleton et al.
-2002). \hfill\break
+(Pagel 1999). The (\(\lambda\)) parameter transforms the lengths of the
+internal branches of the phylogeny to improve a fit of data to the
+phylogeny via maximum likelihood (Pagel 1999; Freckleton et al. 2002).
+Pagel's \(\lambda\) ranges from \(0\to1\), with larger values signifying
+a greater dependence of observed trait variation on the phylogeny.
+Pagel's \(\lambda\) also has the appeal that it may be included in
+phylogenetic generalized least-squares regression (PGLS) to account for
+the degree of phylogenetic signal in comparative analyses (see
+Freckleton et al. 2002). \hfill\break
 
-Evolutionary biologists commonly seek to describe the relative strength
-of phylogenetic signal in phenotypic traits, to determine the extent to
-which shared evolutionary history has influenced trait covariation among
-taxa. This is often accomplished by interpreting empirical estimates of
-\(\lambda\); with smaller values signifying `weak' phylogenetic signal,
-while larger values are interpreted as `strong' phylogenetic signal
+In addition to functioning as a parameter that is tuned for appropriate
+analysis, \(\lambda\) can function as a descriptive statistic to
+describe the relative strength of phylogenetic signal in phenotypic
+traits, to determine the extent to which shared evolutionary history has
+influenced trait covariation among taxa. The appeal of \(\lambda\) as a
+descriptive statistic for evolutionary biologists is a basis for
+interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e., small
+versus large values of \(\lambda\), respectively, in a comparative sense
 (e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
-Other approaches for interpreting \(\lambda\) are more statistical. For
-instance, some have evaluated whether the observed \(\lambda\) differs
-from some expected value through the use of confidence intervals
-(Vandelook et al. 2019) or by performing likelihood ratio tests that
-compare the observed model fit to that obtained when \(\lambda=0\) or
-\(\lambda=1\) (Freckleton et al. 2002; Cooper et al. 2010; Bose et al.
-2019). Additionally, qualitative comparisons of \(\lambda\) estimates
-obtained from multiple phenotypic traits have been used to infer whether
-the strength of phylogenetic signal is greater in one trait as compared
-to another (e.g., Liu et al. 2019; Bai et al. 2019). Indeed, statements
-regarding the strength of phylogenetic signal based on \(\lambda\) are
-rather common in the evolutionary literature. For instance, of the 204
-papers published in 2019 that estimated and reported Pagel's \(\lambda\)
-(found in Google.scholar), 40\% interpreted the strength of phylogenetic
-signal for at least one phenotypic trait. Additionally, because nearly
-half of the 1,572 \(\lambda\) values reported were near the limits of
-the parameter (Figure 1), this percentage is even higher, as biological
-interpretation of phylogenetic signal at the limits of \(\lambda\) are
-known. \hfill\break
+Indeed, statements regarding the strength of phylogenetic signal based
+on \(\lambda\) are rather common in the evolutionary literature. For
+instance, of the 204 papers published in 2019 that estimated and
+reported Pagel's \(\lambda\) (found from a literature survey we
+conducted in Google.scholar), 40\% interpreted the strength of
+phylogenetic signal for at least one phenotypic trait. Further, because
+nearly half of the 1,572 \(\lambda\) values reported were near 0 or 1
+(Figure 1) where the biological interpretation of \(\lambda\) is known,
+this percentage is even higher. \hfill\break
 
 {[}insert Figure 1 here{]} \hfill\break
+
+Various other approaches use \(\lambda\) as a parameter that can be
+varied for inferences akin to sensitivity analysis. For instance, some
+have performed likelihood ratio tests that compare observed model fits
+to those obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
+2002; Cooper et al. 2010; Bose et al. 2019) or evaluated whether
+observed \(\lambda\) differs from an expected \(\lambda\), based on
+confidence intervals generated for the expected value (Vandelook et al.
+2019). Qualitative comparisons of \(\lambda\) estimates have also been
+performed for multiple traits on the same phylogenetic tree to infer
+whether the strength of phylogenetic signal is greater in one trait as
+compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
+\hfill\break
 
@@
 are often more precise at greater sample sizes (Cohen 1988). However,
-this conclusion also assumes that the precision of \(\lambda\) remains
+this conclusion also implies that the precision of \(\lambda\) remains
 constant across its range (\(\lambda = 0 \to 1\)); an assumption that to
 date, has not been verified. Thus, despite widespread use of Pagel's
-(1999) \(\lambda\) in macroevolutionary studies, at present, we still
-lack a general understanding of the precision with which \(\lambda\) can
+(1999) \(\lambda\) in macroevolutionary studies, at present, we lack a
+general understanding of the precision with which \(\lambda\) can
 estimate levels of phylogenetic signal in phenotypic datasets.
@@
 \(\lambda\) must be a statistical effect size. Effect sizes are a
-measure the magnitude of a statistical effect in data, represented on a
-common scale (Glass 1976; Cohen 1988). Effect sizes have widespread use
-in many areas of the quantiative sciences, as they represent measures
-that may be readily summarized across datasets as in meta-analysis
-(Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster 1995), or
-compared among datasets (e.g., Adams and Collyer 2016, 2019a).
-Unfortunatley, not all model parameters and test statistics are effect
-sizes, and thus many summary measures must first be converted to
-standardized units (i.e., an effect size) for meaningful comparison (see
-Rosenthal 1994). As a consequence, it follows that only if \(\lambda\)
-is a statistical effect size can comparisons of estimates across
-datasets be interpretable. For the case of \(\lambda\), this has not yet
-been explored. \hfill\break
+measure of the magnitude of a statistical effect in data, represented on
+a common scale (Glass 1976; Cohen 1988). Effect sizes have widespread
+use in many areas of the quantiative sciences, as they represent
+measures that may be readily summarized across datasets as in
+meta-analyses (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster
+1995), or compared among datasets (e.g., Adams and Collyer 2016, 2019a).
+Unfortunatley, not all model parameters and descriptive statistics are
+effect sizes, and thus many summary measures must first be converted to
+statistics with standardized units (i.e., conversion to an effect size)
+for meaningful comparison (see Rosenthal 1994). As a consequence, it
+follows that only if \(\lambda\) is a statistical effect size can
+comparisons of estimates across datasets be interpretable. For the case
+of \(\lambda\), this has not yet been explored. \hfill\break
 
@@
 from datasets containing vastly different input levels of phylogenetic
-signal. Thus, \(\lambda\) is not a reliable estimate of the strength of
+signal. Thus, \(\lambda\) is not a reliable indicator of the strength of
 phylogenetic signal in phenotypic data. We then describe a standardized
@@
 phenotypic datasets, and apply the concept to two common measures of
-phylogenetic signal: \(\lambda\) and \(Kappa\). Through simulations
-across a wide range of conditions, we find that the precision of effect
-sizes based on \(\lambda\) (\(Z_{\lambda}\)) are less reliable than that
-those based on \(Kappa\) (\(Z_K\)), implying that \(Z_K\) is a more
-robust effect size measure. We also propose a two-sample test statistic
-that may be used to compare the strength of phylogenetic signal among
-datasets, and provide an empirical example to demonstrate its use. We
-conclude that estimates of phylogenetic signal using Pagel's \(\lambda\)
-are often inaccurate, and thus interpreting strength of phylogenetic
-signal in phenotypic datasets based on this measure is compromised. By
-contrast, effect sizes obtained from \(Kappa\) hold promise for
-characterizing phylogenetic signal, and for comparing the strength of
-phylogenetic signal across datasets.
+phylogenetic signal: \(\lambda\) and \(\kappa\). Through simulations we
+find that the precision of effect sizes based on \(\lambda\)
+(\(Z_{\lambda}\)) are less reliable than that those based on \(\kappa\)
+(\(Z_\kappa\)), implying that \(Z_\kappa\) is a more robust effect size
+measure. We also propose a two-sample test statistic that may be used to
+compare the strength of phylogenetic signal among datasets, and provide
+an empirical example to demonstrate its use. We conclude that estimates
+of phylogenetic signal using Pagel's \(\lambda\) are often inaccurate,
+and thus interpreting strength of phylogenetic signal in phenotypic
+datasets based on this measure is compromised. By contrast, effect sizes
+obtained from \(\kappa\) hold promise for characterizing phylogenetic
+signal, and for comparing the strength of phylogenetic signal across
+datasets.
 
@@
 ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
-phylogenetic signal corresponding reflecting Brownian motion (when
+phylogenetic signal reflecting Brownian motion (when
 \(\lambda_{in} =1\)). For each dataset we then estimated phylogenetic
-signal (\(\lambda_{est}\)), and calculated the precision of \(\lambda\)
-using the variance (\(\sigma^2_\lambda\)) across datasets at each input
-level of phylogenetic signal and level of species richness. \hfill\break
+signal (\(\lambda_{est}\)), and calculated the variance of \(\lambda\)
+(\(\sigma^2_\lambda\)) across datasets at each input level of
+phylogenetic signal and level of species richness as an estimate of
+precision. We verified that the variance of traits simulated had no
+effect on phylogenetic signal estimation. \hfill\break
 
@@
 regression and ANOVA (i.e., \(Y\sim{X}\)). Here, an independent variable
-\(X\) was simulated on each phylogeny under a Brownian motion model of
-evolution (for PGLS regression). For phylogenetic ANOVA, random groups
-(\(X\)) were obtained by simulating a discrete (binary) character on
-each phylogeny. Next, the dependent variable was simulated in such a
-manner as to contain a known relationship with \(X\) plus random error
-containing phylogenetic signal. This was accomplished as:
-\(Y=\beta{X}+\epsilon\). Here, the association between \(Y\) and \(X\)
-was modeled using a range of values:
-\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
-to contain phylogenetic signal simulated under a Brownian motion model
-of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
-Revell 2010 for a similar simulation design). The fit of the
-phylogenetic regression was estimated using maximum likelihood, and
-parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
-obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
-of phylogenetic signal and level of species richness were then observed.
-\hfill\break
+\(X\) was simulated on each rescaled phylogeny under a Brownian motion
+model of evolution (for PGLS regression). For phylogenetic ANOVA, random
+groups (\(X\)) were obtained by simulating a discrete (binary, 0 or 1)
+character on each phylogeny. Next, the dependent variable was simulated
+in such a manner as to contain a known relationship with \(X\) plus
+random error containing phylogenetic signal. This was accomplished as:
+\(Y=\beta{X}+\epsilon\). The association between \(Y\) and \(X\) was
+modeled using a range of values: \(\beta=(0.0,0.25, 0.5, 0.75,1.0)\),
+and the residual error (\(\epsilon\)) was modeled to contain
+phylogenetic signal simulated under a Brownian motion model of evolution
+on each rescaled phylogeny:
+\(\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\): (see Revell
+2010 for a similar simulation design). The fit of the phylogenetic
+regression was estimated using maximum likelihood, and parameter
+estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were obtained. We then
+calculated precision estimates (\(\sigma^2_\lambda\)) at each input
+level of phylogenetic signal and level of species richness. We verified
+that the amount of residual variance simulated had no effect on
+\(\sigma^2_\lambda\) but did influence the precision of coefficients
+estimated from the linear model (precision increased with smaller
+\(\epsilon\), as expected). \hfill\break
 
@@
 literature survey. Over one quarter of the \(\lambda\) estimates
-obtained in empirical studies (421 of 1,572) were between
+published in empirical studies (421 of 1,572) were between
 \(\lambda=0.25\) and \(\lambda=0.75\) (Figure 1). This range reflected
@@
 in terms of accurately characterizing levels of phylogenetic signal, yet
-30\% of these empirical estimates were explicitly interpreted in terms
-of the strength of phylogenetic signal that they represented (i.e.,
-weak, intermediate, strong phylogenetic signal). \hfill\break
+30\% of these mid-range empirical estimates were explicitly interpreted
+in terms of the strength of phylogenetic signal that they represented
+(i.e., weak, intermediate, strong phylogenetic signal). \hfill\break
 
 Finally, when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression, the results of our simulations were quite similar.
-Here, regression parameters (\(\beta\)) were accurately estimated,
+PGLS regression and ANOVA, the results of our simulations were quite
+similar. Regression parameters (\(\beta\)) were accurately estimated,
 confirming earlier findings of Revell (2010) (see Supporting
 Information). However, estimates of phylogenetic signal (\(\lambda\))
-were less precise (Figure 3), and the spread of \(\lambda_{est}\) was
-similar to that observed when \(\lambda\) was estimated for only the
-dependent variable, as in Figure 2. Taken together, these findings
-reveal that \(\lambda_{est}\) does not precisely characterize observed
-levels of phylogenetic signal in phenotypic datasets, and that
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) may be highly inaccurate. \hfill\break
+were less precise (Figure 3; see also Supporting Information), and the
+spread of \(\lambda_{est}\) was similar to that observed when
+\(\lambda\) was estimated for only the dependent variable, as in Figure
+2. Taken together, these findings reveal that \(\lambda_{est}\) does not
+precisely characterize observed levels of phylogenetic signal in
+phenotypic datasets, and that biological interpretations of the strength
+of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
+\hfill\break
 
@@
 derived from parametric theory. However, recent advances in resampling
-theory (Collyer et al. 2015; Adams and Collyer 2016, 2019a) have shown
-that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from an
-empirical sampling distribution of \(\theta\) obtained from permutation
-procedures. \hfill\break
+theory (Collyer et al. 2015; Adams and Collyer 2016, 2019a; Collyer and
+Adams 2018) have shown that \(E(\theta)\) and \(\sigma_\theta\) may also
+be obtained from an empirical sampling distribution of \(\theta\)
+obtained from permutation procedures. \hfill\break
 
 Adams and Collyer (2019b) suggested that the strength of phylogenetic
-signal could be represented as an effect size, based on the \(Kappa\)
+signal could be represented as an effect size based on the \(\kappa\)
 statistic and its empirical sampling distribution from permutation. Here
@@
 \begin{align}
-    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
+    Z_\kappa=\frac{log(\kappa_{obs})-\hat\mu_{log(\kappa)}}{\hat\sigma_{log(\kappa)}}
 \end{align}
 
-where \(K_{obs}\) is the observed phylogenetic signal, and \(\hat\mu_K\)
-and \(\hat\sigma_K\) are the mean and standard deviation of the
-empirical sampling distribution of \(Kappa\) obtained via permutation.
-Similarly, an effect size based on \(\lambda\) could be envisioned as:
+where \(\kappa_{obs}\) is the observed phylogenetic signal, and
+\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
+deviation of the empirical sampling distribution of \(log(\kappa)\)
+obtained via permutation. Note that the logarithm was used because
+\(\kappa\) takes only positive values (\(0\to\infty\)) and its sampling
+distribution is log-normally distributed (for a similar transformation
+when calculating multivariate effect sizes see: Appendix 1 of Collyer
+and Adams 2018). \hfill\break
+
+An effect size based on \(\lambda\) could be envisioned, which is found
+as:
 
@@
 In this case, \(\lambda_{obs}\) and \(\hat\sigma_\lambda\) are
-empirically derived using maximum likelihood. Note that under the null
-hypothesis, no phylogenetic signal is expected (Freckleton et al. 2002),
-and thus \(E(\lambda)=0\) under this condition. \hfill\break
+empirically derived using maximum likelihood, as permutation approaches
+have not been developed for evaluating \(\lambda\). Note also that under
+the null hypothesis, no phylogenetic signal is expected (Freckleton et
+al. 2002), and thus \(E(\lambda)=0\) under this condition. \hfill\break
 
-To evaluate the utility of \(Z_K\) and \(Z_\lambda\) we calculated both
-effect sizes for the simulated datasets generated above, and summarized
-the precision of each using its variance (\(\sigma^2_{Z_K}\) and
-\(\sigma^2_{Z_\lambda}\)). Results are found in Figure 4 (additional
-results are found in the Supporting Information). Here two things are
-evident. First, estimates of \(Z_K\) track the input phylogenetic signal
-in a more linear fashion than do estimates of \(Z_\lambda\) (Figure
-4A,B). Thus, actual changes in the strength of phylogenetic signal are
+To evaluate the utility of \(Z_\kappa\) and \(Z_\lambda\) we calculated
+both effect sizes for the simulated datasets generated above, and
+summarized the precision of each using its variance
+(\(\sigma^2_{Z_\kappa}\) and \(\sigma^2_{Z_\lambda}\), Figure 4:
+additional results in the Supporting Information). Here two things are
+evident. First, estimates of \(Z_\kappa\) linearly track the input
+phylogenetic signal whereas estimates of \(Z_\lambda\) do not (Figure
+4A, B). Thus, actual changes in the strength of phylogenetic signal are
 reflected more evenly in the corresponding values of the effect size
-\(Z_K\). Second, the precision of \(Z_K\) is considerably more stable as
-compared with \(Z_\lambda\). This may be seen by calculating the
-coefficients of variation for the set of precision estimates (i.e.,
-\(\sigma^2_{Z_K}\) and \(\sigma^2_{Z_\lambda}\)) across input levels of
-phylogenetic signal. Here coefficients of variation in the precision of
-\(Z_K\) were an order of magnitude smaller for than for \(Z_\lambda\)
-(Figure 4C,D). This implied that estimates of the strength of
-phylogenetic signal were more reliable and robust when using \(Z_K\) as
-compared with \(Z_\lambda\). \hfill\break
+\(Z_\kappa\). Second, the precision of \(Z_\kappa\) is considerably more
+stable as compared with \(Z_\lambda\). This may be seen by calculating
+the coefficients of variation for the set of precision estimates (i.e.,
+\(\sigma^2_{Z_\kappa}\) and \(\sigma^2_{Z_\lambda}\)) across input
+levels of phylogenetic signal. Coefficients of variation in the
+precision of \(Z_\kappa\) were up to an order of magnitude smaller for
+than for \(Z_\lambda\) (Figure 4C), implying that estimates of the
+strength of phylogenetic signal were more reliable and robust when using
+\(Z_\kappa\). \hfill\break
 
@@
 Once the magnitude of phylogenetic signal is characterized using
-\(Z_K\), one may wish to compare such measures across datasets, to
+\(Z_\kappa\), one may wish to compare such measures across datasets, to
 determine whether the strength of phylogenetic signal is significantly
-greater in one phenotypic trait as compared with another. As with other
-effect sizes derived from permutation distributions (e.g., Adams and
-Collyer 2016, 2019a), a two-sample test statistic may be found as:
+greater in one phenotypic trait than another. As with other effect sizes
+derived from permutation distributions (e.g., Adams and Collyer 2016,
+2019a), a two-sample test statistic may be calculated as:
 
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}}
+  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
 \end{align}
 
-where \(K_1\), \(K_2\), \(\hat\mu_{K_1}\), \(\hat\mu_{K_2}\),
-\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above for
-equation 2. Estimates of significance of \(\hat{Z}_{12}\) may be
+where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
+\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
+\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
+side of the equation illustrates that if \(Z_\kappa\) has already been
+calculated for two sampling distributions as in equation 2, the sampling
+distributions have unit variance for each of the \(Z_\kappa\)
+statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
 obtained from a standard normal distribution. Typically,
@@
 
-One important advantage of the approach advocated here is that the
-resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
-measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
-2012). As a consequence, the effect sizes represent the strength of
-phylogenetic signal on a common and comparable scale, measured in
-standard deviation units, regardless of the initial units and scale of
-the original phenotypic variables under investigation. This means that
-conceivably, the strength of phylogenetic signal may be compared across
-datsets for phenotypic traits in different units and scale (e.g.,
-behavioral and morphological traits), so long as they are continuous
-variables. Note that this is directly related to the common practice of
-converting data to their standard normal deviates prior to their use in
-multivariate analysis, if the original variables are expressed in
-different units or scales (see Everitt and Hothorn 2011; Legendre and
-Legendre 2012; Adams and Collyer 2019b).
-
 \hypertarget{empirical-example}{%
@@
 
-To demonstrate the utility of \(\hat{Z}_{12}\) we performed an analysis
-of the strength of phylogenetic signal in two phenotypic datasets from
-species of plethodontid salamander. The data were part of a series of
-studies examining macroevolutionary trends in phenotypic diversification
-in this group (Baken and Adams 2019; Baken et al. 2020). Our dataset
-contained surface area to volume ratios (SA:V) and relative body width
-(\(\frac{BW}{SVL}\)) for 305 species (Figure 5A). For SA:V, 11 linear
-body measurements from 2,781 individuals were taken, from which
-estimates of the surface area and volume of the head, body, and tail
-were calculated and subsequently combined to arrive at the SA:V for each
-indiviual (for mathematical details see Baken et al. 2020). Species
+To demonstrate the utility of \(\hat{Z}_{12}\) we quantified and
+compared the strength of phylogenetic signal of two phenotypic traits
+across species of plethodontid salamander. The data were part of a
+series of studies examining macroevolutionary trends in phenotypic
+diversification in this group (Baken and Adams 2019; Baken et al. 2020).
+Our dataset contained surface area to volume ratios (SA:V) and relative
+body width (\(\frac{BW}{SVL}\)) for 305 species (Figure 5A). For SA:V,
+11 linear body measurements from 2,781 individuals were taken, from
+which estimates of the surface area and volume of the head, body, and
+tail were calculated and subsequently combined to arrive at the SA:V for
+each indiviual (for mathematical details see Baken et al. 2020). Species
 means were then obtained. Likewise, body size (SVL) and body width (BW)
@@
 containing 305 species. The phylogenetic signal in each trait was then
-characterized using \(Kappa\), which was converted to its effect size
-(\(Z_K\)) using \texttt{geomorph} 3.2.1 (Adams and Otárola-Castillo
+characterized using \(\kappa\), which was converted to its effect size
+(\(Z_\kappa\)) using \texttt{geomorph} 3.2.1 (Adams and Otárola-Castillo
 2013; Adams et al. 2020). Finally, the strength of phylogenetic signal
@@
 \emph{Results.} Both SA:V and relative body width displayed significant
-phylogenetic signal (\(Kappa_{SA:V}=0.7608\); \(P=0.001\);
-\(Kappa_{BW/SVL}=0.2515\); \(P=0.001\)). For both phenotypic traits,
-\(K_{obs}\) differed markedly from their corresponding permutation
+phylogenetic signal (\(\kappa_{SA:V}=0.7608\); \(P=0.001\);
+\(\kappa_{BW/SVL}=0.2515\); \(P=0.001\)). For both phenotypic traits,
+\(\kappa_{obs}\) differed markedly from their corresponding permutation
 distributions, which were found to overlap almost perfectly (Figure 5B).
@@
 display greater variation in SA:V which covaries with disparity in their
-climatic niches (Baken et al. 2020). Thus, because of this
-macroevolutionary association, strong phylogenetic signal in SA:V is to
-be expected.
+climatic niches (Baken et al. 2020). We hypothesize that because of this
+macroevolutionary association, strong phylogenetic signal in SA:V is
+observed.
 
-\hypertarget{conclusions-and-implications}{%
-\section{Conclusions and
-Implications}\label{conclusions-and-implications}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-1: summary paragraph
+It is common in comparative evolutionary studies to characterize the
+phylogenetic signal in phenotypic traits to determine the extent to
+which shared evolutionary history has generated trait covariation among
+taxa. However, while numerous analytical approaches may be used to
+quantify phylogenetic signal (e.g., Abouheif 1999; Gittleman and Kot
+1990; Pagel 1999; Blomberg et al. 2003; Adams 2014a), methods that
+explicitly measure the strength of phylogenetic signal, or facilitate
+comparisons among datasets, have remained underdeveloped. In this study,
+we evaluated the precision of one common measure, Pagel's \(\lambda\),
+and explored its efficacy for characterizing the strength of
+phylogenetic signal in phenotypic data. Using computer simulations, we
+found that the precision of \(\lambda\) increased with increasing sample
+sizes; a pattern noted previously (Boettiger et al. 2012), and one that
+conformed with parametric statistical theory (Cohen 1988). However, we
+also found that vastly different \(\lambda\) estimates could be obtained
+from data containing the same level of phylogenetic signal, and that
+similar \(\lambda\) estimates may be obtained from data containing
+differing levels of phylogenetic signal. Further, the precision of
+\(\lambda\) varied with the strength of phylogenetic signal, where lower
+precision was observed when in data whose phylogenetic signal was of
+intemediate strength. From these findings we conclude that \(\lambda\)
+is not a reliable indicator of the observed strength of phylogenetic
+signal in phenoytpic datasets, and that biological interpretations of
+the strength of signal based on this parameter may innacurately
+characterize such effects. \hfill\break
 
-2: expand on Lambda.. lambda innacurate, not precise, level of precision
-varies with input physig (worse in mid-range). NEW RESULT. We are first
-to show this. NOTE: pattern is obvious with reflection. Since it is a
-`bounded' parameter estimation should be best at the extremes\ldots{}
-(state this?).. hmm.
+As an alternative, we described a standardized effect size (\(Z\)) for
+assessing the strength of phylogenetic signal. \(Z\) expresses the
+magnitude of phylogenetic signal as a standard normal deviate, which is
+easily interpretable as the strength of phylogenetic signal relative to
+the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
+and found that \(Z_\kappa\) was a better estimate of the strength of
+phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
+precise than \(Z_{\lambda}\), and precision was more consistent across
+the range of input levels of phylogenetic signal. Additionally, values
+of \(Z_\kappa\) more accurately tracked known levels of phylogenetic
+signal, with changes in the actual strength of phylogenetic signal
+reflected in a more linear fashion by concomitant changes in the values
+of \(Z_\kappa\). Thus, \(Z_\kappa\) holds promise as a measure of the
+relative strength of phylogenetic signal that reflects the magnitude of
+this effect in phenotypic data. We therefore recommend that future
+studies interested in the strength of phylogenetic signal incorporate
+\(Z_\kappa\) as a statistical measure of this effect. \hfill\break
 
-Patterns worse with PGLS, though beta still estimated properly.
-Conclusion, lambda not overly useful.
-
-3: By contrast, effect size Z-K useful, equally precise across range of
-values. Can be used to characterize the strength of physignal, and
-because robust to input levels, etc. may be used to compare across
-datasets.
-
-Somewhere, recognize that this is somewhat `backwards' from prior
-recommendations where Kappa had somewhat lower performance in terms of
-type I and type II error (which?? I forget). However, recall that those
-studies did not examine the precision of the estimates. Nor was Z-k
-included, because it was not yet invented. So Use of Z-k should make
-good sense here.
-
-Closing paragraph.
-
+Based on the effect size \(Z_\kappa\), we then proposed a two-sample
+test, which provides means of determining whether the strength of
+phylogenetic signal is greater in one phenotypic trait as compared to
+another, via a hypothesis test. Prior studies have summarized patterns
+of variation in phylogenetic signal across datasets using summary test
+values, such as \(\kappa\) (e.g., Blomberg et al. 2003). However,
+\(\kappa\) does not scale linearly with input levels of phylogenetic
+signal, and its variance increases (i.e., precision decreases) with
+increasing strength of phylogenetic signal (Munkemuller et al. 2012;
+Diniz-Filho et al. 2012: see also Supporting Information). Thus,
+\(\kappa\) should not be considered an effect size that measures the
+strength of phylogenetic signal on a common scale. By contrast,
+standardizing \(\kappa\) (\(Z_\kappa\), via equation 2) alleviates these
+concerns, and facilitates formal statistical comparisons of the strength
+of signal across datasets. Thus when viewed from this perspective, the
+approach developed here aligns well with other statistical approaches
+such as meta-analysis (sensu Hedges and Olkin 1985; Glass 1976; Arnqvist
+and Wooster 1995), where summary statistics across datasets are
+converted to standardized effect sizes for subsequent ``higher order''
+statistical summaries or comparisons. As such our approach enables
+evolutionary biologists to quantitatively examine the relative strength
+of phylogenetic signal across a wide range of phenotypic traits, and
+thus opens the door for future discoveries that inform on how phenotypic
+diversity accumulates in macroevolutionary time across the tree of life.
 \hfill\break
 
-More discussion paragraphs
+One important advantage of the approach advocated here is that the
+resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
+measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
+2012). Thus, \(Z_\kappa\) represents the strength of phylogenetic signal
+on a common and comparable scale -- measured in standard deviation --
+regardless of the initial units and original scale of the phenotypic
+variables under investigation. This means that the strength of
+phylogenetic signal may be compared across datsets for continuous
+phenotypic traits measured in different units and scale, because those
+units have been standardized through their conversion to \(Z_\kappa\).
+For example, our approach could be utilized to determine whether the
+strength of phylogenetic signal (say, in response to ecological
+differentiation) is stronger in morphological traits (linear traits:
+\(mm\)), physiological traits (metabolic rate: \(\frac{O^2}{min}\)), or
+behavioral traits (aggression: \(\frac{\#{displays}}{second}\)). In
+fact, our empirical example provided such a comparison, as SA:V is
+represented in \(mm^{-1}\) while relative body size is a unitless ratio
+(\(\frac{BW}{SVL}\)). Additionally, our method is capable of comparing
+the strength of phylogenetic signal in traits of different
+dimensionality, as estimates of phylogenetic signal using \(\kappa\)
+have been generalized for multivariate data (\(K_{mult}\): see Adams
+2014a). Furthermore, tests based on \(\hat{Z}_{12}\) may be utilized for
+comparing the strength of phylogenetic signal among datasets containing
+a different number of species, and even for phenotypes obtained from
+species in different lineages, because their phylogenetic
+non-independence and observed variation are taken into account in the
+generation of the empirical sampling distribution via permutation.
+\hfill\break
+
+Phylogenetic signal can be thought of as both an attribute to be
+measured in the data and a paramter that can be tuned to account for the
+phylogenetic non-independence among observations, for analysis of the
+data. As such, \(\lambda\) is appealing, as a statistic that potentially
+fulfills both roles. However, the inability to estimate phylogenetic
+signal with \(\lambda\) for data simulated with known phylogenetic
+signal is troublesome, and we recommend evolutionary biologists refrain
+from viewing it as a useful statistic to describe the amount of
+phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
+standardized to an effect size \(Z_\kappa\) -- is a better statistic for
+measuring the amount of phylogenetic signal in data simulated with
+respect to known levels of \(\lambda\). Although \(\lambda\) might be
+viewed as an important parameter for modifying the the conditional
+estimation of linear model coefficients with respect to phylogeny, it is
+neither a statistic that has meaningful comparative value as a measure
+of phylogenetic signal nor a statistic that lends itself well to
+reliable calculation of a test statistic. By contrast, \(\kappa\) has
+been shown here to be a reliable statistic, but only when standardized
+by the mean and standard deviation of its empirical sampling
+distribution (i.e., when converted to the effect size, \(Z_\kappa\)).
+Because one has control over the number of permutations used in
+analysis, one can be assured with many permutations that the empirical
+sampling distribution is representative of true probability
+distributions (Adams and Collyer 2018). With low coefficients of
+variation for \(Z_\kappa\) (Figure 4), it is difficult to imagine that a
+hypothesis test can improve equation 4 for efficiently comparing
+phylogenetic signal for different traits, different trees, or a
+combination of both.
 
@@
 
+\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
+Collyer, M. L., and D. C. Adams. 2018. RRPP: An r package for fitting
+linear models to high-dimensional data using residual randomization.
+Methods in Ecology and Evolution 9:1772--1779.
+
 \leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
@@
 35:673--679.
-
-\leavevmode\hypertarget{ref-Everitt2011}{}%
-Everitt, B., and T. Hothorn. 2011. An introduction to applied
-multivariate analysis with r. Springer Science \& Business Media.
 
@@
 biology 59:245--261.
-
-\leavevmode\hypertarget{ref-LegendreLegendre2012}{}%
-Legendre, P., and L. Legendre. 2012. Numerical ecology. 3rd ed.
-Elsevier, Amsterdam.
 
@@
 
-\textbf{Figure 4}. Variation in estimates of phylogenetic signal across
-input levels of phylogenetic signal. (A) Estimates of Pagel's
-\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
-(B) Estimates of \(Z_K\) for data simulated on phylogenies with 128 taxa
-(\(n=128\)), (C) Variance in the variation of \(\lambda_{est}\) across
-input levels of phylogenetic signal, estimated on phylogenies containing
-differing numbers of species. (D) Variance in the variation of \(Z_K\)
-across input levels of phylogenetic signal, estimated on phylogenies
-containing differing numbers of species.
+\textbf{Figure 4}. Variation in effect size estimates of phylogenetic
+signal across input levels of phylogenetic signal. (A) Estimates
+\(Z_\lambda\) for data simulated on phylogenies with 32 taxa (\(n=32\)),
+(B) Estimates of \(Z_\kappa\) for data simulated on phylogenies with 32
+taxa (\(n=32\)), (C) Coefficients of variation of precision estimates of
+\(Z_\lambda\) and \(Z_\kappa\) across input levels of phylogenetic
+signal, estimated on phylogenies containing differing numbers of
+species.
+
+\textbf{Figure 5}. (A) Linear measures for relative body size, and
+regions of the body used to estimate surface area to volume (SA:V)
+ratios. (B) Permutation distributions of phylogenetic signal for SA:V
+and \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
+95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).
 
@@
 
-\singlespacing \textbf{Figure 4}. Variation in estimates of phylogenetic
-signal across input levels of phylogenetic signal. (A) Estimates of
-Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
-(\(n=128\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
-with 128 taxa (\(n=128\)), (C) Variance in the variation of
-\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
-on phylogenies containing differing numbers of species. (D) Variance in
-the variation of \(Z_K\) across input levels of phylogenetic signal,
-estimated on phylogenies containing differing numbers of species.
+\singlespacing \textbf{Figure 4}. Variation in effect size estimates of
+phylogenetic signal across input levels of phylogenetic signal. (A)
+Estimates \(Z_\lambda\) for data simulated on phylogenies with 32 taxa
+(\(n=32\)), (B) Estimates of \(Z_\kappa\) for data simulated on
+phylogenies with 32 taxa (\(n=32\)), (C) Coefficients of variation of
+precision estimates of \(Z_\lambda\) and \(Z_\kappa\) across input
+levels of phylogenetic signal, estimated on phylogenies containing
+differing numbers of species.
+
+\includegraphics[width=0.95\linewidth]{Fig5}
+
+\singlespacing \textbf{Figure 5}. (A) Linear measures for relative body
+size, and regions of the body used to estimate surface area to volume
+(SA:V) ratios. (B) Permutation distributions of phylogenetic signal for
+SA:V and \(\frac{BW}{SVL}\), with observed values shown as vertical
+bars. (C) Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\),
+with their 95\% confidence intervals (CI not standardized by
+\(\sqrt(n)\)).
 

### 2020-05-09 VersionDiff.r

@@ -4,5 +4,5 @@
 # NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
 
 library(latexdiffr)
-latexdiff("2020-Evol-MultiRR_Vsn6.Rmd", "2020-Evol-MultiRR_Vsn6_AK.Rmd", clean = TRUE)
+latexdiff("Manuscript-5.Rmd", "Manuscript-4-EKB.Rmd", clean = TRUE)
 

### 2020-05-09 Manuscript/SupportingInfo-MLC.Rmd
SI update

--- 2020-05-06 Manuscript/SupportingInfo.Rmd
@@
 
-## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_K$)
+## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_{\kappa}$)
 
 We proposed the use of effect sizes (Z-scores) as a measure of the strength of phylogenetic signal.
-These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $Kappa$.
-The precision of effect sizes $Z_\lambda$ and $Z_K$ are summarized in the main text, primarily for simulations where $n=32$.
+These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $\kappa$.
+The precision of effect sizes $Z_\lambda$ and $Z_{\kappa}$ are summarized in the main text, primarily for simulations where $n=32$.
 Here we present results from additional simulations.
@@
 Further, the precision of $Z_\lambda$ varied widely across input levels of phylogenetic signal, with decidely greater variation (less precision) at stronger input levels of phylogenetic signal (Figure S5).
-By contrast, effect sizes for *Kappa* ($Z_K$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6).
-Additionally, the precision of $Z_K$ remained constant across input levels of phylogenetic signal (Figure S6).
-Together these results imply that $Z_K$ is a more reliable estimate of the strength of phylogenetic signal.
+By contrast, effect sizes for $\kappa$ ($Z_{\kappa}$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6).
+Additionally, the precision of $Z_{\kappa}$ remained constant across input levels of phylogenetic signal (Figure S6).
+Together these results imply that $Z_{\kappa}$ is a more reliable estimate of the strength of phylogenetic signal.
 
@@
 
-\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break
+\textbf{Figure S6}. Variation in effect size (Z score) of $\kappa$ across input $\lambda$ values on various tree sizes. \hfill\break
 
@@
 \textbf{Figure S7}.
-Estimates of phylogenetic signal using $Kappa$ across known lambda inputs on various tree sizes.
-Plot shows that while $Kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies.
-Thus, conversion to a standardized effect size ($Z_K$) is preferred.
+Estimates of phylogenetic signal using $\kappa$ across known lambda inputs on various tree sizes.
+Plot shows that while $\kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies.
+Thus, conversion to a standardized effect size ($Z_{\kappa}$) is preferred.

### 2020-05-09 Manuscript/SupportingInfo-MLC.tex
SI update

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{supporting-information-for-a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}{%
\section{Supporting Information for: A Standardized Effect Size for
Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not
Appropriate}\label{supporting-information-for-a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}}

Here we provide additional supporting information referenced in the main
document: additional analyses, and simulation results across a wider set
of input conditions.

\hypertarget{simulations-on-differently-shaped-phylogenies}{%
\subsection{Simulations on differently shaped
phylogenies}\label{simulations-on-differently-shaped-phylogenies}}

In addition to using pure-birth phylogenies, we explored the effect of
tree shape on our findings using both balanced and pectinate trees. As
before, simulations were conducted on a range of tree sizes
(\(n=2^5 - 2^{10}\)), and across a range of input levels of phylogenetic
signal (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05 units).
For each \(n\) and \(\lambda_{in}\) combination, 50 replicates of a
continuous trait were simulated using a Brownian motion model of
evolution. Using these, we estimated the degree of phylogenetic signal
using \(\lambda\). \hfill\break

\emph{Results}. As found with pure-birth phylogenies, estimates of
\(\lambda\) varied most dramatically in simulations with fewer species
and at intermediate values of lambda (Fig. S1, S2). Pectinate trees
showed an interesting tendency to underestimate \(\lambda\) across input
values. Most dramatically, some simulations on the largest pectinate
phylogeny (\(n = 1024\)) estimated \(\lambda = 0\) for input values as
high as 0.5 (Fig. S2).

\includegraphics[width=0.95\linewidth]{FigS1}

\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known
lambda inputs on various tree sizes. Results obtained using balanced
trees. \hfill\break

\includegraphics[width=0.95\linewidth]{FigS2}

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known
lambda inputs on various tree sizes. Results obtained using pectinate
trees.

\hypertarget{simulations-of-phylogenetic-regression-and-anova}{%
\subsection{Simulations of phylogenetic regression and
ANOVA}\label{simulations-of-phylogenetic-regression-and-anova}}

We analyzed the precision of parameter estimation when parameters were
obtained in phylogenetic regression and ANOVA. This involved simulations
of dependent variables with \(\lambda\) values ranging from 0 to 1 (21
intervals of 0.05 units) across pure-birth trees (\(n=2^5 - 2^{10}\)).
For each \(\lambda\) input value, independent variables were then
generated with a known relationship to the simulated dependent variable
(\(\beta = 0, 0.25, 0.5, 0.75, 1.0\)), from which \(\lambda\) was then
estimated.

\emph{Results} As with the results from the regression analyses in the
main text, we found poor precision of \(\lambda\) estimation in smaller
trees and at intermediate levels of phylogenetic signal with
phylogenetic ANOVA (Fig. S3). Slope was reliably estimated across tree
sizes and known \(\lambda\) values (Fig. S4).

\includegraphics[width=0.95\linewidth]{FigS3}

\textbf{Figure S3}. Precision of Pagel's \(\lambda\) when incorporated
in phylogenetic ANOVA. Results are from input \(\beta\) values of
\(0.5\). \hfill\break

\includegraphics[width=0.95\linewidth]{FigS4}

\textbf{Figure S4}. Precision of slope estimates across known input
\(\beta\) values of phylogenetic regression. Results shown include all
input \(\lambda\) values.

\hypertarget{effect-sizes-of-pagels-lambda-z_lambda-and-kappa-z_kappa}{%
\subsection{\texorpdfstring{Effect Sizes of Pagel's Lambda
(\(Z_\lambda\)) and Kappa
(\(Z_{\kappa}\))}{Effect Sizes of Pagel's Lambda (Z\_\textbackslash{}lambda) and Kappa (Z\_\{\textbackslash{}kappa\})}}\label{effect-sizes-of-pagels-lambda-z_lambda-and-kappa-z_kappa}}

We proposed the use of effect sizes (Z-scores) as a measure of the
strength of phylogenetic signal. These were obtained from summary
parameters of phylogenetic signal: \(\lambda\) and \(\kappa\). The
precision of effect sizes \(Z_\lambda\) and \(Z_{\kappa}\) are
summarized in the main text, primarily for simulations where \(n=32\).
Here we present results from additional simulations.

\emph{Results} Effect sizes for Pagel's \(\lambda\) (\(Z_\lambda\))
scale nonlinearly with input levels of phylogenetic signal, and this
pattern is consistent across all tree sizes examined (Figure S5).
Further, the precision of \(Z_\lambda\) varied widely across input
levels of phylogenetic signal, with decidely greater variation (less
precision) at stronger input levels of phylogenetic signal (Figure S5).
By contrast, effect sizes for \(\kappa\) (\(Z_{\kappa}\)) displayed a
linear increase with increasing input levels of phylogenetic signal,
which remained consistent across all tree sizes examined (Figure S6).
Additionally, the precision of \(Z_{\kappa}\) remained constant across
input levels of phylogenetic signal (Figure S6). Together these results
imply that \(Z_{\kappa}\) is a more reliable estimate of the strength of
phylogenetic signal.

\includegraphics[width=0.95\linewidth]{FigS5}

\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's
\(\lambda\) across input \(\lambda\) values on various tree sizes.
\hfill\break

\includegraphics[width=0.95\linewidth]{FigS6}

\textbf{Figure S6}. Variation in effect size (Z score) of \(\kappa\)
across input \(\lambda\) values on various tree sizes. \hfill\break

\includegraphics[width=0.95\linewidth]{FigS7}

\textbf{Figure S7}. Estimates of phylogenetic signal using \(\kappa\)
across known lambda inputs on various tree sizes. Plot shows that while
\(\kappa\) increases with increasing phylogenetic signal, the precision
of this estimate also varies. Thus, conversion to a standardized effect
size (\(Z_{\kappa}\)) is preferred.

\end{document}

### 2020-05-10 Manuscript/Manuscript-6-DCA.Rmd
added paragraph mentioned by Dean

@@
 
+This study is not the first to compare $\lambda$ and $\kappa$ for their ability as statistics to measure phylogenetic signal.
+Our results for $\lambda$ and $\kappa$ values are consistent with those found in the simulations performed by Münkemüller et al.
+[-@Munkemuller_et_al2012], but that study investigated type I error rates and statistical power, finding that $\lambda$ performed better in both regards, irrespective of species number in trees.
+Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent $\kappa$ at intermediate phylogenetic signal strengths was observed [see Fig.
+2, @Munkemuller_et_al2012].
+Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
+Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
+An interesting question is whether an inflated $\kappa$ value leads to an inflated $Z_\kappa$ or does a tendency of a particular tree to inflate estimates of $\kappa$ also inflate the values in random permutations of a test, in which case $Z_\kappa$ is robust to polytomies?
+We repeated the analyses in Figure 4, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (see Supporting Information).
+This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size, and because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, if the sampling distributions also shift because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+\hfill\break
+
+***Note from Mike*** I did not use Boettiger's blog as ammo because he also speculates that $\kappa$ is problematic and I did not want to point a gun at our bow.
+I think we need to add a section to the SI, which I will do, but I did a quick simulation using Molina-Venegas and Rodríguez's method for making a tree polytomy-rich and compared $\kappa$ and $Z$ from physignal analyses for $\lambda = 0$ and $\lambda = 1$ to the plots in Fig.
+4 and there is no difference.
+So I wrote the paragraph before having the full results but will obtain those results in a daya or two.
+Take-home point, any distortion of $kappa$ because of trees does not translate to distortion of $Z_{\kappa}$!
+\hfill\break
+
 Phylogenetic signal can be thought of as both an attribute to be measured in the data and a paramter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.

### 2020-05-10 Manuscript/Manuscript-6-DCA.tex
added paragraph mentioned by Dean

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[
-]{article}
+\documentclass[]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
-% Use upquote if available, for straight quotes in verbatim environments
+% use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
+% use microtype if available
+\IfFileExists{microtype.sty}{%
 \usepackage[]{microtype}
@@
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
@@
 \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
 \hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage[margin=1in]{geometry}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% set default figure placement to htbp
 \makeatletter
@@
 \makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+
 \usepackage{setspace}\doublespacing
@@
 have also been investigated to determine when phylogenetic signal can be
-detected and under what conditions (e.g., Munkemuller et al. 2012;
+detected and under what conditions (e.g., Münkemüller et al. 2012;
 Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
-Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008; Revell
+Molina-Venegas and Rodríguez 2017; see also Revell et al. 2008; Revell
 2010). One of the most widely used methods for characterizing
@@
 \subsection{\texorpdfstring{\emph{The Precision of \(\lambda\) is
-Variable}}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
+Variable}}{The Precision of \textbackslash{}lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
 
@@
 the number of species increased (Figure 2). This confirmed earlier
-findings of Boettiger et al.~(2012), and adhered to parametric
+findings of Boettiger et al. (2012), and adhered to parametric
 statistical theory. However, in many cases the set of \(\lambda_{est}\)
@@
 signal, and its variance increases (i.e., precision decreases) with
-increasing strength of phylogenetic signal (Munkemuller et al. 2012;
+increasing strength of phylogenetic signal (Münkemüller et al. 2012;
 Diniz-Filho et al. 2012: see also Supporting Information). Thus,
@@
 
+This study is not the first to compare \(\lambda\) and \(\kappa\) for
+their ability as statistics to measure phylogenetic signal. Our results
+for \(\lambda\) and \(\kappa\) values are consistent with those found in
+the simulations performed by Münkemüller et al. (2012), but that study
+investigated type I error rates and statistical power, finding that
+\(\lambda\) performed better in both regards, irrespective of species
+number in trees. Although not the central focus of their study, the same
+tendency for variable \(\lambda\) and consistent \(\kappa\) at
+intermediate phylogenetic signal strengths was observed (see Fig. 2,
+Münkemüller et al. 2012). Recent work by Molina-Venegas and Rodríguez
+(2017) found that \(\kappa\) but not \(lambda\) tended to inflate the
+estimate of phylogenetic signal, leading to moderate type I and type II
+biases, if polytomic chronograms were used. Their work more thoroughly
+addressed previous observations of inflated \(\kappa\) for incompletely
+resolved phylogenetic trees (Davies et al. 2012; Münkemüller et al.
+2012). An interesting question is whether an inflated \(\kappa\) value
+leads to an inflated \(Z_\kappa\) or does a tendency of a particular
+tree to inflate estimates of \(\kappa\) also inflate the values in
+random permutations of a test, in which case \(Z_\kappa\) is robust to
+polytomies? We repeated the analyses in Figure 4, adjusting trees to
+have 50\% collapsed nodes, per the technique of Molina-Venegas and
+Rodríguez (2017), and found results were consistent (see Supporting
+Information). This confirms that any tendency of incompletely resolved
+trees to inflate \(\kappa\) as a descriptive statistic does not inflate
+\(Z_\kappa\) as an effect size, and because comparison of effect sizes
+in a test is a comparison of locations of observed values in their
+sampling distributions, if the sampling distributions also shift because
+of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
+be robust in spite of unresolved trees. \hfill\break
+
+\textbf{\emph{Note from Mike}} I did not use Boettiger's blog as ammo
+because he also speculates that \(\kappa\) is problematic and I did not
+want to point a gun at our bow. I think we need to add a section to the
+SI, which I will do, but I did a quick simulation using Molina-Venegas
+and Rodríguez's method for making a tree polytomy-rich and compared
+\(\kappa\) and \(Z\) from physignal analyses for \(\lambda = 0\) and
+\(\lambda = 1\) to the plots in Fig. 4 and there is no difference. So I
+wrote the paragraph before having the full results but will obtain those
+results in a daya or two. Take-home point, any distortion of \(kappa\)
+because of trees does not translate to distortion of \(Z_{\kappa}\)!
+\hfill\break
+
 Phylogenetic signal can be thought of as both an attribute to be
@@
 
+\leavevmode\hypertarget{ref-Davies2012}{}%
+Davies, T. J., N. J. Kraft, N. Salamin, and E. M. Wolkovich. 2012.
+Incompletely resolved phylogenetic trees inflate estimates of
+phylogenetic conservatism. Ecology 93:242--247. Wiley Online Library.
+
 \leavevmode\hypertarget{ref-DeMeester2019}{}%
@@
 \leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
-Molina-Venegas, R., and M. A. Rodriguez. 2017. Revisiting phylogenetic
+Molina-Venegas, R., and M. A. Rodríguez. 2017. Revisiting phylogenetic
 signal; strong or negligible impacts of polytomies and branch length
@@
 \leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
-Munkemuller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
+Münkemüller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
 Schiffers, and W. Thuiller. 2012. How to measure and test phylogenetic
@@
 \leavevmode\hypertarget{ref-Orme2013}{}%
-Orme, D., R. P. Freckleton, G. H. Thomas, T. Petzoldt, S. A. Fritz, and
-N. Isaac. 2013. CAPER: Comparative analyses of phylogenetics and
-evolution in r. Methods in Ecology and Evolution 3:145--151.
+Orme, D., R. Freckleton, G. Thomas, T. Petzoldt, S. Fritz, and N. Isaac.
+2013. CAPER: Comparative analyses of phylogenetics and evolution in r.
+Methods in Ecology and Evolution 3:145--151.
 

### 2020-05-10 Manuscript/Manuscript-6-DCA.Rmd
6

@@
 
-***Note from Mike*** I did not use Boettiger's blog as ammo because he also speculates that $\kappa$ is problematic and I did not want to point a gun at our bow.
-I think we need to add a section to the SI, which I will do, but I did a quick simulation using Molina-Venegas and Rodríguez's method for making a tree polytomy-rich and compared $\kappa$ and $Z$ from physignal analyses for $\lambda = 0$ and $\lambda = 1$ to the plots in Fig.
-4 and there is no difference.
-So I wrote the paragraph before having the full results but will obtain those results in a daya or two.
-Take-home point, any distortion of $kappa$ because of trees does not translate to distortion of $Z_{\kappa}$!
-\hfill\break
-
 Phylogenetic signal can be thought of as both an attribute to be measured in the data and a paramter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.

### 2020-05-10 Manuscript/Manuscript-6-DCA.tex
6

--- 2020-05-09 Manuscript/Manuscript-6-DCA.tex
@@
 
+This study is not the first to compare \(\lambda\) and \(\kappa\) for
+their ability as statistics to measure phylogenetic signal. Our results
+for \(\lambda\) and \(\kappa\) values are consistent with those found in
+the simulations performed by Münkemüller et al.~(2012), but that study
+investigated type I error rates and statistical power, finding that
+\(\lambda\) performed better in both regards, irrespective of species
+number in trees. Although not the central focus of their study, the same
+tendency for variable \(\lambda\) and consistent \(\kappa\) at
+intermediate phylogenetic signal strengths was observed (see Fig. 2,
+Münkemüller et al. 2012). Recent work by Molina-Venegas and Rodríguez
+(2017) found that \(\kappa\) but not \(lambda\) tended to inflate the
+estimate of phylogenetic signal, leading to moderate type I and type II
+biases, if polytomic chronograms were used. Their work more thoroughly
+addressed previous observations of inflated \(\kappa\) for incompletely
+resolved phylogenetic trees (Davies et al. 2012; Münkemüller et al.
+2012). An interesting question is whether an inflated \(\kappa\) value
+leads to an inflated \(Z_\kappa\) or does a tendency of a particular
+tree to inflate estimates of \(\kappa\) also inflate the values in
+random permutations of a test, in which case \(Z_\kappa\) is robust to
+polytomies? We repeated the analyses in Figure 4, adjusting trees to
+have 50\% collapsed nodes, per the technique of Molina-Venegas and
+Rodríguez (2017), and found results were consistent (see Supporting
+Information). This confirms that any tendency of incompletely resolved
+trees to inflate \(\kappa\) as a descriptive statistic does not inflate
+\(Z_\kappa\) as an effect size, and because comparison of effect sizes
+in a test is a comparison of locations of observed values in their
+sampling distributions, if the sampling distributions also shift because
+of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
+be robust in spite of unresolved trees. \hfill\break
+
 Phylogenetic signal can be thought of as both an attribute to be

### 2020-05-11 Manuscript/Manuscript-6-EKB.Rmd
EKB's edits on Manuscript 6 (based on manuscript-6-dca.rmd)

--- 2020-05-10 Manuscript/Manuscript-6-DCA.Rmd
@@
 Tests based on $Z_\kappa$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
-Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scale.
+Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
 
@@
 One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999].
-The ($\lambda$) parameter transforms the lengths of the internal branches of the phylogeny to improve a fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
+The ~~($\lambda$)~~ parameter ($\lambda$) transforms the lengths of the internal branches of the phylogeny to improve ~~a~~ THE fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
 Pagel's $\lambda$ ranges from $0\to1$, with larger values signifying a greater dependence of observed trait variation on the phylogeny.
@@
 
-In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-The appeal of $\lambda$ as a descriptive statistic for evolutionary biologists is a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\lambda$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
+In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic ~~to describe~~ CHARACTERIZING the relative strength of phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
+~~The appeal~~ THIS DESCRIPTIVE UTILITY of $\lambda$ ~~as a descriptive statistic for evolutionary biologists~~ is ~~a~~ THE basis for interpreting "weak" versus "strong" phylogenetic signal~~,~~ (i.e., small versus large values of $\lambda$, respectively)~~,~~ in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% EXPLICITLY interpreted the strength of phylogenetic signal for at least one phenotypic trait.
 Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
@@
 
-Various other approaches use $\lambda$ as a parameter that can be varied for inferences akin to sensitivity analysis.
+Various other approaches use $\lambda$ as a parameter that can be ~~varied~~ ADJUSTED for inferences akin to sensitivity analysis.
 For instance, some have performed likelihood ratio tests that compare observed model fits to those obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or evaluated whether observed $\lambda$ differs from an expected $\lambda$, based on confidence intervals generated for the expected value [@Vandelook2019].
@@
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
-From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
+From this observation it was concluded that insufficient data (i.e., ~~the~~ LOW number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
 Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
@@
 As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
-For the case of $\lambda$, this has not yet been explored.
+~~For the case of $\lambda$, this~~ HOWEVER, THE CALCULATION AND STATISTICAL BEHAVIOR OF $\lambda$ AS AN EFFECT SIZE has not yet been explored.
 \hfill\break
@@
 Thus, $\lambda$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data.
-We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $\kappa$.
+We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $\kappa$.
 Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $\kappa$ ($Z_\kappa$), implying that $Z_\kappa$ is a more robust effect size measure.
@@
 However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; see also Supporting Information), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
-Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
+Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize ~~observed~~ KNOWN levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
 \hfill\break
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_\kappa}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Coefficients of variation in the precision of $Z_\kappa$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_\kappa$.
+Coefficients of variation in the precision of $Z_\kappa$ were up to an order of magnitude smaller ~~for~~ than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_\kappa$.
 \hfill\break
@@
 Likewise, body size (SVL) and body width (BW) measurements were taken from 3,371 individuals, and species means of relative body width ($\frac{BW}{SVL}$) were calculated [data from @Baken2019].
-A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the dataset, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species.
-The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
+A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the PHENOTYPIC dataset.
+~~, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species.~~ <!-- number of species already mentioned earlier in this paragraph--> The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
 Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**).
@@
 We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
-First, $Z_\kappa$ was more precise than $Z_{\lambda}$, and precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_\kappa$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_\kappa$.
+First, $Z_\kappa$ was more precise than $Z_\lambda$, and precision was more consistent across the range of input levels of phylogenetic signal.
+Additionally, values of $Z_\kappa$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_\kappa$ <!-- The phrasing in this sentence is a little rough.
+Perhaps: "Additionally, values of $Z_\kappa$ more accurately tracked changes in magnitude of known phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa and $\lambda_{in}$." -->.
 Thus, $Z_\kappa$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
@@
 Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
-As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
+As such, our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
 \hfill\break
@@
 One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation <!--deviations?--> -- regardless of the initial units and original scale of the phenotypic variables under investigation.
 This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
@@
 2, @Munkemuller_et_al2012].
-Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
+Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
 Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
@@
 We repeated the analyses in Figure 4, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (see Supporting Information).
-This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size, and because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, if the sampling distributions also shift because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size~~, and~~.
+FURTHERMORE, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, ~~if the sampling distributions also~~ WHICH WOULD shift EQUALLY <!-- not sure 'equally' is the right word here.
+Alternatives: "similarly" "concomitantly" "correspondingly" --> because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
 \hfill\break

### 2020-05-11 Manuscript/Manuscript-6-EKB.tex
EKB's edits on Manuscript 6 (based on manuscript-6-dca.rmd)

--- 2020-05-10 Manuscript/Manuscript-6-DCA.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-%
 \documentclass[]{article}
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
 \else % if luatex or xelatex
-  \usepackage{unicode-math}
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
@@
 }{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
+\usepackage[unicode=true]{hyperref}
 \hypersetup{
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\usepackage[normalem]{ulem}
+% avoid problems with \sout in headers with hyperref:
+\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
 
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
 \usepackage{setspace}\doublespacing
 \usepackage{lineno}\linenumbers
-
-\author{}
-\date{\vspace{-2.5em}}
+% https://github.com/rstudio/rmarkdown/issues/337
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+% https://github.com/rstudio/rmarkdown/pull/252
+\usepackage{titling}
+\setlength{\droptitle}{-2em}
+
+\pretitle{\vspace{\droptitle}\centering\huge}
+\posttitle{\par}
+
+\preauthor{\centering\large\emph}
+\postauthor{\par}
+
+\predate{\centering\large\emph}
+\postdate{\par}
+
+\date{}
 
@@
 
-\hypertarget{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}{%
 \section{A Standardized Effect Size for Evaluating the Strength of
 Phylogenetic Signal, and Why Lambda is not
-Appropriate}\label{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}}
+Appropriate}\label{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}
 
@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\section{Abstract}\label{abstract}
 
@@
 even when those traits are found in different evolutionary lineages or
-have different units or scale.
+have different units or scales.
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 have also been investigated to determine when phylogenetic signal can be
-detected and under what conditions (e.g., Münkemüller et al. 2012;
-Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
-Molina-Venegas and Rodríguez 2017; see also Revell et al. 2008; Revell
-2010). One of the most widely used methods for characterizing
+detected and under what conditions (see also Revell et al. 2008; Revell
+2010; e.g., Münkemüller et al. 2012; Pavoine and Ricotta 2012;
+Diniz-Filho et al. 2012; Adams 2014a; Molina-Venegas and Rodríguez
+2017). One of the most widely used methods for characterizing
 phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
-(Pagel 1999). The (\(\lambda\)) parameter transforms the lengths of the
-internal branches of the phylogeny to improve a fit of data to the
-phylogeny via maximum likelihood (Pagel 1999; Freckleton et al. 2002).
-Pagel's \(\lambda\) ranges from \(0\to1\), with larger values signifying
-a greater dependence of observed trait variation on the phylogeny.
-Pagel's \(\lambda\) also has the appeal that it may be included in
-phylogenetic generalized least-squares regression (PGLS) to account for
-the degree of phylogenetic signal in comparative analyses (see
-Freckleton et al. 2002). \hfill\break
+(Pagel 1999). The \sout{(\(\lambda\))} parameter (\(\lambda\))
+transforms the lengths of the internal branches of the phylogeny to
+improve \sout{a} THE fit of data to the phylogeny via maximum likelihood
+(Pagel 1999; Freckleton et al. 2002). Pagel's \(\lambda\) ranges from
+\(0\to1\), with larger values signifying a greater dependence of
+observed trait variation on the phylogeny. Pagel's \(\lambda\) also has
+the appeal that it may be included in phylogenetic generalized
+least-squares regression (PGLS) to account for the degree of
+phylogenetic signal in comparative analyses (see Freckleton et al.
+2002). \hfill\break
 
 In addition to functioning as a parameter that is tuned for appropriate
-analysis, \(\lambda\) can function as a descriptive statistic to
-describe the relative strength of phylogenetic signal in phenotypic
-traits, to determine the extent to which shared evolutionary history has
-influenced trait covariation among taxa. The appeal of \(\lambda\) as a
-descriptive statistic for evolutionary biologists is a basis for
-interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e., small
-versus large values of \(\lambda\), respectively, in a comparative sense
-(e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
-Indeed, statements regarding the strength of phylogenetic signal based
-on \(\lambda\) are rather common in the evolutionary literature. For
-instance, of the 204 papers published in 2019 that estimated and
-reported Pagel's \(\lambda\) (found from a literature survey we
-conducted in Google.scholar), 40\% interpreted the strength of
-phylogenetic signal for at least one phenotypic trait. Further, because
-nearly half of the 1,572 \(\lambda\) values reported were near 0 or 1
-(Figure 1) where the biological interpretation of \(\lambda\) is known,
-this percentage is even higher. \hfill\break
+analysis, \(\lambda\) can function as a descriptive statistic \sout{to
+describe} CHARACTERIZING the relative strength of phylogenetic signal in
+phenotypic traits to determine the extent to which shared evolutionary
+history has influenced trait covariation among taxa. \sout{The appeal}
+THIS DESCRIPTIVE UTILITY of \(\lambda\) \sout{as a descriptive statistic
+for evolutionary biologists} is \sout{a} THE basis for interpreting
+``weak'' versus ``strong'' phylogenetic signal\sout{,} (i.e., small
+versus large values of \(\lambda\), respectively)\sout{,} in a
+comparative sense (e.g., De Meester et al. 2019; Pintanel et al. 2019;
+Su et al. 2019). Indeed, statements regarding the strength of
+phylogenetic signal based on \(\lambda\) are rather common in the
+evolutionary literature. For instance, of the 204 papers published in
+2019 that estimated and reported Pagel's \(\lambda\) (found from a
+literature survey we conducted in Google.scholar), 40\% EXPLICITLY
+interpreted the strength of phylogenetic signal for at least one
+phenotypic trait. Further, because nearly half of the 1,572 \(\lambda\)
+values reported were near 0 or 1 (Figure 1) where the biological
+interpretation of \(\lambda\) is known, this percentage is even higher.
+\hfill\break
 
@@
 Various other approaches use \(\lambda\) as a parameter that can be
-varied for inferences akin to sensitivity analysis. For instance, some
-have performed likelihood ratio tests that compare observed model fits
-to those obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
-2002; Cooper et al. 2010; Bose et al. 2019) or evaluated whether
-observed \(\lambda\) differs from an expected \(\lambda\), based on
-confidence intervals generated for the expected value (Vandelook et al.
-2019). Qualitative comparisons of \(\lambda\) estimates have also been
-performed for multiple traits on the same phylogenetic tree to infer
-whether the strength of phylogenetic signal is greater in one trait as
-compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
-\hfill\break
+\sout{varied} ADJUSTED for inferences akin to sensitivity analysis. For
+instance, some have performed likelihood ratio tests that compare
+observed model fits to those obtained when \(\lambda=0\) or
+\(\lambda=1\) (Freckleton et al. 2002; Cooper et al. 2010; Bose et al.
+2019) or evaluated whether observed \(\lambda\) differs from an expected
+\(\lambda\), based on confidence intervals generated for the expected
+value (Vandelook et al. 2019). Qualitative comparisons of \(\lambda\)
+estimates have also been performed for multiple traits on the same
+phylogenetic tree to infer whether the strength of phylogenetic signal
+is greater in one trait as compared to another (e.g., Liu et al. 2019;
+Bai et al. 2019). \hfill\break
 
@@
 phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
-observation it was concluded that insufficient data (i.e., the number of
-species) was the underlying cause of the increased variation across
-parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
-common with statistical estimators, as summary statistics and parameters
-are often more precise at greater sample sizes (Cohen 1988). However,
-this conclusion also implies that the precision of \(\lambda\) remains
-constant across its range (\(\lambda = 0 \to 1\)); an assumption that to
-date, has not been verified. Thus, despite widespread use of Pagel's
-(1999) \(\lambda\) in macroevolutionary studies, at present, we lack a
-general understanding of the precision with which \(\lambda\) can
-estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
+observation it was concluded that insufficient data (i.e., \sout{the}
+LOW number of species) was the underlying cause of the increased
+variation across parameter estimates (Boettiger et al. 2012). Indeed,
+such a pattern is common with statistical estimators, as summary
+statistics and parameters are often more precise at greater sample sizes
+(Cohen 1988). However, this conclusion also implies that the precision
+of \(\lambda\) remains constant across its range
+(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
+verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
+macroevolutionary studies, at present, we lack a general understanding
+of the precision with which \(\lambda\) can estimate levels of
+phylogenetic signal in phenotypic datasets. \hfill\break
 
@@
 follows that only if \(\lambda\) is a statistical effect size can
-comparisons of estimates across datasets be interpretable. For the case
-of \(\lambda\), this has not yet been explored. \hfill\break
+comparisons of estimates across datasets be interpretable. \sout{For the
+case of \(\lambda\), this} HOWEVER, THE CALCULATION AND STATISTICAL
+BEHAVIOR OF \(\lambda\) AS AN EFFECT SIZE has not yet been explored.
+\hfill\break
 
@@
 effect size for measuring the strength of phylogenetic signal in
-phenotypic datasets, and apply the concept to two common measures of
+phenotypic datasets and apply the concept to two common measures of
 phylogenetic signal: \(\lambda\) and \(\kappa\). Through simulations we
@@
 
-\hypertarget{methods-and-results}{%
-\section{Methods and Results}\label{methods-and-results}}
-
-\hypertarget{the-precision-of-lambda-is-variable}{%
+\section{Methods and Results}\label{methods-and-results}
+
 \subsection{\texorpdfstring{\emph{The Precision of \(\lambda\) is
-Variable}}{The Precision of \textbackslash{}lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
+Variable}}{The Precision of \textbackslash{}lambda is Variable}}\label{the-precision-of-lambda-is-variable}
 
@@
 2. Taken together, these findings reveal that \(\lambda_{est}\) does not
-precisely characterize observed levels of phylogenetic signal in
-phenotypic datasets, and that biological interpretations of the strength
-of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
-\hfill\break
+precisely characterize \sout{observed} KNOWN levels of phylogenetic
+signal in phenotypic datasets, and that biological interpretations of
+the strength of phylogenetic signal based on \(\lambda\) may be highly
+inaccurate. \hfill\break
 
@@
 
-\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
 \subsection{\texorpdfstring{\emph{A Standardized Effect Size for
 Phylogenetic
-Signal}}{A Standardized Effect Size for Phylogenetic Signal}}\label{a-standardized-effect-size-for-phylogenetic-signal}}
+Signal}}{A Standardized Effect Size for Phylogenetic Signal}}\label{a-standardized-effect-size-for-phylogenetic-signal}
 
@@
 levels of phylogenetic signal. Coefficients of variation in the
-precision of \(Z_\kappa\) were up to an order of magnitude smaller for
-than for \(Z_\lambda\) (Figure 4C), implying that estimates of the
-strength of phylogenetic signal were more reliable and robust when using
-\(Z_\kappa\). \hfill\break
+precision of \(Z_\kappa\) were up to an order of magnitude smaller
+\sout{for} than for \(Z_\lambda\) (Figure 4C), implying that estimates
+of the strength of phylogenetic signal were more reliable and robust
+when using \(Z_\kappa\). \hfill\break
 
@@
 
-\hypertarget{statistical-comparisons-of-phylogenetic-signal}{%
 \subsection{\texorpdfstring{\emph{Statistical Comparisons of
 Phylogenetic
-Signal}}{Statistical Comparisons of Phylogenetic Signal}}\label{statistical-comparisons-of-phylogenetic-signal}}
+Signal}}{Statistical Comparisons of Phylogenetic Signal}}\label{statistical-comparisons-of-phylogenetic-signal}
 
@@
 
-\hypertarget{empirical-example}{%
 \subsection{\texorpdfstring{\emph{Empirical
-Example}}{Empirical Example}}\label{empirical-example}}
+Example}}{Empirical Example}}\label{empirical-example}
 
@@
 (Bonett and Blair 2017) was then pruned to match the species in the
-dataset, resulting in a phylogeny and corresponding phenotypic dataset
-containing 305 species. The phylogenetic signal in each trait was then
-characterized using \(\kappa\), which was converted to its effect size
-(\(Z_\kappa\)) using \texttt{geomorph} 3.2.1 (Adams and Otárola-Castillo
-2013; Adams et al. 2020). Finally, the strength of phylogenetic signal
-was compared across traits using \(\hat{Z}_{12}\) as described above
-(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
-acceptance}). \hfill\break
+PHENOTYPIC dataset. \sout{, resulting in a phylogeny and corresponding
+phenotypic dataset containing 305 species.} The phylogenetic signal in
+each trait was then characterized using \(\kappa\), which was converted
+to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.2.1 (Adams
+and Otárola-Castillo 2013; Adams et al. 2020). Finally, the strength of
+phylogenetic signal was compared across traits using \(\hat{Z}_{12}\) as
+described above (\textbf{to be incorporated in \texttt{geomorph} upon
+manuscript acceptance}). \hfill\break
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 taxa. However, while numerous analytical approaches may be used to
-quantify phylogenetic signal (e.g., Abouheif 1999; Gittleman and Kot
-1990; Pagel 1999; Blomberg et al. 2003; Adams 2014a), methods that
+quantify phylogenetic signal (Gittleman and Kot 1990; e.g., Abouheif
+1999; Pagel 1999; Blomberg et al. 2003; Adams 2014a), methods that
 explicitly measure the strength of phylogenetic signal, or facilitate
@@
 phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
-precise than \(Z_{\lambda}\), and precision was more consistent across
-the range of input levels of phylogenetic signal. Additionally, values
-of \(Z_\kappa\) more accurately tracked known levels of phylogenetic
+precise than \(Z_\lambda\), and precision was more consistent across the
+range of input levels of phylogenetic signal. Additionally, values of
+\(Z_\kappa\) more accurately tracked known levels of phylogenetic
 signal, with changes in the actual strength of phylogenetic signal
 reflected in a more linear fashion by concomitant changes in the values
-of \(Z_\kappa\). Thus, \(Z_\kappa\) holds promise as a measure of the
+of \(Z_\kappa\) . Thus, \(Z_\kappa\) holds promise as a measure of the
 relative strength of phylogenetic signal that reflects the magnitude of
@@
 approach developed here aligns well with other statistical approaches
-such as meta-analysis (sensu Hedges and Olkin 1985; Glass 1976; Arnqvist
+such as meta-analysis (Glass 1976; sensu Hedges and Olkin 1985; Arnqvist
 and Wooster 1995), where summary statistics across datasets are
 converted to standardized effect sizes for subsequent ``higher order''
-statistical summaries or comparisons. As such our approach enables
+statistical summaries or comparisons. As such, our approach enables
 evolutionary biologists to quantitatively examine the relative strength
@@
 Münkemüller et al. 2012). Recent work by Molina-Venegas and Rodríguez
-(2017) found that \(\kappa\) but not \(lambda\) tended to inflate the
+(2017) found that \(\kappa\) but not \(\lambda\) tended to inflate the
 estimate of phylogenetic signal, leading to moderate type I and type II
@@
 trees to inflate \(\kappa\) as a descriptive statistic does not inflate
-\(Z_\kappa\) as an effect size, and because comparison of effect sizes
-in a test is a comparison of locations of observed values in their
-sampling distributions, if the sampling distributions also shift because
-of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
-be robust in spite of unresolved trees. \hfill\break
-
-\textbf{\emph{Note from Mike}} I did not use Boettiger's blog as ammo
-because he also speculates that \(\kappa\) is problematic and I did not
-want to point a gun at our bow. I think we need to add a section to the
-SI, which I will do, but I did a quick simulation using Molina-Venegas
-and Rodríguez's method for making a tree polytomy-rich and compared
-\(\kappa\) and \(Z\) from physignal analyses for \(\lambda = 0\) and
-\(\lambda = 1\) to the plots in Fig. 4 and there is no difference. So I
-wrote the paragraph before having the full results but will obtain those
-results in a daya or two. Take-home point, any distortion of \(kappa\)
-because of trees does not translate to distortion of \(Z_{\kappa}\)!
-\hfill\break
+\(Z_\kappa\) as an effect size\sout{, and}. FURTHERMORE, because
+comparison of effect sizes in a test is a comparison of locations of
+observed values in their sampling distributions, \sout{if the sampling
+distributions also} WHICH WOULD shift EQUALLY because of this tendency,
+the \(Z_{12}\) test statistic in equation 4 appears to be robust in
+spite of unresolved trees. \hfill\break
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abouheif1999}{}%
+\hypertarget{ref-Abouheif1999}{}
 Abouheif, E. 1999. A method for testing the assumption of phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Adams2014a}{}%
+\hypertarget{ref-Adams2014a}{}
 Adams, D. C. 2014a. A generalized Kappa statistic for estimating
@@
 
-\leavevmode\hypertarget{ref-Adams2014b}{}%
+\hypertarget{ref-Adams2014b}{}
 Adams, D. C. 2014b. A method for assessing phylogenetic least squares
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
+\hypertarget{ref-AdamsCollyer2019b}{}
 Adams, D. C., and M. L. Collyer. 2019a. Comparing the strength of
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
+\hypertarget{ref-AdamsCollyer2016}{}
 Adams, D. C., and M. L. Collyer. 2016. On the comparison of the strength
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
+\hypertarget{ref-AdamsCollyer2018b}{}
 Adams, D. C., and M. L. Collyer. 2018. Phylogenetic anova: Group-clade
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
+\hypertarget{ref-AdamsCollyer2019}{}
 Adams, D. C., and M. L. Collyer. 2019b. Phylogenetic comparative methods
@@
 
-\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
-Software for geometric morphometric analyses. R package version 3.2.1.
-
-\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
+\hypertarget{ref-AdamsOtarola2013}{}
 Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
@@
 
-\leavevmode\hypertarget{ref-Arnqvist1995}{}%
+\hypertarget{ref-AdamsGeomorph}{}
+Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
+Software for geometric morphometric analyses. r package version 3.2.1.
+
+\hypertarget{ref-Arnqvist1995}{}
 Arnqvist, G., and D. Wooster. 1995. Meta-analysis: Synthesizing research
@@
 
-\leavevmode\hypertarget{ref-Bai2019}{}%
+\hypertarget{ref-Bai2019}{}
 Bai, K., S. Lv, S. Ning, D. Zeng, Y. Guo, and B. Wang. 2019. Leaf
@@
 
-\leavevmode\hypertarget{ref-Baken2019}{}%
+\hypertarget{ref-Baken2019}{}
 Baken, E. K., and D. C. Adams. 2019. Macroevolution of arboreality in
@@
 
-\leavevmode\hypertarget{ref-Baken2020}{}%
+\hypertarget{ref-Baken2020}{}
 Baken, E. K., L. E. Mellenthin, and D. C. Adams. 2020. Macroevolution of
@@
 
-\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
+\hypertarget{ref-Beaulieu_et_al2012}{}
 Beaulieu, J. M., D. C. Jhwueng, C. Boettiger, and B. C. O'Meara. 2012.
@@
 
-\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
+\hypertarget{ref-Blomberg_et_al2003}{}
 Blomberg, S. P., T. Garland, and A. R. Ives. 2003. Testing for
@@
 
-\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
+\hypertarget{ref-Boettiger_et_al2012}{}
 Boettiger, C., G. Coop, and P. Ralph. 2012. Is your phylogeny
@@
 
-\leavevmode\hypertarget{ref-Bonett2017}{}%
+\hypertarget{ref-Bonett2017}{}
 Bonett, R. M., and A. L. Blair. 2017. Evidence for complex life cycle
@@
 
-\leavevmode\hypertarget{ref-Bose2019}{}%
+\hypertarget{ref-Bose2019}{}
 Bose, R., B. R. Ramesh, R. Pélissier, and F. Munoz. 2019. Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-ButlerKing2004}{}%
+\hypertarget{ref-ButlerKing2004}{}
 Butler, M. A., and A. A. King. 2004. Phylogenetic comparative analysis:
@@
 
-\leavevmode\hypertarget{ref-Cohen1988}{}%
+\hypertarget{ref-Cohen1988}{}
 Cohen, J. 1988. Statistical power analysis for the behavioral sciences.
@@
 
-\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
+\hypertarget{ref-CollyerAdams2018}{}
 Collyer, M. L., and D. C. Adams. 2018. RRPP: An r package for fitting
@@
 
-\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
+\hypertarget{ref-Collyer_et_al2015a}{}
 Collyer, M. L., D. J. Sekora, and D. C. Adams. 2015. A method for
@@
 
-\leavevmode\hypertarget{ref-Cooper2010}{}%
+\hypertarget{ref-Cooper2010}{}
 Cooper, N., W. Jetz, and R. P. Freckleton. 2010. Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Davies2012}{}%
+\hypertarget{ref-Davies2012}{}
 Davies, T. J., N. J. Kraft, N. Salamin, and E. M. Wolkovich. 2012.
@@
 
-\leavevmode\hypertarget{ref-DeMeester2019}{}%
+\hypertarget{ref-DeMeester2019}{}
 De Meester, G., K. Huyghe, and R. Van Damme. 2019. Brain size, ecology
@@
 
-\leavevmode\hypertarget{ref-DinizFilho2012}{}%
+\hypertarget{ref-DinizFilho2012}{}
 Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
@@
 
-\leavevmode\hypertarget{ref-Felsenstein1985}{}%
+\hypertarget{ref-Felsenstein1985}{}
 Felsenstein, J. 1985. Phylogenies and the comparative method. American
@@
 
-\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
+\hypertarget{ref-Freckleton_et_al2002}{}
 Freckleton, R. P., P. H. Harvey, and M. Pagel. 2002. Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-GarlandIves2000}{}%
+\hypertarget{ref-GarlandIves2000}{}
 Garland, T. J., and A. R. Ives. 2000. Using the past to predict the
@@
 
-\leavevmode\hypertarget{ref-Gittleman1990}{}%
+\hypertarget{ref-Gittleman1990}{}
 Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
@@
 
-\leavevmode\hypertarget{ref-Glass1976}{}%
+\hypertarget{ref-Glass1976}{}
 Glass, G. V. 1976. Primary, secondary, and meta-analysis of research.
@@
 
-\leavevmode\hypertarget{ref-Grafen1989}{}%
+\hypertarget{ref-Grafen1989}{}
 Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
@@
 
-\leavevmode\hypertarget{ref-Harmon2008}{}%
+\hypertarget{ref-Harmon2008}{}
 Harmon, L. J., J. T. Weir, C. D. Brock, R. E. Glor, and W. Challenger.
@@
 
-\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
+\hypertarget{ref-HarveyPagel1991}{}
 Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
@@
 
-\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
+\hypertarget{ref-HedgesOlkin1985}{}
 Hedges, L. V., and I. Olkin. 1985. Statistical methods for
@@
 
-\leavevmode\hypertarget{ref-Kelley2012}{}%
+\hypertarget{ref-Kelley2012}{}
 Kelley, K., and K. J. Preacher. 2012. On effect size. Psychological
@@
 
-\leavevmode\hypertarget{ref-Klingenberg2010}{}%
+\hypertarget{ref-Klingenberg2010}{}
 Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
@@
 
-\leavevmode\hypertarget{ref-Liu2019}{}%
+\hypertarget{ref-Liu2019}{}
 Liu, H., C. P. Osborne, D. Yin, R. P. Freckleton, G. Jiang, and M. Liu.
@@
 
-\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
+\hypertarget{ref-MartinsHansen1997}{}
 Martins, E. P., and T. F. Hansen. 1997. Phylogenies and the comparative
@@
 
-\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
+\hypertarget{ref-MolinaVenegas2017}{}
 Molina-Venegas, R., and M. A. Rodríguez. 2017. Revisiting phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
+\hypertarget{ref-Munkemuller_et_al2012}{}
 Münkemüller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
@@
 
-\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
+\hypertarget{ref-Orme2013}{}
+Orme, D., R. Freckleton, G. Thomas, T. Petzoldt, S. Fritz, and N. Isaac.
+2013. CAPER: Comparative analyses of phylogenetics and evolution in r.
+Methods in Ecology and Evolution 3:145--151.
+
+\hypertarget{ref-OMeara_et_al2006}{}
 O'Meara, B. C., C. Ane, M. J. Sanderson, and P. C. Wainwright. 2006.
@@
 
-\leavevmode\hypertarget{ref-Orme2013}{}%
-Orme, D., R. Freckleton, G. Thomas, T. Petzoldt, S. Fritz, and N. Isaac.
-2013. CAPER: Comparative analyses of phylogenetics and evolution in r.
-Methods in Ecology and Evolution 3:145--151.
-
-\leavevmode\hypertarget{ref-Pagel1999}{}%
+\hypertarget{ref-Pagel1999}{}
 Pagel, M. D. 1999. Inferring the historical patterns of biological
@@
 
-\leavevmode\hypertarget{ref-Pavoine2012}{}%
+\hypertarget{ref-Pavoine2012}{}
 Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
@@
 
-\leavevmode\hypertarget{ref-Pintanel2019}{}%
+\hypertarget{ref-Pintanel2019}{}
 Pintanel, P., M. Tejedo, S. R. Ron, G. A. Llorente, and A.
@@
 
-\leavevmode\hypertarget{ref-RCT}{}%
+\hypertarget{ref-RCT}{}
 R Core Team. 2019. R: A language and environment for statistical
@@
 
-\leavevmode\hypertarget{ref-Revell2010}{}%
+\hypertarget{ref-Revell2010}{}
 Revell, L. J. 2010. Phylogenetic signal and linear regression on species
@@
 
-\leavevmode\hypertarget{ref-Revell2012}{}%
+\hypertarget{ref-Revell2012}{}
 Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
@@
 
-\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
+\hypertarget{ref-RevellHarmon2008}{}
 Revell, L. J., and L. J. Harmon. 2008. Testing quantitative genetic
@@
 
-\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
+\hypertarget{ref-Revell_et_al2008}{}
 Revell, L. J., L. J. Harmon, and D. C. Collar. 2008. Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Rohlf2001}{}%
+\hypertarget{ref-Rohlf2001}{}
 Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
@@
 
-\leavevmode\hypertarget{ref-Rosenthal1994}{}%
+\hypertarget{ref-Rosenthal1994}{}
 Rosenthal, R. 1994. The handbook of research synthesis. Pp. 231--244
@@
 
-\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
+\hypertarget{ref-SokalRohlf2012}{}
 Sokal, R. R., and F. J. Rohlf. 2012. Biometry. 4th ed. W.H. Freeman \&
@@
 
-\leavevmode\hypertarget{ref-Su2019}{}%
+\hypertarget{ref-Su2019}{}
 Su, G., S. Villéger, and S. Brosse. 2019. Morphological diversity of
@@
 
-\leavevmode\hypertarget{ref-Vandelook2019}{}%
+\hypertarget{ref-Vandelook2019}{}
 Vandelook, F., S. Janssens, P. Gijbels, E. Fischer, W. Van den Ende, O.
@@
 
-\hypertarget{figure-legends}{%
-\section{Figure Legends}\label{figure-legends}}
+\section{Figure Legends}\label{figure-legends}
 

### 2020-05-11 Data_Analyses/Analysis_script.R
Editing figures to use $\kappa$ instead of K

@@ -362,9 +362,9 @@ colnames(Nulls) <- c("SAV.k.null", "BWL.k.null")
 Interval<-abs(qnorm(sav_vs_bwl$sample.r.sd.pk.stand)) # CI are represented without 
 
 png("Figures/Fig5_BC.png", width = 1200, height = 600, res = 150)
-par(mfrow=c(1,2), mar = c(5,4,2,2))
+par(mfrow=c(1,2), mar = c(5,5,2,2))
  # null distribution of kappas with observed kappas
-hist(Nulls[,1], col = alpha("red",0.5), xlim = c(0,.9), border = "red", breaks = 15, yaxs = "i", xlab = "Null and Observed Kappas", main = "")
+hist(Nulls[,1], col = alpha("red",0.5), xlim = c(0,.9), border = "red", breaks = 15, yaxs = "i", xlab = latex2exp::TeX('Null and Observed $\\kappa$s'), main = "")
 hist(Nulls[,2], col = alpha("black", 0.5), border = "black", xlim = c(0,.9), breaks = 15, yaxs = "i", add = T)
 abline(v = kappa_sav$random.K[1], col = "red", lwd = 3)
 abline(v = kappa_bwl$random.K[1], col = "black", lwd = 3)
@@ -373,7 +373,7 @@ box()
 
 plotCI(x = c(1,2), y = sav_vs_bwl$sample.z, xlim = c(.8, 2.2), 
        xlab = "Morphological Traits", yaxs="i", axes = F,
-       ylim = c(0,10), pch = 19, cex = 2, ylab = latex2exp::TeX('Effect Sizes ($\\Z_K$)'),
+       ylim = c(0,10), pch = 19, cex = 2, ylab = latex2exp::TeX('Effect Sizes ($\\Z_\\kappa$)'),
        uiw = Interval, col = c("red", "black"))
 axis(2, at=0:10, labels=0:10)
 axis(1, at=c(1,2), labels = c("SA:V", "Relative Body Width"))

### 2020-05-11 Figures/Figures_Script.R

@@ -15,21 +15,21 @@ published.lambdas.df <- as.data.frame(published.lambdas.pruned)
 colnames(published.lambdas.df) <- c("Published_Lambdas", "Taxa")
 
 published.lambdas.df$Taxa_Bins <- as.character(published.lambdas.df$Taxa)
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa<=tree.sizes[[1]])] <- "Up To 32"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[1]] & published.lambdas.df$Taxa<=tree.sizes[[2]])] <- "33 to 64"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[2]] & published.lambdas.df$Taxa<=tree.sizes[[3]])] <- "65 to 128"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[3]] & published.lambdas.df$Taxa<=tree.sizes[[4]])] <- "129 to 256"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[4]] & published.lambdas.df$Taxa<=tree.sizes[[5]])] <- "257 to 512"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[5]] & published.lambdas.df$Taxa<=tree.sizes[[6]])] <- "513 to 1024"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[6]])] <- "Over 1024"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa<=tree.sizes[[1]])] <- "< 33"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[1]] & published.lambdas.df$Taxa<=tree.sizes[[2]])] <- "33 - 64"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[2]] & published.lambdas.df$Taxa<=tree.sizes[[3]])] <- "65 - 128"
+published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[3]] & published.lambdas.df$Taxa<=tree.sizes[[4]])] <- "129 - 256"
[…]

### 2020-05-11 Manuscript/Manuscript-7.tex
vsn 7

--- 2020-05-10 Manuscript/Manuscript-6-DCA.tex
@@
 phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
-(Pagel 1999). The (\(\lambda\)) parameter transforms the lengths of the
-internal branches of the phylogeny to improve a fit of data to the
+(Pagel 1999). The parameter (\(\lambda\)) transforms the lengths of the
+internal branches of the phylogeny to improve the fit of data to the
 phylogeny via maximum likelihood (Pagel 1999; Freckleton et al. 2002).
@@
 In addition to functioning as a parameter that is tuned for appropriate
-analysis, \(\lambda\) can function as a descriptive statistic to
-describe the relative strength of phylogenetic signal in phenotypic
-traits, to determine the extent to which shared evolutionary history has
-influenced trait covariation among taxa. The appeal of \(\lambda\) as a
-descriptive statistic for evolutionary biologists is a basis for
-interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e., small
-versus large values of \(\lambda\), respectively, in a comparative sense
-(e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
-Indeed, statements regarding the strength of phylogenetic signal based
-on \(\lambda\) are rather common in the evolutionary literature. For
-instance, of the 204 papers published in 2019 that estimated and
+analysis, \(\lambda\) can function as a descriptive statistic
+characterizing the relative strength of phylogenetic signal in
+phenotypic traits to determine the extent to which shared evolutionary
+history has influenced trait covariation among taxa. The appeal of
+\(\lambda\) as a descriptive statistic is that it serves as the basis
+for interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e.,
+small versus large values of \(\lambda\), respectively, in a comparative
+sense (e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al.
+2019). Indeed, statements regarding the strength of phylogenetic signal
+based on \(\lambda\) are rather common in the evolutionary literature.
+For instance, of the 204 papers published in 2019 that estimated and
 reported Pagel's \(\lambda\) (found from a literature survey we
-conducted in Google.scholar), 40\% interpreted the strength of
-phylogenetic signal for at least one phenotypic trait. Further, because
-nearly half of the 1,572 \(\lambda\) values reported were near 0 or 1
-(Figure 1) where the biological interpretation of \(\lambda\) is known,
+conducted in Google.scholar), 40\% explicitly interpreted the strength
+of phylogenetic signal for at least one phenotypic trait. Further,
+because nearly half of the 1,572 \(\lambda\) values reported were near 0
+or 1 (Figure 1) where the biological interpretation of \(\lambda\) is
+known (\(\lambda=0\) represents no phylogenetic signal, while
+\(\lambda=1\) is phylogenetic signal as expected under Brownian motion),
 this percentage is even higher. \hfill\break
@@
 Various other approaches use \(\lambda\) as a parameter that can be
-varied for inferences akin to sensitivity analysis. For instance, some
-have performed likelihood ratio tests that compare observed model fits
-to those obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
-2002; Cooper et al. 2010; Bose et al. 2019) or evaluated whether
-observed \(\lambda\) differs from an expected \(\lambda\), based on
-confidence intervals generated for the expected value (Vandelook et al.
-2019). Qualitative comparisons of \(\lambda\) estimates have also been
-performed for multiple traits on the same phylogenetic tree to infer
-whether the strength of phylogenetic signal is greater in one trait as
-compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
-\hfill\break
+adjusted for inferences that are akin to sensitivity analysis. For
+instance, some have performed likelihood ratio tests that compare
+observed model fits to those obtained when \(\lambda=0\) or
+\(\lambda=1\) (Freckleton et al. 2002; Cooper et al. 2010; Bose et al.
+2019) or evaluated whether observed \(\lambda\) differs from an expected
+\(\lambda\), based on confidence intervals generated for the expected
+value (Vandelook et al. 2019). Qualitative comparisons of \(\lambda\)
+estimates have also been performed for multiple traits on the same
+phylogenetic tree to infer whether the strength of phylogenetic signal
+is greater in one trait as compared to another (e.g., Liu et al. 2019;
+Bai et al. 2019). \hfill\break
 
@@
 bounded scale (\(0\to1\)) for which interpretation of its extremal
-points are understood. Specifically, \(\lambda=0\) represents no
-phylogenetic signal, while \(\lambda=1\) is phylogenetic signal as
-expected under Brownian motion. However, equating values of \(\lambda\)
-directly to the strength of phylogenetic signal presumes two important
+points are understood. However, equating values of \(\lambda\) directly
+to the strength of phylogenetic signal presumes two important
 statistical properties that have not been fully explored. First, it
@@
 phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
-observation it was concluded that insufficient data (i.e., the number of
+observation it was concluded that insufficient data (i.e., low number of
 species) was the underlying cause of the increased variation across
@@
 1995), or compared among datasets (e.g., Adams and Collyer 2016, 2019a).
-Unfortunatley, not all model parameters and descriptive statistics are
+Unfortunately, not all model parameters and descriptive statistics are
 effect sizes, and thus many summary measures must first be converted to
@@
 follows that only if \(\lambda\) is a statistical effect size can
-comparisons of estimates across datasets be interpretable. For the case
-of \(\lambda\), this has not yet been explored. \hfill\break
+comparisons of estimates across datasets be interpretable. However, the
+calculation and statisticl behavior of \(\lambda\) as an effect size has
+not yet been explored. \hfill\break
 
@@
 
-Finally, when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression and ANOVA, the results of our simulations were quite
-similar. Regression parameters (\(\beta\)) were accurately estimated,
-confirming earlier findings of Revell (2010) (see Supporting
-Information). However, estimates of phylogenetic signal (\(\lambda\))
-were less precise (Figure 3; see also Supporting Information), and the
-spread of \(\lambda_{est}\) was similar to that observed when
-\(\lambda\) was estimated for only the dependent variable, as in Figure
-2. Taken together, these findings reveal that \(\lambda_{est}\) does not
-precisely characterize observed levels of phylogenetic signal in
-phenotypic datasets, and that biological interpretations of the strength
-of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
-\hfill\break
+When phylogenetic signal was investigated on balanced or pectinate
+trees, patterns in the precision of \(\lambda\) were largely the same,
+with decreased precision at intermediate levels of phylogenetic signal
+(Supporting Information). Likewise, when \(\lambda\) was co-estimated
+with regression parameters in PGLS regression and ANOVA, the results of
+our simulations were quite similar. Regression parameters (\(\beta\))
+were accurately estimated, confirming earlier findings of Revell (2010)
+(see Supporting Information). However, estimates of phylogenetic signal
+(\(\lambda\)) were less precise (Figure 3; see also Supporting
+Information), and the spread of \(\lambda_{est}\) was similar to that
+observed when \(\lambda\) was estimated for only the dependent variable,
+as in Figure 2. Taken together, these findings reveal that
+\(\lambda_{est}\) does not precisely characterize known levels of
+phylogenetic signal in phenotypic datasets, and that biological
+interpretations of the strength of phylogenetic signal based on
+\(\lambda\) may be highly inaccurate. \hfill\break
 
@@
 levels of phylogenetic signal. Coefficients of variation in the
-precision of \(Z_\kappa\) were up to an order of magnitude smaller for
-than for \(Z_\lambda\) (Figure 4C), implying that estimates of the
-strength of phylogenetic signal were more reliable and robust when using
+precision of \(Z_\kappa\) were up to an order of magnitude smaller than
+for \(Z_\lambda\) (Figure 4C), implying that estimates of the strength
+of phylogenetic signal were more reliable and robust when using
 \(Z_\kappa\). \hfill\break
@@
 (Bonett and Blair 2017) was then pruned to match the species in the
-dataset, resulting in a phylogeny and corresponding phenotypic dataset
-containing 305 species. The phylogenetic signal in each trait was then
+phenotypic dataset. The phylogenetic signal in each trait was then
 characterized using \(\kappa\), which was converted to its effect size
@@
 phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
-precise than \(Z_{\lambda}\), and precision was more consistent across
-the range of input levels of phylogenetic signal. Additionally, values
-of \(Z_\kappa\) more accurately tracked known levels of phylogenetic
-signal, with changes in the actual strength of phylogenetic signal
-reflected in a more linear fashion by concomitant changes in the values
-of \(Z_\kappa\). Thus, \(Z_\kappa\) holds promise as a measure of the
-relative strength of phylogenetic signal that reflects the magnitude of
-this effect in phenotypic data. We therefore recommend that future
-studies interested in the strength of phylogenetic signal incorporate
-\(Z_\kappa\) as a statistical measure of this effect. \hfill\break
+precise than \(Z_\lambda\), and precision was more consistent across the
+range of input levels of phylogenetic signal. Additionally, values of
+\(Z_\kappa\) more accurately tracked known changes in the magnitude of
+phylogenetic signal, as demonstrated by the linear relationship between
+\(Z_\kappa\) and \(\lambda_{in}\). Thus, \(Z_\kappa\) holds promise as a
+measure of the relative strength of phylogenetic signal that reflects
+the magnitude of this effect in phenotypic data. We therefore recommend
+that future studies interested in the strength of phylogenetic signal
+incorporate \(Z_\kappa\) as a statistical measure of this effect.
+\hfill\break
 
@@
 converted to standardized effect sizes for subsequent ``higher order''
-statistical summaries or comparisons. As such our approach enables
+statistical summaries or comparisons. As such, our approach enables
 evolutionary biologists to quantitatively examine the relative strength
@@
 2012). Thus, \(Z_\kappa\) represents the strength of phylogenetic signal
-on a common and comparable scale -- measured in standard deviation --
+on a common and comparable scale -- measured in standard deviations --
 regardless of the initial units and original scale of the phenotypic
@@
 dimensionality, as estimates of phylogenetic signal using \(\kappa\)
-have been generalized for multivariate data (\(K_{mult}\): see Adams
-2014a). Furthermore, tests based on \(\hat{Z}_{12}\) may be utilized for
-comparing the strength of phylogenetic signal among datasets containing
-a different number of species, and even for phenotypes obtained from
-species in different lineages, because their phylogenetic
-non-independence and observed variation are taken into account in the
-generation of the empirical sampling distribution via permutation.
-\hfill\break
+have been generalized for multivariate data (\(\kappa_{mult}\): see
+Adams 2014a). Furthermore, tests based on \(\hat{Z}_{12}\) may be
+utilized for comparing the strength of phylogenetic signal among
+datasets containing a different number of species, and even for
+phenotypes obtained from species in different lineages, because their
+phylogenetic non-independence and observed variation are taken into
+account in the generation of the empirical sampling distribution via
+permutation. \hfill\break
 
@@
 trees to inflate \(\kappa\) as a descriptive statistic does not inflate
-\(Z_\kappa\) as an effect size, and because comparison of effect sizes
-in a test is a comparison of locations of observed values in their
-sampling distributions, if the sampling distributions also shift because
+\(Z_\kappa\) as an effect size. Furthermore, because comparison of
+effect sizes in a test is a comparison of locations of observed values
+in their sampling distributions, which would shift concomitantly because
 of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
@@
 Phylogenetic signal can be thought of as both an attribute to be
-measured in the data and a paramter that can be tuned to account for the
-phylogenetic non-independence among observations, for analysis of the
-data. As such, \(\lambda\) is appealing, as a statistic that potentially
-fulfills both roles. However, the inability to estimate phylogenetic
-signal with \(\lambda\) for data simulated with known phylogenetic
-signal is troublesome, and we recommend evolutionary biologists refrain
-from viewing it as a useful statistic to describe the amount of
-phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
+measured in the data and a parameter that can be tuned to account for
+the phylogenetic non-independence among observations, for analysis of
+the data. As such, \(\lambda\) is appealing, as a statistic that
+potentially fulfills both roles. However, the inability to estimate
+phylogenetic signal with \(\lambda\) for data simulated with known
+phylogenetic signal is troublesome, and we recommend evolutionary
+biologists refrain from viewing it as a statistic to describe the amount
+of phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
 standardized to an effect size \(Z_\kappa\) -- is a better statistic for

### 2020-05-11 Manuscript/VersionDiff.r

@@ -4,5 +4,5 @@
 # NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
 
 library(latexdiffr)
-latexdiff("Manuscript-5.Rmd", "Manuscript-6.Rmd", clean = TRUE, output = "V5-V6.diff.pdf")
+latexdiff("Manuscript-6-DCA.Rmd", "Manuscript-6-EKB.Rmd", clean = TRUE, output = "V6-V6EB.diff.pdf")
 

### 2020-07-12 Manuscript/Manuscript-7-MLC.tex
version 8 started

--- 2020-05-11 Manuscript/Manuscript-7.tex
@@
 
-In addition to functioning as a parameter that is tuned for appropriate
-analysis, \(\lambda\) can function as a descriptive statistic
-characterizing the relative strength of phylogenetic signal in
-phenotypic traits to determine the extent to which shared evolutionary
-history has influenced trait covariation among taxa. The appeal of
-\(\lambda\) as a descriptive statistic is that it serves as the basis
-for interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e.,
-small versus large values of \(\lambda\), respectively, in a comparative
-sense (e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al.
-2019). Indeed, statements regarding the strength of phylogenetic signal
-based on \(\lambda\) are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and
-reported Pagel's \(\lambda\) (found from a literature survey we
-conducted in Google.scholar), 40\% explicitly interpreted the strength
-of phylogenetic signal for at least one phenotypic trait. Further,
-because nearly half of the 1,572 \(\lambda\) values reported were near 0
-or 1 (Figure 1) where the biological interpretation of \(\lambda\) is
-known (\(\lambda=0\) represents no phylogenetic signal, while
-\(\lambda=1\) is phylogenetic signal as expected under Brownian motion),
-this percentage is even higher. \hfill\break
+In PGLS analysis, \(\lambda\) is a parameter that is tuned via
+log-likelihood profiling for its estimation. After obtaining an
+optimized \(\lambda\) value, \(\hat{\lambda}\), likelihood ratio tests
+can be performed to compare observed model fits using \(\hat{\lambda}\)
+to those obtained when \(\lambda=0\) or \(\lambda=1\) to infer whether
+phylogenetic signal differs from no signal or a Brownian motion (BM)
+model of evolutionary divergence, respectively (Freckleton et al. 2002;
+Cooper et al. 2010; Bose et al. 2019). Confidence limits on
+\(\hat{\lambda}\) can be also estimated based on percentiles of a
+\(\chi^2\) distribution and where these values are mapped onto a
+log-likelihood profile (\emph{sensu} Freckleton et al. 2002). With
+respect to confidence limits, one can ascertain whether \(\lambda=0\) or
+\(\lambda=1\) is contained within an interval of \(\hat{\lambda}\), as
+an analog to likelihood ratio tests (see, e.g., Vandelook et al. 2019).
+It is, therefore, tempting to regard \(\hat{\lambda}\) as a descriptive
+statistic that measures the relative strength of phylogenetic signal,
+for inferring the extent to which shared evolutionary history has
+influenced trait covariation among taxa, is common. The appeal of
+\(\hat{\lambda}\) as a descriptive statistic is that it serves as the
+basis for interpreting ``weak'' versus ``strong'' phylogenetic signal;
+i.e., small versus large values of \(\hat{\lambda}\), respectively, in a
+comparative sense (e.g., De Meester et al. 2019; Pintanel et al. 2019;
+Su et al. 2019). For example, qualitative comparisons of
+\(\hat{\lambda}\) for multiple traits on the same phylogenetic tree
+provide a basis for inferring whether the strength of phylogenetic
+signal is greater in one trait as compared to another (e.g., Liu et al.
+2019; Bai et al. 2019). Indeed, statements regarding the strength of
+phylogenetic signal based on \(\hat{\lambda}\) are rather common in the
+evolutionary literature. For instance, of the 204 papers published in
+2019 that estimated and reported Pagel's \(\lambda\) (found from a
+literature survey we conducted in Google.scholar), 40\% explicitly
+interpreted the strength of phylogenetic signal for at least one
+phenotypic trait. Further, because nearly half of the 1,572
+\(\hat{\lambda}\) values reported were near 0 or 1 (Figure 1) where the
+biological interpretation of \(\lambda\) is known (\(\lambda=0\)
+represents no phylogenetic signal, while \(\lambda=1\) is phylogenetic
+signal as expected under BM), this percentage is even higher.
+\hfill\break
 
@@
 
-Various other approaches use \(\lambda\) as a parameter that can be
-adjusted for inferences that are akin to sensitivity analysis. For
-instance, some have performed likelihood ratio tests that compare
-observed model fits to those obtained when \(\lambda=0\) or
-\(\lambda=1\) (Freckleton et al. 2002; Cooper et al. 2010; Bose et al.
-2019) or evaluated whether observed \(\lambda\) differs from an expected
-\(\lambda\), based on confidence intervals generated for the expected
-value (Vandelook et al. 2019). Qualitative comparisons of \(\lambda\)
-estimates have also been performed for multiple traits on the same
-phylogenetic tree to infer whether the strength of phylogenetic signal
-is greater in one trait as compared to another (e.g., Liu et al. 2019;
-Bai et al. 2019). \hfill\break
-
 It seems intuitive to interpret the strength of phylogenetic signal
-based on the value of \(\lambda\), as \(\lambda\) is a parameter on a
-bounded scale (\(0\to1\)) for which interpretation of its extremal
-points are understood. However, equating values of \(\lambda\) directly
-to the strength of phylogenetic signal presumes two important
+based on the value of \(\hat{\lambda}\), as \(\lambda\) is a parameter
+on a bounded scale (\(0\to1\)) for which interpretation of its extremal
+points are understood. However, equating values of \(\hat{\lambda}\)
+directly to the strength of phylogenetic signal presumes two important
 statistical properties that have not been fully explored. First, it
-presumes that values of \(\lambda\) can be precisely estimated, as
+presumes that values of \(\hat{\lambda}\) can be precisely estimated, as
 biological inferences regarding the strength of phylogenetic signal
-depend on high accuracy in its estimation. Therefore, understanding the
-precision in estimating \(\lambda\) is paramount. One study (Boettiger
-et al. 2012) found that estimates of Pagel's \(\lambda\) displayed less
-variation (i.e., greater precision) when data were simulated on a large
-phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
-observation it was concluded that insufficient data (i.e., low number of
-species) was the underlying cause of the increased variation across
-parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
-common with statistical estimators, as summary statistics and parameters
-are often more precise at greater sample sizes (Cohen 1988). However,
-this conclusion also implies that the precision of \(\lambda\) remains
-constant across its range (\(\lambda = 0 \to 1\)); an assumption that to
-date, has not been verified. Thus, despite widespread use of Pagel's
-(1999) \(\lambda\) in macroevolutionary studies, at present, we lack a
-general understanding of the precision with which \(\lambda\) can
-estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
+depend on high accuracy in its estimation. (In order to be precise, the
+log-likelihood profile would have to be peaked rather than flat.)
+Therefore, understanding the precision in estimating \(\hat{\lambda}\)
+is paramount. One study (Boettiger et al. 2012) found that estimates of
+Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
+when data were simulated on a large phylogeny (\(N=281\)) as compared to
+a small one (\(N=13\)). From this observation it was concluded that
+insufficient data (i.e., low number of species) was the underlying cause
+of the increased variation across parameter estimates (Boettiger et al.
+2012). Indeed, such a pattern is common with statistical estimators, as
+summary statistics and parameters are often more precise at greater
+sample sizes (Cohen 1988). However, this conclusion could imply that the
+shape of log-likelihood profiles are assumed consistent with species
+number. Such an assumption and the importance of tree shape have to date
+not been verified. Thus, despite widespread use of Pagel's (1999)
+\(\lambda\) in macroevolutionary studies, at present, we lack a general
+understanding of the precision with which \(\hat{\lambda}\) can estimate
+levels of phylogenetic signal in phenotypic datasets. \hfill\break
 
@@
 (\(0\to1\)), this does not \emph{de-facto} imply that the estimated
-values of this parameter correspond to the actual strength of the
-underlying input signal in the data. For this to be the case,
-\(\lambda\) must be a statistical effect size. Effect sizes are a
-measure of the magnitude of a statistical effect in data, represented on
-a common scale (Glass 1976; Cohen 1988). Effect sizes have widespread
-use in many areas of the quantiative sciences, as they represent
-measures that may be readily summarized across datasets as in
+values, \(\hat{\lambda}\), of this parameter correspond to the actual
+strength of the underlying input signal in the data. For this to be the
+case, \(\hat{\lambda}\) must be a statistical effect size. Effect sizes
+are a measure of the magnitude of a statistical effect in data,
+represented on a common scale (Glass 1976; Cohen 1988). Effect sizes
+have widespread use in many areas of the quantiative sciences, as they
+represent measures that may be readily summarized across datasets as in
 meta-analyses (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster
@@
 Unfortunately, not all model parameters and descriptive statistics are
-effect sizes, and thus many summary measures must first be converted to
-statistics with standardized units (i.e., conversion to an effect size)
-for meaningful comparison (see Rosenthal 1994). As a consequence, it
-follows that only if \(\lambda\) is a statistical effect size can
-comparisons of estimates across datasets be interpretable. However, the
-calculation and statisticl behavior of \(\lambda\) as an effect size has
-not yet been explored. \hfill\break
-
-In this study, we evaluate the precision of Pagel's \(\lambda\) for
-estimating known levels of phylogenetic signal in phenotypic data. We
-use computer simulations with differing numbers of species, differently
-shaped phylogenies, and differing input levels of phylogenetic signal,
-to explore the degree to which \(\lambda\) correctly identifies known
-levels of phylogenetic signal, and under what circumstances. We find
-that estimates of \(\lambda\) vary widely for a given input value of
-phylogenetic signal, and that the precision in estimating \(\lambda\) is
-not constant across its range. Rather, there is decreased precision when
-input levels of phylogenetic signal are of intermediate strength.
-Additionally, the same estimated values of \(\lambda\) may be obtained
-from datasets containing vastly different input levels of phylogenetic
-signal. Thus, \(\lambda\) is not a reliable indicator of the strength of
-phylogenetic signal in phenotypic data. We then describe a standardized
-effect size for measuring the strength of phylogenetic signal in
-phenotypic datasets and apply the concept to two common measures of
-phylogenetic signal: \(\lambda\) and \(\kappa\). Through simulations we
-find that the precision of effect sizes based on \(\lambda\)
-(\(Z_{\lambda}\)) are less reliable than that those based on \(\kappa\)
-(\(Z_\kappa\)), implying that \(Z_\kappa\) is a more robust effect size
-measure. We also propose a two-sample test statistic that may be used to
-compare the strength of phylogenetic signal among datasets, and provide
-an empirical example to demonstrate its use. We conclude that estimates
-of phylogenetic signal using Pagel's \(\lambda\) are often inaccurate,
-and thus interpreting strength of phylogenetic signal in phenotypic
-datasets based on this measure is compromised. By contrast, effect sizes
-obtained from \(\kappa\) hold promise for characterizing phylogenetic
-signal, and for comparing the strength of phylogenetic signal across
-datasets.
+effect sizes, directly, and thus many summary measures must first be
+converted to statistics with standardized units (i.e., conversion to an
+effect size) for meaningful comparison (see Rosenthal 1994). As a
+consequence, it follows that only if \(\hat{\lambda}\) is a statistical
+effect size can comparisons of estimates across datasets be
+interpretable. However, the calculation and statistical behavior of
+\(\hat{\lambda}\) as an effect size has not yet been explored.
+\hfill\break
+
+As a prortional random variable bounded by 0 and 1, we might expect that
+\(\hat{\lambda}\) follows A Bernoulli distribution; i.e., branch lengths
+in a tree are scaled proportionally to the probability that data arise
+from a BM process. As such, we would also expect in simulation
+experiment that given a known \(\lambda\) value used to generate random
+data on a tree: the mean of an empirical sampling distribution of
+\(\hat{\lambda}\) would approximately equal \(\lambda\); the standard
+error of \(\hat{\lambda}\) would be largest at intermediate values of
+\(\lambda\), and predictable over the range of \(\lambda\) with respect
+to treesize; the distribution of \(\hat{\lambda}\) would be symmetric at
+intermediate values of \(\lambda\) and more skewed toward values of 0 or
+1; and that the distribution of \(\hat{\lambda}\) will be more
+platykurtic at intermediate values of \(\lambda\), becoming more
+leptokurtic toward 0 and 1. These properties seemed to be
+superficiallyreasonable from the simulations perfomed by Münkemüller et
+al.(2012, see Fig. 2). Their study simulated strength of phylogenetic
+signal as a varied weighted-average of data simulated on trees with
+\(\lambda=0\) and \(\lambda=1\), rather than from prescribed values of
+\(\lambda\). Nonetheless, their results revealed that as a random
+variable, \(\hat{\lambda}\) resembled a Bernoulli random variable based
+superficially on statistical moments (mean, variance, skewness, and
+kurtosis; see their Fig. 2) with respect to strength of phylogenetic
+signal, for a given tree size. However, for comparisons among tree sizes
+for a given strength of phylogenetic signal, median values of
+\(\hat{\lambda}\) were slightly, positively associated with the number
+of species in a tree, especially at weaker strength of signals,
+suggesting \(\hat{\lambda}\) might not be an accurate statistic across a
+range of tree sizes. In order for \(\hat{\lambda}\) to be considered a
+reliable statistic as a measure of phylogenetic signal, as a minimum, it
+should be accurate and precise, properties that could be measured from
+simulation experiments for known values of \(\lambda\).
+
+In this study, we evaluate the precision of \(\hat{\lambda}\) as an
+optimized parameter, used as a statistic, for estimating known levels of
+phylogenetic signal in phenotypic data (data simulated with known
+\(\lambda\)). We use computer simulations with differing numbers of
+species, differently shaped phylogenies, and differing input levels of
+\(\lambda\), to explore the degree to which \(\hat{\lambda}\) correctly
+identifies known levels of phylogenetic signal, \(\lambda\), and under
+which circumstances. We find that estimates of \(\hat{\lambda}\) vary
+widely for a given input value of \(\lambda\), and that the precision of
+\(\hat{\lambda}\) is not constant across its range. Rather, there is
+decreased precision when input levels of phylogenetic signal are of
+intermediate strength. Additionally, the same values of
+\(\hat{\lambda}\) may be obtained from datasets containing vastly
+different input levels of \(\lambda\). Thus, \(\hat{\lambda}\) is not a
+reliable indicator of the strength of phylogenetic signal in phenotypic
+data. We then describe a standardized effect size for measuring the
+strength of phylogenetic signal in phenotypic datasets and apply the
+concept to two common measures of phylogenetic signal: \(\hat{\lambda}\)
+and \(\kappa\). Through simulations we find that the precision of effect
+sizes based on \(\hat{\lambda}\) (\(Z_{\lambda}\)) are less reliable
+than that those based on \(\kappa\) (\(Z_\kappa\)), implying that
+\(Z_\kappa\) is a more robust effect size measure. We also propose a
+two-sample test statistic that may be used to compare the strength of
+phylogenetic signal among datasets, and provide an empirical example to
+demonstrate its use. We conclude that estimates of phylogenetic signal
+using Pagel's \(\lambda\) (as a descriptive statistic, not as a PGLS
+parameter) are often inaccurate, and thus interpreting strength of
+phylogenetic signal in phenotypic datasets based on this measure is
+compromised. By contrast, effect sizes obtained from \(\kappa\) hold
+promise for characterizing phylogenetic signal, and for comparing the
+strength of phylogenetic signal across datasets.
 
@@
 
-\hypertarget{the-precision-of-lambda-is-variable}{%
-\subsection{\texorpdfstring{\emph{The Precision of \(\lambda\) is
-Variable}}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}
+\hypertarget{the-precision-of-hatlambda-is-variable}{%
+\subsection{\texorpdfstring{\emph{The Precision of \(\hat{\lambda}\) is
+Variable}}{The Precision of \textbackslash{}hat\{\textbackslash{}lambda\} is Variable}}\label{the-precision-of-hatlambda-is-variable}}
 
 We conducted a series of computer simulations to evaluate the precision
-of Pagel's \(\lambda\). Our primary simulations were based on pure-birth
+of \(\hat{\lambda}\) based on varied input of Pagel's \(\lambda\) to
+generate data. Our primary simulations were based on pure-birth
 phylogenies; however, we also evaluated patterns on both balanced and
@@
 phylogenies at each of six different tree sizes, ranging from 32 to 1024
-taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the simulated phylogenies
-by multiplying the internal branches by \(\lambda_{in}\), using 21
-intervals of 0.05 units across its range
+taxa (\(n=2^5 - 2^{10}\)). For each simulation run, we rescaled the
+simulated phylogenies by multiplying the internal branches by
+\(\lambda_{in}\), using 21 intervals of 0.05 units across its range
 (\(\lambda_{in} = 0.0 \to 1.0\)), resulting in 1050 scaled phylogenies
 at each level of species richness (\(n\)). Continuous traits were then
-simulated on each phylogeny under a Brownian motion model of evolution
-to obtain datasets with differing levels of phylogenetic signal, that
-ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
-phylogenetic signal reflecting Brownian motion (when
-\(\lambda_{in} =1\)). For each dataset we then estimated phylogenetic
-signal (\(\lambda_{est}\)), and calculated the variance of \(\lambda\)
-(\(\sigma^2_\lambda\)) across datasets at each input level of
-phylogenetic signal and level of species richness as an estimate of
-precision. We verified that the variance of traits simulated had no
-effect on phylogenetic signal estimation. \hfill\break
-
-We also evaluated the precision of \(\lambda\) when estimated in PGLS
-regression and ANOVA (i.e., \(Y\sim{X}\)). Here, an independent variable
-\(X\) was simulated on each rescaled phylogeny under a Brownian motion
-model of evolution (for PGLS regression). For phylogenetic ANOVA, random
-groups (\(X\)) were obtained by simulating a discrete (binary, 0 or 1)
-character on each phylogeny. Next, the dependent variable was simulated
-in such a manner as to contain a known relationship with \(X\) plus
-random error containing phylogenetic signal. This was accomplished as:
-\(Y=\beta{X}+\epsilon\). The association between \(Y\) and \(X\) was
-modeled using a range of values: \(\beta=(0.0,0.25, 0.5, 0.75,1.0)\),
-and the residual error (\(\epsilon\)) was modeled to contain
-phylogenetic signal simulated under a Brownian motion model of evolution
-on each rescaled phylogeny:
-\(\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\): (see Revell
-2010 for a similar simulation design). The fit of the phylogenetic
-regression was estimated using maximum likelihood, and parameter
-estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were obtained. We then
-calculated precision estimates (\(\sigma^2_\lambda\)) at each input
-level of phylogenetic signal and level of species richness. We verified
-that the amount of residual variance simulated had no effect on
-\(\sigma^2_\lambda\) but did influence the precision of coefficients
-estimated from the linear model (precision increased with smaller
-\(\epsilon\), as expected). \hfill\break
+simulated on each phylogeny under a BM model of evolution to obtain
+datasets with differing levels of phylogenetic signal, that ranged from
+no phylogenetic signal (when \(\lambda_{in} =0\)), to phylogenetic
+signal reflecting a BM model of evolution (when \(\lambda_{in} =1\)).
+For each dataset we then estimated phylogenetic signal
+(\(\hat{\lambda}\)), and calculated its variance
+(\(\sigma^2_{\hat{\lambda}}\)) at each input level of phylogenetic
+signal and level of species richness as an estimate of precision. Random
+data samples were generated from a normal distribution, i.e.,
+\(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\). We verified that the
+level of \(\sigma^2\) had no effect on \(\hat{\lambda}\) or
+\(\sigma^2_{\hat{\lambda}}\), but the results we present used
+\(\sigma^2=1\). \hfill\break
+
+We also evaluated the precision of \(\hat{\lambda}\) when estimated in
+PGLS regression and ANOVA (i.e., \(Y\sim{X}\)). For these simulations,
+\(Y=\beta{X}+\epsilon\) was used to generate data with varied
+phylogenetic signal and variable associations (\emph{sensu} Revell
+2010). In the case of PGLS regression, \(X\) was a variable simulated
+under a BM model of evolution, constrained by rescaling phylogeny
+according to \(\lambda\), and in the case of ANOVA, \(X\) was a balanced
+binary variable (values of 0 and 1), randomly ordered. The strength of
+variable associations ranged as \(\beta=0.0,0.25, 0.5, 0.75\) and
+\(1.0\). Residuals, \(\epsilon\), were simulated under a BM model of
+evolution, constrained by rescaling phylogeny according to \(\lambda\).
+Simulated independent variables or residuals were generated from a
+normal distribution, i.e.,
+\(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\), with the phylogenetic
+covariance matrix, \(\mathbf{C}\), implicitly varied for each tree by
+the level of \(\lambda\). The fit of the phylogenetic regression was
+estimated using maximum likelihood, and parameter estimates
+(\(\hat{\lambda}\) and \(\hat{\beta}\)) were obtained. We then
+calculated precision estimates (\(\sigma^2_{\hat{\lambda}}\)) at each
+input level of phylogenetic signal and level of species richness. We
+verified that the level of \(\sigma^2\) had no effect on
+\(\sigma^2_{\hat{\lambda}}\) but did influence the precision of
+\(\hat{\beta}\) estimated from the linear model (precision increased
+with smaller \(\epsilon\), as expected). The results we present used
+\(\sigma^2=1\). \hfill\break
 
@@
 \begin{align}
-    Z_\kappa=\frac{log(\kappa_{obs})-\hat\mu_{log(\kappa)}}{\hat\sigma_{log(\kappa)}}
+    Z_\kappa=\frac{log(\kappa_{obs})-\hat\mu_{log(\kappa)}}{\hat\sigma_{log(\kappa)}},
 \end{align}
@@
 
-An effect size based on \(\lambda\) could be envisioned, which is found
-as:
+An effect size based on \(\lambda\) is a little more challenging
+because, as a statistic, it does not have a sampling distribution on
+which to measure standard error. Confidence intervals for \(\lambda\)
+are not generated from standard error calculations, but rather for the
+values of lambda that intersect the log-likehihood profile for
+corresponding percentiles of the \(\chi^2\) distribution used to compare
+the compare the putative appropriate model to a null model with
+\(\lambda = 0\) {[}citation needed, think Boettiger has a paper on
+this{]}. The simplest similar effect size is
 
 \begin{align}
-    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
+   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
 \end{align}
+
+where, \(\hat{\lambda}\) is the maximized likelihood value of
+\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
+statistic for the value. Hypothetically, this equation would be better
+represented as \(Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}\), where
+\(d\) is a binary value \((-1,1)\) to indicate a direction based on
+whether \(\hat{\lambda}\) is below or above a critical value of
+\(\lambda\), for a quantile from a \(\chi^{2}\) distribution at a
+probability of 0.5. The critical value would correspond to a value of
+\(Z = 0\); i.e., the strength of phylogenetic signal that might be
+observed at the 50th percentile in a distribution of random outcomes,
+generated from a model where \(\lambda = 0\). However, in some
+preliminary simulations, we found that mapping
+\(\chi^2_{\hat{\lambda}}\) values this way did not produce effect sizes
+that were symmetrical about \(Z = 0\), as the mapping is not linear and,
+especially for small trees, the log-likelihood profiles can be rather
+flat, making estimation of \(\hat{\lambda}\) rather imprecise (as we
+elaborate more below.) We will heneforth use equation 3, recognizing
+that large values of \(Z\) would likely be positive, anyway, based on
+the reference model for the likelihood ratio test.
 
@@
 for \(\lambda\) and \(\kappa\) values are consistent with those found in
-the simulations performed by Münkemüller et al.~(2012), but that study
+the simulations performed by Münkemüller et al. (2012), but that study
 investigated type I error rates and statistical power, finding that

### 2020-07-12 Manuscript/Manuscript-8.Rmd
version 8 started

---
title: A Standardized Effect Size for Evaluating the Comparing the Strength of Phylogenetic Signal

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: pnas.csl
bibliography: Lambda-Refs.bib

header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---

\begin{center}
\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
\end{center}

\begin{center}`r format(Sys.time(), '%d %B, %Y')`\end{center}

$^{1}$Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.

$^{2}$Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.

$^{*}$Correspondence: Dean C. Adams <dcadams@iastate.edu> 
\hfill\break

**Keywords**: phylogenetic signal, lambda, kappa \hfill\break

**Short Title**: Effect size for phylogenetic signal \hfill\break

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(stringr)
library(tidyverse)

RmdWords <- function(file) {

  # Creates a string of text
  file_string <- file %>%
    readLines() %>%
    paste0(collapse = " ") %>%
    # Remove YAML header
    str_replace_all("^<--- .*?--- ", "") %>%    
    str_replace_all("^--- .*?--- ", "") %>%
    # Remove code
    str_replace_all("```.*?```", "") %>%
    str_replace_all("`.*?`", "") %>%
    # Remove LaTeX
    str_replace_all("[^\\\\]\\$\\$.*?[^\\\\]\\$\\$", "") %>%
    str_replace_all("[^\\\\]\\$.*?[^\\\\]\\$", "") %>%
    # Deletes text between tags
    str_replace_all("TC:ignore.*?TC:endignore", "") %>%
    str_replace_all("[[:punct:]]", " ") %>%
    str_replace_all("  ", "") %>%
    str_replace_all("<", "") %>%
    str_replace_all(">", "")

  # Save several different results
  word_count <- str_count(file_string, "\\S+")
  char_count <- str_replace_all(string = file_string, " ", "") %>% str_count()

   return(list(num_words = word_count, num_char = char_count, word_list = file_string))
}

words <- RmdWords("Manuscript-8.Rmd")
```

```{r, echo = FALSE}
count <- c(words = words$num_words, characters = words$num_char)
knitr::kable(t(count))
```

\newpage

# Abstract

Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits. However, analytical tools for such comparisons have largely remained underdeveloped. In this study, we evaluated the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal. We find that the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised. We then propose a standardized effect size based on $\kappa$ ($Z_\kappa$), which measures the strength of phylogenetic signal, and places it on a common scale for statistical comparison. Tests based on $Z_\kappa$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

\newpage

# Introduction

Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates an assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991]. Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004]. The past several decades have witnessed an impressive expansion in the development of PCMs to address an ever-growing set of macroevolutionary hypotheses [@MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @Adams2014a; @AdamsCollyer2018b]. These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003]. Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [see @Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008]. \hfill\break

Several analytical tools have been developed to quantify phylogenetic signal in phenotypic datasets, including measures of serial independence [$\mathbf{C}$: @Abouheif1999], autocorrelation estimates [$I$: @Gittleman1990], statistical ratios of trait variation relative to what is expected given the phylogeny [$\kappa$: @Blomberg_et_al2003; @Adams2014a], and scaling parameters used in maximum likelihood fitting of the data to the phylogeny [$\lambda$: @Pagel1999], among others [e.g., @Klingenberg2010]. The statistical properties of these methods -- namely type I error rates and statistical power -- have also been investigated to determine under what conditions phylogenetic signal can be detected [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @Adams2014a; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010]. One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999]. The parameter ($\lambda$) transforms the lengths of the internal branches of the phylogeny to improve the fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002]. Blomberg's $\kappa$ [@Blomberg_et_al2003] is also widely used.  This statistic measures the ratio of trait variation to the amount expected under a Brownian motion (BM) model of evolution.  Pagel's $\lambda$ and Blomberg's $\kappa$ have some similar intuitive appeal, namely because for both a value of $0$ corresponds to no phyloegentic signal in the data and a value of $1$ corresponds to the amount of phylogenetic signal one would expect with a BM model of evolution.  However, as putative measures (statistics) of the amount of phylogenetic signal, they are calculated quite differently.  \hfill\break

In PGLS analysis, Pagel's $\lambda$ is a parameter that ranges from $0$ to $1$ and is optimized via log-likelihood profiling.  After obtaining an optimized $\lambda$ value, $\hat{\lambda}$, likelihood ratio tests can be performed to compare observed model fits using $\hat{\lambda}$ to those obtained when $\lambda=0$ or $\lambda=1$ to infer whether phylogenetic signal differs from no signal or a BM model of evolutionary divergence, respectively [@Freckleton_et_al2002; @Cooper2010; @Bose2019].  Confidence limits on $\hat{\lambda}$ can also be estimated based on percentiles of a $\chi^2$ distribution and where these values are mapped onto a log-likelihood profile [*sensu* @Freckleton_et_al2002].  With respect to confidence limits, one can ascertain whether $\lambda=0$ or $\lambda=1$ is contained within an interval of $\hat{\lambda}$, as an analog to likelihood ratio tests [see, e.g., @Vandelook2019].  It is, therefore, tempting to regard $\hat{\lambda}$ as a descriptive statistic that measures the relative strength of phylogenetic signal, for inferring the extent to which shared evolutionary history has influenced trait covariation among taxa.  By contrast, Blomberg's $\kappa$ is a descriptive statistic that measures the ratio of obsrved trait variation to the amount expected under a BM model of evolution.  Like Pagel's $\lambda$, the range of $0$ to $1$ defines a greater dependence of observed trait variation on the phylogeny but values greater than 1 are possible when phylogenetic signal is especially strong.  Additionally, Blomberg's $\kappa$ can be a test statistic, by employing a permutation test to generate its sampling distribution [@Adams2014a].  Therefore, both statistics have a descriptive meaning, plus an associated test for the null hypothesis of no phlogenetic signal.  An important question that has yet to considered is whether these statistics are or can be converted to effect sizes for comparative analyses?  \hfill\break

The appeal of either Pagel's $\lambda$ or Blomberg's $\kappa$ as descriptive statistics is that they provide a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$ or $\kappa$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. As statistics, they should have reliable distributional properties, which could be revealed with simulation experiments.  As a proportional random variable bounded by 0 and 1, we might expect that $\hat{\lambda}$ follows a Bernoulli distribution (**add ref**); i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.  As such, we would also expect in a simulation experiment that given a known $\lambda$ value used to generate random data on a tree that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, and predictable over the range of $\lambda$ with respect to treesize; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 (**add same ref**). These properties seemed to be reasonable from the simulations perfomed by Münkemüller et al.[-@Munkemuller_et_al2012, see Fig. 2].  Their study simulated "strength of Brownian motion" as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$, rather than from prescribed values of $\lambda$.  Nonetheless, across the range of BM strength, $\hat{\lambda}$ resembled a Bernoulli random variable, based superficially on statistical moments (mean, variance, skewness, and kurtosis; see their Fig. 2), for a given tree size. \hfill\break

By contrast, with Blomberg's $\kappa$, which is positively ubounded, we might expect that $\hat{\lambda}$ follows a normal distribution (**add ref**).  A simulation experiment should reveal that $\kappa$ is symmetrically distributed across different strengths of phylogenetic signal, for any $\lambda$ used to generate data.  This attribute seemed less reasonable based on the simulations performed by Münkemüller et al. [-@Munkemuller_et_al2012], as distibutions were postively skewed, suggesting the Blomberg's $\kappa$ might not behave as a statistic that follows a noraml distribution.  However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations).  For both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.  Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.  Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparative anlyses. \hfill\break

In this study, we used simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and Blomberg's $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.  We present results based on pure-birth trees, but results from blanced or pectinate trees were qualitatively no different (see Supporting Information).  Results from trees with polytomies, or from linear regression or phylogenetic ANOVA, were also qualitatively no different (see Supporting Information).

# Results

Our simulations involved tree sizes of $2^5$, $2^6$, ..., $2^{10}$ and $\lambda$ of 0, 0.05, 0.10, ..., 1.0, with 100 random trees for each intersection of tree size and $\lambda$.  For each $\lambda$ within each tree size, we measured the mean values of $\hat{\lambda}$ and Blomberg's $\kappa$, their standard deviation, and a Shapiro-Wilk  $W$ statistic as a departure from normality (symmetry), with a value of $1.0$ indicating normally distributed values.  Departures from $1.0$ indicate skewness.  Kurtosis was not measured directly but could be qualitatively evaluated from plots. \hfill\break

In general, the following six results were clear (Fig. 1).  First, for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.  The mean value of $\hat{\lambda}$ increased as $\lambda$ increased; the standard deviation of $\hat{\lambda}$ was largest at intermediate values of $\lambda$ and smallest at extreme values; and the distributions of $\hat{\lambda}$ were increasing skewed at extreme values of $\lambda$ but rather normal at intermediate values.  (For small tree sizes, it was also clear that distributions were more platykurtik at intermeidate values of $\hat{\lambda}$; see Fig. 1).  However, second, $\hat{\lambda}$ was a negatively-biased estimate of $\lambda$ for small tree sizes, as the mean value was consistently less than the input $\lambda$ value.  Third, standard deviations of $\hat{\lambda}$ values were negatively associated with tree size.  In fact, for tress of 128 species or less, $\hat{\lambda}$ was quite variable, except for cases when $\lambda$ was near or equal to 1.  Fourth, for $\kappa$, mean values increased consistently with $\lambda$, irespective of tree size; standard deviation of $\kappa$ was consistent across tree sizes; and $\kappa$ was normally distributed regarless of $\kappa$ or tree size, in general (Fig. 2).  Fifth, standard deviation of $\kappa$ increased with $\lambda$, somewhat tracking the mean, but were always less than corresponding means.  This is perhaps unsurprising, as $\kappa$ is bound by 0, and was never large for small values of $\lambda$.  Finally, sixth, the only evidence of skewing of $\kappa$ distributions was at small tree sizes and $\lambda$ near or equal to 1, which differs from the results of Münkemüller et al. [-@Munkemuller_et_al2012], who used weighted combinations of simulated data.  The skewing observed in their simulation results appears to be from combining random values generated independently.  \hfill\break

In terms of effect sizes, $Z$-scores were associated with both phylogenetic signal strength ($\lambda$) and tree size, for both statistics.  Effect sizes from $\hat{\lambda}$ made little sense.  Effect sizes were exceedingly associated with tree size, as much or more so than  phylogenetic signal strength (Fig. 3).  By contrast, standard deviates from $\kappa$ were much more consistent accross tree sizes, and exhibited only a slight positive association with tree size despite a clear positive association with phylogenetic signal strength.  Between the two statistics, effect sizes from $\kappa$ would be reliable for cross-study comparisons but effect sizes from $\hat{\lambda}$ would require comparisons limited to phylogenies with similar numbers of species.

# Discussion

\newpage

Real figures will have 6 panels for lambda and kappa and probably only a subset for z.  Simulations need to be re-run.

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.1.temp.png")

```

\singlespacing \textbf{Figure 1}.  Temporary Fig. 1 showing lambda patterns

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.2.temp.png")

```

\singlespacing \textbf{Figure 2}.  Temporary Fig. 2 showing kappa patterns

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.3.temp.png")

```

\singlespacing \textbf{Figure 3}.  Temporary Fig. 3 showing z patterns

\newpage

### 2020-07-12 Manuscript/Manuscript-8.tex
version 8 started

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={A Standardized Effect Size for Evaluating the Comparing the Strength of Phylogenetic Signal},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

\title{A Standardized Effect Size for Evaluating the Comparing the Strength of
Phylogenetic Signal}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{center}
\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
\end{center}

\begin{center}12 July, 2020\end{center}

\(^{1}\)Department of Ecology, Evolution, and Organismal Biology, Iowa
State University, Ames, Iowa, USA.

\(^{2}\)Department of Science, Chatham University, Pittsburgh,
Pennsylvania, USA.

\(^{*}\)Correspondence: Dean C. Adams
\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}
\hfill\break

\textbf{Keywords}: phylogenetic signal, lambda, kappa \hfill\break

\textbf{Short Title}: Effect size for phylogenetic signal \hfill\break

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
1751 & 10974\tabularnewline
\bottomrule
\end{longtable}

\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. In this study, we evaluated the
efficacy of one commonly used parameter (Pagel's \(\lambda\)) to
estimate the strength of phylogenetic signal in phenotypic traits, and
evaluate the degree to which \(\lambda\) correctly identifies known
levels of phylogenetic signal. We find that the precision of \(\lambda\)
in estimating actual levels of phylogenetic signal is often inaccurate,
and that biological interpretations of the strength of phylogenetic
signal based on \(\lambda\) are therefore compromised. We then propose a
standardized effect size based on \(\kappa\) (\(Z_\kappa\)), which
measures the strength of phylogenetic signal, and places it on a common
scale for statistical comparison. Tests based on \(Z_\kappa\) provide a
mechanism for formally comparing the strength of phylogenetic signal
across datasets, in much the same manner as effect sizes may be used to
summarize patterns in quantitative meta-analysis. Our approach extends
the phylogenetic comparative toolkit to address hypotheses that compare
the strength of phylogenetic signal between various phenotypic traits,
even when those traits are found in different evolutionary lineages or
have different units or scales.

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates an assumption of independence among trait values that is common
for statistical tests (1, 2). Accounting for this evolutionary
non-independence is the purview of \emph{phylogenetic comparative
methods} (PCMs): a suite of analytical tools that condition trends in
the data on the phylogenetic relatedness of observations (3--6). The
past several decades have witnessed an impressive expansion in the
development of PCMs to address an ever-growing set of macroevolutionary
hypotheses (7--13). These methods are predicated on the notion that
phylogenetic signal -- the tendancy for closely related species to
display similar trait values -- is present in cross-species datasets (1,
14, 15). Indeed, under numerous evolutionary models, phylogenetic signal
is to be expected, as stochastic character change along the hierarchical
structure of the tree of life generates trait covaration among related
taxa (1, 15, 16). \hfill\break

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets, including measures of serial independence
(17), autocorrelation estimates (18), statistical ratios of trait
variation relative to what is expected given the phylogeny (12, 15), and
scaling parameters used in maximum likelihood fitting of the data to the
phylogeny (14), among others (19). The statistical properties of these
methods -- namely type I error rates and statistical power -- have also
been investigated to determine under what conditions phylogenetic signal
can be detected (12, 16, 20--24). One of the most widely used methods
for characterizing phylogenetic signal in macroevolutionary studies is
Pagel's \(\lambda\) (14). The parameter (\(\lambda\)) transforms the
lengths of the internal branches of the phylogeny to improve the fit of
data to the phylogeny via maximum likelihood (14, 25). Blomberg's
\(\kappa\) (15) is also widely used. This statistic measures the ratio
of trait variation to the amount expected under a Brownian motion (BM)
model of evolution. Pagel's \(\lambda\) and Blomberg's \(\kappa\) have
some similar intuitive appeal, namely because for both a value of \(0\)
corresponds to no phyloegentic signal in the data and a value of \(1\)
corresponds to the amount of phylogenetic signal one would expect with a
BM model of evolution. However, as putative measures (statistics) of the
amount of phylogenetic signal, they are calculated quite differently.
\hfill\break

In PGLS analysis, Pagel's \(\lambda\) is a parameter that ranges from
\(0\) to \(1\) and is optimized via log-likelihood profiling. After
obtaining an optimized \(\lambda\) value, \(\hat{\lambda}\), likelihood
ratio tests can be performed to compare observed model fits using
\(\hat{\lambda}\) to those obtained when \(\lambda=0\) or \(\lambda=1\)
to infer whether phylogenetic signal differs from no signal or a BM
model of evolutionary divergence, respectively (25--27). Confidence
limits on \(\hat{\lambda}\) can also be estimated based on percentiles
of a \(\chi^2\) distribution and where these values are mapped onto a
log-likelihood profile (25). With respect to confidence limits, one can
ascertain whether \(\lambda=0\) or \(\lambda=1\) is contained within an
interval of \(\hat{\lambda}\), as an analog to likelihood ratio tests
(28). It is, therefore, tempting to regard \(\hat{\lambda}\) as a
descriptive statistic that measures the relative strength of
phylogenetic signal, for inferring the extent to which shared
evolutionary history has influenced trait covariation among taxa. By
contrast, Blomberg's \(\kappa\) is a descriptive statistic that measures
the ratio of obsrved trait variation to the amount expected under a BM
model of evolution. Like Pagel's \(\lambda\), the range of \(0\) to
\(1\) defines a greater dependence of observed trait variation on the
phylogeny but values greater than 1 are possible when phylogenetic
signal is especially strong. Additionally, Blomberg's \(\kappa\) can be
a test statistic, by employing a permutation test to generate its
sampling distribution (12). Therefore, both statistics have a
descriptive meaning, plus an associated test for the null hypothesis of
no phlogenetic signal. An important question that has yet to considered
is whether these statistics are or can be converted to effect sizes for
comparative analyses? \hfill\break

The appeal of either Pagel's \(\lambda\) or Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (29--31). As statistics, they should have reliable
distributional properties, which could be revealed with simulation
experiments. As a proportional random variable bounded by 0 and 1, we
might expect that \(\hat{\lambda}\) follows a Bernoulli distribution
(\textbf{add ref}); i.e., branch lengths in a tree are scaled
proportionally to the probability that data arise from a BM process. As
such, we would also expect in a simulation experiment that given a known
\(\lambda\) value used to generate random data on a tree that the mean
of an empirical sampling distribution of \(\hat{\lambda}\) would
approximately equal \(\lambda\); the dispersion of \(\hat{\lambda}\)
would be largest at intermediate values of \(\lambda\), and predictable
over the range of \(\lambda\) with respect to treesize; the distribution
of \(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). These properties seemed to be reasonable
from the simulations perfomed by Münkemüller et al.(20). Their study
simulated ``strength of Brownian motion'' as a varied weighted-average
of data simulated on trees with \(\lambda=0\) and \(\lambda=1\), rather
than from prescribed values of \(\lambda\). Nonetheless, across the
range of BM strength, \(\hat{\lambda}\) resembled a Bernoulli random
variable, based superficially on statistical moments (mean, variance,
skewness, and kurtosis; see their Fig. 2), for a given tree size.
\hfill\break

By contrast, with Blomberg's \(\kappa\), which is positively ubounded,
we might expect that \(\hat{\lambda}\) follows a normal distribution
(\textbf{add ref}). A simulation experiment should reveal that
\(\kappa\) is symmetrically distributed across different strengths of
phylogenetic signal, for any \(\lambda\) used to generate data. This
attribute seemed less reasonable based on the simulations performed by
Münkemüller et al. (20), as distibutions were postively skewed,
suggesting the Blomberg's \(\kappa\) might not behave as a statistic
that follows a noraml distribution. However, because their simulations
used a weighted combination of simulated phylogenetic signal strengths,
strong inferences are not possible (and distributional attributes were
not the intended result of their simulations). For both Pagel's
\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
across a range of \(\lambda\) used to generate data would be valuable
for adjudicating the relibalilty of these statistics. Furthermore, these
are statistics that appear to have expected values that vary with tree
size (20), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores in the same
simulation experiments would allow evaluation of the efficacy of each
statistic to yield effect sizes that could be used for comparative
anlyses. \hfill\break

In this study, we used simulation experiments to compare the
distributional attributes of \(\hat{\lambda}\) and Blomberg's
\(\kappa\), plus their effect sizes (\(Z\)-scores), across a range of
tree size and phylogenetic signal strength. We present results based on
pure-birth trees, but results from blanced or pectinate trees were
qualitatively no different (see Supporting Information). Results from
trees with polytomies, or from linear regression or phylogenetic ANOVA,
were also qualitatively no different (see Supporting Information).

\hypertarget{results}{%
\section{Results}\label{results}}

Our simulations involved tree sizes of \(2^5\), \(2^6\), \ldots{},
\(2^{10}\) and \(\lambda\) of 0, 0.05, 0.10, \ldots{}, 1.0, with 100
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, we measured the mean values of
\(\hat{\lambda}\) and Blomberg's \(\kappa\), their standard deviation,
and a Shapiro-Wilk \(W\) statistic as a departure from normality
(symmetry), with a value of \(1.0\) indicating normally distributed
values. Departures from \(1.0\) indicate skewness. Kurtosis was not
measured directly but could be qualitatively evaluated from plots.
\hfill\break

In general, the following six results were clear (Fig. 1). First, for
\(\hat{\lambda}\), the distributional expectations of a Bernoulli
variable were mostly upheld. The mean value of \(\hat{\lambda}\)
increased as \(\lambda\) increased; the standard deviation of
\(\hat{\lambda}\) was largest at intermediate values of \(\lambda\) and
smallest at extreme values; and the distributions of \(\hat{\lambda}\)
were increasing skewed at extreme values of \(\lambda\) but rather
normal at intermediate values. (For small tree sizes, it was also clear
that distributions were more platykurtik at intermeidate values of
\(\hat{\lambda}\); see Fig. 1). However, second, \(\hat{\lambda}\) was a
negatively-biased estimate of \(\lambda\) for small tree sizes, as the
mean value was consistently less than the input \(\lambda\) value.
Third, standard deviations of \(\hat{\lambda}\) values were negatively
associated with tree size. In fact, for tress of 128 species or less,
\(\hat{\lambda}\) was quite variable, except for cases when \(\lambda\)
was near or equal to 1. Fourth, for \(\kappa\), mean values increased
consistently with \(\lambda\), irespective of tree size; standard
deviation of \(\kappa\) was consistent across tree sizes; and \(\kappa\)
was normally distributed regarless of \(\kappa\) or tree size, in
general (Fig. 2). Fifth, standard deviation of \(\kappa\) increased with
\(\lambda\), somewhat tracking the mean, but were always less than
corresponding means. This is perhaps unsurprising, as \(\kappa\) is
bound by 0, and was never large for small values of \(\lambda\).
Finally, sixth, the only evidence of skewing of \(\kappa\) distributions
was at small tree sizes and \(\lambda\) near or equal to 1, which
differs from the results of Münkemüller et al. (20), who used weighted
combinations of simulated data. The skewing observed in their simulation
results appears to be from combining random values generated
independently. \hfill\break

In terms of effect sizes, \(Z\)-scores were associated with both
phylogenetic signal strength (\(\lambda\)) and tree size, for both
statistics. Effect sizes from \(\hat{\lambda}\) made little sense.
Effect sizes were exceedingly associated with tree size, as much or more
so than phylogenetic signal strength (Fig. 3). By contrast, standard
deviates from \(\kappa\) were much more consistent accross tree sizes,
and exhibited only a slight positive association with tree size despite
a clear positive association with phylogenetic signal strength. Between
the two statistics, effect sizes from \(\kappa\) would be reliable for
cross-study comparisons but effect sizes from \(\hat{\lambda}\) would
require comparisons limited to phylogenies with similar numbers of
species.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

Real figures will have 6 panels for lambda and kappa and probably only a
subset for z. Simulations need to be re-run.

\includegraphics[width=0.95\linewidth]{new.fig.1.temp}

\singlespacing \textbf{Figure 1}. Temporary Fig. 1 showing lambda
patterns

\newpage

\includegraphics[width=0.95\linewidth]{new.fig.2.temp}

\singlespacing \textbf{Figure 2}. Temporary Fig. 2 showing kappa
patterns

\newpage

\includegraphics[width=0.95\linewidth]{new.fig.3.temp}

\singlespacing \textbf{Figure 3}. Temporary Fig. 3 showing z patterns

\newpage

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Felsenstein1985}{}%
1. J. Felsenstein, Phylogenies and the comparative method.
\emph{American Naturalist} \textbf{125}, 1--15 (1985).

\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
2. P. H. Harvey, M. D. Pagel, \emph{The comparative method in
evolutionary biology} (Oxford University Press, 1991).

\leavevmode\hypertarget{ref-Grafen1989}{}%
3. A. Grafen, The phylogenetic regression. \emph{Philosophical
Transactions of the Royal Society of London B, Biological Sciences}
\textbf{326}, 119--157 (1989).

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
4. T. J. Garland, A. R. Ives, Using the past to predict the present:
Confidence intervals for regression equations in phylogenetic
comparative methods. \emph{American Naturalist} \textbf{155}, 346--364
(2000).

\leavevmode\hypertarget{ref-Rohlf2001}{}%
5. F. J. Rohlf, Comparative methods for the analysis of continuous
variables: Geometric interpretations. \emph{Evolution} \textbf{55},
2143--2160 (2001).

\leavevmode\hypertarget{ref-ButlerKing2004}{}%
6. M. A. Butler, A. A. King, Phylogenetic comparative analysis: A
modeling approach for adaptive evolution. \emph{American Naturalist}
\textbf{164}, 683--695 (2004).

\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
7. E. P. Martins, T. F. Hansen, Phylogenies and the comparative method:
A general approach to incorporating phylogenetic information into the
analysis of interspecific data. \emph{American Naturalist} \textbf{149},
646--667 (1997).

\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
8. B. C. O'Meara, C. Ane, M. J. Sanderson, P. C. Wainwright, Testing for
different rates of continuous trait evolution using likelihood.
\emph{Evolution} \textbf{60}, 922--933 (2006).

\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
9. L. J. Revell, L. J. Harmon, Testing quantitative genetic hypotheses
about the evolutionary rate matrix for continuous characters.
\emph{Evolutionary Ecology Research} \textbf{10}, 311--331 (2008).

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
10. J. M. Beaulieu, D. C. Jhwueng, C. Boettiger, B. C. O'Meara, Modeling
stabilizing selection: Expanding the ornstein-uhlenbeck model of
adaptive evolution. \emph{Evolution} \textbf{66}, 2369--2383 (2012).

\leavevmode\hypertarget{ref-Adams2014b}{}%
11. D. C. Adams, A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution} \textbf{68}, 2675--2688 (2014).

\leavevmode\hypertarget{ref-Adams2014a}{}%
12. D. C. Adams, A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. \emph{Systematic Biology} \textbf{63}, 685--697 (2014).

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
13. D. C. Adams, M. L. Collyer, Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution} \textbf{72}, 1204--1215 (2018).

\leavevmode\hypertarget{ref-Pagel1999}{}%
14. M. D. Pagel, Inferring the historical patterns of biological
evolution. \emph{Nature} \textbf{401}, 877--884 (1999).

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
15. S. P. Blomberg, T. Garland, A. R. Ives, Testing for phylogenetic
signal in comparative data: Behavioral traits are more labile.
\emph{Evolution} \textbf{57}, 717--745 (2003).

\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
16. L. J. Revell, L. J. Harmon, D. C. Collar, Phylogenetic signal,
evolutionary process, and rate. \emph{Systematic Biology} \textbf{57},
591--601 (2008).

\leavevmode\hypertarget{ref-Abouheif1999}{}%
17. E. Abouheif, A method for testing the assumption of phylogenetic
independence in comparative data. \emph{Evolutionary Ecology Research}
\textbf{1}, 895--909 (1999).

\leavevmode\hypertarget{ref-Gittleman1990}{}%
18. J. L. Gittleman, M. Kot, Adaptation: Statistics and a null model for
estimating phylogenetic effects. \emph{Systematic Zoology} \textbf{39},
227--241 (1990).

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
19. C. P. Klingenberg, N. A. Gidaszewski, Testing and quantifying
phylogenetic signals and homoplasy in morphometric data.
\emph{Systematic biology} \textbf{59}, 245--261 (2010).

\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
20. T. Münkemüller\emph{et al.}, How to measure and test phylogenetic
signal. \emph{Methods in Ecology and Evolution} \textbf{3}, 743--756
(2012).

\leavevmode\hypertarget{ref-Pavoine2012}{}%
21. S. Pavoine, C. Ricotta, Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics.
\emph{Evolution: International Journal of Organic Evolution}
\textbf{67}, 828--840 (2012).

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
22. J. A. F. Diniz-Filho, T. Santos, T. F. Rangel, L. M. Bini, A
comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. \emph{Genetics and Molecular Biology}
\textbf{35}, 673--679 (2012).

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
23. R. Molina-Venegas, M. A. Rodríguez, Revisiting phylogenetic signal;
strong or negligible impacts of polytomies and branch length
information? \emph{BMC evolutionary biology} \textbf{17}, 53 (2017).

\leavevmode\hypertarget{ref-Revell2010}{}%
24. L. J. Revell, Phylogenetic signal and linear regression on species
data. \emph{Methods in Ecology and Evolution} \textbf{1}, 319--329
(2010).

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
25. R. P. Freckleton, P. H. Harvey, M. Pagel, Phylogenetic analysis and
comparative data: A test and review of evidence. \emph{American
Naturalist} \textbf{160}, 712--726 (2002).

\leavevmode\hypertarget{ref-Cooper2010}{}%
26. N. Cooper, W. Jetz, R. P. Freckleton, Phylogenetic comparative
approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology} \textbf{23}, 2529--2539 (2010).

\leavevmode\hypertarget{ref-Bose2019}{}%
27. R. Bose, B. R. Ramesh, R. Pélissier, F. Munoz, Phylogenetic
diversity in the western ghats biodiversity hotspot reflects
environmental filtering and past niche diversification of trees.
\emph{Journal of Biogeography} \textbf{46}, 145--157 (2019).

\leavevmode\hypertarget{ref-Vandelook2019}{}%
28. F. Vandelook\emph{et al.}, Nectar traits differ between pollination
syndromes in balsaminaceae. \emph{Annals of Botany} \textbf{124},
269--279 (2019).

\leavevmode\hypertarget{ref-DeMeester2019}{}%
29. G. De Meester, K. Huyghe, R. Van Damme, Brain size, ecology and
sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society} \textbf{126}, 381--391 (2019).

\leavevmode\hypertarget{ref-Pintanel2019}{}%
30. P. Pintanel, M. Tejedo, S. R. Ron, G. A. Llorente, A. Merino-Viteri,
Elevational and microclimatic drivers of thermal tolerance in andean
pristimantis frogs. \emph{Journal of Biogeography} \textbf{46},
1664--1675 (2019).

\leavevmode\hypertarget{ref-Su2019}{}%
31. G. Su, S. Villéger, S. Brosse, Morphological diversity of freshwater
fishes differs between realms, but morphologically extreme species are
widespread. \emph{Global ecology and biogeography} \textbf{28}, 211--221
(2019).

\end{document}

### 2020-07-12 Data_Analyses/Simulation_script-PB-KappaComparison_for_polytomies.R

@@ -1,0 +1,167 @@
+# Simulating pgls inference errors from estimated lambdas
+
+# Set up
+# setwd("/Users/ericabaken/Documents/School/Projects/Lambda_Eval/Manuscript/2020-Adams_and_Baken/")
+
+library(phytools)
+library(geomorph)
+library(geiger)
+library(caper)
+
+ # extract kappa values for comparison
+
+treesizes <-  2^(5:10) # 32,64,128,256,512,1024
+nsim <- 200 # number of simulations
+lambdas <- seq(0,1,.05) # starting with just 3 lambda values
+lambdas_expanded <- rep(lambdas, each = nsim)
+fullsimlength <- length(lambdas_expanded)
+
+n <- treesizes[2] # number of tips: adjust each iteration
+
+Perc <- 0 # Percentage of the nodes of the given “true” chronogram to be collapsed
+Total<- n - 2
+To_collapse <- round((Total*Perc)/100)
+
+gg <- function(tree, n){ # From Molina-Venegas and Rodríguez (2017)
+  if(n > 0) {
+    for(i in 1:n){
+      ii <- sample(2:tree$Nnode,1) + length(tree$tip.label)
+      ll <- tree$edge.length[which(tree$edge[,2] == ii)]
+      tree$edge.length[which(tree$edge[,1] == ii)]<-
+        tree$edge.length[which(tree$edge[,1] == ii)] + ll
+      tree$edge.length[which(tree$edge[,2] == ii)] <- 0
+      tree<-di2multi(tree)
+    }
+  }
+  tree
+}
+
+
+
+lambda.Z <- function(pgls.mod) {
+  FP <- pgls.mod$param
+  pars <- pgls.likelihood(optimPar = FP, fixedPar = pgls.mod$param, 
+                          y = pgls.mod$y, x = pgls.mod$x, V = pgls.mod$Vt, 
+                          optim.output = FALSE)
+  FP[2] <- 0
+  pars0 <- pgls.likelihood(optimPar = FP, fixedPar = FP, 
+                           y = pgls.mod$y, x = pgls.mod$x, V = pgls.mod$Vt, 
+                           optim.output = FALSE)
+  
+  Q0 <- -2*(pars0$ll - pars$ll)
+  Z <- sqrt(abs(Q0))
+
+  
+  Z
+  
+}
+
+# New Rep #####
[…]

### 2020-07-12 Data_Analyses/updated_simulations/simulation.analyses.R

@@ -2,11 +2,14 @@
 
 # general phy-signal analyses
 
-#### pure-birth tree analyses
+library(readxl)
+library(latex2exp)
 
-nsim <- 100
-treesizes <- 2^(5:10)
-labmdas <- seq(0, 1, 0.05)
+#### pure-birth tree analyses
+source("simulation.source.R")
+nsim <- 50
+treesizes <- 2^(5:8)
+lambdas <- seq(0, 1, 0.05)
 location <- getwd() # should be same as this file location
 destination <- paste(getwd(), "/results", sep = "")
 
@@ -14,10 +17,12 @@ pb.args <- list(treesize = treesizes[1], tree.type = "pb",
                 nsim = nsim, lambdas = lambdas, poly.per = 0,
                 beta = 1, mod.type = "mean")
 
+setwd(destination)
+
 pb.sets <- lapply(1:length(treesizes), function(j){
   pb.args$treesize = treesizes[j]
   R <- do.call(effects.simulations.PP, pb.args)
-
+  write.csv(R$Result, file = R$file.name, row.names = FALSE)
   cat(j, "tree of", length(treesizes), "trees, completed\n")
   R
 })
@@ -25,11 +30,6 @@ pb.sets <- lapply(1:length(treesizes), function(j){
 names(pb.sets) <- treesizes
 
 # save files
-setwd(destination)
-lapply(1:length(pb.sets), function(j){
-  R <- pb.sets[[j]]
-  write.csv(R$Result, file = R$file.name, row.names = FALSE)
-})
 
 save(pb.sets, file = "pb.sets.Rdata")
 setwd(location)
@@ -39,12 +39,105 @@ setwd(location)
 par(mfrow = c(2, length(treesizes)/2))
 for(i in 1:length(treesizes)) {
   R <- pb.sets[[i]]
-  lambda.plot(R, pch = 19, 
+  Rr <- R$Result
+  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Signal ($\\lambda_{est}$)"),
[…]

### 2020-07-12 Data_Analyses/updated_simulations/simulation.source.R

@@ -25,7 +25,8 @@ effects.simulations <- function(treesize , tree.type = c("bal", "pect", "pb"),
   
   n <- treesize
 
-  x <- if(mod.type == "mean" || type == "reg") matrix(1, n) else matrix(rep(c(0, 1), each  = n/2)) 
+  x <- if(mod.type == "mean" || type == "reg") matrix(1, n) else 
+    sample(matrix(rep(c(0, 1), each  = n/2)))
   
   stat.args <- list(tree = 0, x = x, beta = beta[1], lambda = lambdas[1],
                     type = mod.type)
@@ -218,19 +219,19 @@ kappa.plot <- function(R, ...) {
   
 }
 
-lambda.z.plot <- function(R, var.col = 2, var.lwd = 2, ...) {
+lambda.z.plot <- function(R, mean.col = 2, mean.lwd = 2, ...) {
   df <- R$Result
   plot(df$lambda.in, df$lambda.z, ...)
-  vars <- by(df$lambda.z, df$lambda.in, var)
-  points(unique(df$lambda.in), vars, type = "l", col = var.col, lwd = var.lwd)
+  means <- by(df$lambda.z, df$lambda.in, mean)
+  points(unique(df$lambda.in), means, type = "l", col = mean.col, lwd = mean.lwd)
   
 }
 
-kappa.z.plot <- function(R, var.col = 2, var.lwd = 2, ...) {
+kappa.z.plot <- function(R, mean.col = 2, mean.lwd = 2, ...) {
   df <- R$Result
   plot(df$lambda.in, df$kappa.z, ...)
-  vars <- by(df$kappa.z, df$lambda.in, var)
-  points(unique(df$lambda.in), vars, type = "l", col = var.col, lwd = var.lwd)
+  means <- by(df$kappa.z, df$lambda.in, mean)
+  points(unique(df$lambda.in), means, type = "l", col = mean.col, lwd = mean.lwd)
   
 }
 

### 2020-07-12 Manuscript/VersionDiff.r

@@ -4,5 +4,5 @@
 # NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
 
 library(latexdiffr)
-latexdiff("Manuscript-6-DCA.Rmd", "Manuscript-6-EKB.Rmd", clean = TRUE, output = "V6-V6EB.diff.pdf")
+latexdiff("Manuscript-5.Rmd", "Manuscript-6.Rmd", clean = TRUE, output = "V5-V6.diff.pdf")
 

### 2020-07-12 Manuscript/Manuscript-8.tex
typos

@@
 By contrast, with Blomberg's \(\kappa\), which is positively ubounded,
-we might expect that \(\hat{\lambda}\) follows a normal distribution
+we might expect that \(\kappa\) follows a normal distribution
 (\textbf{add ref}). A simulation experiment should reveal that

### 2020-07-15 Manuscript/OldVsns/Manuscript-1-DCA.tex
pnas vsn

--- 2020-04-21 Manuscript/Manuscript-1-DCA.tex
@@
 the ratio of observed trait variation to the trait variation conditioned
-on the phylogeny; scaled by what is expected under Brownian motion given
-the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
-Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
-likelihood to fit the data to the phylogeny under some model of
-evolutionary change (typically Brownian motion). The inclusion of a
-scaling parameter, \(\lambda\), transforms the lengths of the internal
-branches of the phylogeny to improve the fit, and this parameter
-describes the degree of phylogenetic signal in the dataset (Pagel 1999;
-Freckleton et al. 2002). Pagel's \(\lambda\) also has the appeal that it
-may be included when estimating the association of traits in a
-phylogenetic context, meaning that one may account for the degree of
-phylogenetic signal while conducting phylogenetic regression or ANOVA
+on the phylogeny; which is then scaled by what is expected under
+Brownian motion given the phylogeny's size and shape (Blomberg et al.
+2003; also Adams 2014a). Another approach, Pagel's \(\lambda\) (Pagel
+1999), uses maximum likelihood to fit the data to the phylogeny under
+some model of evolutionary change (typically Brownian motion). The
+inclusion of a scaling parameter, \(\lambda\), transforms the lengths of
+the internal branches of the phylogeny to improve the fit, and this
+parameter describes the degree of phylogenetic signal in the dataset
+(Pagel 1999; Freckleton et al. 2002). Pagel's \(\lambda\) also has the
+appeal that it may be included when estimating the association of traits
+in a phylogenetic context, meaning that one may account for the degree
+of phylogenetic signal while conducting phylogenetic regression or ANOVA
 (see Freckleton et al. 2002). \hfill\break
@@
 not directly comparable to the measures of phylogenetic signal being
-compared (see Munkemuller et al. 2012). One study (Boettiger et al.
-2012) found that estimates of Pagel's \(\lambda\) displayed less
-variation when data were simulated on a large phylogeny (\(N=281\)) as
-compared to a small one (\(N=13\)), concluding that insufficient data
-(i.e., the number of species) was the underlying cause of the increased
-variation across parameter estimates. However, this conclusion assumes
-that the precision in estimating \(\lambda\) remains constant across its
-range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
-verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
-macroevolutionary studies, at present, we still lack a general
-understanding of the precision with which \(\lambda\) can estimate
-levels of phylogenetic signal in phenotypic datasets. \hfill\break
+compared (see Munkemuller et al. 2012). Indeed, understanding the
+precision in estimating levels of phylogenetic signal is paramount,
+because biological inferences regarding the strength of phylogenetic
+signal in phenotypic datasets requires high precision of its estimation.
+One study (Boettiger et al. 2012) found that estimates of Pagel's
+\(\lambda\) displayed less variation when data were simulated on a large
+phylogeny (\(N=281\)) as compared to a small one (\(N=13\)), concluding
+that insufficient data (i.e., the number of species) was the underlying
+cause of the increased variation across parameter estimates. However,
+this conclusion assumes that the precision in estimating \(\lambda\)
+remains constant across its range (\(\lambda = 0 \to 1\)); an assumption
+that to date, has not been verified. Thus, despite widespread use of
+Pagel's (1999) \(\lambda\) in macroevolutionary studies, at present, we
+still lack a general understanding of the precision with which
+\(\lambda\) can estimate levels of phylogenetic signal in phenotypic
+datasets. \hfill\break
 
@@
 signal, with decreased precision when phylogenetic signal is of
-intermediate strength. In addition, the same \(\lambda_{est}\) may be
+intermediate strength. Additionally, the same \(\lambda_{est}\) may be
 obtained from datasets containing vastly different input levels of
@@
 that estimates of phylogenetic signal using Pagel's \(\lambda\) are
-often innacurate, and thus interpreting strength of phylogenetic signal
+often inaccurate, and thus interpreting strength of phylogenetic signal
 in phenotypic datasets based on this measure is compromised. By

### 2020-07-15 Manuscript/OldVsns/Manuscript-3-EKB.Rmd
pnas vsn

--- 2020-04-28 Manuscript/Manuscript-3.Rmd
@@
 Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
-Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
-Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.
+Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature; ~~.
+For instance,~~ of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% EXPLICITLY interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+~~Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.~~ FURTHERMORE, THIS PERCENTAGE IS EVEN HIGHER WHEN CONSIDERING THE IMPLICIT BIOLOGICAL INTERPRETATION OF PHYLOGENETIC SIGNAL FOR $\lambda$s NEAR 0 OR 1, AS IS THE CASE FOR NEARLY HALF OF THE 1,572 REPORTED $\lambda$ VALUES (FIGURE 1).
 \hfill\break
@@
 However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
-Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
+Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we ~~still~~ lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
 \hfill\break
@@
 For this to be the case, $\lambda$ must be a statistical effect size.
-Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
-Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analysis [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
-Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
-As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
+Effect sizes are a measure OF the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
+Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in ~~meta-analysis~~ META-ANALYSES [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
+~~Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
+As a consequence,~~ THUS, it follows that only ~~if $\lambda$ is a statistical effect size~~ BY CONVERTING $\lambda$ TO AN EFFECT SIZE can comparisons of estimates across datasets be interpretable.
 For the case of $\lambda$, this has not yet been explored.
@@
 Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data.
-We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
-Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
-We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
+We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$.
+Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than ~~that~~ those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure.
+We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets and provide an empirical example to demonstrate its use.
 We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
@@
 Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
-Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal corresponding reflecting Brownian motion (when $\lambda_{in} =1$).
-For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the precision of $\lambda$ using the variance ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness.
+Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal ~~corresponding~~ reflecting Brownian motion (when $\lambda_{in} =1$).
+For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the ~~precision~~ VARIANCE of $\lambda$ ~~using the variance~~ ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness AS A MEASURE OF ESTIMATE PRECISION.
 \hfill\break
@@
 The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
-Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness were then observed.
-\hfill\break
+WE THEN CALCULATED Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness.
+~~were then observed.~~ \hfill\break
 
@@
 Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
-This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
+This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these MID-RANGE empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
 \hfill\break
 
-Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression, the results of our simulations were quite similar.
+Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression AND ANOVA, the results of our simulations were quite similar.
 Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
-However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
+However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; FIGURE S1), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
 Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C,D).
+Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C~~,D~~).
 This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared with another.
-As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared ~~with~~ TO another.
+As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be ~~found~~ CALCULATED as:
 
@@
 One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation.
+As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and ORIGINAL scale of the ~~original~~ phenotypic variables under investigation.
 This means that conceivably, the strength of phylogenetic signal may be compared across datsets for phenotypic traits in different units and scale (e.g., behavioral and morphological traits), so long as they are continuous variables.
@@
 
-Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
+Patterns ~~worse~~ STILL BAD with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.
 

### 2020-07-15 Manuscript/OldVsns/Manuscript-3.tex
pnas vsn

--- 2020-04-28 Manuscript/Manuscript-3.tex
@@
 
-\textbf{Figure 4}. Variation in estimates of phylogenetic signal across
-input levels of phylogenetic signal. (A) Estimates of Pagel's
-\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
-(B) Estimates of \(Z_K\) for data simulated on phylogenies with 128 taxa
-(\(n=128\)), (C) Variance in the variation of \(\lambda_{est}\) across
-input levels of phylogenetic signal, estimated on phylogenies containing
-differing numbers of species. (D) Variance in the variation of \(Z_K\)
-across input levels of phylogenetic signal, estimated on phylogenies
-containing differing numbers of species.
+\textbf{Figure 4}. Variation in effect size estimates of phylogenetic
+signal across input levels of phylogenetic signal. (A) Estimates
+\(Z_\lambda\) for data simulated on phylogenies with 32 taxa (\(n=32\)),
+(B) Estimates of \(Z_K\) for data simulated on phylogenies with 32 taxa
+(\(n=32\)), (C) Coefficients of variation of precision estimates of
+\(Z_\lambda\) and \(Z_K\) across input levels of phylogenetic signal,
+estimated on phylogenies containing differing numbers of species.
+
+\textbf{Figure 5}. (A) Linear measures for relative body size, and
+regions of the body used to estimate surface area to volume (SA:V)
+ratios. (B) Permutation distributions of phylogenetic signal for SA:V
+and \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
+confidence intervals (CI not standardized by \(\sqrt(n)\)).
 
@@
 
-\singlespacing \textbf{Figure 4}. Variation in estimates of phylogenetic
-signal across input levels of phylogenetic signal. (A) Estimates of
-Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
-(\(n=128\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
-with 128 taxa (\(n=128\)), (C) Variance in the variation of
-\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
-on phylogenies containing differing numbers of species. (D) Variance in
-the variation of \(Z_K\) across input levels of phylogenetic signal,
-estimated on phylogenies containing differing numbers of species.
+\singlespacing \textbf{Figure 4}. Variation in effect size estimates of
+phylogenetic signal across input levels of phylogenetic signal. (A)
+Estimates \(Z_\lambda\) for data simulated on phylogenies with 32 taxa
+(\(n=32\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
+with 32 taxa (\(n=32\)), (C) Coefficients of variation of precision
+estimates of \(Z_\lambda\) and \(Z_K\) across input levels of
+phylogenetic signal, estimated on phylogenies containing differing
+numbers of species.
+
+\includegraphics[width=0.95\linewidth]{Fig5}
+
+\singlespacing \textbf{Figure 5}. (A) Linear measures for relative body
+size, and regions of the body used to estimate surface area to volume
+(SA:V) ratios. (B) Permutation distributions of phylogenetic signal for
+SA:V and \(\frac{BW}{SVL}\), with observed values shown as vertical
+bars. (C) Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with
+their 95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).
 

### 2020-07-15 Manuscript/OldVsns/Manuscript-4-EKB.Rmd
pnas vsn

--- 2020-04-30 Manuscript/Manuscript-4.Rmd
@@
 It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
-Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion.
-However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
+Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion, However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes ~~two important statistical properties that have not been fully explored.
 First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
-Therefore, understanding the precision in estimating $\lambda$ is paramount.
-One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
-From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
+Therefore, understanding the precision in estimating $\lambda$ is paramount.~~ THE IMPORTANT STATISTICAL PROPERTY OF PRECISION.
+One study BRIEFLY EXPLORED THIS CONCEPT [@Boettiger_et_al2012] AND found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
+From this observation, it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
 Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
 However, this conclusion also implies that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
-Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
+Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.~~ [EKB: I suggested the rewrite above because I feel like the two presumptions of precision and consistent precision across the range can be simplified.
+As always, feel free to ignore what you dislike.]   \hfill\break
 
@@
 For this to be the case, $\lambda$ must be a statistical effect size.
-Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
+Effect sizes are a measure OF the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
 Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994].
-As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
+As a consequence, it follows that only if $\lambda$ is CONVERTED TO a statistical effect size can comparisons of estimates across datasets be interpretable.
 For the case of $\lambda$, this has not yet been explored.
@@
 Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$).
-These findings were particularly unsettling when considered in light of our literature survey.
-Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
+These findings were particularly unsettling when considered in light of our literature survey [WE DON'T EXPLICITLY SAY THAT WE CONDUCTED A LITERATURE SURVEY BEFORE THIS POINT, JUST MENTION SOME RESULTS FROM IT IN THE INTRO...
+IS THAT OK OR COULD WE ADD SOME PHRASING TO THAT SECTION OF THE INTRO SO AT LEAST THE PHRASE 'LITERATURE SURVEY' APPEARS BEFORE HERE?].
+Over one quarter of the $\lambda$ estimates ~~obtained~~ PUBLISHED in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1).
 This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal).
@@
 
-To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
-Results are found in Figure 4 (additional results are found in the Supporting Information).
-Here two things are evident.
-First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B).
+To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$~~).
+Results are found in Figure 4 (~~, FIGURE 4, additional results ~~are found~~ in the Supporting Information).
+Here, two things are evident.
+First, estimates of $Z_K$ LINEARLY track the input phylogenetic signal ~~in a more linear fashion than do~~ WHEREAS estimates of $Z_\lambda$ DO NOT (Figure 4A,B).
 Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$.
@@
 
-Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared to another.
+Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait ~~as compared to~~ THAN another [EKB: changed wording to avoid repetition of word "compare"].
 As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:
@@
 
-To demonstrate the utility of $\hat{Z}_{12}$ we performed an analysis of the strength of phylogenetic signal in two phenotypic datasets from species of plethodontid salamander.
+To demonstrate the utility of $\hat{Z}_{12}$ we ~~performed an analysis of~~ QUANTIFIED AND COMPARED the strength of phylogenetic signal ~~in~~ OF two phenotypic ~~datasets~~ TRAITS WITH DIFFERENT UNITS ACROSS ~~from~~ species of plethodontid salamander.
 The data were part of a series of studies examining macroevolutionary trends in phenotypic diversification in this group [@Baken2019; @Baken2020].
-Our dataset contained surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) for 305 species (Figure 5A).
+Our dataset contained surface area to volume ratios (SA:V; $mm^{-1}$) and relative body width ($\frac{BW}{SVL}$; NO UNITS) for 305 species (Figure 5A).
 For SA:V, 11 linear body measurements from 2,781 individuals were taken, from which estimates of the surface area and volume of the head, body, and tail were calculated and subsequently combined to arrive at the SA:V for each indiviual [for mathematical details see @Baken2020].
@@
 First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomite changes in the values of $Z_K$.
+Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by ~~concomite~~ CONCOMITANT changes in the values of $Z_K$.
 Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
@@
 Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation.
-This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$.
+This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale because those units have been standardized through their conversion to $Z_K$.
 For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
 Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
-Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.
+Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, ~~and~~ even for phenotypes obtained from species in different lineages because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.
 \hfill\break

### 2020-07-15 Manuscript/OldVsns/Manuscript-6.Rmd
pnas vsn

--- 2020-05-06 Manuscript/Manuscript-5.Rmd
@@
 Tests based on $Z_K$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
-Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal various phenotypic traits, even when those traits are found in different evolutionary lineages or of in different units or scale.
+Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scale.
 
@@
 # Introduction
-Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species generates statistical non-independence [@Felsenstein1985; @HarveyPagel1991].
-Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs); a suite of analytical tools that condition the data on the phylogeny through the course of statisical evaluations of phenotypic trends [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004].
+Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates an assumption of independence among trait valuesthat is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
+Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [e.g., @Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004].
 The past several decades have witnessed a rapid expansion in the development of PCMs to address an ever-growing set of macroevolutionary hypotheses [@MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @Adams2014a; @AdamsCollyer2018b].
@@
 One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999].
-Here, maximum likelihood is used to fit the data to the phylogeny under a Brownian motion model of evolution.
-A parameter ($\lambda$) is included, which transforms the lengths of the internal branches of the phylogeny to improve the fit [@Pagel1999; @Freckleton_et_al2002].
+The ($\lambda$) parameter stransforms the lengths of the internal branches of the phylogeny to improve a fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
 Pagel's $\lambda$ ranges from $0\to1$, with larger values signifying a greater dependence of observed trait variation on the phylogeny.
-Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic regression  (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
+Pagel's $\lambda$ also has the appeal that it may be included in phylogenetic generalized least-squares regression (PGLS) to account for the degree of phylogenetic signal in comparative analyses [see @Freckleton_et_al2002].
 \hfill\break
 
-Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
-Other approaches for interpreting $\lambda$ are more statistical.
-For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
-Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic to describe the relative strength of phylogentic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
+The appeal of $\lambda$ as a descriptive statistic for evolutionary biologists is a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\lambda$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
@@
 [insert Figure 1 here] \hfill\break
+
+Various other approaches use $\lambda$ as a parameter that can be varied for inferences akin to sensitivty analysis.
+For instance, some have performed likelihood ratio tests that compare observed model fits to those obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or evaluated whether observed $\lambda$ differs from an expected $\lambda$, based on confidence intervals generated for the expected value [@Vandelook2019].
+Qualitative comparisons of $\lambda$ estimates have also been performed for multiple traits on the same phylogenetic tree to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
 
@@
 
+***note from Mike***  Throughout the ms thus far $Kappa$ is used.  Why not $\kappa$, the actual greek symbol for $Kappa$?
+
 # Methods and Results
@@
 Here two things are evident.
-First, estimates of $Z_K$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A,B).
+First, estimates of $Z_K$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A, B).
 Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$.
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Here coefficients of variation in the precision of $Z_K$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$.
+Coefficients of variation in the precision of $Z_K$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$.
 \hfill\break
@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} = \frac{\lvert Z_{K_1} - Z_{K_2}\rvert}{\sqrt{2}}
 \end{align}
@@
 where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2.
+The right side of the equation illustrates that if $Z_K$ has already been calculated for two sampling distributions as in equation 2, the sampling distributions have unit variance for each of the $Z_K$ statistics.
 Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
@@
 We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
-First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal.
+First, $Z_K$ was more precise than $Z_{\lambda}$, and precision was more consistent across the range of input levels of phylogenetic signal.
 Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_K$.
@@
 
-Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another.
+Based on the effect size $Z_K$, we then proposed a two-sample test, which provides means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another, via a hypothesis test.
 Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003].
 However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
-Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale.
-By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
-Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons.
+Thus, $Kappa$ should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
+By contrast, standardizing $Kappa$ ($Z_K$, via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
 As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
@@
 One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation -- regardless of the initial units and original scale of the phenotypic variables under investigation.
 This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$.
@@
 
-**Finally... Need Closing paragraph.**
+Phylogenetic signal can be thought of as both an attribute to be measured in the data and a paramter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
+As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
+However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a useful  statistic to describe the amount of phylogenetic signal in the data.
+Interestingly, $Kappa$ is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
+Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
+By contrast, $Kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution.
+Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
+With low coefficients of variation for $Z_K$ (Figure 4), it is difficult to imagine that a hypothesis test can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
 

### 2020-07-15 Manuscript/OldVsns/Manuscript-7-MLC.Rmd
pnas vsn

--- 2020-05-11 Manuscript/Manuscript-6-EKB.Rmd
@@
 One of the most widely used methods for characterizing phylogenetic signal in macroevolutionary studies is Pagel's $\lambda$ [@Pagel1999].
-The ~~($\lambda$)~~ parameter ($\lambda$) transforms the lengths of the internal branches of the phylogeny to improve ~~a~~ THE fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
+The parameter ($\lambda$) transforms the lengths of the internal branches of the phylogeny to improve the fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
 Pagel's $\lambda$ ranges from $0\to1$, with larger values signifying a greater dependence of observed trait variation on the phylogeny.
@@
 
-In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic ~~to describe~~ CHARACTERIZING the relative strength of phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-~~The appeal~~ THIS DESCRIPTIVE UTILITY of $\lambda$ ~~as a descriptive statistic for evolutionary biologists~~ is ~~a~~ THE basis for interpreting "weak" versus "strong" phylogenetic signal~~,~~ (i.e., small versus large values of $\lambda$, respectively)~~,~~ in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
-Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% EXPLICITLY interpreted the strength of phylogenetic signal for at least one phenotypic trait.
-Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
+In PGLS analysis, $\lambda$ is a parameter that is tuned via log-likelihood profiling for its estimation.
+After obtaining an optimized $\lambda$ value, $\hat{\lambda}$, likelihood ratio tests can be performed to compare observed model fits using $\hat{\lambda}$ to those obtained when $\lambda=0$ or $\lambda=1$ to infer whether phylogenetic signal differs from no signal or a Brownian motion (BM) model of evolutionary divergence, respectively [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
+Confidence limits on $\hat{\lambda}$ can also be estimated based on percentiles of a $\chi^2$ distribution and where these values are mapped onto a log-likelihood profile [*sensu* @Freckleton_et_al2002].
+With respect to confidence limits, one can ascertain whether $\lambda=0$ or $\lambda=1$ is contained within an interval of $\hat{\lambda}$, as an analog to likelihood ratio tests [see, e.g., @Vandelook2019].
+It is, therefore, tempting to regard $\hat{\lambda}$ as a descriptive statistic that measures the relative strength of phylogenetic signal, for inferring the extent to which shared evolutionary history has influenced trait covariation among taxa.
+The appeal of $\hat{\lambda}$ as a descriptive statistic is that it serves as the basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
+For example, qualitative comparisons of $\hat{\lambda}$ for multiple traits on the same phylogenetic tree provide a basis for inferring whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+Indeed, statements regarding the strength of phylogenetic signal based on $\hat{\lambda}$ are rather common in the evolutionary literature.
+For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% explicitly interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+Further, because nearly half of the 1,572 $\hat{\lambda}$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under BM), this percentage is even higher.
 \hfill\break
@@
 
-Various other approaches use $\lambda$ as a parameter that can be ~~varied~~ ADJUSTED for inferences akin to sensitivity analysis.
-For instance, some have performed likelihood ratio tests that compare observed model fits to those obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or evaluated whether observed $\lambda$ differs from an expected $\lambda$, based on confidence intervals generated for the expected value [@Vandelook2019].
-Qualitative comparisons of $\lambda$ estimates have also been performed for multiple traits on the same phylogenetic tree to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+# As a statistic that measures phylogenetic signal, how should $\hat{\lambda}$ behave?
+
+It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\hat{\lambda}$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
+However, equating values of $\hat{\lambda}$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
+First, it presumes that values of $\hat{\lambda}$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
+(In order to be precise, the log-likelihood profile would have to be peaked rather than flat.)  Therefore, understanding the precision in estimating $\hat{\lambda}$ is paramount.
+As a prortional random variable bounded by 0 and 1, we might expect that $\hat{\lambda}$ follows A Bernoulli distribution; i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
+As such, we would also expect in simulation experiment that given a known $\lambda$ value used to generate random data on a tree: the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the standard error of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, and predictable over the range of $\lambda$ with respect to treesize; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1.
+These properties seemed to be superficially reasonable from the simulations perfomed by Münkemüller et al.[-@Munkemuller_et_al2012, see Fig.
+2].
+Their study simulated strength of phylogenetic signal as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$, rather than from prescribed values of $\lambda$.
+Nonetheless, their results revealed that as a random variable, $\hat{\lambda}$ resembled a Bernoulli random variable based superficially on statistical moments (mean, variance, skewness, and kurtosis; see their Fig.
+2) with respect to strength of phylogenetic signal, for a given tree size.
+However, for comparisons among tree sizes for a given strength of phylogenetic signal, median values of $\hat{\lambda}$ were slightly, positively associated with the number of species in a tree, especially at weaker strength of signals, suggesting $\hat{\lambda}$ might not be an accurate statistic across a range of tree sizes.
 \hfill\break
 
-It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
-Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion.
-However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
-First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
-Therefore, understanding the precision in estimating $\lambda$ is paramount.
-One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
-From this observation it was concluded that insufficient data (i.e., ~~the~~ LOW number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
+Another study [@Boettiger_et_al2012] found that $\hat{\lambda}$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
+From this observation it was concluded that insufficient data (i.e., low number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
 Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
-However, this conclusion also implies that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified.
-Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.
+However, this conclusion could imply that the shape of log-likelihood profiles are assumed consistent with species number.
+Such an assumption and the importance of tree shape have to date not been verified.
+Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\hat{\lambda}$ can estimate levels of phylogenetic signal in phenotypic datasets.
 \hfill\break
 
-Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data.
-For this to be the case, $\lambda$ must be a statistical effect size.
+Second, while $\lambda$ varies over a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values, $\hat{\lambda}$, of this parameter correspond to the actual strength of the underlying input signal in the data.
+For this to be the case, $\hat{\lambda}$ must be a statistical effect size.
 Effect sizes are a measure of the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988].
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
-Unfortunatley, not all model parameters and descriptive statistics are effect sizes, and thus many summary measures must first be converted to statistics with standardized units (i.e., conversion to an effect size) for meaningful comparison [see @Rosenthal1994].
-As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
-~~For the case of $\lambda$, this~~ HOWEVER, THE CALCULATION AND STATISTICAL BEHAVIOR OF $\lambda$ AS AN EFFECT SIZE has not yet been explored.
+Unfortunately, not all model parameters and descriptive statistics are effect sizes, directly, and thus many summary measures must first be converted to statistics with standardized units (i.e., conversion to an effect size) for meaningful comparison [see @Rosenthal1994].
+As a consequence, it follows that only if $\hat{\lambda}$ can be cobverted to a statistical effect size can comparisons of estimates across datasets be interpretable.
+However, the calculation and statistical behavior of $\hat{\lambda}$ as an effect size has not yet been explored.
 \hfill\break
 
-In this study, we evaluate the precision of Pagel's $\lambda$ for estimating known levels of phylogenetic signal in phenotypic data.
-We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances.
-We find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across its range.
+In this study, we evaluate the precision of $\hat{\lambda}$ as an optimized parameter, used as a statistic, for estimating known levels of phylogenetic signal in phenotypic data (data simulated with known $\lambda$).
+We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of $\lambda$, to explore the degree to which $\hat{\lambda}$ correctly identifies known levels of phylogenetic signal, $\lambda$, and under which circumstances.
+We find that estimates of $\hat{\lambda}$ vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range.
 Rather, there is decreased precision when input levels of phylogenetic signal are of intermediate strength.
-Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal.
-Thus, $\lambda$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data.
-We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $\kappa$.
-Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $\kappa$ ($Z_\kappa$), implying that $Z_\kappa$ is a more robust effect size measure.
+Additionally, the same values of $\hat{\lambda}$ may be obtained from datasets containing vastly different input levels of $\lambda$.
+Thus, $\hat{\lambda}$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data.
+We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets and apply the concept to two common measures of phylogenetic signal: $\hat{\lambda}$ and $\kappa$.
+Through simulations we find that the precision of effect sizes based on $\hat{\lambda}$ ($Z_{\lambda}$) are less reliable than that those based on $\kappa$ ($Z_\kappa$), implying that $Z_\kappa$ is a more robust effect size measure.
 We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use.
-We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
+We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ (as a descriptive statistic, not as a PGLS parameter) are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised.
 By contrast, effect sizes obtained from $\kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.
@@
 
-## *The Precision of $\lambda$ is Variable*
+## *The accuracy and precision of $\hat{\lambda}$ is Variable*
 
-We conducted a series of computer simulations to evaluate the precision of Pagel's $\lambda$.
+We conducted a series of computer simulations to evaluate the precision of $\hat{\lambda}$ based on varied input of Pagel's $\lambda$ to generate data.
 Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information).
 First we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$).
-Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
-Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal reflecting Brownian motion (when $\lambda_{in} =1$).
-For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the variance of $\lambda$ ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness as an estimate of precision.
-We verified that the variance of traits simulated had no effect on phylogenetic signal estimation.
+For each simulation run, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$).
+Continuous traits were then simulated on each phylogeny under a BM model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal reflecting a BM model of evolution (when $\lambda_{in} =1$).
+For each dataset we then estimated phylogenetic signal ($\hat{\lambda}$), and calculated the standard deviation of ${\hat{\lambda}$ ($\sigma_{\hat{\lambda}}$) at each input level of phylogenetic signal and level of species richness as an estimate of precision.
+Random data samples were generated from a normal distribution, i.e.,  $\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$.
+We verified that the level of $\sigma^2$ had no effect on $\hat{\lambda}$ or $\sigma^2_{\hat{\lambda}}$, but the results we present used $\sigma^2=1$.
 \hfill\break
 
-We also evaluated the precision of $\lambda$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$).
-Here, an independent variable $X$ was simulated on each rescaled phylogeny under a Brownian motion model of evolution (for PGLS regression).
-For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary, 0 or 1) character on each phylogeny.
-Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal.
-This was accomplished as: $Y=\beta{X}+\epsilon$.
-The association between $Y$ and $X$ was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error ($\epsilon$) was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution on each rescaled phylogeny: $\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$: [see @Revell2010 for a similar simulation design].
-The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained.
-We then calculated precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness.
-We verified that the amount of residual variance simulated had no effect on $\sigma^2_\lambda$ but did influence the precision of coefficients estimated from the linear model (precision increased with smaller $\epsilon$, as expected).
+We also evaluated the precision of $\hat{\lambda}$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$).
+For these simulations, $Y=\beta{X}+\epsilon$ was used to generate data with varied phylogenetic signal and variable associations [*sensu* @Revell2010].
+In the case of PGLS regression, $X$ was a variable simulated under a BM model of evolution, constrained by rescaling phylogeny according to $\lambda$, and in the case of ANOVA, $X$ was a balanced binary variable (values of 0 and 1), randomly ordered.
+The strength of variable associations ranged as $\beta=0.0,0.25, 0.5, 0.75$ and $1.0$.
+Residuals, $\epsilon$, were simulated under a BM model of evolution, constrained by rescaling phylogeny according to $\lambda$.
+Simulated independent variables or residuals were generated from a normal distribution, i.e.,  $\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$, with the phylogenetic covariance matrix, $\mathbf{C}$, implicitly varied for each tree by the level of $\lambda$.
+The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\hat{\lambda}$ and $\hat{\beta}$) were obtained.
+We then calculated precision estimates ($\sigma^2_{\hat{\lambda}}$) at each input level of phylogenetic signal and level of species richness.
+We verified that the level of $\sigma^2$ had no effect on $\sigma^2_{\hat{\lambda}}$ but did influence the precision of $\hat{\beta}$ estimated from the linear model (precision increased with smaller $\epsilon$, as expected).
+The results we present used $\sigma^2=1$.
 \hfill\break
@@
 
-*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions.
+*Results.*
+
+superifical Bernoulli traits
+innacuracy
+precision
+
+We found that the precision of $\lambda_{est}$ varied widely across simulation conditions.
 Predictably, precision improved as the number of species increased (Figure 2).
@@
 
-Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression and ANOVA, the results of our simulations were quite similar.
+When phylogenetic signal was investigated on balanced or pectinate trees, patterns in the precision of $\lambda$ were largely the same, with decreased precision at intermediate levels of phylogenetic signal (Supporting Information).
+Likewise, when $\lambda$ was co-estimated with regression parameters in PGLS regression and ANOVA, the results of our simulations were quite similar.
 Regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information).
 However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; see also Supporting Information), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2.
-Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize ~~observed~~ KNOWN levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
+Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize known levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate.
 \hfill\break
@@
 
-An effect size based on $\lambda$ could be envisioned, which is found as:
+An effect size based on $\lambda$ is a little more challenging because, as a statistic, it does not have a sampling distribution on which to measure standard error.
+Confidence intervals for $\lambda$ are not generated from standard error calculations, but rather for the values of lambda that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the compare the putative appropriate model to a null model with $\lambda = 0$ [citation needed, think Boettiger has a paper on this].
+The simplest similar effect size is
 
 \begin{align}
-    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
+   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
 \end{align}
+
+where, $\hat{\lambda}$ is the maximized likelihood value of $\lambda$ and $\chi^{2}_{\hat{\lambda}}$ is the likelihood ratio statistic for the value.
+Hypothetically, this equation would be better represented as $Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}$, where $d$ is a binary value $(-1,1)$ to indicate a direction based on whether $\hat{\lambda}$ is below or above a critical value of $\lambda$, for a quantile from a $\chi^{2}$ distribution at a probability of 0.5.
+The critical value would correspond to a value of $Z = 0$; i.e., the strength of phylogenetic signal that might be observed at the 50th percentile in a distribution of random outcomes, generated from a  model where $\lambda = 0$.
+However, in some preliminary simulations, we found that mapping $\chi^2_{\hat{\lambda}}$ values this way did not produce effect sizes that were symmetrical about $Z = 0$, as the mapping is not linear and, especially for small trees, the log-likelihood profiles can be rather flat, making estimation of $\hat{\lambda}$ rather imprecise (as we elaborate more below.)  We will heneforth use equation 3, recognizing that large values of $Z$ would likely be positive, anyway, based on the reference model for the likelihood ratio test.
 
@@
 This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_\kappa}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.
-Coefficients of variation in the precision of $Z_\kappa$ were up to an order of magnitude smaller ~~for~~ than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_\kappa$.
+Coefficients of variation in the precision of $Z_\kappa$ were up to an order of magnitude smaller than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_\kappa$.
 \hfill\break
@@
 Likewise, body size (SVL) and body width (BW) measurements were taken from 3,371 individuals, and species means of relative body width ($\frac{BW}{SVL}$) were calculated [data from @Baken2019].
-A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the PHENOTYPIC dataset.
-~~, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species.~~ <!-- number of species already mentioned earlier in this paragraph--> The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
+A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the phenotypic dataset.
+The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph].
 Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**).
@@
 First, $Z_\kappa$ was more precise than $Z_\lambda$, and precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_\kappa$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_\kappa$ <!-- The phrasing in this sentence is a little rough.
-Perhaps: "Additionally, values of $Z_\kappa$ more accurately tracked changes in magnitude of known phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa and $\lambda_{in}$." -->.
+Additionally, values of $Z_\kappa$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa$ and $\lambda_{in}$.
 Thus, $Z_\kappa$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
@@
 One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation <!--deviations?--> -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
 This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
@@
 In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
-Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a].
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
 Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
@@
 We repeated the analyses in Figure 4, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (see Supporting Information).
-This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size~~, and~~.
-FURTHERMORE, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, ~~if the sampling distributions also~~ WHICH WOULD shift EQUALLY <!-- not sure 'equally' is the right word here.
-Alternatives: "similarly" "concomitantly" "correspondingly" --> because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size.
+Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
 \hfill\break
 
-Phylogenetic signal can be thought of as both an attribute to be measured in the data and a paramter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
+Phylogenetic signal can be thought of as both an attribute to be measured in the data and a parameter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
 As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
-However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a useful  statistic to describe the amount of phylogenetic signal in the data.
+However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a statistic to describe the amount of phylogenetic signal in the data.
 Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.

### 2020-07-15 Manuscript/OldVsns/Manuscript-7-MLC.tex
pnas vsn

--- 2020-07-12 Manuscript/Manuscript-7-MLC.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[]{article}
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{setspace}\doublespacing
@@
 Cooper et al. 2010; Bose et al. 2019). Confidence limits on
-\(\hat{\lambda}\) can be also estimated based on percentiles of a
+\(\hat{\lambda}\) can also be estimated based on percentiles of a
 \(\chi^2\) distribution and where these values are mapped onto a
@@
 for inferring the extent to which shared evolutionary history has
-influenced trait covariation among taxa, is common. The appeal of
-\(\hat{\lambda}\) as a descriptive statistic is that it serves as the
-basis for interpreting ``weak'' versus ``strong'' phylogenetic signal;
-i.e., small versus large values of \(\hat{\lambda}\), respectively, in a
-comparative sense (e.g., De Meester et al. 2019; Pintanel et al. 2019;
-Su et al. 2019). For example, qualitative comparisons of
-\(\hat{\lambda}\) for multiple traits on the same phylogenetic tree
-provide a basis for inferring whether the strength of phylogenetic
-signal is greater in one trait as compared to another (e.g., Liu et al.
-2019; Bai et al. 2019). Indeed, statements regarding the strength of
-phylogenetic signal based on \(\hat{\lambda}\) are rather common in the
-evolutionary literature. For instance, of the 204 papers published in
-2019 that estimated and reported Pagel's \(\lambda\) (found from a
-literature survey we conducted in Google.scholar), 40\% explicitly
-interpreted the strength of phylogenetic signal for at least one
-phenotypic trait. Further, because nearly half of the 1,572
-\(\hat{\lambda}\) values reported were near 0 or 1 (Figure 1) where the
-biological interpretation of \(\lambda\) is known (\(\lambda=0\)
-represents no phylogenetic signal, while \(\lambda=1\) is phylogenetic
-signal as expected under BM), this percentage is even higher.
-\hfill\break
+influenced trait covariation among taxa. The appeal of \(\hat{\lambda}\)
+as a descriptive statistic is that it serves as the basis for
+interpreting ``weak'' versus ``strong'' phylogenetic signal; i.e., small
+versus large values of \(\hat{\lambda}\), respectively, in a comparative
+sense (e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al.
+2019). For example, qualitative comparisons of \(\hat{\lambda}\) for
+multiple traits on the same phylogenetic tree provide a basis for
+inferring whether the strength of phylogenetic signal is greater in one
+trait as compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
+Indeed, statements regarding the strength of phylogenetic signal based
+on \(\hat{\lambda}\) are rather common in the evolutionary literature.
+For instance, of the 204 papers published in 2019 that estimated and
+reported Pagel's \(\lambda\) (found from a literature survey we
+conducted in Google.scholar), 40\% explicitly interpreted the strength
+of phylogenetic signal for at least one phenotypic trait. Further,
+because nearly half of the 1,572 \(\hat{\lambda}\) values reported were
+near 0 or 1 (Figure 1) where the biological interpretation of
+\(\lambda\) is known (\(\lambda=0\) represents no phylogenetic signal,
+while \(\lambda=1\) is phylogenetic signal as expected under BM), this
+percentage is even higher. \hfill\break
 
 {[}insert Figure 1 here{]} \hfill\break
+
+\hypertarget{as-a-statistic-that-measures-phylogenetic-signal-how-should-hatlambda-behave}{%
+\section{\texorpdfstring{As a statistic that measures phylogenetic
+signal, how should \(\hat{\lambda}\)
+behave?}{As a statistic that measures phylogenetic signal, how should \textbackslash hat\{\textbackslash lambda\} behave?}}\label{as-a-statistic-that-measures-phylogenetic-signal-how-should-hatlambda-behave}}
 
@@
 Therefore, understanding the precision in estimating \(\hat{\lambda}\)
-is paramount. One study (Boettiger et al. 2012) found that estimates of
-Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
-when data were simulated on a large phylogeny (\(N=281\)) as compared to
-a small one (\(N=13\)). From this observation it was concluded that
-insufficient data (i.e., low number of species) was the underlying cause
-of the increased variation across parameter estimates (Boettiger et al.
-2012). Indeed, such a pattern is common with statistical estimators, as
-summary statistics and parameters are often more precise at greater
-sample sizes (Cohen 1988). However, this conclusion could imply that the
-shape of log-likelihood profiles are assumed consistent with species
-number. Such an assumption and the importance of tree shape have to date
-not been verified. Thus, despite widespread use of Pagel's (1999)
-\(\lambda\) in macroevolutionary studies, at present, we lack a general
-understanding of the precision with which \(\hat{\lambda}\) can estimate
-levels of phylogenetic signal in phenotypic datasets. \hfill\break
-
-Second, while estimates of \(\lambda\) are within a bounded scale
-(\(0\to1\)), this does not \emph{de-facto} imply that the estimated
-values, \(\hat{\lambda}\), of this parameter correspond to the actual
-strength of the underlying input signal in the data. For this to be the
-case, \(\hat{\lambda}\) must be a statistical effect size. Effect sizes
-are a measure of the magnitude of a statistical effect in data,
-represented on a common scale (Glass 1976; Cohen 1988). Effect sizes
-have widespread use in many areas of the quantiative sciences, as they
-represent measures that may be readily summarized across datasets as in
-meta-analyses (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster
-1995), or compared among datasets (e.g., Adams and Collyer 2016, 2019a).
-Unfortunately, not all model parameters and descriptive statistics are
-effect sizes, directly, and thus many summary measures must first be
-converted to statistics with standardized units (i.e., conversion to an
-effect size) for meaningful comparison (see Rosenthal 1994). As a
-consequence, it follows that only if \(\hat{\lambda}\) is a statistical
-effect size can comparisons of estimates across datasets be
-interpretable. However, the calculation and statistical behavior of
-\(\hat{\lambda}\) as an effect size has not yet been explored.
-\hfill\break
-
-As a prortional random variable bounded by 0 and 1, we might expect that
-\(\hat{\lambda}\) follows A Bernoulli distribution; i.e., branch lengths
-in a tree are scaled proportionally to the probability that data arise
-from a BM process. As such, we would also expect in simulation
-experiment that given a known \(\lambda\) value used to generate random
-data on a tree: the mean of an empirical sampling distribution of
-\(\hat{\lambda}\) would approximately equal \(\lambda\); the standard
-error of \(\hat{\lambda}\) would be largest at intermediate values of
-\(\lambda\), and predictable over the range of \(\lambda\) with respect
-to treesize; the distribution of \(\hat{\lambda}\) would be symmetric at
-intermediate values of \(\lambda\) and more skewed toward values of 0 or
-1; and that the distribution of \(\hat{\lambda}\) will be more
-platykurtic at intermediate values of \(\lambda\), becoming more
-leptokurtic toward 0 and 1. These properties seemed to be
-superficiallyreasonable from the simulations perfomed by Münkemüller et
-al.(2012, see Fig. 2). Their study simulated strength of phylogenetic
-signal as a varied weighted-average of data simulated on trees with
-\(\lambda=0\) and \(\lambda=1\), rather than from prescribed values of
-\(\lambda\). Nonetheless, their results revealed that as a random
-variable, \(\hat{\lambda}\) resembled a Bernoulli random variable based
+is paramount. As a prortional random variable bounded by 0 and 1, we
+might expect that \(\hat{\lambda}\) follows A Bernoulli distribution;
+i.e., branch lengths in a tree are scaled proportionally to the
+probability that data arise from a BM process. As such, we would also
+expect in simulation experiment that given a known \(\lambda\) value
+used to generate random data on a tree: the mean of an empirical
+sampling distribution of \(\hat{\lambda}\) would approximately equal
+\(\lambda\); the standard error of \(\hat{\lambda}\) would be largest at
+intermediate values of \(\lambda\), and predictable over the range of
+\(\lambda\) with respect to treesize; the distribution of
+\(\hat{\lambda}\) would be symmetric at intermediate values of
+\(\lambda\) and more skewed toward values of 0 or 1; and that the
+distribution of \(\hat{\lambda}\) will be more platykurtic at
+intermediate values of \(\lambda\), becoming more leptokurtic toward 0
+and 1. These properties seemed to be superficially reasonable from the
+simulations perfomed by Münkemüller et al.(2012, see Fig. 2). Their
+study simulated strength of phylogenetic signal as a varied
+weighted-average of data simulated on trees with \(\lambda=0\) and
+\(\lambda=1\), rather than from prescribed values of \(\lambda\).
+Nonetheless, their results revealed that as a random variable,
+\(\hat{\lambda}\) resembled a Bernoulli random variable based
 superficially on statistical moments (mean, variance, skewness, and
@@
 suggesting \(\hat{\lambda}\) might not be an accurate statistic across a
-range of tree sizes. In order for \(\hat{\lambda}\) to be considered a
-reliable statistic as a measure of phylogenetic signal, as a minimum, it
-should be accurate and precise, properties that could be measured from
-simulation experiments for known values of \(\lambda\).
+range of tree sizes. \hfill\break
+
+Another study (Boettiger et al. 2012) found that \(\hat{\lambda}\)
+displayed less variation (i.e., greater precision) when data were
+simulated on a large phylogeny (\(N=281\)) as compared to a small one
+(\(N=13\)). From this observation it was concluded that insufficient
+data (i.e., low number of species) was the underlying cause of the
+increased variation across parameter estimates (Boettiger et al. 2012).
+Indeed, such a pattern is common with statistical estimators, as summary
+statistics and parameters are often more precise at greater sample sizes
+(Cohen 1988). However, this conclusion could imply that the shape of
+log-likelihood profiles are assumed consistent with species number. Such
+an assumption and the importance of tree shape have to date not been
+verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
+macroevolutionary studies, at present, we lack a general understanding
+of the precision with which \(\hat{\lambda}\) can estimate levels of
+phylogenetic signal in phenotypic datasets. \hfill\break
+
+Second, while \(\lambda\) varies over a bounded scale (\(0\to1\)), this
+does not \emph{de-facto} imply that the estimated values,
+\(\hat{\lambda}\), of this parameter correspond to the actual strength
+of the underlying input signal in the data. For this to be the case,
+\(\hat{\lambda}\) must be a statistical effect size. Effect sizes are a
+measure of the magnitude of a statistical effect in data, represented on
+a common scale (Glass 1976; Cohen 1988). Effect sizes have widespread
+use in many areas of the quantiative sciences, as they represent
+measures that may be readily summarized across datasets as in
+meta-analyses (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster
+1995), or compared among datasets (e.g., Adams and Collyer 2016, 2019a).
+Unfortunately, not all model parameters and descriptive statistics are
+effect sizes, directly, and thus many summary measures must first be
+converted to statistics with standardized units (i.e., conversion to an
+effect size) for meaningful comparison (see Rosenthal 1994). As a
+consequence, it follows that only if \(\hat{\lambda}\) can be cobverted
+to a statistical effect size can comparisons of estimates across
+datasets be interpretable. However, the calculation and statistical
+behavior of \(\hat{\lambda}\) as an effect size has not yet been
+explored. \hfill\break
 
@@
 
-\hypertarget{the-precision-of-hatlambda-is-variable}{%
-\subsection{\texorpdfstring{\emph{The Precision of \(\hat{\lambda}\) is
-Variable}}{The Precision of \textbackslash{}hat\{\textbackslash{}lambda\} is Variable}}\label{the-precision-of-hatlambda-is-variable}}
+\hypertarget{the-accuracy-and-precision-of-hatlambda-is-variable}{%
+\subsection{\texorpdfstring{\emph{The accuracy and precision of
+\(\hat{\lambda}\) is
+Variable}}{The accuracy and precision of \textbackslash hat\{\textbackslash lambda\} is Variable}}\label{the-accuracy-and-precision-of-hatlambda-is-variable}}
 
@@
 For each dataset we then estimated phylogenetic signal
-(\(\hat{\lambda}\)), and calculated its variance
-(\(\sigma^2_{\hat{\lambda}}\)) at each input level of phylogenetic
-signal and level of species richness as an estimate of precision. Random
-data samples were generated from a normal distribution, i.e.,
-\(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\). We verified that the
-level of \(\sigma^2\) had no effect on \(\hat{\lambda}\) or
-\(\sigma^2_{\hat{\lambda}}\), but the results we present used
-\(\sigma^2=1\). \hfill\break
+(\(\hat{\lambda}\)), and calculated the standard deviation of
+\({\hat{\lambda}\) (\(\sigma_{\hat{\lambda}}\)) at each input level of
+phylogenetic signal and level of species richness as an estimate of
+precision. Random data samples were generated from a normal
+distribution, i.e., \(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\). We
+verified that the level of \(\sigma^2\) had no effect on
+\(\hat{\lambda}\) or \(\sigma^2_{\hat{\lambda}}\), but the results we
+present used \(\sigma^2=1\). \hfill\break
 
@@
 
-\emph{Results.} We found that the precision of \(\lambda_{est}\) varied
-widely across simulation conditions. Predictably, precision improved as
-the number of species increased (Figure 2). This confirmed earlier
-findings of Boettiger et al. (2012), and adhered to parametric
-statistical theory. However, in many cases the set of \(\lambda_{est}\)
-spanned nearly the entire range of possible values (e.g., \(n=32\);
-\(\lambda_{in}=0.5\): \(\lambda_{est}= 0.0\to 0.985\)), revealing that
-estimates of \(\lambda\) were not a reliable indicator of input
-phylogenetic signal. Importantly, the precision of \(\lambda_{est}\) was
-not uniform across all levels of phylogenetic signal. The worst
-precision was observed at intermediate levels of phylogenetic signal
-(\(\lambda_{in}\approx0.5\)), while precision improved as input levels
-approached the extremes of \(\lambda\)'s range (i.e.,
-\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
-estimates of \(\lambda\) were least reflective of the true input signal
-at intermediate values. Additionally, even at large levels of species
-richness, we found that the range of \(\lambda_{est}\) still encompassed
-a substantial portion of possible values (e.g., \(n=512\);
-\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
-same \(\lambda_{est}\) could be obtained from datasets containing vastly
-different input levels of phylogenetic signal (e.g., \(n=512\);
-\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). These
-findings were particularly unsettling when considered in light of our
-literature survey. Over one quarter of the \(\lambda\) estimates
-published in empirical studies (421 of 1,572) were between
-\(\lambda=0.25\) and \(\lambda=0.75\) (Figure 1). This range reflected
-the region that our simulations identified as being the least reliable
-in terms of accurately characterizing levels of phylogenetic signal, yet
-30\% of these mid-range empirical estimates were explicitly interpreted
-in terms of the strength of phylogenetic signal that they represented
-(i.e., weak, intermediate, strong phylogenetic signal). \hfill\break
+\emph{Results.}
+
+superifical Bernoulli traits innacuracy precision
+
+We found that the precision of \(\lambda_{est}\) varied widely across
+simulation conditions. Predictably, precision improved as the number of
+species increased (Figure 2). This confirmed earlier findings of
+Boettiger et al.~(2012), and adhered to parametric statistical theory.
+However, in many cases the set of \(\lambda_{est}\) spanned nearly the
+entire range of possible values (e.g., \(n=32\); \(\lambda_{in}=0.5\):
+\(\lambda_{est}= 0.0\to 0.985\)), revealing that estimates of
+\(\lambda\) were not a reliable indicator of input phylogenetic signal.
+Importantly, the precision of \(\lambda_{est}\) was not uniform across
+all levels of phylogenetic signal. The worst precision was observed at
+intermediate levels of phylogenetic signal (\(\lambda_{in}\approx0.5\)),
+while precision improved as input levels approached the extremes of
+\(\lambda\)'s range (i.e., \(\lambda_{in}\rightarrow0\) \&
+\(\lambda_{in}\rightarrow1\)). Thus, estimates of \(\lambda\) were least
+reflective of the true input signal at intermediate values.
+Additionally, even at large levels of species richness, we found that
+the range of \(\lambda_{est}\) still encompassed a substantial portion
+of possible values (e.g., \(n=512\); \(\lambda_{in}=0.5\):
+\(\lambda_{est} = 0.32\to 0.68\)). Likewise, the same \(\lambda_{est}\)
+could be obtained from datasets containing vastly different input levels
+of phylogenetic signal (e.g., \(n=512\); \(\lambda_{est} = 0.5\);
+\(\lambda_{in} = 0.25 \to 0.65\)). These findings were particularly
+unsettling when considered in light of our literature survey. Over one
+quarter of the \(\lambda\) estimates published in empirical studies (421
+of 1,572) were between \(\lambda=0.25\) and \(\lambda=0.75\) (Figure 1).
+This range reflected the region that our simulations identified as being
+the least reliable in terms of accurately characterizing levels of
+phylogenetic signal, yet 30\% of these mid-range empirical estimates
+were explicitly interpreted in terms of the strength of phylogenetic
+signal that they represented (i.e., weak, intermediate, strong
+phylogenetic signal). \hfill\break
 
@@
 for \(\lambda\) and \(\kappa\) values are consistent with those found in
-the simulations performed by Münkemüller et al. (2012), but that study
+the simulations performed by Münkemüller et al.~(2012), but that study
 investigated type I error rates and statistical power, finding that
@@
 Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
-Software for geometric morphometric analyses. R package version 3.2.1.
+Software for geometric morphometric analyses. R package version 3.3.1.
 
@@
 \leavevmode\hypertarget{ref-Orme2013}{}%
-Orme, D., R. Freckleton, G. Thomas, T. Petzoldt, S. Fritz, and N. Isaac.
-2013. CAPER: Comparative analyses of phylogenetics and evolution in r.
-Methods in Ecology and Evolution 3:145--151.
+Orme, D., R. P. Freckleton, G. H. Thomas, T. Petzoldt, S. A. Fritz, and
+N. Isaac. 2013. CAPER: Comparative analyses of phylogenetics and
+evolution in r. Methods in Ecology and Evolution 3:145--151.
 

### 2020-07-15 Manuscript/OldVsns/Manuscript-7.Rmd
pnas vsn

--- 2020-05-11 Manuscript/Manuscript-6-EKB.Rmd
@@
 
-In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic ~~to describe~~ CHARACTERIZING the relative strength of phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
-~~The appeal~~ THIS DESCRIPTIVE UTILITY of $\lambda$ ~~as a descriptive statistic for evolutionary biologists~~ is ~~a~~ THE basis for interpreting "weak" versus "strong" phylogenetic signal~~,~~ (i.e., small versus large values of $\lambda$, respectively)~~,~~ in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
+In addition to functioning as a parameter that is tuned for appropriate analysis, $\lambda$ can function as a descriptive statistic characterizing the relative strength of phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has influenced trait covariation among taxa.
+The appeal of $\lambda$ as a descriptive statistic is that it serves as the basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\lambda$, respectively, in a comparative sense [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
 Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
-For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% EXPLICITLY interpreted the strength of phylogenetic signal for at least one phenotypic trait.
-Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.
+For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% explicitly interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion), this percentage is even higher.
 \hfill\break
@@
 
-Various other approaches use $\lambda$ as a parameter that can be ~~varied~~ ADJUSTED for inferences akin to sensitivity analysis.
+Various other approaches use $\lambda$ as a parameter that can be adjusted for inferences that are akin to sensitivity analysis.
 For instance, some have performed likelihood ratio tests that compare observed model fits to those obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or evaluated whether observed $\lambda$ differs from an expected $\lambda$, based on confidence intervals generated for the expected value [@Vandelook2019].
@@
 It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
-Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion.
 However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
@@
 One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$).
-From this observation it was concluded that insufficient data (i.e., ~~the~~ LOW number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
+From this observation it was concluded that insufficient data (i.e., low number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012].
 Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988].
@@
 Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b].
-Unfortunatley, not all model parameters and descriptive statistics are effect sizes, and thus many summary measures must first be converted to statistics with standardized units (i.e., conversion to an effect size) for meaningful comparison [see @Rosenthal1994].
+Unfortunately, not all model parameters and descriptive statistics are effect sizes, and thus many summary measures must first be converted to statistics with standardized units (i.e., conversion to an effect size) for meaningful comparison [see @Rosenthal1994].
 As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable.
-~~For the case of $\lambda$, this~~ HOWEVER, THE CALCULATION AND STATISTICAL BEHAVIOR OF $\lambda$ AS AN EFFECT SIZE has not yet been explored.
+However, the calculation and statisticl behavior of $\lambda$ as an effect size has not yet been explored.
 \hfill\break

### 2020-07-15 Manuscript/OldVsns/Manuscript-8-DCA.Rmd
pnas vsn

---
title:

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: pnas.csl
bibliography: Lambda-Refs.bib

header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---

**Classification:**  Biological Sciences -- Evolution
\hfill\break

\begin{center}
\textbf{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}
\end{center}

\hfill\break

\begin{center}
\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
\end{center}

\hfill\break

$^{1}$Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.

$^{2}$Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.

$^{*}$Correspondence: Dean C. Adams. Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA. 515-294-3834. <dcadams@iastate.edu>

\hfill\break

\begin{center}`r format(Sys.time(), '%d %B, %Y')`\end{center}

\hfill\break

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(stringr)
library(tidyverse)

RmdWords <- function(file) {

  # Creates a string of text
  file_string <- file %>%
    readLines() %>%
    paste0(collapse = " ") %>%
    # Remove YAML header
    str_replace_all("^<--- .*?--- ", "") %>%    
    str_replace_all("^--- .*?--- ", "") %>%
    # Remove code
    str_replace_all("```.*?```", "") %>%
    str_replace_all("`.*?`", "") %>%
    # Remove LaTeX
    str_replace_all("[^\\\\]\\$\\$.*?[^\\\\]\\$\\$", "") %>%
    str_replace_all("[^\\\\]\\$.*?[^\\\\]\\$", "") %>%
    # Deletes text between tags
    str_replace_all("TC:ignore.*?TC:endignore", "") %>%
    str_replace_all("[[:punct:]]", " ") %>%
    str_replace_all("  ", "") %>%
    str_replace_all("<", "") %>%
    str_replace_all(">", "")

  # Save several different results
  word_count <- str_count(file_string, "\\S+")
  char_count <- str_replace_all(string = file_string, " ", "") %>% str_count()

   return(list(num_words = word_count, num_char = char_count, word_list = file_string))
}

words <- RmdWords("Manuscript-8-DCA.Rmd")
```

```{r, echo = FALSE}
count <- c(words = words$num_words, characters = words$num_char)
knitr::kable(t(count))
```

\newpage

# Abstract

Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits. However, analytical tools for such comparisons have largely remained underdeveloped. Here we evaluate the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal. We find that $\lambda$ behaves as a Bernoulli random variable, and that estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal. Further, the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised. We propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison. We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales. \hfill\break

**Keywords**: phylogenetic signal, macroevolution, lambda, kappa

\newpage

# Significance Statement
Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across traits, but analytical tools for these comparisons are generally lacking. Here we develop a standardized effect size based on $\kappa$ ($Z_\kappa$), which measures the strength of phylogenetic signal on a common statistical scale, and provides a mechanism for formally comparing the strength of phylogenetic signal across datasets. Additionally, we find that a commonly used parameter (Pagel's $\lambda$) is insuitable for this purpose. Our procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

\newpage

# Introduction

Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991]. Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004; @MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b]. These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003]. Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008]. \hfill\break

Several analytical tools have been developed to quantify phylogenetic signal in phenotypic datasets [@Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999; @Klingenberg2010], and their statistical properties -- namely type I error rates and statistical power -- have been investigated to determine under what conditions phylogenetic signal can be detected [@Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @Adams2014a; @MolinaVenegas2017; @Revell_et_al2008; @Revell2010; @Boettiger_et_al2012]. One of the most widely used methods for characterizing phylogenetic signal is Pagel's $\lambda$ [@Pagel1999], which transforms the lengths of the internal branches of the phylogeny to improve the fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002]. When incorporated in PGLS, $\lambda$ serves as a tuning parameter (ranging from $0\rightarrow1$) which is optimized via log-likelihood profiling while evaluating the covariation between the dependent and independent variables, given the phylogeny [@Pagel1999; @Freckleton_et_al2002]. To infer whether phylogenetic signal differs from no signal or a Brownian motion model of evolutionary divergence, the observed model fit using $\hat\lambda$ may be statistically compared to that using  $\lambda=0$ or $\lambda=1$ via likelihood ratio tests [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or confidence limits [@Vandelook2019]. Another widely used measure is Blomberg's $\kappa$ [@Blomberg_et_al2003], which measures the ratio of observed trait variation to the amount expected under Brownian motion. Blomberg's $\kappa$ can be treated as a test statistic by employing a permutation test to generate its sampling distribution [@Blomberg_et_al2003; @Adams2014a] for determining whether significant phylogenetic signal is present in data. Both $\lambda$ and $\kappa$ seem intuitive to interpret, as a value of $0$ for both corresponds to no phylogenetic signal, while a value of $1$ corresponds to the amount of phylogenetic signal expected under Brownian motion. Thus, it is tempting to regard both $\lambda$ and $\kappa$ as descriptive statistics that measure the relative strength of phylogenetic signal, providing an estimate of its magnitude for comparison.  \hfill\break

The appeal of Pagel's $\lambda$ and Blomberg's $\kappa$ as descriptive statistics is that they provide a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$ or $\kappa$, respectively, in a comparative sense [@DeMeester2019; @Pintanel2019;@Su2019]. Nonetheless, an important question that has yet to be considered is whether these statistics are or can be converted to effect sizes for comparative analyses across datasets? To be statistics representing phylogenetic signal, they should have reliable distributional properties, which could be revealed with simulation experiments. For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ follows a Bernoulli distribution (**add ref**); i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process. Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to treesize; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 (**add same ref**). Prior work [@Munkemuller_et_al2012] seems to support some of these conjectures, based superficially on statistical moments for a given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of [@Munkemuller_et_al2012]). However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging. \hfill\break

By contrast, Blomberg's $\kappa$, which is positively ubounded, might be expected to follow a normal distribution (**add ref**). Thus we might expect that $\kappa$ is symmetrically distributed across different strengths of phylogenetic signal, for any $\lambda$ used to generate data. This attribute seemed less reasonable based on the simulations performed by Münkemüller et al. [-@Munkemuller_et_al2012], as distibutions were postively skewed, suggesting the Blomberg's $\kappa$ might not behave as a statistic that follows a noraml distribution. However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations). For both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.  Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging. Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparative anlyses. \hfill\break

Here we used simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and Blomberg's $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.  We find that estimates of $\hat{\lambda}$ are increasingly skewed at larger and smaller input levels of phylogenetic signal, vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range. By contrast, estimates of $\kappa$ are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$. We then propose an effect size based on $\kappa$ ($Z_{\kappa}$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.

# Results

**Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations revealed that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld. First, the mean value of $\hat{\lambda}$ increased as $\lambda$ increased (Fig. 1 black line); the standard deviation of $\hat{\lambda}$ was largest at intermediate values of $\lambda$ and smallest at extreme values (Fig. 1 red line); and the distributions of $\hat{\lambda}$ were increasing skewed at extreme values of $\lambda$ but rather normal at intermediate values (Fig. 1 blue line). For small tree sizes, it was also clear that distributions were more platykurtik at intermeidate values of $\hat{\lambda}$. Second, $\hat{\lambda}$ was a negatively-biased estimate of $\lambda$ for small tree sizes, as the mean value was consistently less than the input $\lambda$ value. Third, standard deviations of $\hat{\lambda}$ were negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ was quite variable, except for cases when $\lambda$ was near or equal to $1$. As such its statistical properties make it insuitable as an effect size for measuring the strength of phylogenetic signal. \hfill\break

**Kappa ($\kappa$) estimates of phylogenetic signal are more stable.** Results for $\kappa$ demonstrated better statistical properties. First, mean values increased consistently with $\lambda$ irespective of tree size (Fig. 2 black line); the standard deviation of $\kappa$ was consistent across tree sizes (Fig. 2 red line); and $\kappa$ was normally distributed regardless of $\lambda$ or tree size (Fig. 2 blue line).  The standard deviation of $\kappa$ increased with $\lambda$, somewhat tracking the mean, but was always less than the corresponding means. This is perhaps unsurprising, as $\kappa$ is bound by 0, and was never large for small values of $\lambda$.  Finally, distributions of $\kappa$ displayed slight skewing only at small tree sizes and when $\lambda$ was near or equal to 1. This differs from the results of [-@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently. \hfill\break

**Effect sizes from $\kappa$ ($Z_{\kappa}$) better characterize phylogenetic signal.**
We derived effect sizes (Z-scores) for both $\lambda$ and $\kappa$ (see Methods) and evaluated them relative to known phylogenetic signal. Both $Z_{\lambda}$ and $Z_{\kappa}$ were associated with input phylogenetic signal ($\lambda$), indicating that both statistics captured the observed signal (Fig. 3). However, effect sizes from $\hat{\lambda}$ made little sense, as they were more strongly associated with tree size than the actual phylogenetic signal in the data (Fig. 3). By contrast, $Z_{\kappa}$ was much more consistent accross tree sizes, and exhibited a much stronger association with phylogenetic signal strength as compared to tree size (Fig. 3). Thus between the two statistics, $Z_{\kappa}$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets. \hfill\break

**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** We derived a two-sample test statistic ($\hat{Z}_{12}$) for statistically comparing the strength of phylogenetic signal across datasets (see Methods). To demonstrate its utility, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig. 4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020]. While both contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V (Fig. 4). Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020].

# Discussion

address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales

# Methods

**Derivation of effect sizes.** Statistically, a standardized effect size may be found as:

\begin{align}
    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
\end{align}

where $\theta_{obs}$ is the observed test statistic, $E(\theta)$ is its expected value under the null hypothesis, and $\sigma_\theta$ is its standard error [@Glass1976; @Cohen1988;@Rosenthal1994]. Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory. However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures. \hfill\break

Formalizing the suggestion of Adams and Collyer [@AdamsCollyer2019],  an effect size for $\kappa$ may be found as:

\begin{align}
    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
\end{align}

where $\kappa_{obs}$ is the observed phylogenetic signal, and $\hat\mu_\kappa$ and $\hat\sigma_\kappa$ are the mean and standard deviation of the empirical sampling distribution of $\kappa$ obtained via permutation. The empirical sampling distribution of $\kappa$ is first transformed via Box-Cox to better adhere to the assumption of normality. \hfill\break

For $\lambda$ deriving an effect size is more challenging, it does not have a sampling distribution from which the standard error (and thus confidence intervals) may be obtained. Thus, its confidence intervals are generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the compare the putative appropriate model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].  As such, an effect size for $\lambda$ may be found as:

\begin{align}
   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
\end{align}

where, $\hat{\lambda}$ is the maximized likelihood value of $\lambda$ and $\chi^{2}_{\hat{\lambda}}$ is the likelihood ratio statistic for the value. Note that alternative formulations could be envisioned: $Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}$, where $d$ is a binary value $(-1,1)$ to indicate a direction based on whether $\hat{\lambda}$ is below or above a critical value of $\lambda$, for a quantile from a $\chi^{2}$ distribution at a probability of 0.5. However, preliminary investigations found that mapping $\chi^2_{\hat{\lambda}}$ values in this manner did not produce effect sizes that were symmetrical about $Z = 0$, as the mapping was not linear and the log-likelihood profiles can be rather flat for small trees. \hfill\break

**Derivation of two-sample test statistic.** To compare the strength of phylogenetic signal across datasets, a two-sample test statistic may be calculated as:

\begin{align}
  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
\end{align}

where $\kappa_1$, $\kappa_2$, $\hat\mu_{\kappa_1}$, $\hat\mu_{\kappa_2}$, $\hat\sigma_{\kappa_1}$, and $\hat\sigma_{\kappa_2}$ are as defined above for equation 2. The right side of the equation illustrates that if $Z_\kappa$ has already been calculated for two sampling distributions as in equation 2, the sampling distributions have unit variance for each of the $Z_\kappa$ statistics. Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution. Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b]. \hfill\break

**Simulations.** Simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylognetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$). We generated 100 random trees for each intersection of tree size and $\lambda$. For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution. For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and $\kappa$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry). For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness. Simulations were then repeated for both bananced and pectinate trees, which yielded qualitatively similar results (see Supporting Information). Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information). \hfill\break

**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from individuals of 305 species (2,781 and 3,371 respectively), from which species means were obtained [@Baken2019; @Baken2020]. A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the phenotypic dataset. The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph].

# Acknowledgements

This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.).

\newpage
# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>

\newpage

Real figures will have 6 panels for lambda and kappa and probably only a subset for z.  Simulations need to be re-run.

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.1.temp.png")

```

\singlespacing \textbf{Figure 1}.  Temporary Fig. 1 showing lambda patterns

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.2.temp.png")

```

\singlespacing \textbf{Figure 2}.  Temporary Fig. 2 showing kappa patterns

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.3.temp.png")

```

\singlespacing \textbf{Figure 3}.  Temporary Fig. 3 showing z patterns

\newpage

```{r, echo = FALSE, out.width="95%"}
include_graphics("new.fig.4.png")

```

\singlespacing \textbf{Figure 4}.  (A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios. (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars. (C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).

### 2020-07-15 Manuscript/PNAS-Adams-etal.Format.Rmd
pnas vsn

--- 2020-07-15 Manuscript/OldVsns/Manuscript-8-DCA.Rmd
@@
 ---
-title:
+title: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal
 
-output: 
-  pdf_document:
-    fig_caption: yes
-    number_sections: no  #note: change to 'no' for most journals
-    keep_tex: true
-    pandoc_args:
+author:
+  - name: Dean C. Adams
+    affiliation: a
+  - name: Erica K. Baken
+    affiliation: a,b
+  - name: Michael L. Collyer
+    affiliation: b
+address:
+  - code: a
+    address: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.
+  - code: b
+    address: Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.
 
-csl: pnas.csl
-bibliography: Lambda-Refs.bib
+# For footer text  TODO(fold into template, allow free form two-authors)
+lead_author_surname: Adams et. al
 
-header-includes:
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{lineno}\linenumbers
-
----
-
-**Classification:**  Biological Sciences -- Evolution
-\hfill\break
-
-\begin{center}
-\textbf{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}
-\end{center}
-
-\hfill\break
-
-\begin{center}
-\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
-\end{center}
-
-\hfill\break
-
-$^{1}$Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.
-
-$^{2}$Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.
-
-$^{*}$Correspondence: Dean C. Adams. Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA. 515-294-3834. <dcadams@iastate.edu>
-
-\hfill\break
-
-\begin{center}`r format(Sys.time(), '%d %B, %Y')`\end{center}
-
-\hfill\break
-
-```{r setup, include=FALSE}
-  library(knitr)
-  knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r, include = FALSE}
-library(stringr)
-library(tidyverse)
-
-RmdWords <- function(file) {
-
-  # Creates a string of text
-  file_string <- file %>%
-    readLines() %>%
-    paste0(collapse = " ") %>%
-    # Remove YAML header
-    str_replace_all("^<--- .*?--- ", "") %>%    
-    str_replace_all("^--- .*?--- ", "") %>%
-    # Remove code
-    str_replace_all("```.*?```", "") %>%
-    str_replace_all("`.*?`", "") %>%
-    # Remove LaTeX
-    str_replace_all("[^\\\\]\\$\\$.*?[^\\\\]\\$\\$", "") %>%
-    str_replace_all("[^\\\\]\\$.*?[^\\\\]\\$", "") %>%
-    # Deletes text between tags
-    str_replace_all("TC:ignore.*?TC:endignore", "") %>%
-    str_replace_all("[[:punct:]]", " ") %>%
-    str_replace_all("  ", "") %>%
-    str_replace_all("<", "") %>%
-    str_replace_all(">", "")
-
-  # Save several different results
-  word_count <- str_count(file_string, "\\S+")
-  char_count <- str_replace_all(string = file_string, " ", "") %>% str_count()
-
-   return(list(num_words = word_count, num_char = char_count, word_list = file_string))
-}
-
-words <- RmdWords("Manuscript-8-DCA.Rmd")
-```
-
-```{r, echo = FALSE}
-count <- c(words = words$num_words, characters = words$num_char)
-knitr::kable(t(count))
-```
-
-\newpage
+# Place DOI URL or CRAN Package URL here
+doi_footer: "doi: TBD"
 
 # Abstract
-
+abstract: |
   Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits.
@@
 Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
-\hfill\break
 
-**Keywords**: phylogenetic signal, macroevolution, lambda, kappa
-
-\newpage
-
-# Significance Statement
+# Significance
+significance: |
   Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across traits, but analytical tools for these comparisons are generally lacking.
@@
 
-\newpage
+# Optional: Acknowledgements
+acknowledgements: |
+   This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.).
+
+# Optional: One or more keywords
+keywords: phylogenetic signal, macroevolution, lambda, kappa
+
+# Paper size for the document, values of letter and a4
+papersize: letter
+
+# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
+fontsize: 9pt
+
+# Optional: Force one-column layout, default is two-column
+#one_column: true
+
+# Optional: Enables lineo mode, but only if one_column mode is also true
+lineno: true
+
+# Optional: Enable one-sided layout, default is two-sided
+#one_sided: true
+
+# Optional: Enable section numbering, default is unnumbered
+# numbersections: true
+
+# Optional: Specify the depth of section number, default is 5
+secnumdepth: 3
+
+# Optional: Bibliography 
+csl: pnas.csl
+bibliography: Lambda-Refs.bib
+
+# Optional: include-after
+#include-after: somefile.tex
+
+# Optional: Skip inserting final break between acknowledgements, default is false
+skip_final_break: true
+
+# Optional: Enable a 'Draft' watermark on the document
+watermark: true
+
+# Produce a pinp document
+# 
+# Other options that can be set here, shown with their defaults
+#   keep_tex: TRUE
+#   collapse: FALSE
+#   citation_package: natbib
+#
+# Any other arguments to rmarkdown::pdf_document could be set 
+# here as well
+#
+output:
+  pinp::pinp:
+#    includes:
+#      after_body: bib.tex
+
+---
 
@@
 Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
-\hfill\break
 
@@
 Thus, it is tempting to regard both $\lambda$ and $\kappa$ as descriptive statistics that measure the relative strength of phylogenetic signal, providing an estimate of its magnitude for comparison.
-\hfill\break
 
@@
 However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
-\hfill\break
 
@@
 Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparative anlyses.
-\hfill\break
 
@@
 As such its statistical properties make it insuitable as an effect size for measuring the strength of phylogenetic signal.
-\hfill\break
+
+![Patterns in $\lambda$.<span data-label="fig:f1"></span>](new.fig.1.temp.png)
 
@@
 This differs from the results of [-@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently.
-\hfill\break
+
+![Patterns in $\kappa$.<span data-label="fig:f2"></span>](new.fig.2.temp.png)
 
@@
 Thus between the two statistics, $Z_{\kappa}$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets.
-\hfill\break
+
+![Patterns in Z.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
 
@@
 Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020].
+
+![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](new.fig.4.png)
 
@@
 However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures.
-\hfill\break
 
@@
 However, preliminary investigations found that mapping $\chi^2_{\hat{\lambda}}$ values in this manner did not produce effect sizes that were symmetrical about $Z = 0$, as the mapping was not linear and the log-likelihood profiles can be rather flat for small trees.
-\hfill\break
 
@@
 Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b].
-\hfill\break
 
@@
 Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information).
-\hfill\break
 
@@
 
-# Acknowledgements
+# Code
 
-This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.).
+Pandoc segments are highlighted as usual, and per a convention in this
+template also shown in a framed and slightly shaded box as seen here and
+above.
 
-\newpage
-# References
+Another example from Python:
 
- \setlength{\parindent}{-0.25in}
- \setlength{\leftskip}{0.25in}
- \setlength{\parskip}{8pt}
- \noindent
-
-<div id="refs"></div>
-
-\newpage
-
-Real figures will have 6 panels for lambda and kappa and probably only a subset for z.  Simulations need to be re-run.
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("new.fig.1.temp.png")
+```py
+# A Python example
+def printSomething(str):
+   "This prints the string passed in"
+   print str
 
@@
 
-\singlespacing \textbf{Figure 1}.  Temporary Fig. 1 showing lambda patterns
-
-\newpage
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("new.fig.2.temp.png")
-
-```
-
-\singlespacing \textbf{Figure 2}.  Temporary Fig. 2 showing kappa patterns
-
-\newpage
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("new.fig.3.temp.png")
-
-```
-
-\singlespacing \textbf{Figure 3}.  Temporary Fig. 3 showing z patterns
-
-\newpage
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("new.fig.4.png")
-
-```
-
-\singlespacing \textbf{Figure 4}.
-(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
-(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
+<!-- Acknowledgements, and References, are automatically typeset after this -->

### 2020-07-15 Manuscript/PNAS-Adams-rticle.Rmd
pnas vsn

--- 2020-07-15 Manuscript/PNAS-Adams-etal.Format.Rmd
@@
 
+# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
 author:
@@
 
-# For footer text  TODO(fold into template, allow free form two-authors)
-lead_author_surname: Adams et. al
+corresponding_author:
+  - code: 2
+    text: "To whom correspondence should be addressed. E-mail: dcadams@iastate.edu"
 
-# Place DOI URL or CRAN Package URL here
-doi_footer: "doi: TBD"
+# For footer text
+lead_author_surname: Adams et al.
 
-# Abstract
+author_contributions: |
+  D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the research and wrote the paper.
+
+## Remove this if not required
+conflict_of_interest: |
+  The authors declare no conflict of interest.
+
 abstract: |
@@
 
-# Significance
 significance: |
@@
 
-# Optional: Acknowledgements
 acknowledgements: |
-   This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.).
+  We thank E. Glynne and B. Juarez for comments on early drafts of the manuscript. This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break
 
-# Optional: One or more keywords
-keywords: phylogenetic signal, macroevolution, lambda, kappa
+keywords:
+  - phylogenetic signal
+  - macroevolution
+  - lambda
+  - kappa
 
-# Paper size for the document, values of letter and a4
-papersize: letter
+## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
+pnas_type: pnasresearcharticle
 
-# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
-fontsize: 9pt
+bibliography: Lambda-Refs.bib
+csl: pnas.csl
 
-# Optional: Force one-column layout, default is two-column
-#one_column: true
-
-# Optional: Enables lineo mode, but only if one_column mode is also true
+## change to true to add optional line numbering
 lineno: true
 
-# Optional: Enable one-sided layout, default is two-sided
-#one_sided: true
-
-# Optional: Enable section numbering, default is unnumbered
-# numbersections: true
-
-# Optional: Specify the depth of section number, default is 5
-secnumdepth: 3
-
-# Optional: Bibliography 
-csl: pnas.csl
-bibliography: Lambda-Refs.bib
-
-# Optional: include-after
-#include-after: somefile.tex
-
-# Optional: Skip inserting final break between acknowledgements, default is false
-skip_final_break: true
-
-# Optional: Enable a 'Draft' watermark on the document
-watermark: true
-
-# Produce a pinp document
-# 
-# Other options that can be set here, shown with their defaults
-#   keep_tex: TRUE
-#   collapse: FALSE
-#   citation_package: natbib
-#
-# Any other arguments to rmarkdown::pdf_document could be set 
-# here as well
-#
-output:
-  pinp::pinp:
-#    includes:
-#      after_body: bib.tex
-
+output: rticles::pnas_article
 ---
-
-# Introduction
-
 Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
@@
 
-address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales
+**To be edited.** address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
+
+It is common in comparative evolutionary studies to characterize the phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has generated trait covariation among taxa.
+However, while numerous analytical approaches may be used to quantify phylogenetic signal [e.g., @Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999], methods that explicitly measure the strength of phylogenetic signal, or facilitate comparisons among datasets, have remained underdeveloped.
+In this study, we evaluated the precision of one common measure, Pagel's $\lambda$, and explored its efficacy for characterizing the strength of phylogenetic signal in phenotypic data.
+Using computer simulations, we found that the precision of $\lambda$ increased with increasing sample sizes; a pattern noted previously [@Boettiger_et_al2012], and one that conformed with parametric statistical theory [@Cohen1988].
+However, we also found that vastly different $\lambda$ estimates could be obtained from data containing the same level of phylogenetic signal, and that similar $\lambda$ estimates may be obtained from data containing differing levels of phylogenetic signal.
+Further, the precision of $\lambda$ varied with the strength of phylogenetic signal, where lower precision was observed when in data whose phylogenetic signal was of intemediate strength.
+From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and that biological interpretations of the strength of signal based on this parameter may innacurately characterize such effects.
+
+As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal.
+$Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
+We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
+First, $Z_\kappa$ was more precise than $Z_\lambda$, and precision was more consistent across the range of input levels of phylogenetic signal.
+Additionally, values of $Z_\kappa$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa$ and $\lambda_{in}$.
+Thus, $Z_\kappa$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
+We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_\kappa$ as a statistical measure of this effect.
+
+Based on the effect size $Z_\kappa$, we then proposed a two-sample test, which provides means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another, via a hypothesis test.
+Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $\kappa$ [e.g.,  @Blomberg_et_al2003].
+However, $\kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
+Thus, $\kappa$ should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
+By contrast, standardizing $\kappa$ ($Z_\kappa$, via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
+As such, our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
+
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
+Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
+For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
+In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
+Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
+
+This study is not the first to compare $\lambda$ and $\kappa$ for their ability as statistics to measure phylogenetic signal.
+Our results for $\lambda$ and $\kappa$ values are consistent with those found in the simulations performed by Münkemüller et al.
+[-@Munkemuller_et_al2012], but that study investigated type I error rates and statistical power, finding that $\lambda$ performed better in both regards, irrespective of species number in trees.
+Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent $\kappa$ at intermediate phylogenetic signal strengths was observed [see Fig.
+2, @Munkemuller_et_al2012].
+Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
+Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
+An interesting question is whether an inflated $\kappa$ value leads to an inflated $Z_\kappa$ or does a tendency of a particular tree to inflate estimates of $\kappa$ also inflate the values in random permutations of a test, in which case $Z_\kappa$ is robust to polytomies?
+We repeated the analyses in Figure 4, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (see Supporting Information).
+This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size.
+Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+
+Phylogenetic signal can be thought of as both an attribute to be measured in the data and a parameter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
+As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
+However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a statistic to describe the amount of phylogenetic signal in the data.
+Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
+Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
+By contrast, $\kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_\kappa$).
+Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
+With low coefficients of variation for $Z_\kappa$ (Figure 4), it is difficult to imagine that a hypothesis test can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
 
@@
 
-# Code
-
-Pandoc segments are highlighted as usual, and per a convention in this
-template also shown in a framed and slightly shaded box as seen here and
-above.
-
-Another example from Python:
-
-```py
-# A Python example
-def printSomething(str):
-   "This prints the string passed in"
-   print str
-
-```
-
-<!-- Acknowledgements, and References, are automatically typeset after this -->
+<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
+\showmatmethods
+\showacknow
+\pnasbreak

### 2020-07-15 Manuscript/pnas_article-Template/skeleton/PNAS-Adams-rticle.Rmd
pnas vsn

---
title: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Dean C. Adams
    affiliation: a
  - name: Erica K. Baken
    affiliation: a,b
  - name: Michael L. Collyer
    affiliation: b
address:
  - code: a
    address: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.
  - code: b
    address: Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.

corresponding_author:
  - code: 2
    text: "To whom correspondence should be addressed. E-mail: dcadams@iastate.edu"

# For footer text
lead_author_surname: Adams et al.

author_contributions: |
  D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the research and wrote the paper.

## Remove this if not required
conflict_of_interest: |
  The authors declare no conflict of interest.

abstract: |
  Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits. However, analytical tools for such comparisons have largely remained underdeveloped. Here we evaluate the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal. We find that $\lambda$ behaves as a Bernoulli random variable, and that estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal. Further, the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised. We propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison. We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

significance: |
  Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across traits, but analytical tools for these comparisons are generally lacking. Here we develop a standardized effect size based on $\kappa$ ($Z_\kappa$), which measures the strength of phylogenetic signal on a common statistical scale, and provides a mechanism for formally comparing the strength of phylogenetic signal across datasets. Additionally, we find that a commonly used parameter (Pagel's $\lambda$) is insuitable for this purpose. Our procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

acknowledgements: |
  We thank E. Glynne and B. Juarez for comments on early drafts of the manuscript. This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.).

keywords:
  - phylogenetic signal
  - macroevolution
  - lambda
  - kappa

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: Lambda-Refs.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

output: rticles::pnas_article
---

This PNAS journal template is provided to help you write your work in the
correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
"Introduction" as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

Guide to using this template {.unnumbered}
========================================

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the [PNAS Information for
Authors](http://www.pnas.org/site/authors/format.xhtml).

Author Affiliations {#author-affiliations .unnumbered}
-------------------

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

Submitting Manuscripts {#submitting-manuscripts .unnumbered}
----------------------

All authors must submit their articles at
[PNAScentral](http://www.pnascentral.org/cgi-bin/main.plex). If you are
using Overleaf to write your article, you can use the "Submit to PNAS"
option in the top bar of the editor window.

Format {#format .unnumbered}
------

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

Manuscript Length {#manuscript-length .unnumbered}
-----------------

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for PNAS PLUS.

References {#references .unnumbered}
----------

References should be cited in numerical order as they appear in text;
this will be done automatically via bibtex, e.g.  All references, including
for the SI, should be included in the main manuscript file. References
appearing in both sections should not be duplicated. SI references
included in tables should be included with the main reference section.

Data Archival {#data-archival .unnumbered}
-------------

PNAS must be able to archive the data essential to a published article.
Where such archiving is not possible, deposition of data in public
databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
and others outlined in the Information for Authors, is acceptable.

Language-Editing Services {#language-editing-services .unnumbered}
-------------------------

Prior to submission, authors who believe their manuscripts would benefit
from professional editing are encouraged to use a language-editing
service (see list at www.pnas.org/site/authors/language-editing.xhtml).
PNAS does not take responsibility for or endorse these services, and
their use has no bearing on acceptance of a manuscript for publication.

![Placeholder image of a frog with a long example caption to show
justification setting.<span data-label="fig:frog"></span>](frog.png)

Digital Figures {#sec:figures .unnumbered}
---------------

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for
figures that will appear in the main text, and images must be final
size. Authors may submit U3D or PRC files for 3D images; these must be
accompanied by 2D representations in TIFF, EPS, or high-resolution PDF
format. Color images must be in RGB (red, green, blue) mode. Include the
font files for any text.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

Figure \[fig:frog\] shows an example of how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. Figures wider than one
column should be sized to 11.4 cm or 17.8 cm wide.

Single column equations {#single-column-equations .unnumbered}
-----------------------

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \[eqn:example\] below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

$$\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}$$

<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

Supporting Information (SI) {#supporting-information-si .unnumbered}
---------------------------

The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak

### 2020-07-15 Manuscript/VersionDiffNew.r

@@ -1,0 +1,34 @@
+# Script to compare two Rmd files
+
+# step by step, as 'latexdiffr' no longer working (breaks in the system2(latexdiff call)
+
+#1: generate tex from Rmd
+rmarkdown::render(input =  "2020-Evol-MultiRR_Submitted.Rmd")
+rmarkdown::render(input = "2020-Evol-MultiRR_Rvsn.Rmd")
+
+#2: latexdiff of tex files
+shell(cmd="latexdiff.exe 2020-Evol-MultiRR_Submitted.tex 2020-Evol-MultiRR_Rvsn.tex > diff.tex")
+
+#3: PDF of diff.tex
+tinytex::latexmk("diff.tex") 
+
+
+
+
+
+
+
+
+
+
+
+
+###########
+#Note, requires installation of 'latexdiff' in your tex components & a version of Perl
+
+# NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
+
+##OLD WAY: not working now
+#library(latexdiffr)
+#latexdiff("2020-Evol-MultiRR_Submitted.Rmd", "2020-Evol-MultiRR_Rvsn.Rmd", clean = TRUE)
+

### 2020-07-16 Data_Analyses/updated_simulations/simulation.source.R
Update simulation.source.R

@@ -196,7 +196,10 @@ get.stats <- function(tree, x, beta, lambda, type) {
     try(pgls(y ~ x, data = df, lambda='ML'), silent = T)
   kappa_est <- physignal(cbind(y,0), phy = tree, iter = 999, print.progress = F)
   if(!inherits(pgls_est, "try-error")) {
-    res <- c(unlist(pgls_est$model), unlist(pgls_est$param.CI)[c(1,6,7)],
+    summ <- summary(pgls_est)
+    Pval <- unlist(summ$coefficients[,4])
+    names(Pval) <- paste("Pval", names(Pval), sep = ".")
+    res <- c(unlist(pgls_est$model), Pval, unlist(pgls_est$param.CI)[c(1,6,7)],
              lambda.z = lambda.Z(pgls_est), 
              unlist(kappa_est[c("phy.signal", "Z")]))
     names(res)[length(res)] <- "kappa.z"

### 2020-07-16 Manuscript/OldVsns/Manuscript-2.Rmd

--- 2020-04-24 Manuscript/Manuscript-2.Rmd
@@
 This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019].
-Other approaches for interpreting $\lambda$ are more statistical, through the use of confidence intervals [@Vandelook2019] or likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
-Likewise, qualitative comparisons of $\lambda$ across multiple phenotypic traits have also been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
-Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion).
-However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
+Other approaches for interpreting $\lambda$ are more statistical.
+For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019].
+Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019].
+Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature.
+We conducted a literature survey in Google.scholar and found that of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$, 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait.
+Additionally, nearly 30% of the 421 $\lambda$ estimates between 0.25 and 0.75 were assigned a strength of signal, and because nearly half of the 1,572 values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known.
 \hfill\break
 
+[insert Figure 1 here] \hfill\break
+
+It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood.
+Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion.
+However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored.
 First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation.
@@
 Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$).
-Results were similar when $\lambda$ was co-estimated with regression parameters in PGLS regression (Figure 2).
+Results were similar when $\lambda$ was co-estimated with regression parameters in PGLS regression (Figure 3).
 Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell 2010 [-@Revell2010] (see Supporting Information).
@@
 
-Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented an effect size, obtained from $Kappa$ and its empirical sampling distribution from permutation.
+Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation.
 Here we formalize that suggestion, and find an effect size as:
@@
 \begin{align}
-    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
+    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
 \end{align}
 
-Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002]. \hfill\break
+In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood.
+Note also that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002].
+\hfill\break
 
 To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$).
-Results are found in Figure 3.
+Results are found in Figure 4.
 Here two things are evident.
-First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than to estimates of $Z_\lambda$.
-Second, the precision of $Z_K$ is considerably more stable, as coefficients of variation for the set of $\sigma^2_{Z_K}$ across input levels of phylogenetic signal for the same sample size was an order of magnitude smaller than that observed for $\sigma^2_{Z_\lambda}$ (Figure 3).
-This implies that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
+First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$.
+Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$, as coefficients of variation for the set of $\sigma^2_{Z_K}$ across input levels of phylogenetic signal were an order of magnitude smaller for than was observed for $\sigma^2_{Z_\lambda}$ (Figure 4).
+This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$.
 \hfill\break
 
-[insert Figure 3 here]
+[insert Figure 4 here]
 

### 2020-07-16 Manuscript/OldVsns/Manuscript-3-EKB.tex

--- 2020-07-15 Manuscript/OldVsns/Manuscript-3.tex
@@
 \makeatother
+\usepackage[normalem]{ulem}
+% Avoid problems with \sout in headers with hyperref
+\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 regarding the strength of phylogenetic signal based on \(\lambda\) are
-rather common in the evolutionary literature. For instance, of the 204
-papers published in 2019 that estimated and reported Pagel's \(\lambda\)
-(found in Google.scholar), 40\% interpreted the strength of phylogenetic
-signal for at least one phenotypic trait. Additionally, because nearly
-half of the 1,572 \(\lambda\) values reported were near the limits of
-the parameter (Figure 1), this percentage is even higher, as biological
-interpretation of phylogenetic signal at the limits of \(\lambda\) are
-known. \hfill\break
+rather common in the evolutionary literature; \sout{. For instance,} of
+the 204 papers published in 2019 that estimated and reported Pagel's
+\(\lambda\) (found in Google.scholar), 40\% EXPLICITLY interpreted the
+strength of phylogenetic signal for at least one phenotypic trait.
+\sout{Additionally, because nearly half of the 1,572 \(\lambda\) values
+reported were near the limits of the parameter (Figure 1), this
+percentage is even higher, as biological interpretation of phylogenetic
+signal at the limits of \(\lambda\) are known.} FURTHERMORE, THIS
+PERCENTAGE IS EVEN HIGHER WHEN CONSIDERING THE IMPLICIT BIOLOGICAL
+INTERPRETATION OF PHYLOGENETIC SIGNAL FOR \(\lambda\)s NEAR 0 OR 1, AS
+IS THE CASE FOR NEARLY HALF OF THE 1,572 REPORTED \(\lambda\) VALUES
+(FIGURE 1). \hfill\break
 
@@
 date, has not been verified. Thus, despite widespread use of Pagel's
-(1999) \(\lambda\) in macroevolutionary studies, at present, we still
-lack a general understanding of the precision with which \(\lambda\) can
-estimate levels of phylogenetic signal in phenotypic datasets.
-\hfill\break
+(1999) \(\lambda\) in macroevolutionary studies, at present, we
+\sout{still} lack a general understanding of the precision with which
+\(\lambda\) can estimate levels of phylogenetic signal in phenotypic
+datasets. \hfill\break
 
@@
 \(\lambda\) must be a statistical effect size. Effect sizes are a
-measure the magnitude of a statistical effect in data, represented on a
-common scale (Glass 1976; Cohen 1988). Effect sizes have widespread use
-in many areas of the quantiative sciences, as they represent measures
-that may be readily summarized across datasets as in meta-analysis
-(Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster 1995), or
-compared among datasets (e.g., Adams and Collyer 2016, 2019a).
-Unfortunatley, not all model parameters and test statistics are effect
-sizes, and thus many summary measures must first be converted to
-standardized units (i.e., an effect size) for meaningful comparison (see
-Rosenthal 1994). As a consequence, it follows that only if \(\lambda\)
-is a statistical effect size can comparisons of estimates across
-datasets be interpretable. For the case of \(\lambda\), this has not yet
-been explored. \hfill\break
+measure OF the magnitude of a statistical effect in data, represented on
+a common scale (Glass 1976; Cohen 1988). Effect sizes have widespread
+use in many areas of the quantiative sciences, as they represent
+measures that may be readily summarized across datasets as in
+\sout{meta-analysis} META-ANALYSES (Glass 1976; Hedges and Olkin 1985;
+Arnqvist and Wooster 1995), or compared among datasets (e.g., Adams and
+Collyer 2016, 2019a). \sout{Unfortunatley, not all model parameters and
+test statistics are effect sizes, and thus many summary measures must
+first be converted to standardized units (i.e., an effect size) for
+meaningful comparison (see Rosenthal 1994). As a consequence,} THUS, it
+follows that only \sout{if \(\lambda\) is a statistical effect size} BY
+CONVERTING \(\lambda\) TO AN EFFECT SIZE can comparisons of estimates
+across datasets be interpretable. For the case of \(\lambda\), this has
+not yet been explored. \hfill\break
 
@@
 effect size for measuring the strength of phylogenetic signal in
-phenotypic datasets, and apply the concept to two common measures of
+phenotypic datasets and apply the concept to two common measures of
 phylogenetic signal: \(\lambda\) and \(Kappa\). Through simulations
 across a wide range of conditions, we find that the precision of effect
-sizes based on \(\lambda\) (\(Z_{\lambda}\)) are less reliable than that
-those based on \(Kappa\) (\(Z_K\)), implying that \(Z_K\) is a more
-robust effect size measure. We also propose a two-sample test statistic
-that may be used to compare the strength of phylogenetic signal among
-datasets, and provide an empirical example to demonstrate its use. We
-conclude that estimates of phylogenetic signal using Pagel's \(\lambda\)
-are often inaccurate, and thus interpreting strength of phylogenetic
-signal in phenotypic datasets based on this measure is compromised. By
-contrast, effect sizes obtained from \(Kappa\) hold promise for
-characterizing phylogenetic signal, and for comparing the strength of
-phylogenetic signal across datasets.
+sizes based on \(\lambda\) (\(Z_{\lambda}\)) are less reliable than
+\sout{that} those based on \(Kappa\) (\(Z_K\)), implying that \(Z_K\) is
+a more robust effect size measure. We also propose a two-sample test
+statistic that may be used to compare the strength of phylogenetic
+signal among datasets and provide an empirical example to demonstrate
+its use. We conclude that estimates of phylogenetic signal using Pagel's
+\(\lambda\) are often inaccurate, and thus interpreting strength of
+phylogenetic signal in phenotypic datasets based on this measure is
+compromised. By contrast, effect sizes obtained from \(Kappa\) hold
+promise for characterizing phylogenetic signal, and for comparing the
+strength of phylogenetic signal across datasets.
 
@@
 ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
-phylogenetic signal corresponding reflecting Brownian motion (when
-\(\lambda_{in} =1\)). For each dataset we then estimated phylogenetic
-signal (\(\lambda_{est}\)), and calculated the precision of \(\lambda\)
-using the variance (\(\sigma^2_\lambda\)) across datasets at each input
-level of phylogenetic signal and level of species richness. \hfill\break
+phylogenetic signal \sout{corresponding} reflecting Brownian motion
+(when \(\lambda_{in} =1\)). For each dataset we then estimated
+phylogenetic signal (\(\lambda_{est}\)), and calculated the
+\sout{precision} VARIANCE of \(\lambda\) \sout{using the variance}
+(\(\sigma^2_\lambda\)) across datasets at each input level of
+phylogenetic signal and level of species richness AS A MEASURE OF
+ESTIMATE PRECISION. \hfill\break
 
@@
 parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
-obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
-of phylogenetic signal and level of species richness were then observed.
-\hfill\break
+obtained. WE THEN CALCULATED Precision estimates (\(\sigma^2_\lambda\))
+at each input level of phylogenetic signal and level of species
+richness. \sout{were then observed.} \hfill\break
 
@@
 in terms of accurately characterizing levels of phylogenetic signal, yet
-30\% of these empirical estimates were explicitly interpreted in terms
-of the strength of phylogenetic signal that they represented (i.e.,
-weak, intermediate, strong phylogenetic signal). \hfill\break
+30\% of these MID-RANGE empirical estimates were explicitly interpreted
+in terms of the strength of phylogenetic signal that they represented
+(i.e., weak, intermediate, strong phylogenetic signal). \hfill\break
 
 Finally, when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression, the results of our simulations were quite similar.
-Here, regression parameters (\(\beta\)) were accurately estimated,
-confirming earlier findings of Revell (2010) (see Supporting
+PGLS regression AND ANOVA, the results of our simulations were quite
+similar. Here, regression parameters (\(\beta\)) were accurately
+estimated, confirming earlier findings of Revell (2010) (see Supporting
 Information). However, estimates of phylogenetic signal (\(\lambda\))
-were less precise (Figure 3), and the spread of \(\lambda_{est}\) was
-similar to that observed when \(\lambda\) was estimated for only the
-dependent variable, as in Figure 2. Taken together, these findings
-reveal that \(\lambda_{est}\) does not precisely characterize observed
-levels of phylogenetic signal in phenotypic datasets, and that
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) may be highly inaccurate. \hfill\break
+were less precise (Figure 3; FIGURE S1), and the spread of
+\(\lambda_{est}\) was similar to that observed when \(\lambda\) was
+estimated for only the dependent variable, as in Figure 2. Taken
+together, these findings reveal that \(\lambda_{est}\) does not
+precisely characterize observed levels of phylogenetic signal in
+phenotypic datasets, and that biological interpretations of the strength
+of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
+\hfill\break
 
@@
 \(Z_K\) were an order of magnitude smaller for than for \(Z_\lambda\)
-(Figure 4C,D). This implied that estimates of the strength of
+(Figure 4C\sout{,D}). This implied that estimates of the strength of
 phylogenetic signal were more reliable and robust when using \(Z_K\) as
@@
 determine whether the strength of phylogenetic signal is significantly
-greater in one phenotypic trait as compared with another. As with other
-effect sizes derived from permutation distributions (e.g., Adams and
-Collyer 2016, 2019a), a two-sample test statistic may be found as:
+greater in one phenotypic trait as compared \sout{with} TO another. As
+with other effect sizes derived from permutation distributions (e.g.,
+Adams and Collyer 2016, 2019a), a two-sample test statistic may be
+\sout{found} CALCULATED as:
 
@@
 phylogenetic signal on a common and comparable scale, measured in
-standard deviation units, regardless of the initial units and scale of
-the original phenotypic variables under investigation. This means that
-conceivably, the strength of phylogenetic signal may be compared across
-datsets for phenotypic traits in different units and scale (e.g.,
-behavioral and morphological traits), so long as they are continuous
-variables. Note that this is directly related to the common practice of
-converting data to their standard normal deviates prior to their use in
-multivariate analysis, if the original variables are expressed in
-different units or scales (see Everitt and Hothorn 2011; Legendre and
-Legendre 2012; Adams and Collyer 2019b).
+standard deviation units, regardless of the initial units and ORIGINAL
+scale of the \sout{original} phenotypic variables under investigation.
+This means that conceivably, the strength of phylogenetic signal may be
+compared across datsets for phenotypic traits in different units and
+scale (e.g., behavioral and morphological traits), so long as they are
+continuous variables. Note that this is directly related to the common
+practice of converting data to their standard normal deviates prior to
+their use in multivariate analysis, if the original variables are
+expressed in different units or scales (see Everitt and Hothorn 2011;
+Legendre and Legendre 2012; Adams and Collyer 2019b).
 
@@
 
-Patterns worse with PGLS, though beta still estimated properly.
-Conclusion, lambda not overly useful.
+Patterns \sout{worse} STILL BAD with PGLS, though beta still estimated
+properly. Conclusion, lambda not overly useful.
 

### 2020-07-16 Manuscript/OldVsns/Manuscript-4.tex

--- 2020-07-15 Manuscript/OldVsns/Manuscript-3.tex
@@
 
+Macroevolutionary studies frequently characterize the phylogenetic
+signal in phenotypes, and wish to compare the strength of that signal
+across traits. However, analytical tools for such comparisons have
+largely remained underdeveloped. In this study, we evaluated the
+efficacy of one commonly used parameter (Pagel's \(\lambda\)) to
+estimate the strength of phylogenetic signal in phenotypic traits, and
+evaluate the degree to which \(\lambda\) correctly identifies known
+levels of phylogenetic signal. We find that the precision of \(\lambda\)
+in estimating actual levels of phylogenetic signal is often inaccurate,
+and that biological interpretations of the strength of phylogenetic
+signal based on \(\lambda\) are therefore compromised. We then propose a
+standardized effect size based on \(Kappa\) (\(Z_K\)), which measures
+the strength of phylogenetic signal, and places it on a common scale for
+statistical comparison. Tests based on \(Z_K\) provide a mechanism for
+formally comparing the strength of phylogenetic signal across datasets,
+in much the same manner as effect sizes may be used to summarize
+patterns in quantitative meta-analysis. Our approach extends the
+phylogenetic comparative toolkit to address hypotheses that compare the
+strength of phylogenetic signal various phenotypic traits, even when
+those traits are found in different evolutionary lineages or of in
+different units or scale.
+
 \newpage
@@
 (found in Google.scholar), 40\% interpreted the strength of phylogenetic
-signal for at least one phenotypic trait. Additionally, because nearly
-half of the 1,572 \(\lambda\) values reported were near the limits of
-the parameter (Figure 1), this percentage is even higher, as biological
-interpretation of phylogenetic signal at the limits of \(\lambda\) are
-known. \hfill\break
+signal for at least one phenotypic trait. Further, because nearly half
+of the 1,572 \(\lambda\) values reported were near 0 or 1 (Figure 1)
+where the biological interpretation of \(\lambda\) is known, this
+percentage is even higher.
+
+\hfill\break
 
@@
 in many areas of the quantiative sciences, as they represent measures
-that may be readily summarized across datasets as in meta-analysis
+that may be readily summarized across datasets as in meta-analyses
 (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster 1995), or
@@
 phenotypic datasets, and apply the concept to two common measures of
-phylogenetic signal: \(\lambda\) and \(Kappa\). Through simulations
-across a wide range of conditions, we find that the precision of effect
-sizes based on \(\lambda\) (\(Z_{\lambda}\)) are less reliable than that
-those based on \(Kappa\) (\(Z_K\)), implying that \(Z_K\) is a more
-robust effect size measure. We also propose a two-sample test statistic
-that may be used to compare the strength of phylogenetic signal among
-datasets, and provide an empirical example to demonstrate its use. We
-conclude that estimates of phylogenetic signal using Pagel's \(\lambda\)
-are often inaccurate, and thus interpreting strength of phylogenetic
-signal in phenotypic datasets based on this measure is compromised. By
-contrast, effect sizes obtained from \(Kappa\) hold promise for
-characterizing phylogenetic signal, and for comparing the strength of
-phylogenetic signal across datasets.
+phylogenetic signal: \(\lambda\) and \(Kappa\). Through simulations we
+find that the precision of effect sizes based on \(\lambda\)
+(\(Z_{\lambda}\)) are less reliable than that those based on \(Kappa\)
+(\(Z_K\)), implying that \(Z_K\) is a more robust effect size measure.
+We also propose a two-sample test statistic that may be used to compare
+the strength of phylogenetic signal among datasets, and provide an
+empirical example to demonstrate its use. We conclude that estimates of
+phylogenetic signal using Pagel's \(\lambda\) are often inaccurate, and
+thus interpreting strength of phylogenetic signal in phenotypic datasets
+based on this measure is compromised. By contrast, effect sizes obtained
+from \(Kappa\) hold promise for characterizing phylogenetic signal, and
+for comparing the strength of phylogenetic signal across datasets.
 
@@
 ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
-phylogenetic signal corresponding reflecting Brownian motion (when
+phylogenetic signal reflecting Brownian motion (when
 \(\lambda_{in} =1\)). For each dataset we then estimated phylogenetic
-signal (\(\lambda_{est}\)), and calculated the precision of \(\lambda\)
-using the variance (\(\sigma^2_\lambda\)) across datasets at each input
-level of phylogenetic signal and level of species richness. \hfill\break
+signal (\(\lambda_{est}\)), and calculated the variance of \(\lambda\)
+(\(\sigma^2_\lambda\)) across datasets at each input level of
+phylogenetic signal and level of species richness as an estimate of
+precision. \hfill\break
 
@@
 regression and ANOVA (i.e., \(Y\sim{X}\)). Here, an independent variable
-\(X\) was simulated on each phylogeny under a Brownian motion model of
-evolution (for PGLS regression). For phylogenetic ANOVA, random groups
-(\(X\)) were obtained by simulating a discrete (binary) character on
-each phylogeny. Next, the dependent variable was simulated in such a
+\(X\) was simulated on each rescaled phylogeny under a Brownian motion
+model of evolution (for PGLS regression). For phylogenetic ANOVA, random
+groups (\(X\)) were obtained by simulating a discrete (binary) character
+on each phylogeny. Next, the dependent variable was simulated in such a
 manner as to contain a known relationship with \(X\) plus random error
@@
 to contain phylogenetic signal simulated under a Brownian motion model
-of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
-Revell 2010 for a similar simulation design). The fit of the
-phylogenetic regression was estimated using maximum likelihood, and
-parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
-obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
-of phylogenetic signal and level of species richness were then observed.
-\hfill\break
+of evolution on each rescaled phylogeny:
+\(\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\): (see Revell
+2010 for a similar simulation design). The fit of the phylogenetic
+regression was estimated using maximum likelihood, and parameter
+estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were obtained. We then
+calculated precision estimates (\(\sigma^2_\lambda\)) at each input
+level of phylogenetic signal and level of species richness. \hfill\break
 
@@
 in terms of accurately characterizing levels of phylogenetic signal, yet
-30\% of these empirical estimates were explicitly interpreted in terms
-of the strength of phylogenetic signal that they represented (i.e.,
-weak, intermediate, strong phylogenetic signal). \hfill\break
+30\% of these mid-range empirical estimates were explicitly interpreted
+in terms of the strength of phylogenetic signal that they represented
+(i.e., weak, intermediate, strong phylogenetic signal). \hfill\break
 
 Finally, when \(\lambda\) was co-estimated with regression parameters in
-PGLS regression, the results of our simulations were quite similar.
-Here, regression parameters (\(\beta\)) were accurately estimated,
-confirming earlier findings of Revell (2010) (see Supporting
+PGLS regression and ANOVA, the results of our simulations were quite
+similar. Here, regression parameters (\(\beta\)) were accurately
+estimated, confirming earlier findings of Revell (2010) (see Supporting
 Information). However, estimates of phylogenetic signal (\(\lambda\))
-were less precise (Figure 3), and the spread of \(\lambda_{est}\) was
-similar to that observed when \(\lambda\) was estimated for only the
-dependent variable, as in Figure 2. Taken together, these findings
-reveal that \(\lambda_{est}\) does not precisely characterize observed
-levels of phylogenetic signal in phenotypic datasets, and that
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) may be highly inaccurate. \hfill\break
+were less precise (Figure 3; see also Supporting Information), and the
+spread of \(\lambda_{est}\) was similar to that observed when
+\(\lambda\) was estimated for only the dependent variable, as in Figure
+2. Taken together, these findings reveal that \(\lambda_{est}\) does not
+precisely characterize observed levels of phylogenetic signal in
+phenotypic datasets, and that biological interpretations of the strength
+of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
+\hfill\break
 
@@
 Adams and Collyer (2019b) suggested that the strength of phylogenetic
-signal could be represented as an effect size, based on the \(Kappa\)
+signal could be represented as an effect size based on the \(Kappa\)
 statistic and its empirical sampling distribution from permutation. Here
@@
 \begin{align}
-    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
+    Z_{K}=\frac{log(K_{obs})-\hat\mu_{log(K)}}{\hat\sigma_{log(K)}}
 \end{align}
@@
 and \(\hat\sigma_K\) are the mean and standard deviation of the
-empirical sampling distribution of \(Kappa\) obtained via permutation.
+empirical sampling distribution of \(log(Kappa)\) obtained via
+permutation. Here the logarithm was used because \(Kappa\) takes only
+positive values (\(0\to\infty\)) and its sampling distribution is
+log-normally distributed (for a similar transformation when calculating
+multivariate effect sizes see: Appendix 1 of Collyer and Adams 2018).
 Similarly, an effect size based on \(\lambda\) could be envisioned as:
@@
 In this case, \(\lambda_{obs}\) and \(\hat\sigma_\lambda\) are
-empirically derived using maximum likelihood. Note that under the null
-hypothesis, no phylogenetic signal is expected (Freckleton et al. 2002),
-and thus \(E(\lambda)=0\) under this condition. \hfill\break
+empirically derived using maximum likelihood, as permutation approaches
+have not been developed for evaluating \(\lambda\). Note also that under
+the null hypothesis, no phylogenetic signal is expected (Freckleton et
+al. 2002), and thus \(E(\lambda)=0\) under this condition. \hfill\break
 
@@
 \(Z_K\) were an order of magnitude smaller for than for \(Z_\lambda\)
-(Figure 4C,D). This implied that estimates of the strength of
-phylogenetic signal were more reliable and robust when using \(Z_K\) as
-compared with \(Z_\lambda\). \hfill\break
+(Figure 4C). This implied that estimates of the strength of phylogenetic
+signal were more reliable and robust when using \(Z_K\) as compared with
+\(Z_\lambda\). \hfill\break
 
@@
 determine whether the strength of phylogenetic signal is significantly
-greater in one phenotypic trait as compared with another. As with other
+greater in one phenotypic trait as compared to another. As with other
 effect sizes derived from permutation distributions (e.g., Adams and
-Collyer 2016, 2019a), a two-sample test statistic may be found as:
+Collyer 2016, 2019a), a two-sample test statistic may be calculated as:
 
@@
 require it (see Adams and Collyer 2016, 2019a). \hfill\break
-
-One important advantage of the approach advocated here is that the
-resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
-measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
-2012). As a consequence, the effect sizes represent the strength of
-phylogenetic signal on a common and comparable scale, measured in
-standard deviation units, regardless of the initial units and scale of
-the original phenotypic variables under investigation. This means that
-conceivably, the strength of phylogenetic signal may be compared across
-datsets for phenotypic traits in different units and scale (e.g.,
-behavioral and morphological traits), so long as they are continuous
-variables. Note that this is directly related to the common practice of
-converting data to their standard normal deviates prior to their use in
-multivariate analysis, if the original variables are expressed in
-different units or scales (see Everitt and Hothorn 2011; Legendre and
-Legendre 2012; Adams and Collyer 2019b).
 
@@
 
-\hypertarget{conclusions-and-implications}{%
-\section{Conclusions and
-Implications}\label{conclusions-and-implications}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-1: summary paragraph
+It is common in comparative evolutionary studies to characterize the
+phylogenetic signal in phenotypic traits to determine the extent to
+which shared evolutionary history has generated trait covariation among
+taxa. However, while numerous analytical approaches may be used to
+quantify phylogenetic signal (e.g., Abouheif 1999; Gittleman and Kot
+1990; Pagel 1999; Blomberg et al. 2003; Adams 2014a), methods that
+explicitly measure the strength of phylogenetic signal, or facilitate
+comparisons among datasets, have remained underdeveloped. In this study,
+we evaluated the precision of one common measure, Pagel's \(\lambda\),
+and explored its efficacy for characterizing the strength of
+phylogenetic signal in phenotypic data. Using computer simulations, we
+found that the precision of \(\lambda\) increased with increasing sample
+sizes; a pattern noted previously (Boettiger et al. 2012), and one that
+conformed with parametric statistical theory (Cohen 1988). However, we
+also found that vastly different \(\lambda\) estimates could be obtained
+from data containing the same level of phylogenetic signal, and that
+similar \(\lambda\) estimates may be obtained from data containing
+differing levels of phylogenetic signal. Further, the precision of
+\(\lambda\) varied with the strength of phylogenetic signal, where lower
+precision was observed when in data whose phylogenetic signal was of
+intemediate strength. From these findings we conclude that \(\lambda\)
+is not a reliable indicator of the observed strength of phylogenetic
+signal in phenoytpic datasets, and that biological interpretations of
+the strength of signal based on this parameter may innacurately
+characterize such effects. \hfill\break
 
-2: expand on Lambda.. lambda innacurate, not precise, level of precision
-varies with input physig (worse in mid-range). NEW RESULT. We are first
-to show this. NOTE: pattern is obvious with reflection. Since it is a
-`bounded' parameter estimation should be best at the extremes\ldots{}
-(state this?).. hmm.
+As an alternative, we described a standardized effect size (\(Z\)) for
+assessing the strength of phylogenetic signal. \(Z\) expresses the
+magnitude of phylogenetic signal as a standard normal deviate, which is
+easily interpretable as the strength of phylogenetic signal relative to
+the mean. We applied this concept to both \(\lambda\) and \(Kappa\), and
+found that \(Z_K\) was a better estimate of the strength of phylogenetic
+signal in phenotypic data. First, \(Z_K\) was more precise than
+\(Z_{\lambda}\), and variation in this precision was more consistent
+across the range of input levels of phylogenetic signal. Additionally,
+values of \(Z_K\) more accurately tracked known levels of phylogenetic
+signal, with changes in the actual strength of phylogenetic signal
+reflected in a more linear fashion by concomite changes in the values of
+\(Z_K\). Thus, \(Z_K\) holds promise as a measure of the relative
+strength of phylogenetic signal that reflects the magnitude of this
+effect in phenotypic data. We therefore recommend that future studies
+interested in the strength of phylogenetic signal incorporate \(Z_K\) as
+a statistical measure of this effect. \hfill\break
 
-Patterns worse with PGLS, though beta still estimated properly.
-Conclusion, lambda not overly useful.
-
-3: By contrast, effect size Z-K useful, equally precise across range of
-values. Can be used to characterize the strength of physignal, and
-because robust to input levels, etc. may be used to compare across
-datasets.
-
-Somewhere, recognize that this is somewhat `backwards' from prior
-recommendations where Kappa had somewhat lower performance in terms of
-type I and type II error (which?? I forget). However, recall that those
-studies did not examine the precision of the estimates. Nor was Z-k
-included, because it was not yet invented. So Use of Z-k should make
-good sense here.
-
-Closing paragraph.
-
+Based on the effect size \(Z_K\), we then proposed a two-sample test,
+which provides a statistical means of determining whether the strength
+of phylogenetic signal is greater in one phenotypic trait as compared to
+another. Prior studies have summarized patterns of variation in
+phylogenetic signal across datasets using summary test values, such as
+\(Kappa\) (e.g., Blomberg et al. 2003). However, \(Kappa\) does not
+scale linearly with input levels of phylogenetic signal, and its
+variance increases (i.e., precision decreases) with increasing strength
+of phylogenetic signal (Munkemuller et al. 2012; Diniz-Filho et al.
+2012: see also Supporting Information). Thus, \(Kappa\) should not be
+considered a standardized effect size that measures the strength of
+phylogenetic signal on a common scale. By contrast, converting \(Kappa\)
+to \(Z_K\) (via equation 2) alleviates these concerns, and facilitates
+formal statistical comparisons of the strength of signal across
+datasets. Thus when viewed from this perspective, the approach developed
+here is statistically aligned with other statistical approaches such as
+meta-analysis (sensu Hedges and Olkin 1985; Glass 1976; Arnqvist and
+Wooster 1995), where summary statistics across datasets are converted to
+standardized effect sizes for subsequent `higher order' statistical
+summaries or comparisons. As such our approach enables evolutionary
+biologists to quantitatively examine the relative strength of
+phylogenetic signal across a wide range of phenotypic traits, and thus
+opens the door for future discoveries that inform on how phenotypic
+diversity accumulates in macroevolutionary time across the tree of life.
 \hfill\break
 
-More discussion paragraphs
+One important advantage of the approach advocated here is that the
+resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
+measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
+2012). Thus, \(Z_K\) represents the strength of phylogenetic signal on a
+common and comparable scale -- measured in standard deviation units --
+regardless of the initial units and original scale of the phenotypic
+variables under investigation. This means that the strength of
+phylogenetic signal may be compared across datsets for continuous
+phenotypic traits measured in different units and scale, because those
+units have been standardized through their conversion to \(Z_K\). For
+example, our approach could be utilized to determine whether the
+strength of phylogenetic signal (say, in response to ecological
+differentiation) is stronger in morphological traits (linear traits:
+\(mm\)), physiological traits (metabolic rate: \(\frac{O^2}{min}\)), or
+behavioral traits (aggression: \(\frac{\#{displays}}{second}\)).
+Additionally, our method is capable of comparing the strength of
+phylogenetic signal in traits of different dimensionality, as estimates
+of phylogenetic signal using \(Kappa\) have been generalized for
+multivariate data (\(K_{mult}\): see Adams 2014a). Likewise, as \(Z_K\)
+is a standard deviation unit (which accounts for variation, sample size,
+and the phylogeny), tests based on \(\hat{Z}_{12}\) may be utilized for
+comparing the strength of phylogenetic signal across datasets containing
+a different number of species, and even for phenotypes obtained from
+species in different lineages, because their phylogenetic
+non-independence, and observed variation, are taken into account in the
+generation of the empirical sampling distribution via permutation.
+\hfill\break
+
+** Finally\ldots{} Need Closing paragraph. **
 

### 2020-07-16 Manuscript/OldVsns/Manuscript-8-DCA.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\textbf{Classification:} Biological Sciences -- Evolution \hfill\break

\begin{center}
\textbf{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}
\end{center}

\hfill\break

\begin{center}
\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
\end{center}

\hfill\break

\(^{1}\)Department of Ecology, Evolution, and Organismal Biology, Iowa
State University, Ames, Iowa, USA.

\(^{2}\)Department of Science, Chatham University, Pittsburgh,
Pennsylvania, USA.

\(^{*}\)Correspondence: Dean C. Adams. Department of Ecology, Evolution,
and Organismal Biology, Iowa State University, Ames, Iowa, USA.
515-294-3834.
\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}

\hfill\break

\begin{center}15 July, 2020\end{center}

\hfill\break

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
2553 & 16162\tabularnewline
\bottomrule
\end{longtable}

\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. Here we evaluate the efficacy of one
commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
of phylogenetic signal in phenotypic traits, and evaluate the degree to
which \(\lambda\) correctly identifies known levels of phylogenetic
signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
and that estimates are increasingly skewed at larger and smaller input
levels of phylogenetic signal. Further, the precision of \(\lambda\) in
estimating actual levels of phylogenetic signal is often inaccurate, and
biological interpretations of the strength of phylogenetic signal based
on \(\lambda\) are therefore compromised. We propose a standardized
effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
strength of phylogenetic signal more reliably than does \(\lambda\), and
places that signal on a common scale for statistical comparison. We
develop tests based on \(Z_\kappa\) to provide a mechanism for formally
comparing the strength of phylogenetic signal across datasets, in much
the same manner as effect sizes may be used to summarize patterns in
quantitative meta-analysis. Our approach extends the phylogenetic
comparative toolkit to address hypotheses that compare the strength of
phylogenetic signal between various phenotypic traits, even when those
traits are found in different evolutionary lineages or have different
units or scales. \hfill\break

\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa

\newpage

\hypertarget{significance-statement}{%
\section{Significance Statement}\label{significance-statement}}

Evolutionary biologists wish to quantify and compare the strength of
phylogenetic signal across traits, but analytical tools for these
comparisons are generally lacking. Here we develop a standardized effect
size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
phylogenetic signal on a common statistical scale, and provides a
mechanism for formally comparing the strength of phylogenetic signal
across datasets. Additionally, we find that a commonly used parameter
(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
enables biologists to quantitatively address hypotheses that compare the
strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have
different units or scales.

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates the assumption of independence among trait values that is
common for statistical tests (1, 2). Accounting for this evolutionary
non-independence is the purview of \emph{phylogenetic comparative
methods} (PCMs): a suite of analytical tools that condition trends in
the data on the phylogenetic relatedness of observations (3--12). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (1, 13, 14). Indeed, under numerous
evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (1, 14, 15).
\hfill\break

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets (13, 14, 16--19), and their statistical
properties -- namely type I error rates and statistical power -- have
been investigated to determine under what conditions phylogenetic signal
can be detected (15, 18, 20--25). One of the most widely used methods
for characterizing phylogenetic signal is Pagel's \(\lambda\) (13),
which transforms the lengths of the internal branches of the phylogeny
to improve the fit of data to the phylogeny via maximum likelihood (13,
26). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
(ranging from \(0\rightarrow1\)) which is optimized via log-likelihood
profiling while evaluating the covariation between the dependent and
independent variables, given the phylogeny (13, 26). To infer whether
phylogenetic signal differs from no signal or a Brownian motion model of
evolutionary divergence, the observed model fit using \(\hat\lambda\)
may be statistically compared to that using \(\lambda=0\) or
\(\lambda=1\) via likelihood ratio tests (26--28) or confidence limits
(29). Another widely used measure is Blomberg's \(\kappa\) (14), which
measures the ratio of observed trait variation to the amount expected
under Brownian motion. Blomberg's \(\kappa\) can be treated as a test
statistic by employing a permutation test to generate its sampling
distribution (14, 18) for determining whether significant phylogenetic
signal is present in data. Both \(\lambda\) and \(\kappa\) seem
intuitive to interpret, as a value of \(0\) for both corresponds to no
phylogenetic signal, while a value of \(1\) corresponds to the amount of
phylogenetic signal expected under Brownian motion. Thus, it is tempting
to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
measure the relative strength of phylogenetic signal, providing an
estimate of its magnitude for comparison. \hfill\break

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (30--32). Nonetheless, an important question that has
yet to be considered is whether these statistics are or can be converted
to effect sizes for comparative analyses across datasets? To be
statistics representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
tree are scaled proportionally to the probability that data arise from a
BM process. Given a known \(\lambda\) value used to generate random data
on a tree, we would also expect that the mean of an empirical sampling
distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
the dispersion of \(\hat{\lambda}\) would be largest at intermediate
values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
range of \(\lambda\) with respect to treesize; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). Prior work (20) seems to support some of
these conjectures, based superficially on statistical moments for a
given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
(20)). However, because the ``strength of Brownian motion'' was
simulated as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\) and not prescribed values of \(\lambda\)
(20), interpretation of these patterns is challenging. \hfill\break

By contrast, Blomberg's \(\kappa\), which is positively ubounded, might
be expected to follow a normal distribution (\textbf{add ref}). Thus we
might expect that \(\kappa\) is symmetrically distributed across
different strengths of phylogenetic signal, for any \(\lambda\) used to
generate data. This attribute seemed less reasonable based on the
simulations performed by Münkemüller et al.~(20), as distibutions were
postively skewed, suggesting the Blomberg's \(\kappa\) might not behave
as a statistic that follows a noraml distribution. However, because
their simulations used a weighted combination of simulated phylogenetic
signal strengths, strong inferences are not possible (and distributional
attributes were not the intended result of their simulations). For both
Pagel's \(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical
moments across a range of \(\lambda\) used to generate data would be
valuable for adjudicating the relibalilty of these statistics.
Furthermore, these are statistics that appear to have expected values
that vary with tree size (20), making comparisons across studies
challenging. Therefore, transformation of these statistics into
\(Z\)-scores in the same simulation experiments would allow evaluation
of the efficacy of each statistic to yield effect sizes that could be
used for comparative anlyses. \hfill\break

Here we used simulation experiments to compare the distributional
attributes of \(\hat{\lambda}\) and Blomberg's \(\kappa\), plus their
effect sizes (\(Z\)-scores), across a range of tree size and
phylogenetic signal strength. We find that estimates of
\(\hat{\lambda}\) are increasingly skewed at larger and smaller input
levels of phylogenetic signal, vary widely for a given input value of
\(\lambda\), and that the precision of \(\hat{\lambda}\) is not constant
across its range. By contrast, estimates of \(\kappa\) are more
consistent across tree sizes, and are normally distributed across the
range of input levels of \(\lambda\). We then propose an effect size
based on \(\kappa\) (\(Z_{\kappa}\)), which provides consistent
estimates of the strength of phylogenetic signal across tree sizes, and
facilitates quantitative comparisons of the relative strength of
phylogenetic signal across datasets.

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations revealed that for \(\hat{\lambda}\),
the distributional expectations of a Bernoulli variable were mostly
upheld. First, the mean value of \(\hat{\lambda}\) increased as
\(\lambda\) increased (Fig. 1 black line); the standard deviation of
\(\hat{\lambda}\) was largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig. 1 red line); and the distributions of
\(\hat{\lambda}\) were increasing skewed at extreme values of
\(\lambda\) but rather normal at intermediate values (Fig. 1 blue line).
For small tree sizes, it was also clear that distributions were more
platykurtik at intermeidate values of \(\hat{\lambda}\). Second,
\(\hat{\lambda}\) was a negatively-biased estimate of \(\lambda\) for
small tree sizes, as the mean value was consistently less than the input
\(\lambda\) value. Third, standard deviations of \(\hat{\lambda}\) were
negatively associated with tree size, and for tress of 128 species or
less, \(\hat{\lambda}\) was quite variable, except for cases when
\(\lambda\) was near or equal to \(1\). As such its statistical
properties make it insuitable as an effect size for measuring the
strength of phylogenetic signal. \hfill\break

\textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
stable.} Results for \(\kappa\) demonstrated better statistical
properties. First, mean values increased consistently with \(\lambda\)
irespective of tree size (Fig. 2 black line); the standard deviation of
\(\kappa\) was consistent across tree sizes (Fig. 2 red line); and
\(\kappa\) was normally distributed regardless of \(\lambda\) or tree
size (Fig. 2 blue line). The standard deviation of \(\kappa\) increased
with \(\lambda\), somewhat tracking the mean, but was always less than
the corresponding means. This is perhaps unsurprising, as \(\kappa\) is
bound by 0, and was never large for small values of \(\lambda\).
Finally, distributions of \(\kappa\) displayed slight skewing only at
small tree sizes and when \(\lambda\) was near or equal to 1. This
differs from the results of (20), where skewing appears to be from
combining random values generated independently. \hfill\break

\textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
characterize phylogenetic signal.} We derived effect sizes (Z-scores)
for both \(\lambda\) and \(\kappa\) (see Methods) and evaluated them
relative to known phylogenetic signal. Both \(Z_{\lambda}\) and
\(Z_{\kappa}\) were associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics captured the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they were more strongly associated with tree size than
the actual phylogenetic signal in the data (Fig. 3). By contrast,
\(Z_{\kappa}\) was much more consistent accross tree sizes, and
exhibited a much stronger association with phylogenetic signal strength
as compared to tree size (Fig. 3). Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets. \hfill\break

\textbf{A test statistic (\(\hat{Z}_{12}\)) allows meaningful
comparisons across datasets.} We derived a two-sample test statistic
(\(\hat{Z}_{12}\)) for statistically comparing the strength of
phylogenetic signal across datasets (see Methods). To demonstrate its
utility, we compared \(Z_{\kappa}\) for two ecologically-relevant traits
in plethodontid salamander (Fig. 4): surface area to volume ratios
(SA:V) and relative body width (\(\frac{BW}{SVL}\)) (33, 34). While both
contained significant phylogenetic signal, tests based on
\(\hat{Z}_{12}\) revealed that the degree of phylogenetic signal was
significantly stronger in SA:V (Fig. 4). Biologically, this observation
corresponds with the fact that tropical species -- which form a
monophyletic group within plethodontids -- display greater variation in
SA:V which covaries with disparity in their climatic niches (34).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\textbf{Derivation of effect sizes.} Statistically, a standardized
effect size may be found as:

\begin{align}
    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
\end{align}

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (35--37). Typically, \(\theta_{obs}\) and
\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
obtained from the distribution of \(\theta\) derived from parametric
theory. However, recent advances in resampling theory (38--41) have
shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
an empirical sampling distribution of \(\theta\) obtained from
permutation procedures. \hfill\break

Formalizing the suggestion of Adams and Collyer (42), an effect size for
\(\kappa\) may be found as:

\begin{align}
    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
\end{align}

where \(\kappa_{obs}\) is the observed phylogenetic signal, and
\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
deviation of the empirical sampling distribution of \(\kappa\) obtained
via permutation. The empirical sampling distribution of \(\kappa\) is
first transformed via Box-Cox to better adhere to the assumption of
normality. \hfill\break

For \(\lambda\) deriving an effect size is more challenging, it does not
have a sampling distribution from which the standard error (and thus
confidence intervals) may be obtained. Thus, its confidence intervals
are generated for the values of \(\lambda\) that intersect the
log-likehihood profile for corresponding percentiles of the \(\chi^2\)
distribution used to compare the compare the putative appropriate model
to a null model with \(\lambda = 0\) {[}\textbf{add ref: MLC thinks
Boettiger paper?}{]}. As such, an effect size for \(\lambda\) may be
found as:

\begin{align}
   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
\end{align}

where, \(\hat{\lambda}\) is the maximized likelihood value of
\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
statistic for the value. Note that alternative formulations could be
envisioned: \(Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}\), where
\(d\) is a binary value \((-1,1)\) to indicate a direction based on
whether \(\hat{\lambda}\) is below or above a critical value of
\(\lambda\), for a quantile from a \(\chi^{2}\) distribution at a
probability of 0.5. However, preliminary investigations found that
mapping \(\chi^2_{\hat{\lambda}}\) values in this manner did not produce
effect sizes that were symmetrical about \(Z = 0\), as the mapping was
not linear and the log-likelihood profiles can be rather flat for small
trees. \hfill\break

\textbf{Derivation of two-sample test statistic.} To compare the
strength of phylogenetic signal across datasets, a two-sample test
statistic may be calculated as:

\begin{align}
  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
\end{align}

where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
side of the equation illustrates that if \(Z_\kappa\) has already been
calculated for two sampling distributions as in equation 2, the sampling
distributions have unit variance for each of the \(Z_\kappa\)
statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
obtained from a standard normal distribution. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (39, 41). \hfill\break

\textbf{Simulations.} Simulations were conducted by generating
pure-birth phylogenies at each of six different tree sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylognetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, continuous traits were then
simulated on each phylogeny under a BM model of evolution. For each set
of 100 trees we measured the mean values of \(\hat{\lambda}\) and
\(\kappa\), their standard deviation, and calculated the Shapiro-Wilk
\(W\) statistic as a departure from normality (symmetry). For the
latter, a value of \(1.0\) indicates normally distributed values, while
departures from \(1.0\) indicate skewness. Simulations were then
repeated for both bananced and pectinate trees, which yielded
qualitatively similar results (see Supporting Information). Trees
containing polytomies, and an evaluation of \(\hat{\lambda}\) from
models of linear regression and phylogenetic ANOVA, were also
investigated, and results were qualitatively similar to those reported
above (see Supporting Information). \hfill\break

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
individuals of 305 species (2,781 and 3,371 respectively), from which
species means were obtained (33, 34). A time-dated molecular phylogeny
for the group (43) was then pruned to match the species in the
phenotypic dataset. The phylogenetic signal in each trait was then
characterized using \(\kappa\), which was converted to its effect size
(\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (44, 45).

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and
DBI-1902694 (to M.L.C.).

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
\setlength{\parskip}{8pt} \noindent

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Felsenstein1985}{}%
1. J. Felsenstein, Phylogenies and the comparative method.
\emph{American Naturalist} \textbf{125}, 1--15 (1985).

\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
2. P. H. Harvey, M. D. Pagel, \emph{The comparative method in
evolutionary biology} (Oxford University Press, 1991).

\leavevmode\hypertarget{ref-Grafen1989}{}%
3. A. Grafen, The phylogenetic regression. \emph{Philosophical
Transactions of the Royal Society of London B, Biological Sciences}
\textbf{326}, 119--157 (1989).

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
4. T. J. Garland, A. R. Ives, Using the past to predict the present:
Confidence intervals for regression equations in phylogenetic
comparative methods. \emph{American Naturalist} \textbf{155}, 346--364
(2000).

\leavevmode\hypertarget{ref-Rohlf2001}{}%
5. F. J. Rohlf, Comparative methods for the analysis of continuous
variables: Geometric interpretations. \emph{Evolution} \textbf{55},
2143--2160 (2001).

\leavevmode\hypertarget{ref-ButlerKing2004}{}%
6. M. A. Butler, A. A. King, Phylogenetic comparative analysis: A
modeling approach for adaptive evolution. \emph{American Naturalist}
\textbf{164}, 683--695 (2004).

\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
7. E. P. Martins, T. F. Hansen, Phylogenies and the comparative method:
A general approach to incorporating phylogenetic information into the
analysis of interspecific data. \emph{American Naturalist} \textbf{149},
646--667 (1997).

\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
8. B. C. O'Meara, C. Ane, M. J. Sanderson, P. C. Wainwright, Testing for
different rates of continuous trait evolution using likelihood.
\emph{Evolution} \textbf{60}, 922--933 (2006).

\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
9. L. J. Revell, L. J. Harmon, Testing quantitative genetic hypotheses
about the evolutionary rate matrix for continuous characters.
\emph{Evolutionary Ecology Research} \textbf{10}, 311--331 (2008).

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
10. J. M. Beaulieu, D. C. Jhwueng, C. Boettiger, B. C. O'Meara, Modeling
stabilizing selection: Expanding the ornstein-uhlenbeck model of
adaptive evolution. \emph{Evolution} \textbf{66}, 2369--2383 (2012).

\leavevmode\hypertarget{ref-Adams2014b}{}%
11. D. C. Adams, A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution} \textbf{68}, 2675--2688 (2014).

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
12. D. C. Adams, M. L. Collyer, Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution} \textbf{72}, 1204--1215 (2018).

\leavevmode\hypertarget{ref-Pagel1999}{}%
13. M. D. Pagel, Inferring the historical patterns of biological
evolution. \emph{Nature} \textbf{401}, 877--884 (1999).

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
14. S. P. Blomberg, T. Garland, A. R. Ives, Testing for phylogenetic
signal in comparative data: Behavioral traits are more labile.
\emph{Evolution} \textbf{57}, 717--745 (2003).

\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
15. L. J. Revell, L. J. Harmon, D. C. Collar, Phylogenetic signal,
evolutionary process, and rate. \emph{Systematic Biology} \textbf{57},
591--601 (2008).

\leavevmode\hypertarget{ref-Abouheif1999}{}%
16. E. Abouheif, A method for testing the assumption of phylogenetic
independence in comparative data. \emph{Evolutionary Ecology Research}
\textbf{1}, 895--909 (1999).

\leavevmode\hypertarget{ref-Gittleman1990}{}%
17. J. L. Gittleman, M. Kot, Adaptation: Statistics and a null model for
estimating phylogenetic effects. \emph{Systematic Zoology} \textbf{39},
227--241 (1990).

\leavevmode\hypertarget{ref-Adams2014a}{}%
18. D. C. Adams, A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. \emph{Systematic Biology} \textbf{63}, 685--697 (2014).

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
19. C. P. Klingenberg, N. A. Gidaszewski, Testing and quantifying
phylogenetic signals and homoplasy in morphometric data.
\emph{Systematic biology} \textbf{59}, 245--261 (2010).

\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
20. T. Münkemüller\emph{et al.}, How to measure and test phylogenetic
signal. \emph{Methods in Ecology and Evolution} \textbf{3}, 743--756
(2012).

\leavevmode\hypertarget{ref-Pavoine2012}{}%
21. S. Pavoine, C. Ricotta, Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics.
\emph{Evolution: International Journal of Organic Evolution}
\textbf{67}, 828--840 (2012).

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
22. J. A. F. Diniz-Filho, T. Santos, T. F. Rangel, L. M. Bini, A
comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. \emph{Genetics and Molecular Biology}
\textbf{35}, 673--679 (2012).

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
23. R. Molina-Venegas, M. A. Rodríguez, Revisiting phylogenetic signal;
strong or negligible impacts of polytomies and branch length
information? \emph{BMC evolutionary biology} \textbf{17}, 53 (2017).

\leavevmode\hypertarget{ref-Revell2010}{}%
24. L. J. Revell, Phylogenetic signal and linear regression on species
data. \emph{Methods in Ecology and Evolution} \textbf{1}, 319--329
(2010).

\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
25. C. Boettiger, G. Coop, P. Ralph, Is your phylogeny informative?
Measuring the power of comparative methods. \emph{Evolution}
\textbf{67}, 2240--2251 (2012).

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
26. R. P. Freckleton, P. H. Harvey, M. Pagel, Phylogenetic analysis and
comparative data: A test and review of evidence. \emph{American
Naturalist} \textbf{160}, 712--726 (2002).

\leavevmode\hypertarget{ref-Cooper2010}{}%
27. N. Cooper, W. Jetz, R. P. Freckleton, Phylogenetic comparative
approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology} \textbf{23}, 2529--2539 (2010).

\leavevmode\hypertarget{ref-Bose2019}{}%
28. R. Bose, B. R. Ramesh, R. Pélissier, F. Munoz, Phylogenetic
diversity in the western ghats biodiversity hotspot reflects
environmental filtering and past niche diversification of trees.
\emph{Journal of Biogeography} \textbf{46}, 145--157 (2019).

\leavevmode\hypertarget{ref-Vandelook2019}{}%
29. F. Vandelook\emph{et al.}, Nectar traits differ between pollination
syndromes in balsaminaceae. \emph{Annals of Botany} \textbf{124},
269--279 (2019).

\leavevmode\hypertarget{ref-DeMeester2019}{}%
30. G. De Meester, K. Huyghe, R. Van Damme, Brain size, ecology and
sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society} \textbf{126}, 381--391 (2019).

\leavevmode\hypertarget{ref-Pintanel2019}{}%
31. P. Pintanel, M. Tejedo, S. R. Ron, G. A. Llorente, A. Merino-Viteri,
Elevational and microclimatic drivers of thermal tolerance in andean
pristimantis frogs. \emph{Journal of Biogeography} \textbf{46},
1664--1675 (2019).

\leavevmode\hypertarget{ref-Su2019}{}%
32. G. Su, S. Villéger, S. Brosse, Morphological diversity of freshwater
fishes differs between realms, but morphologically extreme species are
widespread. \emph{Global ecology and biogeography} \textbf{28}, 211--221
(2019).

\leavevmode\hypertarget{ref-Baken2019}{}%
33. E. K. Baken, D. C. Adams, Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution} \textbf{9}, 7005--7016 (2019).

\leavevmode\hypertarget{ref-Baken2020}{}%
34. E. K. Baken, L. E. Mellenthin, D. C. Adams, Macroevolution of
desiccation‐related morphology in plethodontid salamanders as inferred
from a novel surface area to volume ratio estimation approach.
\emph{Evolution} \textbf{74}, 476--486 (2020).

\leavevmode\hypertarget{ref-Glass1976}{}%
35. G. V. Glass, Primary, secondary, and meta-analysis of research.
\emph{Educational Researcher} \textbf{5}, 3--8 (1976).

\leavevmode\hypertarget{ref-Cohen1988}{}%
36. J. Cohen, \emph{Statistical power analysis for the behavioral
sciences} (Routledge, 1988).

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
37. R. Rosenthal, ``The handbook of research synthesis'' in L. V. Cooper
H Hedges, Ed. (Russell Sage Foundation, 1994), pp. 231--244.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
38. M. L. Collyer, D. J. Sekora, D. C. Adams, A method for analysis of
phenotypic change for phenotypes described by high-dimensional data.
\emph{Heredity} \textbf{115}, 357--365 (2015).

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
39. D. C. Adams, M. L. Collyer, On the comparison of the strength of
morphological integration across morphometric datasets. \emph{Evolution}
\textbf{70}, 2623--2631 (2016).

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
40. M. L. Collyer, D. C. Adams, RRPP: An r package for fitting linear
models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution} \textbf{9}, 1772--1779 (2018).

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
41. D. C. Adams, M. L. Collyer, Comparing the strength of modular
signal, and evaluating alternative modular hypotheses, using covariance
ratio effect sizes with morphometric data. \emph{Evolution} \textbf{73},
2352--2367 (2019).

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
42. D. C. Adams, M. L. Collyer, Phylogenetic comparative methods and the
evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
Evolution, and Systematics} \textbf{50}, 405--425 (2019).

\leavevmode\hypertarget{ref-Bonett2017}{}%
43. R. M. Bonett, A. L. Blair, Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, U.S.A.} \textbf{114}, 9936--9941
(2017).

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
44. D. C. Adams, E. Otárola-Castillo, Geomorph: An r package for the
collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution} \textbf{4}, 393--399 (2013).

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
45. D. C. Adams, M. L. Collyer, A. Kaliontzopoulou, Geomorph: Software
for geometric morphometric analyses. R package version 3.3.1. (2020).

\newpage

Real figures will have 6 panels for lambda and kappa and probably only a
subset for z. Simulations need to be re-run.

\includegraphics[width=0.95\linewidth]{new.fig.1.temp}

\singlespacing \textbf{Figure 1}. Temporary Fig. 1 showing lambda
patterns

\newpage

\includegraphics[width=0.95\linewidth]{new.fig.2.temp}

\singlespacing \textbf{Figure 2}. Temporary Fig. 2 showing kappa
patterns

\newpage

\includegraphics[width=0.95\linewidth]{new.fig.3.temp}

\singlespacing \textbf{Figure 3}. Temporary Fig. 3 showing z patterns

\newpage

\includegraphics[width=0.95\linewidth]{new.fig.4}

\singlespacing \textbf{Figure 4}. (A) Linear measures for relative body
size, and regions of the body used to estimate surface area to volume
(SA:V) ratios. (B) Permutation distributions of phylogenetic signal for
SA:V and \(\frac{BW}{SVL}\), with observed values shown as vertical
bars. (C) Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\),
with their 95\% confidence intervals (CI not standardized by
\(\sqrt(n)\)).

\end{document}

### 2020-07-16 Manuscript/OldVsns/Manuscript.Rmd

--- 2020-04-08 Manuscript/Manuscript.tex
@@
-\documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+---
+title:
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+output: 
+  pdf_document:
+    fig_caption: yes
+    number_sections: no  #note: change to 'no' for most journals
+    keep_tex: true
+    pandoc_args:
 
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\usepackage{setspace}\doublespacing
-\usepackage{lineno}\linenumbers
-% https://github.com/rstudio/rmarkdown/issues/337
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
+csl: evolution.csl
+bibliography: bibliography.bib
 
-% https://github.com/rstudio/rmarkdown/pull/252
-\usepackage{titling}
-\setlength{\droptitle}{-2em}
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
 
-\pretitle{\vspace{\droptitle}\centering\huge}
-\posttitle{\par}
-
-\preauthor{\centering\large\emph}
-\postauthor{\par}
-
-\predate{\centering\large\emph}
-\postdate{\par}
-
-\date{}
-
-\begin{document}
-
-\section{Pagel's Lambda Estimates are Often
-Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}
+---
+# Pagel's Lambda Estimates are Often Inaccurate
 
@@
 
-\textbf{Keywords}: Pagel's lambda, phylogenetic signal \hfill\break
+**Keywords**: Pagel's lambda, phylogenetic signal \hfill\break
 
-\textbf{Short Title}: Inaccuracies in Pagel's Lambda \hfill\break
+**Short Title**: Inaccuracies in Pagel's Lambda \hfill\break
 
-\section{Abstract}\label{abstract}
+```{r setup, include=FALSE}
+  library(knitr)
+  knitr::opts_chunk$set(echo = TRUE)
+```
 
-\{conclusion holds: interpreting the regression is not appreciably
-different (in terms of slopes and f values)\}
+# Abstract
+
+{conclusion holds: interpreting the regression is not appreciably different (in terms of slopes and f values)}
 
@@
 
-\section{Introduction}\label{introduction}
-
-Investigating macroevolutionary patterns requires a phylogenetic
-approach as species are non-independent by nature of their shared
-ancestry. Since the first appropriate method was introduced by
-Felsenstein (phylogenetic independent contrasts; Felsenstein (1985)),
-dozens of other methods have been developed and applied to increasingly
-complex questions in macroevolutionary biology (e.g. Grafen (1989);
-Harvey and Pagel (1991); Rohlf (2001); ({\textbf{???}})). A particularly
-useful measure in this field is phylogenetic signal, a quantification of
-a trait's correspondence with the clade's phylogenic history.
-Understanding the degree of phylogenetic signal present in a dataset is
-paramount and identifies the mode under which a trait has evolved; high
-measures of phylogenetic signal indicate a Brownian motion process,
-whereas lower levels of phylogenetic signal indicate natural selection
-or some other evolutionary force has influenced the traits evolutionary
-history. \hfill\break
-
-Several approaches to quantify phylogenetic signal exist. For continuous
-data, the most common parameters used in the literature include Pagel's
-lambda ({\textbf{???}}) and Blomberg's kappa ({\textbf{???}}). Pagel's
-lambda has the advantage of being nested in the likelihood framework and
-thus has also been utilized to simultaneously estimate and account for
-phylogenetic signal while doing phylogenetic regressions or ANOVAs. At
-the development of these estimation methods, clear recommendations on
-the use and interpretation of this parameter were made based on data
-simulations and error evaluations (Revell 2010). However, the accuracy
-of the lambda estiamtion methods have not been fully evaluated, and
-consequently many studies have substantially misused these methods.
+# Introduction
+Investigating macroevolutionary patterns requires a phylogenetic approach as species are non-independent by nature of their shared ancestry.
+Since the first appropriate method was introduced by Felsenstein (phylogenetic independent contrasts; @Felsenstein1985), dozens of other methods have been developed and applied to increasingly complex questions in macroevolutionary biology (e.g.
+@Grafen1989; @HarveyPagel1991; @Rohlf2001; @AdamsNason2019).
+A particularly useful measure in this field is phylogenetic signal, a quantification of a trait's correspondence with the clade's phylogenic history.
+Understanding the degree of phylogenetic signal present in a dataset is paramount and identifies the mode under which a trait has evolved; high measures of phylogenetic signal indicate a Brownian motion process, whereas lower levels of phylogenetic signal indicate natural selection or some other evolutionary force has influenced the traits evolutionary history.
 \hfill\break
 
-An earlier study ({\textbf{???}}), breifly addressed this topic by
-showing how uninformative smaller phylogenies are when estimating
-various evolutionary parameters, such as Pagel's lambda. However, it
-remains unknown the degree to which lambda estimates accurately
-represent degree of phylogenetic signal, and thus we fully explicate the
-extent to which Pagel's lambda estimation methods can and have been
-misused in recent macroevolutionary studies. Here we take a
-comprehensive approach to demonstrate the scenarios under which
-estimated Pagel's lambdas accurately reflect known lambdas. We then
-demonstrate the effect of these, at times, dubious estimation methods on
-significance testing when used in a PGLS framework. Finally, we perform
-a meta-analysis of research published in 2019 employing Pagel's lambda
-estimation methods to demonstrate the uses and misuses of this method in
-the literature. This investigation provides evidence for the
-circumstances under which estimating Pagel's lambda can be informative,
-and which interpretations or uses are misleading with the hope of moving
-the field towards more appropriate employment of phylogenetic
-comparative methods.
-
-\section{Methods and Results}\label{methods-and-results}
-
-\subsection{\texorpdfstring{\emph{Simulated
-trait}}{Simulated trait}}\label{simulated-trait}
-
-To assess the accuracy of Pagel's lambda estimations, we simulated
-pure-birth phylogenies of variable size, ranging in tip number from 32
-to 1024. We then scaled the simulated phylogenies by lambda values
-ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree
-size) with which we generated trait data with known lambdas by
-simulating a continuous variable on each scaled phylogeny under Brownian
-motion. We then estimated lambda values from these data using
-phylogenetic generalized least squares (PGLS) to compare against the
-known lambda values. \hfill\break
-
-To visualize the accuracy of the lambda estimation methods, we first
-plotted known lambdas (input lambdas) against the estimated lambdas
-(Figure 1). This plot demonstrates the rampant inaccuracy of estimating
-lambdas on phylogenies with fewer than 200 tips, as the spread of data
-in the upper panels in Figure 1 is remarkably wide. We also see that the
-widest spread of estimated lambdas is observed near the center of each
-plot corresponding to intermediate values of known lambda. Lastly, we
-see a slight tendency for the PGLS estimation methods to underestimate
-lambda, especially for known lambda values below 0.5 as can be seen by
-the numerous data points along the x axis for the smaller phylogenies
-analyzed (n tips \textless{} 200). \hfill\break
-
-{[}insert Figure 1 here{]}
-
-\subsection{\texorpdfstring{\emph{Simulated ANOVA and
-Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}
-
-To measure the statistical performance of PGLS lambda estimation methods
-when applied to ANOVA and regression analyses, we used the above
-generated data (independent variables) to simulate a second set of trait
-variables (depedendent variables) across the range of correlation
-strengths (betas ranging from 0 to 1 at intervales of 0.25). We then
-used PGLS to estimate phylogenetic signal of the dependent variable and
-the slope coefficient for the regression between the dependent and
-independent variables. We also calculated F- and p-values from the
-regression analyses. Finally, we again fit the dependent variable to the
-independent variable while holding the lambda value at 0 to approximate
-an ordinary least squares (OLS) apprach. \hfill\break
-
-We compared the estimated slope coefficients across variable input beta
-values to assess the ability of PGLS to identify significant
-correlations (Figure 2). All distribution groups center around the 1:1
-relationship between input beta values and estimated slopes. However the
-variance around that mean is substantial, with estimates ranging from
-approximately -2 to 4 in datasets with 32 taxa and a known relationship
-of (beta = 1). This variance of the estimated slopes only becomes
-reasonable when phylogenies have over 200 tips, similarly to what we saw
-in the first simulation analysis in this study. \hfill\break
-
-{[}insert Figure 2 here{]} \hfill\break
-
-Surprisingly, the estimated lambda values of the dependent variable do
-not correspond with input lambda values that characterize the
-independent variable (Figure S1). This disconnect then resulted in no
-appreciable difference of slope estimates across input lambda values
-(Figure S2), nor do we see substantial differences in slope estimates
-between the PGLS and OLS analyses (results not shown). Thus, we show
-that phylogenetic signal present in one variable does not translate to
-phylogenetic signal in a second, even highly correlated, dependent
-variable. \hfill\break
-
-Scaling and trait generation procedures for all simulation methods were
-repeated with symmetrical and ladder phylogenies of variable size.
-Results generated using these variable phylogeny shapes were consistent
-with the pure-birth phylogeny results presented above and can be found
-in the Supplemental Materials. All analyses were performed in R v3.6.0
-({\textbf{???}}) using the packages \texttt{geiger} ({\textbf{???}}) and
-\texttt{caper} ({\textbf{???}}), and the corresponding scripts can be
-found in the Supplemental Materials.
-
-\subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
-Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}
-
-To understand the extent of this problematic estimation method in
-application, we performed a meta-analysis of studies published in 2019
-that cited Pagel's 1997 manuscript ({\textbf{???}}). The list of
-manuscripts was compiled through Google Scholar on Jan 23, 2020 and
-totaled 341 manuscripts. For each study, we extracted any published
-lambda estimates, along with the size of the phylogeny used in the
-analysis. We also noted whether authors reported confidence intervals,
-significance tests assessing difference of the lambda estimate from 0 or
-1, and whether authors interpreted biological meaning from the magnitude
-of the estimated lambda. For studies that reported more than one lambda
-estimate, we also noted if the authors compared the lambdas against one
-another, and whether that was accompanied with an appropriate
-statistical test between the estimated lambda values. \hfill\break
-
-We found 182 manuscripts from 2019 that estimated and reported Pagel's
-lambda values using PGLS methods. These papers averaged 8.527 lambda
-values, ranging from a single lambda estimate up to 71 estimated
-lambdas. Almost exactly half of the published lambda estimates were
-either below 0.05 (25.32\%) or above 0.9 (24.74\%; Figure 3). 73.32\% of
-the published lambdas were estimated using phylogenies with fewer than
-200 tips, and 348 lambda estimates (8.57\% of all published estimates)
-came from phylogenies with fewer than 30 tips. \hfill\break
-
-{[}insert Figure 3 here{]} \hfill\break
-
-Many of the reviewed manuscripts liberally interpreted the magnitude of
-the estimate lambda, using phrases such as ``strong'' or ``weak''
-phylogentic signal when statistically, all that was clear was a
-difference between the estimated lambda and 0 or 1 respectively. We
-estimated that about 20.49\% of the manuscripts revealed some sort of
-biological interpretation of the magnitude of estimated phylogenetic
-signal that overreached the statistical findings. We also identified
-seven manuscripts as having inappropriately interpreted differences in
-lambda values, indicating that some traits had stronger or weaker signal
-than other traits without the appropriate statistical tests.
+Several approaches to quantify phylogenetic signal exist.
+For continuous data, the most common parameters used in the literature include Pagel's lambda [@Pagel1997] and Blomberg's kappa [@Blomberg1900].
+Pagel's lambda has the advantage of being nested in the likelihood framework and thus has also been utilized to simultaneously estimate and account for phylogenetic signal while doing phylogenetic regressions or ANOVAs.
+At the development of these estimation methods, clear recommendations on the use and interpretation of this parameter were made based on data simulations and error evaluations [@Revell2010].
+However, the accuracy of the lambda estiamtion methods have not been fully evaluated, and consequently many studies have substantially misused these methods.
 \hfill\break
 
-As is evidenced by macroevolutionary papers published in 2019 papers,
-Pagel's lambda estimation methods are often misused and
-over-interpretted. Despite the urging of Boettiger and colleagues to
-publish confidence intervals with all lambda parameter estimates, only
-18\% of papers published in 2019 do so.
+An earlier study [@BoettigerEtAl2012], breifly addressed this topic by showing how uninformative smaller phylogenies are when  estimating various evolutionary parameters, such as Pagel's lambda.
+However, it remains unknown the degree to which lambda estimates accurately represent degree of phylogenetic signal, and thus we fully explicate the extent to which Pagel's lambda estimation methods can and have been misused in recent macroevolutionary studies.
+Here we take a comprehensive approach to demonstrate the scenarios under which estimated Pagel's lambdas accurately reflect known lambdas.
+We then demonstrate the effect of these, at times, dubious estimation methods on significance testing when used in a PGLS framework.
+Finally, we perform a meta-analysis of research published in 2019 employing Pagel's lambda estimation methods to demonstrate the uses and misuses of this method in the literature.
+This investigation provides evidence for the circumstances under which estimating Pagel's lambda can be informative, and which interpretations or uses are misleading with the hope of moving the field towards more appropriate employment of phylogenetic comparative methods.
 
-\section{Discussion}\label{discussion}
+# Methods and Results
 
-{[}This part is obviously not written yet{]} \hfill\break
+## *Simulated trait*
 
-General conclusions :Using the estimated lambda values from pgls are not
-useful. The questions of whether or not signal exists is appropriate,
-but inferring more from lambda \emph{magnitude} is inappropriate.
+To assess the accuracy of Pagel's lambda estimations, we simulated pure-birth phylogenies of variable size, ranging in tip number from 32 to 1024.
+We then scaled the simulated phylogenies by lambda values ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree size) with which we generated trait data with known lambdas by simulating a continuous variable on each scaled phylogeny under Brownian motion.
+We then estimated lambda values from these data using phylogenetic generalized least squares (PGLS)  to compare against the known lambda values.
+\hfill\break
+
+To visualize the accuracy of the lambda estimation methods, we first plotted known lambdas (input lambdas) against the estimated lambdas (Figure 1).
+This plot demonstrates the rampant inaccuracy of estimating lambdas on phylogenies with fewer than 200 tips, as the spread of data in the upper panels in Figure 1 is remarkably wide.
+We also see that the widest spread of estimated lambdas is observed near the center of each plot corresponding to intermediate values of known lambda.
+Lastly, we see a slight tendency for the PGLS estimation methods to underestimate lambda, especially for known lambda values below 0.5 as can be seen by the numerous data points along the x axis for the smaller phylogenies analyzed (n tips < 200).
+\hfill\break
+
+[insert Figure 1 here]
+
+## *Simulated ANOVA and Regressions*
+
+To measure the statistical performance of PGLS lambda estimation methods when applied to ANOVA and regression analyses, we used the above generated data (independent variables) to simulate a second set of trait variables (depedendent variables) across the range of correlation strengths (betas ranging from 0 to 1 at intervales of 0.25).
+We then used PGLS to estimate phylogenetic signal of the dependent variable and the slope coefficient for the regression between the dependent and independent variables.
+We also calculated F- and p-values from the regression analyses.
+Finally, we again fit the dependent variable to the independent variable while holding the lambda value at 0 to approximate an ordinary least squares (OLS) apprach.
+\hfill\break
+
+We compared the estimated slope coefficients across variable input beta values to assess the ability of PGLS to identify significant correlations (Figure 2).
+All distribution groups center around the 1:1 relationship between input beta values and estimated slopes.
+However the variance around that mean is substantial, with estimates ranging from approximately -2 to 4 in datasets with 32 taxa and a known relationship of (beta = 1).
+This variance of the estimated slopes only becomes reasonable when phylogenies have over 200 tips, similarly to what we saw in the first simulation analysis in this study.
+\hfill\break
+
+[insert Figure 2 here] \hfill\break
+
+Surprisingly, the estimated lambda values of the dependent variable do not correspond with input lambda values that characterize the independent variable (Figure S1).
+This disconnect then resulted in no appreciable difference of slope estimates across input lambda values (Figure S2), nor do we see substantial differences in slope estimates between the PGLS and OLS analyses (results not shown).
+Thus, we show that phylogenetic signal present in one variable does not translate to phylogenetic signal in a second, even highly correlated, dependent variable.
+\hfill\break
+
+Scaling and trait generation procedures for all simulation methods were repeated with symmetrical and ladder phylogenies of variable size.
+Results generated using these variable phylogeny shapes were consistent with the pure-birth phylogeny results presented above and can be found in the Supplemental Materials.
+All analyses were performed in R v3.6.0 [@RCoreTeam] using the packages `geiger` [@HarmonEtAl2019] and `caper` [@Orme2011], and the corresponding scripts can be found in the Supplemental Materials.
+
+## *Meta-Analysis of Empirical Results*
+
+To understand the extent of this problematic estimation method in application, we performed a meta-analysis of studies published in 2019 that cited Pagel's 1997 manuscript [@Pagel1997].
+The list of manuscripts was compiled through Google Scholar on Jan 23, 2020 and totaled 341 manuscripts.
+For each study, we extracted any published lambda estimates, along with the size of the phylogeny used in the analysis.
+We also noted whether authors reported confidence intervals, significance tests assessing difference of the lambda estimate from 0 or 1, and whether authors interpreted biological meaning from the magnitude of the estimated lambda.
+For studies that reported more than one lambda estimate, we also noted if the authors compared the lambdas against one another, and whether that was accompanied with an appropriate statistical test between the estimated lambda values.
+\hfill\break
+
+We found 182 manuscripts from 2019 that estimated and reported Pagel's lambda values using PGLS methods.
+These papers averaged 8.527 lambda values, ranging from a single lambda estimate up to 71 estimated lambdas.
+Almost exactly half of the published lambda estimates were either below 0.05 (25.32%) or above 0.9 (24.74%; Figure 3).
+73.32% of the published lambdas were estimated using phylogenies with fewer than 200 tips, and 348 lambda estimates (8.57% of all published estimates) came from phylogenies with fewer than 30 tips.
+\hfill\break
+
+[insert Figure 3 here] \hfill\break
+
+Many of the reviewed manuscripts liberally interpreted the magnitude of the estimate lambda, using phrases such as "strong" or "weak" phylogentic signal when statistically, all that was clear was a difference between the estimated lambda and 0 or 1 respectively.
+We estimated that about 20.49% of the manuscripts revealed some sort of biological interpretation of the magnitude of estimated phylogenetic signal that overreached the statistical findings.
+We also identified seven manuscripts as having inappropriately interpreted differences in lambda values, indicating that some traits had stronger or weaker signal than other traits without the appropriate statistical tests.
+\hfill\break
+
+As is evidenced by macroevolutionary papers published in 2019 papers, Pagel's lambda estimation methods are often misused and over-interpretted.
+Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so.
+
+# Discussion
+
+[This part is obviously not written yet] \hfill\break
+
+General conclusions :Using the estimated lambda values from pgls are not useful.
+The questions of whether or not signal exists is appropriate, but inferring more from lambda *magnitude* is inappropriate.
 \hfill\break
@@
 
-\section{References}\label{references}
+# References
 
-\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
-\setlength{\parskip}{8pt} \noindent
+ \setlength{\parindent}{-0.25in}
+ \setlength{\leftskip}{0.25in}
+ \setlength{\parskip}{8pt}
+ \noindent
 
-\hypertarget{refs}{}
-\hypertarget{ref-Felsenstein1985}{}
-Felsenstein, J. 1985. Phylogenies and the comparative method. American
-Naturalist 125:1--15.
-
-\hypertarget{ref-Grafen1989}{}
-Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
-of the Royal Society of London B, Biological Sciences 326:119--157.
-
-\hypertarget{ref-HarveyPagel1991}{}
-Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
-evolutionary biology. Oxford University Press, Oxford.
-
-\hypertarget{ref-Revell2010}{}
-Revell, L. J. 2010. Phylogenetic signal and linear regression on species
-data. Methods in Ecology and Evolution 1:319--329.
-
-\hypertarget{ref-Rohlf2001}{}
-Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
-variables: Geometric interpretations. Evolution 55:2143--2160.
+<div id="refs"></div>
 
@@
 
-\section{Figure Legends}\label{figure-legends}
+# Figure Legends
 
-\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. As trees increase in size, the
-estimates more closely resemble the input lambdas, however considerable
-and concerning variation is apparent in trees smaller than those with
-200 tips. \hfill\break
+\textbf{Figure 1}.
+Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
+As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips.
+\hfill\break
 
-\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes,
-the mean estimated slope matches the input slope, and as trees increase
-in size, the variance around this mean estimate decreases. However, for
-trees with fewer than 200 tips, the error around the estimated slope is
-considerable, where these analyses frequently estimate slopes in the
-opposite direction of the known pattern. \hfill\break
+\textbf{Figure 2}.
+Estimated ANOVA slopes under PGLS.
+Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
+However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
+\hfill\break
 
-\textbf{Figure 3}. Frequency of estimated lambda values published in
-manuscripts in 2019. The majority of these values were close to 0 or 1,
-and from phylogenies with fewer than 200 taxa.
+\textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.
 
@@
 
-\includegraphics[width=0.95\linewidth]{Fig1}
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig1.png")
 
-\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations
-across known lambda inputs on various tree sizes. As trees increase in
-size, the estimates more closely resemble the input lambdas, however
-considerable and concerning variation is apparent in trees smaller than
-those with 200 tips. \hfill\break
+```
+
+\singlespacing \textbf{Figure 1}.
+Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes.
+As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips.
+\hfill\break
 
 \newpage
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig2.png")
 
-\includegraphics[width=0.95\linewidth]{Fig2}
+```
 
-\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS.
-Across tree sizes, the mean estimated slope matches the input slope, and
-as trees increase in size, the variance around this mean estimate
-decreases. However, for trees with fewer than 200 tips, the error around
-the estimated slope is considerable, where these analyses frequently
-estimate slopes in the opposite direction of the known pattern.
+\singlespacing \textbf{Figure 2}.
+Estimated ANOVA slopes under PGLS.
+Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases.
+However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern.
 \hfill\break
@@
 
-\includegraphics[width=0.95\linewidth]{Fig3}
+```{r, echo = FALSE, out.width="95%"}
+include_graphics("Fig3.png")
 
-\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
-published in manuscripts in 2019. The majority of these values were
-close to 0 or 1, and from phylogenies with fewer than 200 taxa.
+```
 
-\end{document}
+\singlespacing \textbf{Figure 3}.
+Frequency of estimated lambda values published in manuscripts in 2019.
+The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

### 2020-07-16 Manuscript/OldVsns/Manuscript.tex

--- 2020-04-08 Manuscript/Manuscript.tex
@@
 
+\hypertarget{pagels-lambda-estimates-are-often-inaccurate}{%
 \section{Pagel's Lambda Estimates are Often
-Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}
+Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}}
 
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 dozens of other methods have been developed and applied to increasingly
-complex questions in macroevolutionary biology (e.g. Grafen (1989);
+complex questions in macroevolutionary biology (e.g.~Grafen (1989);
 Harvey and Pagel (1991); Rohlf (2001); ({\textbf{???}})). A particularly
@@
 
-\section{Methods and Results}\label{methods-and-results}
+\hypertarget{methods-and-results}{%
+\section{Methods and Results}\label{methods-and-results}}
 
+\hypertarget{simulated-trait}{%
 \subsection{\texorpdfstring{\emph{Simulated
-trait}}{Simulated trait}}\label{simulated-trait}
+trait}}{Simulated trait}}\label{simulated-trait}}
 
@@
 
+\hypertarget{meta-analysis-of-empirical-results}{%
 \subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
-Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}
+Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}}
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Felsenstein1985}{}
+\leavevmode\hypertarget{ref-Felsenstein1985}{}%
 Felsenstein, J. 1985. Phylogenies and the comparative method. American
@@
 
-\hypertarget{ref-Grafen1989}{}
+\leavevmode\hypertarget{ref-Grafen1989}{}%
 Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
@@
 
-\hypertarget{ref-HarveyPagel1991}{}
+\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
 Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
@@
 
-\hypertarget{ref-Revell2010}{}
+\leavevmode\hypertarget{ref-Revell2010}{}%
 Revell, L. J. 2010. Phylogenetic signal and linear regression on species
@@
 
-\hypertarget{ref-Rohlf2001}{}
+\leavevmode\hypertarget{ref-Rohlf2001}{}%
 Rohlf, F. J. 2001. Comparative methods for the analysis of continuous

### 2020-07-16 Manuscript/PNAS-Adams-rticle.tex

--- 2020-07-16 Manuscript/OldVsns/Manuscript-8-DCA.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+
+%% Some pieces required from the pandoc template
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{setspace}\doublespacing
-\usepackage{lineno}\linenumbers
 
-\author{}
-\date{\vspace{-2.5em}}
+\templatetype{pnasresearcharticle}  % Choose template
 
-\begin{document}
+\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
+Phylogenetic Signal}
 
-\textbf{Classification:} Biological Sciences -- Evolution \hfill\break
+\author[a]{Dean C. Adams}
+\author[a,b]{Erica K. Baken}
+\author[b]{Michael L. Collyer}
 
-\begin{center}
-\textbf{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}
-\end{center}
+  \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
+University, Ames, Iowa, USA.}
+  \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
+USA.}
 
-\hfill\break
+% Please give the surname of the lead author for the running footer
+\leadauthor{Adams et al.}
 
-\begin{center}
-\textbf{Dean C. Adams$^{1 *}$, Erica K. Baken$^{1,2}$,  and Michael L. Collyer$^{2}$}
-\end{center}
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Evolutionary biologists wish to quantify and compare the strength of
+phylogenetic signal across traits, but analytical tools for these
+comparisons are generally lacking. Here we develop a standardized effect
+size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
+phylogenetic signal on a common statistical scale, and provides a
+mechanism for formally comparing the strength of phylogenetic signal
+across datasets. Additionally, we find that a commonly used parameter
+(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
+enables biologists to quantitatively address hypotheses that compare the
+strength of phylogenetic signal between various phenotypic traits, even
+when those traits are found in different evolutionary lineages or have
+different units or scales.}
 
-\hfill\break
+\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
+research and wrote the paper.}
 
-\(^{1}\)Department of Ecology, Evolution, and Organismal Biology, Iowa
-State University, Ames, Iowa, USA.
+\authordeclaration{The authors declare no conflict of interest.}
 
-\(^{2}\)Department of Science, Chatham University, Pittsburgh,
-Pennsylvania, USA.
+\correspondingauthor{\textsuperscript{} }
 
-\(^{*}\)Correspondence: Dean C. Adams. Department of Ecology, Evolution,
-and Organismal Biology, Iowa State University, Ames, Iowa, USA.
-515-294-3834.
-\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+ \keywords{  phylogenetic signal |  macroevolution |  lambda |  kappa  }
 
-\hfill\break
-
-\begin{center}15 July, 2020\end{center}
-
-\hfill\break
-
-\begin{longtable}[]{@{}rr@{}}
-\toprule
-words & characters\tabularnewline
-\midrule
-\endhead
-2553 & 16162\tabularnewline
-\bottomrule
-\end{longtable}
-
-\newpage
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
+\begin{abstract}
 Macroevolutionary studies frequently characterize the phylogenetic
@@
 traits are found in different evolutionary lineages or have different
-units or scales. \hfill\break
+units or scales.
+\end{abstract}
 
-\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
-\newpage
+\begin{document}
 
-\hypertarget{significance-statement}{%
-\section{Significance Statement}\label{significance-statement}}
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
 
-Evolutionary biologists wish to quantify and compare the strength of
-phylogenetic signal across traits, but analytical tools for these
-comparisons are generally lacking. Here we develop a standardized effect
-size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
-phylogenetic signal on a common statistical scale, and provides a
-mechanism for formally comparing the strength of phylogenetic signal
-across datasets. Additionally, we find that a commonly used parameter
-(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
-enables biologists to quantitatively address hypotheses that compare the
-strength of phylogenetic signal between various phenotypic traits, even
-when those traits are found in different evolutionary lineages or have
-different units or scales.
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\newpage
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the
+manuscript. This work was supported in part by NSF grant DBI-1902511 (to
+D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break}
 
@@
 of life generates trait covaration among related taxa (1, 14, 15).
-\hfill\break
 
@@
 measure the relative strength of phylogenetic signal, providing an
-estimate of its magnitude for comparison. \hfill\break
+estimate of its magnitude for comparison.
 
@@
 \(\lambda=0\) and \(\lambda=1\) and not prescribed values of \(\lambda\)
-(20), interpretation of these patterns is challenging. \hfill\break
+(20), interpretation of these patterns is challenging.
 
@@
 of the efficacy of each statistic to yield effect sizes that could be
-used for comparative anlyses. \hfill\break
+used for comparative anlyses.
 
@@
 properties make it insuitable as an effect size for measuring the
-strength of phylogenetic signal. \hfill\break
+strength of phylogenetic signal.
+
+\begin{figure}
+\centering
+\includegraphics{new.fig.1.temp.png}
+\caption{Patterns in \(\lambda\).{}}
+\end{figure}
 
@@
 differs from the results of (20), where skewing appears to be from
-combining random values generated independently. \hfill\break
+combining random values generated independently.
+
+\begin{figure}
+\centering
+\includegraphics{new.fig.2.temp.png}
+\caption{Patterns in \(\kappa\).{}}
+\end{figure}
 
@@
 phylogenetic signal, and may be used to compare levels of phylogenetic
-signal across datasets. \hfill\break
+signal across datasets.
+
+\begin{figure}
+\centering
+\includegraphics{new.fig.3.temp.png}
+\caption{Patterns in Z.{}}
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{new.fig.4.png}
+\caption{(A) Linear measures for relative body size, and regions of the
+body used to estimate surface area to volume (SA:V) ratios. (B)
+Permutation distributions of phylogenetic signal for SA:V and
+\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
+95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
+\end{figure}
+
 \hypertarget{discussion}{%
@@
 
-address hypotheses that compare the strength of phylogenetic signal
-between various phenotypic traits, even when those traits are found in
-different evolutionary lineages or have different units or scales
+\textbf{To be edited.} address hypotheses that compare the strength of
+phylogenetic signal between various phenotypic traits, even when those
+traits are found in different evolutionary lineages or have different
+units or scales.
+
+It is common in comparative evolutionary studies to characterize the
+phylogenetic signal in phenotypic traits to determine the extent to
+which shared evolutionary history has generated trait covariation among
+taxa. However, while numerous analytical approaches may be used to
+quantify phylogenetic signal (13, 14, 16--18), methods that explicitly
+measure the strength of phylogenetic signal, or facilitate comparisons
+among datasets, have remained underdeveloped. In this study, we
+evaluated the precision of one common measure, Pagel's \(\lambda\), and
+explored its efficacy for characterizing the strength of phylogenetic
+signal in phenotypic data. Using computer simulations, we found that the
+precision of \(\lambda\) increased with increasing sample sizes; a
+pattern noted previously (25), and one that conformed with parametric
+statistical theory (35). However, we also found that vastly different
+\(\lambda\) estimates could be obtained from data containing the same
+level of phylogenetic signal, and that similar \(\lambda\) estimates may
+be obtained from data containing differing levels of phylogenetic
+signal. Further, the precision of \(\lambda\) varied with the strength
+of phylogenetic signal, where lower precision was observed when in data
+whose phylogenetic signal was of intemediate strength. From these
+findings we conclude that \(\lambda\) is not a reliable indicator of the
+observed strength of phylogenetic signal in phenoytpic datasets, and
+that biological interpretations of the strength of signal based on this
+parameter may innacurately characterize such effects.
+
+As an alternative, we described a standardized effect size (\(Z\)) for
+assessing the strength of phylogenetic signal. \(Z\) expresses the
+magnitude of phylogenetic signal as a standard normal deviate, which is
+easily interpretable as the strength of phylogenetic signal relative to
+the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
+and found that \(Z_\kappa\) was a better estimate of the strength of
+phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
+precise than \(Z_\lambda\), and precision was more consistent across the
+range of input levels of phylogenetic signal. Additionally, values of
+\(Z_\kappa\) more accurately tracked known changes in the magnitude of
+phylogenetic signal, as demonstrated by the linear relationship between
+\(Z_\kappa\) and \(\lambda_{in}\). Thus, \(Z_\kappa\) holds promise as a
+measure of the relative strength of phylogenetic signal that reflects
+the magnitude of this effect in phenotypic data. We therefore recommend
+that future studies interested in the strength of phylogenetic signal
+incorporate \(Z_\kappa\) as a statistical measure of this effect.
+
+Based on the effect size \(Z_\kappa\), we then proposed a two-sample
+test, which provides means of determining whether the strength of
+phylogenetic signal is greater in one phenotypic trait as compared to
+another, via a hypothesis test. Prior studies have summarized patterns
+of variation in phylogenetic signal across datasets using summary test
+values, such as \(\kappa\) (14). However, \(\kappa\) does not scale
+linearly with input levels of phylogenetic signal, and its variance
+increases (i.e., precision decreases) with increasing strength of
+phylogenetic signal (20, 22). Thus, \(\kappa\) should not be considered
+an effect size that measures the strength of phylogenetic signal on a
+common scale. By contrast, standardizing \(\kappa\) (\(Z_\kappa\), via
+equation 2) alleviates these concerns, and facilitates formal
+statistical comparisons of the strength of signal across datasets. Thus
+when viewed from this perspective, the approach developed here aligns
+well with other statistical approaches such as meta-analysis (36--38),
+where summary statistics across datasets are converted to standardized
+effect sizes for subsequent ``higher order'' statistical summaries or
+comparisons. As such, our approach enables evolutionary biologists to
+quantitatively examine the relative strength of phylogenetic signal
+across a wide range of phenotypic traits, and thus opens the door for
+future discoveries that inform on how phenotypic diversity accumulates
+in macroevolutionary time across the tree of life.
+
+One important advantage of the approach advocated here is that the
+resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
+measurement cancel out during the calculation of \(Z\) (39). Thus,
+\(Z_\kappa\) represents the strength of phylogenetic signal on a common
+and comparable scale -- measured in standard deviations -- regardless of
+the initial units and original scale of the phenotypic variables under
+investigation. This means that the strength of phylogenetic signal may
+be compared across datsets for continuous phenotypic traits measured in
+different units and scale, because those units have been standardized
+through their conversion to \(Z_\kappa\). For example, our approach
+could be utilized to determine whether the strength of phylogenetic
+signal (say, in response to ecological differentiation) is stronger in
+morphological traits (linear traits: \(mm\)), physiological traits
+(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
+\(\frac{\#{displays}}{second}\)). In fact, our empirical example
+provided such a comparison, as SA:V is represented in \(mm^{-1}\) while
+relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
+Additionally, our method is capable of comparing the strength of
+phylogenetic signal in traits of different dimensionality, as estimates
+of phylogenetic signal using \(\kappa\) have been generalized for
+multivariate data (18). Furthermore, tests based on \(\hat{Z}_{12}\) may
+be utilized for comparing the strength of phylogenetic signal among
+datasets containing a different number of species, and even for
+phenotypes obtained from species in different lineages, because their
+phylogenetic non-independence and observed variation are taken into
+account in the generation of the empirical sampling distribution via
+permutation.
+
+This study is not the first to compare \(\lambda\) and \(\kappa\) for
+their ability as statistics to measure phylogenetic signal. Our results
+for \(\lambda\) and \(\kappa\) values are consistent with those found in
+the simulations performed by Münkemüller et al.~(20), but that study
+investigated type I error rates and statistical power, finding that
+\(\lambda\) performed better in both regards, irrespective of species
+number in trees. Although not the central focus of their study, the same
+tendency for variable \(\lambda\) and consistent \(\kappa\) at
+intermediate phylogenetic signal strengths was observed (20). Recent
+work by Molina-Venegas and Rodríguez (23) found that \(\kappa\) but not
+\(\lambda\) tended to inflate the estimate of phylogenetic signal,
+leading to moderate type I and type II biases, if polytomic chronograms
+were used. Their work more thoroughly addressed previous observations of
+inflated \(\kappa\) for incompletely resolved phylogenetic trees (20,
+40). An interesting question is whether an inflated \(\kappa\) value
+leads to an inflated \(Z_\kappa\) or does a tendency of a particular
+tree to inflate estimates of \(\kappa\) also inflate the values in
+random permutations of a test, in which case \(Z_\kappa\) is robust to
+polytomies? We repeated the analyses in Figure 4, adjusting trees to
+have 50\% collapsed nodes, per the technique of Molina-Venegas and
+Rodríguez (23), and found results were consistent (see Supporting
+Information). This confirms that any tendency of incompletely resolved
+trees to inflate \(\kappa\) as a descriptive statistic does not inflate
+\(Z_\kappa\) as an effect size. Furthermore, because comparison of
+effect sizes in a test is a comparison of locations of observed values
+in their sampling distributions, which would shift concomitantly because
+of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
+be robust in spite of unresolved trees.
+
+Phylogenetic signal can be thought of as both an attribute to be
+measured in the data and a parameter that can be tuned to account for
+the phylogenetic non-independence among observations, for analysis of
+the data. As such, \(\lambda\) is appealing, as a statistic that
+potentially fulfills both roles. However, the inability to estimate
+phylogenetic signal with \(\lambda\) for data simulated with known
+phylogenetic signal is troublesome, and we recommend evolutionary
+biologists refrain from viewing it as a statistic to describe the amount
+of phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
+standardized to an effect size \(Z_\kappa\) -- is a better statistic for
+measuring the amount of phylogenetic signal in data simulated with
+respect to known levels of \(\lambda\). Although \(\lambda\) might be
+viewed as an important parameter for modifying the the conditional
+estimation of linear model coefficients with respect to phylogeny, it is
+neither a statistic that has meaningful comparative value as a measure
+of phylogenetic signal nor a statistic that lends itself well to
+reliable calculation of a test statistic. By contrast, \(\kappa\) has
+been shown here to be a reliable statistic, but only when standardized
+by the mean and standard deviation of its empirical sampling
+distribution (i.e., when converted to the effect size, \(Z_\kappa\)).
+Because one has control over the number of permutations used in
+analysis, one can be assured with many permutations that the empirical
+sampling distribution is representative of true probability
+distributions (12). With low coefficients of variation for \(Z_\kappa\)
+(Figure 4), it is difficult to imagine that a hypothesis test can
+improve equation 4 for efficiently comparing phylogenetic signal for
+different traits, different trees, or a combination of both.
 
@@
 its expected value under the null hypothesis, and \(\sigma_\theta\) is
-its standard error (35--37). Typically, \(\theta_{obs}\) and
+its standard error (35, 37, 41). Typically, \(\theta_{obs}\) and
 \(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
 obtained from the distribution of \(\theta\) derived from parametric
-theory. However, recent advances in resampling theory (38--41) have
+theory. However, recent advances in resampling theory (42--45) have
 shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
 an empirical sampling distribution of \(\theta\) obtained from
-permutation procedures. \hfill\break
+permutation procedures.
 
-Formalizing the suggestion of Adams and Collyer (42), an effect size for
+Formalizing the suggestion of Adams and Collyer (46), an effect size for
 \(\kappa\) may be found as:
@@
 not linear and the log-likelihood profiles can be rather flat for small
-trees. \hfill\break
+trees.
 
@@
 (one-tailed) tests may be specified should the empirical situation
-require it (39, 41). \hfill\break
+require it (43, 45).
 
@@
 investigated, and results were qualitatively similar to those reported
-above (see Supporting Information). \hfill\break
+above (see Supporting Information).
 
@@
 species means were obtained (33, 34). A time-dated molecular phylogeny
-for the group (43) was then pruned to match the species in the
+for the group (47) was then pruned to match the species in the
 phenotypic dataset. The phylogenetic signal in each trait was then
 characterized using \(\kappa\), which was converted to its effect size
-(\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (44, 45).
+(\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (48, 49).
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and
-DBI-1902694 (to M.L.C.).
-
-\newpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
-\setlength{\parskip}{8pt} \noindent
+\showmatmethods
+\showacknow
+\pnasbreak
 
@@
 \leavevmode\hypertarget{ref-Felsenstein1985}{}%
-1. J. Felsenstein, Phylogenies and the comparative method.
-\emph{American Naturalist} \textbf{125}, 1--15 (1985).
+1. Felsenstein J (1985) Phylogenies and the comparative method.
+\emph{American Naturalist} 125(1):1--15.
 
 \leavevmode\hypertarget{ref-HarveyPagel1991}{}%
-2. P. H. Harvey, M. D. Pagel, \emph{The comparative method in
-evolutionary biology} (Oxford University Press, 1991).
+2. Harvey PH, Pagel MD (1991) \emph{The comparative method in
+evolutionary biology} (Oxford University Press, Oxford).
 
 \leavevmode\hypertarget{ref-Grafen1989}{}%
-3. A. Grafen, The phylogenetic regression. \emph{Philosophical
+3. Grafen A (1989) The phylogenetic regression. \emph{Philosophical
 Transactions of the Royal Society of London B, Biological Sciences}
-\textbf{326}, 119--157 (1989).
+326:119--157.
 
 \leavevmode\hypertarget{ref-GarlandIves2000}{}%
-4. T. J. Garland, A. R. Ives, Using the past to predict the present:
+4. Garland TJ, Ives AR (2000) Using the past to predict the present:
 Confidence intervals for regression equations in phylogenetic
-comparative methods. \emph{American Naturalist} \textbf{155}, 346--364
-(2000).
+comparative methods. \emph{American Naturalist} 155:346--364.
 
 \leavevmode\hypertarget{ref-Rohlf2001}{}%
-5. F. J. Rohlf, Comparative methods for the analysis of continuous
-variables: Geometric interpretations. \emph{Evolution} \textbf{55},
-2143--2160 (2001).
+5. Rohlf FJ (2001) Comparative methods for the analysis of continuous
+variables: Geometric interpretations. \emph{Evolution} 55:2143--2160.
 
 \leavevmode\hypertarget{ref-ButlerKing2004}{}%
-6. M. A. Butler, A. A. King, Phylogenetic comparative analysis: A
+6. Butler MA, King AA (2004) Phylogenetic comparative analysis: A
 modeling approach for adaptive evolution. \emph{American Naturalist}
-\textbf{164}, 683--695 (2004).
+164:683--695.
 
 \leavevmode\hypertarget{ref-MartinsHansen1997}{}%
-7. E. P. Martins, T. F. Hansen, Phylogenies and the comparative method:
+7. Martins EP, Hansen TF (1997) Phylogenies and the comparative method:
 A general approach to incorporating phylogenetic information into the
-analysis of interspecific data. \emph{American Naturalist} \textbf{149},
-646--667 (1997).
+analysis of interspecific data. \emph{American Naturalist} 149:646--667.
 
 \leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
-8. B. C. O'Meara, C. Ane, M. J. Sanderson, P. C. Wainwright, Testing for
+8. O'Meara BC, Ane C, Sanderson MJ, Wainwright PC (2006) Testing for
 different rates of continuous trait evolution using likelihood.
-\emph{Evolution} \textbf{60}, 922--933 (2006).
+\emph{Evolution} 60:922--933.
 
 \leavevmode\hypertarget{ref-RevellHarmon2008}{}%
-9. L. J. Revell, L. J. Harmon, Testing quantitative genetic hypotheses
+9. Revell LJ, Harmon LJ (2008) Testing quantitative genetic hypotheses
 about the evolutionary rate matrix for continuous characters.
-\emph{Evolutionary Ecology Research} \textbf{10}, 311--331 (2008).
+\emph{Evolutionary Ecology Research} 10:311--331.
 
 \leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
-10. J. M. Beaulieu, D. C. Jhwueng, C. Boettiger, B. C. O'Meara, Modeling
+10. Beaulieu JM, Jhwueng DC, Boettiger C, O'Meara BC (2012) Modeling
 stabilizing selection: Expanding the ornstein-uhlenbeck model of
-adaptive evolution. \emph{Evolution} \textbf{66}, 2369--2383 (2012).
+adaptive evolution. \emph{Evolution} 66:2369--2383.
 
 \leavevmode\hypertarget{ref-Adams2014b}{}%
-11. D. C. Adams, A method for assessing phylogenetic least squares
+11. Adams DC (2014) A method for assessing phylogenetic least squares
 models for shape and other high-dimensional multivariate data.
-\emph{Evolution} \textbf{68}, 2675--2688 (2014).
+\emph{Evolution} 68:2675--2688.
 
 \leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
-12. D. C. Adams, M. L. Collyer, Phylogenetic anova: Group-clade
+12. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
 aggregation, biological challenges, and a refined permutation procedure.
-\emph{Evolution} \textbf{72}, 1204--1215 (2018).
+\emph{Evolution} 72(6):1204--1215.
 
 \leavevmode\hypertarget{ref-Pagel1999}{}%
-13. M. D. Pagel, Inferring the historical patterns of biological
-evolution. \emph{Nature} \textbf{401}, 877--884 (1999).
+13. Pagel MD (1999) Inferring the historical patterns of biological
+evolution. \emph{Nature} 401:877--884.
 
 \leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
-14. S. P. Blomberg, T. Garland, A. R. Ives, Testing for phylogenetic
+14. Blomberg SP, Garland T, Ives AR (2003) Testing for phylogenetic
 signal in comparative data: Behavioral traits are more labile.
-\emph{Evolution} \textbf{57}, 717--745 (2003).
+\emph{Evolution} 57:717--745.
 
 \leavevmode\hypertarget{ref-Revell_et_al2008}{}%
-15. L. J. Revell, L. J. Harmon, D. C. Collar, Phylogenetic signal,
-evolutionary process, and rate. \emph{Systematic Biology} \textbf{57},
-591--601 (2008).
+15. Revell LJ, Harmon LJ, Collar DC (2008) Phylogenetic signal,
+evolutionary process, and rate. \emph{Systematic Biology} 57:591--601.
 
 \leavevmode\hypertarget{ref-Abouheif1999}{}%
-16. E. Abouheif, A method for testing the assumption of phylogenetic
-independence in comparative data. \emph{Evolutionary Ecology Research}
-\textbf{1}, 895--909 (1999).
+16. Abouheif E (1999) A method for testing the assumption of
+phylogenetic independence in comparative data. \emph{Evolutionary
+Ecology Research} 1:895--909.
 
 \leavevmode\hypertarget{ref-Gittleman1990}{}%
-17. J. L. Gittleman, M. Kot, Adaptation: Statistics and a null model for
-estimating phylogenetic effects. \emph{Systematic Zoology} \textbf{39},
-227--241 (1990).
+17. Gittleman JL, Kot M (1990) Adaptation: Statistics and a null model
+for estimating phylogenetic effects. \emph{Systematic Zoology}
+39(3):227--241.
 
 \leavevmode\hypertarget{ref-Adams2014a}{}%
-18. D. C. Adams, A generalized Kappa statistic for estimating
+18. Adams DC (2014) A generalized Kappa statistic for estimating
 phylogenetic signal from shape and other high-dimensional dultivariate
-data. \emph{Systematic Biology} \textbf{63}, 685--697 (2014).
+data. \emph{Systematic Biology} 63:685--697.
 
 \leavevmode\hypertarget{ref-Klingenberg2010}{}%
-19. C. P. Klingenberg, N. A. Gidaszewski, Testing and quantifying
+19. Klingenberg CP, Gidaszewski NA (2010) Testing and quantifying
 phylogenetic signals and homoplasy in morphometric data.
-\emph{Systematic biology} \textbf{59}, 245--261 (2010).
+\emph{Systematic biology} 59(3):245--261.
 
 \leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
-20. T. Münkemüller\emph{et al.}, How to measure and test phylogenetic
-signal. \emph{Methods in Ecology and Evolution} \textbf{3}, 743--756
-(2012).
+20. Münkemüller T, et al. (2012) How to measure and test phylogenetic
+signal. \emph{Methods in Ecology and Evolution} 3:743--756.
 
 \leavevmode\hypertarget{ref-Pavoine2012}{}%
-21. S. Pavoine, C. Ricotta, Testing for phylogenetic signal in
+21. Pavoine S, Ricotta C (2012) Testing for phylogenetic signal in
 biological traits: The ubiquity of cross-product statistics.
 \emph{Evolution: International Journal of Organic Evolution}
-\textbf{67}, 828--840 (2012).
+67(3):828--840.
 
 \leavevmode\hypertarget{ref-DinizFilho2012}{}%
-22. J. A. F. Diniz-Filho, T. Santos, T. F. Rangel, L. M. Bini, A
-comparison of metrics for estimating phylogenetic signal under
-alternative evolutionary models. \emph{Genetics and Molecular Biology}
-\textbf{35}, 673--679 (2012).
+22. Diniz-Filho JAF, Santos T, Rangel TF, Bini LM (2012) A comparison of
+metrics for estimating phylogenetic signal under alternative
+evolutionary models. \emph{Genetics and Molecular Biology}
+35(3):673--679.
 
 \leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
-23. R. Molina-Venegas, M. A. Rodríguez, Revisiting phylogenetic signal;
-strong or negligible impacts of polytomies and branch length
-information? \emph{BMC evolutionary biology} \textbf{17}, 53 (2017).
+23. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic
+signal; strong or negligible impacts of polytomies and branch length
+information? \emph{BMC evolutionary biology} 17(1):53.
 
 \leavevmode\hypertarget{ref-Revell2010}{}%
-24. L. J. Revell, Phylogenetic signal and linear regression on species
-data. \emph{Methods in Ecology and Evolution} \textbf{1}, 319--329
-(2010).
+24. Revell LJ (2010) Phylogenetic signal and linear regression on
+species data. \emph{Methods in Ecology and Evolution} 1:319--329.
 
 \leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
-25. C. Boettiger, G. Coop, P. Ralph, Is your phylogeny informative?
+25. Boettiger C, Coop G, Ralph P (2012) Is your phylogeny informative?
 Measuring the power of comparative methods. \emph{Evolution}
-\textbf{67}, 2240--2251 (2012).
+67:2240--2251.
 
 \leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
-26. R. P. Freckleton, P. H. Harvey, M. Pagel, Phylogenetic analysis and
+26. Freckleton RP, Harvey PH, Pagel M (2002) Phylogenetic analysis and
 comparative data: A test and review of evidence. \emph{American
-Naturalist} \textbf{160}, 712--726 (2002).
+Naturalist} 160:712--726.
 
 \leavevmode\hypertarget{ref-Cooper2010}{}%
-27. N. Cooper, W. Jetz, R. P. Freckleton, Phylogenetic comparative
+27. Cooper N, Jetz W, Freckleton RP (2010) Phylogenetic comparative
 approaches for studying niche conservatism. \emph{Journal of
-Evolutionary Biology} \textbf{23}, 2529--2539 (2010).
+Evolutionary Biology} 23(12):2529--2539.
 
 \leavevmode\hypertarget{ref-Bose2019}{}%
-28. R. Bose, B. R. Ramesh, R. Pélissier, F. Munoz, Phylogenetic
+28. Bose R, Ramesh BR, Pélissier R, Munoz F (2019) Phylogenetic
 diversity in the western ghats biodiversity hotspot reflects
 environmental filtering and past niche diversification of trees.
-\emph{Journal of Biogeography} \textbf{46}, 145--157 (2019).
+\emph{Journal of Biogeography} 46(1):145--157.
 
 \leavevmode\hypertarget{ref-Vandelook2019}{}%
-29. F. Vandelook\emph{et al.}, Nectar traits differ between pollination
-syndromes in balsaminaceae. \emph{Annals of Botany} \textbf{124},
-269--279 (2019).
+29. Vandelook F, et al. (2019) Nectar traits differ between pollination
+syndromes in balsaminaceae. \emph{Annals of Botany} 124(2):269--279.
 
 \leavevmode\hypertarget{ref-DeMeester2019}{}%
-30. G. De Meester, K. Huyghe, R. Van Damme, Brain size, ecology and
+30. De Meester G, Huyghe K, Van Damme R (2019) Brain size, ecology and
 sociality: A reptilian perspective. \emph{Biological Journal of the
-Linnean Society} \textbf{126}, 381--391 (2019).
+Linnean Society} 126(3):381--391.
 
 \leavevmode\hypertarget{ref-Pintanel2019}{}%
-31. P. Pintanel, M. Tejedo, S. R. Ron, G. A. Llorente, A. Merino-Viteri,
+31. Pintanel P, Tejedo M, Ron SR, Llorente GA, Merino-Viteri A (2019)
 Elevational and microclimatic drivers of thermal tolerance in andean
-pristimantis frogs. \emph{Journal of Biogeography} \textbf{46},
-1664--1675 (2019).
+pristimantis frogs. \emph{Journal of Biogeography} 46(8):1664--1675.
 
 \leavevmode\hypertarget{ref-Su2019}{}%
-32. G. Su, S. Villéger, S. Brosse, Morphological diversity of freshwater
-fishes differs between realms, but morphologically extreme species are
-widespread. \emph{Global ecology and biogeography} \textbf{28}, 211--221
-(2019).
+32. Su G, Villéger S, Brosse S (2019) Morphological diversity of
+freshwater fishes differs between realms, but morphologically extreme
+species are widespread. \emph{Global ecology and biogeography}
+28(2):211--221.
 
 \leavevmode\hypertarget{ref-Baken2019}{}%
-33. E. K. Baken, D. C. Adams, Macroevolution of arboreality in
-salamanders. \emph{Ecology and Evolution} \textbf{9}, 7005--7016 (2019).
+33. Baken EK, Adams DC (2019) Macroevolution of arboreality in
+salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.
 
 \leavevmode\hypertarget{ref-Baken2020}{}%
-34. E. K. Baken, L. E. Mellenthin, D. C. Adams, Macroevolution of
+34. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
 desiccation‐related morphology in plethodontid salamanders as inferred
 from a novel surface area to volume ratio estimation approach.
-\emph{Evolution} \textbf{74}, 476--486 (2020).
+\emph{Evolution} 74:476--486.
+
+\leavevmode\hypertarget{ref-Cohen1988}{}%
+35. Cohen J (1988) \emph{Statistical power analysis for the behavioral
+sciences} (Routledge).
+
+\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
+36. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
+meta-analysis} (Elsevier).
 
 \leavevmode\hypertarget{ref-Glass1976}{}%
-35. G. V. Glass, Primary, secondary, and meta-analysis of research.
-\emph{Educational Researcher} \textbf{5}, 3--8 (1976).
+37. Glass GV (1976) Primary, secondary, and meta-analysis of research.
+\emph{Educational Researcher} 5:3--8.
 
-\leavevmode\hypertarget{ref-Cohen1988}{}%
-36. J. Cohen, \emph{Statistical power analysis for the behavioral
-sciences} (Routledge, 1988).
+\leavevmode\hypertarget{ref-Arnqvist1995}{}%
+38. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
+findings in ecology and evolution. \emph{Trends in Ecology and
+Evolution} 10:236--240.
+
+\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
+39. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
+San Francisco). 4th Ed.
+
+\leavevmode\hypertarget{ref-Davies2012}{}%
+40. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
+resolved phylogenetic trees inflate estimates of phylogenetic
+conservatism. \emph{Ecology} 93(2):242--247.
 
 \leavevmode\hypertarget{ref-Rosenthal1994}{}%
-37. R. Rosenthal, ``The handbook of research synthesis'' in L. V. Cooper
-H Hedges, Ed. (Russell Sage Foundation, 1994), pp. 231--244.
+41. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
+H Hedges (Russell Sage Foundation), pp 231--244.
 
 \leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
-38. M. L. Collyer, D. J. Sekora, D. C. Adams, A method for analysis of
+42. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
 phenotypic change for phenotypes described by high-dimensional data.
-\emph{Heredity} \textbf{115}, 357--365 (2015).
+\emph{Heredity} 115:357--365.
 
 \leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
-39. D. C. Adams, M. L. Collyer, On the comparison of the strength of
+43. Adams DC, Collyer ML (2016) On the comparison of the strength of
 morphological integration across morphometric datasets. \emph{Evolution}
-\textbf{70}, 2623--2631 (2016).
+70:2623--2631.
 
 \leavevmode\hypertarget{ref-CollyerAdams2018}{}%
-40. M. L. Collyer, D. C. Adams, RRPP: An r package for fitting linear
+44. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
 models to high-dimensional data using residual randomization.
-\emph{Methods in Ecology and Evolution} \textbf{9}, 1772--1779 (2018).
+\emph{Methods in Ecology and Evolution} 9:1772--1779.
 
 \leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
-41. D. C. Adams, M. L. Collyer, Comparing the strength of modular
+45. Adams DC, Collyer ML (2019) Comparing the strength of modular
 signal, and evaluating alternative modular hypotheses, using covariance
-ratio effect sizes with morphometric data. \emph{Evolution} \textbf{73},
-2352--2367 (2019).
+ratio effect sizes with morphometric data. \emph{Evolution}
+73(12):2352--2367.
 
 \leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
-42. D. C. Adams, M. L. Collyer, Phylogenetic comparative methods and the
+46. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
 evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
-Evolution, and Systematics} \textbf{50}, 405--425 (2019).
+Evolution, and Systematics} 50:405--425.
 
 \leavevmode\hypertarget{ref-Bonett2017}{}%
-43. R. M. Bonett, A. L. Blair, Evidence for complex life cycle
+47. Bonett RM, Blair AL (2017) Evidence for complex life cycle
 constraints on salamander body form diversification. \emph{Proceedings
-of the National Academy of Sciences, U.S.A.} \textbf{114}, 9936--9941
-(2017).
+of the National Academy of Sciences, USA} 114:9936--9941.
 
 \leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
-44. D. C. Adams, E. Otárola-Castillo, Geomorph: An r package for the
+48. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
 collection and analysis of geometric morphometric shape data.
-\emph{Methods in Ecology and Evolution} \textbf{4}, 393--399 (2013).
+\emph{Methods in Ecology and Evolution} 4:393--399.
 
 \leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-45. D. C. Adams, M. L. Collyer, A. Kaliontzopoulou, Geomorph: Software
-for geometric morphometric analyses. R package version 3.3.1. (2020).
+49. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
+for geometric morphometric analyses. R package version 3.3.1. Available
+at: \url{https://cran.r-project.org/package=geomorph}.
 
-\newpage
-
-Real figures will have 6 panels for lambda and kappa and probably only a
-subset for z. Simulations need to be re-run.
-
-\includegraphics[width=0.95\linewidth]{new.fig.1.temp}
-
-\singlespacing \textbf{Figure 1}. Temporary Fig. 1 showing lambda
-patterns
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{new.fig.2.temp}
-
-\singlespacing \textbf{Figure 2}. Temporary Fig. 2 showing kappa
-patterns
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{new.fig.3.temp}
-
-\singlespacing \textbf{Figure 3}. Temporary Fig. 3 showing z patterns
-
-\newpage
-
-\includegraphics[width=0.95\linewidth]{new.fig.4}
-
-\singlespacing \textbf{Figure 4}. (A) Linear measures for relative body
-size, and regions of the body used to estimate surface area to volume
-(SA:V) ratios. (B) Permutation distributions of phylogenetic signal for
-SA:V and \(\frac{BW}{SVL}\), with observed values shown as vertical
-bars. (C) Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\),
-with their 95\% confidence intervals (CI not standardized by
-\(\sqrt(n)\)).
+% Bibliography
+% \bibliography{pnas-sample}
 

### 2020-07-16 Manuscript/pnas_article-Template/skeleton/PNAS-Adams-rticle.tex

\documentclass[9pt,twocolumn,twoside,]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
Phylogenetic Signal}

\author[a]{Dean C. Adams}
\author[a,b]{Erica K. Baken}
\author[b]{Michael L. Collyer}

  \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
University, Ames, Iowa, USA.}
  \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
USA.}

% Please give the surname of the lead author for the running footer
\leadauthor{Adams et al.}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Evolutionary biologists wish to quantify and compare the strength of
phylogenetic signal across traits, but analytical tools for these
comparisons are generally lacking. Here we develop a standardized effect
size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
phylogenetic signal on a common statistical scale, and provides a
mechanism for formally comparing the strength of phylogenetic signal
across datasets. Additionally, we find that a commonly used parameter
(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
enables biologists to quantitatively address hypotheses that compare the
strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have
different units or scales.}

\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
research and wrote the paper.}

\authordeclaration{The authors declare no conflict of interest.}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  phylogenetic signal |  macroevolution |  lambda |  kappa  }

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. Here we evaluate the efficacy of one
commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
of phylogenetic signal in phenotypic traits, and evaluate the degree to
which \(\lambda\) correctly identifies known levels of phylogenetic
signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
and that estimates are increasingly skewed at larger and smaller input
levels of phylogenetic signal. Further, the precision of \(\lambda\) in
estimating actual levels of phylogenetic signal is often inaccurate, and
biological interpretations of the strength of phylogenetic signal based
on \(\lambda\) are therefore compromised. We propose a standardized
effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
strength of phylogenetic signal more reliably than does \(\lambda\), and
places that signal on a common scale for statistical comparison. We
develop tests based on \(Z_\kappa\) to provide a mechanism for formally
comparing the strength of phylogenetic signal across datasets, in much
the same manner as effect sizes may be used to summarize patterns in
quantitative meta-analysis. Our approach extends the phylogenetic
comparative toolkit to address hypotheses that compare the strength of
phylogenetic signal between various phenotypic traits, even when those
traits are found in different evolutionary lineages or have different
units or scales.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the
manuscript. This work was supported in part by NSF grant DBI-1902511 (to
D.C.A.) and DBI-1902694 (to M.L.C.).}

This PNAS journal template is provided to help you write your work in
the correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
``Introduction'' as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

\hypertarget{guide-to-using-this-template}{%
\section*{Guide to using this
template}\label{guide-to-using-this-template}}
\addcontentsline{toc}{section}{Guide to using this template}

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the
\href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information
for Authors}.

\hypertarget{author-affiliations}{%
\subsection*{Author Affiliations}\label{author-affiliations}}
\addcontentsline{toc}{subsection}{Author Affiliations}

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

\hypertarget{submitting-manuscripts}{%
\subsection*{Submitting Manuscripts}\label{submitting-manuscripts}}
\addcontentsline{toc}{subsection}{Submitting Manuscripts}

All authors must submit their articles at
\href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you
are using Overleaf to write your article, you can use the ``Submit to
PNAS'' option in the top bar of the editor window.

\hypertarget{format}{%
\subsection*{Format}\label{format}}
\addcontentsline{toc}{subsection}{Format}

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

\hypertarget{manuscript-length}{%
\subsection*{Manuscript Length}\label{manuscript-length}}
\addcontentsline{toc}{subsection}{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for PNAS PLUS.

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

References should be cited in numerical order as they appear in text;
this will be done automatically via bibtex, e.g.~All references,
including for the SI, should be included in the main manuscript file.
References appearing in both sections should not be duplicated. SI
references included in tables should be included with the main reference
section.

\hypertarget{data-archival}{%
\subsection*{Data Archival}\label{data-archival}}
\addcontentsline{toc}{subsection}{Data Archival}

PNAS must be able to archive the data essential to a published article.
Where such archiving is not possible, deposition of data in public
databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
and others outlined in the Information for Authors, is acceptable.

\hypertarget{language-editing-services}{%
\subsection*{Language-Editing
Services}\label{language-editing-services}}
\addcontentsline{toc}{subsection}{Language-Editing Services}

Prior to submission, authors who believe their manuscripts would benefit
from professional editing are encouraged to use a language-editing
service (see list at www.pnas.org/site/authors/language-editing.xhtml).
PNAS does not take responsibility for or endorse these services, and
their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}
\centering
\includegraphics{frog.png}
\caption{Placeholder image of a frog with a long example caption to show
justification setting.{}}
\end{figure}

\hypertarget{sec:figures}{%
\subsection*{Digital Figures}\label{sec:figures}}
\addcontentsline{toc}{subsection}{Digital Figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for
figures that will appear in the main text, and images must be final
size. Authors may submit U3D or PRC files for 3D images; these must be
accompanied by 2D representations in TIFF, EPS, or high-resolution PDF
format. Color images must be in RGB (red, green, blue) mode. Include the
font files for any text.

Figures and Tables should be labelled and referenced in the standard way
using the \texttt{\textbackslash{}label\{\}} and
\texttt{\textbackslash{}ref\{\}} commands.

Figure \[fig:frog\] shows an example of how to insert a column-wide
figure. To insert a figure wider than one column, please use the
\texttt{\textbackslash{}begin\{figure*\}...\textbackslash{}end\{figure*\}}
environment. Figures wider than one column should be sized to 11.4 cm or
17.8 cm wide.

\hypertarget{single-column-equations}{%
\subsection*{Single column equations}\label{single-column-equations}}
\addcontentsline{toc}{subsection}{Single column equations}

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
\texttt{\textbackslash{}begin\{figure*\}...\textbackslash{}end\{figure*\}}
environment mentioned above for figures, or to use the
\texttt{\textbackslash{}begin\{widetext\}...\textbackslash{}end\{widetext\}}
environment as shown in equation \[eqn:example\] below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted
package documentation}. In the case of problems with footnotes, it may
be possible to correct the situation using commands
\texttt{\textbackslash{}footnotemark} and
\texttt{\textbackslash{}footnotetext}.

\[\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}\]

\hypertarget{supporting-information-si}{%
\subsection*{Supporting Information
(SI)}\label{supporting-information-si}}
\addcontentsline{toc}{subsection}{Supporting Information (SI)}

The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

\hypertarget{si-text}{%
\subsubsection*{SI Text}\label{si-text}}
\addcontentsline{toc}{subsubsection}{SI Text}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

\hypertarget{si-figures}{%
\subsubsection*{SI Figures}\label{si-figures}}
\addcontentsline{toc}{subsubsection}{SI Figures}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

\hypertarget{d-figures}{%
\subsubsection*{3D Figures}\label{d-figures}}
\addcontentsline{toc}{subsubsection}{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

\hypertarget{si-tables}{%
\subsubsection*{SI Tables}\label{si-tables}}
\addcontentsline{toc}{subsubsection}{SI Tables}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

\hypertarget{si-datasets}{%
\subsubsection*{SI Datasets}\label{si-datasets}}
\addcontentsline{toc}{subsubsection}{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

\hypertarget{si-movies}{%
\subsubsection*{SI Movies}\label{si-movies}}
\addcontentsline{toc}{subsubsection}{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

\hypertarget{still-images}{%
\subsubsection*{Still images}\label{still-images}}
\addcontentsline{toc}{subsubsection}{Still images}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

\hypertarget{appendices}{%
\subsubsection*{Appendices}\label{appendices}}
\addcontentsline{toc}{subsubsection}{Appendices}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.

\showmatmethods
\showacknow
\pnasbreak

% Bibliography
% \bibliography{pnas-sample}

\end{document}

### 2020-07-16 Manuscript/pnas_article-Template/skeleton/skeleton.tex

--- 2020-07-16 Manuscript/pnas_article-Template/skeleton/PNAS-Adams-rticle.tex
@@
 
-\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
-Phylogenetic Signal}
+\title{Template for preparing your research report submission to PNAS using
+RMarkdown}
 
-\author[a]{Dean C. Adams}
-\author[a,b]{Erica K. Baken}
-\author[b]{Michael L. Collyer}
+\author[a,1,2]{Alice Anonymous}
+\author[a,b]{Bob Security}
 
-  \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
-University, Ames, Iowa, USA.}
-  \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
-USA.}
+  \affil[a]{Some Institute of Technology, Department, Street, City, State, Zip}
+  \affil[b]{Another University Department, Street, City, State, Zip}
 
 % Please give the surname of the lead author for the running footer
-\leadauthor{Adams et al.}
+\leadauthor{Anonymous}
 
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Evolutionary biologists wish to quantify and compare the strength of
-phylogenetic signal across traits, but analytical tools for these
-comparisons are generally lacking. Here we develop a standardized effect
-size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
-phylogenetic signal on a common statistical scale, and provides a
-mechanism for formally comparing the strength of phylogenetic signal
-across datasets. Additionally, we find that a commonly used parameter
-(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
-enables biologists to quantitatively address hypotheses that compare the
-strength of phylogenetic signal between various phenotypic traits, even
-when those traits are found in different evolutionary lineages or have
-different units or scales.}
+\significancestatement{Authors must submit a 120-word maximum statement about the significance
+of their research paper written at a level understandable to an
+undergraduate educated scientist outside their field of speciality. The
+primary goal of the Significance Statement is to explain the relevance
+of the work in broad context to a broad readership. The Significance
+Statement appears in the paper itself and is required for all research
+papers.}
 
-\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
-research and wrote the paper.}
+\authorcontributions{Please provide details of author contributions here.}
 
-\authordeclaration{The authors declare no conflict of interest.}
+\authordeclaration{Please declare any conflict of interest here.}
+
+\equalauthors{\textsuperscript{} }
 
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
- \keywords{  phylogenetic signal |  macroevolution |  lambda |  kappa  }
+ \keywords{  one |  two |  optional |  optional |  optional  }
 
 \begin{abstract}
-Macroevolutionary studies frequently characterize the phylogenetic
-signal in phenotypes, and wish to compare the strength of that signal
-across traits. However, analytical tools for such comparisons have
-largely remained underdeveloped. Here we evaluate the efficacy of one
-commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
-of phylogenetic signal in phenotypic traits, and evaluate the degree to
-which \(\lambda\) correctly identifies known levels of phylogenetic
-signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
-and that estimates are increasingly skewed at larger and smaller input
-levels of phylogenetic signal. Further, the precision of \(\lambda\) in
-estimating actual levels of phylogenetic signal is often inaccurate, and
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) are therefore compromised. We propose a standardized
-effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
-strength of phylogenetic signal more reliably than does \(\lambda\), and
-places that signal on a common scale for statistical comparison. We
-develop tests based on \(Z_\kappa\) to provide a mechanism for formally
-comparing the strength of phylogenetic signal across datasets, in much
-the same manner as effect sizes may be used to summarize patterns in
-quantitative meta-analysis. Our approach extends the phylogenetic
-comparative toolkit to address hypotheses that compare the strength of
-phylogenetic signal between various phenotypic traits, even when those
-traits are found in different evolutionary lineages or have different
-units or scales.
+Please provide an abstract of no more than 250 words in a single
+paragraph. Abstracts should explain to the general reader the major
+contributions of the article. References in the abstract must be cited
+in full within the abstract itself and cited in the text.
 \end{abstract}
@@
 
-\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the
-manuscript. This work was supported in part by NSF grant DBI-1902511 (to
-D.C.A.) and DBI-1902694 (to M.L.C.).}
+\acknow{Please include your acknowledgments here, set in a single paragraph.
+Please do not include any acknowledgments in the Supporting Information,
+or anywhere else in the manuscript.}
 
@@
 References should be cited in numerical order as they appear in text;
-this will be done automatically via bibtex, e.g.~All references,
-including for the SI, should be included in the main manuscript file.
-References appearing in both sections should not be duplicated. SI
-references included in tables should be included with the main reference
-section.
+this will be done automatically via bibtex, e.g.~(1) and (2, 3). All
+references, including for the SI, should be included in the main
+manuscript file. References appearing in both sections should not be
+duplicated. SI references included in tables should be included with the
+main reference section.
 
@@
 
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-belkin2002using}{}%
+1. Belkin M, Niyogi P (2002) Using manifold stucture for partially
+labeled classification. \emph{Advances in Neural Information Processing
+Systems}, pp 929--936.
+
+\leavevmode\hypertarget{ref-berard1994embedding}{}%
+2. Bérard P, Besson G, Gallot S (1994) Embedding riemannian manifolds by
+their heat kernel. \emph{Geometric \& Functional Analysis GAFA}
+4(4):373--398.
+
+\leavevmode\hypertarget{ref-coifman2005geometric}{}%
+3. Coifman RR, et al. (2005) Geometric diffusions as a tool for harmonic
+analysis and structure definition of data: Diffusion maps.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America} 102(21):7426--7431.
+
 % Bibliography

### 2020-07-16 Manuscript/PNAS-Adams-rticle.Rmd
11

@@
   - name: Dean C. Adams
-    affiliation: a
+    affiliation: a,1
   - name: Erica K. Baken
@@
   - code: a
-    address: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.
+    address: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, 50010. USA.
   - code: b
-    address: Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.
+    address: Department of Science, Chatham University, Pittsburgh, Pennsylvania, 15232. USA.
 
 corresponding_author:
-  - code: 2
+  - code: 1
     text: "To whom correspondence should be addressed. E-mail: dcadams@iastate.edu"
@@
 conflict_of_interest: |
-  The authors declare no conflict of interest.
+  The authors declare no conflict of interest.  \hfill\break
+
+  Data deposition: Data for the empirical example may be found on DRYAD: doi:10.5061/dryad.b554m44 and doi:10.5061/dryad.59zw3r23m.
+R-scripts for simulation tests are found on Github: XXX.
+Computer code for implementing the two-sample comparison of effect sizes is found in geomorph: https://cran.r-project.org/web/packages/geomorph/index.html \hfill\break
+
+  ^1^To whom correspondence should be addressed. E-mail: dcadams@iastate.edu
 
@@
 Further, the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised.
-We propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
+As an alternative, we propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
 We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
@@
   Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across traits, but analytical tools for these comparisons are generally lacking.
-Here we develop a standardized effect size based on $\kappa$ ($Z_\kappa$), which measures the strength of phylogenetic signal on a common statistical scale, and provides a mechanism for formally comparing the strength of phylogenetic signal across datasets.
-Additionally, we find that a commonly used parameter (Pagel's $\lambda$) is insuitable for this purpose.
+Here we develop a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal on a common statistical scale, and provides a mechanism for formally comparing the strength of phylogenetic signal across datasets.
+Additionally, we find that a commonly used parameter (Pagel's $\lambda$) is unsuitable for this purpose.
 Our procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
@@
 One of the most widely used methods for characterizing phylogenetic signal is Pagel's $\lambda$ [@Pagel1999], which transforms the lengths of the internal branches of the phylogeny to improve the fit of data to the phylogeny via maximum likelihood [@Pagel1999; @Freckleton_et_al2002].
-When incorporated in PGLS, $\lambda$ serves as a tuning parameter (ranging from $0\rightarrow1$) which is optimized via log-likelihood profiling while evaluating the covariation between the dependent and independent variables, given the phylogeny [@Pagel1999; @Freckleton_et_al2002].
+When incorporated in PGLS, $\lambda$ serves as a tuning parameter which is optimized via log-likelihood profiling while evaluating the covariation between the dependent and independent variables, given the phylogeny [@Pagel1999; @Freckleton_et_al2002].
 To infer whether phylogenetic signal differs from no signal or a Brownian motion model of evolutionary divergence, the observed model fit using $\hat\lambda$ may be statistically compared to that using  $\lambda=0$ or $\lambda=1$ via likelihood ratio tests [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or confidence limits [@Vandelook2019].
@@
 Prior work [@Munkemuller_et_al2012] seems to support some of these conjectures, based superficially on statistical moments for a given tree size (mean, variance, skewness, and kurtosis; see Fig.
-2 of [@Munkemuller_et_al2012]).
-However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
+2 of ref.
+[@Munkemuller_et_al2012]).
+However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not as prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
 
@@
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
-[-@Munkemuller_et_al2012], as distibutions were postively skewed, suggesting the Blomberg's $\kappa$ might not behave as a statistic that follows a noraml distribution.
+[-@Munkemuller_et_al2012], which suggested that distributions were positively skewed and that Blomberg's $\kappa$ might not behave as a statistic that follows a normal distribution.
 However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations).
-For both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.
+Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.
 Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
-Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparative anlyses.
+Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and clades.
 
-Here we used simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and Blomberg's $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
-We find that estimates of $\hat{\lambda}$ are increasingly skewed at larger and smaller input levels of phylogenetic signal, vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range.
-By contrast, estimates of $\kappa$ are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$.
-We then propose an effect size based on $\kappa$ ($Z_{\kappa}$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.
+Here we used simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
+We find that estimates of $\hat{\lambda}$ are increasingly skewed at larger and smaller input levels of phylogenetic signal and at smaller tree sizes, vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range.
+By contrast, estimates of $\kappa$ are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$, making $\kappa$ a more reliable statistic.
+We then propose an effect size based on $\kappa$, ($Z_{\kappa}$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes and signal strength, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.
 
@@
 **Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations revealed that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
-First, the mean value of $\hat{\lambda}$ increased as $\lambda$ increased (Fig.
-1 black line); the standard deviation of $\hat{\lambda}$ was largest at intermediate values of $\lambda$ and smallest at extreme values (Fig.
-1 red line); and the distributions of $\hat{\lambda}$ were increasing skewed at extreme values of $\lambda$ but rather normal at intermediate values (Fig.
+First, the mean value of $\hat{\lambda}$ did increase as $\lambda$ increased, but it was negatively-biased (particularly for small tree sizes), and was consistently less than the input $\lambda$ value across most of its range (Fig.
+1 black line).
+Second, the standard deviation of $\hat{\lambda}$ was largest at intermediate values of $\lambda$ and smallest at extreme values, implying that the precision in estimating $\lambda$ varied across the range of input values (Fig.
+1 red line).
+Additionally, standard deviations of $\hat{\lambda}$ were negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ was quite variable, except for cases when $\lambda$ was near or equal to $1$.
+Third, the distributions of $\hat{\lambda}$ were not normal across its range, but became increasingly skewed at more extreme values of $\lambda$ (Fig.
 1 blue line).
-For small tree sizes, it was also clear that distributions were more platykurtik at intermeidate values of $\hat{\lambda}$.
-Second, $\hat{\lambda}$ was a negatively-biased estimate of $\lambda$ for small tree sizes, as the mean value was consistently less than the input $\lambda$ value.
-Third, standard deviations of $\hat{\lambda}$ were negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ was quite variable, except for cases when $\lambda$ was near or equal to $1$.
-As such its statistical properties make it insuitable as an effect size for measuring the strength of phylogenetic signal.
+For small tree sizes, it was also clear that distributions were more platykurtic at intermeidate values of $\hat{\lambda}$.
+Taken together these results reveal that $\hat\lambda$ inconsistently estimated  phylogenetic signal, both across tree sizes and across the range of input values.
+Additional simulations (Supplemental Information) revealed that incorporating $\lambda$ in PGLS anova and regression did not adversely affect the statistical properties of model fitting (type I error, power, parameter estimation); thus, it is reasonable to include $\lambda$ in PGLS as a parameter for tuning the degree of phylogenetic signal in the dependent variables during the analysis.
+However, the statistical properties revealed in Fig.
+1 demonstrate that $\lambda$ is unsuitable as an effect size for measuring the strength of phylogenetic signal in data.
 
@@
 
-**Kappa ($\kappa$) estimates of phylogenetic signal are more stable.** Results for $\kappa$ demonstrated better statistical properties.
-First, mean values increased consistently with $\lambda$ irespective of tree size (Fig.
-2 black line); the standard deviation of $\kappa$ was consistent across tree sizes (Fig.
-2 red line); and $\kappa$ was normally distributed regardless of $\lambda$ or tree size (Fig.
+**Kappa ($\kappa$) estimates of phylogenetic signal are more stable.** Simulation results for $\kappa$ demonstrated that this measure displayed better statistical properties.
+First, as expected, mean values of $\kappa$ increased consistently with increasing signal ($\lambda$) irrespective of tree size (Fig.
+2 black line).
+Additionally, the standard deviation of $\kappa$ was consistent across tree sizes (Fig.
+2 red line), and while it increased with $\lambda$, it was always less than the corresponding mean.
+This finding is perhaps unsurprising, as $\kappa$ is bound by 0, and was never large for small values of $\lambda$.
+Importantly, $\kappa$ was normally distributed across the range of input $\lambda$, and remained consistent in this pattern regardless of tree size (Fig.
 2 blue line).
-The standard deviation of $\kappa$ increased with $\lambda$, somewhat tracking the mean, but was always less than the corresponding means.
-This is perhaps unsurprising, as $\kappa$ is bound by 0, and was never large for small values of $\lambda$.
-Finally, distributions of $\kappa$ displayed slight skewing only at small tree sizes and when $\lambda$ was near or equal to 1.
-This differs from the results of [-@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently.
+In fact, distributions of $\kappa$ displayed slight skewing only at small tree sizes and when $\lambda$ was near or equal to $1$.
+This result differs from those of [@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently.
+These findings reveal that $\kappa$ is more reliable as an estimate of phylogenetic signal.
 
@@
 **Effect sizes from $\kappa$ ($Z_{\kappa}$) better characterize phylogenetic signal.**
-We derived effect sizes (Z-scores) for both $\lambda$ and $\kappa$ (see Methods) and evaluated them relative to known phylogenetic signal.
+To measure the strength of phylogenetic signal on a common scale, we derived effect sizes (Z-scores) for both $\lambda$ and $\kappa$ (see Methods) and evaluated them relative to known phylogenetic signal.
 Both $Z_{\lambda}$ and $Z_{\kappa}$ were associated with input phylogenetic signal ($\lambda$), indicating that both statistics captured the observed signal (Fig.
 3).
-However, effect sizes from $\hat{\lambda}$ made little sense, as they were more strongly associated with tree size than the actual phylogenetic signal in the data (Fig.
+However, effect sizes from $\hat{\lambda}$ made little sense, as they were more strongly associated with tree size than they were with the actual phylogenetic signal in the data (Fig.
 3).
@@
 
-**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** We derived a two-sample test statistic ($\hat{Z}_{12}$) for statistically comparing the strength of phylogenetic signal across datasets (see Methods).
-To demonstrate its utility, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
+**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** To statistically compare the strength of phylogenetic signal across datasets we derived a two-sample test statistic ($\hat{Z}_{12}$: see Methods).
+As an example of its utility, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
 4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
-While both contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V (Fig.
+While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V (Fig.
 4).
-Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020].
+Biologically, this observation may be interpreted by the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
+Thus, greater phylogenetic signal in SA:V is to be expected.
 
@@
 The empirical sampling distribution of $\kappa$ is first transformed via Box-Cox to better adhere to the assumption of normality.
-\hfill\break
 
-For $\lambda$ deriving an effect size is more challenging, it does not have a sampling distribution from which the standard error (and thus confidence intervals) may be obtained.
-Thus, its confidence intervals are generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the compare the putative appropriate model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].
-As such, an effect size for $\lambda$ may be found as:
+For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error (and thus confidence intervals) may be obtained.
+Confidence intervals are tehrefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].
+Thus, an effect size for $\lambda$ may be found as:
 
@@
 However, preliminary investigations found that mapping $\chi^2_{\hat{\lambda}}$ values in this manner did not produce effect sizes that were symmetrical about $Z = 0$, as the mapping was not linear and the log-likelihood profiles can be rather flat for small trees.
+**Is alternative  needed here?**
 
@@
 
-**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from individuals of 305 species (2,781 and 3,371 respectively), from which species means were obtained [@Baken2019; @Baken2020].
-A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the phenotypic dataset.
-The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph].
+**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from as species means from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
+A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.
+The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph], and routines by the authors (**to be incorporated in `geomorph` upon manuscript acceptance**).
 

### 2020-07-16 Manuscript/PNAS-Adams-rticle.tex
11

@@
 
-\author[a]{Dean C. Adams}
+\author[a,1]{Dean C. Adams}
 \author[a,b]{Erica K. Baken}
@@
   \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
-University, Ames, Iowa, USA.}
+University, Ames, Iowa, 50010. USA.}
   \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
-USA.}
+15232. USA.}
 
@@
 comparisons are generally lacking. Here we develop a standardized effect
-size based on \(\kappa\) (\(Z_\kappa\)), which measures the strength of
+size based on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
 phylogenetic signal on a common statistical scale, and provides a
@@
 across datasets. Additionally, we find that a commonly used parameter
-(Pagel's \(\lambda\)) is insuitable for this purpose. Our procedure
+(Pagel's \(\lambda\)) is unsuitable for this purpose. Our procedure
 enables biologists to quantitatively address hypotheses that compare the
@@
 
-\authordeclaration{The authors declare no conflict of interest.}
+\authordeclaration{The authors declare no conflict of interest. \hfill\break
+
+Data deposition: Data for the empirical example may be found on DRYAD:
+\url{doi:10.5061/dryad.b554m44} and \url{doi:10.5061/dryad.59zw3r23m}.
+R-scripts for simulation tests are found on Github: XXX. Computer code
+for implementing the two-sample comparison of effect sizes is found in
+geomorph:
+\url{https://cran.r-project.org/web/packages/geomorph/index.html}
+\hfill\break
+
+\textsuperscript{1}To whom correspondence should be addressed. E-mail:
+\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}}
 
@@
 biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) are therefore compromised. We propose a standardized
-effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the
-strength of phylogenetic signal more reliably than does \(\lambda\), and
-places that signal on a common scale for statistical comparison. We
-develop tests based on \(Z_\kappa\) to provide a mechanism for formally
-comparing the strength of phylogenetic signal across datasets, in much
-the same manner as effect sizes may be used to summarize patterns in
-quantitative meta-analysis. Our approach extends the phylogenetic
-comparative toolkit to address hypotheses that compare the strength of
-phylogenetic signal between various phenotypic traits, even when those
-traits are found in different evolutionary lineages or have different
-units or scales.
+on \(\lambda\) are therefore compromised. As an alternative, we propose
+a standardized effect size based on \(\kappa\), (\(Z_\kappa\)), which
+measures the strength of phylogenetic signal more reliably than does
+\(\lambda\), and places that signal on a common scale for statistical
+comparison. We develop tests based on \(Z_\kappa\) to provide a
+mechanism for formally comparing the strength of phylogenetic signal
+across datasets, in much the same manner as effect sizes may be used to
+summarize patterns in quantitative meta-analysis. Our approach extends
+the phylogenetic comparative toolkit to address hypotheses that compare
+the strength of phylogenetic signal between various phenotypic traits,
+even when those traits are found in different evolutionary lineages or
+have different units or scales.
 \end{abstract}
@@
 26). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
-(ranging from \(0\rightarrow1\)) which is optimized via log-likelihood
-profiling while evaluating the covariation between the dependent and
-independent variables, given the phylogeny (13, 26). To infer whether
-phylogenetic signal differs from no signal or a Brownian motion model of
-evolutionary divergence, the observed model fit using \(\hat\lambda\)
-may be statistically compared to that using \(\lambda=0\) or
-\(\lambda=1\) via likelihood ratio tests (26--28) or confidence limits
-(29). Another widely used measure is Blomberg's \(\kappa\) (14), which
-measures the ratio of observed trait variation to the amount expected
-under Brownian motion. Blomberg's \(\kappa\) can be treated as a test
-statistic by employing a permutation test to generate its sampling
-distribution (14, 18) for determining whether significant phylogenetic
-signal is present in data. Both \(\lambda\) and \(\kappa\) seem
-intuitive to interpret, as a value of \(0\) for both corresponds to no
-phylogenetic signal, while a value of \(1\) corresponds to the amount of
-phylogenetic signal expected under Brownian motion. Thus, it is tempting
-to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
-measure the relative strength of phylogenetic signal, providing an
-estimate of its magnitude for comparison.
+which is optimized via log-likelihood profiling while evaluating the
+covariation between the dependent and independent variables, given the
+phylogeny (13, 26). To infer whether phylogenetic signal differs from no
+signal or a Brownian motion model of evolutionary divergence, the
+observed model fit using \(\hat\lambda\) may be statistically compared
+to that using \(\lambda=0\) or \(\lambda=1\) via likelihood ratio tests
+(26--28) or confidence limits (29). Another widely used measure is
+Blomberg's \(\kappa\) (14), which measures the ratio of observed trait
+variation to the amount expected under Brownian motion. Blomberg's
+\(\kappa\) can be treated as a test statistic by employing a permutation
+test to generate its sampling distribution (14, 18) for determining
+whether significant phylogenetic signal is present in data. Both
+\(\lambda\) and \(\kappa\) seem intuitive to interpret, as a value of
+\(0\) for both corresponds to no phylogenetic signal, while a value of
+\(1\) corresponds to the amount of phylogenetic signal expected under
+Brownian motion. Thus, it is tempting to regard both \(\lambda\) and
+\(\kappa\) as descriptive statistics that measure the relative strength
+of phylogenetic signal, providing an estimate of its magnitude for
+comparison.
 
@@
 given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
-(20)). However, because the ``strength of Brownian motion'' was
+ref. (20)). However, because the ``strength of Brownian motion'' was
 simulated as a varied weighted-average of data simulated on trees with
-\(\lambda=0\) and \(\lambda=1\) and not prescribed values of \(\lambda\)
-(20), interpretation of these patterns is challenging.
+\(\lambda=0\) and \(\lambda=1\) and not as prescribed values of
+\(\lambda\) (20), interpretation of these patterns is challenging.
 
@@
 generate data. This attribute seemed less reasonable based on the
-simulations performed by Münkemüller et al.~(20), as distibutions were
-postively skewed, suggesting the Blomberg's \(\kappa\) might not behave
-as a statistic that follows a noraml distribution. However, because
-their simulations used a weighted combination of simulated phylogenetic
-signal strengths, strong inferences are not possible (and distributional
-attributes were not the intended result of their simulations). For both
-Pagel's \(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical
-moments across a range of \(\lambda\) used to generate data would be
-valuable for adjudicating the relibalilty of these statistics.
-Furthermore, these are statistics that appear to have expected values
-that vary with tree size (20), making comparisons across studies
-challenging. Therefore, transformation of these statistics into
-\(Z\)-scores in the same simulation experiments would allow evaluation
-of the efficacy of each statistic to yield effect sizes that could be
-used for comparative anlyses.
+simulations performed by Münkemüller et al.~(20), which suggested that
+distributions were positively skewed and that Blomberg's \(\kappa\)
+might not behave as a statistic that follows a normal distribution.
+However, because their simulations used a weighted combination of
+simulated phylogenetic signal strengths, strong inferences are not
+possible (and distributional attributes were not the intended result of
+their simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's
+\(\kappa\), evaluation of statistical moments across a range of
+\(\lambda\) used to generate data would be valuable for adjudicating the
+relibalilty of these statistics. Furthermore, these are statistics that
+appear to have expected values that vary with tree size (20), making
+comparisons across studies challenging. Therefore, transformation of
+these statistics into \(Z\)-scores in the same simulation experiments
+would allow evaluation of the efficacy of each statistic to yield effect
+sizes that could be used for comparisons of the strength of phylogenetic
+signal across traits and clades.
 
 Here we used simulation experiments to compare the distributional
-attributes of \(\hat{\lambda}\) and Blomberg's \(\kappa\), plus their
-effect sizes (\(Z\)-scores), across a range of tree size and
-phylogenetic signal strength. We find that estimates of
-\(\hat{\lambda}\) are increasingly skewed at larger and smaller input
-levels of phylogenetic signal, vary widely for a given input value of
-\(\lambda\), and that the precision of \(\hat{\lambda}\) is not constant
-across its range. By contrast, estimates of \(\kappa\) are more
-consistent across tree sizes, and are normally distributed across the
-range of input levels of \(\lambda\). We then propose an effect size
-based on \(\kappa\) (\(Z_{\kappa}\)), which provides consistent
-estimates of the strength of phylogenetic signal across tree sizes, and
-facilitates quantitative comparisons of the relative strength of
-phylogenetic signal across datasets.
+attributes of \(\hat{\lambda}\) and \(\kappa\), plus their effect sizes
+(\(Z\)-scores), across a range of tree size and phylogenetic signal
+strength. We find that estimates of \(\hat{\lambda}\) are increasingly
+skewed at larger and smaller input levels of phylogenetic signal and at
+smaller tree sizes, vary widely for a given input value of \(\lambda\),
+and that the precision of \(\hat{\lambda}\) is not constant across its
+range. By contrast, estimates of \(\kappa\) are more consistent across
+tree sizes, and are normally distributed across the range of input
+levels of \(\lambda\), making \(\kappa\) a more reliable statistic. We
+then propose an effect size based on \(\kappa\), (\(Z_{\kappa}\)), which
+provides consistent estimates of the strength of phylogenetic signal
+across tree sizes and signal strength, and facilitates quantitative
+comparisons of the relative strength of phylogenetic signal across
+datasets.
 
@@
 the distributional expectations of a Bernoulli variable were mostly
-upheld. First, the mean value of \(\hat{\lambda}\) increased as
-\(\lambda\) increased (Fig. 1 black line); the standard deviation of
-\(\hat{\lambda}\) was largest at intermediate values of \(\lambda\) and
-smallest at extreme values (Fig. 1 red line); and the distributions of
-\(\hat{\lambda}\) were increasing skewed at extreme values of
-\(\lambda\) but rather normal at intermediate values (Fig. 1 blue line).
-For small tree sizes, it was also clear that distributions were more
-platykurtik at intermeidate values of \(\hat{\lambda}\). Second,
-\(\hat{\lambda}\) was a negatively-biased estimate of \(\lambda\) for
-small tree sizes, as the mean value was consistently less than the input
-\(\lambda\) value. Third, standard deviations of \(\hat{\lambda}\) were
+upheld. First, the mean value of \(\hat{\lambda}\) did increase as
+\(\lambda\) increased, but it was negatively-biased (particularly for
+small tree sizes), and was consistently less than the input \(\lambda\)
+value across most of its range (Fig. 1 black line). Second, the standard
+deviation of \(\hat{\lambda}\) was largest at intermediate values of
+\(\lambda\) and smallest at extreme values, implying that the precision
+in estimating \(\lambda\) varied across the range of input values (Fig.
+1 red line). Additionally, standard deviations of \(\hat{\lambda}\) were
 negatively associated with tree size, and for tress of 128 species or
 less, \(\hat{\lambda}\) was quite variable, except for cases when
-\(\lambda\) was near or equal to \(1\). As such its statistical
-properties make it insuitable as an effect size for measuring the
-strength of phylogenetic signal.
+\(\lambda\) was near or equal to \(1\). Third, the distributions of
+\(\hat{\lambda}\) were not normal across its range, but became
+increasingly skewed at more extreme values of \(\lambda\) (Fig. 1 blue
+line). For small tree sizes, it was also clear that distributions were
+more platykurtic at intermeidate values of \(\hat{\lambda}\). Taken
+together these results reveal that \(\hat\lambda\) inconsistently
+estimated phylogenetic signal, both across tree sizes and across the
+range of input values. Additional simulations (Supplemental Information)
+revealed that incorporating \(\lambda\) in PGLS anova and regression did
+not adversely affect the statistical properties of model fitting (type I
+error, power, parameter estimation); thus, it is reasonable to include
+\(\lambda\) in PGLS as a parameter for tuning the degree of phylogenetic
+signal in the dependent variables during the analysis. However, the
+statistical properties revealed in Fig. 1 demonstrate that \(\lambda\)
+is unsuitable as an effect size for measuring the strength of
+phylogenetic signal in data.
 
@@
 \textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
-stable.} Results for \(\kappa\) demonstrated better statistical
-properties. First, mean values increased consistently with \(\lambda\)
-irespective of tree size (Fig. 2 black line); the standard deviation of
-\(\kappa\) was consistent across tree sizes (Fig. 2 red line); and
-\(\kappa\) was normally distributed regardless of \(\lambda\) or tree
-size (Fig. 2 blue line). The standard deviation of \(\kappa\) increased
-with \(\lambda\), somewhat tracking the mean, but was always less than
-the corresponding means. This is perhaps unsurprising, as \(\kappa\) is
-bound by 0, and was never large for small values of \(\lambda\).
-Finally, distributions of \(\kappa\) displayed slight skewing only at
-small tree sizes and when \(\lambda\) was near or equal to 1. This
-differs from the results of (20), where skewing appears to be from
-combining random values generated independently.
+stable.} Simulation results for \(\kappa\) demonstrated that this
+measure displayed better statistical properties. First, as expected,
+mean values of \(\kappa\) increased consistently with increasing signal
+(\(\lambda\)) irrespective of tree size (Fig. 2 black line).
+Additionally, the standard deviation of \(\kappa\) was consistent across
+tree sizes (Fig. 2 red line), and while it increased with \(\lambda\),
+it was always less than the corresponding mean. This finding is perhaps
+unsurprising, as \(\kappa\) is bound by 0, and was never large for small
+values of \(\lambda\). Importantly, \(\kappa\) was normally distributed
+across the range of input \(\lambda\), and remained consistent in this
+pattern regardless of tree size (Fig. 2 blue line). In fact,
+distributions of \(\kappa\) displayed slight skewing only at small tree
+sizes and when \(\lambda\) was near or equal to \(1\). This result
+differs from those of (20), where skewing appears to be from combining
+random values generated independently. These findings reveal that
+\(\kappa\) is more reliable as an estimate of phylogenetic signal.
 
@@
 \textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
-characterize phylogenetic signal.} We derived effect sizes (Z-scores)
-for both \(\lambda\) and \(\kappa\) (see Methods) and evaluated them
-relative to known phylogenetic signal. Both \(Z_{\lambda}\) and
-\(Z_{\kappa}\) were associated with input phylogenetic signal
-(\(\lambda\)), indicating that both statistics captured the observed
-signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
-little sense, as they were more strongly associated with tree size than
-the actual phylogenetic signal in the data (Fig. 3). By contrast,
-\(Z_{\kappa}\) was much more consistent accross tree sizes, and
-exhibited a much stronger association with phylogenetic signal strength
-as compared to tree size (Fig. 3). Thus between the two statistics,
-\(Z_{\kappa}\) is a more reliable measure of the strength of
-phylogenetic signal, and may be used to compare levels of phylogenetic
-signal across datasets.
+characterize phylogenetic signal.} To measure the strength of
+phylogenetic signal on a common scale, we derived effect sizes
+(Z-scores) for both \(\lambda\) and \(\kappa\) (see Methods) and
+evaluated them relative to known phylogenetic signal. Both
+\(Z_{\lambda}\) and \(Z_{\kappa}\) were associated with input
+phylogenetic signal (\(\lambda\)), indicating that both statistics
+captured the observed signal (Fig. 3). However, effect sizes from
+\(\hat{\lambda}\) made little sense, as they were more strongly
+associated with tree size than they were with the actual phylogenetic
+signal in the data (Fig. 3). By contrast, \(Z_{\kappa}\) was much more
+consistent accross tree sizes, and exhibited a much stronger association
+with phylogenetic signal strength as compared to tree size (Fig. 3).
+Thus between the two statistics, \(Z_{\kappa}\) is a more reliable
+measure of the strength of phylogenetic signal, and may be used to
+compare levels of phylogenetic signal across datasets.
 
@@
 \textbf{A test statistic (\(\hat{Z}_{12}\)) allows meaningful
-comparisons across datasets.} We derived a two-sample test statistic
-(\(\hat{Z}_{12}\)) for statistically comparing the strength of
-phylogenetic signal across datasets (see Methods). To demonstrate its
-utility, we compared \(Z_{\kappa}\) for two ecologically-relevant traits
-in plethodontid salamander (Fig. 4): surface area to volume ratios
-(SA:V) and relative body width (\(\frac{BW}{SVL}\)) (33, 34). While both
+comparisons across datasets.} To statistically compare the strength of
+phylogenetic signal across datasets we derived a two-sample test
+statistic (\(\hat{Z}_{12}\): see Methods). As an example of its utility,
+we compared \(Z_{\kappa}\) for two ecologically-relevant traits in
+plethodontid salamander (Fig. 4): surface area to volume ratios (SA:V)
+and relative body width (\(\frac{BW}{SVL}\)) (33, 34). While both traits
 contained significant phylogenetic signal, tests based on
@@
 significantly stronger in SA:V (Fig. 4). Biologically, this observation
-corresponds with the fact that tropical species -- which form a
+may be interpreted by the fact that tropical species -- which form a
 monophyletic group within plethodontids -- display greater variation in
-SA:V which covaries with disparity in their climatic niches (34).
+SA:V, which covaries with disparity in their climatic niches (34). Thus,
+greater phylogenetic signal in SA:V is to be expected.
 
@@
 first transformed via Box-Cox to better adhere to the assumption of
-normality. \hfill\break
+normality.
 
-For \(\lambda\) deriving an effect size is more challenging, it does not
-have a sampling distribution from which the standard error (and thus
-confidence intervals) may be obtained. Thus, its confidence intervals
-are generated for the values of \(\lambda\) that intersect the
-log-likehihood profile for corresponding percentiles of the \(\chi^2\)
-distribution used to compare the compare the putative appropriate model
-to a null model with \(\lambda = 0\) {[}\textbf{add ref: MLC thinks
-Boettiger paper?}{]}. As such, an effect size for \(\lambda\) may be
+For \(\lambda\), deriving an effect size is more challenging, as
+\(\lambda\) does not have a sampling distribution from which the
+standard error (and thus confidence intervals) may be obtained.
+Confidence intervals are tehrefore generated for the values of
+\(\lambda\) that intersect the log-likehihood profile for corresponding
+percentiles of the \(\chi^2\) distribution used to compare the putative
+model to a null model with \(\lambda = 0\) {[}\textbf{add ref: MLC
+thinks Boettiger paper?}{]}. Thus, an effect size for \(\lambda\) may be
 found as:
@@
 not linear and the log-likelihood profiles can be rather flat for small
-trees.
+trees. \textbf{Is alternative needed here?}
 
@@
 \textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
-relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
-individuals of 305 species (2,781 and 3,371 respectively), from which
-species means were obtained (33, 34). A time-dated molecular phylogeny
-for the group (47) was then pruned to match the species in the
-phenotypic dataset. The phylogenetic signal in each trait was then
-characterized using \(\kappa\), which was converted to its effect size
-(\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (48, 49).
+relative body width (\(\frac{BW}{SVL}\)) measures were obtained from as
+species means from individuals of 305 species, from which species means
+were obtained (33, 34). A time-dated molecular phylogeny for the group
+(47) was pruned to match the species in the phenotypic dataset. The
+phylogenetic signal in each trait was then characterized using
+\(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
+\texttt{geomorph} 3.3.1 (48, 49), and routines by the authors
+(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
+acceptance}).
 

### 2020-07-16 Manuscript/PNAS-Adams-01.Rmd
pnas-01

--- 2020-07-16 Manuscript/PNAS-Adams-rticle.Rmd
@@
 abstract: |
-  Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits.
-However, analytical tools for such comparisons have largely remained underdeveloped.
-Here we evaluate the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal.
-We find that $\lambda$ behaves as a Bernoulli random variable, and that estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
-Further, the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised.
+  Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped.
+Here we evaluate the efficacy of of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
+We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
+Further, the precision of $\lambda$ varies with input signal.
+Another measure, Blomberg's $\kappa$, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal.
+However, that relationship is decidedly nonlinear.
+Thus, neither $\lambda$ nor $\kappa$ are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets.
 As an alternative, we propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
@@
 significance: |
-  Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across traits, but analytical tools for these comparisons are generally lacking.
-Here we develop a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal on a common statistical scale, and provides a mechanism for formally comparing the strength of phylogenetic signal across datasets.
-Additionally, we find that a commonly used parameter (Pagel's $\lambda$) is unsuitable for this purpose.
-Our procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
+  Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking.
+Here we develop a standardized effect size, $Z_\kappa$, which measures the strength of phylogenetic signal on a common statistical scale.
+We also provide a test statistic, $\hat{Z}_{12}$, for comparing the strength of phylogenetic signal across datasets.
+We find that two commonly used parameters (Pagel's $\lambda$ and Blomberg's $\kappa$), not converted to effect sizes, are unsuitable for this purpose.
+Our effect-size procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
 
@@
 Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
-Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @ButlerKing2004; @MartinsHansen1997; @OMeara_et_al2006; @RevellHarmon2008; @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b].
+Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @MartinsHansen1997; @OMeara_et_al2006;  @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b].
 These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
@@
 To infer whether phylogenetic signal differs from no signal or a Brownian motion model of evolutionary divergence, the observed model fit using $\hat\lambda$ may be statistically compared to that using  $\lambda=0$ or $\lambda=1$ via likelihood ratio tests [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or confidence limits [@Vandelook2019].
-Another widely used measure is Blomberg's $\kappa$ [@Blomberg_et_al2003], which measures the ratio of observed trait variation to the amount expected under Brownian motion.
+
+Another widely used measure of phylogenetic signal is Blomberg's $\kappa$ [@Blomberg_et_al2003], which characterizes phylogenetic signal as the ratio of observed trait variation to the amount of variation expected under Brownian motion.
 Blomberg's $\kappa$ can be treated as a test statistic by employing a permutation test to generate its sampling distribution [@Blomberg_et_al2003; @Adams2014a] for determining whether significant phylogenetic signal is present in data.
@@
 For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ follows a Bernoulli distribution (**add ref**); i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
-Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to treesize; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 (**add same ref**).
+Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to tree size; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 (**add same ref**).
 Prior work [@Munkemuller_et_al2012] seems to support some of these conjectures, based superficially on statistical moments for a given tree size (mean, variance, skewness, and kurtosis; see Fig.
@@
 
-By contrast, Blomberg's $\kappa$, which is positively ubounded, might be expected to follow a normal distribution (**add ref**).
-Thus we might expect that $\kappa$ is symmetrically distributed across different strengths of phylogenetic signal, for any $\lambda$ used to generate data.
+By contrast, for Blomberg's $\kappa$, which is positively ubounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
@@
 Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
-Therefore, transformation of these statistics into $Z$-scores in the same simulation experiments would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and clades.
+Therefore, transformation of these statistics into $Z$-scores would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and lineages.
 
-Here we used simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
+Here we use simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
 We find that estimates of $\hat{\lambda}$ are increasingly skewed at larger and smaller input levels of phylogenetic signal and at smaller tree sizes, vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range.
@@
 
-**Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations revealed that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
-First, the mean value of $\hat{\lambda}$ did increase as $\lambda$ increased, but it was negatively-biased (particularly for small tree sizes), and was consistently less than the input $\lambda$ value across most of its range (Fig.
+**Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
+First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases, but it is negatively-biased (particularly for small tree sizes), and is consistently less than the input $\lambda$ value across most of its range (Fig.
 1 black line).
-Second, the standard deviation of $\hat{\lambda}$ was largest at intermediate values of $\lambda$ and smallest at extreme values, implying that the precision in estimating $\lambda$ varied across the range of input values (Fig.
+Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values, implying that the precision in estimating $\lambda$ varies across the range of input values (Fig.
 1 red line).
-Additionally, standard deviations of $\hat{\lambda}$ were negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ was quite variable, except for cases when $\lambda$ was near or equal to $1$.
-Third, the distributions of $\hat{\lambda}$ were not normal across its range, but became increasingly skewed at more extreme values of $\lambda$ (Fig.
+Additionally, standard deviations of $\hat{\lambda}$ are negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ are quite variable, except for cases when $\lambda$ is near or equal to $1$.
+Third, the distributions of $\hat{\lambda}$ are not normal across its range, but become increasingly skewed at more extreme values of $\lambda$ (Fig.
 1 blue line).
-For small tree sizes, it was also clear that distributions were more platykurtic at intermeidate values of $\hat{\lambda}$.
-Taken together these results reveal that $\hat\lambda$ inconsistently estimated  phylogenetic signal, both across tree sizes and across the range of input values.
-Additional simulations (Supplemental Information) revealed that incorporating $\lambda$ in PGLS anova and regression did not adversely affect the statistical properties of model fitting (type I error, power, parameter estimation); thus, it is reasonable to include $\lambda$ in PGLS as a parameter for tuning the degree of phylogenetic signal in the dependent variables during the analysis.
-However, the statistical properties revealed in Fig.
-1 demonstrate that $\lambda$ is unsuitable as an effect size for measuring the strength of phylogenetic signal in data.
+For small tree sizes, it is also clear that distributions are more platykurtic at intermediate values of $\hat{\lambda}$.
+Taken together these results reveal that $\hat\lambda$ inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
+Additional simulations (Supplemental Information) reveal that incorporating $\hat\lambda$ in PGLS anova and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
+Thus, it is reasonable to incorporate $\hat\lambda$ in PGLS as a parameter for tuning the degree of phylogenetic signal in the dependent variables during the analysis.
+However, the statistical properties shown in Fig.
+1 demonstrate that $\lambda$ is unsuitable as an effect size for measuring the strength of phylogenetic signal in data, or for comparing that signal across datasets.
 
-![Patterns in $\lambda$.<span data-label="fig:f1"></span>](new.fig.1.temp.png)
+![Response of Pagel's $\lambda$ to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$.
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](new.fig.1.temp.png)
 
-**Kappa ($\kappa$) estimates of phylogenetic signal are more stable.** Simulation results for $\kappa$ demonstrated that this measure displayed better statistical properties.
-First, as expected, mean values of $\kappa$ increased consistently with increasing signal ($\lambda$) irrespective of tree size (Fig.
+**Kappa ($\kappa$) estimates of phylogenetic signal are more reliable.** Simulation results for $\hat\kappa$ demonstrate that $\hat\kappa$ displays better statistical properties.
+First, as expected, mean values of $\hat\kappa$ increase with increasing signal ($\lambda$) irrespective of tree size, though the increase does not scale linearly with input levels of phylogenetic signal (Fig.
 2 black line).
-Additionally, the standard deviation of $\kappa$ was consistent across tree sizes (Fig.
-2 red line), and while it increased with $\lambda$, it was always less than the corresponding mean.
-This finding is perhaps unsurprising, as $\kappa$ is bound by 0, and was never large for small values of $\lambda$.
-Importantly, $\kappa$ was normally distributed across the range of input $\lambda$, and remained consistent in this pattern regardless of tree size (Fig.
+Additionally, the standard deviation of $\hat\kappa$ is consistent across tree sizes (Fig.
+2 red line), and while it increases with $\lambda$, it is always less than the corresponding mean.
+This finding is perhaps unsurprising, as $\hat\kappa$ is lower-bounded by 0, and is never large for small values of $\lambda$.
+Importantly, $\hat\kappa$ is normally distributed across the range of input $\lambda$, and remains consistent in this pattern regardless of tree size (Fig.
 2 blue line).
-In fact, distributions of $\kappa$ displayed slight skewing only at small tree sizes and when $\lambda$ was near or equal to $1$.
+In fact, distributions of $\hat\kappa$ display slight skewing only at small tree sizes and when $\lambda$ is near or equal to $1$.
 This result differs from those of [@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently.
-These findings reveal that $\kappa$ is more reliable as an estimate of phylogenetic signal.
+These findings reveal that while $\kappa$ is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
 
-![Patterns in $\kappa$.<span data-label="fig:f2"></span>](new.fig.2.temp.png)
+![Response of Blomberg's $\hat\kappa$ to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$.
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](new.fig.2.temp.png)
 
 **Effect sizes from $\kappa$ ($Z_{\kappa}$) better characterize phylogenetic signal.**
-To measure the strength of phylogenetic signal on a common scale, we derived effect sizes (Z-scores) for both $\lambda$ and $\kappa$ (see Methods) and evaluated them relative to known phylogenetic signal.
-Both $Z_{\lambda}$ and $Z_{\kappa}$ were associated with input phylogenetic signal ($\lambda$), indicating that both statistics captured the observed signal (Fig.
-3).
-However, effect sizes from $\hat{\lambda}$ made little sense, as they were more strongly associated with tree size than they were with the actual phylogenetic signal in the data (Fig.
-3).
-By contrast, $Z_{\kappa}$ was much more consistent accross tree sizes, and exhibited a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
-3).
-Thus between the two statistics, $Z_{\kappa}$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets.
-
-![Patterns in Z.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
-
-**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** To statistically compare the strength of phylogenetic signal across datasets we derived a two-sample test statistic ($\hat{Z}_{12}$: see Methods).
-As an example of its utility, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
-4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
-While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V (Fig.
-4).
-Biologically, this observation may be interpreted by the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
-Thus, greater phylogenetic signal in SA:V is to be expected.
-
-![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
-(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](new.fig.4.png)
-
-# Discussion
-
-**To be edited.** address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
-
-It is common in comparative evolutionary studies to characterize the phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has generated trait covariation among taxa.
-However, while numerous analytical approaches may be used to quantify phylogenetic signal [e.g., @Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999], methods that explicitly measure the strength of phylogenetic signal, or facilitate comparisons among datasets, have remained underdeveloped.
-In this study, we evaluated the precision of one common measure, Pagel's $\lambda$, and explored its efficacy for characterizing the strength of phylogenetic signal in phenotypic data.
-Using computer simulations, we found that the precision of $\lambda$ increased with increasing sample sizes; a pattern noted previously [@Boettiger_et_al2012], and one that conformed with parametric statistical theory [@Cohen1988].
-However, we also found that vastly different $\lambda$ estimates could be obtained from data containing the same level of phylogenetic signal, and that similar $\lambda$ estimates may be obtained from data containing differing levels of phylogenetic signal.
-Further, the precision of $\lambda$ varied with the strength of phylogenetic signal, where lower precision was observed when in data whose phylogenetic signal was of intemediate strength.
-From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and that biological interpretations of the strength of signal based on this parameter may innacurately characterize such effects.
-
-As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal.
-$Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
-We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
-First, $Z_\kappa$ was more precise than $Z_\lambda$, and precision was more consistent across the range of input levels of phylogenetic signal.
-Additionally, values of $Z_\kappa$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa$ and $\lambda_{in}$.
-Thus, $Z_\kappa$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data.
-We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_\kappa$ as a statistical measure of this effect.
-
-Based on the effect size $Z_\kappa$, we then proposed a two-sample test, which provides means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another, via a hypothesis test.
-Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $\kappa$ [e.g.,  @Blomberg_et_al2003].
-However, $\kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information].
-Thus, $\kappa$ should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
-By contrast, standardizing $\kappa$ ($Z_\kappa$, via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
-Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
-As such, our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
-
-One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
-This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
-For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
-In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
-Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
-Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
-
-This study is not the first to compare $\lambda$ and $\kappa$ for their ability as statistics to measure phylogenetic signal.
-Our results for $\lambda$ and $\kappa$ values are consistent with those found in the simulations performed by Münkemüller et al.
-[-@Munkemuller_et_al2012], but that study investigated type I error rates and statistical power, finding that $\lambda$ performed better in both regards, irrespective of species number in trees.
-Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent $\kappa$ at intermediate phylogenetic signal strengths was observed [see Fig.
-2, @Munkemuller_et_al2012].
-Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
-Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
-An interesting question is whether an inflated $\kappa$ value leads to an inflated $Z_\kappa$ or does a tendency of a particular tree to inflate estimates of $\kappa$ also inflate the values in random permutations of a test, in which case $Z_\kappa$ is robust to polytomies?
-We repeated the analyses in Figure 4, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (see Supporting Information).
-This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size.
-Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
-
-Phylogenetic signal can be thought of as both an attribute to be measured in the data and a parameter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
-As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
-However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a statistic to describe the amount of phylogenetic signal in the data.
-Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
-Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
-By contrast, $\kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_\kappa$).
-Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
-With low coefficients of variation for $Z_\kappa$ (Figure 4), it is difficult to imagine that a hypothesis test can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
-
-# Methods
-
-**Derivation of effect sizes.** Statistically, a standardized effect size may be found as:
+To measure the strength of phylogenetic signal on a common scale, here we propose effect sizes (Z-scores) for both $\lambda$ and $\kappa$. Statistically, a standardized effect size may be found as:
 
@@
 
-For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error (and thus confidence intervals) may be obtained.
-Confidence intervals are tehrefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].
+For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error and confidence intervals may be obtained.
+Confidence intervals are therefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].
 Thus, an effect size for $\lambda$ may be found as:
@@
 where, $\hat{\lambda}$ is the maximized likelihood value of $\lambda$ and $\chi^{2}_{\hat{\lambda}}$ is the likelihood ratio statistic for the value.
-Note that alternative formulations could be envisioned: $Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}$, where $d$ is a binary value $(-1,1)$ to indicate a direction based on whether $\hat{\lambda}$ is below or above a critical value of $\lambda$, for a quantile from a $\chi^{2}$ distribution at a probability of 0.5.
-However, preliminary investigations found that mapping $\chi^2_{\hat{\lambda}}$ values in this manner did not produce effect sizes that were symmetrical about $Z = 0$, as the mapping was not linear and the log-likelihood profiles can be rather flat for small trees.
-**Is alternative  needed here?**
 
-**Derivation of two-sample test statistic.** To compare the strength of phylogenetic signal across datasets, a two-sample test statistic may be calculated as:
+Here we evaluate the ability of $Z_{\lambda}$ and $Z_{\kappa}$ to characterize known levels of phylogenetic signal.
+Both $Z_{\lambda}$ and $Z_{\kappa}$ are associated with input phylogenetic signal ($\lambda$), indicating that both statistics capture the observed signal (Fig.
+3).
+However, effect sizes from $\hat{\lambda}$ made little sense, as they are more strongly associated with tree size than they are with the actual phylogenetic signal in the data (Fig.
+3).
+By contrast, $Z_{\kappa}$ is much more consistent accross tree sizes, and increases more linearly with incrasing levels of phylogenetic signal.
+Additionally, $Z_{\kappa}$ exhibits a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
+3), and its standard deviation across input signal is more consistent.
+This implies that similar levels of precision are found with $Z_{\kappa}$ across the range of input values.
+Thus between the two statistics, $Z_{\kappa}$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets.
+
+![Response of effect sizes $Z_\lambda$ and $Z_\kappa$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
+
+**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** To statistically compare the strength of phylogenetic signal across datasets we propose a two-sample test statistic ($\hat{Z}_{12}$).
+Based on statistical theory, a two-sample test statistic may be calculated as:
 
@@
 Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b].
+
+To demonstrate the utility of $\hat{Z}_12$, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
+4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
+While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V ($\hat{Z}_{12}=4.13$; $P=0.000036$: Fig.
+4).
+Biologically, this observation may be interpreted by the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
+Thus, greater phylogenetic signal in SA:V is to be expected.
+
+![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](new.fig.4.png)
+
+# Discussion
+
+It is common in comparative evolutionary studies to characterize the phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has generated trait covariation among taxa.
+However, while numerous analytical approaches may be used to quantify phylogenetic signal [e.g., @Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999], methods that explicitly measure the strength of phylogenetic signal, or facilitate comparisons among datasets, have remained underdeveloped.
+We evaluated the precision of one common measure, Pagel's $\lambda$, and explored its efficacy for characterizing the strength of phylogenetic signal in phenotypic data.
+Using computer simulations, we found that $\lambda$ behaves as a Bernoulli random variable, with estimates that are increasingly skewed at larger and smaller input levels of phylogenetic signal.
+Further, the precision of $\lambda$ in estimating actual levels of phylogenetic signal varies with both tree size (see also ref.
+@Boettiger_et_al2012) and input levels of phylogentetic signal.
+From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and should not be used as an effect size for comparing the degree of phylogenetic signal between datasets.
+
+As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal.
+$Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
+We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
+First, values of $Z_\kappa$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa$ and input signal.
+Additionally, the precision of $Z_{\kappa}$ was more consistent across the range of input levels of phylogenetic signal.
+Thus, $Z_\kappa$ is a more reliable measure of the relative strength of phylogenetic signal, and places that effect on a common and comparable scale.
+We therefore recommend that future studies interested in evaluating the strength of phylogenetic signal incorporate $Z_\kappa$ as a statistical measure of this effect.
+
+Next we proposed a two-sample test ($\hat{Z}_{12}$), which provides a formal statistical procedure for determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another.
+Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $\kappa$ [e.g.,  @Blomberg_et_al2003].
+However, because $\kappa$ does not scale linearly with input levels of phylogenetic signal (Fig.
+2), and its variance increases with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012], it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
+By contrast, standardizing $\kappa$ to $Z_\kappa$ via equation 2 alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
+As such, our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
+
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
+Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+This means that the strength of phylogenetic signal may be compared across datasets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
+For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
+In fact, our empirical example provided just such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
+Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
+
+This study is not the first to compare $\lambda$ and $\kappa$ for their ability as statistics to measure phylogenetic signal.
+Our results for $\lambda$ and $\kappa$ values are consistent with those found in the simulations performed by Münkemüller et al.
+[-@Munkemuller_et_al2012], but that study investigated type I error rates and statistical power, finding that $\lambda$ performed better in both regards, irrespective of species number in trees.
+Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent $\kappa$ at intermediate phylogenetic signal strengths was observed (Fig.
+2 of ref.
+[@Munkemuller_et_al2012]).
+Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
+Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
+An interesting question is whether an inflated $\kappa$ value leads to an inflated $Z_\kappa$ or does a tendency of a particular tree to inflate estimates of $\kappa$ also inflate the values in random permutations of a test, in which case $Z_\kappa$ is robust to polytomies?
+We repeated the analyses in Figs.
+1 & 2, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (Supporting Information).
+This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size.
+Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+
+Phylogenetic signal can be thought of as both an attribute to be measured in the data and a parameter that can be tuned to account for the phylogenetic non-independence among observations, for analysis of the data.
+As such, $\lambda$ is appealing, as a statistic that potentially fulfills both roles.
+However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a statistic to describe the amount of phylogenetic signal in the data.
+Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
+Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
+By contrast, $\kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_\kappa$).
+Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
+Given the greater consistency in estimates of $Z_\kappa$ across tree sizes and input signal, it is difficult to imagine a hypothesis test that can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
+
+# Methods
 

### 2020-07-16 Manuscript/PNAS-Adams-01.tex
pnas-01

--- 2020-07-16 Manuscript/PNAS-Adams-rticle.tex
@@
 \significancestatement{Evolutionary biologists wish to quantify and compare the strength of
-phylogenetic signal across traits, but analytical tools for these
+phylogenetic signal across datasets, but analytical tools for these
 comparisons are generally lacking. Here we develop a standardized effect
-size based on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
-phylogenetic signal on a common statistical scale, and provides a
-mechanism for formally comparing the strength of phylogenetic signal
-across datasets. Additionally, we find that a commonly used parameter
-(Pagel's \(\lambda\)) is unsuitable for this purpose. Our procedure
-enables biologists to quantitatively address hypotheses that compare the
+size, \(Z_\kappa\), which measures the strength of phylogenetic signal
+on a common statistical scale. We also provide a test statistic,
+\(\hat{Z}_{12}\), for comparing the strength of phylogenetic signal
+across datasets. We find that two commonly used parameters (Pagel's
+\(\lambda\) and Blomberg's \(\kappa\)), not converted to effect sizes,
+are unsuitable for this purpose. Our effect-size procedure enables
+biologists to quantitatively address hypotheses that compare the
 strength of phylogenetic signal between various phenotypic traits, even
@@
 Macroevolutionary studies frequently characterize the phylogenetic
-signal in phenotypes, and wish to compare the strength of that signal
-across traits. However, analytical tools for such comparisons have
-largely remained underdeveloped. Here we evaluate the efficacy of one
-commonly used parameter (Pagel's \(\lambda\)) to estimate the strength
-of phylogenetic signal in phenotypic traits, and evaluate the degree to
-which \(\lambda\) correctly identifies known levels of phylogenetic
-signal. We find that \(\lambda\) behaves as a Bernoulli random variable,
-and that estimates are increasingly skewed at larger and smaller input
-levels of phylogenetic signal. Further, the precision of \(\lambda\) in
-estimating actual levels of phylogenetic signal is often inaccurate, and
-biological interpretations of the strength of phylogenetic signal based
-on \(\lambda\) are therefore compromised. As an alternative, we propose
-a standardized effect size based on \(\kappa\), (\(Z_\kappa\)), which
-measures the strength of phylogenetic signal more reliably than does
-\(\lambda\), and places that signal on a common scale for statistical
-comparison. We develop tests based on \(Z_\kappa\) to provide a
-mechanism for formally comparing the strength of phylogenetic signal
-across datasets, in much the same manner as effect sizes may be used to
-summarize patterns in quantitative meta-analysis. Our approach extends
-the phylogenetic comparative toolkit to address hypotheses that compare
-the strength of phylogenetic signal between various phenotypic traits,
-even when those traits are found in different evolutionary lineages or
-have different units or scales.
+signal in phenotypes, however, analytical tools for comparing the
+strength of that signal across traits remain largely underdeveloped.
+Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly
+estimate the strength of phylogenetic signal in phenotypic traits across
+a range of input values. We find that \(\lambda\) behaves as a Bernoulli
+random variable, where estimates are increasingly skewed at larger and
+smaller input levels of phylogenetic signal. Further, the precision of
+\(\lambda\) varies with input signal. Another measure, Blomberg's
+\(\kappa\), is more consistent across a range of tree sizes, and
+exhibits a positive relationship with input levels of phylogenetic
+signal. However, that relationship is decidedly nonlinear. Thus, neither
+\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
+the strength of phylogenetic signal, and comparing that signal across
+datasets. As an alternative, we propose a standardized effect size based
+on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
+phylogenetic signal more reliably than does \(\lambda\), and places that
+signal on a common scale for statistical comparison. We develop tests
+based on \(Z_\kappa\) to provide a mechanism for formally comparing the
+strength of phylogenetic signal across datasets, in much the same manner
+as effect sizes may be used to summarize patterns in quantitative
+meta-analysis. Our approach extends the phylogenetic comparative toolkit
+to address hypotheses that compare the strength of phylogenetic signal
+between various phenotypic traits, even when those traits are found in
+different evolutionary lineages or have different units or scales.
 \end{abstract}
@@
 methods} (PCMs): a suite of analytical tools that condition trends in
-the data on the phylogenetic relatedness of observations (3--12). These
+the data on the phylogenetic relatedness of observations (3--10). These
 methods are predicated on the notion that phylogenetic signal -- the
 tendancy for closely related species to display similar trait values --
-is present in cross-species datasets (1, 13, 14). Indeed, under numerous
+is present in cross-species datasets (1, 11, 12). Indeed, under numerous
 evolutionary models, phylogenetic signal is to be expected, as
 stochastic character change along the hierarchical structure of the tree
-of life generates trait covaration among related taxa (1, 14, 15).
+of life generates trait covaration among related taxa (1, 12, 13).
 
 Several analytical tools have been developed to quantify phylogenetic
-signal in phenotypic datasets (13, 14, 16--19), and their statistical
+signal in phenotypic datasets (11, 12, 14--17), and their statistical
 properties -- namely type I error rates and statistical power -- have
 been investigated to determine under what conditions phylogenetic signal
-can be detected (15, 18, 20--25). One of the most widely used methods
-for characterizing phylogenetic signal is Pagel's \(\lambda\) (13),
+can be detected (13, 16, 18--23). One of the most widely used methods
+for characterizing phylogenetic signal is Pagel's \(\lambda\) (11),
 which transforms the lengths of the internal branches of the phylogeny
-to improve the fit of data to the phylogeny via maximum likelihood (13,
-26). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
+to improve the fit of data to the phylogeny via maximum likelihood (11,
+24). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
 which is optimized via log-likelihood profiling while evaluating the
 covariation between the dependent and independent variables, given the
-phylogeny (13, 26). To infer whether phylogenetic signal differs from no
+phylogeny (11, 24). To infer whether phylogenetic signal differs from no
 signal or a Brownian motion model of evolutionary divergence, the
@@
 to that using \(\lambda=0\) or \(\lambda=1\) via likelihood ratio tests
-(26--28) or confidence limits (29). Another widely used measure is
-Blomberg's \(\kappa\) (14), which measures the ratio of observed trait
-variation to the amount expected under Brownian motion. Blomberg's
-\(\kappa\) can be treated as a test statistic by employing a permutation
-test to generate its sampling distribution (14, 18) for determining
-whether significant phylogenetic signal is present in data. Both
-\(\lambda\) and \(\kappa\) seem intuitive to interpret, as a value of
-\(0\) for both corresponds to no phylogenetic signal, while a value of
-\(1\) corresponds to the amount of phylogenetic signal expected under
-Brownian motion. Thus, it is tempting to regard both \(\lambda\) and
-\(\kappa\) as descriptive statistics that measure the relative strength
-of phylogenetic signal, providing an estimate of its magnitude for
-comparison.
+(24--26) or confidence limits (27).
+
+Another widely used measure of phylogenetic signal is Blomberg's
+\(\kappa\) (12), which characterizes phylogenetic signal as the ratio of
+observed trait variation to the amount of variation expected under
+Brownian motion. Blomberg's \(\kappa\) can be treated as a test
+statistic by employing a permutation test to generate its sampling
+distribution (12, 16) for determining whether significant phylogenetic
+signal is present in data. Both \(\lambda\) and \(\kappa\) seem
+intuitive to interpret, as a value of \(0\) for both corresponds to no
+phylogenetic signal, while a value of \(1\) corresponds to the amount of
+phylogenetic signal expected under Brownian motion. Thus, it is tempting
+to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
+measure the relative strength of phylogenetic signal, providing an
+estimate of its magnitude for comparison.
 
@@
 values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
-comparative sense (30--32). Nonetheless, an important question that has
+comparative sense (28--30). Nonetheless, an important question that has
 yet to be considered is whether these statistics are or can be converted
@@
 values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
-range of \(\lambda\) with respect to treesize; the distribution of
+range of \(\lambda\) with respect to tree size; the distribution of
 \(\hat{\lambda}\) would be symmetric at intermediate values of
@@
 intermediate values of \(\lambda\), becoming more leptokurtic toward 0
-and 1 (\textbf{add same ref}). Prior work (20) seems to support some of
+and 1 (\textbf{add same ref}). Prior work (18) seems to support some of
 these conjectures, based superficially on statistical moments for a
 given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
-ref. (20)). However, because the ``strength of Brownian motion'' was
+ref. (18)). However, because the ``strength of Brownian motion'' was
 simulated as a varied weighted-average of data simulated on trees with
 \(\lambda=0\) and \(\lambda=1\) and not as prescribed values of
-\(\lambda\) (20), interpretation of these patterns is challenging.
+\(\lambda\) (18), interpretation of these patterns is challenging.
 
-By contrast, Blomberg's \(\kappa\), which is positively ubounded, might
-be expected to follow a normal distribution (\textbf{add ref}). Thus we
-might expect that \(\kappa\) is symmetrically distributed across
-different strengths of phylogenetic signal, for any \(\lambda\) used to
-generate data. This attribute seemed less reasonable based on the
-simulations performed by Münkemüller et al.~(20), which suggested that
-distributions were positively skewed and that Blomberg's \(\kappa\)
-might not behave as a statistic that follows a normal distribution.
-However, because their simulations used a weighted combination of
-simulated phylogenetic signal strengths, strong inferences are not
-possible (and distributional attributes were not the intended result of
-their simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's
-\(\kappa\), evaluation of statistical moments across a range of
-\(\lambda\) used to generate data would be valuable for adjudicating the
-relibalilty of these statistics. Furthermore, these are statistics that
-appear to have expected values that vary with tree size (20), making
-comparisons across studies challenging. Therefore, transformation of
-these statistics into \(Z\)-scores in the same simulation experiments
-would allow evaluation of the efficacy of each statistic to yield effect
-sizes that could be used for comparisons of the strength of phylogenetic
-signal across traits and clades.
+By contrast, for Blomberg's \(\kappa\), which is positively ubounded, we
+might expect that for any \(\lambda\) used to generate data, estimates
+of \(\kappa\) might follow a normal distribution, with values
+distributed symmetrically about the input value. This attribute seemed
+less reasonable based on the simulations performed by Münkemüller et
+al.~(18), which suggested that distributions were positively skewed and
+that Blomberg's \(\kappa\) might not behave as a statistic that follows
+a normal distribution. However, because their simulations used a
+weighted combination of simulated phylogenetic signal strengths, strong
+inferences are not possible (and distributional attributes were not the
+intended result of their simulations). Thus, for both Pagel's
+\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
+across a range of \(\lambda\) used to generate data would be valuable
+for adjudicating the relibalilty of these statistics. Furthermore, these
+are statistics that appear to have expected values that vary with tree
+size (18), making comparisons across studies challenging. Therefore,
+transformation of these statistics into \(Z\)-scores would allow
+evaluation of the efficacy of each statistic to yield effect sizes that
+could be used for comparisons of the strength of phylogenetic signal
+across traits and lineages.
 
-Here we used simulation experiments to compare the distributional
+Here we use simulation experiments to compare the distributional
 attributes of \(\hat{\lambda}\) and \(\kappa\), plus their effect sizes
@@
 \textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
-innacurate.} Computer simulations revealed that for \(\hat{\lambda}\),
-the distributional expectations of a Bernoulli variable were mostly
-upheld. First, the mean value of \(\hat{\lambda}\) did increase as
-\(\lambda\) increased, but it was negatively-biased (particularly for
-small tree sizes), and was consistently less than the input \(\lambda\)
-value across most of its range (Fig. 1 black line). Second, the standard
-deviation of \(\hat{\lambda}\) was largest at intermediate values of
-\(\lambda\) and smallest at extreme values, implying that the precision
-in estimating \(\lambda\) varied across the range of input values (Fig.
-1 red line). Additionally, standard deviations of \(\hat{\lambda}\) were
-negatively associated with tree size, and for tress of 128 species or
-less, \(\hat{\lambda}\) was quite variable, except for cases when
-\(\lambda\) was near or equal to \(1\). Third, the distributions of
-\(\hat{\lambda}\) were not normal across its range, but became
-increasingly skewed at more extreme values of \(\lambda\) (Fig. 1 blue
-line). For small tree sizes, it was also clear that distributions were
-more platykurtic at intermeidate values of \(\hat{\lambda}\). Taken
-together these results reveal that \(\hat\lambda\) inconsistently
-estimated phylogenetic signal, both across tree sizes and across the
-range of input values. Additional simulations (Supplemental Information)
-revealed that incorporating \(\lambda\) in PGLS anova and regression did
-not adversely affect the statistical properties of model fitting (type I
-error, power, parameter estimation); thus, it is reasonable to include
-\(\lambda\) in PGLS as a parameter for tuning the degree of phylogenetic
-signal in the dependent variables during the analysis. However, the
-statistical properties revealed in Fig. 1 demonstrate that \(\lambda\)
-is unsuitable as an effect size for measuring the strength of
-phylogenetic signal in data.
+innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
+distributional expectations of a Bernoulli variable were mostly upheld.
+First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
+increases, but it is negatively-biased (particularly for small tree
+sizes), and is consistently less than the input \(\lambda\) value across
+most of its range (Fig. 1 black line). Second, the standard deviation of
+\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
+smallest at extreme values, implying that the precision in estimating
+\(\lambda\) varies across the range of input values (Fig. 1 red line).
+Additionally, standard deviations of \(\hat{\lambda}\) are negatively
+associated with tree size, and for tress of 128 species or less,
+\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
+is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
+are not normal across its range, but become increasingly skewed at more
+extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
+it is also clear that distributions are more platykurtic at intermediate
+values of \(\hat{\lambda}\). Taken together these results reveal that
+\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
+across tree sizes and across the range of input values. Additional
+simulations (Supplemental Information) reveal that incorporating
+\(\hat\lambda\) in PGLS anova and regression does not adversely affect
+the statistical properties of PGLS parameter estimation or model
+evaluation (type I error, power, bias in coefficients). Thus, it is
+reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
+tuning the degree of phylogenetic signal in the dependent variables
+during the analysis. However, the statistical properties shown in Fig. 1
+demonstrate that \(\lambda\) is unsuitable as an effect size for
+measuring the strength of phylogenetic signal in data, or for comparing
+that signal across datasets.
 
@@
 \includegraphics{new.fig.1.temp.png}
-\caption{Patterns in \(\lambda\).{}}
+\caption{Response of Pagel's \(\lambda\) to increasing strength of
+Brownian motion. Gray line signifies the 1:1 line where the input value
+matches the estimate \(\hat\lambda\). At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).{}}
 \end{figure}
@@
 \textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
-stable.} Simulation results for \(\kappa\) demonstrated that this
-measure displayed better statistical properties. First, as expected,
-mean values of \(\kappa\) increased consistently with increasing signal
-(\(\lambda\)) irrespective of tree size (Fig. 2 black line).
-Additionally, the standard deviation of \(\kappa\) was consistent across
-tree sizes (Fig. 2 red line), and while it increased with \(\lambda\),
-it was always less than the corresponding mean. This finding is perhaps
-unsurprising, as \(\kappa\) is bound by 0, and was never large for small
-values of \(\lambda\). Importantly, \(\kappa\) was normally distributed
-across the range of input \(\lambda\), and remained consistent in this
-pattern regardless of tree size (Fig. 2 blue line). In fact,
-distributions of \(\kappa\) displayed slight skewing only at small tree
-sizes and when \(\lambda\) was near or equal to \(1\). This result
-differs from those of (20), where skewing appears to be from combining
-random values generated independently. These findings reveal that
-\(\kappa\) is more reliable as an estimate of phylogenetic signal.
+reliable.} Simulation results for \(\hat\kappa\) demonstrate that
+\(\hat\kappa\) displays better statistical properties. First, as
+expected, mean values of \(\hat\kappa\) increase with increasing signal
+(\(\lambda\)) irrespective of tree size, though the increase does not
+scale linearly with input levels of phylogenetic signal (Fig. 2 black
+line). Additionally, the standard deviation of \(\hat\kappa\) is
+consistent across tree sizes (Fig. 2 red line), and while it increases
+with \(\lambda\), it is always less than the corresponding mean. This
+finding is perhaps unsurprising, as \(\hat\kappa\) is lower-bounded by
+0, and is never large for small values of \(\lambda\). Importantly,
+\(\hat\kappa\) is normally distributed across the range of input
+\(\lambda\), and remains consistent in this pattern regardless of tree
+size (Fig. 2 blue line). In fact, distributions of \(\hat\kappa\)
+display slight skewing only at small tree sizes and when \(\lambda\) is
+near or equal to \(1\). This result differs from those of (18), where
+skewing appears to be from combining random values generated
+independently. These findings reveal that while \(\kappa\) is more
+reliable as an estimate of phylogenetic signal, the non-linear scaling
+with input signal implies that it should not be considered an effect
+size that measures the strength of phylogenetic signal on a common scale
+for comparison across datasets.
 
@@
 \includegraphics{new.fig.2.temp.png}
-\caption{Patterns in \(\kappa\).{}}
+\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
+Brownian motion. Gray line signifies the 1:1 line where the input value
+matches the estimate \(\hat\kappa\). At each input level, the dark black
+line represents the empirically derived expected value (mean) of
+\(\hat\kappa\), the red line is the standard deviation of
+\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
+\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).{}}
 \end{figure}
@@
 characterize phylogenetic signal.} To measure the strength of
-phylogenetic signal on a common scale, we derived effect sizes
-(Z-scores) for both \(\lambda\) and \(\kappa\) (see Methods) and
-evaluated them relative to known phylogenetic signal. Both
-\(Z_{\lambda}\) and \(Z_{\kappa}\) were associated with input
-phylogenetic signal (\(\lambda\)), indicating that both statistics
-captured the observed signal (Fig. 3). However, effect sizes from
-\(\hat{\lambda}\) made little sense, as they were more strongly
-associated with tree size than they were with the actual phylogenetic
-signal in the data (Fig. 3). By contrast, \(Z_{\kappa}\) was much more
-consistent accross tree sizes, and exhibited a much stronger association
-with phylogenetic signal strength as compared to tree size (Fig. 3).
-Thus between the two statistics, \(Z_{\kappa}\) is a more reliable
-measure of the strength of phylogenetic signal, and may be used to
-compare levels of phylogenetic signal across datasets.
+phylogenetic signal on a common scale, here we propose effect sizes
+(Z-scores) for both \(\lambda\) and \(\kappa\). Statistically, a
+standardized effect size may be found as:
+
+\begin{align}
+    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
+\end{align}
+
+where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
+its expected value under the null hypothesis, and \(\sigma_\theta\) is
+its standard error (31--33). Typically, \(\theta_{obs}\) and
+\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
+obtained from the distribution of \(\theta\) derived from parametric
+theory. However, recent advances in resampling theory (34--37) have
+shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
+an empirical sampling distribution of \(\theta\) obtained from
+permutation procedures.
+
+Formalizing the suggestion of Adams and Collyer (38), an effect size for
+\(\kappa\) may be found as:
+
+\begin{align}
+    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
+\end{align}
+
+where \(\kappa_{obs}\) is the observed phylogenetic signal, and
+\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
+deviation of the empirical sampling distribution of \(\kappa\) obtained
+via permutation. The empirical sampling distribution of \(\kappa\) is
+first transformed via Box-Cox to better adhere to the assumption of
+normality.
+
+For \(\lambda\), deriving an effect size is more challenging, as
+\(\lambda\) does not have a sampling distribution from which the
+standard error and confidence intervals may be obtained. Confidence
+intervals are therefore generated for the values of \(\lambda\) that
+intersect the log-likehihood profile for corresponding percentiles of
+the \(\chi^2\) distribution used to compare the putative model to a null
+model with \(\lambda = 0\) {[}\textbf{add ref: MLC thinks Boettiger
+paper?}{]}. Thus, an effect size for \(\lambda\) may be found as:
+
+\begin{align}
+   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
+\end{align}
+
+where, \(\hat{\lambda}\) is the maximized likelihood value of
+\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
+statistic for the value.
+
+Here we evaluate the ability of \(Z_{\lambda}\) and \(Z_{\kappa}\) to
+characterize known levels of phylogenetic signal. Both \(Z_{\lambda}\)
+and \(Z_{\kappa}\) are associated with input phylogenetic signal
+(\(\lambda\)), indicating that both statistics capture the observed
+signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
+little sense, as they are more strongly associated with tree size than
+they are with the actual phylogenetic signal in the data (Fig. 3). By
+contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
+increases more linearly with incrasing levels of phylogenetic signal.
+Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
+phylogenetic signal strength as compared to tree size (Fig. 3), and its
+standard deviation across input signal is more consistent. This implies
+that similar levels of precision are found with \(Z_{\kappa}\) across
+the range of input values. Thus between the two statistics,
+\(Z_{\kappa}\) is a more reliable measure of the strength of
+phylogenetic signal, and may be used to compare levels of phylogenetic
+signal across datasets.
 
@@
 \includegraphics{new.fig.3.temp.png}
-\caption{Patterns in Z.{}}
+\caption{Response of effect sizes \(Z_\lambda\) and \(Z_\kappa\) to
+increasing strength of Brownian motion.{}}
 \end{figure}
@@
 comparisons across datasets.} To statistically compare the strength of
-phylogenetic signal across datasets we derived a two-sample test
-statistic (\(\hat{Z}_{12}\): see Methods). As an example of its utility,
-we compared \(Z_{\kappa}\) for two ecologically-relevant traits in
-plethodontid salamander (Fig. 4): surface area to volume ratios (SA:V)
-and relative body width (\(\frac{BW}{SVL}\)) (33, 34). While both traits
-contained significant phylogenetic signal, tests based on
-\(\hat{Z}_{12}\) revealed that the degree of phylogenetic signal was
-significantly stronger in SA:V (Fig. 4). Biologically, this observation
-may be interpreted by the fact that tropical species -- which form a
-monophyletic group within plethodontids -- display greater variation in
-SA:V, which covaries with disparity in their climatic niches (34). Thus,
-greater phylogenetic signal in SA:V is to be expected.
+phylogenetic signal across datasets we propose a two-sample test
+statistic (\(\hat{Z}_{12}\)). Based on statistical theory, a two-sample
+test statistic may be calculated as:
+
+\begin{align}
+  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
+\end{align}
+
+where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
+\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
+\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
+side of the equation illustrates that if \(Z_\kappa\) has already been
+calculated for two sampling distributions as in equation 2, the sampling
+distributions have unit variance for each of the \(Z_\kappa\)
+statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
+obtained from a standard normal distribution. Typically,
+\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
+(one-tailed) tests may be specified should the empirical situation
+require it (35, 37).
+
+To demonstrate the utility of \(\hat{Z}_12\), we compared \(Z_{\kappa}\)
+for two ecologically-relevant traits in plethodontid salamander (Fig.
+4): surface area to volume ratios (SA:V) and relative body width
+(\(\frac{BW}{SVL}\)) (39, 40). While both traits contained significant
+phylogenetic signal, tests based on \(\hat{Z}_{12}\) revealed that the
+degree of phylogenetic signal was significantly stronger in SA:V
+(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
+observation may be interpreted by the fact that tropical species --
+which form a monophyletic group within plethodontids -- display greater
+variation in SA:V, which covaries with disparity in their climatic
+niches (40). Thus, greater phylogenetic signal in SA:V is to be
+expected.
 
@@
 
-\textbf{To be edited.} address hypotheses that compare the strength of
-phylogenetic signal between various phenotypic traits, even when those
-traits are found in different evolutionary lineages or have different
-units or scales.
-
 It is common in comparative evolutionary studies to characterize the
@@
 taxa. However, while numerous analytical approaches may be used to
-quantify phylogenetic signal (13, 14, 16--18), methods that explicitly
+quantify phylogenetic signal (11, 12, 14--16), methods that explicitly
 measure the strength of phylogenetic signal, or facilitate comparisons
-among datasets, have remained underdeveloped. In this study, we
-evaluated the precision of one common measure, Pagel's \(\lambda\), and
-explored its efficacy for characterizing the strength of phylogenetic
-signal in phenotypic data. Using computer simulations, we found that the
-precision of \(\lambda\) increased with increasing sample sizes; a
-pattern noted previously (25), and one that conformed with parametric
-statistical theory (35). However, we also found that vastly different
-\(\lambda\) estimates could be obtained from data containing the same
-level of phylogenetic signal, and that similar \(\lambda\) estimates may
-be obtained from data containing differing levels of phylogenetic
-signal. Further, the precision of \(\lambda\) varied with the strength
-of phylogenetic signal, where lower precision was observed when in data
-whose phylogenetic signal was of intemediate strength. From these
-findings we conclude that \(\lambda\) is not a reliable indicator of the
-observed strength of phylogenetic signal in phenoytpic datasets, and
-that biological interpretations of the strength of signal based on this
-parameter may innacurately characterize such effects.
+among datasets, have remained underdeveloped. We evaluated the precision
+of one common measure, Pagel's \(\lambda\), and explored its efficacy
+for characterizing the strength of phylogenetic signal in phenotypic
+data. Using computer simulations, we found that \(\lambda\) behaves as a
+Bernoulli random variable, with estimates that are increasingly skewed
+at larger and smaller input levels of phylogenetic signal. Further, the
+precision of \(\lambda\) in estimating actual levels of phylogenetic
+signal varies with both tree size (see also ref. (23)) and input levels
+of phylogentetic signal. From these findings we conclude that
+\(\lambda\) is not a reliable indicator of the observed strength of
+phylogenetic signal in phenoytpic datasets, and should not be used as an
+effect size for comparing the degree of phylogenetic signal between
+datasets.
 
@@
 and found that \(Z_\kappa\) was a better estimate of the strength of
-phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
-precise than \(Z_\lambda\), and precision was more consistent across the
-range of input levels of phylogenetic signal. Additionally, values of
-\(Z_\kappa\) more accurately tracked known changes in the magnitude of
-phylogenetic signal, as demonstrated by the linear relationship between
-\(Z_\kappa\) and \(\lambda_{in}\). Thus, \(Z_\kappa\) holds promise as a
-measure of the relative strength of phylogenetic signal that reflects
-the magnitude of this effect in phenotypic data. We therefore recommend
-that future studies interested in the strength of phylogenetic signal
-incorporate \(Z_\kappa\) as a statistical measure of this effect.
+phylogenetic signal in phenotypic data. First, values of \(Z_\kappa\)
+more accurately tracked known changes in the magnitude of phylogenetic
+signal, as demonstrated by the linear relationship between \(Z_\kappa\)
+and input signal. Additionally, the precision of \(Z_{\kappa}\) was more
+consistent across the range of input levels of phylogenetic signal.
+Thus, \(Z_\kappa\) is a more reliable measure of the relative strength
+of phylogenetic signal, and places that effect on a common and
+comparable scale. We therefore recommend that future studies interested
+in evaluating the strength of phylogenetic signal incorporate
+\(Z_\kappa\) as a statistical measure of this effect.
 
-Based on the effect size \(Z_\kappa\), we then proposed a two-sample
-test, which provides means of determining whether the strength of
+Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
+formal statistical procedure for determining whether the strength of
 phylogenetic signal is greater in one phenotypic trait as compared to
-another, via a hypothesis test. Prior studies have summarized patterns
-of variation in phylogenetic signal across datasets using summary test
-values, such as \(\kappa\) (14). However, \(\kappa\) does not scale
-linearly with input levels of phylogenetic signal, and its variance
-increases (i.e., precision decreases) with increasing strength of
-phylogenetic signal (20, 22). Thus, \(\kappa\) should not be considered
-an effect size that measures the strength of phylogenetic signal on a
-common scale. By contrast, standardizing \(\kappa\) (\(Z_\kappa\), via
-equation 2) alleviates these concerns, and facilitates formal
-statistical comparisons of the strength of signal across datasets. Thus
-when viewed from this perspective, the approach developed here aligns
-well with other statistical approaches such as meta-analysis (36--38),
-where summary statistics across datasets are converted to standardized
-effect sizes for subsequent ``higher order'' statistical summaries or
-comparisons. As such, our approach enables evolutionary biologists to
-quantitatively examine the relative strength of phylogenetic signal
-across a wide range of phenotypic traits, and thus opens the door for
-future discoveries that inform on how phenotypic diversity accumulates
-in macroevolutionary time across the tree of life.
+another. Prior studies have summarized patterns of variation in
+phylogenetic signal across datasets using summary test values, such as
+\(\kappa\) (12). However, because \(\kappa\) does not scale linearly
+with input levels of phylogenetic signal (Fig. 2), and its variance
+increases with increasing strength of phylogenetic signal (18, 20), it
+should not be considered an effect size that measures the strength of
+phylogenetic signal on a common scale. By contrast, standardizing
+\(\kappa\) to \(Z_\kappa\) via equation 2 alleviates these concerns, and
+facilitates formal statistical comparisons of the strength of signal
+across datasets. Thus when viewed from this perspective, the approach
+developed here aligns well with other statistical approaches such as
+meta-analysis (31, 41, 42), where summary statistics across datasets are
+converted to standardized effect sizes for subsequent ``higher order''
+statistical summaries or comparisons. As such, our approach enables
+evolutionary biologists to quantitatively examine the relative strength
+of phylogenetic signal across a wide range of phenotypic traits, and
+thus opens the door for future discoveries that inform on how phenotypic
+diversity accumulates in macroevolutionary time across the tree of life.
 
@@
 resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
-measurement cancel out during the calculation of \(Z\) (39). Thus,
+measurement cancel out during the calculation of \(Z\) (43). Thus,
 \(Z_\kappa\) represents the strength of phylogenetic signal on a common
@@
 investigation. This means that the strength of phylogenetic signal may
-be compared across datsets for continuous phenotypic traits measured in
+be compared across datasets for continuous phenotypic traits measured in
 different units and scale, because those units have been standardized
@@
 \(\frac{\#{displays}}{second}\)). In fact, our empirical example
-provided such a comparison, as SA:V is represented in \(mm^{-1}\) while
-relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
+provided just such a comparison, as SA:V is represented in \(mm^{-1}\)
+while relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
 Additionally, our method is capable of comparing the strength of
@@
 of phylogenetic signal using \(\kappa\) have been generalized for
-multivariate data (18). Furthermore, tests based on \(\hat{Z}_{12}\) may
+multivariate data (16). Furthermore, tests based on \(\hat{Z}_{12}\) may
 be utilized for comparing the strength of phylogenetic signal among
@@
 for \(\lambda\) and \(\kappa\) values are consistent with those found in
-the simulations performed by Münkemüller et al.~(20), but that study
+the simulations performed by Münkemüller et al.~(18), but that study
 investigated type I error rates and statistical power, finding that
@@
 tendency for variable \(\lambda\) and consistent \(\kappa\) at
-intermediate phylogenetic signal strengths was observed (20). Recent
-work by Molina-Venegas and Rodríguez (23) found that \(\kappa\) but not
-\(\lambda\) tended to inflate the estimate of phylogenetic signal,
-leading to moderate type I and type II biases, if polytomic chronograms
-were used. Their work more thoroughly addressed previous observations of
-inflated \(\kappa\) for incompletely resolved phylogenetic trees (20,
-40). An interesting question is whether an inflated \(\kappa\) value
-leads to an inflated \(Z_\kappa\) or does a tendency of a particular
-tree to inflate estimates of \(\kappa\) also inflate the values in
-random permutations of a test, in which case \(Z_\kappa\) is robust to
-polytomies? We repeated the analyses in Figure 4, adjusting trees to
-have 50\% collapsed nodes, per the technique of Molina-Venegas and
-Rodríguez (23), and found results were consistent (see Supporting
-Information). This confirms that any tendency of incompletely resolved
-trees to inflate \(\kappa\) as a descriptive statistic does not inflate
-\(Z_\kappa\) as an effect size. Furthermore, because comparison of
-effect sizes in a test is a comparison of locations of observed values
-in their sampling distributions, which would shift concomitantly because
-of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
-be robust in spite of unresolved trees.
+intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
+(18)). Recent work by Molina-Venegas and Rodríguez (21) found that
+\(\kappa\) but not \(\lambda\) tended to inflate the estimate of
+phylogenetic signal, leading to moderate type I and type II biases, if
+polytomic chronograms were used. Their work more thoroughly addressed
+previous observations of inflated \(\kappa\) for incompletely resolved
+phylogenetic trees (18, 44). An interesting question is whether an
+inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
+tendency of a particular tree to inflate estimates of \(\kappa\) also
+inflate the values in random permutations of a test, in which case
+\(Z_\kappa\) is robust to polytomies? We repeated the analyses in Figs.
+1 \& 2, adjusting trees to have 50\% collapsed nodes, per the technique
+of Molina-Venegas and Rodríguez (21), and found results were consistent
+(Supporting Information). This confirms that any tendency of
+incompletely resolved trees to inflate \(\kappa\) as a descriptive
+statistic does not inflate \(Z_\kappa\) as an effect size. Furthermore,
+because comparison of effect sizes in a test is a comparison of
+locations of observed values in their sampling distributions, which
+would shift concomitantly because of this tendency, the \(Z_{12}\) test
+statistic in equation 4 appears to be robust in spite of unresolved
+trees.
 
@@
 sampling distribution is representative of true probability
-distributions (12). With low coefficients of variation for \(Z_\kappa\)
-(Figure 4), it is difficult to imagine that a hypothesis test can
-improve equation 4 for efficiently comparing phylogenetic signal for
-different traits, different trees, or a combination of both.
+distributions (10). Given the greater consistency in estimates of
+\(Z_\kappa\) across tree sizes and input signal, it is difficult to
+imagine a hypothesis test that can improve equation 4 for efficiently
+comparing phylogenetic signal for different traits, different trees, or
+a combination of both.
 
@@
 \section{Methods}\label{methods}}
-
-\textbf{Derivation of effect sizes.} Statistically, a standardized
-effect size may be found as:
-
-\begin{align}
-    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
-\end{align}
-
-where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
-its expected value under the null hypothesis, and \(\sigma_\theta\) is
-its standard error (35, 37, 41). Typically, \(\theta_{obs}\) and
-\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
-obtained from the distribution of \(\theta\) derived from parametric
-theory. However, recent advances in resampling theory (42--45) have
-shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
-an empirical sampling distribution of \(\theta\) obtained from
-permutation procedures.
-
-Formalizing the suggestion of Adams and Collyer (46), an effect size for
-\(\kappa\) may be found as:
-
-\begin{align}
-    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
-\end{align}
-
-where \(\kappa_{obs}\) is the observed phylogenetic signal, and
-\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
-deviation of the empirical sampling distribution of \(\kappa\) obtained
-via permutation. The empirical sampling distribution of \(\kappa\) is
-first transformed via Box-Cox to better adhere to the assumption of
-normality.
-
-For \(\lambda\), deriving an effect size is more challenging, as
-\(\lambda\) does not have a sampling distribution from which the
-standard error (and thus confidence intervals) may be obtained.
-Confidence intervals are tehrefore generated for the values of
-\(\lambda\) that intersect the log-likehihood profile for corresponding
-percentiles of the \(\chi^2\) distribution used to compare the putative
-model to a null model with \(\lambda = 0\) {[}\textbf{add ref: MLC
-thinks Boettiger paper?}{]}. Thus, an effect size for \(\lambda\) may be
-found as:
-
-\begin{align}
-   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
-\end{align}
-
-where, \(\hat{\lambda}\) is the maximized likelihood value of
-\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
-statistic for the value. Note that alternative formulations could be
-envisioned: \(Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}\), where
-\(d\) is a binary value \((-1,1)\) to indicate a direction based on
-whether \(\hat{\lambda}\) is below or above a critical value of
-\(\lambda\), for a quantile from a \(\chi^{2}\) distribution at a
-probability of 0.5. However, preliminary investigations found that
-mapping \(\chi^2_{\hat{\lambda}}\) values in this manner did not produce
-effect sizes that were symmetrical about \(Z = 0\), as the mapping was
-not linear and the log-likelihood profiles can be rather flat for small
-trees. \textbf{Is alternative needed here?}
-
-\textbf{Derivation of two-sample test statistic.} To compare the
-strength of phylogenetic signal across datasets, a two-sample test
-statistic may be calculated as:
-
-\begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
-\end{align}
-
-where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
-\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
-\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
-side of the equation illustrates that if \(Z_\kappa\) has already been
-calculated for two sampling distributions as in equation 2, the sampling
-distributions have unit variance for each of the \(Z_\kappa\)
-statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
-obtained from a standard normal distribution. Typically,
-\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
-(one-tailed) tests may be specified should the empirical situation
-require it (43, 45).
 
@@
 species means from individuals of 305 species, from which species means
-were obtained (33, 34). A time-dated molecular phylogeny for the group
-(47) was pruned to match the species in the phenotypic dataset. The
+were obtained (39, 40). A time-dated molecular phylogeny for the group
+(45) was pruned to match the species in the phenotypic dataset. The
 phylogenetic signal in each trait was then characterized using
 \(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
-\texttt{geomorph} 3.3.1 (48, 49), and routines by the authors
+\texttt{geomorph} 3.3.1 (46, 47), and routines by the authors
 (\textbf{to be incorporated in \texttt{geomorph} upon manuscript
@@
 
-\leavevmode\hypertarget{ref-ButlerKing2004}{}%
-6. Butler MA, King AA (2004) Phylogenetic comparative analysis: A
-modeling approach for adaptive evolution. \emph{American Naturalist}
-164:683--695.
-
 \leavevmode\hypertarget{ref-MartinsHansen1997}{}%
-7. Martins EP, Hansen TF (1997) Phylogenies and the comparative method:
+6. Martins EP, Hansen TF (1997) Phylogenies and the comparative method:
 A general approach to incorporating phylogenetic information into the
@@
 \leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
-8. O'Meara BC, Ane C, Sanderson MJ, Wainwright PC (2006) Testing for
+7. O'Meara BC, Ane C, Sanderson MJ, Wainwright PC (2006) Testing for
 different rates of continuous trait evolution using likelihood.
@@
 
-\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
-9. Revell LJ, Harmon LJ (2008) Testing quantitative genetic hypotheses
-about the evolutionary rate matrix for continuous characters.
-\emph{Evolutionary Ecology Research} 10:311--331.
-
 \leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
-10. Beaulieu JM, Jhwueng DC, Boettiger C, O'Meara BC (2012) Modeling
+8. Beaulieu JM, Jhwueng DC, Boettiger C, O'Meara BC (2012) Modeling
 stabilizing selection: Expanding the ornstein-uhlenbeck model of
@@
 \leavevmode\hypertarget{ref-Adams2014b}{}%
-11. Adams DC (2014) A method for assessing phylogenetic least squares
+9. Adams DC (2014) A method for assessing phylogenetic least squares
 models for shape and other high-dimensional multivariate data.
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
-12. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
+10. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
 aggregation, biological challenges, and a refined permutation procedure.
@@
 \leavevmode\hypertarget{ref-Pagel1999}{}%
-13. Pagel MD (1999) Inferring the historical patterns of biological
+11. Pagel MD (1999) Inferring the historical patterns of biological
 evolution. \emph{Nature} 401:877--884.
@@
 \leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
-14. Blomberg SP, Garland T, Ives AR (2003) Testing for phylogenetic
+12. Blomberg SP, Garland T, Ives AR (2003) Testing for phylogenetic
 signal in comparative data: Behavioral traits are more labile.
@@
 \leavevmode\hypertarget{ref-Revell_et_al2008}{}%
-15. Revell LJ, Harmon LJ, Collar DC (2008) Phylogenetic signal,
+13. Revell LJ, Harmon LJ, Collar DC (2008) Phylogenetic signal,
 evolutionary process, and rate. \emph{Systematic Biology} 57:591--601.
@@
 \leavevmode\hypertarget{ref-Abouheif1999}{}%
-16. Abouheif E (1999) A method for testing the assumption of
+14. Abouheif E (1999) A method for testing the assumption of
 phylogenetic independence in comparative data. \emph{Evolutionary
@@
 \leavevmode\hypertarget{ref-Gittleman1990}{}%
-17. Gittleman JL, Kot M (1990) Adaptation: Statistics and a null model
+15. Gittleman JL, Kot M (1990) Adaptation: Statistics and a null model
 for estimating phylogenetic effects. \emph{Systematic Zoology}
@@
 \leavevmode\hypertarget{ref-Adams2014a}{}%
-18. Adams DC (2014) A generalized Kappa statistic for estimating
+16. Adams DC (2014) A generalized Kappa statistic for estimating
 phylogenetic signal from shape and other high-dimensional dultivariate
@@
 \leavevmode\hypertarget{ref-Klingenberg2010}{}%
-19. Klingenberg CP, Gidaszewski NA (2010) Testing and quantifying
+17. Klingenberg CP, Gidaszewski NA (2010) Testing and quantifying
 phylogenetic signals and homoplasy in morphometric data.
@@
 \leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
-20. Münkemüller T, et al. (2012) How to measure and test phylogenetic
+18. Münkemüller T, et al. (2012) How to measure and test phylogenetic
 signal. \emph{Methods in Ecology and Evolution} 3:743--756.
@@
 \leavevmode\hypertarget{ref-Pavoine2012}{}%
-21. Pavoine S, Ricotta C (2012) Testing for phylogenetic signal in
+19. Pavoine S, Ricotta C (2012) Testing for phylogenetic signal in
 biological traits: The ubiquity of cross-product statistics.
@@
 \leavevmode\hypertarget{ref-DinizFilho2012}{}%
-22. Diniz-Filho JAF, Santos T, Rangel TF, Bini LM (2012) A comparison of
+20. Diniz-Filho JAF, Santos T, Rangel TF, Bini LM (2012) A comparison of
 metrics for estimating phylogenetic signal under alternative
@@
 \leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
-23. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic
+21. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic
 signal; strong or negligible impacts of polytomies and branch length
@@
 \leavevmode\hypertarget{ref-Revell2010}{}%
-24. Revell LJ (2010) Phylogenetic signal and linear regression on
+22. Revell LJ (2010) Phylogenetic signal and linear regression on
 species data. \emph{Methods in Ecology and Evolution} 1:319--329.
@@
 \leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
-25. Boettiger C, Coop G, Ralph P (2012) Is your phylogeny informative?
+23. Boettiger C, Coop G, Ralph P (2012) Is your phylogeny informative?
 Measuring the power of comparative methods. \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
-26. Freckleton RP, Harvey PH, Pagel M (2002) Phylogenetic analysis and
+24. Freckleton RP, Harvey PH, Pagel M (2002) Phylogenetic analysis and
 comparative data: A test and review of evidence. \emph{American
@@
 \leavevmode\hypertarget{ref-Cooper2010}{}%
-27. Cooper N, Jetz W, Freckleton RP (2010) Phylogenetic comparative
+25. Cooper N, Jetz W, Freckleton RP (2010) Phylogenetic comparative
 approaches for studying niche conservatism. \emph{Journal of
@@
 \leavevmode\hypertarget{ref-Bose2019}{}%
-28. Bose R, Ramesh BR, Pélissier R, Munoz F (2019) Phylogenetic
+26. Bose R, Ramesh BR, Pélissier R, Munoz F (2019) Phylogenetic
 diversity in the western ghats biodiversity hotspot reflects
@@
 \leavevmode\hypertarget{ref-Vandelook2019}{}%
-29. Vandelook F, et al. (2019) Nectar traits differ between pollination
+27. Vandelook F, et al. (2019) Nectar traits differ between pollination
 syndromes in balsaminaceae. \emph{Annals of Botany} 124(2):269--279.
@@
 \leavevmode\hypertarget{ref-DeMeester2019}{}%
-30. De Meester G, Huyghe K, Van Damme R (2019) Brain size, ecology and
+28. De Meester G, Huyghe K, Van Damme R (2019) Brain size, ecology and
 sociality: A reptilian perspective. \emph{Biological Journal of the
@@
 \leavevmode\hypertarget{ref-Pintanel2019}{}%
-31. Pintanel P, Tejedo M, Ron SR, Llorente GA, Merino-Viteri A (2019)
+29. Pintanel P, Tejedo M, Ron SR, Llorente GA, Merino-Viteri A (2019)
 Elevational and microclimatic drivers of thermal tolerance in andean
@@
 \leavevmode\hypertarget{ref-Su2019}{}%
-32. Su G, Villéger S, Brosse S (2019) Morphological diversity of
+30. Su G, Villéger S, Brosse S (2019) Morphological diversity of
 freshwater fishes differs between realms, but morphologically extreme
@@
 
-\leavevmode\hypertarget{ref-Baken2019}{}%
-33. Baken EK, Adams DC (2019) Macroevolution of arboreality in
-salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.
-
-\leavevmode\hypertarget{ref-Baken2020}{}%
-34. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
-desiccation‐related morphology in plethodontid salamanders as inferred
-from a novel surface area to volume ratio estimation approach.
-\emph{Evolution} 74:476--486.
+\leavevmode\hypertarget{ref-Glass1976}{}%
+31. Glass GV (1976) Primary, secondary, and meta-analysis of research.
+\emph{Educational Researcher} 5:3--8.
 
 \leavevmode\hypertarget{ref-Cohen1988}{}%
-35. Cohen J (1988) \emph{Statistical power analysis for the behavioral
+32. Cohen J (1988) \emph{Statistical power analysis for the behavioral
 sciences} (Routledge).
 
-\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
-36. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
-meta-analysis} (Elsevier).
-
-\leavevmode\hypertarget{ref-Glass1976}{}%
-37. Glass GV (1976) Primary, secondary, and meta-analysis of research.
-\emph{Educational Researcher} 5:3--8.
-
-\leavevmode\hypertarget{ref-Arnqvist1995}{}%
-38. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
-findings in ecology and evolution. \emph{Trends in Ecology and
-Evolution} 10:236--240.
-
-\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
-39. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
-San Francisco). 4th Ed.
-
-\leavevmode\hypertarget{ref-Davies2012}{}%
-40. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
-resolved phylogenetic trees inflate estimates of phylogenetic
-conservatism. \emph{Ecology} 93(2):242--247.
-
 \leavevmode\hypertarget{ref-Rosenthal1994}{}%
-41. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
+33. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
 H Hedges (Russell Sage Foundation), pp 231--244.
@@
 \leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
-42. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
+34. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
 phenotypic change for phenotypes described by high-dimensional data.
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
-43. Adams DC, Collyer ML (2016) On the comparison of the strength of
+35. Adams DC, Collyer ML (2016) On the comparison of the strength of
 morphological integration across morphometric datasets. \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-CollyerAdams2018}{}%
-44. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
+36. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
 models to high-dimensional data using residual randomization.
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
-45. Adams DC, Collyer ML (2019) Comparing the strength of modular
+37. Adams DC, Collyer ML (2019) Comparing the strength of modular
 signal, and evaluating alternative modular hypotheses, using covariance
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
-46. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
+38. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
 evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
@@
 
+\leavevmode\hypertarget{ref-Baken2019}{}%
+39. Baken EK, Adams DC (2019) Macroevolution of arboreality in
+salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.
+
+\leavevmode\hypertarget{ref-Baken2020}{}%
+40. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
+desiccation‐related morphology in plethodontid salamanders as inferred
+from a novel surface area to volume ratio estimation approach.
+\emph{Evolution} 74:476--486.
+
+\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
+41. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
+meta-analysis} (Elsevier).
+
+\leavevmode\hypertarget{ref-Arnqvist1995}{}%
+42. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
+findings in ecology and evolution. \emph{Trends in Ecology and
+Evolution} 10:236--240.
+
+\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
+43. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
+San Francisco). 4th Ed.
+
+\leavevmode\hypertarget{ref-Davies2012}{}%
+44. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
+resolved phylogenetic trees inflate estimates of phylogenetic
+conservatism. \emph{Ecology} 93(2):242--247.
+
 \leavevmode\hypertarget{ref-Bonett2017}{}%
-47. Bonett RM, Blair AL (2017) Evidence for complex life cycle
+45. Bonett RM, Blair AL (2017) Evidence for complex life cycle
 constraints on salamander body form diversification. \emph{Proceedings
@@
 \leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
-48. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
+46. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
 collection and analysis of geometric morphometric shape data.
@@
 \leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-49. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
+47. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
 for geometric morphometric analyses. R package version 3.3.1. Available

### 2020-07-17 Manuscript/PNAS-Adams-01.Rmd
vsn 1

@@
 
-For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error and confidence intervals may be obtained.
+For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error and confidence intervals may be obtained, and estimates from the Hessian matrix from PGLS are unreliable [@Boettiger_et_al2012].
 Confidence intervals are therefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R
simulation scripts ready

@@ -1,0 +1,160 @@
+# Simulation analyses
+
+# general phy-signal analyses
+
+library(readxl)
+library(latex2exp)
+source("simulation.source.R")
+
+#### Pure-birth tree analyses ------------------------------------------
+
+### Do this now: set working directory to same location as this file
+### Can do that with drop-drown in R Studio
+
+### Parameters
+
+nsim <- 10 # change to 50 for final runs
+treesizes <- 2^(4:7) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.1) for final runs
+
+### Prepare results
+
+location <- getwd() 
+destination <- paste(getwd(), "/results", sep = "")
+
+### Arguments for pure-birth tree analyses
+
+sim.args <- list(treesize = treesizes[1], tree.type = "pb",
+                nsim = nsim, lambdas = lambdas, poly.per = 0,
+                beta = 1, mod.type = "mean")
+
+setwd(destination)
+
+### RUNS!  change effects.simulations.PP to effects.simulations below to avoid parallel processing
+
+sim.sets <- lapply(1:length(treesizes), function(j){
+  sim.args$treesize = treesizes[j]
+  R <- do.call(effects.simulations.PP, sim.args)
+  write.csv(R$Result, file = R$file.name, row.names = FALSE)
+  cat(j, "tree of", length(treesizes), "trees, completed\n")
+  R
+})
+
+names(sim.sets) <- treesizes
+
+####  save files as Rdata (optional) / .csv files already saved
+
+nm <- paste(sim.args$tree.type, ".sets.Rdata", sep = "")
+save(sim.sets, file = nm)
+setwd(location)
+
+
+### plot lambdas
+
+par(mfrow = c(2, length(treesizes)/2))
+for(i in 1:length(treesizes)) {
+  R <- sim.sets[[i]]
+  Rr <- R$Result
+  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+              xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -1,0 +1,160 @@
+# Simulation analyses
+
+# general phy-signal analyses
+
+library(readxl)
+library(latex2exp)
+source("simulation.source.R")
+
+#### Balanced tree analyses ------------------------------------------
+
+### Do this now: set working directory to same location as this file
+### Can do that with drop-drown in R Studio
+
+### Parameters
+
+nsim <- 10 # change to 50 for final runs
+treesizes <- 2^(4:7) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.1) for final runs
+
+### Prepare results
+
+location <- getwd() 
+destination <- paste(getwd(), "/results", sep = "")
+
+### Arguments for balanced tree analyses
+
+sim.args <- list(treesize = treesizes[1], tree.type = "bal",
+                nsim = nsim, lambdas = lambdas, poly.per = 0,
+                beta = 1, mod.type = "mean")
+
+setwd(destination)
+
+### RUNS!  change effects.simulations.PP to effects.simulations below to avoid parallel processing
+
+sim.sets <- lapply(1:length(treesizes), function(j){
+  sim.args$treesize = treesizes[j]
+  R <- do.call(effects.simulations.PP, sim.args)
+  write.csv(R$Result, file = R$file.name, row.names = FALSE)
+  cat(j, "tree of", length(treesizes), "trees, completed\n")
+  R
+})
+
+names(sim.sets) <- treesizes
+
+####  save files as Rdata (optional) / .csv files already saved
+
+nm <- paste(sim.args$tree.type, ".sets.Rdata", sep = "")
+save(sim.sets, file = nm)
+setwd(location)
+
+
+### plot lambdas
+
+par(mfrow = c(2, length(treesizes)/2))
+for(i in 1:length(treesizes)) {
+  R <- sim.sets[[i]]
+  Rr <- R$Result
+  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+              xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.pectinate.runs.R

@@ -4,44 +4,59 @@
 
 library(readxl)
 library(latex2exp)
-
-#### pure-birth tree analyses
 source("simulation.source.R")
-nsim <- 50
-treesizes <- 2^(5:8)
-lambdas <- seq(0, 1, 0.05)
-location <- getwd() # should be same as this file location
+
+#### Pectinate tree analyses ------------------------------------------
+
+### Do this now: set working directory to same location as this file
+### Can do that with drop-drown in R Studio
+
+### Parameters
+
+nsim <- 10 # change to 50 for final runs
+treesizes <- 2^(4:7) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.1) for final runs
+
+### Prepare results
+
+location <- getwd() 
 destination <- paste(getwd(), "/results", sep = "")
 
-pb.args <- list(treesize = treesizes[1], tree.type = "pb",
+### Arguments for pectinate tree analyses
+
+sim.args <- list(treesize = treesizes[1], tree.type = "pect",
                 nsim = nsim, lambdas = lambdas, poly.per = 0,
                 beta = 1, mod.type = "mean")
 
 setwd(destination)
 
-pb.sets <- lapply(1:length(treesizes), function(j){
-  pb.args$treesize = treesizes[j]
-  R <- do.call(effects.simulations.PP, pb.args)
+### RUNS!  change effects.simulations.PP to effects.simulations below to avoid parallel processing
+
+sim.sets <- lapply(1:length(treesizes), function(j){
+  sim.args$treesize = treesizes[j]
+  R <- do.call(effects.simulations.PP, sim.args)
   write.csv(R$Result, file = R$file.name, row.names = FALSE)
   cat(j, "tree of", length(treesizes), "trees, completed\n")
   R
 })
 
-names(pb.sets) <- treesizes
+names(sim.sets) <- treesizes
 
-# save files
+####  save files as Rdata (optional) / .csv files already saved
 
-save(pb.sets, file = "pb.sets.Rdata")
+nm <- paste(sim.args$tree.type, ".sets.Rdata", sep = "")
[…]

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R

@@ -1,0 +1,160 @@
+# Simulation analyses
+
+# general phy-signal analyses
+
+library(readxl)
+library(latex2exp)
+source("simulation.source.R")
+
+#### Pure-birth with polytomies tree analyses ------------------------------------------
+
+### Do this now: set working directory to same location as this file
+### Can do that with drop-drown in R Studio
+
+### Parameters
+
+nsim <- 10 # change to 50 for final runs
+treesizes <- 2^(4:7) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.1) for final runs
+
+### Prepare results
+
+location <- getwd() 
+destination <- paste(getwd(), "/results", sep = "")
+
+### Arguments for PB tree analyses with polytomies 20% but can change %
+
+sim.args <- list(treesize = treesizes[1], tree.type = "pb",
+                nsim = nsim, lambdas = lambdas, poly.per = 0.2,
+                beta = 1, mod.type = "mean")
+
+setwd(destination)
+
+### RUNS!  change effects.simulations.PP to effects.simulations below to avoid parallel processing
+
+sim.sets <- lapply(1:length(treesizes), function(j){
+  sim.args$treesize = treesizes[j]
+  R <- do.call(effects.simulations.PP, sim.args)
+  write.csv(R$Result, file = R$file.name, row.names = FALSE)
+  cat(j, "tree of", length(treesizes), "trees, completed\n")
+  R
+})
+
+names(sim.sets) <- treesizes
+
+####  save files as Rdata (optional) / .csv files already saved
+
+nm <- paste(sim.args$tree.type, ".sets.Rdata", sep = "")
+save(sim.sets, file = nm)
+setwd(location)
+
+
+### plot lambdas
+
+par(mfrow = c(2, length(treesizes)/2))
+for(i in 1:length(treesizes)) {
+  R <- sim.sets[[i]]
+  Rr <- R$Result
+  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+              xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-07-17 Manuscript/PNAS-Adams-01.tex

@@
 \(\lambda\) does not have a sampling distribution from which the
-standard error and confidence intervals may be obtained. Confidence
+standard error and confidence intervals may be obtained, and estimates
+from the Hessian matrix from PGLS are unreliable (23). Confidence
 intervals are therefore generated for the values of \(\lambda\) that

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R
slight typo fix in simulation scripts

@@ -15,7 +15,7 @@ source("simulation.source.R")
 
 nsim <- 10 # change to 50 for final runs
 treesizes <- 2^(4:7) # change to 5:10 for final runs
-lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.1) for final runs
+lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.05) for final runs
 
 ### Prepare results
 

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R
fixing simulation bugs

@@ -6,7 +6,7 @@ library(readxl)
 library(latex2exp)
 source("simulation.source.R")
 
-#### Pure-birth tree analyses ------------------------------------------
+#### Pure-birth tree analyses  with Regression ---------------------------------------
 
 ### Do this now: set working directory to same location as this file
 ### Can do that with drop-drown in R Studio
@@ -14,7 +14,7 @@ source("simulation.source.R")
 ### Parameters
 
 nsim <- 10 # change to 50 for final runs
-treesizes <- 2^(4:7) # change to 5:10 for final runs
+treesizes <- 2^(3:4) # change to 5:10 for final runs
 lambdas <- seq(0, 1, 0.1) # change to seq(0, 1, 0.05) for final runs
 
 ### Prepare results
@@ -26,7 +26,8 @@ destination <- paste(getwd(), "/results", sep = "")
 
 sim.args <- list(treesize = treesizes[1], tree.type = "pb",
                 nsim = nsim, lambdas = lambdas, poly.per = 0,
-                beta = 1, mod.type = "mean")
+                beta = c(0, .25, .5, .75, 1), 
+                mod.type = "reg")
 
 setwd(destination)
 
@@ -49,7 +50,7 @@ save(sim.sets, file = nm)
 setwd(location)
 
 
-### plot lambdas
+### plot lambdas for beta = 0.5
 
 par(mfrow = c(2, length(treesizes)/2))
 for(i in 1:length(treesizes)) {
@@ -78,7 +79,7 @@ for(i in 1:length(treesizes)) {
 }
 
 
-### plot kappas
+### plot kappas for beta = 0.5
 
 par(mfrow = c(2, length(treesizes)/2))
 for(i in 1:length(treesizes)) {
@@ -115,7 +116,7 @@ for(i in 1:2) {
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
-  lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  lambda.z.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
[…]

### 2020-07-17 Data_Analyses/updated_simulations/simulation.source.R

@@ -25,7 +25,7 @@ effects.simulations <- function(treesize , tree.type = c("bal", "pect", "pb"),
   
   n <- treesize
 
-  x <- if(mod.type == "mean" || type == "reg") matrix(1, n) else 
+  x <- if(mod.type == "mean" || mod.type == "reg") matrix(1, n) else 
     sample(matrix(rep(c(0, 1), each  = n/2)))
   
   stat.args <- list(tree = 0, x = x, beta = beta[1], lambda = lambdas[1],
@@ -34,12 +34,12 @@ effects.simulations <- function(treesize , tree.type = c("bal", "pect", "pb"),
  
   Result <- lapply(1:nsim, function(j) {
     
-    result <- sapply(1:length(lambdas), function(jj){
+    result <- lapply(1:length(lambdas), function(jj){
       lambda <- lambdas[jj]
       stat.args$lambda <- tree.args$lambda <- lambda
       stat.args$tree  <- do.call(make.tree, tree.args)
     
-      stats <- lapply(1:length(beta), function(jjj){
+      stats <- sapply(1:length(beta), function(jjj){
         TE <- TRUE
         stat.args$beta <- beta[jjj]
         while(TE) {
@@ -47,15 +47,18 @@ effects.simulations <- function(treesize , tree.type = c("bal", "pect", "pb"),
           TE <- inherits(res, "try-error")
         }
         res <- c(lambda.in = stat.args$lambda, beta.in = stat.args$beta, res)
+        res
       })
-      stats
+      t(stats)
     })
     
+    result <- do.call(rbind, result)
     setTxtProgressBar(pb, j)
-
-    t(simplify2array(result))
+    result
     
   })
+  
+  Result <- do.call(rbind, Result)
   
   close(pb)
   endtime <- Sys.time()
@@ -64,7 +67,7 @@ effects.simulations <- function(treesize , tree.type = c("bal", "pect", "pb"),
   cat(diff, attr(diff, "units"))
   cat("\n")
 
- list(Result = as.data.frame(do.call(rbind, Result)), file.name = file.name)
+  list(Result = as.data.frame(Result), file.name = file.name)
   
 }
 
@@ -88,7 +91,7 @@ effects.simulations.PP <- function(treesize , tree.type = c("bal", "pect", "pb")
   
   n <- treesize
   
[…]

### 2020-07-17 Manuscript/PNAS-Adams-01.Rmd
fig mucking

@@
   - name: Dean C. Adams
-    affiliation: a,1
+    affiliation: a,2
   - name: Erica K. Baken
@@
 corresponding_author:
-  - code: 1
+  - code: 2
     text: "To whom correspondence should be addressed. E-mail: dcadams@iastate.edu"
@@
 
-  ^1^To whom correspondence should be addressed. E-mail: dcadams@iastate.edu
+  ^2^To whom correspondence should be addressed. E-mail: dcadams\@iastate.edu
 
@@
 
-![Response of effect sizes $Z_\lambda$ and $Z_\kappa$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
+![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
 

### 2020-07-17 Manuscript/PNAS-LaTex-Template/PNAS-template-Adams.tex
fig mucking

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,2]{Dean C. Adams}
\author[a,b]{Erica K. Baken}
\author[b]{Michael L. Collyer}

\affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
University, Ames, Iowa, 50010. USA.}
\affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
15232. USA.}

% Please give the surname of the lead author for the running footer
\leadauthor{Adams et al.}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking. Here we develop a standardized effect size, \(Z_\kappa\), which measures the strength of phylogenetic signal on a common statistical scale. We also provide a test statistic, \(\hat{Z}_{12}\), for comparing the strength of phylogenetic signal across datasets. We find that two commonly used parameters (Pagel's \(\lambda\) and Blomberg's \(\kappa\)), not converted to effect sizes, are unsuitable for this purpose. Our effect-size procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have different units or scales.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
research and wrote the paper.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: dcadams\@iastate.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{phylogenetic signal |  macroevolution |  lambda |  kappa }

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely nderdeveloped.
Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values. We find that \(\lambda\) behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal. Further, the precision of \(\lambda\) varies with input signal. Another measure, Blomberg's \(\kappa\), is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal. However, that relationship is decidedly nonlinear. Thus, neither \(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that signal on a common scale for statistical comparison. We develop tests based on \(Z_\kappa\) to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references should be included in the main manuscript file.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:frog}
\end{figure}

\begin{SCfigure*}[\sidecaptionrelwidth][t]
\centering
\includegraphics[width=11.4cm,height=11.4cm]{frog}
\caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
\end{SCfigure*}

\subsection*{Digital Figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\subsection*{Tables}
In addition to including your tables within this manuscript file, PNAS requires that each table be uploaded to the submission separately as a “Table” file.  Please ensure that each table .tex file contains a preamble, the \verb|\begin{document}| command, and the \verb|\end{document}| command. This is necessary so that the submission system can convert each file to PDF.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.

Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.

\begin{figure*}[bt!]
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{figure*}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information (SI)}

Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.  PNAS will publish SI uncomposed, as the authors have provided it.  Additional details can be found here: \href{http://www.pnas.org/page/authors/journal-policies}{policy on SI}.  For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.  The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.  Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.

Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

\subsubsection*{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the manuscript. This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R
silly shapiro.test limitation fixed

@@ -55,7 +55,7 @@ par(mfrow = c(2, length(treesizes)/2))
 for(i in 1:length(treesizes)) {
   R <- sim.sets[[i]]
   Rr <- R$Result
-  lambda.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
               ylab = latex2exp::TeX("Estimated Phylogenetic Signal ($\\hat{lambda}$), $s$, or $W$"),
               ylim = c(0, 1.1))
@@ -69,7 +69,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
+  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) if(length(unique(x)) == 1) 1 else
+    shapiro.test(x)$statistic)
   
   points(lambdas, Ws,
          type = "l", lwd = 3, col = 4)
@@ -84,7 +85,7 @@ par(mfrow = c(2, length(treesizes)/2))
 for(i in 1:length(treesizes)) {
   R <- sim.sets[[i]]
   Rr <- R$Result
-  kappa.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  kappa.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal($\\lambda_{in}$)"),
               ylab = latex2exp::TeX("Estimated Phylogenetic Signal ($\\kappa$),, $s$, or $W$"),
               ylim = c(0, 2))
@@ -97,7 +98,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$phy.signal, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
[…]

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R

@@ -70,7 +70,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
+  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) if(length(unique(x)) == 1) 1 else
+    shapiro.test(x)$statistic)
   
   points(lambdas, Ws,
          type = "l", lwd = 3, col = 4)
@@ -98,7 +99,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$phy.signal, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
+  Ws <- by(Rr$phy.signal, Rr$lambda.in, function(x) if(length(unique(x)) == 1) 1 else
+    shapiro.test(x)$statistic)
   
   points(lambdas, Ws,
          type = "l", lwd = 3, col = 4)

### 2020-07-17 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -69,7 +69,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
+  Ws <- by(Rr$lambda.opt.lambda, Rr$lambda.in, function(x) if(length(unique(x)) == 1) 1 else
+    shapiro.test(x)$statistic)
   
   points(lambdas, Ws,
          type = "l", lwd = 3, col = 4)
@@ -97,7 +98,8 @@ for(i in 1:length(treesizes)) {
   points(lambdas, sds,
          type = "l", lwd = 3, col = 2)
   
-  Ws <- by(Rr$phy.signal, Rr$lambda.in, function(x) shapiro.test(x)$statistic)
+  Ws <- by(Rr$phy.signal, Rr$lambda.in, function(x) if(length(unique(x)) == 1) 1 else
+    shapiro.test(x)$statistic)
   
   points(lambdas, Ws,
          type = "l", lwd = 3, col = 4)

### 2020-07-17 Manuscript/PNAS-Adams-01.tex

@@
 
-\author[a,1]{Dean C. Adams}
+\author[a,2]{Dean C. Adams}
 \author[a,b]{Erica K. Baken}
@@
 
-\textsuperscript{1}To whom correspondence should be addressed. E-mail:
-\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}}
+\textsuperscript{2}To whom correspondence should be addressed. E-mail:
+dcadams@iastate.edu}
 
@@
 \includegraphics{new.fig.3.temp.png}
-\caption{Response of effect sizes \(Z_\lambda\) and \(Z_\kappa\) to
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to
 increasing strength of Brownian motion.{}}

### 2020-08-07 Data_Analyses/simulation.analyses.Regression.runs.R
Almost all simulations (Data_Analysis/updated_simulations) done and results/ folder updated. Only missing results from 1024 tree for ANOVA and reg runs.

@@ -177,3 +177,5 @@ for(i in 1:length(treesizes)) {
 
 
 par(mfcol = c(1,1))
+
+

### 2020-08-07 Data_Analyses/updated_simulations/server_tests/packages.R

@@ -1,0 +1,56 @@
+install.packages("caper", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("latex2exp", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("readxl", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("ape", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+
+install.packages("magick", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("igraph", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("animation", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("phangorn", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("mnormt", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+install.packages("phytools", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+
+#install.packages("processx", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("BH", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("callr", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("yaml", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("base64enc", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("evaluate", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("highr", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
+#install.packages("markdown", lib="/home/ebaken/Rlibs", repos="https://mirror.las.iastate.edu/CRAN")
[…]

### 2020-08-07 Data_Analyses/updated_simulations/server_tests/test.R

@@ -1,0 +1,12 @@
+library(geomorph)
+
+a<-3
+b<-66
+c <- a*b
+d<-seq(a,b)
+
+plot(d, (d*2 + rnorm(length(d))), pch = 1)
+
+source("/work/LAS/dcadams-lab/baken/updated_simulations/test_func.R")
+pth <- test_func(6,9)
+write.csv(pth, "test_pth.csv")

### 2020-08-07 Data_Analyses/updated_simulations/server_tests/test_func.R

@@ -1,0 +1,5 @@
+test_func <- function(x,y) {
+  a <- rnorm(x)
+  b <- rnorm(y)
+  length(a) - max(b)
+}

### 2020-08-07 Manuscript/PNAS-Adams-01-EKB.Rmd
EKB's edits to PNAS manuscript added to 01-ekb.rmd file

--- 2020-07-17 Manuscript/PNAS-Adams-01.Rmd
@@
   Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy of of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
+Here we evaluate the efficacy --of-- of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
 We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
@@
 Nonetheless, an important question that has yet to be considered is whether these statistics are or can be converted to effect sizes for comparative analyses across datasets?
-To be statistics representing phylogenetic signal, they should have reliable distributional properties, which could be revealed with simulation experiments.
+<!-- I think this sentence should be broken into two thoughts: "Nonetheless, an important question that has yet to be considered is whether this comparison is analytically appropriate.
+To be so, these statistics should be converted to effect sizes to avoid issues of incommensorate units or variation due to different phylogenetic branching patterns." or something like that, identifying why the conversion to effect size is necessary.
+Otherwise, I think it makes the issue less relatable.
+I bet few people have considered that question of whether lambda and kappa can be converted to effect sizes, but a much more relatable question is whether people *should* compare the raw values against each other.
+--> To be statistics representing phylogenetic signal, they should have reliable distributional properties, which could be revealed with simulation experiments.
 For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ follows a Bernoulli distribution (**add ref**); i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
@@
 [@Munkemuller_et_al2012]).
-However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not as prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
+However, because the "strength of Brownian motion" <!-- unclear what this quote is referring to.
+Is it in the Munkemuller paper?
+--> was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not as prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
 
-By contrast, for Blomberg's $\kappa$, which is positively ubounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
+By contrast, for Blomberg's $\kappa$, which is positively uNbounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
@@
 However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations).
-Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.
-Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
+Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the --relibalilty-- RELIABILITY of these statistics.
+Furthermore, --these are statistics that appear to have expected values that vary with tree size-- the expected values of these statistics appear to vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
 Therefore, transformation of these statistics into $Z$-scores would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and lineages.
@@
 1 black line).
-Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values, implying that the precision in estimating $\lambda$ varies across the range of input values (Fig.
+Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values <!-- I think the reference to (Fig.1 red line) should go here.
+As it is written, you'd expect the red line to represent an estimate of precision rather than SD -->, implying that the precision in estimating $\lambda$ varies across the range of input values (Fig.
 1 red line).
@@
 1 blue line).
-For small tree sizes, it is also clear that distributions are more platykurtic at intermediate values of $\hat{\lambda}$.
+<!-- Maybe add a quick explanation that lower values of blue line indicate higher skew.
+Without the explanation, a reader might glance at the figure and think "the blue line shows *lower* values at 0 and 1, but the text said it's *more* skewed at 0 and 1".
+I know it's in the caption, but that was my first thought when reading.
+--> For small tree sizes, it is also clear that distributions are more platykurtic at intermediate values of $\hat{\lambda}$.
 Taken together these results reveal that $\hat\lambda$ inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
@@
 3).
-By contrast, $Z_{\kappa}$ is much more consistent accross tree sizes, and increases more linearly with incrasing levels of phylogenetic signal.
+By contrast, $Z_{\kappa}$ is much more consistent accross tree sizes, and increases more linearly with incrEasing levels of phylogenetic signal.
 Additionally, $Z_{\kappa}$ exhibits a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
@@
 For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
-Simulations were then repeated for both bananced and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
+Simulations were then repeated for both --bananced-- BALANCED and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
 Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information).
 
-**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from as species means from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
+**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained --from as species means-- from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
 A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.

### 2020-08-12 Manuscript/PNAS-Adams-01-EKB.tex
vsn 2

--- 2020-07-17 Manuscript/PNAS-Adams-01.tex
@@
 strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly
+Here we evaluate the efficacy --of-- of Pagel's \(\lambda\) to correctly
 estimate the strength of phylogenetic signal in phenotypic traits across
@@
 
-By contrast, for Blomberg's \(\kappa\), which is positively ubounded, we
-might expect that for any \(\lambda\) used to generate data, estimates
-of \(\kappa\) might follow a normal distribution, with values
+By contrast, for Blomberg's \(\kappa\), which is positively uNbounded,
+we might expect that for any \(\lambda\) used to generate data,
+estimates of \(\kappa\) might follow a normal distribution, with values
 distributed symmetrically about the input value. This attribute seemed
@@
 across a range of \(\lambda\) used to generate data would be valuable
-for adjudicating the relibalilty of these statistics. Furthermore, these
-are statistics that appear to have expected values that vary with tree
-size (18), making comparisons across studies challenging. Therefore,
-transformation of these statistics into \(Z\)-scores would allow
-evaluation of the efficacy of each statistic to yield effect sizes that
-could be used for comparisons of the strength of phylogenetic signal
-across traits and lineages.
+for adjudicating the --relibalilty-- RELIABILITY of these statistics.
+Furthermore, --these are statistics that appear to have expected values
+that vary with tree size-- the expected values of these statistics
+appear to vary with tree size (18), making comparisons across studies
+challenging. Therefore, transformation of these statistics into
+\(Z\)-scores would allow evaluation of the efficacy of each statistic to
+yield effect sizes that could be used for comparisons of the strength of
+phylogenetic signal across traits and lineages.
 
@@
 \(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
-smallest at extreme values, implying that the precision in estimating
+smallest at extreme values , implying that the precision in estimating
 \(\lambda\) varies across the range of input values (Fig. 1 red line).
@@
 \centering
-\includegraphics{new.fig.1.temp.png}
+\includegraphics{fig.1.png}
 \caption{Response of Pagel's \(\lambda\) to increasing strength of
@@
 \centering
-\includegraphics{new.fig.2.temp.png}
+\includegraphics{fig.2.png}
 \caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
@@
 contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
-increases more linearly with incrasing levels of phylogenetic signal.
+increases more linearly with incrEasing levels of phylogenetic signal.
 Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
@@
 \centering
-\includegraphics{new.fig.3.temp.png}
+\includegraphics{fig.3.png}
 \caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to
@@
 \centering
-\includegraphics{new.fig.4.png}
+\includegraphics{fig.4.png}
 \caption{(A) Linear measures for relative body size, and regions of the
@@
 departures from \(1.0\) indicate skewness. Simulations were then
-repeated for both bananced and pectinate trees, which yielded
-qualitatively similar results (see Supporting Information). Trees
-containing polytomies, and an evaluation of \(\hat{\lambda}\) from
+repeated for both --bananced-- BALANCED and pectinate trees, which
+yielded qualitatively similar results (see Supporting Information).
+Trees containing polytomies, and an evaluation of \(\hat{\lambda}\) from
 models of linear regression and phylogenetic ANOVA, were also
@@
 \textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
-relative body width (\(\frac{BW}{SVL}\)) measures were obtained from as
-species means from individuals of 305 species, from which species means
-were obtained (39, 40). A time-dated molecular phylogeny for the group
-(45) was pruned to match the species in the phenotypic dataset. The
-phylogenetic signal in each trait was then characterized using
+relative body width (\(\frac{BW}{SVL}\)) measures were obtained --from
+as species means-- from individuals of 305 species, from which species
+means were obtained (39, 40). A time-dated molecular phylogeny for the
+group (45) was pruned to match the species in the phenotypic dataset.
+The phylogenetic signal in each trait was then characterized using
 \(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using

### 2020-08-12 Manuscript/PNAS-Adams-02.Rmd
vsn 2

--- 2020-08-07 Manuscript/PNAS-Adams-01-EKB.Rmd
@@
   Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy --of-- of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
+Here we evaluate the efficacy of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
 We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
@@
 Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$.
-At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](new.fig.1.temp.png)
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](fig.1.png)
 
@@
 Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$.
-At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](new.fig.2.temp.png)
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](fig.2.png)
 
@@
 
-![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](new.fig.3.temp.png)
+![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](fig.3.png)
 
@@
 (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](new.fig.4.png)
+(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](fig.4.png)
 

### 2020-08-12 Manuscript/PNAS-Adams-02.tex
vsn 2

--- 2020-08-12 Manuscript/PNAS-Adams-01-EKB.tex
@@
 strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy --of-- of Pagel's \(\lambda\) to correctly
+Here we evaluate the efficacy of Pagel's \(\lambda\) to correctly
 estimate the strength of phylogenetic signal in phenotypic traits across

### 2020-08-12 Manuscript/VersionDiffNew.r

@@ -7,7 +7,7 @@ rmarkdown::render(input =  "2020-Evol-MultiRR_Submitted.Rmd")
 rmarkdown::render(input = "2020-Evol-MultiRR_Rvsn.Rmd")
 
 #2: latexdiff of tex files
-shell(cmd="latexdiff.exe 2020-Evol-MultiRR_Submitted.tex 2020-Evol-MultiRR_Rvsn.tex > diff.tex")
+shell(cmd="latexdiff.exe PNAS-Adams-01.tex PNAS-Adams-01-EKB.tex > diff.tex")
 
 #3: PDF of diff.tex
 tinytex::latexmk("diff.tex") 

### 2020-08-12 Manuscript/PNAS-Adams-02.Rmd
vsn 2

--- 2020-07-17 Manuscript/PNAS-Adams-01.Rmd
@@
   Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy of of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
+Here we evaluate the efficacy of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
 We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
@@
 The appeal of Pagel's $\lambda$ and Blomberg's $\kappa$ as descriptive statistics is that they provide a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$ or $\kappa$, respectively, in a comparative sense [@DeMeester2019; @Pintanel2019;@Su2019].
-Nonetheless, an important question that has yet to be considered is whether these statistics are or can be converted to effect sizes for comparative analyses across datasets?
+Nonetheless, an important question that has yet to be considered is whether such comparisons are analytically appropriate, and whether these statistics are, or can be, converted to effect sizes for comparative analyses across datasets.
 To be statistics representing phylogenetic signal, they should have reliable distributional properties, which could be revealed with simulation experiments.
@@
 
-By contrast, for Blomberg's $\kappa$, which is positively ubounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
+By contrast, for Blomberg's $\kappa$, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
@@
 However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations).
-Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the relibalilty of these statistics.
-Furthermore, these are statistics that appear to have expected values that vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
+Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the reliability of these statistics as effect sizes.
+Furthermore, the expected values of these statistics appear to vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
 Therefore, transformation of these statistics into $Z$-scores would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and lineages.
@@
 1 black line).
-Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values, implying that the precision in estimating $\lambda$ varies across the range of input values (Fig.
-1 red line).
+Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values (Fig.1 red line), implying that the precision in estimating $\lambda$ varies across the range of input values.
 Additionally, standard deviations of $\hat{\lambda}$ are negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ are quite variable, except for cases when $\lambda$ is near or equal to $1$.
@@
 However, the statistical properties shown in Fig.
-1 demonstrate that $\lambda$ is unsuitable as an effect size for measuring the strength of phylogenetic signal in data, or for comparing that signal across datasets.
+1 demonstrate that $\lambda$ is unsuitable as an effect size for measuring the strength of phylogenetic signal in data, and thus $\lambda$ should not be used for comparing phylogenetic signal across datasets.
 
@@
 2 blue line).
-In fact, distributions of $\hat\kappa$ display slight skewing only at small tree sizes and when $\lambda$ is near or equal to $1$.
-This result differs from those of [@Munkemuller_et_al2012], where skewing appears to be from combining random values generated independently.
-These findings reveal that while $\kappa$ is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
+This result differs from those of [@Munkemuller_et_al2012], where the skewing appears to be the result of combining random values generated independently, rather than being a property of $\kappa$ itself.
+Overall, these findings reveal that while $\kappa$ is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
 
@@
 Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$.
-At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](new.fig.2.temp.png)
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](fig.2.png)
 
 **Effect sizes from $\kappa$ ($Z_{\kappa}$) better characterize phylogenetic signal.**
-To measure the strength of phylogenetic signal on a common scale, here we propose effect sizes (Z-scores) for both $\lambda$ and $\kappa$. Statistically, a standardized effect size may be found as:
+To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and $\kappa$. Statistically, a standardized effect size may be found as:
 
@@
 3).
-By contrast, $Z_{\kappa}$ is much more consistent accross tree sizes, and increases more linearly with incrasing levels of phylogenetic signal.
+By contrast, $Z_{\kappa}$ is much more consistent across tree sizes, and increases more linearly with increasing levels of phylogenetic signal.
 Additionally, $Z_{\kappa}$ exhibits a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
@@
 For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
-Simulations were then repeated for both bananced and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
+Simulations were then repeated for both balanced and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
 Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information).
 
-**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from as species means from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
+**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
 A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.

### 2020-08-12 Manuscript/PNAS-Adams-02.tex
vsn 2

--- 2020-07-17 Manuscript/PNAS-Adams-01.tex
@@
 strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly
+Here we evaluate the efficacy of Pagel's \(\lambda\) to correctly
 estimate the strength of phylogenetic signal in phenotypic traits across
@@
 comparative sense (28--30). Nonetheless, an important question that has
-yet to be considered is whether these statistics are or can be converted
-to effect sizes for comparative analyses across datasets? To be
-statistics representing phylogenetic signal, they should have reliable
+yet to be considered is whether such comparisons are analytically
+appropriate, and whether these statistics are, or can be, converted to
+effect sizes for comparative analyses across datasets. To be statistics
+representing phylogenetic signal, they should have reliable
 distributional properties, which could be revealed with simulation
@@
 
-By contrast, for Blomberg's \(\kappa\), which is positively ubounded, we
-might expect that for any \(\lambda\) used to generate data, estimates
-of \(\kappa\) might follow a normal distribution, with values
+By contrast, for Blomberg's \(\kappa\), which is positively unbounded,
+we might expect that for any \(\lambda\) used to generate data,
+estimates of \(\kappa\) might follow a normal distribution, with values
 distributed symmetrically about the input value. This attribute seemed
@@
 across a range of \(\lambda\) used to generate data would be valuable
-for adjudicating the relibalilty of these statistics. Furthermore, these
-are statistics that appear to have expected values that vary with tree
-size (18), making comparisons across studies challenging. Therefore,
-transformation of these statistics into \(Z\)-scores would allow
-evaluation of the efficacy of each statistic to yield effect sizes that
-could be used for comparisons of the strength of phylogenetic signal
-across traits and lineages.
+for adjudicating the reliability of these statistics as effect sizes.
+Furthermore, the expected values of these statistics appear to vary with
+tree size (18), making comparisons across studies challenging.
+Therefore, transformation of these statistics into \(Z\)-scores would
+allow evaluation of the efficacy of each statistic to yield effect sizes
+that could be used for comparisons of the strength of phylogenetic
+signal across traits and lineages.
 
@@
 \(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
-smallest at extreme values, implying that the precision in estimating
-\(\lambda\) varies across the range of input values (Fig. 1 red line).
+smallest at extreme values (Fig.1 red line), implying that the precision
+in estimating \(\lambda\) varies across the range of input values.
 Additionally, standard deviations of \(\hat{\lambda}\) are negatively
@@
 demonstrate that \(\lambda\) is unsuitable as an effect size for
-measuring the strength of phylogenetic signal in data, or for comparing
-that signal across datasets.
+measuring the strength of phylogenetic signal in data, and thus
+\(\lambda\) should not be used for comparing phylogenetic signal across
+datasets.
 
@@
 \(\lambda\), and remains consistent in this pattern regardless of tree
-size (Fig. 2 blue line). In fact, distributions of \(\hat\kappa\)
-display slight skewing only at small tree sizes and when \(\lambda\) is
-near or equal to \(1\). This result differs from those of (18), where
-skewing appears to be from combining random values generated
-independently. These findings reveal that while \(\kappa\) is more
+size (Fig. 2 blue line). This result differs from those of (18), where
+the skewing appears to be the result of combining random values
+generated independently, rather than being a property of \(\kappa\)
+itself. Overall, these findings reveal that while \(\kappa\) is more
 reliable as an estimate of phylogenetic signal, the non-linear scaling
@@
 characterize phylogenetic signal.} To measure the strength of
-phylogenetic signal on a common scale, here we propose effect sizes
+phylogenetic signal on a common scale, we propose effect sizes
 (Z-scores) for both \(\lambda\) and \(\kappa\). Statistically, a
@@
 they are with the actual phylogenetic signal in the data (Fig. 3). By
-contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
-increases more linearly with incrasing levels of phylogenetic signal.
+contrast, \(Z_{\kappa}\) is much more consistent across tree sizes, and
+increases more linearly with increasing levels of phylogenetic signal.
 Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
@@
 departures from \(1.0\) indicate skewness. Simulations were then
-repeated for both bananced and pectinate trees, which yielded
+repeated for both balanced and pectinate trees, which yielded
 qualitatively similar results (see Supporting Information). Trees
@@
 \textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
-relative body width (\(\frac{BW}{SVL}\)) measures were obtained from as
-species means from individuals of 305 species, from which species means
-were obtained (39, 40). A time-dated molecular phylogeny for the group
-(45) was pruned to match the species in the phenotypic dataset. The
-phylogenetic signal in each trait was then characterized using
-\(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
-\texttt{geomorph} 3.3.1 (46, 47), and routines by the authors
-(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
-acceptance}).
+relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
+individuals of 305 species, from which species means were obtained (39,
+40). A time-dated molecular phylogeny for the group (45) was pruned to
+match the species in the phenotypic dataset. The phylogenetic signal in
+each trait was then characterized using \(\kappa\), which was converted
+to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (46,
+47), and routines by the authors (\textbf{to be incorporated in
+\texttt{geomorph} upon manuscript acceptance}).
 

### 2020-08-14 Manuscript/PNAS-Adams-02.Rmd
vsn 2

@@
 output: rticles::pnas_article
+
 ---
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=8.25cm]{fig.1.png}
+\caption{Response of Pagel's \(\lambda\) to increasing strength of
+Brownian motion. Gray line signifies the 1:1 line where the input value
+matches the estimate \(\hat\lambda\). At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).{}}
+\end{figure}
+
 **Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
@@
 
-![Response of Pagel's $\lambda$ to increasing strength of Brownian motion.
-Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$.
-At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](fig.1.png)
+\begin{figure}
+\centering
+\includegraphics[width=8.25cm]{fig.2.png}
+\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate \(\hat\kappa\).
+At each input level, the dark black line represents the empirically derived expected value (mean) of
+\(\hat\kappa\), the red line is the standard deviation of
+\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
+\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).{}}
+\end{figure}
 
@@
 Overall, these findings reveal that while $\kappa$ is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
-
-![Response of Blomberg's $\hat\kappa$ to increasing strength of Brownian motion.
-Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$.
-At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](fig.2.png)
 
@@
 
-![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](fig.3.png)
+\begin{figure}
+\centering
+\includegraphics[width=8.25cm]{fig.3.png}
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.{}}
+\end{figure}
 

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.ANOVA.runs.R
Editing scripts in the Data_Analyses/updated_simulations/ folder to match the new plotting outputs (decided week of Aug 15, 2020).
Last simulations finished running, munged, and included in Data_Analyses/updated_simulations/results folder

Corresponding SupportingInfo_Figs generated. Not yet converted into high quality png, tiff, and pdf format. EKB will do so once Dean and Mike look at and approve current state.

@@ -51,15 +51,33 @@ setwd(location)
 
 
 ### plot lambdas for beta = 0.5
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pb.32.50.sim.0.poly.anova.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pb.64.50.sim.0.poly.anova.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pb.128.50.sim.0.poly.anova.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pb.256.50.sim.0.poly.anova.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pb.512.50.sim.0.poly.anova.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.sim.0.poly.anova.type.csv")
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
 
-par(mfrow = c(2, length(treesizes)/2))
+png(filename = "Manuscript/SupportingInfo_Figs/fig.1.ANOVA.png",width = 800, height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3, 2), 
+    mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
-  R <- sim.sets[[i]]
+
+  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
   lambda.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Signal ($\\hat{lambda}$), $s$, or $W$"),
[…]

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R

@@ -25,8 +25,8 @@ destination <- paste(getwd(), "/results", sep = "")
 ### Arguments for pure-birth tree analyses
 
 sim.args <- list(treesize = treesizes[1], tree.type = "pb",
-                nsim = nsim, lambdas = lambdas, poly.per = 0,
-                beta = 1, mod.type = "mean")
+                 nsim = nsim, lambdas = lambdas, poly.per = 0,
+                 beta = 1, mod.type = "mean")
 
 setwd(destination)
 
@@ -34,8 +34,8 @@ setwd(destination)
 
 sim.sets <- lapply(1:length(treesizes), function(j){
   sim.args$treesize = treesizes[j]
-  R <- do.call(effects.simulations.PP, sim.args)
-  write.csv(R$Result, file = R$file.name, row.names = FALSE)
+  R <- do.call(effects.simulations, sim.args)
+  #write.csv(R$Result, file = R$file.name, row.names = FALSE)
   cat(j, "tree of", length(treesizes), "trees, completed\n")
   R
 })
@@ -50,14 +50,32 @@ setwd(location)
 
 
 ### plot lambdas
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pb.32.50.sim.0.poly.mean.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pb.64.50.sim.0.poly.mean.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pb.128.50.sim.0.poly.mean.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pb.256.50.sim.0.poly.mean.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pb.512.50.sim.0.poly.mean.type.csv")
[…]

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R

@@ -52,13 +52,31 @@ setwd(location)
 
 ### plot lambdas for beta = 0.5
 
-par(mfrow = c(2, length(treesizes)/2))
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pb.32.50.sim.0.poly.reg.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pb.64.50.sim.0.poly.reg.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pb.128.50.sim.0.poly.reg.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pb.256.50.sim.0.poly.reg.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pb.512.50.sim.0.poly.reg.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.sim.0.poly.reg.type.csv")
+
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
+
+png(filename = "Manuscript/SupportingInfo_Figs/fig.1.reg.png",width = 800, height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3, 2), 
+    mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
-  R <- sim.sets[[i]]
+  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
   lambda.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Signal ($\\hat{lambda}$), $s$, or $W$"),
[…]

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -50,14 +50,31 @@ setwd(location)
 
 
 ### plot lambdas
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/bal.32.50.sim.0.poly.mean.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/bal.64.50.sim.0.poly.mean.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/bal.128.50.sim.0.poly.mean.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/bal.256.50.sim.0.poly.mean.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/bal.512.50.sim.0.poly.mean.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/bal.1024.50.sim.0.poly.mean.type.csv")
 
-par(mfrow = c(2, length(treesizes)/2))
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
+
+png(filename = "Manuscript/SupportingInfo_Figs/fig.1.bal.png",width = 800, height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3, 2), 
+    mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
-  R <- sim.sets[[i]]
+  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
-  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.pectinate.runs.R

@@ -50,14 +50,31 @@ setwd(location)
 
 
 ### plot lambdas
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pect.32.50.sim.0.poly.mean.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pect.64.50.sim.0.poly.mean.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pect.128.50.sim.0.poly.mean.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pect.256.50.sim.0.poly.mean.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pect.512.50.sim.0.poly.mean.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pect.1024.50.sim.0.poly.mean.type.csv")
 
-par(mfrow = c(2, length(treesizes)/2))
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
+
+png(filename = "Manuscript/SupportingInfo_Figs/fig.1.pect.png",width = 800, height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3, 2), 
+    mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
-  R <- sim.sets[[i]]
+  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
-  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R

@@ -50,14 +50,31 @@ setwd(location)
 
 
 ### plot lambdas
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pb.32.50.sim.0.2.poly.mean.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pb.64.50.sim.0.2.poly.mean.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pb.128.50.sim.0.2.poly.mean.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pb.256.50.sim.0.2.poly.mean.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pb.512.50.sim.0.2.poly.mean.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.sim.0.2.poly.mean.type.csv")
 
-par(mfrow = c(2, length(treesizes)/2))
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
+
+png(filename = "Manuscript/SupportingInfo_Figs/fig.1.poly.png",width = 800, height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3, 2), 
+    mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
-  R <- sim.sets[[i]]
+  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
-  lambda.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
+  lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-20 Manuscript/PNAS-Adams-02.Rmd
SI

@@
 
-**Simulations.** Simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylognetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+**Simulations.** Simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
 We generated 100 random trees for each intersection of tree size and $\lambda$.

### 2020-08-20 Manuscript/PNAS-Adams-02.tex
vsn 2
SI

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=8.25cm]{fig.1.png}
+\caption{Response of Pagel's \(\lambda\) to increasing strength of
+Brownian motion. Gray line signifies the 1:1 line where the input value
+matches the estimate \(\hat\lambda\). At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).{}}
+\end{figure}
+
 \textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
@@
 \centering
-\includegraphics{fig.1.png}
-\caption{Response of Pagel's \(\lambda\) to increasing strength of
-Brownian motion. Gray line signifies the 1:1 line where the input value
-matches the estimate \(\hat\lambda\). At each input level, the dark
-black line represents the empirically derived expected value (mean) of
-\(\hat\lambda\), the red line is the standard deviation of
-\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
-\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+\includegraphics[width=8.25cm]{fig.2.png}
+\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate \(\hat\kappa\).
+At each input level, the dark black line represents the empirically derived expected value (mean) of
+\(\hat\kappa\), the red line is the standard deviation of
+\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
+\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
 skewed distributions).{}}
@@
 for comparison across datasets.
-
-\begin{figure}
-\centering
-\includegraphics{fig.2.png}
-\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
-Brownian motion. Gray line signifies the 1:1 line where the input value
-matches the estimate \(\hat\kappa\). At each input level, the dark black
-line represents the empirically derived expected value (mean) of
-\(\hat\kappa\), the red line is the standard deviation of
-\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
-\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
-skewed distributions).{}}
-\end{figure}
 
@@
 \centering
-\includegraphics{fig.3.png}
-\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to
-increasing strength of Brownian motion.{}}
+\includegraphics[width=8.25cm]{fig.3.png}
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.{}}
 \end{figure}
@@
 (\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
-phylognetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100
-random trees for each intersection of tree size and \(\lambda\). For
+phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated
+100 random trees for each intersection of tree size and \(\lambda\). For
 each \(\lambda\) within each tree size, continuous traits were then

### 2020-08-20 Manuscript/SupportingInfo.Rmd
SI

--- 2020-04-20 Manuscript/SupportingInfo.Rmd
@@
 ---
-title:
+title: "Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal"
+
+author: "Dean C. Adams, Erica K. Baken, and Michael L. Collyer"
 
@@
     fig_caption: yes
-    number_sections: no  #note: change to 'no' for most journals
     keep_tex: true
@@
 
-csl: evolution.csl
+csl: pnas.csl
 bibliography: Lambda-Refs.bib
-
 ---
-# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate
-
-**FOR DCA**
-
-1: balanced/symmetric trees
-2: Kappa and K_Z
-3: Beta from PGLS
-
-Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.
-
-## Simulations on differently shaped phylogenies
 
@@
 
-In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees.
-As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units).
-For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution.
-Using these, we estimated the degree of phylogenetic signal using $\lambda$.
+Here we provide additional supporting information referenced in the main document, which include additional analyses, and simulation results across a wider set of input conditions.
+As before, simulations were conducted on six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+We generated 100 random trees for each intersection of tree size and $\lambda$.
+For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
+For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and $\kappa$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
+
+## Simulations on Pectinate Phylogenies
+
+Results from simulations on pectinate phylogenies largely mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S1: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S1 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S1 blue line).
 \hfill\break
 
-*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).
+For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S2 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S2 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S2 blue line).
+\hfill\break
 
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_{\kappa}$ was more consistent (Fig.
+S3).
+Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
 
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S1.png")  
 ```
 
-\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break
+\textbf{Figure S1}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
 
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
-
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S2.png")  
 ```
 
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break
+\textbf{Figure S2}. Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
 
-## Literature Survey
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S3.png")  
+```
+
+\textbf{Figure S3}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
 
@@
 
-# References
+## Simulations on Balanced Phylogenies
 
- \setlength{\parindent}{-0.25in}
- \setlength{\leftskip}{0.25in}
- \setlength{\parskip}{8pt}
- \noindent
+Results from simulations on balanced phylogenies also mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S4: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S4 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S4 blue line).
+\hfill\break
 
-<div id="refs"></div>
+For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S5 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S5 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S5 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size (fig.
+S6).
+Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, though there was some affect of tree size with balanced phylogenies.
+But as before, its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S4.png")  
+```
+
+\textbf{Figure S4}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S5.png")  
+```
+
+\textbf{Figure S5}. Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S6.png")  
+```
+
+\textbf{Figure S6}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations on Phylogenies Containing Polytomies
+
+We also investigaged the effect of unresolved phylogenies on estimates of phylogenetic signal by adjusting pure-birth trees to have 50% collapsed nodes, following the procedures of [@MolinaVenegas2017].
+Results from simulations on phylogenies containing polytomies mirrored those found on fully resolved trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S7: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S7 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S7 blue line).
+\hfill\break
+
+For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S8 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S8 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$ (Fig.
+S8 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_{\kappa}$ was more consistent (Fig.
+S9).
+Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on $Z_{\kappa}$.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S7.png")  
+```
+
+\textbf{Figure S7}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S8.png")  
+```
+
+\textbf{Figure S8}.
+Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S9.png")  
+```
+
+\textbf{Figure S9}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations of Phylogenetic Regression and ANOVA
+
+We also investigated the effect of incorporating $\lambda$ in PGLS analyses (phylogenetic regression and ANOVA).
+Here, patterns for $\hat\lambda$, $\hat\kappa$, $Z_\lambda$ and $Z_\kappa$ were virtually identical to those found previously.
+In essence, $\hat\lambda$ was not suitable as a test statistic representing phylogenetic signal due to all properties previously shown (Figs.
+S10 & S11), $\hat\kappa$ was more appropriate, but  associated non-linearly with signal strength (Figs.
+S12 & 13), while $Z_\kappa$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S14 & S15).
+\hfill\break
+
+With respect to model performance, input parameters ($\beta$) were well estimated in phylogenetic regression  and ANOVA when $\lambda$ was incorporated (Figs S16 & S17), implying that $\lambda$ may be used as a tuning parameter for these models.
+Additionally, type I error and statistical power were unaffected by the inclusion of $\lambda$ (Figs S18 & S19).
+This latter result confirms earlier findings of [@Revell2010] for phylogenetic regression, and extends them to the case of phylogenetic ANOVA.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S10.png")  
+```
+
+\textbf{Figure S10}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S11.png")  
+```
+
+\textbf{Figure S11}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S12.png")  
+```
+
+\textbf{Figure S12}.
+Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S13.png")  
+```
+
+\textbf{Figure S13}.
+Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S14.png")  
+```
+
+\textbf{Figure S14}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S15.png")  
+```
+
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S16.png")  
+```
+
+\textbf{Figure S16}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S17.png")  
+```
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+
+\newpage
+
+## References

### 2020-08-20 Manuscript/SupportingInfo.tex
SI

@@
 \hypersetup{
+  pdftitle={Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
+  pdfauthor={Dean C. Adams, Erica K. Baken, and Michael L. Collyer},
   hidelinks,
@@
 
-\author{}
-\date{\vspace{-2.5em}}
+\title{Supporting Information for: A Standardized Effect Size for Evaluating
+and Comparing the Strength of Phylogenetic Signal}
+\author{Dean C. Adams, Erica K. Baken, and Michael L. Collyer}
+\date{}
 
 \begin{document}
-
-\hypertarget{supporting-information-for-estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}{%
-\section{Supporting Information for: Estimates of Phylogenetic Signal
-Based on Lambda are Often
-Inaccurate}\label{supporting-information-for-estimates-of-phylogenetic-signal-based-on-lambda-are-often-inaccurate}}
+\maketitle
 
 Here we provide additional supporting information referenced in the main
-document: additional analyses, and simulation results across a wider set
-of input conditions.
+document, which include additional analyses, and simulation results
+across a wider set of input conditions. As before, simulations were
+conducted on six different tree sizes (\(n=2^5, 2^6, \cdots, 2^{10}\)),
+and with differing levels of phylogenetic signal
+(\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100 random trees for
+each intersection of tree size and \(\lambda\). For each \(\lambda\)
+within each tree size, continuous traits were then simulated on each
+phylogeny under a BM model of evolution. For each set of 100 trees we
+measured the mean values of \(\hat{\lambda}\) and \(\kappa\), their
+standard deviation, and calculated the Shapiro-Wilk \(W\) statistic as a
+departure from normality (symmetry). For the latter, a value of \(1.0\)
+indicates normally distributed values, while departures from \(1.0\)
+indicate skewness.
 
-\hypertarget{simulations-on-differently-shaped-phylogenies}{%
-\subsection{Simulations on differently shaped
-phylogenies}\label{simulations-on-differently-shaped-phylogenies}}
+\hypertarget{simulations-on-pectinate-phylogenies}{%
+\subsection{Simulations on Pectinate
+Phylogenies}\label{simulations-on-pectinate-phylogenies}}
 
-In addition to using pure-birth phylogenies, we explored the effect of
-tree shape on our findings using both balanced and pectinate trees. As
-before, simulations were conducted on a range of tree sizes
-(\(n=2^5 - 2^{10}\)), and across a range of input levels of phylogenetic
-signal (\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units).
-For each \(n\) and \(\lambda_{in}\) combination, 50 replicates of a
-continuous trait were simulated using a Brownian motion model of
-evolution. Using these, we estimated the degree of phylogenetic signal
-using \(\lambda\). \hfill\break
+Results from simulations on pectinate phylogenies largely mirrored those
+found on pure-birth trees. For \(\hat{\lambda}\), the mean value
+increased with increasing input signal, but was negatively biased, and
+was less than the input value across most of its range (Fig. S1: black
+line). Additionally, the precision of \(\hat{\lambda}\) varied with
+differing input levels, with the greatest variation found at
+intermediate values of \(\lambda\) (Fig. S1 red line). Finally, the
+distribution of \(\hat{\lambda}\) was not normal, and became more skewed
+at more extreme values of \(\lambda\) (Fig. S1 blue line). \hfill\break
 
-\emph{Results}. As found with pure-birth phylogenies, estimates of
-\(\lambda\) varied \ldots.. ** FINISH THIS** (Fig. S1, S2).
+For \(\hat\kappa\), mean values increased with increasing phylogenetic
+signal, though as was found with pure-birth trees, the increase was
+nonlinear (Fig. S2 black line). Likewise, variation increased with
+increasing phylogenetic signal (Fig. S2 red line), though the
+distribution of \(\hat\kappa\) was more normally distributed throughout
+its range, and across different tree sizes, as compared with
+\(\hat{\lambda}\), thought there was some slight skew for high input
+levels of phylogenetic signal on large phylogenies (Fig. S2 blue line).
+\hfill\break
 
-\includegraphics[width=0.75\linewidth]{Fig1}
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
+increased with increasing input phylogenetic signal, but
+\(\hat{\lambda}\) was strongly affected by tree size, whereas
+\(Z_{\kappa}\) was more consistent (Fig. S3). Also, \(Z_{\kappa}\)
+increased more linearly with increasing levels of phylogenetic signal,
+and its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision.
 
-\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. Results obtained using balanced
-trees. \hfill\break
+\includegraphics[width=0.95\linewidth]{fig.S1}
 
-\includegraphics[width=0.75\linewidth]{Fig1}
+\textbf{Figure S1}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on pectinate trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
 
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. Results obtained using pectinate
-trees. \hfill\break
+\includegraphics[width=0.95\linewidth]{fig.S2}
 
-\hypertarget{literature-survey}{%
-\subsection{Literature Survey}\label{literature-survey}}
+\textbf{Figure S2}. Response of Blomberg's \(\kappa\) to increasing
+strength of Brownian motion on pectinate trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S3}
+
+\textbf{Figure S3}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_{\kappa}\) to increasing strength of Brownian motion.
+
+\newpage
+
+\hypertarget{simulations-on-balanced-phylogenies}{%
+\subsection{Simulations on Balanced
+Phylogenies}\label{simulations-on-balanced-phylogenies}}
+
+Results from simulations on balanced phylogenies also mirrored those
+found on pure-birth trees. For \(\hat{\lambda}\), the mean value
+increased with increasing input signal, but was negatively biased, and
+was less than the input value across most of its range (Fig. S4: black
+line). Additionally, the precision of \(\hat{\lambda}\) varied with
+differing input levels, with the greatest variation found at
+intermediate values of \(\lambda\) (Fig. S4 red line). Finally, the
+distribution of \(\hat{\lambda}\) was not normal, and became more skewed
+at more extreme values of \(\lambda\) (Fig. S4 blue line). \hfill\break
+
+For \(\hat\kappa\), mean values increased with increasing phylogenetic
+signal, though as was found with pure-birth trees, the increase was
+nonlinear (Fig. S5 black line). Likewise, variation increased with
+increasing phylogenetic signal (Fig. S5 red line), though the
+distribution of \(\hat\kappa\) was more normally distributed throughout
+its range, and across different tree sizes, as compared with
+\(\hat{\lambda}\), thought there was some slight skew for high input
+levels of phylogenetic signal on large phylogenies (Fig. S5 blue line).
+\hfill\break
+
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
+increased with increasing input phylogenetic signal, but
+\(\hat{\lambda}\) was strongly affected by tree size (fig.~S6). Also,
+\(Z_{\kappa}\) increased more linearly with increasing levels of
+phylogenetic signal, though there was some affect of tree size with
+balanced phylogenies. But as before, its standard deviation across input
+signal was more even across tree sizes, implying more consistent
+precision.
+
+\includegraphics[width=0.95\linewidth]{fig.S4}
+
+\textbf{Figure S4}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on balanced trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S5}
+
+\textbf{Figure S5}. Response of Blomberg's \(\kappa\) to increasing
+strength of Brownian motion on balanced trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S6}
+
+\textbf{Figure S6}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_{\kappa}\) to increasing strength of Brownian motion.
+
+\newpage
+
+\hypertarget{simulations-on-phylogenies-containing-polytomies}{%
+\subsection{Simulations on Phylogenies Containing
+Polytomies}\label{simulations-on-phylogenies-containing-polytomies}}
+
+We also investigaged the effect of unresolved phylogenies on estimates
+of phylogenetic signal by adjusting pure-birth trees to have 50\%
+collapsed nodes, following the procedures of (1). Results from
+simulations on phylogenies containing polytomies mirrored those found on
+fully resolved trees. For \(\hat{\lambda}\), the mean value increased
+with increasing input signal, but was negatively biased, and was less
+than the input value across most of its range (Fig. S7: black line).
+Additionally, the precision of \(\hat{\lambda}\) varied with differing
+input levels, with the greatest variation found at intermediate values
+of \(\lambda\) (Fig. S7 red line). Finally, the distribution of
+\(\hat{\lambda}\) was not normal, and became more skewed at more extreme
+values of \(\lambda\) (Fig. S7 blue line). \hfill\break
+
+For \(\hat\kappa\), mean values increased with increasing phylogenetic
+signal, though as was found with pure-birth trees, the increase was
+nonlinear (Fig. S8 black line). Likewise, variation increased with
+increasing phylogenetic signal (Fig. S8 red line), though the
+distribution of \(\hat\kappa\) was more normally distributed throughout
+its range, and across different tree sizes, as compared with
+\(\hat{\lambda}\) (Fig. S8 blue line). \hfill\break
+
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
+increased with increasing input phylogenetic signal, but
+\(\hat{\lambda}\) was strongly affected by tree size, whereas
+\(Z_{\kappa}\) was more consistent (Fig. S9). Also, \(Z_{\kappa}\)
+increased more linearly with increasing levels of phylogenetic signal,
+and its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision. Thus, polytomies do not exert
+an appreciable effect on \(Z_{\kappa}\).
+
+\includegraphics[width=0.95\linewidth]{fig.S7}
+
+\textbf{Figure S7}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the
+estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S8}
+
+\textbf{Figure S8}. Response of Blomberg's \(\kappa\) to increasing
+strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the
+estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S9}
+
+\textbf{Figure S9}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_{\kappa}\) to increasing strength of Brownian motion.
+
+\newpage
+
+\hypertarget{simulations-of-phylogenetic-regression-and-anova}{%
+\subsection{Simulations of Phylogenetic Regression and
+ANOVA}\label{simulations-of-phylogenetic-regression-and-anova}}
+
+We also investigated the effect of incorporating \(\lambda\) in PGLS
+analyses (phylogenetic regression and ANOVA). Here, patterns for
+\(\hat\lambda\), \(\hat\kappa\), \(Z_\lambda\) and \(Z_\kappa\) were
+virtually identical to those found previously. In essence,
+\(\hat\lambda\) was not suitable as a test statistic representing
+phylogenetic signal due to all properties previously shown (Figs. S10 \&
+S11), \(\hat\kappa\) was more appropriate, but associated non-linearly
+with signal strength (Figs. S12 \& 13), while \(Z_\kappa\) was found to
+be a superior effect size as compared with \(Z_\lambda\) (Figs S14 \&
+S15). \hfill\break
+
+With respect to model performance, input parameters (\(\beta\)) were
+well estimated in phylogenetic regression and ANOVA when \(\lambda\) was
+incorporated (Figs S16 \& S17), implying that \(\lambda\) may be used as
+a tuning parameter for these models. Additionally, type I error and
+statistical power were unaffected by the inclusion of \(\lambda\) (Figs
+S18 \& S19). This latter result confirms earlier findings of (2) for
+phylogenetic regression, and extends them to the case of phylogenetic
+ANOVA.
+
+\includegraphics[width=0.95\linewidth]{fig.S10}
+
+\textbf{Figure S10}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion in phylogenetic regression (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S11}
+
+\textbf{Figure S11}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S12}
+
+\textbf{Figure S12}. Response of Blomberg's \(\kappa\) to increasing
+strength of Brownian motion in phylogenetic regression (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S13}
+
+\textbf{Figure S13}. Response of Blomberg's \(\kappa\) to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S14}
+
+\textbf{Figure S14}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic
+regression (incorporating \(\lambda\)).
+
+\includegraphics[width=0.95\linewidth]{fig.S15}
+
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic
+ANOVA (incorporating \(\lambda\)).
+
+\includegraphics[width=0.95\linewidth]{fig.S16}
+
+\textbf{Figure S16}. Parameter estimates from phylogenetic regression
+(incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
+the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S17}
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic ANOVA
+(incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
+the estimate.
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
-
-\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
-\setlength{\parskip}{8pt} \noindent
+\subsection*{References}\label{references}}
+\addcontentsline{toc}{subsection}{References}
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
+1. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic signal;
+strong or negligible impacts of polytomies and branch length
+information? \emph{BMC evolutionary biology} 17(1):53.
+
+\leavevmode\hypertarget{ref-Revell2010}{}%
+2. Revell LJ (2010) Phylogenetic signal and linear regression on species
+data. \emph{Methods in Ecology and Evolution} 1:319--329.
 

### 2020-08-20 Figures/Figures_Script.R

@@ -1,580 +1,0 @@
-# Figures
-
-
-# Fig 1: lit hist ####
-
-library(readxl)
-
-tree.sizes <-  treesizes <- 2^(5:10)
-
-review <- read_excel("Literature_Review/1Summary.xlsx", sheet = "Since 2019 Lambda Vals", col_types = "numeric")
-colnames(review)
-published.lambdas <- review[,c(3,4)]
-published.lambdas.pruned <- published.lambdas[-which(published.lambdas > 1 | published.lambdas < 0),]
-published.lambdas.df <- as.data.frame(published.lambdas.pruned)
-colnames(published.lambdas.df) <- c("Published_Lambdas", "Taxa")
-
-published.lambdas.df$Taxa_Bins <- as.character(published.lambdas.df$Taxa)
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa<=tree.sizes[[1]])] <- "< 33"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[1]] & published.lambdas.df$Taxa<=tree.sizes[[2]])] <- "33 - 64"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[2]] & published.lambdas.df$Taxa<=tree.sizes[[3]])] <- "65 - 128"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[3]] & published.lambdas.df$Taxa<=tree.sizes[[4]])] <- "129 - 256"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[4]] & published.lambdas.df$Taxa<=tree.sizes[[5]])] <- "257 - 512"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[5]] & published.lambdas.df$Taxa<=tree.sizes[[6]])] <- "513 - 1024"
-published.lambdas.df$Taxa_Bins[which(published.lambdas.df$Taxa>tree.sizes[[6]])] <- "> 1024"
-
-published.lambdas.df$Taxa_Bins <- as.factor(published.lambdas.df$Taxa_Bins)
-published.lambdas.df$Taxa_Bins <- relevel(published.lambdas.df$Taxa_Bins, "513 - 1024")
[…]

### 2020-08-20 Figures/NewFig2.R

@@ -1,30 +1,0 @@
-# Compare variation across runs
-
-treesizes <-  2^(5:10)
-
-lambda <- read.csv("Data_Analyses/Munged_Data/Lambda_est_vars.csv", header=TRUE)
-Z.k <- read.csv("Data_Analyses/Munged_Data/Kappa_Z_norm_vars.csv", header=TRUE)
- colnames(lambda) <- colnames(Z.k) <- c("lambda.in","32","64","128","256","512","1024")
-kappa_128 <- read.csv("Data_Analyses/Sim_Data/PB_lambda_kappacomparison_128.csv")
-
-#Plot for SM
-library(latex2exp)
-jpeg("Fig2.new.jpeg", res = 120, quality = 100, width = 1000, height = 1000)
-par(mfrow=c(2,2))
-plot(kappa_128$lambda.est ~ kappa_128$lambda.input, 
-     xlab = TeX('$\\lambda_{in}$'), ylab = TeX('$\\lambda_{est}$'), pch = 19)
-title("A", adj=0)
-text(min(kappa_128$lambda.input),max(kappa_128$lambda.est),
-     "n=128",adj= c(0.1,1))
-plot(kappa_128$kappa.z ~ kappa_128$lambda.input, 
-     xlab = TeX('$\\lambda_{in}$'), ylab = TeX('$Z_K$'),pch = 19)
-title("B", adj=0)
-text(min(kappa_128$lambda.input),max(kappa_128$kappa.z),
-     "n=128",adj= c(0.1,1))
-boxplot(lambda[,-1],ylim=c(0,.1),ylab = TeX('$\\sigma^2$'), xlab = "n")
-title("C", adj=0)
-boxplot(Z.k[,-1],ylim=c(0,.1),ylab = TeX('$\\sigma^2$'), xlab = "n")
-title("D", adj=0)
-par(mfrow=c(1,1))
-
-dev.off()

### 2020-08-20 Data_Analyses/updated_simulations/simulation.analyses.ANOVA.runs.R
Adding v1 of figures s18 and s19 (power curves for ANOVA and reg analyses, respectively). Manuscript/fig.S18.png and Manuscript/fig.S19.png

@@ -211,3 +211,32 @@ dev.off()
 par(mfcol = c(1,1))
 
 
+### power curves
+library(dplyr)
+
+png(filename = "Manuscript/fig.S18.png", width = 800, 
+    height = 1000, units = "px", pointsize = 18)
+par(mfrow = c(3,2), 
+    mar = c(5,5,1,2))
+for(i in 1:length(treesizes)) {
+  df <- sim.sets[[i]][,c("beta.in", "lambda.in", "Pval.x")] 
+  df$beta.in <- as.factor(df$beta.in)
+  df$lambda.in <- as.factor(df$lambda.in)
+  new.df <- df %>% group_by(lambda.in, beta.in) %>% summarise(power = length(which(Pval.x < 0.05))/nsim)
+  new.df$beta.in <- as.character(new.df$beta.in)
+  new.df$beta.in <- as.numeric(new.df$beta.in)
+  lambda.cols <- rep(gray.colors(nlevels(new.df$lambda.in), start = 0.2, end = 0.9, alpha = 0.5), each = 5)
+  
+  plot(new.df$beta.in, new.df$power, pch = 19,  col = lambda.cols,
+             xlab = latex2exp::TeX("Input Effect ($\\beta_{in}$)"),
+             ylab = latex2exp::TeX("Statistical Power"),
+             ylim = c(0, 1))
+
+  for(j in 1:21){ points(new.df$beta.in[(j*5-4):(j*5)], new.df$power[(j*5-4):(j*5)], type = "l", 
+                         col = lambda.cols[(j*5-4):(j*5)])}
+
+ legend(title =  latex2exp::TeX("$\\lambda_{in}$"), x = .85, y = .47, title.adj = .5,
+        legend = levels(new.df$lambda.in)[c(21, 15, 11, 7, 1)], fill = unique(lambda.cols)[c(21, 15, 11, 7, 1)], cex = .8)
+ legend("bottomright", paste("N =", treesizes[i]), bty = "n", cex = 1.5)
+}
+dev.off()

### 2020-08-21 Manuscript/PNAS-Adams-02.MLC.Rmd
mlc comments

--- 2020-08-20 Manuscript/PNAS-Adams-02.Rmd
@@
 To be statistics representing phylogenetic signal, they should have reliable distributional properties, which could be revealed with simulation experiments.
-For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ follows a Bernoulli distribution (**add ref**); i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
-Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to tree size; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 (**add same ref**).
+For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ is a random variable that follows the distribution of a Bernoulli probability parameter [@Forbes2011]; i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
+Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to tree size; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 [@Forbes2011].
 Prior work [@Munkemuller_et_al2012] seems to support some of these conjectures, based superficially on statistical moments for a given tree size (mean, variance, skewness, and kurtosis; see Fig.
@@
 
-By contrast, for Blomberg's $\kappa$, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might follow a normal distribution, with values distributed symmetrically about the input value.
+By contrast, for Blomberg's $\kappa$, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might be a random variable that follows a normal distribution, with values distributed symmetrically about the input value [@Forbes2011].
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
@@
 Taken together these results reveal that $\hat\lambda$ inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
-Additional simulations (Supplemental Information) reveal that incorporating $\hat\lambda$ in PGLS anova and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
+Additional simulations (Supplemental Information) reveal that incorporating $\hat\lambda$ in PGLS ANOVA and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
 Thus, it is reasonable to incorporate $\hat\lambda$ in PGLS as a parameter for tuning the degree of phylogenetic signal in the dependent variables during the analysis.
@@
 where $\kappa_{obs}$ is the observed phylogenetic signal, and $\hat\mu_\kappa$ and $\hat\sigma_\kappa$ are the mean and standard deviation of the empirical sampling distribution of $\kappa$ obtained via permutation.
-The empirical sampling distribution of $\kappa$ is first transformed via Box-Cox to better adhere to the assumption of normality.
+The empirical sampling distribution of $\kappa$ is first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
 
@@
 
-To demonstrate the utility of $\hat{Z}_12$, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
+To demonstrate the utility of $\hat{Z}_{12}$, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
 4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].

### 2020-08-21 Manuscript/PNAS-Adams-02.MLC.tex
mlc comments

--- 2020-08-20 Manuscript/PNAS-Adams-02.tex
@@
 experiments. For instance, as a proportional random variable bounded by
-\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
-Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
-tree are scaled proportionally to the probability that data arise from a
-BM process. Given a known \(\lambda\) value used to generate random data
-on a tree, we would also expect that the mean of an empirical sampling
-distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
-the dispersion of \(\hat{\lambda}\) would be largest at intermediate
-values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
-range of \(\lambda\) with respect to tree size; the distribution of
+\(0\) and \(1\), we might expect that \(\hat{\lambda}\) is a random
+variable that follows the distribution of a Bernoulli probability
+parameter (31); i.e., branch lengths in a tree are scaled proportionally
+to the probability that data arise from a BM process. Given a known
+\(\lambda\) value used to generate random data on a tree, we would also
+expect that the mean of an empirical sampling distribution of
+\(\hat{\lambda}\) would approximately equal \(\lambda\); the dispersion
+of \(\hat{\lambda}\) would be largest at intermediate values of
+\(\lambda\), \(\hat{\lambda}\) would be predictable over the range of
+\(\lambda\) with respect to tree size; the distribution of
 \(\hat{\lambda}\) would be symmetric at intermediate values of
@@
 intermediate values of \(\lambda\), becoming more leptokurtic toward 0
-and 1 (\textbf{add same ref}). Prior work (18) seems to support some of
-these conjectures, based superficially on statistical moments for a
-given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
-ref. (18)). However, because the ``strength of Brownian motion'' was
-simulated as a varied weighted-average of data simulated on trees with
-\(\lambda=0\) and \(\lambda=1\) and not as prescribed values of
-\(\lambda\) (18), interpretation of these patterns is challenging.
+and 1 (31). Prior work (18) seems to support some of these conjectures,
+based superficially on statistical moments for a given tree size (mean,
+variance, skewness, and kurtosis; see Fig. 2 of ref. (18)). However,
+because the ``strength of Brownian motion'' was simulated as a varied
+weighted-average of data simulated on trees with \(\lambda=0\) and
+\(\lambda=1\) and not as prescribed values of \(\lambda\) (18),
+interpretation of these patterns is challenging.
 
@@
 we might expect that for any \(\lambda\) used to generate data,
-estimates of \(\kappa\) might follow a normal distribution, with values
-distributed symmetrically about the input value. This attribute seemed
-less reasonable based on the simulations performed by Münkemüller et
-al.~(18), which suggested that distributions were positively skewed and
-that Blomberg's \(\kappa\) might not behave as a statistic that follows
-a normal distribution. However, because their simulations used a
-weighted combination of simulated phylogenetic signal strengths, strong
-inferences are not possible (and distributional attributes were not the
-intended result of their simulations). Thus, for both Pagel's
-\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
-across a range of \(\lambda\) used to generate data would be valuable
-for adjudicating the reliability of these statistics as effect sizes.
-Furthermore, the expected values of these statistics appear to vary with
-tree size (18), making comparisons across studies challenging.
-Therefore, transformation of these statistics into \(Z\)-scores would
-allow evaluation of the efficacy of each statistic to yield effect sizes
-that could be used for comparisons of the strength of phylogenetic
-signal across traits and lineages.
+estimates of \(\kappa\) might be a random variable that follows a normal
+distribution, with values distributed symmetrically about the input
+value (31). This attribute seemed less reasonable based on the
+simulations performed by Münkemüller et al.~(18), which suggested that
+distributions were positively skewed and that Blomberg's \(\kappa\)
+might not behave as a statistic that follows a normal distribution.
+However, because their simulations used a weighted combination of
+simulated phylogenetic signal strengths, strong inferences are not
+possible (and distributional attributes were not the intended result of
+their simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's
+\(\kappa\), evaluation of statistical moments across a range of
+\(\lambda\) used to generate data would be valuable for adjudicating the
+reliability of these statistics as effect sizes. Furthermore, the
+expected values of these statistics appear to vary with tree size (18),
+making comparisons across studies challenging. Therefore, transformation
+of these statistics into \(Z\)-scores would allow evaluation of the
+efficacy of each statistic to yield effect sizes that could be used for
+comparisons of the strength of phylogenetic signal across traits and
+lineages.
 
@@
 simulations (Supplemental Information) reveal that incorporating
-\(\hat\lambda\) in PGLS anova and regression does not adversely affect
+\(\hat\lambda\) in PGLS ANOVA and regression does not adversely affect
 the statistical properties of PGLS parameter estimation or model
@@
 its expected value under the null hypothesis, and \(\sigma_\theta\) is
-its standard error (31--33). Typically, \(\theta_{obs}\) and
+its standard error (32--34). Typically, \(\theta_{obs}\) and
 \(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
 obtained from the distribution of \(\theta\) derived from parametric
-theory. However, recent advances in resampling theory (34--37) have
+theory. However, recent advances in resampling theory (35--38) have
 shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
@@
 
-Formalizing the suggestion of Adams and Collyer (38), an effect size for
+Formalizing the suggestion of Adams and Collyer (39), an effect size for
 \(\kappa\) may be found as:
@@
 via permutation. The empirical sampling distribution of \(\kappa\) is
-first transformed via Box-Cox to better adhere to the assumption of
-normality.
+first transformed via a Box-Cox transformation to better adhere to the
+assumption of normality.
 
@@
 (one-tailed) tests may be specified should the empirical situation
-require it (35, 37).
+require it (36, 38).
 
-To demonstrate the utility of \(\hat{Z}_12\), we compared \(Z_{\kappa}\)
-for two ecologically-relevant traits in plethodontid salamander (Fig.
-4): surface area to volume ratios (SA:V) and relative body width
-(\(\frac{BW}{SVL}\)) (39, 40). While both traits contained significant
-phylogenetic signal, tests based on \(\hat{Z}_{12}\) revealed that the
-degree of phylogenetic signal was significantly stronger in SA:V
-(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
-observation may be interpreted by the fact that tropical species --
-which form a monophyletic group within plethodontids -- display greater
-variation in SA:V, which covaries with disparity in their climatic
-niches (40). Thus, greater phylogenetic signal in SA:V is to be
-expected.
+To demonstrate the utility of \(\hat{Z}_{12}\), we compared
+\(Z_{\kappa}\) for two ecologically-relevant traits in plethodontid
+salamander (Fig. 4): surface area to volume ratios (SA:V) and relative
+body width (\(\frac{BW}{SVL}\)) (40, 41). While both traits contained
+significant phylogenetic signal, tests based on \(\hat{Z}_{12}\)
+revealed that the degree of phylogenetic signal was significantly
+stronger in SA:V (\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4).
+Biologically, this observation may be interpreted by the fact that
+tropical species -- which form a monophyletic group within plethodontids
+-- display greater variation in SA:V, which covaries with disparity in
+their climatic niches (41). Thus, greater phylogenetic signal in SA:V is
+to be expected.
 
@@
 developed here aligns well with other statistical approaches such as
-meta-analysis (31, 41, 42), where summary statistics across datasets are
+meta-analysis (32, 42, 43), where summary statistics across datasets are
 converted to standardized effect sizes for subsequent ``higher order''
@@
 resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
-measurement cancel out during the calculation of \(Z\) (43). Thus,
+measurement cancel out during the calculation of \(Z\) (44). Thus,
 \(Z_\kappa\) represents the strength of phylogenetic signal on a common
@@
 previous observations of inflated \(\kappa\) for incompletely resolved
-phylogenetic trees (18, 44). An interesting question is whether an
+phylogenetic trees (18, 45). An interesting question is whether an
 inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
@@
 relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
-individuals of 305 species, from which species means were obtained (39,
-40). A time-dated molecular phylogeny for the group (45) was pruned to
+individuals of 305 species, from which species means were obtained (40,
+41). A time-dated molecular phylogeny for the group (46) was pruned to
 match the species in the phenotypic dataset. The phylogenetic signal in
 each trait was then characterized using \(\kappa\), which was converted
-to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (46,
-47), and routines by the authors (\textbf{to be incorporated in
+to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (47,
+48), and routines by the authors (\textbf{to be incorporated in
 \texttt{geomorph} upon manuscript acceptance}).
@@
 
+\leavevmode\hypertarget{ref-Forbes2011}{}%
+31. Forbes C, Evans M, Hastings N, Peacock B (2011) \emph{Statistical
+distributions} (John Wiley \& Sons).
+
 \leavevmode\hypertarget{ref-Glass1976}{}%
-31. Glass GV (1976) Primary, secondary, and meta-analysis of research.
+32. Glass GV (1976) Primary, secondary, and meta-analysis of research.
 \emph{Educational Researcher} 5:3--8.
@@
 \leavevmode\hypertarget{ref-Cohen1988}{}%
-32. Cohen J (1988) \emph{Statistical power analysis for the behavioral
+33. Cohen J (1988) \emph{Statistical power analysis for the behavioral
 sciences} (Routledge).
@@
 \leavevmode\hypertarget{ref-Rosenthal1994}{}%
-33. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
+34. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
 H Hedges (Russell Sage Foundation), pp 231--244.
@@
 \leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
-34. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
+35. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
 phenotypic change for phenotypes described by high-dimensional data.
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
-35. Adams DC, Collyer ML (2016) On the comparison of the strength of
+36. Adams DC, Collyer ML (2016) On the comparison of the strength of
 morphological integration across morphometric datasets. \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-CollyerAdams2018}{}%
-36. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
+37. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
 models to high-dimensional data using residual randomization.
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
-37. Adams DC, Collyer ML (2019) Comparing the strength of modular
+38. Adams DC, Collyer ML (2019) Comparing the strength of modular
 signal, and evaluating alternative modular hypotheses, using covariance
@@
 \leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
-38. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
+39. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
 evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
@@
 \leavevmode\hypertarget{ref-Baken2019}{}%
-39. Baken EK, Adams DC (2019) Macroevolution of arboreality in
+40. Baken EK, Adams DC (2019) Macroevolution of arboreality in
 salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.
@@
 \leavevmode\hypertarget{ref-Baken2020}{}%
-40. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
+41. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
 desiccation‐related morphology in plethodontid salamanders as inferred
@@
 \leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
-41. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
+42. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
 meta-analysis} (Elsevier).
@@
 \leavevmode\hypertarget{ref-Arnqvist1995}{}%
-42. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
+43. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
 findings in ecology and evolution. \emph{Trends in Ecology and
@@
 \leavevmode\hypertarget{ref-SokalRohlf2012}{}%
-43. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
+44. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
 San Francisco). 4th Ed.
@@
 \leavevmode\hypertarget{ref-Davies2012}{}%
-44. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
+45. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
 resolved phylogenetic trees inflate estimates of phylogenetic
@@
 \leavevmode\hypertarget{ref-Bonett2017}{}%
-45. Bonett RM, Blair AL (2017) Evidence for complex life cycle
+46. Bonett RM, Blair AL (2017) Evidence for complex life cycle
 constraints on salamander body form diversification. \emph{Proceedings
@@
 \leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
-46. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
+47. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
 collection and analysis of geometric morphometric shape data.
@@
 \leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-47. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
+48. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
 for geometric morphometric analyses. R package version 3.3.1. Available

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.ANOVA.runs.R
editing all figs 3 to match the n = 1024 top option (previously up to n = 256 only).
Fixing order of supinfo figs S18 and S19: now S18 is Regression, S19 is ANOVA

@@ -141,14 +141,14 @@ par(mfrow = c(3,2),
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
 for(i in 1:2) {
-  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
+  R <- list("Result" = sim.sets[[zref[i]]], "file.name" = "filler")
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
   lambda.z.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
@@ -156,7 +156,7 @@ for(i in 1:2) {
   kappa.z.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{K}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   
   abline(h = sqrt(qchisq(0.95,1)))
   
@@ -168,7 +168,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
-plot(lambdas, LZ[,1], ylim = c(-3, 25), 
+plot(lambdas, LZ[,1], ylim = c(-3, 50), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
      ylab = latex2exp::TeX("$Z_{\\lambda}$"), type = "l")
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R

@@ -145,7 +145,7 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
@@ -153,7 +153,7 @@ for(i in 1:2) {
   kappa.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                ylab = latex2exp::TeX("$Z_{K}$"),
-               mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+               mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   
   abline(h = sqrt(qchisq(0.95,1)))
   
@@ -165,7 +165,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
-plot(lambdas, LZ[,1], ylim = c(-3, 25), 
+plot(lambdas, LZ[,1], ylim = c(-3, 50), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
      ylab = latex2exp::TeX("$Z_{\\lambda}$"), type = "l")
 for(i in 2:ncol(LZ)) points(lambdas, LZ[,i], type = "l", lwd = i)
@@ -175,7 +175,7 @@ legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 # Note! change ylim below, if needed.  Too difficult to automate!
 
 KZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$kappa.z, sim.sets[[j]]$lambda.in, mean))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R

@@ -147,7 +147,7 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
@@ -155,7 +155,7 @@ for(i in 1:2) {
   kappa.z.plot(R, pch = 19, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{K}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   
   abline(h = sqrt(qchisq(0.95,1)))
   
@@ -167,7 +167,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
-plot(lambdas, LZ[,1], ylim = c(-3, 25), 
+plot(lambdas, LZ[,1], ylim = c(-3, 50), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
      ylab = latex2exp::TeX("$Z_{\\lambda}$"), type = "l")
 for(i in 2:ncol(LZ)) points(lambdas, LZ[,i], type = "l", lwd = i)
@@ -177,7 +177,7 @@ legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 # Note! change ylim below, if needed.  Too difficult to automate!
 
 KZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$kappa.z, sim.sets[[j]]$lambda.in, mean))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -144,7 +144,7 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 60))
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
@@ -152,7 +152,7 @@ for(i in 1:2) {
   kappa.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                ylab = latex2exp::TeX("$Z_{K}$"),
-               mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+               mean.col = 1, mean.lwd = 3, ylim = c(-3, 60))
   
   abline(h = sqrt(qchisq(0.95,1)))
   
@@ -164,7 +164,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
-plot(lambdas, LZ[,1], ylim = c(-3, 25), 
+plot(lambdas, LZ[,1], ylim = c(-3, 60), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
      ylab = latex2exp::TeX("$Z_{\\lambda}$"), type = "l")
 for(i in 2:ncol(LZ)) points(lambdas, LZ[,i], type = "l", lwd = i)
@@ -174,7 +174,7 @@ legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 # Note! change ylim below, if needed.  Too difficult to automate!
 
 KZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$kappa.z, sim.sets[[j]]$lambda.in, mean))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R

@@ -144,7 +144,7 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
@@ -152,7 +152,7 @@ for(i in 1:2) {
   kappa.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                ylab = latex2exp::TeX("$Z_{K}$"),
-               mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
+               mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
   
   abline(h = sqrt(qchisq(0.95,1)))
   
@@ -164,7 +164,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
-plot(lambdas, LZ[,1], ylim = c(-3, 25), 
+plot(lambdas, LZ[,1], ylim = c(-3, 50), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
      ylab = latex2exp::TeX("$Z_{\\lambda}$"), type = "l")
 for(i in 2:ncol(LZ)) points(lambdas, LZ[,i], type = "l", lwd = i)
@@ -174,7 +174,7 @@ legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 # Note! change ylim below, if needed.  Too difficult to automate!
 
 KZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$kappa.z, sim.sets[[j]]$lambda.in, mean))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.ANOVA.runs.R
editing all figs 3 to match the n = 1024 top option (previously up to n = 256 only).
Fixing order of supinfo figs S18 and S19: now S18 is Regression, S19 is ANOVA.

Editing simulation scripts to match the figure outputs.

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.1.ANOVA.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S11.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -101,7 +101,7 @@ dev.off()
 
 ### plot kappas for beta = 0.5
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.2.ANOVA.png",width = 800, 
+png(filename = "Manuscript/fig.S13.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -110,7 +110,7 @@ for(i in 1:length(treesizes)) {
   Rr <- R$Result
   kappa.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Attributes"),
+              ylab = latex2exp::TeX("Estimated Phylogenetic Signal Attributes"),
               ylim = c(0, 2))
   
   points(lambdas, by(Rr$phy.signal, Rr$lambda.in, mean),
@@ -134,21 +134,21 @@ dev.off()
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.3.ANOVA.png",width = 800, 
+png(filename = "Manuscript/fig.S15.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/new.fig.1.highqual.png", width = 800, 
+png(filename = "Manuscript/fig.1.png", width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
@@ -99,7 +99,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/new.fig.2.highqual.png", width = 800, 
+png(filename = "Manuscript/fig.2.png", width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -131,21 +131,21 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/new.fig.3.highqual.png", width = 800, 
+png(filename = "Manuscript/fig.3.png", width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
 for(i in 1:2) {
-  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
+  R <- list("Result" = sim.sets[[zref[i]]], "file.name" = "filler")
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 25))
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R

@@ -68,7 +68,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.1.reg.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S10.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -76,7 +76,7 @@ for(i in 1:length(treesizes)) {
   Rr <- R$Result
   lambda.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Attributes"),
+              ylab = latex2exp::TeX("Estimated Phylogenetic Signal Attributes"),
               ylim = c(0, 1.1))
   
   points(lambdas, by(Rr$lambda.opt.lambda, Rr$lambda.in, mean),
@@ -100,7 +100,7 @@ dev.off()
 
 ### plot kappas for beta = 0.5
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.2.reg.png",width = 800, 
+png(filename = "Manuscript/fig.S12.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -109,7 +109,7 @@ for(i in 1:length(treesizes)) {
   Rr <- R$Result
   kappa.plot(R, pch = 19, beta = 0.5, col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal($\\lambda_{in}$)"),
-              ylab = latex2exp::TeX("Estimated Phylogenetic Attributes"),
+              ylab = latex2exp::TeX("Estimated Phylogenetic Signal Attributes"),
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/bal.1024.50
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.1.bal.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S4.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/SupportingInfo_Figs/fig.2.bal.png",width = 800, 
+png(filename = "Manuscript/fig.S5.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,21 +130,21 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/SupportingInfo_Figs/fig.3.bal.png",width = 800, 
+png(filename = "Manuscript/fig.S6.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
 for(i in 1:2) {
-  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
+  R <- list("Result" = sim.sets[[zref[i]]], "file.name" = "filler")
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.pectinate.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pect.1024.5
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.1.pect.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S1.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/SupportingInfo_Figs/fig.2.pect.png",width = 800, 
+png(filename = "Manuscript/fig.S2.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,21 +130,21 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/SupportingInfo_Figs/fig.3.pect.png",width = 800, 
+png(filename = "Manuscript/fig.S3.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
 for(i in 1:2) {
-  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
+  R <- list("Result" = sim.sets[[zref[i]]], "file.name" = "filler")
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-21 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/SupportingInfo_Figs/fig.1.poly.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S7.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/SupportingInfo_Figs/fig.2.poly.png",width = 800, 
+png(filename = "Manuscript/fig.S8.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,21 +130,21 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/SupportingInfo_Figs/fig.3.poly.png",width = 800, 
+png(filename = "Manuscript/fig.S9.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
 zref <- c(1, length(sim.sets))
 for(i in 1:2) {
-  R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
+  R <- list("Result" = sim.sets[[zref[i]]], "file.name" = "filler")
   maxZ <- max(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   minZ <- min(na.omit(cbind(R$Result$lambda.z, R$Result$kappa.z)))
   
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-23 Manuscript/PNAS-Adams-03.Rmd
vsn 3

--- 2020-08-21 Manuscript/PNAS-Adams-02.MLC.Rmd
@@
 Further, the precision of $\lambda$ varies with input signal.
-Another measure, Blomberg's $\kappa$, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal.
+Another measure, Blomberg's *K*, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal.
 However, that relationship is decidedly nonlinear.
-Thus, neither $\lambda$ nor $\kappa$ are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets.
-As an alternative, we propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
-We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
+Thus, neither $\lambda$ nor *K* are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets.
+As an alternative, we propose a standardized effect size based on *K*, ($Z_K$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
+We develop tests based on $Z_K$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
 Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
@@
   Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking.
-Here we develop a standardized effect size, $Z_\kappa$, which measures the strength of phylogenetic signal on a common statistical scale.
+Here we develop a standardized effect size, $Z_K$, which measures the strength of phylogenetic signal on a common statistical scale.
 We also provide a test statistic, $\hat{Z}_{12}$, for comparing the strength of phylogenetic signal across datasets.
-We find that two commonly used parameters (Pagel's $\lambda$ and Blomberg's $\kappa$), not converted to effect sizes, are unsuitable for this purpose.
+We find that two commonly used parameters (Pagel's $\lambda$ and Blomberg's *K*), not converted to effect sizes, are unsuitable for this purpose.
 Our effect-size procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
@@
 
-Another widely used measure of phylogenetic signal is Blomberg's $\kappa$ [@Blomberg_et_al2003], which characterizes phylogenetic signal as the ratio of observed trait variation to the amount of variation expected under Brownian motion.
-Blomberg's $\kappa$ can be treated as a test statistic by employing a permutation test to generate its sampling distribution [@Blomberg_et_al2003; @Adams2014a] for determining whether significant phylogenetic signal is present in data.
-Both $\lambda$ and $\kappa$ seem intuitive to interpret, as a value of $0$ for both corresponds to no phylogenetic signal, while a value of $1$ corresponds to the amount of phylogenetic signal expected under Brownian motion.
-Thus, it is tempting to regard both $\lambda$ and $\kappa$ as descriptive statistics that measure the relative strength of phylogenetic signal, providing an estimate of its magnitude for comparison.
+Another widely used measure of phylogenetic signal is Blomberg's *K* [@Blomberg_et_al2003], which characterizes phylogenetic signal as the ratio of observed trait variation to the amount of variation expected under Brownian motion.
+Blomberg's *K* can be treated as a test statistic by employing a permutation test to generate its sampling distribution [@Blomberg_et_al2003; @Adams2014a] for determining whether significant phylogenetic signal is present in data.
+Both $\lambda$ and *K* seem intuitive to interpret, as a value of $0$ for both corresponds to no phylogenetic signal, while a value of $1$ corresponds to the amount of phylogenetic signal expected under Brownian motion.
+Thus, it is tempting to regard both $\lambda$ and *K* as descriptive statistics that measure the relative strength of phylogenetic signal, providing an estimate of its magnitude for comparison.
 
-The appeal of Pagel's $\lambda$ and Blomberg's $\kappa$ as descriptive statistics is that they provide a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$ or $\kappa$, respectively, in a comparative sense [@DeMeester2019; @Pintanel2019;@Su2019].
+The appeal of Pagel's $\lambda$ and Blomberg's *K* as descriptive statistics is that they provide a basis for interpreting "weak" versus "strong" phylogenetic signal; i.e., small versus large values of $\hat{\lambda}$ or *K*, respectively, in a comparative sense [@DeMeester2019; @Pintanel2019;@Su2019].
 Nonetheless, an important question that has yet to be considered is whether such comparisons are analytically appropriate, and whether these statistics are, or can be, converted to effect sizes for comparative analyses across datasets.
@@
 
-By contrast, for Blomberg's $\kappa$, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of $\kappa$ might be a random variable that follows a normal distribution, with values distributed symmetrically about the input value [@Forbes2011].
+By contrast, for Blomberg's *K*, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of *K* might be a random variable that follows a normal distribution, with values distributed symmetrically about the input value [@Forbes2011].
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.
-[-@Munkemuller_et_al2012], which suggested that distributions were positively skewed and that Blomberg's $\kappa$ might not behave as a statistic that follows a normal distribution.
+[-@Munkemuller_et_al2012], which suggested that distributions were positively skewed and that Blomberg's *K* might not behave as a statistic that follows a normal distribution.
 However, because their simulations used a weighted combination of simulated phylogenetic signal strengths, strong inferences are not possible (and distributional attributes were not the intended result of their simulations).
-Thus, for both Pagel's $\lambda$ or Blomberg's $\kappa$, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the reliability of these statistics as effect sizes.
+Thus, for both Pagel's $\lambda$ or Blomberg's *K*, evaluation of statistical moments across a range of $\lambda$ used to generate data would be valuable for adjudicating the reliability of these statistics as effect sizes.
 Furthermore, the expected values of these statistics appear to vary with tree size [@Munkemuller_et_al2012], making comparisons across studies challenging.
@@
 
-Here we use simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and $\kappa$, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
+Here we use simulation experiments to compare the distributional attributes of $\hat{\lambda}$ and *K*, plus their effect sizes ($Z$-scores), across a range of tree size and phylogenetic signal strength.
 We find that estimates of $\hat{\lambda}$ are increasingly skewed at larger and smaller input levels of phylogenetic signal and at smaller tree sizes, vary widely for a given input value of $\lambda$, and that the precision of $\hat{\lambda}$ is not constant across its range.
-By contrast, estimates of $\kappa$ are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$, making $\kappa$ a more reliable statistic.
-We then propose an effect size based on $\kappa$, ($Z_{\kappa}$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes and signal strength, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.
+By contrast, estimates of *K* are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$, making *K* a more reliable statistic.
+We then propose an effect size based on *K*, ($Z_K$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes and signal strength, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.
 
@@
 
-**Lambda ($\lambda$) estimates of phylogenetic signal are innacurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
+**Lambda ($\lambda$) estimates of phylogenetic signal are inaccurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
 First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases, but it is negatively-biased (particularly for small tree sizes), and is consistently less than the input $\lambda$ value across most of its range (Fig.
@@
 Taken together these results reveal that $\hat\lambda$ inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
-Additional simulations (Supplemental Information) reveal that incorporating $\hat\lambda$ in PGLS ANOVA and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
+Additional simulations (Supporting Information) reveal that incorporating $\hat\lambda$ in PGLS ANOVA and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
 Thus, it is reasonable to incorporate $\hat\lambda$ in PGLS as a parameter for tuning the degree of phylogenetic signal in the dependent variables during the analysis.
@@
 
-**Kappa ($\kappa$) estimates of phylogenetic signal are more reliable.** Simulation results for $\hat\kappa$ demonstrate that $\hat\kappa$ displays better statistical properties.
-First, as expected, mean values of $\hat\kappa$ increase with increasing signal ($\lambda$) irrespective of tree size, though the increase does not scale linearly with input levels of phylogenetic signal (Fig.
+**Kappa (*K*) estimates of phylogenetic signal are more reliable.** Simulation results demonstrate that *K* displays better statistical properties.
+First, as expected, mean values of *K* increase with increasing signal ($\lambda$) irrespective of tree size, albeit non-linearly (Fig.
 2 black line).
-Additionally, the standard deviation of $\hat\kappa$ is consistent across tree sizes (Fig.
-2 red line), and while it increases with $\lambda$, it is always less than the corresponding mean.
-This finding is perhaps unsurprising, as $\hat\kappa$ is lower-bounded by 0, and is never large for small values of $\lambda$.
-Importantly, $\hat\kappa$ is normally distributed across the range of input $\lambda$, and remains consistent in this pattern regardless of tree size (Fig.
+Additionally, the standard deviation of *K* is consistent across tree sizes (Fig.
+2 red line), and while it increases with $\lambda$, it is always less than the mean.
+This finding is perhaps unsurprising, as *K* is lower-bounded by 0, and is never large for small values of $\lambda$.
+Importantly, *K* is normally distributed across the range of input $\lambda$; a consistent pattern regardless of tree size (Fig.
 2 blue line).
-This result differs from those of [@Munkemuller_et_al2012], where the skewing appears to be the result of combining random values generated independently, rather than being a property of $\kappa$ itself.
-Overall, these findings reveal that while $\kappa$ is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
+This result differs from those of [@Munkemuller_et_al2012], where the skewing appears to be the result of combining random values generated independently, rather than being a property of *K* itself.
+Overall, these findings reveal that while *K* is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
 
-**Effect sizes from $\kappa$ ($Z_{\kappa}$) better characterize phylogenetic signal.**
-To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and $\kappa$. Statistically, a standardized effect size may be found as:
+**Effect sizes from *K* ($Z_K$) better characterize phylogenetic signal.**
+To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and *K*. Statistically, a standardized effect size may be found as:
 
@@
 
-Formalizing the suggestion of Adams and Collyer [@AdamsCollyer2019],  an effect size for $\kappa$ may be found as:
+Formalizing the suggestion of Adams and Collyer [@AdamsCollyer2019],  an effect size for *K* may be found as:
 
 \begin{align}
-    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
+    Z_K=\frac{K_{obs}-\hat\mu_{K}}{\hat\sigma_{K}},
 \end{align}
 
-where $\kappa_{obs}$ is the observed phylogenetic signal, and $\hat\mu_\kappa$ and $\hat\sigma_\kappa$ are the mean and standard deviation of the empirical sampling distribution of $\kappa$ obtained via permutation.
-The empirical sampling distribution of $\kappa$ is first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
+where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of *K* obtained via permutation.
+The empirical sampling distribution of *K* is first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
 
 For $\lambda$, deriving an effect size is more challenging, as $\lambda$ does not have a sampling distribution from which the standard error and confidence intervals may be obtained, and estimates from the Hessian matrix from PGLS are unreliable [@Boettiger_et_al2012].
-Confidence intervals are therefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [**add ref: MLC thinks Boettiger paper?**].
+Confidence intervals are therefore generated for the values of $\lambda$ that intersect the log-likehihood profile for corresponding percentiles of the $\chi^2$ distribution used to compare the putative model to a null model with $\lambda = 0$ [@Orme2013].
 Thus, an effect size for $\lambda$ may be found as:
@@
 
-Here we evaluate the ability of $Z_{\lambda}$ and $Z_{\kappa}$ to characterize known levels of phylogenetic signal.
-Both $Z_{\lambda}$ and $Z_{\kappa}$ are associated with input phylogenetic signal ($\lambda$), indicating that both statistics capture the observed signal (Fig.
+Simulations reveal that both $Z_{\lambda}$ and $Z_K$ are associated with input phylogenetic signal ($\lambda$), indicating that both statistics capture the observed signal (Fig.
 3).
@@
 3).
-By contrast, $Z_{\kappa}$ is much more consistent across tree sizes, and increases more linearly with increasing levels of phylogenetic signal.
-Additionally, $Z_{\kappa}$ exhibits a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
-3), and its standard deviation across input signal is more consistent.
-This implies that similar levels of precision are found with $Z_{\kappa}$ across the range of input values.
-Thus between the two statistics, $Z_{\kappa}$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets.
+By contrast, $Z_K$ is much more consistent across tree sizes, and increases more linearly with increasing levels of phylogenetic signal.
+Additionally, $Z_K$ exhibits a much stronger association with phylogenetic signal strength as compared to tree size (Fig.
+3), and its standard deviation is more consistent, implying similar levels of precision across the range of input signal (Supporting Information).
+Thus, $Z_K$ is a more reliable measure of the strength of phylogenetic signal, and may be used to compare levels of phylogenetic signal across datasets.
 
@@
 \includegraphics[width=8.25cm]{fig.3.png}
-\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.{}}
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.{}}
 \end{figure}
@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} = \frac{\lvert Z_{K_1} - Z_{K_2}\rvert}{\sqrt{2}}
 \end{align}
 
-where $\kappa_1$, $\kappa_2$, $\hat\mu_{\kappa_1}$, $\hat\mu_{\kappa_2}$, $\hat\sigma_{\kappa_1}$, and $\hat\sigma_{\kappa_2}$ are as defined above for equation 2.
-The right side of the equation illustrates that if $Z_\kappa$ has already been calculated for two sampling distributions as in equation 2, the sampling distributions have unit variance for each of the $Z_\kappa$ statistics.
+where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2.
+The right side of the equation illustrates that if $Z_K$ has already been calculated for two sampling distributions, both have unit variance for each of the $Z_K$ statistics.
 Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
@@
 
-To demonstrate the utility of $\hat{Z}_{12}$, we compared  $Z_{\kappa}$ for two ecologically-relevant traits in plethodontid salamander (Fig.
+To demonstrate the utility of $\hat{Z}_{12}$, we compared  $Z_K$ for two ecologically-relevant traits in plethodontid salamander (Fig.
 4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
@@
 4).
-Biologically, this observation may be interpreted by the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
+Biologically, this observation may be interpreted by the fact that the tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
 Thus, greater phylogenetic signal in SA:V is to be expected.
@@
 (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](fig.4.png)
+(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](fig.4.png)
 
@@
 $Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean.
-We applied this concept to both $\lambda$ and $\kappa$, and found that $Z_\kappa$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
-First, values of $Z_\kappa$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the linear relationship between $Z_\kappa$ and input signal.
-Additionally, the precision of $Z_{\kappa}$ was more consistent across the range of input levels of phylogenetic signal.
-Thus, $Z_\kappa$ is a more reliable measure of the relative strength of phylogenetic signal, and places that effect on a common and comparable scale.
-We therefore recommend that future studies interested in evaluating the strength of phylogenetic signal incorporate $Z_\kappa$ as a statistical measure of this effect.
+We applied this concept to both $\lambda$ and *K*, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data.
+First, values of $Z_K$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the near linear relationship between $Z_K$ and input signal.
+Additionally, the precision of $Z_K$ was more consistent across the range of input levels of phylogenetic signal.
+Thus, $Z_K$ is a more reliable measure of the relative strength of phylogenetic signal, and places that effect on a common and comparable scale.
+We therefore recommend that future studies interested in evaluating the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect.
 
 Next we proposed a two-sample test ($\hat{Z}_{12}$), which provides a formal statistical procedure for determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another.
-Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $\kappa$ [e.g.,  @Blomberg_et_al2003].
-However, because $\kappa$ does not scale linearly with input levels of phylogenetic signal (Fig.
+Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as *K* [e.g.,  @Blomberg_et_al2003].
+However, because *K* does not scale linearly with input levels of phylogenetic signal (Fig.
 2), and its variance increases with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012], it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale.
-By contrast, standardizing $\kappa$ to $Z_\kappa$ via equation 2 alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
+By contrast, standardizing *K* to $Z_K$ via equation 2 alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets.
 Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent "higher order" statistical summaries or comparisons.
@@
 
-One important advantage of the approach advocated here is that the resulting effect sizes ($Z_\kappa$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
-Thus, $Z_\kappa$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
-This means that the strength of phylogenetic signal may be compared across datasets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_\kappa$.
+One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012].
+Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviations -- regardless of the initial units and original scale of the phenotypic variables under investigation.
+This means that the strength of phylogenetic signal may be compared across datasets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$.
 For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$).
 In fact, our empirical example provided just such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$).
-Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $\kappa$ have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
+Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using *K* have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
 Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
 
-This study is not the first to compare $\lambda$ and $\kappa$ for their ability as statistics to measure phylogenetic signal.
-Our results for $\lambda$ and $\kappa$ values are consistent with those found in the simulations performed by Münkemüller et al.
+This study is not the first to compare $\lambda$ and *K* for their ability as statistics to measure phylogenetic signal.
+Our results for $\lambda$ and *K* values are consistent with those found in the simulations performed by Münkemüller et al.
 [-@Munkemuller_et_al2012], but that study investigated type I error rates and statistical power, finding that $\lambda$ performed better in both regards, irrespective of species number in trees.
-Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent $\kappa$ at intermediate phylogenetic signal strengths was observed (Fig.
+Although not the central focus of their study, the same tendency for variable $\lambda$ and consistent *K* at intermediate phylogenetic signal strengths was observed (Fig.
 2 of ref.
 [@Munkemuller_et_al2012]).
-Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that $\kappa$ but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
-Their work more thoroughly addressed previous observations of inflated $\kappa$ for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
-An interesting question is whether an inflated $\kappa$ value leads to an inflated $Z_\kappa$ or does a tendency of a particular tree to inflate estimates of $\kappa$ also inflate the values in random permutations of a test, in which case $Z_\kappa$ is robust to polytomies?
+Recent work by Molina-Venegas and Rodríguez [-@MolinaVenegas2017] found that *K* but not $\lambda$ tended to inflate the estimate of phylogenetic signal, leading to moderate type I and type II biases, if polytomic chronograms were used.
+Their work more thoroughly addressed previous observations of inflated *K* for incompletely resolved phylogenetic trees [@Davies2012; @Munkemuller_et_al2012].
+An interesting question is whether an inflated *K* value leads to an inflated $Z_K$ or does a tendency of a particular tree to inflate estimates of *K* also inflate the values in random permutations of a test, in which case $Z_K$ is robust to polytomies?
 We repeated the analyses in Figs.
 1 & 2, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (Supporting Information).
-This confirms that any tendency of incompletely resolved trees to inflate $\kappa$ as a descriptive statistic does not inflate $Z_\kappa$ as an effect size.
+This confirms that any tendency of incompletely resolved trees to inflate *K* as a descriptive statistic does not inflate $Z_K$ as an effect size.
 Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
@@
 However, the inability to estimate phylogenetic signal with $\lambda$ for data simulated with known phylogenetic signal is troublesome, and we recommend evolutionary biologists refrain from viewing it as a statistic to describe the amount of phylogenetic signal in the data.
-Interestingly, $\kappa$ -- when standardized to an effect size $Z_\kappa$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
+Interestingly, *K* -- when standardized to an effect size $Z_K$ -- is a better statistic for measuring the amount of phylogenetic signal in data simulated with respect to known levels of $\lambda$.
 Although $\lambda$ might be viewed as an important parameter for modifying the the conditional estimation of linear model coefficients with respect to phylogeny, it is neither a statistic that has meaningful comparative value as a measure of phylogenetic signal nor a statistic that lends itself well to reliable calculation of a test statistic.
-By contrast, $\kappa$ has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_\kappa$).
+By contrast, *K* has been shown here to be a reliable statistic, but only when standardized by the mean and standard deviation of its empirical sampling distribution (i.e., when converted to the effect size, $Z_K$).
 Because one has control over the number of permutations used in analysis, one can be assured with many permutations that the empirical sampling distribution is representative of true probability distributions [@AdamsCollyer2018b].
-Given the greater consistency in estimates of $Z_\kappa$ across tree sizes and input signal, it is difficult to imagine a hypothesis test that can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
+Given the greater consistency in estimates of $Z_K$ across tree sizes and input signal, it is difficult to imagine a hypothesis test that can improve equation 4 for efficiently comparing phylogenetic signal for different traits, different trees, or a combination of both.
 
@@
 For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
-For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and $\kappa$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and *K*, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
 For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
@@
 A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.
-The phylogenetic signal in each trait was then characterized using $\kappa$, which was converted to its effect size ($Z_\kappa$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph], and routines by the authors (**to be incorporated in `geomorph` upon manuscript acceptance**).
+The phylogenetic signal in each trait was then characterized using *K*, which was converted to its effect size ($Z_K$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph], and routines by the authors (**to be incorporated in `geomorph` upon acceptance**).
 

### 2020-08-23 Manuscript/PNAS-Adams-03.tex
vsn 3

--- 2020-08-21 Manuscript/PNAS-Adams-02.MLC.tex
@@
 comparisons are generally lacking. Here we develop a standardized effect
-size, \(Z_\kappa\), which measures the strength of phylogenetic signal
-on a common statistical scale. We also provide a test statistic,
+size, \(Z_K\), which measures the strength of phylogenetic signal on a
+common statistical scale. We also provide a test statistic,
 \(\hat{Z}_{12}\), for comparing the strength of phylogenetic signal
 across datasets. We find that two commonly used parameters (Pagel's
-\(\lambda\) and Blomberg's \(\kappa\)), not converted to effect sizes,
-are unsuitable for this purpose. Our effect-size procedure enables
+\(\lambda\) and Blomberg's \emph{K}), not converted to effect sizes, are
+unsuitable for this purpose. Our effect-size procedure enables
 biologists to quantitatively address hypotheses that compare the
@@
 \(\lambda\) varies with input signal. Another measure, Blomberg's
-\(\kappa\), is more consistent across a range of tree sizes, and
-exhibits a positive relationship with input levels of phylogenetic
-signal. However, that relationship is decidedly nonlinear. Thus, neither
-\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
-the strength of phylogenetic signal, and comparing that signal across
+\emph{K}, is more consistent across a range of tree sizes, and exhibits
+a positive relationship with input levels of phylogenetic signal.
+However, that relationship is decidedly nonlinear. Thus, neither
+\(\lambda\) nor \emph{K} are suitable as effect sizes for measuring the
+strength of phylogenetic signal, and comparing that signal across
 datasets. As an alternative, we propose a standardized effect size based
-on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
-phylogenetic signal more reliably than does \(\lambda\), and places that
-signal on a common scale for statistical comparison. We develop tests
-based on \(Z_\kappa\) to provide a mechanism for formally comparing the
-strength of phylogenetic signal across datasets, in much the same manner
-as effect sizes may be used to summarize patterns in quantitative
-meta-analysis. Our approach extends the phylogenetic comparative toolkit
-to address hypotheses that compare the strength of phylogenetic signal
-between various phenotypic traits, even when those traits are found in
-different evolutionary lineages or have different units or scales.
+on \emph{K}, (\(Z_K\)), which measures the strength of phylogenetic
+signal more reliably than does \(\lambda\), and places that signal on a
+common scale for statistical comparison. We develop tests based on
+\(Z_K\) to provide a mechanism for formally comparing the strength of
+phylogenetic signal across datasets, in much the same manner as effect
+sizes may be used to summarize patterns in quantitative meta-analysis.
+Our approach extends the phylogenetic comparative toolkit to address
+hypotheses that compare the strength of phylogenetic signal between
+various phenotypic traits, even when those traits are found in different
+evolutionary lineages or have different units or scales.
 \end{abstract}
@@
 Another widely used measure of phylogenetic signal is Blomberg's
-\(\kappa\) (12), which characterizes phylogenetic signal as the ratio of
+\emph{K} (12), which characterizes phylogenetic signal as the ratio of
 observed trait variation to the amount of variation expected under
-Brownian motion. Blomberg's \(\kappa\) can be treated as a test
-statistic by employing a permutation test to generate its sampling
-distribution (12, 16) for determining whether significant phylogenetic
-signal is present in data. Both \(\lambda\) and \(\kappa\) seem
-intuitive to interpret, as a value of \(0\) for both corresponds to no
-phylogenetic signal, while a value of \(1\) corresponds to the amount of
-phylogenetic signal expected under Brownian motion. Thus, it is tempting
-to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
-measure the relative strength of phylogenetic signal, providing an
-estimate of its magnitude for comparison.
+Brownian motion. Blomberg's \emph{K} can be treated as a test statistic
+by employing a permutation test to generate its sampling distribution
+(12, 16) for determining whether significant phylogenetic signal is
+present in data. Both \(\lambda\) and \emph{K} seem intuitive to
+interpret, as a value of \(0\) for both corresponds to no phylogenetic
+signal, while a value of \(1\) corresponds to the amount of phylogenetic
+signal expected under Brownian motion. Thus, it is tempting to regard
+both \(\lambda\) and \emph{K} as descriptive statistics that measure the
+relative strength of phylogenetic signal, providing an estimate of its
+magnitude for comparison.
 
-The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
-descriptive statistics is that they provide a basis for interpreting
-``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
-values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
-comparative sense (28--30). Nonetheless, an important question that has
-yet to be considered is whether such comparisons are analytically
-appropriate, and whether these statistics are, or can be, converted to
-effect sizes for comparative analyses across datasets. To be statistics
-representing phylogenetic signal, they should have reliable
-distributional properties, which could be revealed with simulation
-experiments. For instance, as a proportional random variable bounded by
-\(0\) and \(1\), we might expect that \(\hat{\lambda}\) is a random
-variable that follows the distribution of a Bernoulli probability
-parameter (31); i.e., branch lengths in a tree are scaled proportionally
-to the probability that data arise from a BM process. Given a known
-\(\lambda\) value used to generate random data on a tree, we would also
-expect that the mean of an empirical sampling distribution of
-\(\hat{\lambda}\) would approximately equal \(\lambda\); the dispersion
-of \(\hat{\lambda}\) would be largest at intermediate values of
-\(\lambda\), \(\hat{\lambda}\) would be predictable over the range of
-\(\lambda\) with respect to tree size; the distribution of
-\(\hat{\lambda}\) would be symmetric at intermediate values of
-\(\lambda\) and more skewed toward values of 0 or 1; and that the
-distribution of \(\hat{\lambda}\) will be more platykurtic at
+The appeal of Pagel's \(\lambda\) and Blomberg's \emph{K} as descriptive
+statistics is that they provide a basis for interpreting ``weak'' versus
+``strong'' phylogenetic signal; i.e., small versus large values of
+\(\hat{\lambda}\) or \emph{K}, respectively, in a comparative sense
+(28--30). Nonetheless, an important question that has yet to be
+considered is whether such comparisons are analytically appropriate, and
+whether these statistics are, or can be, converted to effect sizes for
+comparative analyses across datasets. To be statistics representing
+phylogenetic signal, they should have reliable distributional
+properties, which could be revealed with simulation experiments. For
+instance, as a proportional random variable bounded by \(0\) and \(1\),
+we might expect that \(\hat{\lambda}\) is a random variable that follows
+the distribution of a Bernoulli probability parameter (31); i.e., branch
+lengths in a tree are scaled proportionally to the probability that data
+arise from a BM process. Given a known \(\lambda\) value used to
+generate random data on a tree, we would also expect that the mean of an
+empirical sampling distribution of \(\hat{\lambda}\) would approximately
+equal \(\lambda\); the dispersion of \(\hat{\lambda}\) would be largest
+at intermediate values of \(\lambda\), \(\hat{\lambda}\) would be
+predictable over the range of \(\lambda\) with respect to tree size; the
+distribution of \(\hat{\lambda}\) would be symmetric at intermediate
+values of \(\lambda\) and more skewed toward values of 0 or 1; and that
+the distribution of \(\hat{\lambda}\) will be more platykurtic at
 intermediate values of \(\lambda\), becoming more leptokurtic toward 0
@@
 
-By contrast, for Blomberg's \(\kappa\), which is positively unbounded,
-we might expect that for any \(\lambda\) used to generate data,
-estimates of \(\kappa\) might be a random variable that follows a normal
+By contrast, for Blomberg's \emph{K}, which is positively unbounded, we
+might expect that for any \(\lambda\) used to generate data, estimates
+of \emph{K} might be a random variable that follows a normal
 distribution, with values distributed symmetrically about the input
@@
 simulations performed by Münkemüller et al.~(18), which suggested that
-distributions were positively skewed and that Blomberg's \(\kappa\)
-might not behave as a statistic that follows a normal distribution.
-However, because their simulations used a weighted combination of
-simulated phylogenetic signal strengths, strong inferences are not
-possible (and distributional attributes were not the intended result of
-their simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's
-\(\kappa\), evaluation of statistical moments across a range of
-\(\lambda\) used to generate data would be valuable for adjudicating the
-reliability of these statistics as effect sizes. Furthermore, the
-expected values of these statistics appear to vary with tree size (18),
-making comparisons across studies challenging. Therefore, transformation
-of these statistics into \(Z\)-scores would allow evaluation of the
-efficacy of each statistic to yield effect sizes that could be used for
-comparisons of the strength of phylogenetic signal across traits and
-lineages.
+distributions were positively skewed and that Blomberg's \emph{K} might
+not behave as a statistic that follows a normal distribution. However,
+because their simulations used a weighted combination of simulated
+phylogenetic signal strengths, strong inferences are not possible (and
+distributional attributes were not the intended result of their
+simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's \emph{K},
+evaluation of statistical moments across a range of \(\lambda\) used to
+generate data would be valuable for adjudicating the reliability of
+these statistics as effect sizes. Furthermore, the expected values of
+these statistics appear to vary with tree size (18), making comparisons
+across studies challenging. Therefore, transformation of these
+statistics into \(Z\)-scores would allow evaluation of the efficacy of
+each statistic to yield effect sizes that could be used for comparisons
+of the strength of phylogenetic signal across traits and lineages.
 
 Here we use simulation experiments to compare the distributional
-attributes of \(\hat{\lambda}\) and \(\kappa\), plus their effect sizes
+attributes of \(\hat{\lambda}\) and \emph{K}, plus their effect sizes
 (\(Z\)-scores), across a range of tree size and phylogenetic signal
@@
 and that the precision of \(\hat{\lambda}\) is not constant across its
-range. By contrast, estimates of \(\kappa\) are more consistent across
+range. By contrast, estimates of \emph{K} are more consistent across
 tree sizes, and are normally distributed across the range of input
-levels of \(\lambda\), making \(\kappa\) a more reliable statistic. We
-then propose an effect size based on \(\kappa\), (\(Z_{\kappa}\)), which
-provides consistent estimates of the strength of phylogenetic signal
-across tree sizes and signal strength, and facilitates quantitative
-comparisons of the relative strength of phylogenetic signal across
-datasets.
+levels of \(\lambda\), making \emph{K} a more reliable statistic. We
+then propose an effect size based on \emph{K}, (\(Z_K\)), which provides
+consistent estimates of the strength of phylogenetic signal across tree
+sizes and signal strength, and facilitates quantitative comparisons of
+the relative strength of phylogenetic signal across datasets.
 
@@
 \textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
-innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
+inaccurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
 distributional expectations of a Bernoulli variable were mostly upheld.
@@
 across tree sizes and across the range of input values. Additional
-simulations (Supplemental Information) reveal that incorporating
+simulations (Supporting Information) reveal that incorporating
 \(\hat\lambda\) in PGLS ANOVA and regression does not adversely affect
@@
 
-\textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
-reliable.} Simulation results for \(\hat\kappa\) demonstrate that
-\(\hat\kappa\) displays better statistical properties. First, as
-expected, mean values of \(\hat\kappa\) increase with increasing signal
-(\(\lambda\)) irrespective of tree size, though the increase does not
-scale linearly with input levels of phylogenetic signal (Fig. 2 black
-line). Additionally, the standard deviation of \(\hat\kappa\) is
-consistent across tree sizes (Fig. 2 red line), and while it increases
-with \(\lambda\), it is always less than the corresponding mean. This
-finding is perhaps unsurprising, as \(\hat\kappa\) is lower-bounded by
-0, and is never large for small values of \(\lambda\). Importantly,
-\(\hat\kappa\) is normally distributed across the range of input
-\(\lambda\), and remains consistent in this pattern regardless of tree
-size (Fig. 2 blue line). This result differs from those of (18), where
-the skewing appears to be the result of combining random values
-generated independently, rather than being a property of \(\kappa\)
-itself. Overall, these findings reveal that while \(\kappa\) is more
-reliable as an estimate of phylogenetic signal, the non-linear scaling
-with input signal implies that it should not be considered an effect
-size that measures the strength of phylogenetic signal on a common scale
-for comparison across datasets.
+\textbf{Kappa (\emph{K}) estimates of phylogenetic signal are more
+reliable.} Simulation results demonstrate that \emph{K} displays better
+statistical properties. First, as expected, mean values of \emph{K}
+increase with increasing signal (\(\lambda\)) irrespective of tree size,
+albeit non-linearly (Fig. 2 black line). Additionally, the standard
+deviation of \emph{K} is consistent across tree sizes (Fig. 2 red line),
+and while it increases with \(\lambda\), it is always less than the
+mean. This finding is perhaps unsurprising, as \emph{K} is lower-bounded
+by 0, and is never large for small values of \(\lambda\). Importantly,
+\emph{K} is normally distributed across the range of input \(\lambda\);
+a consistent pattern regardless of tree size (Fig. 2 blue line). This
+result differs from those of (18), where the skewing appears to be the
+result of combining random values generated independently, rather than
+being a property of \emph{K} itself. Overall, these findings reveal that
+while \emph{K} is more reliable as an estimate of phylogenetic signal,
+the non-linear scaling with input signal implies that it should not be
+considered an effect size that measures the strength of phylogenetic
+signal on a common scale for comparison across datasets.
 
-\textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
-characterize phylogenetic signal.} To measure the strength of
-phylogenetic signal on a common scale, we propose effect sizes
-(Z-scores) for both \(\lambda\) and \(\kappa\). Statistically, a
-standardized effect size may be found as:
+\textbf{Effect sizes from \emph{K} (\(Z_K\)) better characterize
+phylogenetic signal.} To measure the strength of phylogenetic signal on
+a common scale, we propose effect sizes (Z-scores) for both \(\lambda\)
+and \emph{K}. Statistically, a standardized effect size may be found as:
 
@@
 Formalizing the suggestion of Adams and Collyer (39), an effect size for
-\(\kappa\) may be found as:
+\emph{K} may be found as:
 
 \begin{align}
-    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
+    Z_K=\frac{K_{obs}-\hat\mu_{K}}{\hat\sigma_{K}},
 \end{align}
 
-where \(\kappa_{obs}\) is the observed phylogenetic signal, and
-\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
-deviation of the empirical sampling distribution of \(\kappa\) obtained
-via permutation. The empirical sampling distribution of \(\kappa\) is
-first transformed via a Box-Cox transformation to better adhere to the
-assumption of normality.
+where \(K_{obs}\) is the observed phylogenetic signal, and \(\hat\mu_K\)
+and \(\hat\sigma_K\) are the mean and standard deviation of the
+empirical sampling distribution of \emph{K} obtained via permutation.
+The empirical sampling distribution of \emph{K} is first transformed via
+a Box-Cox transformation to better adhere to the assumption of
+normality.
 
@@
 the \(\chi^2\) distribution used to compare the putative model to a null
-model with \(\lambda = 0\) {[}\textbf{add ref: MLC thinks Boettiger
-paper?}{]}. Thus, an effect size for \(\lambda\) may be found as:
+model with \(\lambda = 0\) (40). Thus, an effect size for \(\lambda\)
+may be found as:
 
@@
 
-Here we evaluate the ability of \(Z_{\lambda}\) and \(Z_{\kappa}\) to
-characterize known levels of phylogenetic signal. Both \(Z_{\lambda}\)
-and \(Z_{\kappa}\) are associated with input phylogenetic signal
-(\(\lambda\)), indicating that both statistics capture the observed
-signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
-little sense, as they are more strongly associated with tree size than
-they are with the actual phylogenetic signal in the data (Fig. 3). By
-contrast, \(Z_{\kappa}\) is much more consistent across tree sizes, and
-increases more linearly with increasing levels of phylogenetic signal.
-Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
-phylogenetic signal strength as compared to tree size (Fig. 3), and its
-standard deviation across input signal is more consistent. This implies
-that similar levels of precision are found with \(Z_{\kappa}\) across
-the range of input values. Thus between the two statistics,
-\(Z_{\kappa}\) is a more reliable measure of the strength of
-phylogenetic signal, and may be used to compare levels of phylogenetic
-signal across datasets.
+Simulations reveal that both \(Z_{\lambda}\) and \(Z_K\) are associated
+with input phylogenetic signal (\(\lambda\)), indicating that both
+statistics capture the observed signal (Fig. 3). However, effect sizes
+from \(\hat{\lambda}\) made little sense, as they are more strongly
+associated with tree size than they are with the actual phylogenetic
+signal in the data (Fig. 3). By contrast, \(Z_K\) is much more
+consistent across tree sizes, and increases more linearly with
+increasing levels of phylogenetic signal. Additionally, \(Z_K\) exhibits
+a much stronger association with phylogenetic signal strength as
+compared to tree size (Fig. 3), and its standard deviation is more
+consistent, implying similar levels of precision across the range of
+input signal (Supporting Information). Thus, \(Z_K\) is a more reliable
+measure of the strength of phylogenetic signal, and may be used to
+compare levels of phylogenetic signal across datasets.
 
@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} = \frac{\lvert Z_{K_1} - Z_{K_2}\rvert}{\sqrt{2}}
 \end{align}
 
-where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
-\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
-\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
-side of the equation illustrates that if \(Z_\kappa\) has already been
-calculated for two sampling distributions as in equation 2, the sampling
-distributions have unit variance for each of the \(Z_\kappa\)
-statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
-obtained from a standard normal distribution. Typically,
-\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
-(one-tailed) tests may be specified should the empirical situation
-require it (36, 38).
+where \(K_1\), \(K_2\), \(\hat\mu_{K_1}\), \(\hat\mu_{K_2}\),
+\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above for
+equation 2. The right side of the equation illustrates that if \(Z_K\)
+has already been calculated for two sampling distributions, both have
+unit variance for each of the \(Z_K\) statistics. Estimates of
+significance of \(\hat{Z}_{12}\) may be obtained from a standard normal
+distribution. Typically, \(\hat{Z}_{12}\) is considered a two-tailed
+test, however directional (one-tailed) tests may be specified should the
+empirical situation require it (36, 38).
 
-To demonstrate the utility of \(\hat{Z}_{12}\), we compared
-\(Z_{\kappa}\) for two ecologically-relevant traits in plethodontid
-salamander (Fig. 4): surface area to volume ratios (SA:V) and relative
-body width (\(\frac{BW}{SVL}\)) (40, 41). While both traits contained
-significant phylogenetic signal, tests based on \(\hat{Z}_{12}\)
-revealed that the degree of phylogenetic signal was significantly
-stronger in SA:V (\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4).
-Biologically, this observation may be interpreted by the fact that
-tropical species -- which form a monophyletic group within plethodontids
--- display greater variation in SA:V, which covaries with disparity in
-their climatic niches (41). Thus, greater phylogenetic signal in SA:V is
-to be expected.
+To demonstrate the utility of \(\hat{Z}_{12}\), we compared \(Z_K\) for
+two ecologically-relevant traits in plethodontid salamander (Fig. 4):
+surface area to volume ratios (SA:V) and relative body width
+(\(\frac{BW}{SVL}\)) (41, 42). While both traits contained significant
+phylogenetic signal, tests based on \(\hat{Z}_{12}\) revealed that the
+degree of phylogenetic signal was significantly stronger in SA:V
+(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
+observation may be interpreted by the fact that the tropical species --
+which form a monophyletic group within plethodontids -- display greater
+variation in SA:V, which covaries with disparity in their climatic
+niches (42). Thus, greater phylogenetic signal in SA:V is to be
+expected.
 
@@
 \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
-Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
-95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
+Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
+confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
 \end{figure}
@@
 easily interpretable as the strength of phylogenetic signal relative to
-the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
-and found that \(Z_\kappa\) was a better estimate of the strength of
-phylogenetic signal in phenotypic data. First, values of \(Z_\kappa\)
-more accurately tracked known changes in the magnitude of phylogenetic
-signal, as demonstrated by the linear relationship between \(Z_\kappa\)
-and input signal. Additionally, the precision of \(Z_{\kappa}\) was more
-consistent across the range of input levels of phylogenetic signal.
-Thus, \(Z_\kappa\) is a more reliable measure of the relative strength
-of phylogenetic signal, and places that effect on a common and
-comparable scale. We therefore recommend that future studies interested
-in evaluating the strength of phylogenetic signal incorporate
-\(Z_\kappa\) as a statistical measure of this effect.
+the mean. We applied this concept to both \(\lambda\) and \emph{K}, and
+found that \(Z_K\) was a better estimate of the strength of phylogenetic
+signal in phenotypic data. First, values of \(Z_K\) more accurately
+tracked known changes in the magnitude of phylogenetic signal, as
+demonstrated by the near linear relationship between \(Z_K\) and input
+signal. Additionally, the precision of \(Z_K\) was more consistent
+across the range of input levels of phylogenetic signal. Thus, \(Z_K\)
+is a more reliable measure of the relative strength of phylogenetic
+signal, and places that effect on a common and comparable scale. We
+therefore recommend that future studies interested in evaluating the
+strength of phylogenetic signal incorporate \(Z_K\) as a statistical
+measure of this effect.
 
@@
 phylogenetic signal across datasets using summary test values, such as
-\(\kappa\) (12). However, because \(\kappa\) does not scale linearly
-with input levels of phylogenetic signal (Fig. 2), and its variance
-increases with increasing strength of phylogenetic signal (18, 20), it
-should not be considered an effect size that measures the strength of
-phylogenetic signal on a common scale. By contrast, standardizing
-\(\kappa\) to \(Z_\kappa\) via equation 2 alleviates these concerns, and
-facilitates formal statistical comparisons of the strength of signal
-across datasets. Thus when viewed from this perspective, the approach
-developed here aligns well with other statistical approaches such as
-meta-analysis (32, 42, 43), where summary statistics across datasets are
-converted to standardized effect sizes for subsequent ``higher order''
-statistical summaries or comparisons. As such, our approach enables
-evolutionary biologists to quantitatively examine the relative strength
-of phylogenetic signal across a wide range of phenotypic traits, and
-thus opens the door for future discoveries that inform on how phenotypic
+\emph{K} (12). However, because \emph{K} does not scale linearly with
+input levels of phylogenetic signal (Fig. 2), and its variance increases
+with increasing strength of phylogenetic signal (18, 20), it should not
+be considered an effect size that measures the strength of phylogenetic
+signal on a common scale. By contrast, standardizing \emph{K} to \(Z_K\)
+via equation 2 alleviates these concerns, and facilitates formal
+statistical comparisons of the strength of signal across datasets. Thus
+when viewed from this perspective, the approach developed here aligns
+well with other statistical approaches such as meta-analysis (32, 43,
+44), where summary statistics across datasets are converted to
+standardized effect sizes for subsequent ``higher order'' statistical
+summaries or comparisons. As such, our approach enables evolutionary
+biologists to quantitatively examine the relative strength of
+phylogenetic signal across a wide range of phenotypic traits, and thus
+opens the door for future discoveries that inform on how phenotypic
 diversity accumulates in macroevolutionary time across the tree of life.
@@
 One important advantage of the approach advocated here is that the
-resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
-measurement cancel out during the calculation of \(Z\) (44). Thus,
-\(Z_\kappa\) represents the strength of phylogenetic signal on a common
-and comparable scale -- measured in standard deviations -- regardless of
-the initial units and original scale of the phenotypic variables under
+resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
+measurement cancel out during the calculation of \(Z\) (45). Thus,
+\(Z_K\) represents the strength of phylogenetic signal on a common and
+comparable scale -- measured in standard deviations -- regardless of the
+initial units and original scale of the phenotypic variables under
 investigation. This means that the strength of phylogenetic signal may
@@
 different units and scale, because those units have been standardized
-through their conversion to \(Z_\kappa\). For example, our approach
-could be utilized to determine whether the strength of phylogenetic
-signal (say, in response to ecological differentiation) is stronger in
-morphological traits (linear traits: \(mm\)), physiological traits
-(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
+through their conversion to \(Z_K\). For example, our approach could be
+utilized to determine whether the strength of phylogenetic signal (say,
+in response to ecological differentiation) is stronger in morphological
+traits (linear traits: \(mm\)), physiological traits (metabolic rate:
+\(\frac{O^2}{min}\)), or behavioral traits (aggression:
 \(\frac{\#{displays}}{second}\)). In fact, our empirical example
@@
 phylogenetic signal in traits of different dimensionality, as estimates
-of phylogenetic signal using \(\kappa\) have been generalized for
+of phylogenetic signal using \emph{K} have been generalized for
 multivariate data (16). Furthermore, tests based on \(\hat{Z}_{12}\) may
@@
 
-This study is not the first to compare \(\lambda\) and \(\kappa\) for
+This study is not the first to compare \(\lambda\) and \emph{K} for
 their ability as statistics to measure phylogenetic signal. Our results
-for \(\lambda\) and \(\kappa\) values are consistent with those found in
+for \(\lambda\) and \emph{K} values are consistent with those found in
 the simulations performed by Münkemüller et al.~(18), but that study
@@
 number in trees. Although not the central focus of their study, the same
-tendency for variable \(\lambda\) and consistent \(\kappa\) at
+tendency for variable \(\lambda\) and consistent \emph{K} at
 intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
 (18)). Recent work by Molina-Venegas and Rodríguez (21) found that
-\(\kappa\) but not \(\lambda\) tended to inflate the estimate of
+\emph{K} but not \(\lambda\) tended to inflate the estimate of
 phylogenetic signal, leading to moderate type I and type II biases, if
 polytomic chronograms were used. Their work more thoroughly addressed
-previous observations of inflated \(\kappa\) for incompletely resolved
-phylogenetic trees (18, 45). An interesting question is whether an
-inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
-tendency of a particular tree to inflate estimates of \(\kappa\) also
-inflate the values in random permutations of a test, in which case
-\(Z_\kappa\) is robust to polytomies? We repeated the analyses in Figs.
-1 \& 2, adjusting trees to have 50\% collapsed nodes, per the technique
-of Molina-Venegas and Rodríguez (21), and found results were consistent
-(Supporting Information). This confirms that any tendency of
-incompletely resolved trees to inflate \(\kappa\) as a descriptive
-statistic does not inflate \(Z_\kappa\) as an effect size. Furthermore,
-because comparison of effect sizes in a test is a comparison of
-locations of observed values in their sampling distributions, which
-would shift concomitantly because of this tendency, the \(Z_{12}\) test
-statistic in equation 4 appears to be robust in spite of unresolved
-trees.
+previous observations of inflated \emph{K} for incompletely resolved
+phylogenetic trees (18, 46). An interesting question is whether an
+inflated \emph{K} value leads to an inflated \(Z_K\) or does a tendency
+of a particular tree to inflate estimates of \emph{K} also inflate the
+values in random permutations of a test, in which case \(Z_K\) is robust
+to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
+to have 50\% collapsed nodes, per the technique of Molina-Venegas and
+Rodríguez (21), and found results were consistent (Supporting
+Information). This confirms that any tendency of incompletely resolved
+trees to inflate \emph{K} as a descriptive statistic does not inflate
+\(Z_K\) as an effect size. Furthermore, because comparison of effect
+sizes in a test is a comparison of locations of observed values in their
+sampling distributions, which would shift concomitantly because of this
+tendency, the \(Z_{12}\) test statistic in equation 4 appears to be
+robust in spite of unresolved trees.
 
@@
 biologists refrain from viewing it as a statistic to describe the amount
-of phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
-standardized to an effect size \(Z_\kappa\) -- is a better statistic for
+of phylogenetic signal in the data. Interestingly, \emph{K} -- when
+standardized to an effect size \(Z_K\) -- is a better statistic for
 measuring the amount of phylogenetic signal in data simulated with
@@
 of phylogenetic signal nor a statistic that lends itself well to
-reliable calculation of a test statistic. By contrast, \(\kappa\) has
-been shown here to be a reliable statistic, but only when standardized
-by the mean and standard deviation of its empirical sampling
-distribution (i.e., when converted to the effect size, \(Z_\kappa\)).
-Because one has control over the number of permutations used in
-analysis, one can be assured with many permutations that the empirical
-sampling distribution is representative of true probability
-distributions (10). Given the greater consistency in estimates of
-\(Z_\kappa\) across tree sizes and input signal, it is difficult to
-imagine a hypothesis test that can improve equation 4 for efficiently
-comparing phylogenetic signal for different traits, different trees, or
-a combination of both.
+reliable calculation of a test statistic. By contrast, \emph{K} has been
+shown here to be a reliable statistic, but only when standardized by the
+mean and standard deviation of its empirical sampling distribution
+(i.e., when converted to the effect size, \(Z_K\)). Because one has
+control over the number of permutations used in analysis, one can be
+assured with many permutations that the empirical sampling distribution
+is representative of true probability distributions (10). Given the
+greater consistency in estimates of \(Z_K\) across tree sizes and input
+signal, it is difficult to imagine a hypothesis test that can improve
+equation 4 for efficiently comparing phylogenetic signal for different
+traits, different trees, or a combination of both.
 
@@
 of 100 trees we measured the mean values of \(\hat{\lambda}\) and
-\(\kappa\), their standard deviation, and calculated the Shapiro-Wilk
+\emph{K}, their standard deviation, and calculated the Shapiro-Wilk
 \(W\) statistic as a departure from normality (symmetry). For the
@@
 relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
-individuals of 305 species, from which species means were obtained (40,
-41). A time-dated molecular phylogeny for the group (46) was pruned to
+individuals of 305 species, from which species means were obtained (41,
+42). A time-dated molecular phylogeny for the group (47) was pruned to
 match the species in the phenotypic dataset. The phylogenetic signal in
-each trait was then characterized using \(\kappa\), which was converted
-to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (47,
-48), and routines by the authors (\textbf{to be incorporated in
-\texttt{geomorph} upon manuscript acceptance}).
+each trait was then characterized using \emph{K}, which was converted to
+its effect size (\(Z_K\)) using \texttt{geomorph} 3.3.1 (48, 49), and
+routines by the authors (\textbf{to be incorporated in \texttt{geomorph}
+upon acceptance}).
 
@@
 
+\leavevmode\hypertarget{ref-Orme2013}{}%
+40. Orme D, et al. (2013) CAPER: Comparative analyses of phylogenetics
+and evolution in r. \emph{Methods in Ecology and Evolution} 3:145--151.
+
 \leavevmode\hypertarget{ref-Baken2019}{}%
-40. Baken EK, Adams DC (2019) Macroevolution of arboreality in
+41. Baken EK, Adams DC (2019) Macroevolution of arboreality in
 salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.
@@
 \leavevmode\hypertarget{ref-Baken2020}{}%
-41. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
+42. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
 desiccation‐related morphology in plethodontid salamanders as inferred
@@
 \leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
-42. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
+43. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
 meta-analysis} (Elsevier).
@@
 \leavevmode\hypertarget{ref-Arnqvist1995}{}%
-43. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
+44. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
 findings in ecology and evolution. \emph{Trends in Ecology and
@@
 \leavevmode\hypertarget{ref-SokalRohlf2012}{}%
-44. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
+45. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
 San Francisco). 4th Ed.
@@
 \leavevmode\hypertarget{ref-Davies2012}{}%
-45. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
+46. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
 resolved phylogenetic trees inflate estimates of phylogenetic
@@
 \leavevmode\hypertarget{ref-Bonett2017}{}%
-46. Bonett RM, Blair AL (2017) Evidence for complex life cycle
+47. Bonett RM, Blair AL (2017) Evidence for complex life cycle
 constraints on salamander body form diversification. \emph{Proceedings
@@
 \leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
-47. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
+48. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
 collection and analysis of geometric morphometric shape data.
@@
 \leavevmode\hypertarget{ref-AdamsGeomorph}{}%
-48. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
+49. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
 for geometric morphometric analyses. R package version 3.3.1. Available

### 2020-08-23 Manuscript/SupportingInfo.Rmd
vsn 3

@@
 
-## Simulations on Pectinate Phylogenies
+## Additional Results on Pure-Birth Phylogenies
 
-Results from simulations on pectinate phylogenies largely mirrored those found on pure-birth trees.
-For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
-S1: black line).
-Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
-S1 red line).
-Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
-S1 blue line).
-\hfill\break
-
-For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
-S2 black line).
-Likewise, variation increased with increasing phylogenetic signal (Fig.
-S2 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
-S2 blue line).
-\hfill\break
-
-With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_{\kappa}$ was more consistent (Fig.
-S3).
-Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+Results from the simulations on pure-birth trees revealed that both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was m1re consistent (Fig.
+S4).
+Additionally, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+This latter observation is seen in the upper four panels of Fig.
+S1, where the dispersion of simulation results (gray dots) is far more even at each input level of phylogenetic signal for $Z_K$ than it is for $Z_{\lambda}$.
 
@@
 
-\textbf{Figure S1}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations on Pectinate Phylogenies
+
+Results from simulations on pectinate phylogenies largely mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S2: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S2 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S2 blue line).
+\hfill\break
+
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S3 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S3 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S3 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S4).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
 
@@
 
-\textbf{Figure S2}. Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S2}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
 
@@
 
-\textbf{Figure S3}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S3}. Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S4.png")  
+```
+
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
 
@@
 For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
-S4: black line).
+S5: black line).
 Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
-S4 red line).
+S5 red line).
 Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
-S4 blue line).
-\hfill\break
-
-For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
-S5 black line).
-Likewise, variation increased with increasing phylogenetic signal (Fig.
-S5 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
 S5 blue line).
@@
 
-With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size (fig.
-S6).
-Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, though there was some affect of tree size with balanced phylogenies.
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S6 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S6 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S6 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size (fig.
+S7).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, though there was a slight effect of tree size with balanced phylogenies.
 But as before, its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
-
-```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S4.png")  
-```
-
-\textbf{Figure S4}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
 
@@
 
-\textbf{Figure S5}. Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S5}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
 
@@
 
-\textbf{Figure S6}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S6}. Response of Blomberg's \textit{K} to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S7.png")  
+```
+
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
 
@@
 For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
-S7: black line).
+S8: black line).
 Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
-S7 red line).
+S8 red line).
 Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
-S7 blue line).
-\hfill\break
-
-For $\hat\kappa$, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
-S8 black line).
-Likewise, variation increased with increasing phylogenetic signal (Fig.
-S8 red line), though the distribution of $\hat\kappa$ was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$ (Fig.
 S8 blue line).
@@
 
-With respect to effect sizes, both $Z_{\lambda}$ and $Z_{\kappa}$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_{\kappa}$ was more consistent (Fig.
-S9).
-Also, $Z_{\kappa}$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
-Thus, polytomies do not exert an appreciable effect on $Z_{\kappa}$.
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S9 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S9 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$ (Fig.
+S9 blue line).
+\hfill\break
 
-```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S7.png")  
-```
-
-\textbf{Figure S7}.
-Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
-Gray line signifies the 1:1 line where the input value matches the estimate.
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S10).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on $Z_K$.
 
@@
 \textbf{Figure S8}.
-Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
 Gray line signifies the 1:1 line where the input value matches the estimate.
@@
 
-\textbf{Figure S9}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S9}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S10.png")  
+```
+
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
 
@@
 We also investigated the effect of incorporating $\lambda$ in PGLS analyses (phylogenetic regression and ANOVA).
-Here, patterns for $\hat\lambda$, $\hat\kappa$, $Z_\lambda$ and $Z_\kappa$ were virtually identical to those found previously.
+Here, patterns for $\hat\lambda$, *K*, $Z_\lambda$ and $Z_\kappa$ were virtually identical to those found previously.
 In essence, $\hat\lambda$ was not suitable as a test statistic representing phylogenetic signal due to all properties previously shown (Figs.
-S10 & S11), $\hat\kappa$ was more appropriate, but  associated non-linearly with signal strength (Figs.
-S12 & 13), while $Z_\kappa$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S14 & S15).
+S11 & S12), *K* was more appropriate, but  associated non-linearly with signal strength (Figs.
+S13 & 14), while $Z_\kappa$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S15 & S16).
 \hfill\break
 
-With respect to model performance, input parameters ($\beta$) were well estimated in phylogenetic regression  and ANOVA when $\lambda$ was incorporated (Figs S16 & S17), implying that $\lambda$ may be used as a tuning parameter for these models.
-Additionally, type I error and statistical power were unaffected by the inclusion of $\lambda$ (Figs S18 & S19).
-This latter result confirms earlier findings of [@Revell2010] for phylogenetic regression, and extends them to the case of phylogenetic ANOVA.
-
-```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S10.png")  
-```
-
-\textbf{Figure S10}.
-Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
-Gray line signifies the 1:1 line where the input value matches the estimate.
+With respect to model performance, input parameters ($\beta$) were well estimated in phylogenetic regression  and ANOVA when $\lambda$ was incorporated (Figs S17 & S18), implying that $\lambda$ may be used as a tuning parameter for these models.
+Additionally, type I error and statistical power were unaffected by the inclusion of $\lambda$ (Figs S19 & S20), though power was reduced at smaller sample sizes in phylogenetic ANOVA, mirroring prior findings of [@AdamsCollyer2018b].
+These results generally confirmed earlier findings of [@Revell2010] for phylogenetic regression, and extend them to the case of phylogenetic ANOVA.
 
@@
 \textbf{Figure S11}.
-Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
 Gray line signifies the 1:1 line where the input value matches the estimate.
@@
 \textbf{Figure S12}.
-Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
 Gray line signifies the 1:1 line where the input value matches the estimate.
@@
 \textbf{Figure S13}.
-Response of Blomberg's \(\kappa\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
 Gray line signifies the 1:1 line where the input value matches the estimate.
@@
 
-\textbf{Figure S14}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+\textbf{Figure S14}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
 
@@
 
-\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
 
@@
 
-\textbf{Figure S16}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+\textbf{Figure S16}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
 
@@
 
-\textbf{Figure S17}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S18.png")  
+```
+
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S19.png")  
+```
+
+\textbf{Figure S19}. Power curves for phylogenetic regression (incorporating $\lambda$). Type I error is found on the far-left of the plot.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S20.png")  
+```
+
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating $\lambda$). Type I error is found on the far-left of the plot.
 

### 2020-08-23 Manuscript/SupportingInfo.tex
vsn 3

@@
 
+\hypertarget{additional-results-on-pure-birth-phylogenies}{%
+\subsection{Additional Results on Pure-Birth
+Phylogenies}\label{additional-results-on-pure-birth-phylogenies}}
+
+Results from the simulations on pure-birth trees revealed that both
+\(Z_{\lambda}\) and \(Z_K\) increased with increasing input phylogenetic
+signal, but \(\hat{\lambda}\) was strongly affected by tree size,
+whereas \(Z_K\) was m1re consistent (Fig. S4). Additionally, \(Z_K\)
+increased more linearly with increasing levels of phylogenetic signal,
+and its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision. This latter observation is
+seen in the upper four panels of Fig. S1, where the dispersion of
+simulation results (gray dots) is far more even at each input level of
+phylogenetic signal for \(Z_K\) than it is for \(Z_{\lambda}\).
+
+\includegraphics[width=0.95\linewidth]{fig.S1}
+
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion.
+
+\newpage
+
 \hypertarget{simulations-on-pectinate-phylogenies}{%
@@
 increased with increasing input signal, but was negatively biased, and
-was less than the input value across most of its range (Fig. S1: black
+was less than the input value across most of its range (Fig. S2: black
 line). Additionally, the precision of \(\hat{\lambda}\) varied with
 differing input levels, with the greatest variation found at
-intermediate values of \(\lambda\) (Fig. S1 red line). Finally, the
+intermediate values of \(\lambda\) (Fig. S2 red line). Finally, the
 distribution of \(\hat{\lambda}\) was not normal, and became more skewed
-at more extreme values of \(\lambda\) (Fig. S1 blue line). \hfill\break
+at more extreme values of \(\lambda\) (Fig. S2 blue line). \hfill\break
 
-For \(\hat\kappa\), mean values increased with increasing phylogenetic
-signal, though as was found with pure-birth trees, the increase was
-nonlinear (Fig. S2 black line). Likewise, variation increased with
-increasing phylogenetic signal (Fig. S2 red line), though the
-distribution of \(\hat\kappa\) was more normally distributed throughout
-its range, and across different tree sizes, as compared with
-\(\hat{\lambda}\), thought there was some slight skew for high input
-levels of phylogenetic signal on large phylogenies (Fig. S2 blue line).
-\hfill\break
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S3 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S3 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\), thought there
+was some slight skew for high input levels of phylogenetic signal on
+large phylogenies (Fig. S3 blue line). \hfill\break
 
-With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
-increased with increasing input phylogenetic signal, but
-\(\hat{\lambda}\) was strongly affected by tree size, whereas
-\(Z_{\kappa}\) was more consistent (Fig. S3). Also, \(Z_{\kappa}\)
-increased more linearly with increasing levels of phylogenetic signal,
-and its standard deviation across input signal was more even across tree
-sizes, implying more consistent precision.
-
-\includegraphics[width=0.95\linewidth]{fig.S1}
-
-\textbf{Figure S1}. Response of Pagel's \(\lambda\) to increasing
-strength of Brownian motion on pectinate trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size, whereas \(Z_K\) was more consistent
+(Fig. S4). Also, \(Z_K\) increased more linearly with increasing levels
+of phylogenetic signal, and its standard deviation across input signal
+was more even across tree sizes, implying more consistent precision.
 
@@
 
-\textbf{Figure S2}. Response of Blomberg's \(\kappa\) to increasing
+\textbf{Figure S2}. Response of Pagel's \(\lambda\) to increasing
 strength of Brownian motion on pectinate trees. Gray line signifies the
@@
 
-\textbf{Figure S3}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S3}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on pectinate trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S4}
+
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion.
 
@@
 increased with increasing input signal, but was negatively biased, and
-was less than the input value across most of its range (Fig. S4: black
+was less than the input value across most of its range (Fig. S5: black
 line). Additionally, the precision of \(\hat{\lambda}\) varied with
 differing input levels, with the greatest variation found at
-intermediate values of \(\lambda\) (Fig. S4 red line). Finally, the
+intermediate values of \(\lambda\) (Fig. S5 red line). Finally, the
 distribution of \(\hat{\lambda}\) was not normal, and became more skewed
-at more extreme values of \(\lambda\) (Fig. S4 blue line). \hfill\break
+at more extreme values of \(\lambda\) (Fig. S5 blue line). \hfill\break
 
-For \(\hat\kappa\), mean values increased with increasing phylogenetic
-signal, though as was found with pure-birth trees, the increase was
-nonlinear (Fig. S5 black line). Likewise, variation increased with
-increasing phylogenetic signal (Fig. S5 red line), though the
-distribution of \(\hat\kappa\) was more normally distributed throughout
-its range, and across different tree sizes, as compared with
-\(\hat{\lambda}\), thought there was some slight skew for high input
-levels of phylogenetic signal on large phylogenies (Fig. S5 blue line).
-\hfill\break
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S6 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S6 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\), though there
+was some slight skew for high input levels of phylogenetic signal on
+large phylogenies (Fig. S6 blue line). \hfill\break
 
-With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
-increased with increasing input phylogenetic signal, but
-\(\hat{\lambda}\) was strongly affected by tree size (fig.~S6). Also,
-\(Z_{\kappa}\) increased more linearly with increasing levels of
-phylogenetic signal, though there was some affect of tree size with
-balanced phylogenies. But as before, its standard deviation across input
-signal was more even across tree sizes, implying more consistent
-precision.
-
-\includegraphics[width=0.95\linewidth]{fig.S4}
-
-\textbf{Figure S4}. Response of Pagel's \(\lambda\) to increasing
-strength of Brownian motion on balanced trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size (fig.~S7). Also, \(Z_K\) increased more
+linearly with increasing levels of phylogenetic signal, though there was
+a slight effect of tree size with balanced phylogenies. But as before,
+its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision.
 
@@
 
-\textbf{Figure S5}. Response of Blomberg's \(\kappa\) to increasing
+\textbf{Figure S5}. Response of Pagel's \(\lambda\) to increasing
 strength of Brownian motion on balanced trees. Gray line signifies the
@@
 
-\textbf{Figure S6}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S6}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on balanced trees. Gray line signifies the
+1:1 line where the input value matches the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S7}
+
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion.
 
@@
 with increasing input signal, but was negatively biased, and was less
-than the input value across most of its range (Fig. S7: black line).
+than the input value across most of its range (Fig. S8: black line).
 Additionally, the precision of \(\hat{\lambda}\) varied with differing
 input levels, with the greatest variation found at intermediate values
-of \(\lambda\) (Fig. S7 red line). Finally, the distribution of
+of \(\lambda\) (Fig. S8 red line). Finally, the distribution of
 \(\hat{\lambda}\) was not normal, and became more skewed at more extreme
-values of \(\lambda\) (Fig. S7 blue line). \hfill\break
+values of \(\lambda\) (Fig. S8 blue line). \hfill\break
 
-For \(\hat\kappa\), mean values increased with increasing phylogenetic
-signal, though as was found with pure-birth trees, the increase was
-nonlinear (Fig. S8 black line). Likewise, variation increased with
-increasing phylogenetic signal (Fig. S8 red line), though the
-distribution of \(\hat\kappa\) was more normally distributed throughout
-its range, and across different tree sizes, as compared with
-\(\hat{\lambda}\) (Fig. S8 blue line). \hfill\break
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S9 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S9 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\) (Fig. S9 blue
+line). \hfill\break
 
-With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_{\kappa}\)
-increased with increasing input phylogenetic signal, but
-\(\hat{\lambda}\) was strongly affected by tree size, whereas
-\(Z_{\kappa}\) was more consistent (Fig. S9). Also, \(Z_{\kappa}\)
-increased more linearly with increasing levels of phylogenetic signal,
-and its standard deviation across input signal was more even across tree
-sizes, implying more consistent precision. Thus, polytomies do not exert
-an appreciable effect on \(Z_{\kappa}\).
-
-\includegraphics[width=0.95\linewidth]{fig.S7}
-
-\textbf{Figure S7}. Response of Pagel's \(\lambda\) to increasing
-strength of Brownian motion on pure-birth trees containing polytomies.
-Gray line signifies the 1:1 line where the input value matches the
-estimate.
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size, whereas \(Z_K\) was more consistent
+(Fig. S10). Also, \(Z_K\) increased more linearly with increasing levels
+of phylogenetic signal, and its standard deviation across input signal
+was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on \(Z_K\).
 
@@
 
-\textbf{Figure S8}. Response of Blomberg's \(\kappa\) to increasing
+\textbf{Figure S8}. Response of Pagel's \(\lambda\) to increasing
 strength of Brownian motion on pure-birth trees containing polytomies.
@@
 
-\textbf{Figure S9}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_{\kappa}\) to increasing strength of Brownian motion.
+\textbf{Figure S9}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the
+estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S10}
+
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_K\) to increasing strength of Brownian motion.
 
@@
 analyses (phylogenetic regression and ANOVA). Here, patterns for
-\(\hat\lambda\), \(\hat\kappa\), \(Z_\lambda\) and \(Z_\kappa\) were
-virtually identical to those found previously. In essence,
-\(\hat\lambda\) was not suitable as a test statistic representing
-phylogenetic signal due to all properties previously shown (Figs. S10 \&
-S11), \(\hat\kappa\) was more appropriate, but associated non-linearly
-with signal strength (Figs. S12 \& 13), while \(Z_\kappa\) was found to
-be a superior effect size as compared with \(Z_\lambda\) (Figs S14 \&
-S15). \hfill\break
+\(\hat\lambda\), \emph{K}, \(Z_\lambda\) and \(Z_\kappa\) were virtually
+identical to those found previously. In essence, \(\hat\lambda\) was not
+suitable as a test statistic representing phylogenetic signal due to all
+properties previously shown (Figs. S11 \& S12), \emph{K} was more
+appropriate, but associated non-linearly with signal strength (Figs. S13
+\& 14), while \(Z_\kappa\) was found to be a superior effect size as
+compared with \(Z_\lambda\) (Figs S15 \& S16). \hfill\break
 
@@
 well estimated in phylogenetic regression and ANOVA when \(\lambda\) was
-incorporated (Figs S16 \& S17), implying that \(\lambda\) may be used as
+incorporated (Figs S17 \& S18), implying that \(\lambda\) may be used as
 a tuning parameter for these models. Additionally, type I error and
 statistical power were unaffected by the inclusion of \(\lambda\) (Figs
-S18 \& S19). This latter result confirms earlier findings of (2) for
-phylogenetic regression, and extends them to the case of phylogenetic
-ANOVA.
-
-\includegraphics[width=0.95\linewidth]{fig.S10}
-
-\textbf{Figure S10}. Response of Pagel's \(\lambda\) to increasing
-strength of Brownian motion in phylogenetic regression (incorporating
-\(\lambda\)). Gray line signifies the 1:1 line where the input value
-matches the estimate.
+S19 \& S20), though power was reduced at smaller sample sizes in
+phylogenetic ANOVA, mirroring prior findings of (2). These results
+generally confirmed earlier findings of (3) for phylogenetic regression,
+and extend them to the case of phylogenetic ANOVA.
 
@@
 \textbf{Figure S11}. Response of Pagel's \(\lambda\) to increasing
-strength of Brownian motion in phylogenetic ANOVA (incorporating
+strength of Brownian motion in phylogenetic regression (incorporating
 \(\lambda\)). Gray line signifies the 1:1 line where the input value
@@
 
-\textbf{Figure S12}. Response of Blomberg's \(\kappa\) to increasing
-strength of Brownian motion in phylogenetic regression (incorporating
+\textbf{Figure S12}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
 \(\lambda\)). Gray line signifies the 1:1 line where the input value
@@
 
-\textbf{Figure S13}. Response of Blomberg's \(\kappa\) to increasing
-strength of Brownian motion in phylogenetic ANOVA (incorporating
+\textbf{Figure S13}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion in phylogenetic regression (incorporating
 \(\lambda\)). Gray line signifies the 1:1 line where the input value
@@
 
-\textbf{Figure S14}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic
-regression (incorporating \(\lambda\)).
+\textbf{Figure S14}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate.
 
@@
 \textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_{\kappa}\) to increasing strength of Brownian motion in phylogenetic
-ANOVA (incorporating \(\lambda\)).
+\(Z_K\) to increasing strength of Brownian motion in phylogenetic
+regression (incorporating \(\lambda\)).
 
@@
 
-\textbf{Figure S16}. Parameter estimates from phylogenetic regression
+\textbf{Figure S16}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_K\) to increasing strength of Brownian motion in phylogenetic ANOVA
+(incorporating \(\lambda\)).
+
+\includegraphics[width=0.95\linewidth]{fig.S17}
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression
 (incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S17}
+\includegraphics[width=0.95\linewidth]{fig.S18}
 
-\textbf{Figure S17}. Parameter estimates from phylogenetic ANOVA
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA
 (incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
 the estimate.
+
+\includegraphics[width=0.95\linewidth]{fig.S19}
+
+\textbf{Figure S19}. Power curves for phylogenetic regression
+(incorporating \(\lambda\)). Type I error is found on the far-left of
+the plot.
+
+\includegraphics[width=0.95\linewidth]{fig.S20}
+
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating
+\(\lambda\)). Type I error is found on the far-left of the plot.
 
@@
 
+\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
+2. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
+aggregation, biological challenges, and a refined permutation procedure.
+\emph{Evolution} 72(6):1204--1215.
+
 \leavevmode\hypertarget{ref-Revell2010}{}%
-2. Revell LJ (2010) Phylogenetic signal and linear regression on species
+3. Revell LJ (2010) Phylogenetic signal and linear regression on species
 data. \emph{Methods in Ecology and Evolution} 1:319--329.

### 2020-08-24 Manuscript/PNAS-Adams-03-EKB.Rmd
EKB final edits before submission 1

--- 2020-08-23 Manuscript/PNAS-Adams-03.Rmd
@@
 \includegraphics[width=8.25cm]{fig.2.png}
-\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of Brownian motion.
-Gray line signifies the 1:1 line where the input value matches the estimate \(\hat\kappa\).
-At each input level, the dark black line represents the empirically derived expected value (mean) of
-\(\hat\kappa\), the red line is the standard deviation of
-\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
-\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.
+~~Gray line signifies the 1:1 line where the input value matches the estimate \(\hat K\).~~ At each input level, the dark black line represents the empirically derived expected value (mean) of
+*K*, the red line is the standard deviation of
+*K*, and the blue line is Shapiro Wilks statistic of
+*K* (\(W=1.0\) signifies normality, \(W< 1.0\) represent
 skewed distributions).{}}
@@
 First, values of $Z_K$ more accurately tracked known changes in the magnitude of phylogenetic signal, as demonstrated by the near linear relationship between $Z_K$ and input signal.
-Additionally, the precision of $Z_K$ was more consistent across the range of input levels of phylogenetic signal.
+Additionally, the precision of $Z_K$ was more consistent across the range of input levels of phylogenetic signal (Fig S1; Supporting Information).
 Thus, $Z_K$ is a more reliable measure of the relative strength of phylogenetic signal, and places that effect on a common and comparable scale.
@@
 We repeated the analyses in Figs.
-1 & 2, adjusting trees to have 50% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (Supporting Information).
+1 & 2, adjusting trees to have 20% collapsed nodes ~~, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017],~~ and found results were consistent (Supporting Information).
 This confirms that any tendency of incompletely resolved trees to inflate *K* as a descriptive statistic does not inflate $Z_K$ as an effect size.
@@
 **Simulations.** Simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-We generated 100 random trees for each intersection of tree size and $\lambda$.
+We generated 50 random trees for each intersection of tree size and $\lambda$.
 For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
-For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and *K*, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and *K*, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
 For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.

### 2020-08-24 Manuscript/SupportingInfo-EKB.Rmd
EKB final edits before submission 1

--- 2020-05-01 Manuscript/SupportingInfo-EKB.Rmd
@@
 
-csl: evolution.csl
+csl: pnas.csl
 bibliography: Lambda-Refs.bib
-
 ---
-# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate
-
-Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.
-
-## Simulations on differently shaped phylogenies
 
@@
 
-In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees.
-As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units).
-For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution.
-Using these, we estimated the degree of phylogenetic signal using $\lambda$.
-\hfill\break
+Here we provide additional supporting information referenced in the main document, which include additional analyses, and simulation results across a wider set of input conditions.
+As before, simulations were conducted on six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+We generated 50 random trees for each intersection of tree size and $\lambda$.
+For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
+For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and $K$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
 
-*Results*.
-As found with pure-birth phylogenies, estimates of $\lambda$ varied most dramatically in simulations with fewer species and at intermediate values of lambda (Fig.
-S1, S2).
-Pectinate trees showed an interesting tendency to underestimate $/lambda$ as 0 for input $/lambda$ values.
-Most dramatically, some simulations on the largest pectinate phylogeny ($n = 1024$) estimated $/lambda = 0$ for input values as high as 0.75 (Fig.
-S2).
+## Additional Results on Pure-Birth Phylogenies
 
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("FigS1.png")
+Results from the simulations on pure-birth trees revealed that both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S4).
+Additionally, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+This latter observation is seen in the upper four panels of Fig.
+S1, where the dispersion of simulation results (gray dots) is far more even at each input level of phylogenetic signal for $Z_K$ than it is for $Z_{\lambda}$.
 
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S1.png")  
 ```
 
-\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("FigS2.png")
-
-```
-
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break
-
-## Simulations of phylogenetic regression and ANOVA
-
-```{r setup, include=FALSE}
-  library(knitr)
-  knitr::opts_chunk$set(echo = TRUE)
-```
-
-We analyzed the reliability of parameter estimations when used in phylogenetic regression and ANOVA.
-This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$).
-For each $\lambda$ input value, independent variables were then generated with a known relationship ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.
-
-*Results* As with the results from the regression analyses in the main text, we found limmited $\lambda$ estimation precision in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig.
-S3).
-Slope was reliably estimated across tree sizes and known $\lambda$ values (Fig.
-S4).
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("FigS3.png")
-
-```
-
-\textbf{Figure S3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic ANOVA. Results are from input $\beta$ values of $0.5$. \hfill\break
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("FigS4.png")
-
-```
-
-\textbf{Figure S4}. Precision of slope estimates across known input $\beta$ values of phylogenetic regression. Results shown include all input $\lambda$ values. \hfill\break
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
 
@@
 
-# References
+## Simulations on Pectinate Phylogenies
 
- \setlength{\parindent}{-0.25in}
- \setlength{\leftskip}{0.25in}
- \setlength{\parskip}{8pt}
- \noindent
+Results from simulations on pectinate phylogenies largely mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S2: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S2 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S2 blue line).
+\hfill\break
 
-<div id="refs"></div>
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S3 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S3 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S3 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S4).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S2.png")  
+```
+
+\textbf{Figure S2}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S3.png")  
+```
+
+\textbf{Figure S3}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pectinate trees.
+~~Gray line signifies the 1:1 line where the input value matches the estimate.~~ At each input level, the dark black line represents the empirically derived expected value (mean) of
+\textit{K}, the red line is the standard deviation of
+\textit{K}, and the blue line is Shapiro Wilks statistic of
+\textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S4.png")  
+```
+
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations on Balanced Phylogenies
+
+Results from simulations on balanced phylogenies also mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S5: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S5 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S5 blue line).
+\hfill\break
+
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S6 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S6 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S6 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size (fig.
+S7).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, though there was a slight effect of tree size with balanced phylogenies.
+But as before, its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S5.png")  
+```
+
+\textbf{Figure S5}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S6.png")  
+```
+
+\textbf{Figure S6}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on balanced trees.
+~~Gray line signifies the 1:1 line where the input value matches the estimate.~~ At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S7.png")  
+```
+
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations on Phylogenies Containing Polytomies
+
+We also investigaged the effect of unresolved phylogenies on estimates of phylogenetic signal by adjusting pure-birth trees to have 20% collapsed nodes ~~, following the procedures of [@MolinaVenegas2017]~~.
+Results from simulations on phylogenies containing polytomies mirrored those found on fully resolved trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S8: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S8 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S8 blue line).
+\hfill\break
+
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S9 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S9 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$ (Fig.
+S9 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S10).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on $Z_K$.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S8.png")  
+```
+
+\textbf{Figure S8}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S9.png")  
+```
+
+\textbf{Figure S9}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+~~Gray line signifies the 1:1 line where the input value matches the estimate.~~ At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S10.png")  
+```
+
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+
+\newpage
+
+## Simulations of Phylogenetic Regression and ANOVA
+
+We also investigated the effect of incorporating $\lambda$ in PGLS analyses (phylogenetic regression and ANOVA).
+Here, patterns for $\hat\lambda$, *K*, $Z_\lambda$ and $Z_K$ were virtually identical to those found previously.
+In essence, $\hat\lambda$ was not suitable as a test statistic representing phylogenetic signal due to all properties previously shown (Figs.
+S11 & S12), *K* was more appropriate, but  associated non-linearly with signal strength (Figs.
+S13 & 14), while $Z_K$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S15 & S16).
+\hfill\break
+
+With respect to model performance, input parameters ($\beta$) were well estimated in phylogenetic regression  and ANOVA when $\lambda$ was incorporated (Figs S17 & S18), implying that $\lambda$ may be used as a tuning parameter for these models.
+Additionally, type I error and statistical power were unaffected by the inclusion of $\lambda$ (Figs S19 & S20), though power was reduced at smaller sample sizes in phylogenetic ANOVA, mirroring prior findings of [@AdamsCollyer2018b].
+These results generally confirmed earlier findings of [@Revell2010] for phylogenetic regression, and extend them to the case of phylogenetic ANOVA.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S11.png")  
+```
+
+\textbf{Figure S11}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S12.png")  
+```
+
+\textbf{Figure S12}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S13.png")  
+```
+
+\textbf{Figure S13}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+~~Gray line signifies the 1:1 line where the input value matches the estimate.~~ At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S14.png")  
+```
+
+\textbf{Figure S14}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+~~Gray line signifies the 1:1 line where the input value matches the estimate.~~ At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S15.png")  
+```
+
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S16.png")  
+```
+
+\textbf{Figure S16}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S17.png")  
+```
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S18.png")  
+```
+
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S19.png")  
+```
+
+\textbf{Figure S19}. Power curves for phylogenetic regression (incorporating $\lambda$). Type I error is found on the far-left of the plot.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S20.png")  
+```
+
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating $\lambda$). Type I error is found on the far-left of the plot.
+
+\newpage
+
+## References

### 2020-08-24 Data_Analyses/Analysis_script.R

@@ -361,19 +361,19 @@ Nulls <- cbind(kappa_sav$random.K[-1], kappa_bwl$random.K[-1])
 colnames(Nulls) <- c("SAV.k.null", "BWL.k.null")
 Interval<-abs(qnorm(sav_vs_bwl$sample.r.sd.pk.stand)) # CI are represented without 
 
-png("Figures/Fig5_BC.png", width = 1200, height = 600, res = 150)
+png("Data_Analyses/EmpiricalExample/Fig.4.BC.png", width = 1200, height = 600, res = 150)
 par(mfrow=c(1,2), mar = c(5,5,2,2))
  # null distribution of kappas with observed kappas
-hist(Nulls[,1], col = alpha("red",0.5), xlim = c(0,.9), border = "red", breaks = 15, yaxs = "i", xlab = latex2exp::TeX('Null and Observed $\\kappa$s'), main = "")
+hist(Nulls[,1], col = alpha("red",0.5), xlim = c(0,.9), border = "red", breaks = 15, yaxs = "i", xlab = expression(paste("Null and Observed ", italic('K'), "s", sep = "")), main = "")
 hist(Nulls[,2], col = alpha("black", 0.5), border = "black", xlim = c(0,.9), breaks = 15, yaxs = "i", add = T)
 abline(v = kappa_sav$random.K[1], col = "red", lwd = 3)
 abline(v = kappa_bwl$random.K[1], col = "black", lwd = 3)
-legend("topright",c("SA:V","BW Rel"),fill=c("red","black"), inset = .03)
+legend("topright",c("SA:V","Rel BW"),fill=c("red","black"), inset = .03)
 box()
 
 plotCI(x = c(1,2), y = sav_vs_bwl$sample.z, xlim = c(.8, 2.2), 
        xlab = "Morphological Traits", yaxs="i", axes = F,
-       ylim = c(0,10), pch = 19, cex = 2, ylab = latex2exp::TeX('Effect Sizes ($\\Z_\\kappa$)'),
+       ylim = c(0,10), pch = 19, cex = 2, ylab = latex2exp::TeX('Effect Sizes ($\\Z_K$)'),
        uiw = Interval, col = c("red", "black"))
 axis(2, at=0:10, labels=0:10)
 axis(1, at=c(1,2), labels = c("SA:V", "Relative Body Width"))

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.ANOVA.runs.R
EKB's final commit before submission #1 (part 2, had to be split up due to cloud issues)

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/fig.S11.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S12.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -101,7 +101,7 @@ dev.off()
 
 ### plot kappas for beta = 0.5
 
-png(filename = "Manuscript/fig.S13.png",width = 800, 
+png(filename = "Manuscript/fig.S14.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -134,7 +134,7 @@ dev.off()
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
 
-png(filename = "Manuscript/fig.S15.png",width = 800, 
+png(filename = "Manuscript/fig.S16.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -190,7 +190,7 @@ dev.off()
 
 ### regression precision plot
 
-png(filename = "Manuscript/fig.S17.png",width = 800, 
+png(filename = "Manuscript/fig.S18.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -214,7 +214,7 @@ par(mfcol = c(1,1))
 ### power curves
 library(dplyr)
 
-png(filename = "Manuscript/fig.S19.png", width = 800, 
+png(filename = "Manuscript/fig.S20.png", width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R

@@ -71,12 +71,13 @@ png(filename = "Manuscript/fig.1.png", width = 800,
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
+ 
   R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
   lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
               xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
               ylab = latex2exp::TeX("Estimated Phylogenetic Signal Attributes"),
-              ylim = c(0, 1.1))
+              ylim = c(0, 1.1), cex.lab=1.25, cex.axis= 1.25)
   
   points(lambdas, by(Rr$lambda.opt.lambda, Rr$lambda.in, mean),
          type = "l", lwd = 3)
@@ -109,7 +110,7 @@ for(i in 1:length(treesizes)) {
   kappa.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
              xlab = latex2exp::TeX("Input Phylogenetic Signal($\\lambda_{in}$)"),
              ylab = latex2exp::TeX("Estimated Phylogenetic Signal Attributes"),
-             ylim = c(0, 2))
+             ylim = c(0, 2), cex.lab=1.25, cex.axis= 1.25)
   
   points(lambdas, by(Rr$phy.signal, Rr$lambda.in, mean),
          type = "l", lwd = 3)
@@ -131,7 +132,7 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/fig.3?.png", width = 800, 
+png(filename = "Manuscript/fig.3.png", width = 800, 
     height = 330, units = "px", pointsize = 10)
 
 par(mfrow = c(1,2), 
@@ -143,7 +144,7 @@ LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets
 
 plot(lambdas, LZ[,1], ylim = c(-3, 50), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
[…]

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.Regression.runs.R

@@ -68,7 +68,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/fig.S10.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S11.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -100,7 +100,7 @@ dev.off()
 
 ### plot kappas for beta = 0.5
 
-png(filename = "Manuscript/fig.S12.png",width = 800, 
+png(filename = "Manuscript/fig.S13.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -133,7 +133,7 @@ dev.off()
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
 
-png(filename = "Manuscript/fig.S14.png",width = 800, 
+png(filename = "Manuscript/fig.S15.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -188,7 +188,7 @@ legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 dev.off()
 
 ### regression precision plot
-png(filename = "Manuscript/fig.S16.png",width = 800, 
+png(filename = "Manuscript/fig.S17.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -212,7 +212,7 @@ par(mfcol = c(1,1))
 ### power curves
 library(dplyr)
 
-png(filename = "Manuscript/fig.S18.png", width = 800, 
+png(filename = "Manuscript/fig.S19.png", width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.balanced.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/bal.1024.50
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/fig.S4.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S5.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/fig.S5.png",width = 800, 
+png(filename = "Manuscript/fig.S6.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,7 +130,7 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/fig.S6.png",width = 800, 
+png(filename = "Manuscript/fig.S7.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -144,11 +144,7 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-<<<<<<< HEAD
                 mean.col = 1, mean.lwd = 3, ylim = c(-3, 80))
-=======
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 60))
->>>>>>> 4a60e041cdc823625c5b1cef2aea7c11e094e8a2
   abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
[…]

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.pectinate.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pect.1024.5
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/fig.S1.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S2.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/fig.S2.png",width = 800, 
+png(filename = "Manuscript/fig.S3.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,7 +130,7 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/fig.S3.png",width = 800, 
+png(filename = "Manuscript/fig.S4.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 
@@ -144,23 +144,14 @@ for(i in 1:2) {
   lambda.z.plot(R, pch = 19,col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
                 xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
                 ylab = latex2exp::TeX("$Z_{\\lambda}$"),
-<<<<<<< HEAD
                 mean.col = 1, mean.lwd = 3, ylim = c(-3, 35))
-=======
-                mean.col = 1, mean.lwd = 3, ylim = c(-3, 50))
->>>>>>> 4a60e041cdc823625c5b1cef2aea7c11e094e8a2
-  abline(h = sqrt(qchisq(0.95,1)))
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
   
[…]

### 2020-08-24 Data_Analyses/updated_simulations/simulation.analyses.polytomies.runs.R

@@ -66,7 +66,7 @@ sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.
 sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
 names(sim.sets) <- treesizes
 
-png(filename = "Manuscript/fig.S7.png",width = 800, height = 1000, units = "px", pointsize = 18)
+png(filename = "Manuscript/fig.S8.png",width = 800, height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
@@ -98,7 +98,7 @@ dev.off()
 
 
 ### plot kappas
-png(filename = "Manuscript/fig.S8.png",width = 800, 
+png(filename = "Manuscript/fig.S9.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 par(mfrow = c(3,2), 
     mar = c(5,5,1,2))
@@ -130,7 +130,7 @@ dev.off()
 
 ### plot z profiles
 ### only plot all points for smallest and largest, followed by means trends for all
-png(filename = "Manuscript/fig.S9.png",width = 800, 
+png(filename = "Manuscript/fig.S10.png",width = 800, 
     height = 1000, units = "px", pointsize = 18)
 
 par(mfrow = c(3,2), 

### 2020-08-25 Manuscript/PNAS-Adams-03-EKB.tex
03-MLC

--- 2020-08-23 Manuscript/PNAS-Adams-03.tex
@@
 \centering
-\includegraphics[width=8.25cm]{fig.1.png}
+\includegraphics[width=8.25cm]{{fig.1}.png}
 \caption{Response of Pagel's \(\lambda\) to increasing strength of
@@
 \centering
-\includegraphics[width=8.25cm]{fig.2.png}
-\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of Brownian motion.
-Gray line signifies the 1:1 line where the input value matches the estimate \(\hat\kappa\).
-At each input level, the dark black line represents the empirically derived expected value (mean) of
-\(\hat\kappa\), the red line is the standard deviation of
-\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
-\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+\includegraphics[width=8.25cm]{{fig.2}.png}
+\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.
+~~Gray line signifies the 1:1 line where the input value matches the estimate \(\hat K\).~~ At each input level, the dark black line represents the empirically derived expected value (mean) of
+*K*, the red line is the standard deviation of
+*K*, and the blue line is Shapiro Wilks statistic of
+*K* (\(W=1.0\) signifies normality, \(W< 1.0\) represent
 skewed distributions).{}}
@@
 \centering
-\includegraphics[width=8.25cm]{fig.3.png}
+\includegraphics[width=8.25cm]{{fig.3}.png}
 \caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.{}}
@@
 
-\begin{figure}
-\centering
-\includegraphics{fig.4.png}
-\caption{(A) Linear measures for relative body size, and regions of the
-body used to estimate surface area to volume (SA:V) ratios. (B)
-Permutation distributions of phylogenetic signal for SA:V and
-\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
-Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
-confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
-\end{figure}
+{[}(A) Linear measures for relative body size, and regions of the body
+used to estimate surface area to volume (SA:V) ratios. (B) Permutation
+distributions of phylogenetic signal for SA:V and \(\frac{BW}{SVL}\),
+with observed values shown as vertical bars. (C) Effect sizes (\(Z_K\))
+for SA:V and \(\frac{BW}{SVL}\), with their 95\% confidence intervals
+(CI not standardized by \(\sqrt(n)\)).{}{]}\{\{fig.4\}\}.png)
 
@@
 signal. Additionally, the precision of \(Z_K\) was more consistent
-across the range of input levels of phylogenetic signal. Thus, \(Z_K\)
-is a more reliable measure of the relative strength of phylogenetic
-signal, and places that effect on a common and comparable scale. We
-therefore recommend that future studies interested in evaluating the
-strength of phylogenetic signal incorporate \(Z_K\) as a statistical
-measure of this effect.
+across the range of input levels of phylogenetic signal (Fig S1;
+Supporting Information). Thus, \(Z_K\) is a more reliable measure of the
+relative strength of phylogenetic signal, and places that effect on a
+common and comparable scale. We therefore recommend that future studies
+interested in evaluating the strength of phylogenetic signal incorporate
+\(Z_K\) as a statistical measure of this effect.
 
@@
 to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
-to have 50\% collapsed nodes, per the technique of Molina-Venegas and
-Rodríguez (21), and found results were consistent (Supporting
+to have 20\% collapsed nodes \sout{, per the technique of Molina-Venegas
+and Rodríguez (21),} and found results were consistent (Supporting
 Information). This confirms that any tendency of incompletely resolved
@@
 (\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
-phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated
-100 random trees for each intersection of tree size and \(\lambda\). For
+phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 50
+random trees for each intersection of tree size and \(\lambda\). For
 each \(\lambda\) within each tree size, continuous traits were then
 simulated on each phylogeny under a BM model of evolution. For each set
-of 100 trees we measured the mean values of \(\hat{\lambda}\) and
+of 50 trees we measured the mean values of \(\hat{\lambda}\) and
 \emph{K}, their standard deviation, and calculated the Shapiro-Wilk

### 2020-08-25 Manuscript/PNAS-Adams-03-MLC.Rmd
03-MLC

--- 2020-08-24 Manuscript/PNAS-Adams-03-EKB.Rmd
@@
 **Lambda ($\lambda$) estimates of phylogenetic signal are inaccurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
-First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases, but it is negatively-biased (particularly for small tree sizes), and is consistently less than the input $\lambda$ value across most of its range (Fig.
-1 black line).
+First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases.
 Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values (Fig.1 red line), implying that the precision in estimating $\lambda$ varies across the range of input values.
-Additionally, standard deviations of $\hat{\lambda}$ are negatively associated with tree size, and for tress of 128 species or less, $\hat{\lambda}$ are quite variable, except for cases when $\lambda$ is near or equal to $1$.
-Third, the distributions of $\hat{\lambda}$ are not normal across its range, but become increasingly skewed at more extreme values of $\lambda$ (Fig.
+Third, the distributions of $\hat{\lambda}$ tend toward normal ddistributions at intermediate levels of $\lambda$ but become increasingly skewed at more extreme values of $\lambda$ (Fig.
 1 blue line).
 For small tree sizes, it is also clear that distributions are more platykurtic at intermediate values of $\hat{\lambda}$.
-Taken together these results reveal that $\hat\lambda$ inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
+However, the mean value of $\hat{\lambda}$ is negatively-biased (particularly for small tree sizes but also consistently across most of its range; Fig.
+1 black line) and standard deviations of $\hat{\lambda}$ are negatively associated with tree size.
+For tress of 128 species or less, $\hat{\lambda}$ are quite variable, except for cases when $\lambda$ is near or equal to $1$.
+Taken together these results reveal that $\hat\lambda$ is a biased statistic that inconsistently estimates  phylogenetic signal, both across tree sizes and across the range of input values.
 Additional simulations (Supporting Information) reveal that incorporating $\hat\lambda$ in PGLS ANOVA and regression does not adversely affect the statistical properties of PGLS parameter estimation or model evaluation (type I error, power, bias in coefficients).
@@
 \centering
-\includegraphics[width=8.25cm]{fig.2.png}
-\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.
-~~Gray line signifies the 1:1 line where the input value matches the estimate \(\hat K\).~~ At each input level, the dark black line represents the empirically derived expected value (mean) of
+\includegraphics[width=8.25cm]{{fig.2}.png}
+\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.  At each input level, the black line represents the empirically derived expected value (mean) of
 *K*, the red line is the standard deviation of
@@
 Additionally, the standard deviation of *K* is consistent across tree sizes (Fig.
-2 red line), and while it increases with $\lambda$, it is always less than the mean.
+2 red line), and while it increases with $\lambda$, it is always less than the mean (low coefficient of variation).
 This finding is perhaps unsurprising, as *K* is lower-bounded by 0, and is never large for small values of $\lambda$.
@@
 
-**Effect sizes from *K* ($Z_K$) better characterize phylogenetic signal.**
-To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and *K*. Statistically, a standardized effect size may be found as:
+**Effect sizes from *K* ($Z_K$) better characterize phylogenetic signal.**  To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and *K*.
+Statistically, a standardized effect size may be found as:
 
@@
 Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory.
-However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures.
+However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ simulated from permutation procedures.
 
@@
 where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of *K* obtained via permutation.
-The empirical sampling distribution of *K* is first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
+The empirical sampling distribution of *K* can be first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
 
@@
 \centering
-\includegraphics[width=8.25cm]{fig.3.png}
-\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.{}}
+\includegraphics[width=8.25cm]{{fig.3}.png}
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+Means from simulation runs are shown for comparative ease.
+Individual values from each simulation run are available in Supporting Informaton.
+}
 \end{figure}
@@
 We repeated the analyses in Figs.
-1 & 2, adjusting trees to have 20% collapsed nodes ~~, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017],~~ and found results were consistent (Supporting Information).
+1 & 2, adjusting trees to have 20% collapsed nodes, per the technique of Molina-Venegas and Rodríguez [-@MolinaVenegas2017], and found results were consistent (Supporting Information).
 This confirms that any tendency of incompletely resolved trees to inflate *K* as a descriptive statistic does not inflate $Z_K$ as an effect size.

### 2020-08-25 Manuscript/PNAS-Adams-03-MLC.tex
03-MLC

--- 2020-08-25 Manuscript/PNAS-Adams-03-EKB.tex
@@
 First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
-increases, but it is negatively-biased (particularly for small tree
-sizes), and is consistently less than the input \(\lambda\) value across
-most of its range (Fig. 1 black line). Second, the standard deviation of
-\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
-smallest at extreme values (Fig.1 red line), implying that the precision
-in estimating \(\lambda\) varies across the range of input values.
-Additionally, standard deviations of \(\hat{\lambda}\) are negatively
-associated with tree size, and for tress of 128 species or less,
-\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
-is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
-are not normal across its range, but become increasingly skewed at more
-extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
-it is also clear that distributions are more platykurtic at intermediate
-values of \(\hat{\lambda}\). Taken together these results reveal that
-\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
+increases. Second, the standard deviation of \(\hat{\lambda}\) is
+largest at intermediate values of \(\lambda\) and smallest at extreme
+values (Fig.1 red line), implying that the precision in estimating
+\(\lambda\) varies across the range of input values. Third, the
+distributions of \(\hat{\lambda}\) tend toward normal ddistributions at
+intermediate levels of \(\lambda\) but become increasingly skewed at
+more extreme values of \(\lambda\) (Fig. 1 blue line). For small tree
+sizes, it is also clear that distributions are more platykurtic at
+intermediate values of \(\hat{\lambda}\). However, the mean value of
+\(\hat{\lambda}\) is negatively-biased (particularly for small tree
+sizes but also consistently across most of its range; Fig. 1 black line)
+and standard deviations of \(\hat{\lambda}\) are negatively associated
+with tree size. For tress of 128 species or less, \(\hat{\lambda}\) are
+quite variable, except for cases when \(\lambda\) is near or equal to
+\(1\). Taken together these results reveal that \(\hat\lambda\) is a
+biased statistic that inconsistently estimates phylogenetic signal, both
 across tree sizes and across the range of input values. Additional
@@
 deviation of \emph{K} is consistent across tree sizes (Fig. 2 red line),
-and while it increases with \(\lambda\), it is always less than the
-mean. This finding is perhaps unsurprising, as \emph{K} is lower-bounded
-by 0, and is never large for small values of \(\lambda\). Importantly,
-\emph{K} is normally distributed across the range of input \(\lambda\);
-a consistent pattern regardless of tree size (Fig. 2 blue line). This
-result differs from those of (18), where the skewing appears to be the
-result of combining random values generated independently, rather than
-being a property of \emph{K} itself. Overall, these findings reveal that
-while \emph{K} is more reliable as an estimate of phylogenetic signal,
-the non-linear scaling with input signal implies that it should not be
-considered an effect size that measures the strength of phylogenetic
-signal on a common scale for comparison across datasets.
+and while it increases with \(\lambda\), it is always less than the mean
+(low coefficient of variation). This finding is perhaps unsurprising, as
+\emph{K} is lower-bounded by 0, and is never large for small values of
+\(\lambda\). Importantly, \emph{K} is normally distributed across the
+range of input \(\lambda\); a consistent pattern regardless of tree size
+(Fig. 2 blue line). This result differs from those of (18), where the
+skewing appears to be the result of combining random values generated
+independently, rather than being a property of \emph{K} itself. Overall,
+these findings reveal that while \emph{K} is more reliable as an
+estimate of phylogenetic signal, the non-linear scaling with input
+signal implies that it should not be considered an effect size that
+measures the strength of phylogenetic signal on a common scale for
+comparison across datasets.
 
@@
 shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
-an empirical sampling distribution of \(\theta\) obtained from
+an empirical sampling distribution of \(\theta\) simulated from
 permutation procedures.
@@
 empirical sampling distribution of \emph{K} obtained via permutation.
-The empirical sampling distribution of \emph{K} is first transformed via
-a Box-Cox transformation to better adhere to the assumption of
+The empirical sampling distribution of \emph{K} can be first transformed
+via a Box-Cox transformation to better adhere to the assumption of
 normality.
@@
 \includegraphics[width=8.25cm]{{fig.3}.png}
-\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.{}}
+\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+Means from simulation runs are shown for comparative ease.
+Individual values from each simulation run are available in Supporting Informaton.
+}
 \end{figure}
@@
 
-{[}(A) Linear measures for relative body size, and regions of the body
-used to estimate surface area to volume (SA:V) ratios. (B) Permutation
-distributions of phylogenetic signal for SA:V and \(\frac{BW}{SVL}\),
-with observed values shown as vertical bars. (C) Effect sizes (\(Z_K\))
-for SA:V and \(\frac{BW}{SVL}\), with their 95\% confidence intervals
-(CI not standardized by \(\sqrt(n)\)).{}{]}\{\{fig.4\}\}.png)
+\textbackslash begin\{figure\} \centering
+\includegraphics[width=8.25cm]{{fig.4}.png}
+
+\textbackslash caption\{ (A) Linear measures for relative body size, and
+regions of the body used to estimate surface area to volume (SA:V)
+ratios. (B) Permutation distributions of phylogenetic signal for SA:V
+and \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
+confidence intervals (CI not standardized by \(\sqrt(n)\)). \}
+
+\textbackslash end\{figure\}
 
@@
 to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
-to have 20\% collapsed nodes \sout{, per the technique of Molina-Venegas
-and Rodríguez (21),} and found results were consistent (Supporting
+to have 20\% collapsed nodes, per the technique of Molina-Venegas and
+Rodríguez (21), and found results were consistent (Supporting
 Information). This confirms that any tendency of incompletely resolved

### 2020-08-25 Manuscript/SupportingInfo-MLC.Rmd
SI MLC edits

--- 2020-04-20 Manuscript/SupportingInfo.Rmd
@@
 
-In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees.
-As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units).
-For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution.
-Using these, we estimated the degree of phylogenetic signal using $\lambda$.
-\hfill\break
+Here we provide additional supporting information referenced in the main document, which include additional analyses, and simulation results across a wider set of input conditions.
+As before, simulations were conducted on six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+We generated 50 random trees for each intersection of tree size and $\lambda$.
+For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
+For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and $K$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
 
-*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).
+## Additional Results on Pure-Birth Phylogenies
 
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
+Results from the simulations on pure-birth trees revealed that both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S4).
+Additionally, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+This latter observation is seen in the upper four panels of Fig.
+S1, where the dispersion of simulation results (gray dots) is far more even at each input level of phylogenetic signal for $Z_K$ than it is for $Z_{\lambda}$.
 
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S1.png")  
 ```
 
-\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break
-
-```{r, echo = FALSE, out.width="95%"}
-include_graphics("Fig1.png")
-
-```
-
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break
-
-## Literature Survey
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pure-birth trees.
 
@@
 
-# References
+## Simulations on Pectinate Phylogenies
 
- \setlength{\parindent}{-0.25in}
- \setlength{\leftskip}{0.25in}
- \setlength{\parskip}{8pt}
- \noindent
+Results from simulations on pectinate phylogenies largely mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S2: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S2 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S2 blue line).
+\hfill\break
 
-<div id="refs"></div>
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S3 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S3 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S3 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S4).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S2.png")  
+```
+
+\textbf{Figure S2}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S3.png")  
+```
+
+\textbf{Figure S3}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pectinate trees.
+At each input level, the dark black line represents the empirically derived expected value (mean) of
+\textit{K}, the red line is the standard deviation of
+\textit{K}, and the blue line is Shapiro Wilks statistic of
+\textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S4.png")  
+```
+
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pectinate trees.
+
+\newpage
+
+## Simulations on Balanced Phylogenies
+
+Results from simulations on balanced phylogenies also mirrored those found on pure-birth trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S5: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S5 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S5 blue line).
+\hfill\break
+
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S6 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S6 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S6 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size (fig.
+S7).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, though there was a slight effect of tree size with balanced phylogenies.
+But as before, its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S5.png")  
+```
+
+\textbf{Figure S5}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S6.png")  
+```
+
+\textbf{Figure S6}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on balanced trees.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S7.png")  
+```
+
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for balanced trees.
+
+\newpage
+
+## Simulations on Phylogenies Containing Polytomies
+
+We also investigaged the effect of unresolved phylogenies on estimates of phylogenetic signal by adjusting pure-birth trees to have 20% collapsed nodes, following the procedures of [@MolinaVenegas2017].
+Results from simulations on phylogenies containing polytomies mirrored those found on fully resolved trees.
+For $\hat{\lambda}$, the mean value increased with increasing input signal, but was negatively biased, and was less than the input value across most of its range (Fig.
+S8: black line).
+Additionally, the precision of $\hat{\lambda}$ varied with differing input levels, with the greatest variation found at intermediate values of $\lambda$ (Fig.
+S8 red line).
+Finally, the distribution of $\hat{\lambda}$ was not normal, and became more skewed at more extreme values of $\lambda$ (Fig.
+S8 blue line).
+\hfill\break
+
+For *K*, mean values increased with increasing phylogenetic signal, though as was found with pure-birth trees, the increase was nonlinear (Fig.
+S9 black line).
+Likewise, variation increased with increasing phylogenetic signal (Fig.
+S9 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$ (Fig.
+S9 blue line).
+\hfill\break
+
+With respect to effect sizes, both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
+S10).
+Also, $Z_K$ increased more linearly with increasing levels of phylogenetic signal, and its standard deviation across input signal was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on $Z_K$.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S8.png")  
+```
+
+\textbf{Figure S8}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S9.png")  
+```
+
+\textbf{Figure S9}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pure-birth trees containing polytomies.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S10.png")  
+```
+
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pure-birth trees containing polytomies.
+
+\newpage
+
+## Simulations of Phylogenetic Regression and ANOVA
+
+We also investigated the effect of incorporating $\lambda$ in PGLS analyses (phylogenetic regression and ANOVA).
+Here, patterns for $\hat\lambda$, *K*, $Z_\lambda$ and $Z_K$ were virtually identical to those found previously.
+In essence, $\hat\lambda$ was not suitable as a test statistic representing phylogenetic signal due to all properties previously shown (Figs.
+S11 & S12), *K* was more appropriate, but  associated non-linearly with signal strength (Figs.
+S13 & 14), while $Z_K$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S15 & S16).
+\hfill\break
+
+With respect to model performance, input parameters ($\beta$) were well estimated in phylogenetic regression  and ANOVA when $\lambda$ was incorporated (Figs S17 & S18), implying that $\lambda$ may be used as a tuning parameter for these models.
+Additionally, type I error and statistical power were unaffected by the inclusion of $\lambda$ (Figs S19 & S20), though power was reduced at smaller sample sizes in phylogenetic ANOVA, mirroring prior findings of [@AdamsCollyer2018b].
+These results generally confirmed earlier findings of [@Revell2010] for phylogenetic regression, and extend them to the case of phylogenetic ANOVA.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S11.png")  
+```
+
+\textbf{Figure S11}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S12.png")  
+```
+
+\textbf{Figure S12}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S13.png")  
+```
+
+\textbf{Figure S13}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S14.png")  
+```
+
+\textbf{Figure S14}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S15.png")  
+```
+
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S16.png")  
+```
+
+\textbf{Figure S16}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S17.png")  
+```
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S18.png")  
+```
+
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S19.png")  
+```
+
+\textbf{Figure S19}. Power curves for phylogenetic regression (incorporating $\lambda$). Type I error is found on the far-left of the plot (effect = 0).
+
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("fig.S20.png")  
+```
+
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating $\lambda$). Type I error is found on the far-left of the plot (effect = 0).
+
+\newpage
+
+## References

### 2020-08-25 Manuscript/PNAS-Adams-04.Rmd
vsn 4

--- 2020-08-25 Manuscript/PNAS-Adams-03-MLC.Rmd
@@
 ---
-Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry among species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
-Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition trends in the data on the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @MartinsHansen1997; @OMeara_et_al2006;  @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b].
-These methods are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
-Indeed, under numerous evolutionary models, phylogenetic signal is to be expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among related taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
+Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry of species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
+Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition the data by the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @MartinsHansen1997; @OMeara_et_al2006;  @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b].
+PCMs are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
+Indeed, under numerous evolutionary models, phylogenetic signal is expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
 
@@
 
-Another widely used measure of phylogenetic signal is Blomberg's *K* [@Blomberg_et_al2003], which characterizes phylogenetic signal as the ratio of observed trait variation to the amount of variation expected under Brownian motion.
+Another widely used measure is Blomberg's *K* [@Blomberg_et_al2003], which characterizes phylogenetic signal as the ratio of observed trait variation to the amount of variation expected under Brownian motion.
 Blomberg's *K* can be treated as a test statistic by employing a permutation test to generate its sampling distribution [@Blomberg_et_al2003; @Adams2014a] for determining whether significant phylogenetic signal is present in data.
@@
 First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases.
-Second, the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values (Fig.1 red line), implying that the precision in estimating $\lambda$ varies across the range of input values.
-Third, the distributions of $\hat{\lambda}$ tend toward normal ddistributions at intermediate levels of $\lambda$ but become increasingly skewed at more extreme values of $\lambda$ (Fig.
+Second, the precision in estimating $\lambda$ varies across the range of input values, as the standard deviation of $\hat{\lambda}$ is largest at intermediate values of $\lambda$ and smallest at extreme values (Fig.1 red line).
+Third, the distributions of $\hat{\lambda}$ tend toward normal distributions at intermediate levels of $\lambda$ but become increasingly skewed at more extreme values of $\lambda$ (Fig.
 1 blue line).
@@
 2 black line).
-Additionally, the standard deviation of *K* is consistent across tree sizes (Fig.
+Second, the standard deviation of *K* is consistent across tree sizes (Fig.
 2 red line), and while it increases with $\lambda$, it is always less than the mean (low coefficient of variation).
@@
 2 blue line).
-This result differs from those of [@Munkemuller_et_al2012], where the skewing appears to be the result of combining random values generated independently, rather than being a property of *K* itself.
+This differs from results of [@Munkemuller_et_al2012], where the skewing appears to be due to combining random values generated independently, rather than being a property of *K* itself.
 Overall, these findings reveal that while *K* is more reliable as an estimate of phylogenetic signal, the non-linear scaling with input signal implies that it should not be considered an effect size that measures the strength of phylogenetic signal on a common scale for comparison across datasets.
@@
 
-![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
-(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
-(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](fig.4.png)
+\begin{figure}
+\centering
+\includegraphics[width=8.25cm]{{fig.4}.png}
+\caption{(A) Linear measures for relative body size, and regions of the
+body used to estimate surface area to volume (SA:V) ratios. (B)
+Permutation distributions of phylogenetic signal for SA:V and
+\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
+confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
+\end{figure}
 
@@
 Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using *K* have been generalized for multivariate data [$\kappa_{mult}$: see @Adams2014a].
-Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
+Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of variables, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
 

### 2020-08-25 Manuscript/PNAS-Adams-04.tex
vsn 4

--- 2020-08-25 Manuscript/PNAS-Adams-03-MLC.tex
@@
 Investigating macroevolutionary patterns of trait variation requires a
-phylogenetic perspective, because the shared ancestry among species
+phylogenetic perspective, because the shared ancestry of species
 violates the assumption of independence among trait values that is
@@
 non-independence is the purview of \emph{phylogenetic comparative
-methods} (PCMs): a suite of analytical tools that condition trends in
-the data on the phylogenetic relatedness of observations (3--10). These
-methods are predicated on the notion that phylogenetic signal -- the
-tendancy for closely related species to display similar trait values --
-is present in cross-species datasets (1, 11, 12). Indeed, under numerous
-evolutionary models, phylogenetic signal is to be expected, as
-stochastic character change along the hierarchical structure of the tree
-of life generates trait covaration among related taxa (1, 12, 13).
+methods} (PCMs): a suite of analytical tools that condition the data by
+the phylogenetic relatedness of observations (3--10). PCMs are
+predicated on the notion that phylogenetic signal -- the tendancy for
+closely related species to display similar trait values -- is present in
+cross-species datasets (1, 11, 12). Indeed, under numerous evolutionary
+models, phylogenetic signal is expected, as stochastic character change
+along the hierarchical structure of the tree of life generates trait
+covaration among taxa (1, 12, 13).
 
@@
 
-Another widely used measure of phylogenetic signal is Blomberg's
-\emph{K} (12), which characterizes phylogenetic signal as the ratio of
-observed trait variation to the amount of variation expected under
-Brownian motion. Blomberg's \emph{K} can be treated as a test statistic
-by employing a permutation test to generate its sampling distribution
-(12, 16) for determining whether significant phylogenetic signal is
-present in data. Both \(\lambda\) and \emph{K} seem intuitive to
-interpret, as a value of \(0\) for both corresponds to no phylogenetic
-signal, while a value of \(1\) corresponds to the amount of phylogenetic
-signal expected under Brownian motion. Thus, it is tempting to regard
-both \(\lambda\) and \emph{K} as descriptive statistics that measure the
-relative strength of phylogenetic signal, providing an estimate of its
-magnitude for comparison.
+Another widely used measure is Blomberg's \emph{K} (12), which
+characterizes phylogenetic signal as the ratio of observed trait
+variation to the amount of variation expected under Brownian motion.
+Blomberg's \emph{K} can be treated as a test statistic by employing a
+permutation test to generate its sampling distribution (12, 16) for
+determining whether significant phylogenetic signal is present in data.
+Both \(\lambda\) and \emph{K} seem intuitive to interpret, as a value of
+\(0\) for both corresponds to no phylogenetic signal, while a value of
+\(1\) corresponds to the amount of phylogenetic signal expected under
+Brownian motion. Thus, it is tempting to regard both \(\lambda\) and
+\emph{K} as descriptive statistics that measure the relative strength of
+phylogenetic signal, providing an estimate of its magnitude for
+comparison.
 
@@
 First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
-increases. Second, the standard deviation of \(\hat{\lambda}\) is
-largest at intermediate values of \(\lambda\) and smallest at extreme
-values (Fig.1 red line), implying that the precision in estimating
-\(\lambda\) varies across the range of input values. Third, the
-distributions of \(\hat{\lambda}\) tend toward normal ddistributions at
-intermediate levels of \(\lambda\) but become increasingly skewed at
-more extreme values of \(\lambda\) (Fig. 1 blue line). For small tree
-sizes, it is also clear that distributions are more platykurtic at
-intermediate values of \(\hat{\lambda}\). However, the mean value of
+increases. Second, the precision in estimating \(\lambda\) varies across
+the range of input values, as the standard deviation of
+\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
+smallest at extreme values (Fig.1 red line). Third, the distributions of
+\(\hat{\lambda}\) tend toward normal distributions at intermediate
+levels of \(\lambda\) but become increasingly skewed at more extreme
+values of \(\lambda\) (Fig. 1 blue line). For small tree sizes, it is
+also clear that distributions are more platykurtic at intermediate
+values of \(\hat{\lambda}\). However, the mean value of
 \(\hat{\lambda}\) is negatively-biased (particularly for small tree
@@
 \includegraphics[width=8.25cm]{{fig.2}.png}
-\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.
-~~Gray line signifies the 1:1 line where the input value matches the estimate \(\hat K\).~~ At each input level, the dark black line represents the empirically derived expected value (mean) of
+\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.  At each input level, the black line represents the empirically derived expected value (mean) of
 *K*, the red line is the standard deviation of
@@
 increase with increasing signal (\(\lambda\)) irrespective of tree size,
-albeit non-linearly (Fig. 2 black line). Additionally, the standard
-deviation of \emph{K} is consistent across tree sizes (Fig. 2 red line),
-and while it increases with \(\lambda\), it is always less than the mean
-(low coefficient of variation). This finding is perhaps unsurprising, as
+albeit non-linearly (Fig. 2 black line). Second, the standard deviation
+of \emph{K} is consistent across tree sizes (Fig. 2 red line), and while
+it increases with \(\lambda\), it is always less than the mean (low
+coefficient of variation). This finding is perhaps unsurprising, as
 \emph{K} is lower-bounded by 0, and is never large for small values of
@@
 range of input \(\lambda\); a consistent pattern regardless of tree size
-(Fig. 2 blue line). This result differs from those of (18), where the
-skewing appears to be the result of combining random values generated
-independently, rather than being a property of \emph{K} itself. Overall,
-these findings reveal that while \emph{K} is more reliable as an
-estimate of phylogenetic signal, the non-linear scaling with input
-signal implies that it should not be considered an effect size that
-measures the strength of phylogenetic signal on a common scale for
-comparison across datasets.
+(Fig. 2 blue line). This differs from results of (18), where the skewing
+appears to be due to combining random values generated independently,
+rather than being a property of \emph{K} itself. Overall, these findings
+reveal that while \emph{K} is more reliable as an estimate of
+phylogenetic signal, the non-linear scaling with input signal implies
+that it should not be considered an effect size that measures the
+strength of phylogenetic signal on a common scale for comparison across
+datasets.
 
@@
 
-\textbackslash begin\{figure\} \centering
+\begin{figure}
+\centering
 \includegraphics[width=8.25cm]{{fig.4}.png}
-
-\textbackslash caption\{ (A) Linear measures for relative body size, and
-regions of the body used to estimate surface area to volume (SA:V)
-ratios. (B) Permutation distributions of phylogenetic signal for SA:V
-and \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
+\caption{(A) Linear measures for relative body size, and regions of the
+body used to estimate surface area to volume (SA:V) ratios. (B)
+Permutation distributions of phylogenetic signal for SA:V and
+\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
 Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
-confidence intervals (CI not standardized by \(\sqrt(n)\)). \}
-
-\textbackslash end\{figure\}
+confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
+\end{figure}
 
@@
 be utilized for comparing the strength of phylogenetic signal among
-datasets containing a different number of species, and even for
+datasets containing a different number of variables, and even for
 phenotypes obtained from species in different lineages, because their

### 2020-08-25 Manuscript/SupportingInfo.tex
vsn 4

@@
 and with differing levels of phylogenetic signal
-(\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100 random trees for
+(\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 50 random trees for
 each intersection of tree size and \(\lambda\). For each \(\lambda\)
 within each tree size, continuous traits were then simulated on each
-phylogeny under a BM model of evolution. For each set of 100 trees we
-measured the mean values of \(\hat{\lambda}\) and \(\kappa\), their
-standard deviation, and calculated the Shapiro-Wilk \(W\) statistic as a
+phylogeny under a BM model of evolution. For each set of 50 trees we
+measured the mean values of \(\hat{\lambda}\) and \(K\), their standard
+deviation, and calculated the Shapiro-Wilk \(W\) statistic as a
 departure from normality (symmetry). For the latter, a value of \(1.0\)
@@
 signal, but \(\hat{\lambda}\) was strongly affected by tree size,
-whereas \(Z_K\) was m1re consistent (Fig. S4). Additionally, \(Z_K\)
+whereas \(Z_K\) was more consistent (Fig. S4). Additionally, \(Z_K\)
 increased more linearly with increasing levels of phylogenetic signal,
@@
 \textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
-to increasing strength of Brownian motion.
+to increasing strength of Brownian motion for pure-birth trees.
 
@@
 \emph{K} was more normally distributed throughout its range, and across
-different tree sizes, as compared with \(\hat{\lambda}\), thought there
+different tree sizes, as compared with \(\hat{\lambda}\), though there
 was some slight skew for high input levels of phylogenetic signal on
@@
 strength of Brownian motion on pectinate trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+1:1 line where the input value matches the estimate.At each input level,
+the dark black line represents the empirically derived expected value
+(mean) of \(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
@@
 \textbf{Figure S3}. Response of Blomberg's \textit{K} to increasing
-strength of Brownian motion on pectinate trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+strength of Brownian motion on pectinate trees. At each input level, the
+dark black line represents the empirically derived expected value (mean)
+of \textit{K}, the red line is the standard deviation of \textit{K}, and
+the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\)
+signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 \textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
-to increasing strength of Brownian motion.
+to increasing strength of Brownian motion for pectinate trees.
 
@@
 strength of Brownian motion on balanced trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+1:1 line where the input value matches the estimate. At each input
+level, the dark black line represents the empirically derived expected
+value (mean) of \(\hat\lambda\), the red line is the standard deviation
+of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
@@
 \textbf{Figure S6}. Response of Blomberg's \textit{K} to increasing
-strength of Brownian motion on balanced trees. Gray line signifies the
-1:1 line where the input value matches the estimate.
+strength of Brownian motion on balanced trees. At each input level, the
+dark black line represents the empirically derived expected value (mean)
+of \textit{K}, the red line is the standard deviation of \textit{K}, and
+the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\)
+signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 \textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
-to increasing strength of Brownian motion.
+to increasing strength of Brownian motion for balanced trees.
 
@@
 We also investigaged the effect of unresolved phylogenies on estimates
-of phylogenetic signal by adjusting pure-birth trees to have 50\%
+of phylogenetic signal by adjusting pure-birth trees to have 20\%
 collapsed nodes, following the procedures of (1). Results from
@@
 Gray line signifies the 1:1 line where the input value matches the
-estimate.
+estimate. At each input level, the dark black line represents the
+empirically derived expected value (mean) of \(\hat\lambda\), the red
+line is the standard deviation of \(\hat\lambda\), and the blue line is
+Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies
+normality, \(W< 1.0\) represent skewed distributions).
 
@@
 strength of Brownian motion on pure-birth trees containing polytomies.
-Gray line signifies the 1:1 line where the input value matches the
-estimate.
+At each input level, the dark black line represents the empirically
+derived expected value (mean) of \textit{K}, the red line is the
+standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
 
@@
 \textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and
-\(Z_K\) to increasing strength of Brownian motion.
+\(Z_K\) to increasing strength of Brownian motion for pure-birth trees
+containing polytomies.
 
@@
 analyses (phylogenetic regression and ANOVA). Here, patterns for
-\(\hat\lambda\), \emph{K}, \(Z_\lambda\) and \(Z_\kappa\) were virtually
+\(\hat\lambda\), \emph{K}, \(Z_\lambda\) and \(Z_K\) were virtually
 identical to those found previously. In essence, \(\hat\lambda\) was not
@@
 appropriate, but associated non-linearly with signal strength (Figs. S13
-\& 14), while \(Z_\kappa\) was found to be a superior effect size as
-compared with \(Z_\lambda\) (Figs S15 \& S16). \hfill\break
+\& 14), while \(Z_K\) was found to be a superior effect size as compared
+with \(Z_\lambda\) (Figs S15 \& S16). \hfill\break
 
@@
 \(\lambda\)). Gray line signifies the 1:1 line where the input value
-matches the estimate.
+matches the estimate. At each input level, the dark black line
+represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
@@
 strength of Brownian motion in phylogenetic regression (incorporating
-\(\lambda\)). Gray line signifies the 1:1 line where the input value
-matches the estimate.
+\(\lambda\)). At each input level, the dark black line represents the
+empirically derived expected value (mean) of \textit{K}, the red line is
+the standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
 
@@
 strength of Brownian motion in phylogenetic ANOVA (incorporating
-\(\lambda\)). Gray line signifies the 1:1 line where the input value
-matches the estimate.
+\(\lambda\)). At each input level, the dark black line represents the
+empirically derived expected value (mean) of \textit{K}, the red line is
+the standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
 
@@
 \textbf{Figure S17}. Parameter estimates from phylogenetic regression
-(incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
-the estimate.
+(incorporating \(\lambda\)). The red line represents the estimated
+slope, which almost completely overlaps the gray 1:1 line.
 
@@
 \textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA
-(incorporating \(\lambda\)). Gray line is the 1:1 line and red line is
-the estimate.
+(incorporating \(\lambda\)). The red line represents the estimated
+slope, which almost completely overlaps the gray 1:1 line.
 
@@
 (incorporating \(\lambda\)). Type I error is found on the far-left of
-the plot.
+the plot (effect = 0).
 
@@
 \textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating
-\(\lambda\)). Type I error is found on the far-left of the plot.
+\(\lambda\)). Type I error is found on the far-left of the plot (effect
+= 0).
 

### 2020-08-25 Manuscript/SupportingInfo.Rmd
SI

@@
 ---
-title: "Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal"
+title:
 
-author: "Dean C. Adams, Erica K. Baken, and Michael L. Collyer"
+author:
 
@@
 
+```{r, echo = FALSE, out.width="95%", align= "center"}
+include_graphics("SI-Title.png")  
+```
+\hfill\break
+\hfill\break
+
+## Supplementary Information for 
+\hfill\break
+
+A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal \hfill\break
+\hfill\break
+Dean C. Adams, Erica K. Baken, and Michael L. Collyer \hfill\break
+
+\hfill\break
+Dean C. Adams\hfill\break
+Email:  dcadams@iastate.edu\hfill\break
+
+## This PDF file includes:
+\hfill\break
+- Supplementary text 
+- Figures S1 to S20 
+- SI References
+
+\newpage
+
 Here we provide additional supporting information referenced in the main document, which include additional analyses, and simulation results across a wider set of input conditions.
 As before, simulations were conducted on six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-We generated 100 random trees for each intersection of tree size and $\lambda$.
+We generated 50 random trees for each intersection of tree size and $\lambda$.
 For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
-For each set of 100 trees we measured the mean values of $\hat{\lambda}$ and $\kappa$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and $K$, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
 For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
@@
 
-Results from the simulations on pure-birth trees revealed that both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was m1re consistent (Fig.
+Results from the simulations on pure-birth trees revealed that both $Z_{\lambda}$ and $Z_K$ increased with increasing input phylogenetic signal, but $\hat{\lambda}$ was strongly affected by tree size, whereas $Z_K$ was more consistent (Fig.
 S4).
@@
 
-\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pure-birth trees.
 
@@
 Likewise, variation increased with increasing phylogenetic signal (Fig.
-S3 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, thought there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
+S3 red line), though the distribution of *K* was more normally distributed throughout its range, and across different tree sizes, as compared with $\hat{\lambda}$, though there was some slight skew for high input levels of phylogenetic signal on large phylogenies (Fig.
 S3 blue line).
@@
 
-\textbf{Figure S2}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S2}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on pectinate trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.At each input level, the dark
+black line represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
@@
 
-\textbf{Figure S3}. Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pectinate trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S3}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pectinate trees.
+At each input level, the dark black line represents the empirically derived expected value (mean) of
+\textit{K}, the red line is the standard deviation of
+\textit{K}, and the blue line is Shapiro Wilks statistic of
+\textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
@@
 
-\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pectinate trees.
 
@@
 
-\textbf{Figure S5}. Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S5}.
+Response of Pagel's \(\lambda\) to increasing strength of Brownian motion on balanced trees.
+Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 
-\textbf{Figure S6}. Response of Blomberg's \textit{K} to increasing strength of Brownian motion on balanced trees. Gray line signifies the 1:1 line where the input value matches the estimate.
+\textbf{Figure S6}.
+Response of Blomberg's \textit{K} to increasing strength of Brownian motion on balanced trees.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 
-\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for balanced trees.
 
@@
 
-We also investigaged the effect of unresolved phylogenies on estimates of phylogenetic signal by adjusting pure-birth trees to have 50% collapsed nodes, following the procedures of [@MolinaVenegas2017].
+We also investigaged the effect of unresolved phylogenies on estimates of phylogenetic signal by adjusting pure-birth trees to have 20% collapsed nodes, following the procedures of [@MolinaVenegas2017].
 Results from simulations on phylogenies containing polytomies mirrored those found on fully resolved trees.
@@
 Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \(\hat\lambda\), the red line is the standard deviation of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 Response of Blomberg's \textit{K} to increasing strength of Brownian motion on pure-birth trees containing polytomies.
-Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 
-\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion for pure-birth trees containing polytomies.
 
@@
 We also investigated the effect of incorporating $\lambda$ in PGLS analyses (phylogenetic regression and ANOVA).
-Here, patterns for $\hat\lambda$, *K*, $Z_\lambda$ and $Z_\kappa$ were virtually identical to those found previously.
+Here, patterns for $\hat\lambda$, *K*, $Z_\lambda$ and $Z_K$ were virtually identical to those found previously.
 In essence, $\hat\lambda$ was not suitable as a test statistic representing phylogenetic signal due to all properties previously shown (Figs.
 S11 & S12), *K* was more appropriate, but  associated non-linearly with signal strength (Figs.
-S13 & 14), while $Z_\kappa$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S15 & S16).
+S13 & 14), while $Z_K$ was found to be a superior effect size as compared with $Z_\lambda$ (Figs S15 & S16).
 \hfill\break
@@
 Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic regression (incorporating $\lambda$).
-Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 Response of Blomberg's \textit{K} to increasing strength of Brownian motion in phylogenetic ANOVA (incorporating $\lambda$).
-Gray line signifies the 1:1 line where the input value matches the estimate.
+At each input level, the dark black line represents the empirically derived expected value (mean) of \textit{K}, the red line is the standard deviation of \textit{K}, and the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\) represent skewed distributions).
 
@@
 
-\textbf{Figure S17}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
 
@@
 
-\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). Gray line is the 1:1 line and red line is the estimate.
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA (incorporating $\lambda$). The red line represents the estimated slope, which almost completely overlaps the gray 1:1 line.
 
@@
 
-\textbf{Figure S19}. Power curves for phylogenetic regression (incorporating $\lambda$). Type I error is found on the far-left of the plot.
+\textbf{Figure S19}. Power curves for phylogenetic regression (incorporating $\lambda$). Type I error is found on the far-left of the plot (effect = 0).
 
@@
 
-\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating $\lambda$). Type I error is found on the far-left of the plot.
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating $\lambda$). Type I error is found on the far-left of the plot (effect = 0).
 

### 2020-08-26 2020-PNAS-CoverLetter/2020-CoverLetter.Rmd

---
output:
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: false

---

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo = FALSE, fig.align='center', out.width="100%"}
include_graphics("CovLettTop.png")  
```

`r format(Sys.time(), '%d %B, %Y')`
\hfill\break

Prof. May Berenbaum, Editor in Chief\hfill\break
*Proceedings of the National Academy of Sciences*

\hfill\break

Prof. Berenbaum,

Please find attached our manuscript “A standardized effect size for evaluating and comparing the strength of phylogenetic signal” which we are submitting for consideration for publication in *PNAS*. This manuscript is not being considered elsewhere for publication, and both authors are aware of this submission. We are submitting this for consideration in the **Biological Sciences** in the **Evolution** minor category. \hfill\break

Why appropriate for *PNAS*: Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking. Here we develop a standardized effect size, $Z_K$, which measures the strength of phylogenetic signal on a common statistical scale. We also provide a test statistic, $\hat{Z}_{12}$, for comparing the strength of phylogenetic signal across datasets. We find that two commonly used parameters (Pagel's $\lambda$ and Blomberg's *K*), not converted to effect sizes, are unsuitable for this purpose. Our effect-size procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.  \hfill\break

**Editorial Board Members:** As potential Editorial Board Members who could adjudicate our work, we suggest: Dr. Scott Edwards, Dr. Douglas Futuyma, and Dr. Hopi Hoekstra. \hfill\break

**NAS Members:**  NAS Members who could evaluate this work include: Dr. Joseph Felsenstein, Dr. David Hillis, and Dr. Sarah Otto.\hfill\break

**Qualified Reviewers:** Additional qualified reviewers include: Dr. Simon Blomberg, Dr. Anthony Ives, and Dr. Mark Pagel. \hfill\break

Sincerely,

```{r, echo = FALSE, fig.align='left', out.width="30%"}
include_graphics("DCA-DigitalSignature.png")  
```

Dean C. Adams, Professor\hfill\break
Department of Ecology, Evolution, and Organismal Biology\hfill\break
Iowa State University

### 2020-08-26 Manuscript/PNAS-Adams-Submitted.Rmd

--- 2020-08-25 Manuscript/PNAS-Adams-04.Rmd
@@
 Accounting for this evolutionary non-independence is the purview of *phylogenetic comparative methods* (PCMs): a suite of analytical tools that condition the data by the phylogenetic relatedness of observations [@Grafen1989; @GarlandIves2000; @Rohlf2001; @MartinsHansen1997; @OMeara_et_al2006;  @Beaulieu_et_al2012; @Adams2014b; @AdamsCollyer2018b].
-PCMs are predicated on the notion that phylogenetic signal -- the tendancy for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
-Indeed, under numerous evolutionary models, phylogenetic signal is expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covaration among taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
+PCMs are predicated on the notion that phylogenetic signal -- the tendency for closely related species to display similar trait values -- is present in cross-species datasets [@Felsenstein1985; @Pagel1999; @Blomberg_et_al2003].
+Indeed, under numerous evolutionary models, phylogenetic signal is expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covariation among taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
 
@@
 For instance, as a proportional random variable bounded by $0$ and $1$, we might expect that $\hat{\lambda}$ is a random variable that follows the distribution of a Bernoulli probability parameter [@Forbes2011]; i.e., branch lengths in a tree are scaled proportionally to the probability that data arise from a BM process.
-Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to tree size; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ will be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 [@Forbes2011].
+Given a known $\lambda$ value used to generate random data on a tree, we would also expect  that the mean of an empirical sampling distribution of $\hat{\lambda}$ would approximately equal $\lambda$; the dispersion of $\hat{\lambda}$ would be largest at intermediate values of $\lambda$, $\hat{\lambda}$ would be predictable over the range of $\lambda$ with respect to tree size; the distribution of $\hat{\lambda}$ would be symmetric at intermediate values of $\lambda$ and more skewed toward values of 0 or 1; and that the distribution of $\hat{\lambda}$ would be more platykurtic at intermediate values of $\lambda$, becoming more leptokurtic toward 0 and 1 [@Forbes2011].
 Prior work [@Munkemuller_et_al2012] seems to support some of these conjectures, based superficially on statistical moments for a given tree size (mean, variance, skewness, and kurtosis; see Fig.
@@
 
-By contrast, for Blomberg's *K*, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of *K* might be a random variable that follows a normal distribution, with values distributed symmetrically about the input value [@Forbes2011].
+By contrast, for Blomberg's *K*, which is positively unbounded, we might expect that for any $\lambda$ used to generate data, estimates of *K* might be a random variable that follows a normal distribution, with values distributed symmetrically [@Forbes2011].
 This attribute seemed less reasonable based on the simulations performed by Münkemüller et al.

### 2020-08-26 Manuscript/PNAS-Adams-Submitted.tex

--- 2020-08-25 Manuscript/PNAS-Adams-04.tex
@@
 the phylogenetic relatedness of observations (3--10). PCMs are
-predicated on the notion that phylogenetic signal -- the tendancy for
+predicated on the notion that phylogenetic signal -- the tendency for
 closely related species to display similar trait values -- is present in
@@
 along the hierarchical structure of the tree of life generates trait
-covaration among taxa (1, 12, 13).
+covariation among taxa (1, 12, 13).
 
@@
 values of \(\lambda\) and more skewed toward values of 0 or 1; and that
-the distribution of \(\hat{\lambda}\) will be more platykurtic at
+the distribution of \(\hat{\lambda}\) would be more platykurtic at
 intermediate values of \(\lambda\), becoming more leptokurtic toward 0
@@
 of \emph{K} might be a random variable that follows a normal
-distribution, with values distributed symmetrically about the input
-value (31). This attribute seemed less reasonable based on the
-simulations performed by Münkemüller et al.~(18), which suggested that
-distributions were positively skewed and that Blomberg's \emph{K} might
-not behave as a statistic that follows a normal distribution. However,
-because their simulations used a weighted combination of simulated
-phylogenetic signal strengths, strong inferences are not possible (and
-distributional attributes were not the intended result of their
-simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's \emph{K},
-evaluation of statistical moments across a range of \(\lambda\) used to
-generate data would be valuable for adjudicating the reliability of
-these statistics as effect sizes. Furthermore, the expected values of
-these statistics appear to vary with tree size (18), making comparisons
-across studies challenging. Therefore, transformation of these
-statistics into \(Z\)-scores would allow evaluation of the efficacy of
-each statistic to yield effect sizes that could be used for comparisons
-of the strength of phylogenetic signal across traits and lineages.
+distribution, with values distributed symmetrically (31). This attribute
+seemed less reasonable based on the simulations performed by Münkemüller
+et al.~(18), which suggested that distributions were positively skewed
+and that Blomberg's \emph{K} might not behave as a statistic that
+follows a normal distribution. However, because their simulations used a
+weighted combination of simulated phylogenetic signal strengths, strong
+inferences are not possible (and distributional attributes were not the
+intended result of their simulations). Thus, for both Pagel's
+\(\lambda\) or Blomberg's \emph{K}, evaluation of statistical moments
+across a range of \(\lambda\) used to generate data would be valuable
+for adjudicating the reliability of these statistics as effect sizes.
+Furthermore, the expected values of these statistics appear to vary with
+tree size (18), making comparisons across studies challenging.
+Therefore, transformation of these statistics into \(Z\)-scores would
+allow evaluation of the efficacy of each statistic to yield effect sizes
+that could be used for comparisons of the strength of phylogenetic
+signal across traits and lineages.
 

### 2020-08-26 Manuscript/SupportingInfo.tex

@@
 \hypersetup{
-  pdftitle={Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
-  pdfauthor={Dean C. Adams, Erica K. Baken, and Michael L. Collyer},
   hidelinks,
@@
 
-\title{Supporting Information for: A Standardized Effect Size for Evaluating
-and Comparing the Strength of Phylogenetic Signal}
-\author{Dean C. Adams, Erica K. Baken, and Michael L. Collyer}
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
 \begin{document}
-\maketitle
+
+\includegraphics[width=0.95\linewidth]{SI-Title} \hfill\break
+\hfill\break
+
+\hypertarget{supplementary-information-for}{%
+\subsection{Supplementary Information
+for}\label{supplementary-information-for}}
+
+\hfill\break
+
+A Standardized Effect Size for Evaluating and Comparing the Strength of
+Phylogenetic Signal \hfill\break \hfill\break Dean C. Adams, Erica K.
+Baken, and Michael L. Collyer \hfill\break
+
+\hfill\break
+
+Dean C. Adams\hfill\break Email:
+\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}\hfill\break
+
+\hypertarget{this-pdf-file-includes}{%
+\subsection{This PDF file includes:}\label{this-pdf-file-includes}}
+
+\hfill\break
+
+\begin{itemize}
+\tightlist
+\item
+  Supplementary text
+\item
+  Figures S1 to S20
+\item
+  SI References
+\end{itemize}
+
+\newpage
 

### 2020-08-26 Data_Analyses/updated_simulations/simulation.analyses.PB.runs.R
Type1 error initial exploration

@@ -71,7 +71,7 @@ png(filename = "Manuscript/fig.1.png", width = 800,
 par(mfrow = c(3, 2), 
     mar = c(5,5,1,2))
 for(i in 1:length(treesizes)) {
- 
+  
   R <- list("Result" = sim.sets[[i]], "file.name" = "filler")
   Rr <- R$Result
   lambda.plot(R, pch = 19,  col = gray.colors(1, start = 0.6, end = 0.6, alpha = 0.5),
@@ -193,7 +193,7 @@ for(i in 1:2) {
   
   legend("topleft", paste("N =", treesizes[zref[i]]), bty = "n", cex = 1.5)
 }
-  LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets[[j]]$lambda.in, mean))
+LZ <- sapply(1:length(sim.sets), function(j) by(sim.sets[[j]]$lambda.z, sim.sets[[j]]$lambda.in, mean))
 
 # Note! change ylim below, if needed.  Too difficult to automate!
 
@@ -215,3 +215,154 @@ for(i in 2:ncol(KZ)) points(lambdas, KZ[,i], type = "l", lwd = i)
 legend("topleft", as.character(treesizes), lwd = 1:length(treesizes))
 
 dev.off()
+
+# Z_K tree comparison
+treesizes <- 2^(5:10)
+lambdas <- seq(0,1,0.05)
+nsim <- 50
+
+
+treecomp.mat.zk <- matrix(NA, nrow = 21, ncol = 21)
+colnames(treecomp.mat.zk) <- lambdas
+row.names(treecomp.mat.zk) <- c("32vs32", "32vs64", "32vs128", "32vs256", "32vs512", "32vs1024",
+                                "64vs64", "64vs128", "64vs256", "64vs512", "64vs1024",
+                                "128vs128", "128vs256", "128vs512", "128vs1024",
+                                "256vs256", "256vs512", "256vs1024",
+                                "512vs512", "512vs1024",
+                                "1024vs1024")
+treecomp.mat.zl <- matrix(NA, nrow = 21, ncol = 21)
+colnames(treecomp.mat.zl) <- lambdas
+row.names(treecomp.mat.zl) <- c("32vs32", "32vs64", "32vs128", "32vs256", "32vs512", "32vs1024",
[…]

### 2020-08-26 Data_Analyses/updated_simulations/simulation.analyses.PB.type1err.R

@@ -1,0 +1,90 @@
+# simulation.pb.type1error
+
+library(readxl)
+library(latex2exp)
+source("Data_Analyses/updated_simulations/simulation.source.R")
+nsim <- 50 # change to 50 for final runs
+treesizes <- 2^(5:10) # change to 5:10 for final runs
+lambdas <- seq(0, 1, 0.05) # change to seq(0, 1, 0.05) for final runs
+sim.sets_32 <- read.csv("Data_Analyses/updated_simulations/results/pb.32.50.sim.0.poly.mean.type.csv")
+sim.sets_64 <- read.csv("Data_Analyses/updated_simulations/results/pb.64.50.sim.0.poly.mean.type.csv")
+sim.sets_128 <- read.csv("Data_Analyses/updated_simulations/results/pb.128.50.sim.0.poly.mean.type.csv")
+sim.sets_256 <- read.csv("Data_Analyses/updated_simulations/results/pb.256.50.sim.0.poly.mean.type.csv")
+sim.sets_512 <- read.csv("Data_Analyses/updated_simulations/results/pb.512.50.sim.0.poly.mean.type.csv")
+sim.sets_1024 <- read.csv("Data_Analyses/updated_simulations/results/pb.1024.50.sim.0.poly.mean.type.csv")
+
+sim.sets <- list(sim.sets_32, sim.sets_64, sim.sets_128, sim.sets_256, sim.sets_512, sim.sets_1024)
+names(sim.sets) <- treesizes
+
+type1error.zk <- matrix(NA, nrow = 21, ncol = 6)
+colnames(type1error.zk) <-  treesizes
+row.names(type1error.zk) <- lambdas
+
+type1error.zl <- matrix(NA, nrow = 21, ncol = 6)
+colnames(type1error.zl) <-  treesizes
+row.names(type1error.zl) <- lambdas
+
+dist.zk <- matrix(NA, nrow= 50, ncol = 50)
+dist.zl <- matrix(NA, nrow= 50, ncol = 50)
+
+for (i in 1:length(treesizes)) { # for each tree size
+  # i <- 1
+  for (k in 1:length(lambdas)){ # for each lambda input
+    # k <- 20
+    data.i <- sim.sets[[i]][which(as.character(sim.sets[[i]]$lambda.in) == as.character(lambdas[k])),] # extract the 50 sim of that tree size at that lambda input
[…]

### 2020-08-27 Manuscript/PNAS-Adams-Submitted.Rmd
z12

@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} = \frac{\lvert Z_{K_1} - Z_{K_2}\rvert}{\sqrt{2}}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} 
 \end{align}
 
-where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2.
-The right side of the equation illustrates that if $Z_K$ has already been calculated for two sampling distributions, both have unit variance for each of the $Z_K$ statistics.
+where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above.
 Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.

### 2020-08-31 Manuscript/PNAS-Adams-Submitted.Rmd
S21

@@
 where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above.
-Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution.
+Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution, or permutation.
 Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b].
+Simulations reveal that tests based on $\hat{Z}_{12}$ have appropriate type I error rates (~0.05) and reasonable model misspecification rates (~7-12%: Supporting Information).
 

### 2020-08-31 Manuscript/PNAS-Adams-Submitted.tex
z12
S21

@@
 \begin{align}
-  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} = \frac{\lvert Z_{K_1} - Z_{K_2}\rvert}{\sqrt{2}}
+  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} 
 \end{align}
@@
 where \(K_1\), \(K_2\), \(\hat\mu_{K_1}\), \(\hat\mu_{K_2}\),
-\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above for
-equation 2. The right side of the equation illustrates that if \(Z_K\)
-has already been calculated for two sampling distributions, both have
-unit variance for each of the \(Z_K\) statistics. Estimates of
-significance of \(\hat{Z}_{12}\) may be obtained from a standard normal
-distribution. Typically, \(\hat{Z}_{12}\) is considered a two-tailed
-test, however directional (one-tailed) tests may be specified should the
-empirical situation require it (36, 38).
+\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above.
+Estimates of significance of \(\hat{Z}_{12}\) may be obtained from a
+standard normal distribution, or permutation. Typically,
+\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
+(one-tailed) tests may be specified should the empirical situation
+require it (36, 38). Simulations reveal that tests based on
+\(\hat{Z}_{12}\) have appropriate type I error rates
+(\textasciitilde0.05) and reasonable model misspecification rates
+(\textasciitilde7-12\%: Supporting Information).
 

### 2020-08-31 Manuscript/SupportingInfo.Rmd
S21

@@
 
+## Statistical properties of $\hat{Z}_{12}$
+
+Here we examine the statistical properties of tests for comparing the relative strength of phylogenetic signal, using $\hat{Z}_{12}$.
+Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+For each combination of $n$ and $\lambda$, 101 traits were simulated on the phylogeny under a Brownian motion model of evolution.
+Next, the phylogenetic signal of trait 1 was compared to the signal in the remaining 100 traits in pairwise fashion, using $\hat{Z}_{12}$.
+The proportion of significant results provided an estimate of the type I error or model misspecification.
+Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$).
+Likewise, model misspecification was evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$).
+\hfill\break
+
+**Results:** Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig.
+S21).
+Levels of misspecification were slightly higher, and ranged on average between 7% (when comparing traits generated on the same phylogeny: Fig.
+S21) to ~12% (for comparisons of traits from different trees).
+Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.
+
+```{r, echo = FALSE, out.height="50%", align= "center"}
+include_graphics("fig.S21.png")  
+```
+
+\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
+
+\newpage
+
 ## References

### 2020-08-31 Manuscript/SupportingInfo.tex
S21

@@
 
+\hypertarget{statistical-properties-of-hatz_12}{%
+\subsection{\texorpdfstring{Statistical properties of
+\(\hat{Z}_{12}\)}{Statistical properties of \textbackslash hat\{Z\}\_\{12\}}}\label{statistical-properties-of-hatz_12}}
+
+Here we examine the statistical properties of tests for comparing the
+relative strength of phylogenetic signal, using \(\hat{Z}_{12}\).
+Simulations were performed using pure-birth trees of different sizes
+(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
+phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
+combination of \(n\) and \(\lambda\), 101 traits were simulated on the
+phylogeny under a Brownian motion model of evolution. Next, the
+phylogenetic signal of trait 1 was compared to the signal in the
+remaining 100 traits in pairwise fashion, using \(\hat{Z}_{12}\). The
+proportion of significant results provided an estimate of the type I
+error or model misspecification. Specifically, type I error was
+evaluated when traits were simulated using no input phylogenetic signal
+(i.e., \(\lambda = 0\)). Likewise, model misspecification was evaluated
+for data simulated with some known, but equal, level of phylogenetic
+signal (i.e., \(\lambda > 0\)). \hfill\break
+
+\textbf{Results:} Tests revealed that across simulation conditions, the
+average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
+S21). Levels of misspecification were slightly higher, and ranged on
+average between 7\% (when comparing traits generated on the same
+phylogeny: Fig. S21) to \textasciitilde12\% (for comparisons of traits
+from different trees). Both results demonstrate that comparisons of the
+strength of phylogenetic signal using \(\hat{Z}_{12}\) do not display
+high levels of false positives, and are thus appropriate for comparing
+the degree of phylogenetic signal across traits.
+
+\includegraphics[width=28.93in,height=0.5\textheight]{fig.S21}
+
+\textbf{Figure S21}. Type I error rates and model misspecification for
+\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
+(\(\lambda = 0\)).
+
+\newpage
+
 \hypertarget{references}{%

### 2020-09-01 Manuscript/SupportingInfo.Rmd
z12 updates

@@
 Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-For each combination of $n$ and $\lambda$, 101 traits were simulated on the phylogeny under a Brownian motion model of evolution.
-Next, the phylogenetic signal of trait 1 was compared to the signal in the remaining 100 traits in pairwise fashion, using $\hat{Z}_{12}$.
+For each combination of $n$ and $\lambda$, 50 traits were simulated on the phylogeny under a Brownian motion model of evolution.
+Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$.
 The proportion of significant results provided an estimate of the type I error or model misspecification.
@@
 S21).
-Levels of misspecification were slightly higher, and ranged on average between 7% (when comparing traits generated on the same phylogeny: Fig.
-S21) to ~12% (for comparisons of traits from different trees).
+Levels of misspecification were slightly higher, and ranged on average between 7% when comparing traits generated on the same phylogeny (Fig.
+S21 black line) to ~12% for comparisons of traits from different trees.
 Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.
@@
 
-\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
+\textbf{Figure S21}.
+Type I error rates and model misspecification for $\hat{Z}_{12}$.
+Type I error is found on the far-left of the plot ($\lambda = 0$).
+The solid black line represents the average type I error and false discovery rate across all levels of input signal.
 

### 2020-09-17 Manuscript/SupportingInfo.tex

@@
 phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
-combination of \(n\) and \(\lambda\), 101 traits were simulated on the
+combination of \(n\) and \(\lambda\), 50 traits were simulated on the
 phylogeny under a Brownian motion model of evolution. Next, the
-phylogenetic signal of trait 1 was compared to the signal in the
-remaining 100 traits in pairwise fashion, using \(\hat{Z}_{12}\). The
+phylogenetic signal between traits was compared in pairwise fashion for
+all \(50*49/2 = 1225\) combinations, using \(\hat{Z}_{12}\). The
 proportion of significant results provided an estimate of the type I
@@
 S21). Levels of misspecification were slightly higher, and ranged on
-average between 7\% (when comparing traits generated on the same
-phylogeny: Fig. S21) to \textasciitilde12\% (for comparisons of traits
-from different trees). Both results demonstrate that comparisons of the
-strength of phylogenetic signal using \(\hat{Z}_{12}\) do not display
-high levels of false positives, and are thus appropriate for comparing
-the degree of phylogenetic signal across traits.
+average between 7\% when comparing traits generated on the same
+phylogeny (Fig. S21 black line) to \textasciitilde12\% for comparisons
+of traits from different trees. Both results demonstrate that
+comparisons of the strength of phylogenetic signal using
+\(\hat{Z}_{12}\) do not display high levels of false positives, and are
+thus appropriate for comparing the degree of phylogenetic signal across
+traits.
 
@@
 \(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
-(\(\lambda = 0\)).
+(\(\lambda = 0\)). The solid black line represents the average type I
+error and false discovery rate across all levels of input signal.
 

### 2020-09-17 Manuscript/PNAS-Adams-Submitted.Rmd
new Z12 compare

@@
   Data deposition: Data for the empirical example may be found on DRYAD: doi:10.5061/dryad.b554m44 and doi:10.5061/dryad.59zw3r23m.
-R-scripts for simulation tests are found on Github: XXX.
+R-scripts for simulation tests are found at: https://github.com/deanadams/PhySigCompareZ.
 Computer code for implementing the two-sample comparison of effect sizes is found in geomorph: https://cran.r-project.org/web/packages/geomorph/index.html \hfill\break
@@
 Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b].
-Simulations reveal that tests based on $\hat{Z}_{12}$ have appropriate type I error rates (~0.05) and reasonable model misspecification rates (~7-12%: Supporting Information).
+Simulations reveal that tests based on $\hat{Z}_{12}$ have appropriate type I error and model misspecification rates  of approximately 5% (Supporting Information).
 

### 2020-09-17 Manuscript/PNAS-Adams-Submitted.tex
new Z12 compare

@@
 \url{doi:10.5061/dryad.b554m44} and \url{doi:10.5061/dryad.59zw3r23m}.
-R-scripts for simulation tests are found on Github: XXX. Computer code
-for implementing the two-sample comparison of effect sizes is found in
+R-scripts for simulation tests are found at:
+\url{https://github.com/deanadams/PhySigCompareZ}. Computer code for
+implementing the two-sample comparison of effect sizes is found in
 geomorph:
@@
 require it (36, 38). Simulations reveal that tests based on
-\(\hat{Z}_{12}\) have appropriate type I error rates
-(\textasciitilde0.05) and reasonable model misspecification rates
-(\textasciitilde7-12\%: Supporting Information).
+\(\hat{Z}_{12}\) have appropriate type I error and model
+misspecification rates of approximately 5\% (Supporting Information).
 

### 2020-09-17 Manuscript/SupportingInfo.Rmd
new Z12 compare

@@
 Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-For each combination of $n$ and $\lambda$, 50 traits were simulated on the phylogeny under a Brownian motion model of evolution.
+For each combination of $n$ and $\lambda$, 20 pure-birth trees were generated and 50 traits were simulated on each phylogeny under a Brownian motion model of evolution.
 Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$.
@@
 S21).
-Levels of misspecification were slightly higher, and ranged on average between 7% when comparing traits generated on the same phylogeny (Fig.
-S21 black line) to ~12% for comparisons of traits from different trees.
+Likewise, levels of misspecification were also low, and were at or below the nominal 5% level (Fig.
+S21 black line).
 Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.

### 2020-09-17 Manuscript/SupportingInfo.tex
new Z12 compare

--- 2020-08-31 Manuscript/SupportingInfo.tex
@@
 phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
-combination of \(n\) and \(\lambda\), 101 traits were simulated on the
-phylogeny under a Brownian motion model of evolution. Next, the
-phylogenetic signal of trait 1 was compared to the signal in the
-remaining 100 traits in pairwise fashion, using \(\hat{Z}_{12}\). The
-proportion of significant results provided an estimate of the type I
-error or model misspecification. Specifically, type I error was
-evaluated when traits were simulated using no input phylogenetic signal
-(i.e., \(\lambda = 0\)). Likewise, model misspecification was evaluated
-for data simulated with some known, but equal, level of phylogenetic
-signal (i.e., \(\lambda > 0\)). \hfill\break
+combination of \(n\) and \(\lambda\), 20 pure-birth trees were generated
+and 50 traits were simulated on each phylogeny under a Brownian motion
+model of evolution. Next, the phylogenetic signal between traits was
+compared in pairwise fashion for all \(50*49/2 = 1225\) combinations,
+using \(\hat{Z}_{12}\). The proportion of significant results provided
+an estimate of the type I error or model misspecification. Specifically,
+type I error was evaluated when traits were simulated using no input
+phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, model
+misspecification was evaluated for data simulated with some known, but
+equal, level of phylogenetic signal (i.e., \(\lambda > 0\)).
+\hfill\break
 
@@
 average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
-S21). Levels of misspecification were slightly higher, and ranged on
-average between 7\% (when comparing traits generated on the same
-phylogeny: Fig. S21) to \textasciitilde12\% (for comparisons of traits
-from different trees). Both results demonstrate that comparisons of the
-strength of phylogenetic signal using \(\hat{Z}_{12}\) do not display
-high levels of false positives, and are thus appropriate for comparing
-the degree of phylogenetic signal across traits.
+S21). Likewise, levels of misspecification were also low, and were at or
+below the nominal 5\% level (Fig. S21 black line). Both results
+demonstrate that comparisons of the strength of phylogenetic signal
+using \(\hat{Z}_{12}\) do not display high levels of false positives,
+and are thus appropriate for comparing the degree of phylogenetic signal
+across traits.
 
-\includegraphics[width=28.93in,height=0.5\textheight]{fig.S21}
+\includegraphics[width=6.67in,height=0.5\textheight]{fig.S21}
 

### 2020-09-17 Data_Analyses/updated_simulations/simulation.analyses.PB.type1err-DCA.R

@@ -1,0 +1,229 @@
+library(geiger) 
+library(phytools)
+library(geomorph)
+
+##DCA: adjusted BC to not use observed in calculating the transformation,
+
+treesizes <- 2^(8) # 2^(5:7) done, 2^8 being run by EKB, still need to run 2^(9:10)
+lambdas <- seq(0, 1, 0.05)
+ntrees <- 2
+ndatasets <- 50
+physig_iter <- 999
+
+#### functions from develop branch ######
+center <- function(x){
+  if(is.vector(x)) x - mean(x) else {
+    x <- as.matrix(x)
+    dims <- dim(x)
+    fast.center(x, dims[1], dims[2])
+  }
+}
+#Adjust box.cox to use distribution minus observed, then transform observed
+box.cox <- function(y, eps = 0.02) {
+  y.obs <- y[1];   y <- y[-1] #pull out observed
+  if(any(y <= 0)) y = y - min(y) + 0.0001
+  n <- length(y)
+  yy <- y/exp(mean(log(y)))
+  logy <- log(yy)
+  lambda <- seq(-2.4, 2.4, 0.4)
+  m <- length(lambda)
+  
+  loglik <- sapply(1:m, function(j){ # same as MASS::boxcox loglik f
+    la <- lambda[j]
+    yt <- if(abs(la) > eps) yt <- (yy^la - 1)/la else
+      logy * (1 + (la * logy)/2 * (1 + (la * logy)/3 * (1 + (la * logy)/4)))
+    
+    -n/2 * log(sum(center(yt)^2))
+  })
+  
+  lambda.opt <- lambda[which.max(loglik)][[1]]
+  
+  if(abs(lambda.opt) == 2.4) {
+    lambda <- seq(-3.2, 3.2, 0.2)
+    m <- length(lambda)
+    
+    loglik <- sapply(1:m, function(j){ # same as MASS::boxcox loglik f
+      la <- lambda[j]
+      yt <- if(abs(la) > eps) yt <- (yy^la - 1)/la else
+        logy * (1 + (la * logy)/2 * (1 + (la * logy)/3 * (1 + (la * logy)/4)))
+      
+      -n/2 * log(sum(center(yt)^2))
+    })
+    lambda.opt <- lambda[which.max(loglik)][[1]]
+  }
+  
+  sp <- spline(lambda, loglik, n = 100)
+  lambda.opt <- sp$x[which.max(sp$y)]
[…]

### 2020-09-17 Manuscript/SupportingInfo-MLC.tex

@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  pdftitle={Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
+  pdfauthor={Dean C. Adams, Erica K. Baken, and Michael L. Collyer},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\author{}
-\date{\vspace{-2.5em}}
+\title{Supporting Information for: A Standardized Effect Size for Evaluating
+and Comparing the Strength of Phylogenetic Signal}
+\author{Dean C. Adams, Erica K. Baken, and Michael L. Collyer}
+\date{}
 
 \begin{document}
-
-\hypertarget{supporting-information-for-a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}{%
-\section{Supporting Information for: A Standardized Effect Size for
-Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not
-Appropriate}\label{supporting-information-for-a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}}
+\maketitle
 
 Here we provide additional supporting information referenced in the main
-document: additional analyses, and simulation results across a wider set
-of input conditions.
+document, which include additional analyses, and simulation results
+across a wider set of input conditions. As before, simulations were
+conducted on six different tree sizes (\(n=2^5, 2^6, \cdots, 2^{10}\)),
+and with differing levels of phylogenetic signal
+(\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 50 random trees for
+each intersection of tree size and \(\lambda\). For each \(\lambda\)
+within each tree size, continuous traits were then simulated on each
+phylogeny under a BM model of evolution. For each set of 50 trees we
+measured the mean values of \(\hat{\lambda}\) and \(K\), their standard
+deviation, and calculated the Shapiro-Wilk \(W\) statistic as a
+departure from normality (symmetry). For the latter, a value of \(1.0\)
+indicates normally distributed values, while departures from \(1.0\)
+indicate skewness.
 
-\hypertarget{simulations-on-differently-shaped-phylogenies}{%
-\subsection{Simulations on differently shaped
-phylogenies}\label{simulations-on-differently-shaped-phylogenies}}
+\hypertarget{additional-results-on-pure-birth-phylogenies}{%
+\subsection{Additional Results on Pure-Birth
+Phylogenies}\label{additional-results-on-pure-birth-phylogenies}}
 
-In addition to using pure-birth phylogenies, we explored the effect of
-tree shape on our findings using both balanced and pectinate trees. As
-before, simulations were conducted on a range of tree sizes
-(\(n=2^5 - 2^{10}\)), and across a range of input levels of phylogenetic
-signal (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05 units).
-For each \(n\) and \(\lambda_{in}\) combination, 50 replicates of a
-continuous trait were simulated using a Brownian motion model of
-evolution. Using these, we estimated the degree of phylogenetic signal
-using \(\lambda\). \hfill\break
+Results from the simulations on pure-birth trees revealed that both
+\(Z_{\lambda}\) and \(Z_K\) increased with increasing input phylogenetic
+signal, but \(\hat{\lambda}\) was strongly affected by tree size,
+whereas \(Z_K\) was more consistent (Fig. S4). Additionally, \(Z_K\)
+increased more linearly with increasing levels of phylogenetic signal,
+and its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision. This latter observation is
+seen in the upper four panels of Fig. S1, where the dispersion of
+simulation results (gray dots) is far more even at each input level of
+phylogenetic signal for \(Z_K\) than it is for \(Z_{\lambda}\).
 
-\emph{Results}. As found with pure-birth phylogenies, estimates of
-\(\lambda\) varied most dramatically in simulations with fewer species
-and at intermediate values of lambda (Fig. S1, S2). Pectinate trees
-showed an interesting tendency to underestimate \(\lambda\) across input
-values. Most dramatically, some simulations on the largest pectinate
-phylogeny (\(n = 1024\)) estimated \(\lambda = 0\) for input values as
-high as 0.5 (Fig. S2).
+\includegraphics[width=0.95\linewidth]{fig.S1}
 
-\includegraphics[width=0.95\linewidth]{FigS1}
+\textbf{Figure S1}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion for pure-birth trees. All
+results are shown for smallest and largest trees in the top four panels;
+only means of results are shown in the bottom two panels, to facilitate
+comparison with respect to tree size.
 
-\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. Results obtained using balanced
-trees. \hfill\break
+\newpage
 
-\includegraphics[width=0.95\linewidth]{FigS2}
+\hypertarget{simulations-on-pectinate-phylogenies}{%
+\subsection{Simulations on Pectinate
+Phylogenies}\label{simulations-on-pectinate-phylogenies}}
 
-\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known
-lambda inputs on various tree sizes. Results obtained using pectinate
-trees.
+Results from simulations on pectinate phylogenies largely mirrored those
+found on pure-birth trees. For \(\hat{\lambda}\), the mean value
+increased with increasing input signal, but was negatively biased, and
+was less than the input value across most of its range (Fig. S2: black
+line). Additionally, the precision of \(\hat{\lambda}\) varied with
+differing input levels, with the greatest variation found at
+intermediate values of \(\lambda\) (Fig. S2 red line). Finally, the
+distribution of \(\hat{\lambda}\) was not normal, and became more skewed
+at more extreme values of \(\lambda\) (Fig. S2 blue line). \hfill\break
+
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S3 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S3 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\), though there
+was some slight skew for high input levels of phylogenetic signal on
+large phylogenies (Fig. S3 blue line). \hfill\break
+
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size, whereas \(Z_K\) was more consistent
+(Fig. S4). Also, \(Z_K\) increased more linearly with increasing levels
+of phylogenetic signal, and its standard deviation across input signal
+was more even across tree sizes, implying more consistent precision.
+
+\includegraphics[width=0.95\linewidth]{fig.S2}
+
+\textbf{Figure S2}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on pectinate trees. Gray line signifies the
+1:1 line where the input value matches the estimate.At each input level,
+the dark black line represents the empirically derived expected value
+(mean) of \(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S3}
+
+\textbf{Figure S3}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on pectinate trees. At each input level, the
+dark black line represents the empirically derived expected value (mean)
+of \textit{K}, the red line is the standard deviation of \textit{K}, and
+the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\)
+signifies normality, \(W< 1.0\) represent skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S4}
+
+\textbf{Figure S4}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion for pectinate trees. All
+results are shown for smallest and largest trees in the top four panels;
+only means of results are shown in the bottom two panels, to facilitate
+comparison with respect to tree size.
+
+\newpage
+
+\hypertarget{simulations-on-balanced-phylogenies}{%
+\subsection{Simulations on Balanced
+Phylogenies}\label{simulations-on-balanced-phylogenies}}
+
+Results from simulations on balanced phylogenies also mirrored those
+found on pure-birth trees. For \(\hat{\lambda}\), the mean value
+increased with increasing input signal, but was negatively biased, and
+was less than the input value across most of its range (Fig. S5: black
+line). Additionally, the precision of \(\hat{\lambda}\) varied with
+differing input levels, with the greatest variation found at
+intermediate values of \(\lambda\) (Fig. S5 red line). Finally, the
+distribution of \(\hat{\lambda}\) was not normal, and became more skewed
+at more extreme values of \(\lambda\) (Fig. S5 blue line). \hfill\break
+
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S6 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S6 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\), though there
+was some slight skew for high input levels of phylogenetic signal on
+large phylogenies (Fig. S6 blue line). \hfill\break
+
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size (fig.~S7). Also, \(Z_K\) increased more
+linearly with increasing levels of phylogenetic signal, though there was
+a slight effect of tree size with balanced phylogenies. But as before,
+its standard deviation across input signal was more even across tree
+sizes, implying more consistent precision.
+
+\includegraphics[width=0.95\linewidth]{fig.S5}
+
+\textbf{Figure S5}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on balanced trees. Gray line signifies the
+1:1 line where the input value matches the estimate. At each input
+level, the dark black line represents the empirically derived expected
+value (mean) of \(\hat\lambda\), the red line is the standard deviation
+of \(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S6}
+
+\textbf{Figure S6}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on balanced trees. At each input level, the
+dark black line represents the empirically derived expected value (mean)
+of \textit{K}, the red line is the standard deviation of \textit{K}, and
+the blue line is Shapiro Wilks statistic of \textit{K} (\(W=1.0\)
+signifies normality, \(W< 1.0\) represent skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S7}
+
+\textbf{Figure S7}. Response of effect sizes \(Z_{\lambda}\) and \(Z_K\)
+to increasing strength of Brownian motion for balanced trees. All
+results are shown for smallest and largest trees in the top four panels;
+only means of results are shown in the bottom two panels, to facilitate
+comparison with respect to tree size.
+
+\newpage
+
+\hypertarget{simulations-on-phylogenies-containing-polytomies}{%
+\subsection{Simulations on Phylogenies Containing
+Polytomies}\label{simulations-on-phylogenies-containing-polytomies}}
+
+We also investigaged the effect of unresolved phylogenies on estimates
+of phylogenetic signal by adjusting pure-birth trees to have 20\%
+collapsed nodes, following the procedures of (1). Results from
+simulations on phylogenies containing polytomies mirrored those found on
+fully resolved trees. For \(\hat{\lambda}\), the mean value increased
+with increasing input signal, but was negatively biased, and was less
+than the input value across most of its range (Fig. S8: black line).
+Additionally, the precision of \(\hat{\lambda}\) varied with differing
+input levels, with the greatest variation found at intermediate values
+of \(\lambda\) (Fig. S8 red line). Finally, the distribution of
+\(\hat{\lambda}\) was not normal, and became more skewed at more extreme
+values of \(\lambda\) (Fig. S8 blue line). \hfill\break
+
+For \emph{K}, mean values increased with increasing phylogenetic signal,
+though as was found with pure-birth trees, the increase was nonlinear
+(Fig. S9 black line). Likewise, variation increased with increasing
+phylogenetic signal (Fig. S9 red line), though the distribution of
+\emph{K} was more normally distributed throughout its range, and across
+different tree sizes, as compared with \(\hat{\lambda}\) (Fig. S9 blue
+line). \hfill\break
+
+With respect to effect sizes, both \(Z_{\lambda}\) and \(Z_K\) increased
+with increasing input phylogenetic signal, but \(\hat{\lambda}\) was
+strongly affected by tree size, whereas \(Z_K\) was more consistent
+(Fig. S10). Also, \(Z_K\) increased more linearly with increasing levels
+of phylogenetic signal, and its standard deviation across input signal
+was more even across tree sizes, implying more consistent precision.
+Thus, polytomies do not exert an appreciable effect on \(Z_K\).
+
+\includegraphics[width=0.95\linewidth]{fig.S8}
+
+\textbf{Figure S8}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion on pure-birth trees containing polytomies.
+Gray line signifies the 1:1 line where the input value matches the
+estimate. At each input level, the dark black line represents the
+empirically derived expected value (mean) of \(\hat\lambda\), the red
+line is the standard deviation of \(\hat\lambda\), and the blue line is
+Shapiro Wilks statistic of \(\hat\lambda\) (\(W=1.0\) signifies
+normality, \(W< 1.0\) represent skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S9}
+
+\textbf{Figure S9}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion on pure-birth trees containing polytomies.
+At each input level, the dark black line represents the empirically
+derived expected value (mean) of \textit{K}, the red line is the
+standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
+
+\includegraphics[width=0.95\linewidth]{fig.S10}
+
+\textbf{Figure S10}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_K\) to increasing strength of Brownian motion for pure-birth trees
+containing polytomies. All results are shown for smallest and largest
+trees in the top four panels; only means of results are shown in the
+bottom two panels, to facilitate comparison with respect to tree size.
+
+\newpage
 
 \hypertarget{simulations-of-phylogenetic-regression-and-anova}{%
-\subsection{Simulations of phylogenetic regression and
+\subsection{Simulations of Phylogenetic Regression and
 ANOVA}\label{simulations-of-phylogenetic-regression-and-anova}}
 
-We analyzed the precision of parameter estimation when parameters were
-obtained in phylogenetic regression and ANOVA. This involved simulations
-of dependent variables with \(\lambda\) values ranging from 0 to 1 (21
-intervals of 0.05 units) across pure-birth trees (\(n=2^5 - 2^{10}\)).
-For each \(\lambda\) input value, independent variables were then
-generated with a known relationship to the simulated dependent variable
-(\(\beta = 0, 0.25, 0.5, 0.75, 1.0\)), from which \(\lambda\) was then
-estimated.
+We also investigated the effect of incorporating \(\lambda\) in PGLS
+analyses (phylogenetic regression and ANOVA). Here, patterns for
+\(\hat\lambda\), \emph{K}, \(Z_\lambda\) and \(Z_K\) were virtually
+identical to those found previously. In essence, \(\hat\lambda\) was not
+suitable as a test statistic representing phylogenetic signal due to all
+properties previously shown (Figs. S11 \& S12), \emph{K} was more
+appropriate, but associated non-linearly with signal strength (Figs. S13
+\& 14), while \(Z_K\) was found to be a superior effect size as compared
+with \(Z_\lambda\) (Figs S15 \& S16). \hfill\break
 
-\emph{Results} As with the results from the regression analyses in the
-main text, we found poor precision of \(\lambda\) estimation in smaller
-trees and at intermediate levels of phylogenetic signal with
-phylogenetic ANOVA (Fig. S3). Slope was reliably estimated across tree
-sizes and known \(\lambda\) values (Fig. S4).
+With respect to model performance, input parameters (\(\beta\)) were
+well estimated in phylogenetic regression and ANOVA when \(\lambda\) was
+incorporated (Figs S17 \& S18), implying that \(\lambda\) may be used as
+a tuning parameter for these models. Additionally, type I error and
+statistical power were unaffected by the inclusion of \(\lambda\) (Figs
+S19 \& S20), though power was reduced at smaller sample sizes in
+phylogenetic ANOVA, mirroring prior findings of (2). These results
+generally confirmed earlier findings of (3) for phylogenetic regression,
+and extend them to the case of phylogenetic ANOVA.
 
-\includegraphics[width=0.95\linewidth]{FigS3}
+\includegraphics[width=0.95\linewidth]{fig.S11}
 
-\textbf{Figure S3}. Precision of Pagel's \(\lambda\) when incorporated
-in phylogenetic ANOVA. Results are from input \(\beta\) values of
-\(0.5\). \hfill\break
+\textbf{Figure S11}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion in phylogenetic regression (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate. At each input level, the dark black line
+represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
-\includegraphics[width=0.95\linewidth]{FigS4}
+\includegraphics[width=0.95\linewidth]{fig.S12}
 
-\textbf{Figure S4}. Precision of slope estimates across known input
-\(\beta\) values of phylogenetic regression. Results shown include all
-input \(\lambda\) values.
+\textbf{Figure S12}. Response of Pagel's \(\lambda\) to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
+\(\lambda\)). Gray line signifies the 1:1 line where the input value
+matches the estimate. At each input level, the dark black line
+represents the empirically derived expected value (mean) of
+\(\hat\lambda\), the red line is the standard deviation of
+\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
+\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
+skewed distributions).
 
-\hypertarget{effect-sizes-of-pagels-lambda-z_lambda-and-kappa-z_kappa}{%
-\subsection{\texorpdfstring{Effect Sizes of Pagel's Lambda
-(\(Z_\lambda\)) and Kappa
-(\(Z_{\kappa}\))}{Effect Sizes of Pagel's Lambda (Z\_\textbackslash{}lambda) and Kappa (Z\_\{\textbackslash{}kappa\})}}\label{effect-sizes-of-pagels-lambda-z_lambda-and-kappa-z_kappa}}
+\includegraphics[width=0.95\linewidth]{fig.S13}
 
-We proposed the use of effect sizes (Z-scores) as a measure of the
-strength of phylogenetic signal. These were obtained from summary
-parameters of phylogenetic signal: \(\lambda\) and \(\kappa\). The
-precision of effect sizes \(Z_\lambda\) and \(Z_{\kappa}\) are
-summarized in the main text, primarily for simulations where \(n=32\).
-Here we present results from additional simulations.
+\textbf{Figure S13}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion in phylogenetic regression (incorporating
+\(\lambda\)). At each input level, the dark black line represents the
+empirically derived expected value (mean) of \textit{K}, the red line is
+the standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
 
-\emph{Results} Effect sizes for Pagel's \(\lambda\) (\(Z_\lambda\))
-scale nonlinearly with input levels of phylogenetic signal, and this
-pattern is consistent across all tree sizes examined (Figure S5).
-Further, the precision of \(Z_\lambda\) varied widely across input
-levels of phylogenetic signal, with decidely greater variation (less
-precision) at stronger input levels of phylogenetic signal (Figure S5).
-By contrast, effect sizes for \(\kappa\) (\(Z_{\kappa}\)) displayed a
-linear increase with increasing input levels of phylogenetic signal,
-which remained consistent across all tree sizes examined (Figure S6).
-Additionally, the precision of \(Z_{\kappa}\) remained constant across
-input levels of phylogenetic signal (Figure S6). Together these results
-imply that \(Z_{\kappa}\) is a more reliable estimate of the strength of
-phylogenetic signal.
+\includegraphics[width=0.95\linewidth]{fig.S14}
 
-\includegraphics[width=0.95\linewidth]{FigS5}
+\textbf{Figure S14}. Response of Blomberg's \textit{K} to increasing
+strength of Brownian motion in phylogenetic ANOVA (incorporating
+\(\lambda\)). At each input level, the dark black line represents the
+empirically derived expected value (mean) of \textit{K}, the red line is
+the standard deviation of \textit{K}, and the blue line is Shapiro Wilks
+statistic of \textit{K} (\(W=1.0\) signifies normality, \(W< 1.0\)
+represent skewed distributions).
 
-\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's
-\(\lambda\) across input \(\lambda\) values on various tree sizes.
-\hfill\break
+\includegraphics[width=0.95\linewidth]{fig.S15}
 
-\includegraphics[width=0.95\linewidth]{FigS6}
+\textbf{Figure S15}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_K\) to increasing strength of Brownian motion in phylogenetic
+regression (incorporating \(\lambda\)). All results are shown for
+smallest and largest trees in the top four panels; only means of results
+are shown in the bottom two panels, to facilitate comparison with
+respect to tree size.
 
-\textbf{Figure S6}. Variation in effect size (Z score) of \(\kappa\)
-across input \(\lambda\) values on various tree sizes. \hfill\break
+\includegraphics[width=0.95\linewidth]{fig.S16}
 
-\includegraphics[width=0.95\linewidth]{FigS7}
+\textbf{Figure S16}. Response of effect sizes \(Z_{\lambda}\) and
+\(Z_K\) to increasing strength of Brownian motion in phylogenetic ANOVA
+(incorporating \(\lambda\)). All results are shown for smallest and
+largest trees in the top four panels; only means of results are shown in
+the bottom two panels, to facilitate comparison with respect to tree
+size.
 
-\textbf{Figure S7}. Estimates of phylogenetic signal using \(\kappa\)
-across known lambda inputs on various tree sizes. Plot shows that while
-\(\kappa\) increases with increasing phylogenetic signal, the precision
-of this estimate also varies. Thus, conversion to a standardized effect
-size (\(Z_{\kappa}\)) is preferred.
+\includegraphics[width=0.95\linewidth]{fig.S17}
+
+\textbf{Figure S17}. Parameter estimates from phylogenetic regression
+(incorporating \(\lambda\)). The red line represents the estimated
+slope, which almost completely overlaps the gray 1:1 line.
+
+\includegraphics[width=0.95\linewidth]{fig.S18}
+
+\textbf{Figure S18}. Parameter estimates from phylogenetic ANOVA
+(incorporating \(\lambda\)). The red line represents the estimated
+slope, which almost completely overlaps the gray 1:1 line.
+
+\includegraphics[width=0.95\linewidth]{fig.S19}
+
+\textbf{Figure S19}. Power curves for phylogenetic regression
+(incorporating \(\lambda\)). Type I error is found on the far-left of
+the plot (effect = 0).
+
+\includegraphics[width=0.95\linewidth]{fig.S20}
+
+\textbf{Figure S20}. Power curves for phylogenetic ANOVA (incorporating
+\(\lambda\)). Type I error is found on the far-left of the plot (effect
+= 0).
+
+\newpage
+
+\hypertarget{references}{%
+\subsection*{References}\label{references}}
+\addcontentsline{toc}{subsection}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
+1. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic signal;
+strong or negligible impacts of polytomies and branch length
+information? \emph{BMC evolutionary biology} 17(1):53.
+
+\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
+2. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
+aggregation, biological challenges, and a refined permutation procedure.
+\emph{Evolution} 72(6):1204--1215.
+
+\leavevmode\hypertarget{ref-Revell2010}{}%
+3. Revell LJ (2010) Phylogenetic signal and linear regression on species
+data. \emph{Methods in Ecology and Evolution} 1:319--329.
 

### 2020-09-17 Manuscript/PNAS-Adams-Submitted.Rmd
empirical update

@@
 4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
-While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V ($\hat{Z}_{12}=4.13$; $P=0.000036$: Fig.
+While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V ($\hat{Z}_{12}=16.51$; $P< 0.00001$: Fig.
 4).

### 2020-09-17 Data_Analyses/CompareZ.r

@@ -50,13 +50,13 @@ compare.Z <- function(...,two.tailed = TRUE){
   if(length(dots) < 2) stop("At least two objects are needed")
   if(is.null(list.names)) list.names <- paste("Analysis", 1:n, sep = ".")
   names(dots) <- list.names
-  random.val <- lapply(1:length(dots), function(j) { #Check class; log random values as needed
+  random.val <- lapply(1:length(dots), function(j) { 
     if (is(dots[[j]],"lm.rrpp")  && is.null(dots[[j]]$ANOVA$SS))  { rep(0,dim(dots[[j]]$Y)[1])}  #For intercept model
-    else if (is(dots[[j]],"lm.rrpp") && !is.null(dots[[j]]$ANOVA$SS))  {log(dots[[j]]$ANOVA$SS)}  #NOTE: SS  only for now
+    else if (is(dots[[j]],"lm.rrpp") && !is.null(dots[[j]]$ANOVA$SS))  {(dots[[j]]$ANOVA$SS)}  #NOTE: SS  only for now
     else if (is(dots[[j]],"CR")) {t(as.matrix(dots[[j]]$random.CR))} 
     else if (is(dots[[j]],"pls")) {t(as.matrix(dots[[j]]$random.r))} 
-    else if (is(dots[[j]],"physignal")) {t(as.matrix(log(dots[[j]]$random.K)))} 
-    else if (is(dots[[j]],"evolrate")) {t(as.matrix(log(dots[[j]]$random.sigma)))}
+    else if (is(dots[[j]],"physignal")) {t(as.matrix((dots[[j]]$random.K)))} 
+    else if (is(dots[[j]],"evolrate")) {t(as.matrix((dots[[j]]$random.sigma)))}
   })
   sdn <- function(x) sqrt(sum((x-mean(x))^2)/length(x))
   getdrs <- function(fit){  
@@ -71,10 +71,13 @@ compare.Z <- function(...,two.tailed = TRUE){
     if(is.null(fit)) 0 else
       RRPP:::effect.size(as.vector(fit))
   }
+  ## box.cox
+  random.val <- lapply(1:length(random.val), function(j) box.cox(random.val[[j]])$transformed)
+  
   #added unlist b/c could have multiple factors in lm.rrpp
-  list.drs <- unlist(sapply(1:length(random.val), function(j) { apply(random.val[[j]],1,getdrs) }))
[…]

### 2020-09-17 Data_Analyses/EmpiricalExample/CompareZ.r

@@ -1,0 +1,114 @@
+#' Comparisons of Empirically-Derived Effect Sizes from Multiple Analyses
+#'
+#' Function performs an analysis to compare the empirically-generated effect sizes of two or more analyses
+#'
+#' The function statistically compares the effect sizes derived from empirical sampling distributions for 
+#' two or more statistical analyses. The approach is generalized from effect sizes derived from residual 
+#' randomization permutation procedures (RRPP: Collyer et al. 2015: see also Adams and Collyer 2016; 2019). 
+#' The analysis calculates effect sizes as standard deviates, z, from empirically-derived sampling distributions.
+#' These provide a measure of the strength of signal for a particular analysis (e.g., the strength of associaton 
+#' from a partial least squares analysis, the strength of the effect of a factor in a linear model, etc.).
+#' Effect sizes are then statistically compared using a two-sample z-test, based on the pooled 
+#' standard error from the sampling distributions of the analyses. The method follows suggestions in 
+#' Adams and Collyer (2019), which generalized the use of empirically-derived sampling distributions from 
+#' permutation as a means to compare the strength of patterns in multivariate datasets. 
+#' 
+#' To use this function, simply perform a series of analyses in geomorph or RRPP. Any number of objects can be used as 
+#' input. The function will then determine the class of each object: "CR", "evolrate", "lm.rrpp", 
+#' "physignal" or "pls", and will extract the appropriate components. It will then estimate effect sizes for each object
+#' and perform pairwise comparisons of the strength of signal of those objects based on a two-sample z-score (NOTE: 
+#' If one linear model is used as input the function will compare effect sizes across factors within that linear
+#' model). Finally, one may perform the comparison as either a one-tailed or a two-tailed (default) test.
+#' 
+#' @param ... saved analyses of any RRPP or geomorph class
[…]

### 2020-09-17 Data_Analyses/EmpiricalExample/EmpiricalExampleZ12.R

@@ -1,0 +1,38 @@
+# Compare Phylogenetic Signal: SA:V vs. BW/SVL 
+
+# Libraries #### 
+library(stringr)
+library(ape)
+library(geomorph) # Morphological analyses
+library(geiger) # for pruning trees and datasets to each other (function datatree)
+
+Tree <- read.tree("Bonett_and_Blair_Phylo.tre") # Full Tree
+LinMeas <- read.csv("LinMeas.SpeciesSummaries.csv")
+
+row.names(LinMeas) <- LinMeas$Species
+TreeDataPruned <- treedata(Tree, LinMeas) 
+TreePruned <- TreeDataPruned$phy
+LinMeasOrdered <- LinMeas[match(TreePruned$tip.label, LinMeas$Species),]
+SVL <- LinMeasOrdered$SVL; names(SVL) <- rownames(LinMeasOrdered)
+BWL <- LinMeasOrdered$BWL /SVL; names(BWL) <- rownames(LinMeasOrdered)
+
+DF <- read.csv("SAV.Disparity.csv", header = TRUE)
+SAtoV <- DF$SAtoV; names(SAtoV) <-DF$X
+DATA <- treedata(TreePruned, SAtoV, warnings= T)
+
+###### Analyses
+res.BWL <-physignal(A = BWL, phy = TreePruned)
+summary(res.BWL)
+plot(res.BWL)
+length(BWL)            
+
+res.SAToV<- physignal(A = DATA$data, phy = DATA$phy)
+summary(res.SAToV)
+plot(res.SAToV)
+
+### Compare Z12
+source('newBC.r')
+source('CompareZ.r')
+
+compare.Z(res.SAToV,res.BWL)
+

### 2020-09-17 Data_Analyses/EmpiricalExample/newBC.r

@@ -1,0 +1,49 @@
+#### functions from develop branch ######
+center <- function(x){
+  if(is.vector(x)) x - mean(x) else {
+    x <- as.matrix(x)
+    dims <- dim(x)
+    fast.center(x, dims[1], dims[2])
+  }
+}
+#Adjust box.cox to use distribution minus observed, then transform observed
+box.cox <- function(y, eps = 0.02) {
+  y.obs <- y[1];   y <- y[-1] #pull out observed
+  if(any(y <= 0)) y = y - min(y) + 0.0001
+  n <- length(y)
+  yy <- y/exp(mean(log(y)))
+  logy <- log(yy)
+  lambda <- seq(-2.4, 2.4, 0.4)
+  m <- length(lambda)
+  
+  loglik <- sapply(1:m, function(j){ # same as MASS::boxcox loglik f
+    la <- lambda[j]
+    yt <- if(abs(la) > eps) yt <- (yy^la - 1)/la else
+      logy * (1 + (la * logy)/2 * (1 + (la * logy)/3 * (1 + (la * logy)/4)))
+    
+    -n/2 * log(sum(center(yt)^2))
+  })
+  
+  lambda.opt <- lambda[which.max(loglik)][[1]]
+  
+  if(abs(lambda.opt) == 2.4) {
+    lambda <- seq(-3.2, 3.2, 0.2)
+    m <- length(lambda)
+    
+    loglik <- sapply(1:m, function(j){ # same as MASS::boxcox loglik f
+      la <- lambda[j]
+      yt <- if(abs(la) > eps) yt <- (yy^la - 1)/la else
+        logy * (1 + (la * logy)/2 * (1 + (la * logy)/3 * (1 + (la * logy)/4)))
+      
+      -n/2 * log(sum(center(yt)^2))
+    })
+    lambda.opt <- lambda[which.max(loglik)][[1]]
+  }
+  
+  sp <- spline(lambda, loglik, n = 100)
+  lambda.opt <- sp$x[which.max(sp$y)]
+  if(abs(lambda.opt) < eps) lambda.opt <- 0
+  y <- c(y.obs,y) #put back in here
+  res <- if(lambda.opt == 0) log(y) else (y^lambda.opt - 1)/lambda.opt
+  list(opt.lambda = lambda.opt, transformed = res, lambda = sp$x, loglik = sp$y)
+}

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.Rmd
MEE version

--- 2020-09-17 Manuscript/PNAS-Adams-Submitted.Rmd
@@
 ---
-title: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal
+title: "A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal"
 
-# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-author:
-  - name: Dean C. Adams
-    affiliation: a,2
-  - name: Erica K. Baken
-    affiliation: a,b
-  - name: Michael L. Collyer
-    affiliation: b
-address:
-  - code: a
-    address: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, 50010. USA.
-  - code: b
-    address: Department of Science, Chatham University, Pittsburgh, Pennsylvania, 15232. USA.
+output:
+  pdf_document:
+    fig_caption: yes
+    number_sections: yes
+    keep_tex: true
+    pandoc_args:
 
-corresponding_author:
-  - code: 2
-    text: "To whom correspondence should be addressed. E-mail: dcadams@iastate.edu"
+csl: methods-in-ecology-and-evolution.csl
+bibliography: Lambda-Refs.bib
 
-# For footer text
-lead_author_surname: Adams et al.
+header-includes:
+    - \usepackage{setspace}\doublespacing
+    - \usepackage{lineno}\linenumbers
 
-author_contributions: |
-  D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the research and wrote the paper.
+---
+\begin{center}
+\textbf{Dean C. Adams$^{1,*}$, Erica K. Baken$^{1,2}$, and Michael L. Collyer$^2$}
+\end{center}
 
-## Remove this if not required
-conflict_of_interest: |
-  The authors declare no conflict of interest.  \hfill\break
+\begin{center}`r format(Sys.time(), '%d %B, %Y')`\end{center}
 
-  Data deposition: Data for the empirical example may be found on DRYAD: doi:10.5061/dryad.b554m44 and doi:10.5061/dryad.59zw3r23m.
+$^{1}$Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA.
+
+$^{2}$Department of Science, Chatham University, Pittsburgh, Pennsylvania, USA.
+
+$^{*}$Correspondence: Dean C. Adams <dcadams@iastate.edu> 
+\hfill\break
+
+**Keywords**: phylogenetic signal, macroevolution, lambda, kappa \hfill\break
+
+**Short Title**: Effect size for phylogenetic signal \hfill\break
+
+```{r setup, include=FALSE}
+  library(knitr)
+  knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r, include = FALSE}
+library(stringr)
+library(tidyverse)
+
+RmdWords <- function(file) {
+
+  # Creates a string of text
+  file_string <- file %>%
+    readLines() %>%
+    paste0(collapse = " ") %>%
+    # Remove YAML header
+    str_replace_all("^<--- .*?--- ", "") %>%    
+    str_replace_all("^--- .*?--- ", "") %>%
+    # Remove code
+    str_replace_all("```.*?```", "") %>%
+    str_replace_all("`.*?`", "") %>%
+    # Remove LaTeX
+    str_replace_all("[^\\\\]\\$\\$.*?[^\\\\]\\$\\$", "") %>%
+    str_replace_all("[^\\\\]\\$.*?[^\\\\]\\$", "") %>%
+    # Deletes text between tags
+    str_replace_all("TC:ignore.*?TC:endignore", "") %>%
+    str_replace_all("[[:punct:]]", " ") %>%
+    str_replace_all("  ", "") %>%
+    str_replace_all("<", "") %>%
+    str_replace_all(">", "")
+
+  # Save several different results
+  word_count <- str_count(file_string, "\\S+")
+  char_count <- str_replace_all(string = file_string, " ", "") %>% str_count()
+
+   return(list(num_words = word_count, num_char = char_count, word_list = file_string))
+}
+
+words <- RmdWords("2020-MEE-Adams_et_al-Physignal.Rmd")
+```
+
+**Research Article:**
+
+```{r, echo = FALSE}
+count <- c(words = words$num_words, characters = words$num_char)
+knitr::kable(t(count))
+```
+
+**Author Contributions**:  DCA conceived the original idea for this manuscript, and DCA, EKB, and MLC collaboratively developed the concept and contributed to all portions of this manuscript.
+All authors approve of the final product and are willingly accountable for any portion of the content.\hfill\break
+
+**Data Archiving**: Empirical data used in this paper are available on DRYAD (associated with original articles).
 R-scripts for simulation tests are found at: https://github.com/deanadams/PhySigCompareZ.
-Computer code for implementing the two-sample comparison of effect sizes is found in geomorph: https://cran.r-project.org/web/packages/geomorph/index.html \hfill\break
+Computer code for implementing the two-sample comparison of effect sizes is found in geomorph (**upon article acceptance**): https://cran.r-project.org/web/packages/geomorph/index.html \hfill\break
 
-  ^2^To whom correspondence should be addressed. E-mail: dcadams\@iastate.edu
+**Acknowledgments**: We thank E.
+Glynne and B.
+Juarez for comments on early drafts of the manuscript.
+This work was sponsored in part by National Science Foundation Grants DBI-1902511 (to DCA) and DBI-1902694 (to MLC).
+The authors have no known conflicts of interest.
 
-abstract: |
+\newpage
+
+\setcounter{section}{0}
+
+# Abstract {-}
+
+1.
 Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped.
-Here we evaluate the efficacy of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
+
+2.
+In this paper, we evaluate the efficacy of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values.
 We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal.
 Further, the precision of $\lambda$ varies with input signal.
+
+3.
 Another measure, Blomberg's *K*, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal.
@@
 Thus, neither $\lambda$ nor *K* are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets.
+
+4.
 As an alternative, we propose a standardized effect size based on *K*, ($Z_K$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison.
 We develop tests based on $Z_K$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis.
+
+5.
 Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
 
-significance: |
-  Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking.
-Here we develop a standardized effect size, $Z_K$, which measures the strength of phylogenetic signal on a common statistical scale.
-We also provide a test statistic, $\hat{Z}_{12}$, for comparing the strength of phylogenetic signal across datasets.
-We find that two commonly used parameters (Pagel's $\lambda$ and Blomberg's *K*), not converted to effect sizes, are unsuitable for this purpose.
-Our effect-size procedure enables biologists to quantitatively address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.
+\newpage
 
-acknowledgements: |
-  We thank E. Glynne and B. Juarez for comments on early drafts of the manuscript. This work was supported in part by NSF grant DBI-1902511 (to D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break
+# Introduction
 
-keywords:
-  - phylogenetic signal
-  - macroevolution
-  - lambda
-  - kappa
-
-## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
-pnas_type: pnasresearcharticle
-
-bibliography: Lambda-Refs.bib
-csl: pnas.csl
-
-## change to true to add optional line numbering
-lineno: true
-
-output: rticles::pnas_article
-
----
 Investigating macroevolutionary patterns of trait variation requires a phylogenetic perspective, because the shared ancestry of species violates the assumption of independence among trait values that is common for statistical tests [@Felsenstein1985; @HarveyPagel1991].
@@
 Indeed, under numerous evolutionary models, phylogenetic signal is expected, as stochastic character change along the hierarchical structure of the tree of life generates trait covariation among taxa [@Felsenstein1985; @Blomberg_et_al2003; @Revell_et_al2008].
+\hfill\break
 
@@
 When incorporated in PGLS, $\lambda$ serves as a tuning parameter which is optimized via log-likelihood profiling while evaluating the covariation between the dependent and independent variables, given the phylogeny [@Pagel1999; @Freckleton_et_al2002].
-To infer whether phylogenetic signal differs from no signal or a Brownian motion model of evolutionary divergence, the observed model fit using $\hat\lambda$ may be statistically compared to that using  $\lambda=0$ or $\lambda=1$ via likelihood ratio tests [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or confidence limits [@Vandelook2019].
+To infer whether phylogenetic signal differs from no signal or a Brownian motion (BM) model of evolutionary divergence, the observed model fit using $\hat\lambda$ may be statistically compared to that using  $\lambda=0$ or $\lambda=1$ via likelihood ratio tests [@Freckleton_et_al2002; @Cooper2010; @Bose2019] or confidence limits [@Vandelook2019].
+\hfill\break
 
@@
 Thus, it is tempting to regard both $\lambda$ and *K* as descriptive statistics that measure the relative strength of phylogenetic signal, providing an estimate of its magnitude for comparison.
+\hfill\break
 
@@
 However, because the "strength of Brownian motion" was simulated as a varied weighted-average of data simulated on trees with $\lambda=0$ and $\lambda=1$ and not as prescribed values of $\lambda$ [@Munkemuller_et_al2012], interpretation of these patterns is challenging.
+\hfill\break
 
@@
 Therefore, transformation of these statistics into $Z$-scores would allow evaluation of the efficacy of each statistic to yield effect sizes that could be used for comparisons of the strength of phylogenetic signal across traits and lineages.
+\hfill\break
 
@@
 By contrast, estimates of *K* are more consistent across tree sizes, and are normally distributed across the range of input levels of $\lambda$, making *K* a more reliable statistic.
-We then propose an effect size based on *K*, ($Z_K$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes and signal strength, and facilitates quantitative comparisons of the relative strength of phylogenetic signal across datasets.
+However, the relationship between *K* and input levels of phylogenetic signal is decidedly nonlinear.
+Thus, neither $\lambda$ nor *K* are suitable as effect sizes for measuring the strength of phylogenetic signal.
+As an alternative, we propose an effect size based on *K*, ($Z_K$), which provides consistent estimates of the strength of phylogenetic signal across tree sizes and signal strength.
+Further, because $Z_K$ places that phylogenetic signal on a common scale, it facilitates statistical comparisons of the relative strength of phylogenetic signal across datasets.
+We propose a two-sample statistic ($\hat{Z}_{12}$) to accomplish this task, and show that it displays appropriate levels of type I error and model misspecification.
+An empirical example is then provided to illustrate its use.
 
-# Results
+# Simulation methods and results
 
-\begin{figure}
-\centering
-\includegraphics[width=8.25cm]{{fig.1}.png}
-\caption{Response of Pagel's \(\lambda\) to increasing strength of
-Brownian motion. Gray line signifies the 1:1 line where the input value
-matches the estimate \(\hat\lambda\). At each input level, the dark
-black line represents the empirically derived expected value (mean) of
-\(\hat\lambda\), the red line is the standard deviation of
-\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
-\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
-skewed distributions).{}}
-\end{figure}
+Methods for characterizing phylogenetic signal were evaluated by computer simulation.
+Briefly, simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+We generated 50 random trees for each intersection of tree size and $\lambda$.
+For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
+For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and *K*, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
+For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
+Simulations were then repeated for both balanced and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
+Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information).
+\hfill\break
 
-**Lambda ($\lambda$) estimates of phylogenetic signal are inaccurate.** Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
+## Lambda ($\lambda$) estimates of phylogenetic signal are inaccurate
+
+Computer simulations reveal that for $\hat{\lambda}$, the distributional expectations of a Bernoulli variable were mostly upheld.
 First, the mean value of $\hat{\lambda}$ increases as $\lambda$ increases.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=8.25cm]{{fig.2}.png}
-\caption{Response of Blomberg's *K* to increasing strength of Brownian motion.  At each input level, the black line represents the empirically derived expected value (mean) of
-*K*, the red line is the standard deviation of
-*K*, and the blue line is Shapiro Wilks statistic of
-*K* (\(W=1.0\) signifies normality, \(W< 1.0\) represent
-skewed distributions).{}}
-\end{figure}
+## Kappa (*K*) estimates of phylogenetic signal are more reliable
 
-**Kappa (*K*) estimates of phylogenetic signal are more reliable.** Simulation results demonstrate that *K* displays better statistical properties.
+Simulation results demonstrate that *K* displays better statistical properties.
 First, as expected, mean values of *K* increase with increasing signal ($\lambda$) irrespective of tree size, albeit non-linearly (Fig.
@@
 
-**Effect sizes from *K* ($Z_K$) better characterize phylogenetic signal.**  To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and *K*.
-Statistically, a standardized effect size may be found as:
+## Effect sizes from *K* ($Z_K$) better characterize phylogenetic signal
+
+To measure the strength of phylogenetic signal on a common scale, we propose effect sizes (Z-scores) for both $\lambda$ and *K*. Statistically, a standardized effect size may be found as:
 
@@
 However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ simulated from permutation procedures.
+\hfill\break
 
@@
 The empirical sampling distribution of *K* can be first transformed via a Box-Cox transformation to better adhere to the assumption of normality.
+\hfill\break
 
@@
 
-where, $\hat{\lambda}$ is the maximized likelihood value of $\lambda$ and $\chi^{2}_{\hat{\lambda}}$ is the likelihood ratio statistic for the value.
+where, $\hat{\lambda}$ is the maximized likelihood value of $\lambda$ and $\chi^{2}_{\hat{\lambda}}$ is the likelihood ratio statistic for the value. \hfill\break
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=8.25cm]{{fig.3}.png}
-\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_K\) to increasing strength of Brownian motion.
-Means from simulation runs are shown for comparative ease.
-Individual values from each simulation run are available in Supporting Informaton.
-}
-\end{figure}
+## A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets
 
-**A test statistic ($\hat{Z}_{12}$) allows meaningful comparisons across datasets.** To statistically compare the strength of phylogenetic signal across datasets we propose a two-sample test statistic ($\hat{Z}_{12}$).
+To statistically compare the strength of phylogenetic signal across datasets we propose a two-sample test statistic ($\hat{Z}_{12}$).
 Based on statistical theory, a two-sample test statistic may be calculated as:
@@
 Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [@AdamsCollyer2016; @AdamsCollyer2019b].
-Simulations reveal that tests based on $\hat{Z}_{12}$ have appropriate type I error and model misspecification rates  of approximately 5% (Supporting Information).
+\hfill\break
+
+We evaluated the type I error and false discovery rates of tests based on $\hat{Z}_{12}$ using a procedure similar to that above.
+Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
+For each combination of $n$ and $\lambda$, a total of 100 traits were simulated on pure-birth trees (50 traits per tree) under a Brownian motion model of evolution.
+Next, the phylogenetic signal between traits was compared in pairwise fashion for all combinations of traits, using $\hat{Z}_{12}$.
+The proportion of significant results provided an estimate of the type I error or false discovery rate.
+Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$).
+Likewise, false discovery rates were evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$).
+\hfill\break
+
+Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig.
+4).
+Likewise, false discovery rates were also low, and were at or below the nominal 5% level (Fig.
+4).
+Importantly, the statistical performance of $\hat{Z}_{12}$ appears unaffected by phylogeny size.
+Overall, these results reveal that tests based on $\hat{Z}_{12}$ have acceptable type I error and false discovery rates.
+Therefore, $\hat{Z}_{12}$ is an appropriate statistic for comparing the degree of phylogenetic signal across traits.
+
+# Empirical example
 
 To demonstrate the utility of $\hat{Z}_{12}$, we compared  $Z_K$ for two ecologically-relevant traits in plethodontid salamander (Fig.
-4): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
+5): surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) [@Baken2019; @Baken2020].
+For this example, surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
+A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.
+The phylogenetic signal in each trait was then characterized using *K*, which was converted to its effect size ($Z_K$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph], and routines by the authors (**to be incorporated in `geomorph` upon acceptance**).
+\hfill\break
+
 While both traits contained significant phylogenetic signal, tests based on $\hat{Z}_{12}$ revealed that the degree of phylogenetic signal was significantly stronger in SA:V ($\hat{Z}_{12}=16.51$; $P< 0.00001$: Fig.
-4).
+5).
 Biologically, this observation may be interpreted by the fact that the tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V, which covaries with disparity in their climatic niches [@Baken2020].
 Thus, greater phylogenetic signal in SA:V is to be expected.
-
-\begin{figure}
-\centering
-\includegraphics[width=8.25cm]{{fig.4}.png}
-\caption{(A) Linear measures for relative body size, and regions of the
-body used to estimate surface area to volume (SA:V) ratios. (B)
-Permutation distributions of phylogenetic signal for SA:V and
-\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
-Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
-confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
-\end{figure}
 
@@
 From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and should not be used as an effect size for comparing the degree of phylogenetic signal between datasets.
+\hfill\break
 
@@
 We therefore recommend that future studies interested in evaluating the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect.
+\hfill\break
 
@@
 As such, our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.
+\hfill\break
 
@@
 Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of variables, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.
+\hfill\break
 
@@
 Furthermore, because comparison of effect sizes in a test is a comparison of locations of observed values in their sampling distributions, which would shift concomitantly because of this tendency, the $Z_{12}$ test statistic in equation 4 appears to be robust in spite of unresolved trees.
+\hfill\break
 
@@
 
-# Methods
+\newpage
 
-**Simulations.** Simulations were conducted by generating pure-birth phylogenies at each of six different tree sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-We generated 50 random trees for each intersection of tree size and $\lambda$.
-For each $\lambda$ within each tree size, continuous traits were then simulated on each phylogeny under a BM model of evolution.
-For each set of 50 trees we measured the mean values of $\hat{\lambda}$ and *K*, their standard deviation, and calculated the Shapiro-Wilk $W$ statistic as a departure from normality (symmetry).
-For the latter, a value of $1.0$ indicates normally distributed values, while departures from $1.0$ indicate skewness.
-Simulations were then repeated for both balanced and pectinate trees, which yielded qualitatively similar results (see Supporting Information).
-Trees containing polytomies, and an evaluation of $\hat{\lambda}$ from models of linear regression and phylogenetic ANOVA, were also investigated, and results were qualitatively similar to those reported above (see Supporting Information).
+# References {-}
 
-**Empirical Data.** Surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) measures were obtained from individuals of 305 species, from which species means were obtained [@Baken2019; @Baken2020].
-A time-dated molecular phylogeny for the group [@Bonett2017] was pruned to match the species in the phenotypic dataset.
-The phylogenetic signal in each trait was then characterized using *K*, which was converted to its effect size ($Z_K$) using `geomorph` 3.3.1 [@AdamsOtarola2013; @AdamsGeomorph], and routines by the authors (**to be incorporated in `geomorph` upon acceptance**).
+ \setlength{\parindent}{-0.25in}
+ \setlength{\leftskip}{0.25in}
+ \setlength{\parskip}{8pt}
+ \noindent
 
-<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
-\showmatmethods
-\showacknow
-\pnasbreak
+<div id="refs"></div>
+
+\newpage
+
+# Figures
+
+Figure 1.
+Response of Pagel's $\lambda$ to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$.
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).
+\hfill\break
+
+Figure 2.
+Response of Blomberg's $K$ to increasing strength of Brownian motion.
+At each input level, the black line represents the empirically derived expected value (mean) of $K$, the red line is the standard deviation of *K*, and the blue line is Shapiro Wilks statistic of $K$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).
+\hfill\break
+
+Figure 3.
+Response of effect sizes $Z_\lambda$ and $Z_K$ to increasing strength of Brownian motion.
+Means from simulation runs are shown for comparative ease.
+Individual values from each simulation run are available in Supporting Information.
+\hfill\break
+
+Figure 4. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
+
+Figure 5.
+(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).
+\hfill\break
+
+\newpage
+
+```{r, echo = FALSE, out.width="100%", fig.cap = "Response of Pagel's $\\lambda$ to increasing strength of Brownian motion.
+Gray line signifies the 1:1 line where the input value matches the estimate $\\hat\\lambda$.
+At each input level, the dark black line represents the empirically derived expected value (mean) of $\\hat\\lambda$, the red line is the standard deviation of $\\hat\\lambda$, and the blue line is Shapiro Wilks statistic of $\\hat\\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).
+" }
+
+include_graphics("Figs/fig.1.png")
+```
+
+```{r, echo = FALSE, out.width="100%", fig.cap = "Response of Blomberg's $K$ to increasing strength of Brownian motion.
+At each input level, the black line represents the empirically derived expected value (mean) of $K$, the red line is the standard deviation of $K$, and the blue line is Shapiro Wilks statistic of *K* ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions)."}
+
+include_graphics("Figs/fig.2.png")
+```
+
+```{r, echo = FALSE, out.width="100%", fig.cap = "Response of effect sizes $Z_\\lambda$ and $Z_K$ to increasing strength of Brownian motion.
+Means from simulation runs are shown for comparative ease.
+Individual values from each simulation run are available in Supporting Information."}
+
+include_graphics("Figs/fig.3.png")
+```
+
+```{r, echo = FALSE, out.height="75%", fig.cap="Type I error rates and model misspecification for $\\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\\lambda = 0$)."}
+include_graphics("Figs/fig.4.png")  
+```
+
+```{r, echo = FALSE, out.width="90%", fig.cap = "(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios.
+(B) Permutation distributions of phylogenetic signal for SA:V and $\\frac{BW}{SVL}$, with observed values shown as vertical bars.
+(C) Effect sizes ($Z_K$) for SA:V and $\\frac{BW}{SVL}$, with their 95\\% confidence intervals (CI not standardized by $\\sqrt(n)$)."}
+
+include_graphics("Figs/fig.5.png")
+```

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.tex
MEE version

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
Phylogenetic Signal}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{center}
\textbf{Dean C. Adams$^{1,*}$, Erica K. Baken$^{1,2}$, and Michael L. Collyer$^2$}
\end{center}

\begin{center}28 September, 2020\end{center}

\(^{1}\)Department of Ecology, Evolution, and Organismal Biology, Iowa
State University, Ames, Iowa, USA.

\(^{2}\)Department of Science, Chatham University, Pittsburgh,
Pennsylvania, USA.

\(^{*}\)Correspondence: Dean C. Adams
\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}
\hfill\break

\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa
\hfill\break

\textbf{Short Title}: Effect size for phylogenetic signal \hfill\break

\textbf{Research Article:}

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
4201 & 26493\tabularnewline
\bottomrule
\end{longtable}

\textbf{Author Contributions}: DCA conceived the original idea for this
manuscript, and DCA, EKB, and MLC collaboratively developed the concept
and contributed to all portions of this manuscript. All authors approve
of the final product and are willingly accountable for any portion of
the content.\hfill\break

\textbf{Data Archiving}: Empirical data used in this paper are available
on DRYAD (associated with original articles). R-scripts for simulation
tests are found at: \url{https://github.com/deanadams/PhySigCompareZ}.
Computer code for implementing the two-sample comparison of effect sizes
is found in geomorph (\textbf{upon article acceptance}):
\url{https://cran.r-project.org/web/packages/geomorph/index.html}
\hfill\break

\textbf{Acknowledgments}: We thank E. Glynne and B. Juarez for comments
on early drafts of the manuscript. This work was sponsored in part by
National Science Foundation Grants DBI-1902511 (to DCA) and DBI-1902694
(to MLC). The authors have no known conflicts of interest.

\newpage

\setcounter{section}{0}

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Macroevolutionary studies frequently characterize the phylogenetic
  signal in phenotypes, however, analytical tools for comparing the
  strength of that signal across traits remain largely underdeveloped.
\item
  In this paper, we evaluate the efficacy of Pagel's \(\lambda\) to
  correctly estimate the strength of phylogenetic signal in phenotypic
  traits across a range of input values. We find that \(\lambda\)
  behaves as a Bernoulli random variable, where estimates are
  increasingly skewed at larger and smaller input levels of phylogenetic
  signal. Further, the precision of \(\lambda\) varies with input
  signal.
\item
  Another measure, Blomberg's \emph{K}, is more consistent across a
  range of tree sizes, and exhibits a positive relationship with input
  levels of phylogenetic signal. However, that relationship is decidedly
  nonlinear. Thus, neither \(\lambda\) nor \emph{K} are suitable as
  effect sizes for measuring the strength of phylogenetic signal, and
  comparing that signal across datasets.
\item
  As an alternative, we propose a standardized effect size based on
  \emph{K}, (\(Z_K\)), which measures the strength of phylogenetic
  signal more reliably than does \(\lambda\), and places that signal on
  a common scale for statistical comparison. We develop tests based on
  \(Z_K\) to provide a mechanism for formally comparing the strength of
  phylogenetic signal across datasets, in much the same manner as effect
  sizes may be used to summarize patterns in quantitative meta-analysis.
\item
  Our approach extends the phylogenetic comparative toolkit to address
  hypotheses that compare the strength of phylogenetic signal between
  various phenotypic traits, even when those traits are found in
  different evolutionary lineages or have different units or scales.
\end{enumerate}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry of species
violates the assumption of independence among trait values that is
common for statistical tests (Felsenstein, 1985; Harvey \& Pagel, 1991).
Accounting for this evolutionary non-independence is the purview of
\emph{phylogenetic comparative methods} (PCMs): a suite of analytical
tools that condition the data by the phylogenetic relatedness of
observations (Grafen, 1989; Martins \& Hansen, 1997; Garland \& Ives,
2000; Rohlf, 2001; O'Meara, Ane, Sanderson, \& Wainwright, 2006;
Beaulieu, Jhwueng, Boettiger, \& O'Meara, 2012; Adams, 2014b; Adams \&
Collyer, 2018). PCMs are predicated on the notion that phylogenetic
signal -- the tendency for closely related species to display similar
trait values -- is present in cross-species datasets (Felsenstein, 1985;
Pagel, 1999; Blomberg, Garland, \& Ives, 2003). Indeed, under numerous
evolutionary models, phylogenetic signal is expected, as stochastic
character change along the hierarchical structure of the tree of life
generates trait covariation among taxa (Felsenstein, 1985; Blomberg et
al., 2003; Revell, Harmon, \& Collar, 2008). \hfill\break

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets (Gittleman \& Kot, 1990; Abouheif, 1999;
Pagel, 1999; Blomberg et al., 2003; Klingenberg \& Gidaszewski, 2010;
Adams, 2014a), and their statistical properties -- namely type I error
rates and statistical power -- have been investigated to determine under
what conditions phylogenetic signal can be detected (Revell et al.,
2008; Revell, 2010; Boettiger, Coop, \& Ralph, 2012; Diniz-Filho,
Santos, Rangel, \& Bini, 2012; Münkemüller et al., 2012; Pavoine \&
Ricotta, 2012; Adams, 2014a; Molina-Venegas \& Rodríguez, 2017). One of
the most widely used methods for characterizing phylogenetic signal is
Pagel's \(\lambda\) (Pagel, 1999), which transforms the lengths of the
internal branches of the phylogeny to improve the fit of data to the
phylogeny via maximum likelihood (Pagel, 1999; Freckleton, Harvey, \&
Pagel, 2002). When incorporated in PGLS, \(\lambda\) serves as a tuning
parameter which is optimized via log-likelihood profiling while
evaluating the covariation between the dependent and independent
variables, given the phylogeny (Pagel, 1999; Freckleton et al., 2002).
To infer whether phylogenetic signal differs from no signal or a
Brownian motion (BM) model of evolutionary divergence, the observed
model fit using \(\hat\lambda\) may be statistically compared to that
using \(\lambda=0\) or \(\lambda=1\) via likelihood ratio tests
(Freckleton et al., 2002; Cooper, Jetz, \& Freckleton, 2010; Bose,
Ramesh, Pélissier, \& Munoz, 2019) or confidence limits (Vandelook et
al., 2019). \hfill\break

Another widely used measure is Blomberg's \emph{K} (Blomberg et al.,
2003), which characterizes phylogenetic signal as the ratio of observed
trait variation to the amount of variation expected under Brownian
motion. Blomberg's \emph{K} can be treated as a test statistic by
employing a permutation test to generate its sampling distribution
(Blomberg et al., 2003; Adams, 2014a) for determining whether
significant phylogenetic signal is present in data. Both \(\lambda\) and
\emph{K} seem intuitive to interpret, as a value of \(0\) for both
corresponds to no phylogenetic signal, while a value of \(1\)
corresponds to the amount of phylogenetic signal expected under Brownian
motion. Thus, it is tempting to regard both \(\lambda\) and \emph{K} as
descriptive statistics that measure the relative strength of
phylogenetic signal, providing an estimate of its magnitude for
comparison. \hfill\break

The appeal of Pagel's \(\lambda\) and Blomberg's \emph{K} as descriptive
statistics is that they provide a basis for interpreting ``weak'' versus
``strong'' phylogenetic signal; i.e., small versus large values of
\(\hat{\lambda}\) or \emph{K}, respectively, in a comparative sense (De
Meester, Huyghe, \& Van Damme, 2019; Pintanel, Tejedo, Ron, Llorente, \&
Merino-Viteri, 2019; Su, Villéger, \& Brosse, 2019). Nonetheless, an
important question that has yet to be considered is whether such
comparisons are analytically appropriate, and whether these statistics
are, or can be, converted to effect sizes for comparative analyses
across datasets. To be statistics representing phylogenetic signal, they
should have reliable distributional properties, which could be revealed
with simulation experiments. For instance, as a proportional random
variable bounded by \(0\) and \(1\), we might expect that
\(\hat{\lambda}\) is a random variable that follows the distribution of
a Bernoulli probability parameter (Forbes, Evans, Hastings, \& Peacock,
2011); i.e., branch lengths in a tree are scaled proportionally to the
probability that data arise from a BM process. Given a known \(\lambda\)
value used to generate random data on a tree, we would also expect that
the mean of an empirical sampling distribution of \(\hat{\lambda}\)
would approximately equal \(\lambda\); the dispersion of
\(\hat{\lambda}\) would be largest at intermediate values of
\(\lambda\), \(\hat{\lambda}\) would be predictable over the range of
\(\lambda\) with respect to tree size; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) would be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (Forbes et al., 2011). Prior work (Münkemüller et al., 2012) seems
to support some of these conjectures, based superficially on statistical
moments for a given tree size (mean, variance, skewness, and kurtosis;
see Fig. 2 of ref. (Münkemüller et al., 2012)). However, because the
``strength of Brownian motion'' was simulated as a varied
weighted-average of data simulated on trees with \(\lambda=0\) and
\(\lambda=1\) and not as prescribed values of \(\lambda\) (Münkemüller
et al., 2012), interpretation of these patterns is challenging.
\hfill\break

By contrast, for Blomberg's \emph{K}, which is positively unbounded, we
might expect that for any \(\lambda\) used to generate data, estimates
of \emph{K} might be a random variable that follows a normal
distribution, with values distributed symmetrically (Forbes et al.,
2011). This attribute seemed less reasonable based on the simulations
performed by Münkemüller et al.~(2012), which suggested that
distributions were positively skewed and that Blomberg's \emph{K} might
not behave as a statistic that follows a normal distribution. However,
because their simulations used a weighted combination of simulated
phylogenetic signal strengths, strong inferences are not possible (and
distributional attributes were not the intended result of their
simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's \emph{K},
evaluation of statistical moments across a range of \(\lambda\) used to
generate data would be valuable for adjudicating the reliability of
these statistics as effect sizes. Furthermore, the expected values of
these statistics appear to vary with tree size (Münkemüller et al.,
2012), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores would allow
evaluation of the efficacy of each statistic to yield effect sizes that
could be used for comparisons of the strength of phylogenetic signal
across traits and lineages. \hfill\break

Here we use simulation experiments to compare the distributional
attributes of \(\hat{\lambda}\) and \emph{K}, plus their effect sizes
(\(Z\)-scores), across a range of tree size and phylogenetic signal
strength. We find that estimates of \(\hat{\lambda}\) are increasingly
skewed at larger and smaller input levels of phylogenetic signal and at
smaller tree sizes, vary widely for a given input value of \(\lambda\),
and that the precision of \(\hat{\lambda}\) is not constant across its
range. By contrast, estimates of \emph{K} are more consistent across
tree sizes, and are normally distributed across the range of input
levels of \(\lambda\), making \emph{K} a more reliable statistic.
However, the relationship between \emph{K} and input levels of
phylogenetic signal is decidedly nonlinear. Thus, neither \(\lambda\)
nor \emph{K} are suitable as effect sizes for measuring the strength of
phylogenetic signal. As an alternative, we propose an effect size based
on \emph{K}, (\(Z_K\)), which provides consistent estimates of the
strength of phylogenetic signal across tree sizes and signal strength.
Further, because \(Z_K\) places that phylogenetic signal on a common
scale, it facilitates statistical comparisons of the relative strength
of phylogenetic signal across datasets. We propose a two-sample
statistic (\(\hat{Z}_{12}\)) to accomplish this task, and show that it
displays appropriate levels of type I error and model misspecification.
An empirical example is then provided to illustrate its use.

\hypertarget{simulation-methods-and-results}{%
\section{Simulation methods and
results}\label{simulation-methods-and-results}}

Methods for characterizing phylogenetic signal were evaluated by
computer simulation. Briefly, simulations were conducted by generating
pure-birth phylogenies at each of six different tree sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 50
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, continuous traits were then
simulated on each phylogeny under a BM model of evolution. For each set
of 50 trees we measured the mean values of \(\hat{\lambda}\) and
\emph{K}, their standard deviation, and calculated the Shapiro-Wilk
\(W\) statistic as a departure from normality (symmetry). For the
latter, a value of \(1.0\) indicates normally distributed values, while
departures from \(1.0\) indicate skewness. Simulations were then
repeated for both balanced and pectinate trees, which yielded
qualitatively similar results (see Supporting Information). Trees
containing polytomies, and an evaluation of \(\hat{\lambda}\) from
models of linear regression and phylogenetic ANOVA, were also
investigated, and results were qualitatively similar to those reported
above (see Supporting Information). \hfill\break

\hypertarget{lambda-lambda-estimates-of-phylogenetic-signal-are-inaccurate}{%
\subsection{\texorpdfstring{Lambda (\(\lambda\)) estimates of
phylogenetic signal are
inaccurate}{Lambda (\textbackslash lambda) estimates of phylogenetic signal are inaccurate}}\label{lambda-lambda-estimates-of-phylogenetic-signal-are-inaccurate}}

Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases. Second, the precision in estimating \(\lambda\) varies across
the range of input values, as the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig.1 red line). Third, the distributions of
\(\hat{\lambda}\) tend toward normal distributions at intermediate
levels of \(\lambda\) but become increasingly skewed at more extreme
values of \(\lambda\) (Fig. 1 blue line). For small tree sizes, it is
also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). However, the mean value of
\(\hat{\lambda}\) is negatively-biased (particularly for small tree
sizes but also consistently across most of its range; Fig. 1 black line)
and standard deviations of \(\hat{\lambda}\) are negatively associated
with tree size. For tress of 128 species or less, \(\hat{\lambda}\) are
quite variable, except for cases when \(\lambda\) is near or equal to
\(1\). Taken together these results reveal that \(\hat\lambda\) is a
biased statistic that inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supporting Information) reveal that incorporating
\(\hat\lambda\) in PGLS ANOVA and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, and thus
\(\lambda\) should not be used for comparing phylogenetic signal across
datasets.

\hypertarget{kappa-k-estimates-of-phylogenetic-signal-are-more-reliable}{%
\subsection{\texorpdfstring{Kappa (\emph{K}) estimates of phylogenetic
signal are more
reliable}{Kappa (K) estimates of phylogenetic signal are more reliable}}\label{kappa-k-estimates-of-phylogenetic-signal-are-more-reliable}}

Simulation results demonstrate that \emph{K} displays better statistical
properties. First, as expected, mean values of \emph{K} increase with
increasing signal (\(\lambda\)) irrespective of tree size, albeit
non-linearly (Fig. 2 black line). Second, the standard deviation of
\emph{K} is consistent across tree sizes (Fig. 2 red line), and while it
increases with \(\lambda\), it is always less than the mean (low
coefficient of variation). This finding is perhaps unsurprising, as
\emph{K} is lower-bounded by 0, and is never large for small values of
\(\lambda\). Importantly, \emph{K} is normally distributed across the
range of input \(\lambda\); a consistent pattern regardless of tree size
(Fig. 2 blue line). This differs from results of (Münkemüller et al.,
2012), where the skewing appears to be due to combining random values
generated independently, rather than being a property of \emph{K}
itself. Overall, these findings reveal that while \emph{K} is more
reliable as an estimate of phylogenetic signal, the non-linear scaling
with input signal implies that it should not be considered an effect
size that measures the strength of phylogenetic signal on a common scale
for comparison across datasets.

\hypertarget{effect-sizes-from-k-z_k-better-characterize-phylogenetic-signal}{%
\subsection{\texorpdfstring{Effect sizes from \emph{K} (\(Z_K\)) better
characterize phylogenetic
signal}{Effect sizes from K (Z\_K) better characterize phylogenetic signal}}\label{effect-sizes-from-k-z_k-better-characterize-phylogenetic-signal}}

To measure the strength of phylogenetic signal on a common scale, we
propose effect sizes (Z-scores) for both \(\lambda\) and \emph{K}.
Statistically, a standardized effect size may be found as:

\begin{align}
    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
\end{align}

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (Glass, 1976; Cohen, 1988; Rosenthal, 1994).
Typically, \(\theta_{obs}\) and \(\sigma_\theta\) are estimated from the
data, while \(E(\theta)\) is obtained from the distribution of
\(\theta\) derived from parametric theory. However, recent advances in
resampling theory (Collyer, Sekora, \& Adams, 2015; Adams \& Collyer,
2016, 2019a; Collyer \& Adams, 2018) have shown that \(E(\theta)\) and
\(\sigma_\theta\) may also be obtained from an empirical sampling
distribution of \(\theta\) simulated from permutation procedures.
\hfill\break

Formalizing the suggestion of Adams and Collyer (Adams \& Collyer,
2019b), an effect size for \emph{K} may be found as:

\begin{align}
    Z_K=\frac{K_{obs}-\hat\mu_{K}}{\hat\sigma_{K}},
\end{align}

where \(K_{obs}\) is the observed phylogenetic signal, and \(\hat\mu_K\)
and \(\hat\sigma_K\) are the mean and standard deviation of the
empirical sampling distribution of \emph{K} obtained via permutation.
The empirical sampling distribution of \emph{K} can be first transformed
via a Box-Cox transformation to better adhere to the assumption of
normality. \hfill\break

For \(\lambda\), deriving an effect size is more challenging, as
\(\lambda\) does not have a sampling distribution from which the
standard error and confidence intervals may be obtained, and estimates
from the Hessian matrix from PGLS are unreliable (Boettiger et al.,
2012). Confidence intervals are therefore generated for the values of
\(\lambda\) that intersect the log-likehihood profile for corresponding
percentiles of the \(\chi^2\) distribution used to compare the putative
model to a null model with \(\lambda = 0\) (Orme et al., 2013). Thus, an
effect size for \(\lambda\) may be found as:

\begin{align}
   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
\end{align}

where, \(\hat{\lambda}\) is the maximized likelihood value of
\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
statistic for the value. \hfill\break

Simulations reveal that both \(Z_{\lambda}\) and \(Z_K\) are associated
with input phylogenetic signal (\(\lambda\)), indicating that both
statistics capture the observed signal (Fig. 3). However, effect sizes
from \(\hat{\lambda}\) made little sense, as they are more strongly
associated with tree size than they are with the actual phylogenetic
signal in the data (Fig. 3). By contrast, \(Z_K\) is much more
consistent across tree sizes, and increases more linearly with
increasing levels of phylogenetic signal. Additionally, \(Z_K\) exhibits
a much stronger association with phylogenetic signal strength as
compared to tree size (Fig. 3), and its standard deviation is more
consistent, implying similar levels of precision across the range of
input signal (Supporting Information). Thus, \(Z_K\) is a more reliable
measure of the strength of phylogenetic signal, and may be used to
compare levels of phylogenetic signal across datasets.

\hypertarget{a-test-statistic-hatz_12-allows-meaningful-comparisons-across-datasets}{%
\subsection{\texorpdfstring{A test statistic (\(\hat{Z}_{12}\)) allows
meaningful comparisons across
datasets}{A test statistic (\textbackslash hat\{Z\}\_\{12\}) allows meaningful comparisons across datasets}}\label{a-test-statistic-hatz_12-allows-meaningful-comparisons-across-datasets}}

To statistically compare the strength of phylogenetic signal across
datasets we propose a two-sample test statistic (\(\hat{Z}_{12}\)).
Based on statistical theory, a two-sample test statistic may be
calculated as:

\begin{align}
  \hat{Z}_{12}=\frac{\lvert{(K_{1}-\hat\mu_{K_1})-(K_{2}-\hat\mu_{K_2})}\rvert}{\sqrt{\hat\sigma^2_{K_1}+\hat\sigma^2_{K_2}}} 
\end{align}

where \(K_1\), \(K_2\), \(\hat\mu_{K_1}\), \(\hat\mu_{K_2}\),
\(\hat\sigma_{K_1}\), and \(\hat\sigma_{K_2}\) are as defined above.
Estimates of significance of \(\hat{Z}_{12}\) may be obtained from a
standard normal distribution, or permutation. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (Adams \& Collyer, 2016, 2019a). \hfill\break

We evaluated the type I error and false discovery rates of tests based
on \(\hat{Z}_{12}\) using a procedure similar to that above. Simulations
were performed using pure-birth trees of different sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
combination of \(n\) and \(\lambda\), a total of 100 traits were
simulated on pure-birth trees (50 traits per tree) under a Brownian
motion model of evolution. Next, the phylogenetic signal between traits
was compared in pairwise fashion for all combinations of traits, using
\(\hat{Z}_{12}\). The proportion of significant results provided an
estimate of the type I error or false discovery rate. Specifically, type
I error was evaluated when traits were simulated using no input
phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, false discovery
rates were evaluated for data simulated with some known, but equal,
level of phylogenetic signal (i.e., \(\lambda > 0\)). \hfill\break

Tests revealed that across simulation conditions, the average type I
error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig. 4). Likewise,
false discovery rates were also low, and were at or below the nominal
5\% level (Fig. 4). Importantly, the statistical performance of
\(\hat{Z}_{12}\) appears unaffected by phylogeny size. Overall, these
results reveal that tests based on \(\hat{Z}_{12}\) have acceptable type
I error and false discovery rates. Therefore, \(\hat{Z}_{12}\) is an
appropriate statistic for comparing the degree of phylogenetic signal
across traits.

\hypertarget{empirical-example}{%
\section{Empirical example}\label{empirical-example}}

To demonstrate the utility of \(\hat{Z}_{12}\), we compared \(Z_K\) for
two ecologically-relevant traits in plethodontid salamander (Fig. 5):
surface area to volume ratios (SA:V) and relative body width
(\(\frac{BW}{SVL}\)) (Baken \& Adams, 2019; Baken, Mellenthin, \& Adams,
2020). For this example, surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
individuals of 305 species, from which species means were obtained
(Baken \& Adams, 2019; Baken et al., 2020). A time-dated molecular
phylogeny for the group (Bonett \& Blair, 2017) was pruned to match the
species in the phenotypic dataset. The phylogenetic signal in each trait
was then characterized using \emph{K}, which was converted to its effect
size (\(Z_K\)) using \texttt{geomorph} 3.3.1 (Adams \& Otárola-Castillo,
2013; Adams, Collyer, \& Kaliontzopoulou, 2020), and routines by the
authors (\textbf{to be incorporated in \texttt{geomorph} upon
acceptance}). \hfill\break

While both traits contained significant phylogenetic signal, tests based
on \(\hat{Z}_{12}\) revealed that the degree of phylogenetic signal was
significantly stronger in SA:V (\(\hat{Z}_{12}=16.51\); \(P< 0.00001\):
Fig. 5). Biologically, this observation may be interpreted by the fact
that the tropical species -- which form a monophyletic group within
plethodontids -- display greater variation in SA:V, which covaries with
disparity in their climatic niches (Baken et al., 2020). Thus, greater
phylogenetic signal in SA:V is to be expected.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (e.g., Abouheif, 1999; Gittleman \& Kot,
1990; Pagel, 1999; Blomberg et al., 2003; Adams, 2014a), methods that
explicitly measure the strength of phylogenetic signal, or facilitate
comparisons among datasets, have remained underdeveloped. We evaluated
the precision of one common measure, Pagel's \(\lambda\), and explored
its efficacy for characterizing the strength of phylogenetic signal in
phenotypic data. Using computer simulations, we found that \(\lambda\)
behaves as a Bernoulli random variable, with estimates that are
increasingly skewed at larger and smaller input levels of phylogenetic
signal. Further, the precision of \(\lambda\) in estimating actual
levels of phylogenetic signal varies with both tree size (see also ref.
Boettiger et al. (2012)) and input levels of phylogentetic signal. From
these findings we conclude that \(\lambda\) is not a reliable indicator
of the observed strength of phylogenetic signal in phenoytpic datasets,
and should not be used as an effect size for comparing the degree of
phylogenetic signal between datasets. \hfill\break

As an alternative, we described a standardized effect size (\(Z\)) for
assessing the strength of phylogenetic signal. \(Z\) expresses the
magnitude of phylogenetic signal as a standard normal deviate, which is
easily interpretable as the strength of phylogenetic signal relative to
the mean. We applied this concept to both \(\lambda\) and \emph{K}, and
found that \(Z_K\) was a better estimate of the strength of phylogenetic
signal in phenotypic data. First, values of \(Z_K\) more accurately
tracked known changes in the magnitude of phylogenetic signal, as
demonstrated by the near linear relationship between \(Z_K\) and input
signal. Additionally, the precision of \(Z_K\) was more consistent
across the range of input levels of phylogenetic signal (Fig S1;
Supporting Information). Thus, \(Z_K\) is a more reliable measure of the
relative strength of phylogenetic signal, and places that effect on a
common and comparable scale. We therefore recommend that future studies
interested in evaluating the strength of phylogenetic signal incorporate
\(Z_K\) as a statistical measure of this effect. \hfill\break

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\emph{K} (e.g., Blomberg et al., 2003). However, because \emph{K} does
not scale linearly with input levels of phylogenetic signal (Fig. 2),
and its variance increases with increasing strength of phylogenetic
signal (Diniz-Filho et al., 2012; Münkemüller et al., 2012), it should
not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\emph{K} to \(Z_K\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (sensu Hedges \& Olkin, 1985; Glass, 1976; Arnqvist \&
Wooster, 1995), where summary statistics across datasets are converted
to standardized effect sizes for subsequent ``higher order'' statistical
summaries or comparisons. As such, our approach enables evolutionary
biologists to quantitatively examine the relative strength of
phylogenetic signal across a wide range of phenotypic traits, and thus
opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.
\hfill\break

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_K\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (Sokal \& Rohlf,
2012). Thus, \(Z_K\) represents the strength of phylogenetic signal on a
common and comparable scale -- measured in standard deviations --
regardless of the initial units and original scale of the phenotypic
variables under investigation. This means that the strength of
phylogenetic signal may be compared across datasets for continuous
phenotypic traits measured in different units and scale, because those
units have been standardized through their conversion to \(Z_K\). For
example, our approach could be utilized to determine whether the
strength of phylogenetic signal (say, in response to ecological
differentiation) is stronger in morphological traits (linear traits:
\(mm\)), physiological traits (metabolic rate: \(\frac{O^2}{min}\)), or
behavioral traits (aggression: \(\frac{\#{displays}}{second}\)). In
fact, our empirical example provided just such a comparison, as SA:V is
represented in \(mm^{-1}\) while relative body size is a unitless ratio
(\(\frac{BW}{SVL}\)). Additionally, our method is capable of comparing
the strength of phylogenetic signal in traits of different
dimensionality, as estimates of phylogenetic signal using \emph{K} have
been generalized for multivariate data (\(\kappa_{mult}\): see Adams,
2014a). Furthermore, tests based on \(\hat{Z}_{12}\) may be utilized for
comparing the strength of phylogenetic signal among datasets containing
a different number of variables, and even for phenotypes obtained from
species in different lineages, because their phylogenetic
non-independence and observed variation are taken into account in the
generation of the empirical sampling distribution via permutation.
\hfill\break

This study is not the first to compare \(\lambda\) and \emph{K} for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \emph{K} values are consistent with those found in
the simulations performed by Münkemüller et al.~(2012), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \emph{K} at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(Münkemüller et al., 2012)). Recent work by Molina-Venegas and Rodríguez
(2017) found that \emph{K} but not \(\lambda\) tended to inflate the
estimate of phylogenetic signal, leading to moderate type I and type II
biases, if polytomic chronograms were used. Their work more thoroughly
addressed previous observations of inflated \emph{K} for incompletely
resolved phylogenetic trees (Davies, Kraft, Salamin, \& Wolkovich, 2012;
Münkemüller et al., 2012). An interesting question is whether an
inflated \emph{K} value leads to an inflated \(Z_K\) or does a tendency
of a particular tree to inflate estimates of \emph{K} also inflate the
values in random permutations of a test, in which case \(Z_K\) is robust
to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
to have 20\% collapsed nodes, per the technique of Molina-Venegas and
Rodríguez (2017), and found results were consistent (Supporting
Information). This confirms that any tendency of incompletely resolved
trees to inflate \emph{K} as a descriptive statistic does not inflate
\(Z_K\) as an effect size. Furthermore, because comparison of effect
sizes in a test is a comparison of locations of observed values in their
sampling distributions, which would shift concomitantly because of this
tendency, the \(Z_{12}\) test statistic in equation 4 appears to be
robust in spite of unresolved trees. \hfill\break

Phylogenetic signal can be thought of as both an attribute to be
measured in the data and a parameter that can be tuned to account for
the phylogenetic non-independence among observations, for analysis of
the data. As such, \(\lambda\) is appealing, as a statistic that
potentially fulfills both roles. However, the inability to estimate
phylogenetic signal with \(\lambda\) for data simulated with known
phylogenetic signal is troublesome, and we recommend evolutionary
biologists refrain from viewing it as a statistic to describe the amount
of phylogenetic signal in the data. Interestingly, \emph{K} -- when
standardized to an effect size \(Z_K\) -- is a better statistic for
measuring the amount of phylogenetic signal in data simulated with
respect to known levels of \(\lambda\). Although \(\lambda\) might be
viewed as an important parameter for modifying the the conditional
estimation of linear model coefficients with respect to phylogeny, it is
neither a statistic that has meaningful comparative value as a measure
of phylogenetic signal nor a statistic that lends itself well to
reliable calculation of a test statistic. By contrast, \emph{K} has been
shown here to be a reliable statistic, but only when standardized by the
mean and standard deviation of its empirical sampling distribution
(i.e., when converted to the effect size, \(Z_K\)). Because one has
control over the number of permutations used in analysis, one can be
assured with many permutations that the empirical sampling distribution
is representative of true probability distributions (Adams \& Collyer,
2018). Given the greater consistency in estimates of \(Z_K\) across tree
sizes and input signal, it is difficult to imagine a hypothesis test
that can improve equation 4 for efficiently comparing phylogenetic
signal for different traits, different trees, or a combination of both.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
\setlength{\parskip}{8pt} \noindent

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abouheif1999}{}%
Abouheif, E. (1999). A method for testing the assumption of phylogenetic
independence in comparative data. \emph{Evolutionary Ecology Research},
\emph{1}, 895--909. Journal Article.

\leavevmode\hypertarget{ref-Adams2014a}{}%
Adams, D. C. (2014a). A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. \emph{Systematic Biology}, \emph{63}, 685--697.

\leavevmode\hypertarget{ref-Adams2014b}{}%
Adams, D. C. (2014b). A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution}, \emph{68}, 2675--2688.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
Adams, D. C., \& Collyer, M. L. (2016). On the comparison of the
strength of morphological integration across morphometric datasets.
\emph{Evolution}, \emph{70}, 2623--2631. Journal Article.

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution}, \emph{72}(6), 1204--1215.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
Adams, D. C., \& Collyer, M. L. (2019a). Comparing the strength of
modular signal, and evaluating alternative modular hypotheses, using
covariance ratio effect sizes with morphometric data. \emph{Evolution},
\emph{73}(12), 2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
Adams, D. C., \& Collyer, M. L. (2019b). Phylogenetic comparative
methods and the evolution of multivariate phenotypes. \emph{Annual
Review of Ecology, Evolution, and Systematics}, \emph{50}, 405--425.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
Adams, D. C., Collyer, M. L., \& Kaliontzopoulou, A. (2020). Geomorph:
Software for geometric morphometric analyses. R package version 3.3.1.
Retrieved from \url{https://cran.r-project.org/package=geomorph}

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
Adams, D. C., \& Otárola-Castillo, E. (2013). Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution}, \emph{4}, 393--399.

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
Arnqvist, G., \& Wooster, D. (1995). Meta-analysis: Synthesizing
research findings in ecology and evolution. \emph{Trends in Ecology and
Evolution}, \emph{10}, 236--240.

\leavevmode\hypertarget{ref-Baken2019}{}%
Baken, E. K., \& Adams, D. C. (2019). Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution}, \emph{9}(12), 7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
Baken, E. K., Mellenthin, L. E., \& Adams, D. C. (2020). Macroevolution
of desiccation‐related morphology in plethodontid salamanders as
inferred from a novel surface area to volume ratio estimation approach.
\emph{Evolution}, \emph{74}, 476--486.

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
Beaulieu, J. M., Jhwueng, D. C., Boettiger, C., \& O'Meara, B. C.
(2012). Modeling stabilizing selection: Expanding the ornstein-uhlenbeck
model of adaptive evolution. \emph{Evolution}, \emph{66}, 2369--2383.
Journal Article.

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
Blomberg, S. P., Garland, T., \& Ives, A. R. (2003). Testing for
phylogenetic signal in comparative data: Behavioral traits are more
labile. \emph{Evolution}, \emph{57}, 717--745.

\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
Boettiger, C., Coop, G., \& Ralph, P. (2012). Is your phylogeny
informative? Measuring the power of comparative methods.
\emph{Evolution}, \emph{67}, 2240--2251. Journal Article.

\leavevmode\hypertarget{ref-Bonett2017}{}%
Bonett, R. M., \& Blair, A. L. (2017). Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, U.S.A.}, \emph{114}, 9936--9941.
doi:\href{https://doi.org/10.1073/pnas.1703877114}{10.1073/pnas.1703877114}

\leavevmode\hypertarget{ref-Bose2019}{}%
Bose, R., Ramesh, B. R., Pélissier, R., \& Munoz, F. (2019).
Phylogenetic diversity in the western ghats biodiversity hotspot
reflects environmental filtering and past niche diversification of
trees. \emph{Journal of Biogeography}, \emph{46}(1), 145--157.

\leavevmode\hypertarget{ref-Cohen1988}{}%
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral
sciences}. Routledge.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
Collyer, M. L., \& Adams, D. C. (2018). RRPP: An r package for fitting
linear models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution}, \emph{9}, 1772--1779.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
Collyer, M. L., Sekora, D. J., \& Adams, D. C. (2015). A method for
analysis of phenotypic change for phenotypes described by
high-dimensional data. \emph{Heredity}, \emph{115}, 357--365.

\leavevmode\hypertarget{ref-Cooper2010}{}%
Cooper, N., Jetz, W., \& Freckleton, R. P. (2010). Phylogenetic
comparative approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology}, \emph{23}(12), 2529--2539.

\leavevmode\hypertarget{ref-Davies2012}{}%
Davies, T. J., Kraft, N. J., Salamin, N., \& Wolkovich, E. M. (2012).
Incompletely resolved phylogenetic trees inflate estimates of
phylogenetic conservatism. \emph{Ecology}, \emph{93}(2), 242--247.

\leavevmode\hypertarget{ref-DeMeester2019}{}%
De Meester, G., Huyghe, K., \& Van Damme, R. (2019). Brain size, ecology
and sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society}, \emph{126}(3), 381--391.

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
Diniz-Filho, J. A. F., Santos, T., Rangel, T. F., \& Bini, L. M. (2012).
A comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. \emph{Genetics and Molecular Biology},
\emph{35}(3), 673--679.

\leavevmode\hypertarget{ref-Felsenstein1985}{}%
Felsenstein, J. (1985). Phylogenies and the comparative method.
\emph{American Naturalist}, \emph{125}(1), 1--15. Journal Articl.

\leavevmode\hypertarget{ref-Forbes2011}{}%
Forbes, C., Evans, M., Hastings, N., \& Peacock, B. (2011).
\emph{Statistical distributions}. John Wiley \& Sons.

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
Freckleton, R. P., Harvey, P. H., \& Pagel, M. (2002). Phylogenetic
analysis and comparative data: A test and review of evidence.
\emph{American Naturalist}, \emph{160}, 712--726.

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
Garland, T. J., \& Ives, A. R. (2000). Using the past to predict the
present: Confidence intervals for regression equations in phylogenetic
comparative methods. \emph{American Naturalist}, \emph{155}, 346--364.

\leavevmode\hypertarget{ref-Gittleman1990}{}%
Gittleman, J. L., \& Kot, M. (1990). Adaptation: Statistics and a null
model for estimating phylogenetic effects. \emph{Systematic Zoology},
\emph{39}(3), 227--241.

\leavevmode\hypertarget{ref-Glass1976}{}%
Glass, G. V. (1976). Primary, secondary, and meta-analysis of research.
\emph{Educational Researcher}, \emph{5}, 3--8.

\leavevmode\hypertarget{ref-Grafen1989}{}%
Grafen, A. (1989). The phylogenetic regression. \emph{Philosophical
Transactions of the Royal Society of London B, Biological Sciences},
\emph{326}, 119--157.

\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
Harvey, P. H., \& Pagel, M. D. (1991). \emph{The comparative method in
evolutionary biology}. Book, Oxford: Oxford University Press.

\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
Hedges, L. V., \& Olkin, I. (1985). \emph{Statistical methods for
meta-analysis}. Elsevier.

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
Klingenberg, C. P., \& Gidaszewski, N. A. (2010). Testing and
quantifying phylogenetic signals and homoplasy in morphometric data.
\emph{Systematic Biology}, \emph{59}(3), 245--261.

\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
Martins, E. P., \& Hansen, T. F. (1997). Phylogenies and the comparative
method: A general approach to incorporating phylogenetic information
into the analysis of interspecific data. \emph{American Naturalist},
\emph{149}, 646--667.

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
Molina-Venegas, R., \& Rodríguez, M. A. (2017). Revisiting phylogenetic
signal; strong or negligible impacts of polytomies and branch length
information? \emph{BMC Evolutionary Biology}, \emph{17}(1), 53.

\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
Münkemüller, T., Lavergne, S., Bzeznik, B., Dray, S., Jombart, T.,
Schiffers, K., \& Thuiller, W. (2012). How to measure and test
phylogenetic signal. \emph{Methods in Ecology and Evolution}, \emph{3},
743--756. Journal Article.

\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
O'Meara, B. C., Ane, C., Sanderson, M. J., \& Wainwright, P. C. (2006).
Testing for different rates of continuous trait evolution using
likelihood. \emph{Evolution}, \emph{60}, 922--933. Journal Article.

\leavevmode\hypertarget{ref-Orme2013}{}%
Orme, D., Freckleton, R. P., Thomas, G. H., Petzoldt, T., Fritz, S. A.,
\& Isaac, N. (2013). CAPER: Comparative analyses of phylogenetics and
evolution in r. \emph{Methods in Ecology and Evolution}, \emph{3},
145--151.

\leavevmode\hypertarget{ref-Pagel1999}{}%
Pagel, M. D. (1999). Inferring the historical patterns of biological
evolution. \emph{Nature}, \emph{401}, 877--884. Journal Article.

\leavevmode\hypertarget{ref-Pavoine2012}{}%
Pavoine, S., \& Ricotta, C. (2012). Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics.
\emph{Evolution: International Journal of Organic Evolution},
\emph{67}(3), 828--840.

\leavevmode\hypertarget{ref-Pintanel2019}{}%
Pintanel, P., Tejedo, M., Ron, S. R., Llorente, G. A., \& Merino-Viteri,
A. (2019). Elevational and microclimatic drivers of thermal tolerance in
andean pristimantis frogs. \emph{Journal of Biogeography}, \emph{46}(8),
1664--1675.

\leavevmode\hypertarget{ref-Revell2010}{}%
Revell, L. J. (2010). Phylogenetic signal and linear regression on
species data. \emph{Methods in Ecology and Evolution}, \emph{1},
319--329. Journal Article.

\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
Revell, L. J., Harmon, L. J., \& Collar, D. C. (2008). Phylogenetic
signal, evolutionary process, and rate. \emph{Systematic Biology},
\emph{57}, 591--601. Journal Article.

\leavevmode\hypertarget{ref-Rohlf2001}{}%
Rohlf, F. J. (2001). Comparative methods for the analysis of continuous
variables: Geometric interpretations. \emph{Evolution}, \emph{55},
2143--2160. Journal Article.

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
Rosenthal, R. (1994). The handbook of research synthesis. In L. V.
Cooper H Hedges (Ed.) (pp. 231--244). Russell Sage Foundation.

\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
Sokal, R. R., \& Rohlf, F. J. (2012). \emph{Biometry} (4th ed.). San
Francisco: W.H. Freeman \& Co.

\leavevmode\hypertarget{ref-Su2019}{}%
Su, G., Villéger, S., \& Brosse, S. (2019). Morphological diversity of
freshwater fishes differs between realms, but morphologically extreme
species are widespread. \emph{Global Ecology and Biogeography},
\emph{28}(2), 211--221.

\leavevmode\hypertarget{ref-Vandelook2019}{}%
Vandelook, F., Janssens, S., Gijbels, P., Fischer, E., Van den Ende, W.,
Honnay, O., \& Abrahamczyk, S. (2019). Nectar traits differ between
pollination syndromes in balsaminaceae. \emph{Annals of Botany},
\emph{124}(2), 269--279.

\newpage

\hypertarget{figures}{%
\section{Figures}\label{figures}}

Figure 1. Response of Pagel's \(\lambda\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\lambda\). At each input level, the dark
black line represents the empirically derived expected value (mean) of
\(\hat\lambda\), the red line is the standard deviation of
\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions). \hfill\break

Figure 2. Response of Blomberg's \(K\) to increasing strength of
Brownian motion. At each input level, the black line represents the
empirically derived expected value (mean) of \(K\), the red line is the
standard deviation of \emph{K}, and the blue line is Shapiro Wilks
statistic of \(K\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions). \hfill\break

Figure 3. Response of effect sizes \(Z_\lambda\) and \(Z_K\) to
increasing strength of Brownian motion. Means from simulation runs are
shown for comparative ease. Individual values from each simulation run
are available in Supporting Information. \hfill\break

Figure 4. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)).

Figure 5. (A) Linear measures for relative body size, and regions of the
body used to estimate surface area to volume (SA:V) ratios. (B)
Permutation distributions of phylogenetic signal for SA:V and
\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_K\)) for SA:V and \(\frac{BW}{SVL}\), with their 95\%
confidence intervals (CI not standardized by \(\sqrt(n)\)). \hfill\break

\newpage

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/fig.1} \caption{Response of Pagel's $\lambda$ to increasing strength of Brownian motion. Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$. At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions). }\label{fig:unnamed-chunk-3}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/fig.2} \caption{Response of Blomberg's $K$ to increasing strength of Brownian motion.  At each input level, the black line represents the empirically derived expected value (mean) of $K$, the red line is the standard deviation of $K$, and the blue line is Shapiro Wilks statistic of *K* ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).}\label{fig:unnamed-chunk-4}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/fig.3} \caption{Response of effect sizes $Z_\lambda$ and $Z_K$ to increasing strength of Brownian motion.  Means from simulation runs are shown for comparative ease. Individual values from each simulation run are available in Supporting Information.}\label{fig:unnamed-chunk-5}
\end{figure}

\newpage

\begin{figure}
\includegraphics[width=28.51in,height=0.75\textheight]{Figs/fig.4} \caption{Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).}\label{fig:unnamed-chunk-6}
\end{figure}

\newpage

\begin{figure}
\includegraphics[width=0.9\linewidth]{Figs/fig.5} \caption{(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios. (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars. (C) Effect sizes ($Z_K$) for SA:V and $\frac{BW}{SVL}$, with their 95\% confidence intervals (CI not standardized by $\sqrt(n)$).}\label{fig:unnamed-chunk-7}
\end{figure}

\end{document}

### 2020-09-28 Manuscript/Figs/SI.Rmd
MEE version

--- 2020-09-17 Manuscript/SupportingInfo.Rmd
@@
 ---
-title:
-
-author:
+title: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal
+author: Dean C. Adams, Erica K. Baken, and Michael L. Collyer
 
@@
 
-csl: pnas.csl
+csl: methods-in-ecology-and-evolution.csl
 bibliography: Lambda-Refs.bib
@@
 
-```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("SI-Title.png")  
-```
-\hfill\break
-\hfill\break
-
-## Supplementary Information for 
-\hfill\break
-
-A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal \hfill\break
-\hfill\break
-Dean C. Adams, Erica K. Baken, and Michael L. Collyer \hfill\break
-
-\hfill\break
-Dean C. Adams\hfill\break
-Email:  dcadams@iastate.edu\hfill\break
-
-## This PDF file includes:
-\hfill\break
-- Supplementary text 
-- Figures S1 to S20 
-- SI References
-
-\newpage
+# Additional Simulation Results
 
@@
 Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$).
-For each combination of $n$ and $\lambda$, 20 pure-birth trees were generated and 50 traits were simulated on each phylogeny under a Brownian motion model of evolution.
+For each combination of $n$ and $\lambda$, 2 pure-birth trees were generated and 50 traits were simulated on each phylogeny under a Brownian motion model of evolution.
 Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$.

### 2020-09-28 Manuscript/Figs/SI.tex
MEE version

--- 2020-09-17 Manuscript/SupportingInfo.tex
@@
 \hypersetup{
+  pdftitle={A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
+  pdfauthor={Dean C. Adams, Erica K. Baken, and Michael L. Collyer},
   hidelinks,
@@
 
-\author{}
-\date{\vspace{-2.5em}}
+\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
+Phylogenetic Signal}
+\author{Dean C. Adams, Erica K. Baken, and Michael L. Collyer}
+\date{}
 
 \begin{document}
+\maketitle
 
-\includegraphics[width=0.95\linewidth]{SI-Title} \hfill\break
-\hfill\break
-
-\hypertarget{supplementary-information-for}{%
-\subsection{Supplementary Information
-for}\label{supplementary-information-for}}
-
-\hfill\break
-
-A Standardized Effect Size for Evaluating and Comparing the Strength of
-Phylogenetic Signal \hfill\break \hfill\break Dean C. Adams, Erica K.
-Baken, and Michael L. Collyer \hfill\break
-
-\hfill\break
-
-Dean C. Adams\hfill\break Email:
-\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}\hfill\break
-
-\hypertarget{this-pdf-file-includes}{%
-\subsection{This PDF file includes:}\label{this-pdf-file-includes}}
-
-\hfill\break
-
-\begin{itemize}
-\tightlist
-\item
-  Supplementary text
-\item
-  Figures S1 to S20
-\item
-  SI References
-\end{itemize}
-
-\newpage
+\hypertarget{additional-simulation-results}{%
+\section{Additional Simulation
+Results}\label{additional-simulation-results}}
 
@@
 of phylogenetic signal by adjusting pure-birth trees to have 20\%
-collapsed nodes, following the procedures of (1). Results from
-simulations on phylogenies containing polytomies mirrored those found on
-fully resolved trees. For \(\hat{\lambda}\), the mean value increased
-with increasing input signal, but was negatively biased, and was less
-than the input value across most of its range (Fig. S8: black line).
-Additionally, the precision of \(\hat{\lambda}\) varied with differing
-input levels, with the greatest variation found at intermediate values
-of \(\lambda\) (Fig. S8 red line). Finally, the distribution of
-\(\hat{\lambda}\) was not normal, and became more skewed at more extreme
-values of \(\lambda\) (Fig. S8 blue line). \hfill\break
+collapsed nodes, following the procedures of (Molina-Venegas \&
+Rodríguez, 2017). Results from simulations on phylogenies containing
+polytomies mirrored those found on fully resolved trees. For
+\(\hat{\lambda}\), the mean value increased with increasing input
+signal, but was negatively biased, and was less than the input value
+across most of its range (Fig. S8: black line). Additionally, the
+precision of \(\hat{\lambda}\) varied with differing input levels, with
+the greatest variation found at intermediate values of \(\lambda\) (Fig.
+S8 red line). Finally, the distribution of \(\hat{\lambda}\) was not
+normal, and became more skewed at more extreme values of \(\lambda\)
+(Fig. S8 blue line). \hfill\break
 
@@
 S19 \& S20), though power was reduced at smaller sample sizes in
-phylogenetic ANOVA, mirroring prior findings of (2). These results
-generally confirmed earlier findings of (3) for phylogenetic regression,
-and extend them to the case of phylogenetic ANOVA.
+phylogenetic ANOVA, mirroring prior findings of (Adams \& Collyer,
+2018). These results generally confirmed earlier findings of (Revell,
+2010) for phylogenetic regression, and extend them to the case of
+phylogenetic ANOVA.
 
@@
 phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
-combination of \(n\) and \(\lambda\), 20 pure-birth trees were generated
+combination of \(n\) and \(\lambda\), 2 pure-birth trees were generated
 and 50 traits were simulated on each phylogeny under a Brownian motion
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
+Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic anova: Group-clade
+aggregation, biological challenges, and a refined permutation procedure.
+\emph{Evolution}, \emph{72}(6), 1204--1215.
+
 \leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
-1. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic signal;
-strong or negligible impacts of polytomies and branch length
-information? \emph{BMC evolutionary biology} 17(1):53.
-
-\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
-2. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
-aggregation, biological challenges, and a refined permutation procedure.
-\emph{Evolution} 72(6):1204--1215.
+Molina-Venegas, R., \& Rodríguez, M. A. (2017). Revisiting phylogenetic
+signal; strong or negligible impacts of polytomies and branch length
+information? \emph{BMC Evolutionary Biology}, \emph{17}(1), 53.
 
 \leavevmode\hypertarget{ref-Revell2010}{}%
-3. Revell LJ (2010) Phylogenetic signal and linear regression on species
-data. \emph{Methods in Ecology and Evolution} 1:319--329.
+Revell, L. J. (2010). Phylogenetic signal and linear regression on
+species data. \emph{Methods in Ecology and Evolution}, \emph{1},
+319--329. Journal Article.
 

### 2020-09-28 Manuscript/OldVsns/PNAS-Version/PNAS-Adams-Submitted.tex
MEE version

--- 2020-09-17 Manuscript/PNAS-Adams-Submitted.tex
@@
 degree of phylogenetic signal was significantly stronger in SA:V
-(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
+(\(\hat{Z}_{12}=16.51\); \(P< 0.00001\): Fig. 4). Biologically, this
 observation may be interpreted by the fact that the tropical species --

### 2020-09-28 Manuscript/OldVsns/PNAS-Version/SupportingInfo.Rmd
MEE version

--- 2020-09-17 Manuscript/SupportingInfo.Rmd
@@
 - Supplementary text 
-- Figures S1 to S20 
+- Figures S1 to S21 
 - SI References

### 2020-09-28 Manuscript/OldVsns/PNAS-Version/SupportingInfo.tex
MEE version

--- 2020-09-17 Manuscript/SupportingInfo.tex
@@
 \item
-  Figures S1 to S20
+  Figures S1 to S21
 \item

### 2020-09-28 Manuscript/SI.Rmd
MEE version

--- 2020-09-28 Manuscript/Figs/SI.Rmd
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S1.png")  
+include_graphics("Figs/fig.S1.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S2.png")  
+include_graphics("Figs/fig.S2.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S3.png")  
+include_graphics("Figs/fig.S3.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S4.png")  
+include_graphics("Figs/fig.S4.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S5.png")  
+include_graphics("Figs/fig.S5.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S6.png")  
+include_graphics("Figs/fig.S6.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S7.png")  
+include_graphics("Figs/fig.S7.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S8.png")  
+include_graphics("Figs/fig.S8.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S9.png")  
+include_graphics("Figs/fig.S9.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S10.png")  
+include_graphics("Figs/fig.S10.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S11.png")  
+include_graphics("Figs/fig.S11.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S12.png")  
+include_graphics("Figs/fig.S12.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S13.png")  
+include_graphics("Figs/fig.S13.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S14.png")  
+include_graphics("Figs/fig.S14.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S15.png")  
+include_graphics("Figs/fig.S15.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S16.png")  
+include_graphics("Figs/fig.S16.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S17.png")  
+include_graphics("Figs/fig.S17.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S18.png")  
+include_graphics("Figs/fig.S18.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S19.png")  
+include_graphics("Figs/fig.S19.png")  
 ```
@@
 ```{r, echo = FALSE, out.width="95%", align= "center"}
-include_graphics("fig.S20.png")  
+include_graphics("Figs/fig.S20.png")  
 ```
@@
 ```{r, echo = FALSE, out.height="50%", align= "center"}
-include_graphics("fig.S21.png")  
+include_graphics("Figs/fig.S21.png")  
 ```
 
-\textbf{Figure S21}.
-Type I error rates and model misspecification for $\hat{Z}_{12}$.
-Type I error is found on the far-left of the plot ($\lambda = 0$).
-The solid black line represents the average type I error and false discovery rate across all levels of input signal.
+\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
 

### 2020-09-28 Data_Analyses/updated_simulations/simulation.analyses.PB.type1err-DCA.R

@@ -4,10 +4,10 @@ library(geomorph)
 
 ##DCA: adjusted BC to not use observed in calculating the transformation,
 
-treesizes <- 2^(8) # 2^(5:7) done, 2^8 being run by EKB, still need to run 2^(9:10)
+treesizes <- 2^(10) # 2^(5:7) done, 2^8 being run by EKB, still need to run 2^(9:10)
 lambdas <- seq(0, 1, 0.05)
 ntrees <- 2
-ndatasets <- 50
+ndatasets <- 25  #was 50, put back
 physig_iter <- 999
 
 #### functions from develop branch ######
@@ -199,31 +199,33 @@ rownames(t1err_mat) <- treesizes
 errormeans <- colMeans(t1err_mat)
 cis <- lapply(1:length(lambdas), function(i) qnorm(0.975)*(sd(t1err_mat[,i])/sqrt(6)))
 
-library(gplots)
-png("results-DCA/fig.s21.png")
-plotCI(x = lambdas, y = errormeans, gap = 0.5,
-       xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
-       ylab = latex2exp::TeX("Type 1 Error/False Discovery Rates"), 
-       ylim = c(0,1), uiw = unlist(cis), pch = 19)
-points(x = lambdas, y = errormeans, type = "l", col =  "black")
+png("results-DCA/fig.4.png")
+
+plot(x = lambdas, y = unlist(type1error.zk$`32`), 
+     xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
+     ylab = latex2exp::TeX("Type 1 Error/False Discovery Rates"), 
+     type = "l", ylim = c(0,1), col = "purple")
+points(x = lambdas, y = unlist(type1error.zk$`64`), ltw=2, type = "l", col =  "blue")
+points(x = lambdas, y = unlist(type1error.zk$`128`), type = "l", col = "darkgreen")
+points(x = lambdas, y = unlist(type1error.zk$`256`), type = "l", col =  "green")
+points(x = lambdas, y = unlist(type1error.zk$`512`), type = "l", col = "yellow")
+points(x = lambdas, y = unlist(type1error.zk$`1024`), type = "l", col = "orange")
 abline(h = 0.05, col = "red")
[…]

### 2020-09-28 Manuscript/PNAS-LaTex-Template/PDF-of-PNAS-LaTex.r

@@ -1,5 +1,0 @@
-# Script to render PNAS LaTex into PDF
-
-#3: PDF of diff.tex
-tinytex::latexmk("PNAS-template-Adams.tex") 
-

### 2020-09-28 Manuscript/VersionDiff.r

@@ -1,8 +1,0 @@
-# Script to compare two Rmd files
- #Note, requires installation of 'latexdiff' in your tex components & a version of Perl
-
-# NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
-
-library(latexdiffr)
-latexdiff("Manuscript-5.Rmd", "Manuscript-6.Rmd", clean = TRUE, output = "V5-V6.diff.pdf")
-

### 2020-09-28 Manuscript/VersionDiffNew.r

@@ -1,34 +1,0 @@
-# Script to compare two Rmd files
-
-# step by step, as 'latexdiffr' no longer working (breaks in the system2(latexdiff call)
-
-#1: generate tex from Rmd
-rmarkdown::render(input =  "2020-Evol-MultiRR_Submitted.Rmd")
-rmarkdown::render(input = "2020-Evol-MultiRR_Rvsn.Rmd")
-
-#2: latexdiff of tex files
-shell(cmd="latexdiff.exe PNAS-Adams-01.tex PNAS-Adams-01-EKB.tex > diff.tex")
-
-#3: PDF of diff.tex
-tinytex::latexmk("diff.tex") 
-
-
-
-
-
-
-
-
-
-
-
-
-###########
-#Note, requires installation of 'latexdiff' in your tex components & a version of Perl
-
-# NOTE: need to install latexdiffr from: https://github.com/hughjonesd/latexdiffr
-
-##OLD WAY: not working now
-#library(latexdiffr)
-#latexdiff("2020-Evol-MultiRR_Submitted.Rmd", "2020-Evol-MultiRR_Rvsn.Rmd", clean = TRUE)
-

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.Rmd
MEE

@@
 4).
-Likewise, false discovery rates were also low, and were at or below the nominal 5% level (Fig.
+Likewise, average false discovery rates were also low, and were at or below the nominal 5% level (Fig.
 4).
-Importantly, the statistical performance of $\hat{Z}_{12}$ appears unaffected by phylogeny size.
+Importantly, the statistical performance of $\hat{Z}_{12}$ appears unaffected by phylogeny size (Supporting Information).
 Overall, these results reveal that tests based on $\hat{Z}_{12}$ have acceptable type I error and false discovery rates.
@@
 
-Figure 4. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
+Figure 4.
+Type I error rates and model misspecification for $\hat{Z}_{12}$.
+Type I error is found on the far-left of the plot ($\lambda = 0$).
+The black line signifies the average type I error and false discovery rates across simulations of different tree sizes.
 
@@
 
-```{r, echo = FALSE, out.height="75%", fig.cap="Type I error rates and model misspecification for $\\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\\lambda = 0$)."}
+```{r, echo = FALSE, out.height="75%", fig.cap="Type I error rates and model misspecification for $\\hat{Z}_{12}$.
+Type I error is found on the far-left of the plot ($\\lambda = 0$).
+The black line signifies the average type I error and false discovery rates across simulations of different tree sizes."}
 include_graphics("Figs/fig.4.png")  

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.tex
MEE

@@
 \endhead
-4201 & 26493\tabularnewline
+4221 & 26607\tabularnewline
 \bottomrule
@@
 error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig. 4). Likewise,
-false discovery rates were also low, and were at or below the nominal
-5\% level (Fig. 4). Importantly, the statistical performance of
-\(\hat{Z}_{12}\) appears unaffected by phylogeny size. Overall, these
-results reveal that tests based on \(\hat{Z}_{12}\) have acceptable type
-I error and false discovery rates. Therefore, \(\hat{Z}_{12}\) is an
-appropriate statistic for comparing the degree of phylogenetic signal
-across traits.
+average false discovery rates were also low, and were at or below the
+nominal 5\% level (Fig. 4). Importantly, the statistical performance of
+\(\hat{Z}_{12}\) appears unaffected by phylogeny size (Supporting
+Information). Overall, these results reveal that tests based on
+\(\hat{Z}_{12}\) have acceptable type I error and false discovery rates.
+Therefore, \(\hat{Z}_{12}\) is an appropriate statistic for comparing
+the degree of phylogenetic signal across traits.
 
@@
 \(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
-(\(\lambda = 0\)).
+(\(\lambda = 0\)). The black line signifies the average type I error and
+false discovery rates across simulations of different tree sizes.
 
@@
 
-\newpage
-
 \begin{figure}
-\includegraphics[width=28.51in,height=0.75\textheight]{Figs/fig.4} \caption{Type I error rates and model misspecification for $\hat{Z}_{12}$.
-Type I error is found on the far-left of the plot ($\lambda = 0$).}\label{fig:unnamed-chunk-6}
+\includegraphics[width=24.22in,height=0.75\textheight]{Figs/fig.4} \caption{Type I error rates and model misspecification for $\hat{Z}_{12}$.
+Type I error is found on the far-left of the plot ($\lambda = 0$).
+The black line signifies the average type I error and false discovery rates across simulations of different tree sizes.}\label{fig:unnamed-chunk-6}
 \end{figure}
-
-\newpage
 

### 2020-09-28 Manuscript/SI.Rmd
MEE

@@
 Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$.
-The proportion of significant results provided an estimate of the type I error or model misspecification.
+The proportion of significant results provided an estimate of the type I error or false discovery rate.
 Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$).
-Likewise, model misspecification was evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$).
+Likewise, false discovery rates were evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$).
 \hfill\break
@@
 S21).
-Likewise, levels of misspecification were also low, and were at or below the nominal 5% level (Fig.
-S21 black line).
+Likewise, levels of false discovery were also low, and were at or below the nominal 5% level (Fig.
+S21).
 Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.
@@
 
-\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).
+\textbf{Figure S21}.
+Type I error rates and false discovery rates for $\hat{Z}_{12}$.
+Type I error is found on the far-left of the plot ($\lambda = 0$).
+Each colored line represents the curve for a given level of tree size.
 

### 2020-09-28 Manuscript/SI.tex
MEE version
MEE

--- 2020-09-28 Manuscript/Figs/SI.tex
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S1}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S1}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S2}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S2}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S3}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S3}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S4}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S4}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S5}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S5}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S6}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S6}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S7}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S7}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S8}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S8}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S9}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S9}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S10}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S10}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S11}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S11}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S12}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S12}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S13}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S13}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S14}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S14}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S15}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S15}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S16}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S16}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S17}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S17}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S18}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S18}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S19}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S19}
 
@@
 
-\includegraphics[width=0.95\linewidth]{fig.S20}
+\includegraphics[width=0.95\linewidth]{Figs/fig.S20}
 
@@
 using \(\hat{Z}_{12}\). The proportion of significant results provided
-an estimate of the type I error or model misspecification. Specifically,
+an estimate of the type I error or false discovery rate. Specifically,
 type I error was evaluated when traits were simulated using no input
-phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, model
-misspecification was evaluated for data simulated with some known, but
-equal, level of phylogenetic signal (i.e., \(\lambda > 0\)).
-\hfill\break
+phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, false discovery
+rates were evaluated for data simulated with some known, but equal,
+level of phylogenetic signal (i.e., \(\lambda > 0\)). \hfill\break
 
@@
 average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
-S21). Likewise, levels of misspecification were also low, and were at or
-below the nominal 5\% level (Fig. S21 black line). Both results
-demonstrate that comparisons of the strength of phylogenetic signal
-using \(\hat{Z}_{12}\) do not display high levels of false positives,
-and are thus appropriate for comparing the degree of phylogenetic signal
-across traits.
+S21). Likewise, levels of false discovery were also low, and were at or
+below the nominal 5\% level (Fig. S21). Both results demonstrate that
+comparisons of the strength of phylogenetic signal using
+\(\hat{Z}_{12}\) do not display high levels of false positives, and are
+thus appropriate for comparing the degree of phylogenetic signal across
+traits.
 
-\includegraphics[width=6.67in,height=0.5\textheight]{fig.S21}
+\includegraphics[width=6.67in,height=0.5\textheight]{Figs/fig.S21}
 
-\textbf{Figure S21}. Type I error rates and model misspecification for
+\textbf{Figure S21}. Type I error rates and false discovery rates for
 \(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
-(\(\lambda = 0\)). The solid black line represents the average type I
-error and false discovery rate across all levels of input signal.
+(\(\lambda = 0\)). Each colored line represents the curve for a given
+level of tree size.
 

### 2020-09-28 Data_Analyses/updated_simulations/simulation.analyses.PB.type1err-DCA.R

@@ -4,10 +4,10 @@ library(geomorph)
 
 ##DCA: adjusted BC to not use observed in calculating the transformation,
 
-treesizes <- 2^(10) # 2^(5:7) done, 2^8 being run by EKB, still need to run 2^(9:10)
+treesizes <- 2^(5:10) 
 lambdas <- seq(0, 1, 0.05)
 ntrees <- 2
-ndatasets <- 25  #was 50, put back
+ndatasets <- 50  
 physig_iter <- 999
 
 #### functions from develop branch ######
@@ -176,7 +176,6 @@ write.csv(z12_mat, "results-DCA/pb.alltreesizes.20trees.50datasets.z12s.csv", ro
 #### Calculating Type 1 Error & False Discovery Rates ####
 z12_mat <- read.csv("results-DCA/pb.alltreesizes.20trees.50datasets.z12s.csv")
 
-
 z12_mat <- as.data.frame(z12_mat)
 
 type1error.zk <- lapply(1:length(treesizes), function(i){
@@ -197,9 +196,10 @@ colnames(t1err_mat) <- lambdas
 rownames(t1err_mat) <- treesizes
 
 errormeans <- colMeans(t1err_mat)
+
 cis <- lapply(1:length(lambdas), function(i) qnorm(0.975)*(sd(t1err_mat[,i])/sqrt(6)))
 
-png("results-DCA/fig.4.png")
+png("results-DCA/fig.S21.png")
 
 plot(x = lambdas, y = unlist(type1error.zk$`32`), 
      xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
@@ -220,7 +220,7 @@ dev.off()
 
 # 
 library(gplots)
-png("results-DCA/fig.S4-NotUsed.png")
+png("results-DCA/fig.4.png")
 plotCI(x = lambdas, y = errormeans, gap = 0.5,
        xlab = latex2exp::TeX("Input Phylogenetic Signal ($\\lambda_{in}$)"),
        ylab = latex2exp::TeX("Type 1 Error/False Discovery Rates"), 

### 2020-09-29 2020-CoverLetter/2020-CoverLetter.Rmd
mee

--- 2020-08-26 2020-PNAS-CoverLetter/2020-CoverLetter.Rmd
@@
 
-Prof. May Berenbaum, Editor in Chief\hfill\break
-*Proceedings of the National Academy of Sciences*
+Prof. Rob Freckleton, Executive Editor\hfill\break
+*Methods in Ecology and Evolution*
 
@@
 
-Prof. Berenbaum,
+Prof. Freckleton,
 
-Please find attached our manuscript “A standardized effect size for evaluating and comparing the strength of phylogenetic signal” which we are submitting for consideration for publication in *PNAS*.
-This manuscript is not being considered elsewhere for publication, and both authors are aware of this submission.
-We are submitting this for consideration in the **Biological Sciences** in the **Evolution** minor category.
+Please find attached our manuscript “A standardized effect size for evaluating and comparing the strength of phylogenetic signal” which we are submitting for consideration for publication in *Methods in Ecology and Evolution*.
+This manuscript is not being considered elsewhere for publication, and all authors are aware of this submission.
 \hfill\break
 
-Why appropriate for *PNAS*: Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking.
+Why appropriate for *MEE*: Evolutionary biologists wish to quantify and compare the strength of phylogenetic signal across datasets, but analytical tools for these comparisons are generally lacking.
 Here we develop a standardized effect size, $Z_K$, which measures the strength of phylogenetic signal on a common statistical scale.
@@
 
-**Editorial Board Members:** As potential Editorial Board Members who could adjudicate our work, we suggest: Dr. Scott Edwards, Dr. Douglas Futuyma, and Dr. Hopi Hoekstra. \hfill\break
-
-**NAS Members:**  NAS Members who could evaluate this work include: Dr. Joseph Felsenstein, Dr. David Hillis, and Dr. Sarah Otto.\hfill\break
-
-**Qualified Reviewers:** Additional qualified reviewers include: Dr. Simon Blomberg, Dr. Anthony Ives, and Dr. Mark Pagel. \hfill\break
+We hope you find the work suitable for consideration in *Methods in Ecology and Evolution.* \hfill\break
 

### 2021-05-11 Manuscript/2020-MEE-Adams_et_al-Physignal.tex
mee

@@
 ]{article}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
 % Correct order of tables after \paragraph or \subparagraph
@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \usepackage{lineno}\linenumbers
-
-\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
-Phylogenetic Signal}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\title{A Standardized Effect Size for Evaluating and Comparing the
+Strength of Phylogenetic Signal}
 \author{}
@@
 
-\begin{center}28 September, 2020\end{center}
+\begin{center}11 May, 2021\end{center}
 
@@
 
-\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa
-\hfill\break
+\textbf{Keywords}: comparative analysis, macroevolution \hfill\break
 
@@
 \toprule
-words & characters\tabularnewline
+words & characters \\
 \midrule
 \endhead
-4221 & 26607\tabularnewline
+4221 & 26597 \\
 \bottomrule
@@
 2011). This attribute seemed less reasonable based on the simulations
-performed by Münkemüller et al.~(2012), which suggested that
+performed by Münkemüller et al. (2012), which suggested that
 distributions were positively skewed and that Blomberg's \emph{K} might
@@
 taxa. However, while numerous analytical approaches may be used to
-quantify phylogenetic signal (e.g., Abouheif, 1999; Gittleman \& Kot,
-1990; Pagel, 1999; Blomberg et al., 2003; Adams, 2014a), methods that
+quantify phylogenetic signal (Gittleman \& Kot, 1990; e.g., Abouheif,
+1999; Pagel, 1999; Blomberg et al., 2003; Adams, 2014a), methods that
 explicitly measure the strength of phylogenetic signal, or facilitate
@@
 developed here aligns well with other statistical approaches such as
-meta-analysis (sensu Hedges \& Olkin, 1985; Glass, 1976; Arnqvist \&
+meta-analysis (Glass, 1976; sensu Hedges \& Olkin, 1985; Arnqvist \&
 Wooster, 1995), where summary statistics across datasets are converted
@@
 for \(\lambda\) and \emph{K} values are consistent with those found in
-the simulations performed by Münkemüller et al.~(2012), but that study
+the simulations performed by Münkemüller et al. (2012), but that study
 investigated type I error rates and statistical power, finding that
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abouheif1999}{}%
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Abouheif1999}{}}%
 Abouheif, E. (1999). A method for testing the assumption of phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Adams2014a}{}%
-Adams, D. C. (2014a). A generalized Kappa statistic for estimating
+\leavevmode\vadjust pre{\hypertarget{ref-Adams2014a}{}}%
+Adams, D. C. (2014a). {A generalized Kappa statistic for estimating
 phylogenetic signal from shape and other high-dimensional dultivariate
-data. \emph{Systematic Biology}, \emph{63}, 685--697.
-
-\leavevmode\hypertarget{ref-Adams2014b}{}%
+data}. \emph{Systematic Biology}, \emph{63}, 685--697.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Adams2014b}{}}%
 Adams, D. C. (2014b). A method for assessing phylogenetic least squares
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2016}{}}%
 Adams, D. C., \& Collyer, M. L. (2016). On the comparison of the
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
-Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic anova: Group-clade
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2018b}{}}%
+Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic ANOVA: Group-clade
 aggregation, biological challenges, and a refined permutation procedure.
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2019b}{}}%
 Adams, D. C., \& Collyer, M. L. (2019a). Comparing the strength of
@@
 
-\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2019}{}}%
 Adams, D. C., \& Collyer, M. L. (2019b). Phylogenetic comparative
@@
 
-\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsGeomorph}{}}%
 Adams, D. C., Collyer, M. L., \& Kaliontzopoulou, A. (2020). Geomorph:
@@
 
-\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-AdamsOtarola2013}{}}%
 Adams, D. C., \& Otárola-Castillo, E. (2013). Geomorph: An r package for
@@
 
-\leavevmode\hypertarget{ref-Arnqvist1995}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Arnqvist1995}{}}%
 Arnqvist, G., \& Wooster, D. (1995). Meta-analysis: Synthesizing
@@
 
-\leavevmode\hypertarget{ref-Baken2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Baken2019}{}}%
 Baken, E. K., \& Adams, D. C. (2019). Macroevolution of arboreality in
@@
 
-\leavevmode\hypertarget{ref-Baken2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Baken2020}{}}%
 Baken, E. K., Mellenthin, L. E., \& Adams, D. C. (2020). Macroevolution
@@
 
-\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Beaulieu_et_al2012}{}}%
 Beaulieu, J. M., Jhwueng, D. C., Boettiger, C., \& O'Meara, B. C.
@@
 
-\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Blomberg_et_al2003}{}}%
 Blomberg, S. P., Garland, T., \& Ives, A. R. (2003). Testing for
@@
 
-\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Boettiger_et_al2012}{}}%
 Boettiger, C., Coop, G., \& Ralph, P. (2012). Is your phylogeny
@@
 
-\leavevmode\hypertarget{ref-Bonett2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Bonett2017}{}}%
 Bonett, R. M., \& Blair, A. L. (2017). Evidence for complex life cycle
@@
 
-\leavevmode\hypertarget{ref-Bose2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Bose2019}{}}%
 Bose, R., Ramesh, B. R., Pélissier, R., \& Munoz, F. (2019).
@@
 
-\leavevmode\hypertarget{ref-Cohen1988}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Cohen1988}{}}%
 Cohen, J. (1988). \emph{Statistical power analysis for the behavioral
@@
 
-\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-CollyerAdams2018}{}}%
 Collyer, M. L., \& Adams, D. C. (2018). RRPP: An r package for fitting
@@
 
-\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Collyer_et_al2015a}{}}%
 Collyer, M. L., Sekora, D. J., \& Adams, D. C. (2015). A method for
@@
 
-\leavevmode\hypertarget{ref-Cooper2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Cooper2010}{}}%
 Cooper, N., Jetz, W., \& Freckleton, R. P. (2010). Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Davies2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Davies2012}{}}%
 Davies, T. J., Kraft, N. J., Salamin, N., \& Wolkovich, E. M. (2012).
@@
 
-\leavevmode\hypertarget{ref-DeMeester2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-DeMeester2019}{}}%
 De Meester, G., Huyghe, K., \& Van Damme, R. (2019). Brain size, ecology
@@
 
-\leavevmode\hypertarget{ref-DinizFilho2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-DinizFilho2012}{}}%
 Diniz-Filho, J. A. F., Santos, T., Rangel, T. F., \& Bini, L. M. (2012).
@@
 
-\leavevmode\hypertarget{ref-Felsenstein1985}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Felsenstein1985}{}}%
 Felsenstein, J. (1985). Phylogenies and the comparative method.
@@
 
-\leavevmode\hypertarget{ref-Forbes2011}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Forbes2011}{}}%
 Forbes, C., Evans, M., Hastings, N., \& Peacock, B. (2011).
@@
 
-\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Freckleton_et_al2002}{}}%
 Freckleton, R. P., Harvey, P. H., \& Pagel, M. (2002). Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-GarlandIves2000}{}%
-Garland, T. J., \& Ives, A. R. (2000). Using the past to predict the
+\leavevmode\vadjust pre{\hypertarget{ref-GarlandIves2000}{}}%
+Garland, T. Jr., \& Ives, A. R. (2000). Using the past to predict the
 present: Confidence intervals for regression equations in phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Gittleman1990}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Gittleman1990}{}}%
 Gittleman, J. L., \& Kot, M. (1990). Adaptation: Statistics and a null
@@
 
-\leavevmode\hypertarget{ref-Glass1976}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Glass1976}{}}%
 Glass, G. V. (1976). Primary, secondary, and meta-analysis of research.
@@
 
-\leavevmode\hypertarget{ref-Grafen1989}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Grafen1989}{}}%
 Grafen, A. (1989). The phylogenetic regression. \emph{Philosophical
@@
 
-\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-HarveyPagel1991}{}}%
 Harvey, P. H., \& Pagel, M. D. (1991). \emph{The comparative method in
-evolutionary biology}. Book, Oxford: Oxford University Press.
-
-\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
+evolutionary biology}. book, Oxford: Oxford University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-HedgesOlkin1985}{}}%
 Hedges, L. V., \& Olkin, I. (1985). \emph{Statistical methods for
@@
 
-\leavevmode\hypertarget{ref-Klingenberg2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Klingenberg2010}{}}%
 Klingenberg, C. P., \& Gidaszewski, N. A. (2010). Testing and
@@
 
-\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-MartinsHansen1997}{}}%
 Martins, E. P., \& Hansen, T. F. (1997). Phylogenies and the comparative
@@
 
-\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-MolinaVenegas2017}{}}%
 Molina-Venegas, R., \& Rodríguez, M. A. (2017). Revisiting phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Munkemuller_et_al2012}{}}%
 Münkemüller, T., Lavergne, S., Bzeznik, B., Dray, S., Jombart, T.,
@@
 
-\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-OMeara_et_al2006}{}}%
 O'Meara, B. C., Ane, C., Sanderson, M. J., \& Wainwright, P. C. (2006).
@@
 
-\leavevmode\hypertarget{ref-Orme2013}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Orme2013}{}}%
 Orme, D., Freckleton, R. P., Thomas, G. H., Petzoldt, T., Fritz, S. A.,
@@
 
-\leavevmode\hypertarget{ref-Pagel1999}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Pagel1999}{}}%
 Pagel, M. D. (1999). Inferring the historical patterns of biological
@@
 
-\leavevmode\hypertarget{ref-Pavoine2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Pavoine2012}{}}%
 Pavoine, S., \& Ricotta, C. (2012). Testing for phylogenetic signal in
@@
 
-\leavevmode\hypertarget{ref-Pintanel2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Pintanel2019}{}}%
 Pintanel, P., Tejedo, M., Ron, S. R., Llorente, G. A., \& Merino-Viteri,
@@
 
-\leavevmode\hypertarget{ref-Revell2010}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Revell2010}{}}%
 Revell, L. J. (2010). Phylogenetic signal and linear regression on
@@
 
-\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Revell_et_al2008}{}}%
 Revell, L. J., Harmon, L. J., \& Collar, D. C. (2008). Phylogenetic
@@
 
-\leavevmode\hypertarget{ref-Rohlf2001}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Rohlf2001}{}}%
 Rohlf, F. J. (2001). Comparative methods for the analysis of continuous
@@
 
-\leavevmode\hypertarget{ref-Rosenthal1994}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Rosenthal1994}{}}%
 Rosenthal, R. (1994). The handbook of research synthesis. In L. V.
@@
 
-\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-SokalRohlf2012}{}}%
 Sokal, R. R., \& Rohlf, F. J. (2012). \emph{Biometry} (4th ed.). San
@@
 
-\leavevmode\hypertarget{ref-Su2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Su2019}{}}%
 Su, G., Villéger, S., \& Brosse, S. (2019). Morphological diversity of
@@
 
-\leavevmode\hypertarget{ref-Vandelook2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Vandelook2019}{}}%
 Vandelook, F., Janssens, S., Gijbels, P., Fischer, E., Van den Ende, W.,
@@
 \emph{124}(2), 269--279.
+
+\end{CSLReferences}
 
